### 2014-09-12 ms.tex
sketch for new paper

%\documentclass{emulateapj}
\documentclass[letterpaper,12pt,preprint]{aastex}

% packages
\usepackage{amssymb,amsmath,amsbsy}

% commands
\newcommand{\given}{\,|\,}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transpose}[1]{{#1}^{\mathsf{T}}}
\newcommand{\inverse}[1]{{#1}^{-1}}
\newcommand{\msun}{\mathrm{M}_\odot}

\begin{document}

\title{Tidal streams in triaxial systems}
\author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn}}

% Affiliations
\newcommand{\colum}{1}
\newcommand{\adrn}{2}
\altaffiltext{\colum}{Department of Astronomy, 
		              Columbia University, 
		              550 W 120th St., 
		              New York, NY 10027, USA}
\altaffiltext{\adrn}{To whom correspondence should be addressed: adrn@astro.columbia.edu}

\begin{abstract}
% Context
% Aims
% Methods
% Results
% Conclusions
\end{abstract}

\keywords{
}

%\begin{figure*}[!h]
%\begin{center}
%\includegraphics[width=\textwidth]{PATH}
%\caption{ CAPTION }\label{fig:FIGNAME}
%\end{center}
%\end{figure*}

\section{Introduction}\label{sec:introduction}

The halos of galaxies are filled with substructure...

As a satellite galaxy or globular cluster orbits within a parent galaxy, its mass is eroded due to the tidal forces of the host galaxy. For (typically) many orbits, these tidally stripped stars remain spatially coherent and

Of particular interest are such cold, young dynamical structures --- ``tidal streams'' -- which provide useful information about the mass distributions at large distances in the galaxies within which they orbit. [At these distances, disk / classic baryonic components can't help you, mostly dark matter]

The formation and morphology of tidal streams has been studied extensively in a variety of assumed potential forms. [Stream formation

\section{Methods}\label{sec:methods}
\subsection{Potential form}

\section{Experiments}\label{sec:experiments}

\section{Discussion}\label{sec:discussion}

\acknowledgements
APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 
This research made use of Astropy, a community-developed core \texttt{Python} package for Astronomy \citep{astropy13}.
This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance. \\

\bibliographystyle{apj}
\bibliography{refs}

\end{document}

### 2014-10-01 ms.tex
started writing words

@@
 \newcommand{\msun}{\mathrm{M}_\odot}
+\newcommand{\bs}[1]{\boldsymbol{#1}}
 
@@
 \begin{abstract}
-% Context
-% Aims
-% Methods
-% Results
-% Conclusions
+
+Tidal streams form from the steady disruption of stellar systems
+orbiting within the gravitational field of some parent galaxy. Many
+streams and debris structures have been discovered in the halo of the
+Milky Way and have been used to model the potential of the Galaxy. 
+However, few of these models have yet explored the properties of
+tidal debris in triaxial potentials.
+The existence of a variety of orbits, resonances, and chaotic regions in such potentials 
+suggest that the morphologies and dispersal timescales of
+debris could differ significantly from the simpler spherical and oblate cases.
+In this work we use a series of N-body simulations of stellar systems over a range 
+of masses of disruption in triaxial potentials to understand the influence of the nature 
+and types of orbits on debris morphologies.
+Our results suggest that the mere existence of the multitude of thin streams
+already known to orbit the Milky Way provides significant constraints on the
+classes of triaxial potentials that provide a good representation for its dark matter halo.
+
 \end{abstract}
@@
 The halos of galaxies are filled with substructure...
-
 As a satellite galaxy or globular cluster orbits within a parent galaxy, its mass is eroded due to the tidal forces of the host galaxy.
-For (typically) many orbits, these tidally stripped stars remain spatially coherent and
-
-Of particular interest are such cold, young dynamical structures --- ``tidal streams'' -- which provide useful information about the mass distributions at large distances in the galaxies within which they orbit.
+For (typically) many orbits, these tidally stripped stars remain spatially coherent.
+Of particular interest are cold, young dynamical structures --- ``tidal streams'' -- which provide useful information about the mass distributions at large distances in the galaxies within which they orbit.
 [At these distances, disk / classic baryonic components can't help you, mostly dark matter]
-
-The formation and morphology of tidal streams has been studied extensively in a variety of assumed potential forms. [Stream formation
+The formation and morphology of tidal streams has been studied extensively in a variety of assumed potential forms. [Stream formation in oblate, spherical potentials, action-angle, etc.]
 
@@
 
+\subsection{Tidal disruption in action-angle coordinates}
+When a star is tidally stripped from a progenitor system, the difference in angle coordinates, relative to the progenitor, grows linearly with time according to
+\begin{equation}
+	\Delta\bs{\theta} = \Delta\bs{\theta}(t=0) + \Delta\bs{\Omega}t
+\end{equation}
+where $\Delta\bs{\theta}(t=0)$ is the initial offset (at time of stripping, $t=0$) between the star and progenitor, and $\Delta\bs{\Omega}$ is the corresponding difference in fundamental frequencies.
+Once a star has ``entered'' the stream, the constant offset becomes small compared to the term that depends on time, so that
+\begin{equation}
+	\Delta\bs{\theta} \approx \Delta\bs{\Omega}t.
+\end{equation}
+If the spread of actions of the progenitor is small enough --- corresponding to a mass [or mass ratio?] of XXX --- the frequency vector for a given star may be approximated by a Taylor expansion about the frequency vector of the progenitor,
+\begin{equation}
+	\Omega_i \approx \Omega_{0,i} + \frac{\partial^2 H}{\partial J_i J_j}\Delta J_j + \mathcal{O}(\Delta J^2)
+\end{equation}
+where we make use of the Einstein summation convention in the above expression.
+To first order in the difference in actions, the difference in frequencies between a star and the progenitor is approximated by
+\begin{equation}
+	\Delta\Omega_i \approx \frac{\partial^2 H}{\partial J_i J_j}\Delta J_j = D_{ij}\Delta J_j.\label{eq:deltafreq}
+\end{equation}
+where $D_{ij}$ is the Hessian of the potential in action-space.
+For a thin stream to form, the Hessian must be dominated by a single large eigenvalue, where the corresponding eigenvector determines the direction along which the debris will spread in the angles.
+If two or all of the largest eigenvalues are comparable in size, the debris will not form a thin stream and the observability (projected surface density) of the debris structure will be diminished.
+We would like to be able to predict the morphology of a debris structure formed from a disrupting progenitor on a given orbit in a specified potential using this fact.
+
+\subsection{Numerically computing the Hessian in action-space}
+The Hessian can be computed for an orbit in a given potential provided with a numerical scheme for computing the actions, angles, and frequencies of an orbit in an arbitrary potential.
+This is a non-trivial task, and we use a custom implementation of the method presented in \citep{sanders14} to compute these quantities.
+The procedure for computing the Hessian is as follows:
+\begin{enumerate}
+	\item numerically integrate the orbit of interest in the test potential;
+	\item fit a \emph{toy potential} \citep[see][]{sanders14} to the orbit, and create a 3D grid of actions in this toy potential;
+	\item transform the actions and angles of the grid from the toy potential to ordinary phase-space;
+	\item integrate the grid of initial conditions in the test potential;
+	\item solve for actions, angles, and frequencies of this grid;
+	\item use the (now deformed) grid in action-space to linearly interpolate a regular grid about the orbit of interest;
+	\item use a finite-difference scheme to compute first derivatives of the frequency vectors ($D_{ij} = \partial \Omega_i / \partial J_j$).
+\end{enumerate}
+Once we compute the Hessian, we may solve for the eigenvalues and eigenvectors of the matrix.
+
+An alternate approach uses the debris particles themselves to solve for the components of the Hessian at position (in action-space) of the progenitor orbit.
+Equation \ref{eq:deltafreq} demonstrates the relationship between the components of the relative frequency vector to the components of the relative action vector (through the Hessian) for a single particle.
+We now consider a case where we have a set of $K$ particles that we know should obey the equations
+\begin{equation}
+	\Delta\Omega_i^{(k)} = D_{ij}\Delta J_j^{(k)}.
+\end{equation}
+This is analogous to the classic problem in linear algebra of solving an overconstrained set of linear equations, $\mathbf{y} = \mathbf{A}\mathbf{x}$.
+Here, $\mathbf{y}$ is the length-$3K$ (column) vector of frequencies for each star, $\mathbf{x}$ is a length-9 (column) vector containing the ``unravelled'' values of the Hessian, and $\mathbf{A}$ is a $3K\times 9$ matrix containing the relative actions for all stars:
+\begin{align}
+	\mathbf{y} &= \left[ \begin{array}{c} 
+					\Delta\Omega_1^{(1)} \\ \Delta\Omega_2^{(1)} \\ \Delta\Omega_3^{(1)} \\ \vdots \\ 
+					\Delta\Omega_1^{(K)} \\ \Delta\Omega_2^{(K)} \\ \Delta\Omega_3^{(K)}
+				\end{array} \right]\\
+	\mathbf{A} &= \begin{bmatrix} 
+					\Delta J_1^{(1)} & \Delta J_2^{(1)} & \Delta J_3^{(1)} & 0 & 0 & 0 & 0 & 0 & 0  \\
+					0 & 0 & 0 & \Delta J_1^{(1)} & \Delta J_2^{(1)} & \Delta J_3^{(1)} & 0 & 0 & 0  \\
+					0 & 0 & 0 & 0 & 0 & 0 & \Delta J_1^{(1)} & \Delta J_2^{(1)} & \Delta J_3^{(1)}  \\
+					&&&&\vdots \\
+					\Delta J_1^{(K)} & \Delta J_2^{(K)} & \Delta J_3^{(K)} & 0 & 0 & 0 & 0 & 0 & 0  \\
+					0 & 0 & 0 & \Delta J_1^{(K)} & \Delta J_2^{(K)} & \Delta J_3^{(K)} & 0 & 0 & 0  \\
+					0 & 0 & 0 & 0 & 0 & 0 & \Delta J_1^{(K)} & \Delta J_2^{(K)} & \Delta J_3^{(K)}  \\
+			       \end{bmatrix}\\
+	\mathbf{x} &= \left[ \begin{array}{c} D_{11} \\ D_{12} \\ \vdots \\ D_{33} \end{array} \right].
+\end{align}
+The parameter vector, $\mathbf{x}$, is found trivially with ordinary least squares.
+
+% \begin{bmatrix} xz & xw \\ yz & yw \end{bmatrix}
+
 \section{Experiments}\label{sec:experiments}
+
+\subsection{Comparing Hessian solvers}
 

### 2014-12-22 ms.tex
matrices
added poster and ignore illustrator

@@
 \title{Tidal streams in triaxial systems}
-\author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn}}
+\author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
+	    Kathryn V. Johnston\altaffilmark{\colum},
+	    Monica Valluri\altaffilmark{\mich}}
 
@@
 \newcommand{\adrn}{2}
+\newcommand{\mich}{3}
 \altaffiltext{\colum}{Department of Astronomy, 
@@
 \altaffiltext{\adrn}{To whom correspondence should be addressed: adrn@astro.columbia.edu}
+\altaffiltext{\mich}{ }
 
@@
 
+Adding third degree of freedom opens up many new orbits / resonances / etc. -- Valluri and Merritt work...
+
 The halos of galaxies are filled with substructure...
@@
 
+\section{Outline}
+\begin{itemize}
+\item Start by choosing a potential and generating a frequency map for grids of loop orbits in this potential.
+We will use two potentials: a triaxial, NFW halo and a three-component sum of a triaxial NFW halo + disk + bulge.
+We will use the loop orbits to analyze properties of tidal streams along different orbits in these potentials.
+\item Similar to Valluri \& Merritt (1998), we'll generate initial conditions in the $x-z$ plane with $v_x = v_z = y=0$. We'll use fixed energies (from pre-chosen energy shells) to get $v_y$. 
+\item After classifying orbits, special regions, and frequency diffusion rates, we will pick a few representative cases and integrate balls of particles around those initial conditions forward for some integration time, $T_{\rm int}$.
+From this, we can hopefully get an idea of how the morphology of the resultant debris changes with frequency diffusion rate.
+Does the disk matter?
+(previous Valluri work suggests yes).
+What happens to debris on orbits near resonances?
+\item For some of the above cases, run full N-body simulations and look at the debris properties.
+Are there morphological differences?
+(Note: can we trust SCF simulations?
+SCF uses Leapfrog integration, not great for orbits with diffusing frequencies...)
+
+\end{itemize}
+
+{\bf Note: ignore the sections below! They are just notes for myself...}
+
 \section{Methods}\label{sec:methods}
+
+For the debris morphology, ``progenitor'' matters, but also 2nd order potential effects e.g. hessian?
+
 \subsection{Potential form}
@@
 
-\subsection{Comparing Hessian solvers}
+\subsection{Comparing Hessian solutions}
 

### 2015-02-02 ms.tex
working on introduction

@@
 \usepackage{amssymb,amsmath,amsbsy}
+\usepackage{forest}
 
@@
 The formation and morphology of tidal streams has been studied extensively in a variety of assumed potential forms. [Stream formation in oblate, spherical potentials, action-angle, etc.]
+
+For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (d.o.f.), the equations of motion are most simply expressed in angle-action coordinates.
+In these coordinates, the actions, $J_n$, are integrals of the motion, and the conjugate angle variables, $\theta_n$, increase linearly with time
+\begin{align}
+	J_n &= {\rm const.}\\
+	\dot{J}_n &= -\frac{\partial H}{\partial \theta_n} = 0\\
+	\theta_n(t) &= \Omega_n t + \theta_n(0)\\
+	\dot{\theta}_n &= \frac{\partial H}{\partial J_n} = \Omega_n(J_1,...,J_n)
+\end{align}
+where $n=1,2,...N$, and $\Omega_n$ are the fundamental frequencies of the orbit.
+Owing to the involutive property of the actions and the periodic nature of the angle variables, the topology of angle-action space is equivalent to that of the surface of an $N$-dimensional torus, and thus orbits are often referred to in terms of \emph{orbital tori}; each set of actions, $(J_1,J_2,...J_N)$, uniquely determine the torus associated with a given orbit.
+
+If galactic halos are triaxial, their gravitational potentials will almost certainly not be globally integrable.
+For a given halo, if the potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
+\begin{equation}
+	H(\bs{\theta}, \bs{J}) = H_0(\bs{J}) + \epsilon H_1(\bs{\theta}, \bs{J})
+\end{equation}
+then a large number of regular orbits will survive the perturbation (e.g., the KAM theorem; cite Arnold)\footnote{3-vectors are represented with bold symbols}.
+The tori that survive will in general by separated by regions of irregular or chaotic motion, and thus any transformations to angle-action coordinates can only be defined locally.
+The types of motion generic to typical triaxial systems can be split into several sub-categories (summarized in Figure~\ref{fig:orbit-tree}).
+Regular motion can be classified by the number of resonance relations it obeys \citep[e.g.,][]{valluri98}: \emph{quasi-periodic} orbits obey no resonance relations, \emph{uni-resonant} orbits obey a single relation of the form $\bs{m}\cdot\bs{\Omega}=0$, and \emph{bi-resonant} orbits obey two resonance relations, $\bs{m}\cdot\bs{\Omega}=0$ and $\bs{n}\cdot\bs{\Omega}=0$, where $\bs{m}$ and $\bs{n}$ are integer vectors.
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{figures/orbit-tree.pdf}
+\caption{} \label{fig:orbit-tree}
+\end{center}
+\end{figure*}
+
+For small perturbations, quasi-periodic, resonant, and periodic orbits all survive. As the strength of the perturbation increases, the tori associated with quasi-periodic motion
+
+the (bold symbols represent 3-vectors)
+
+\section{Numerical methods}
+
+\subsection{Numerical estimation of fundamental frequencies}
+
+\subsection{Particle balls}
+
+\section{Experiments}
+
+\subsection{Potential choice}
+
+\subsection{Frequency map}
+
+\subsection{Morphology}
+
+\subsection{Morphology}
 

### 2015-02-08 ms.tex
words

--- 2014-10-01 ms.tex
@@
 \usepackage{amssymb,amsmath,amsbsy}
+\usepackage{bbold}
+\usepackage{arydshln}
+\usepackage{algpseudocode}
+\renewcommand\algorithmicthen{}
+\renewcommand\algorithmicdo{}
 
@@
 \newcommand{\bs}[1]{\boldsymbol{#1}}
+\newcommand{\ident}{\mathbb{1}}
+
+\newcommand{\act}{J}
+%\newcommand{\jac}{\bs{\rm J}}
+\newcommand{\jac}{\bs{G}}
 
@@
 
-\section{Methods}\label{sec:methods}
-\subsection{Potential form}
+For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (d.o.f.), the equations of motion are most simply expressed in angle-action coordinates.
+In these coordinates, the actions, $J_n$, are integrals of the motion, and the conjugate angle variables, $\theta_n$, increase linearly with time
+\begin{align}
+	\act_n &= {\rm const.}\\
+	\dot{\act}_n &= -\frac{\partial H}{\partial \theta_n} = 0\\
+	\theta_n(t) &= \Omega_n t + \theta_n(0)\\
+	\dot{\theta}_n &= \frac{\partial H}{\partial \act_n} = \Omega_n(\act_1,...,\act_n)\label{eq:aafreq}
+\end{align}
+where $n=1,2,...N$, and $\Omega_n$ are the fundamental frequencies of the orbit.
+Owing to the involutive property of the actions and the periodic nature of the angle variables, the topology of angle-action space is equivalent to that of the surface of an $N$-dimensional torus, and thus orbits are often referred to in terms of \emph{orbital tori}; each set of actions, $(\act_1,\act_2,...\act_N)$, uniquely determine the torus associated with a given orbit.
 
-\subsection{Tidal disruption in action-angle coordinates}
-When a star is tidally stripped from a progenitor system, the difference in angle coordinates, relative to the progenitor, grows linearly with time according to
+If galactic halos are triaxial, their gravitational potentials will almost certainly not be globally integrable.
+For a given halo, if the potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
 \begin{equation}
-	\Delta\bs{\theta} = \Delta\bs{\theta}(t=0) + \Delta\bs{\Omega}t
+	H(\bs{\theta}, \bs{\act}) = H_0(\bs{\act}) + \epsilon H_1(\bs{\theta}, \bs{\act})
 \end{equation}
-where $\Delta\bs{\theta}(t=0)$ is the initial offset (at time of stripping, $t=0$) between the star and progenitor, and $\Delta\bs{\Omega}$ is the corresponding difference in fundamental frequencies.
-Once a star has ``entered'' the stream, the constant offset becomes small compared to the term that depends on time, so that
+then a large number of regular orbits will survive the perturbation (e.g., the KAM theorem; cite Arnold)\footnote{3-vectors are represented with bold symbols}.
+The tori that survive will in general be separated by regions of irregular or chaotic motion, and thus any transformations to angle-action coordinates can only be defined locally.
+The types of motion generic to typical triaxial systems can be split into several sub-categories (summarized in Figure~\ref{fig:orbit-tree}).
+Regular motion can be classified by the number of resonance relations obeyed \citep[e.g.,][]{lichtenberg83, valluri98}: \emph{conditionally periodic} orbits obey no resonance relations, \emph{uni-resonant} orbits obey a single relation of the form $\bs{m}\cdot\bs{\Omega}=0$, and \emph{bi-resonant} orbits obey two resonance relations, $\bs{m}\cdot\bs{\Omega}=0$ and $\bs{n}\cdot\bs{\Omega}=0$, where $\bs{m}$ and $\bs{n}$ are integer vectors.
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{figures/orbit-tree.pdf}
+\caption{} \label{fig:orbit-tree}
+\end{center}
+\end{figure*}
+
+Resonant orbits are dynamically important in systems with several degrees of freedom as they determine the structure of orbit-space.
+Stable resonant orbits generate families of orbits that behave or appear similarly, whereas unstable resonances typically generate regions of stochasticity.
+Not all orbits can be trivially associated with resonances, however, due to the infinitely complex, hierarchical structure of stable resonances.
+In autonomous, two d.o.f.
+systems, stochastic regions are generally surrounded and separated by regular motion, but in three or more d.o.f.
+the chaotic regions will connect and overlap; this is the origin of the phenomenon of Arnold diffusion \citep{arnold79}.
+The implication of Arnold diffusion is that chaotic orbits may radically change shape by means of traversing the web of unstable resonances and are strictly only bounded to their own energy hypersurface.
+Another way to understand this phenomenon is by means of orbital tori: in two dimensions, if a given orbital torus is destroyed due to perturbations, the orbit is still bounded between the surrounding tori due to energy conservation.
+In three dimensions, the extra degree of freedom allows chaotic orbits to ``escape'' confinement.
+
+For small perturbations, generically, many regular orbits survive and only small chaotic islands are introduced.
+As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed, then the uni-resonant tori are destroyed, and finally the bi-resonant tori -- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.
+The irregular or chaotic orbits that fill the parts of action-space where tori have been destroyed are loosely classified by being either \emph{strongly} or \emph{weakly} chaotic.
+Qualitatively, strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale.
+There are many methods for classifying and measuring the degree of stochasticity of an orbit.
+
+\section{Numerical methods}
+
+\subsection{Potential choice}
+
+The density distributions within dark matter halos formed in cosmological N-body simulations are typically triaxial (e.g., millenium, aquarius, via lactea).
+\citet{jing02} found that a triaxial generalization of the classic NFW density profile \citep{navarro96} generates excellent fits to their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of the many halos formed in their simulations.
+They find median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that \citet{jing02} use the opposite notation so that $c$ is the major and $a$ is the minor axis.} They find significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).
+
+All of these parameters are specified in terms of the \emph{density}; for orbit analysis, we need to determine the form of the potential in terms of these parameters, which, in general, requires numerical integration of the density at each position of interest.
+For computational efficiency, many authors instead express the triaxiality in the form of the potential, however this can lead to unphysical situations where the density becomes negative.
+\citet{lee03} derive a perturbative expansion of the potential integral and show that the expansion is accurate even for modest axis ratios (e.g., the median values shown above).
+
+For this work, we use the triaxial potential expression from \citet{lee03}, parametrized in a slightly different manner.
+In terms of spherical coordinates\footnote{(radius, azimuth, colatitude)} with the radius normalized by the scale radius, $u = r/r_s$
+\begin{align}
+	\Phi(u,\phi,\theta) \approx \frac{v_c^2}{\ln2 - 1/2}\left[F_1(u) + \frac{1}{2}(e_b^2 + e_c^2)F_2(u) + \frac{1}{2} [(e_b\sin\theta \sin\phi)^2 + (e_c\cos\theta)^2] F_3(u) \right]
+\end{align}
+where $v_c$ is the circular velocity at the scale radius, $r_s$ (for the spherical case), $e_b = \sqrt{1 - (b/a)^2}$, and $e_c = \sqrt{1 - (c/a)^2}$.
+The functions $F_i(u)$ are given in the appendix of \cite{lee03}.
+We chose $r_s=20~{\rm kpc}$ and $v_c \approx 200~{\rm km}~{\rm s}^{-1}$ , which [...] (Figure~\ref{fig:potential}).
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=0.75\textwidth]{figures/potential.pdf}
+\caption{Rotation curve along $x$, $y$, and $z$ axes for the triaxial, NFW potential } \label{fig:orbit-tree}
+\end{center}
+\end{figure*}
+
+[see discussion section for adding terms (disk, time evolution) to potential]
+
+\subsection{Orbit integration}
+
+[Use a blah blah integrator...]
+
+\subsection{Lyapunov exponents}
+
+The simplest and perhaps most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent of an orbit.
+Defining the vector $\bs{w} = (q_1,...,q_N,p_1,...,p_N)$ for some set of canonical coordinates $(q_i,p_i)$, we can write Hamilton's equations as
 \begin{equation}
-	\Delta\bs{\theta} \approx \Delta\bs{\Omega}t.
+	\dot{\bs{w}} = \bs{\mathcal{J}} \frac{\partial H}{\partial \bs{w}} = f(\bs{w}) \label{eq:ham}
 \end{equation}
-If the spread of actions of the progenitor is small enough --- corresponding to a mass [or mass ratio?] of XXX --- the frequency vector for a given star may be approximated by a Taylor expansion about the frequency vector of the progenitor,
+where $\bs{\mathcal{J}}$ is the $2N \times 2N$ canonical Poisson tensor (also called the symplectic matrix) defined by
 \begin{equation}
-	\Omega_i \approx \Omega_{0,i} + \frac{\partial^2 H}{\partial J_i J_j}\Delta J_j + \mathcal{O}(\Delta J^2)
+	\bs{\mathcal{J}} = \left( \begin{array}{c:c} 0 & \ident \\ \hdashline -\ident & 0 \end{array} \right)
 \end{equation}
-where we make use of the Einstein summation convention in the above expression.
-To first order in the difference in actions, the difference in frequencies between a star and the progenitor is approximated by
+with $N$-dimensional identity matrices $\ident$.
+If we consider a nearby position separated from $\bs{w}$ by an infinitesimal deviation $\delta \bs{w}$, such that $\bs{w}' = \bs{w} + \delta \bs{w}$.
+We can expand to linear order in the deviation about the parent orbit and write the equations of motion for the deviation as follows
+\begin{align}
+	\dot{\bs{w}}' &= f(\bs{w}')\\
+	\dot{\bs{w}} + \dot{\delta \bs{w}} &= f(\bs{w} +  \delta \bs{w})\\
+	\dot{\bs{w}} + \dot{\delta \bs{w}} &\approx f(\bs{w}) + \jac_{\bs{w}}\cdot\delta \bs{w} + \mathcal{O}(\delta \bs{w}^2)\\
+	\dot{\delta \bs{w}} &\approx \jac_{\bs{w}} \cdot  \delta \bs{w}\label{eq:deviate}
+\end{align}
+where $\jac$ is the Jacobian of Eq.~\ref{eq:ham}, evaluated at the parent orbit.
+For chaotic orbits, the maximum eigenvalue of the solution matrix to Eq.~\ref{eq:deviate} is positive real, leading to exponential divergence of nearby orbits.
+
+Computing the Lyapunov spectrum for a given orbit is often not necessary if one is only interested in characterizing the degree of chaos.
+Instead, it is often more efficient to compute the maximum Lyapunov exponent by estimating the finite-time maximum Lyapunov exponent (FTMLE), defined as
 \begin{equation}
-	\Delta\Omega_i \approx \frac{\partial^2 H}{\partial J_i J_j}\Delta J_j = D_{ij}\Delta J_j.\label{eq:deltafreq}
+	l_{\rm max}(t) = \frac{1}{t}\ln \frac{\|\delta \bs{w}(t)\|}{\|\delta \bs{w}_0\|}.\label{eq:lmax}
 \end{equation}
-where $D_{ij}$ is the Hessian of the potential in action-space.
-For a thin stream to form, the Hessian must be dominated by a single large eigenvalue, where the corresponding eigenvector determines the direction along which the debris will spread in the angles.
-If two or all of the largest eigenvalues are comparable in size, the debris will not form a thin stream and the observability (projected surface density) of the debris structure will be diminished.
-We would like to be able to predict the morphology of a debris structure formed from a disrupting progenitor on a given orbit in a specified potential using this fact.
+The maximum Lyapunov exponent is the limit as $t\rightarrow \infty$ of the FTMLE
+\begin{equation}
+	\lambda_{\rm max} = \lim_{t\rightarrow\infty}l_{\rm max}(t). \label{eq:lyapmax}
+\end{equation}
+Numerically computing this quantity is not trivial because (1) obviously the limit to infinity is not possible and (2) the norm of the deviation vector $\|\delta \bs{w}(t)\|$ is expected to increase exponentially for chaotic orbits, leading to numerical problems.
+To circumvent these issues, it is sufficient to instead start a nearby orbit with some small initial deviation with norm $\delta_0$, integrate for a sufficiently small amount of time, $\tau$, then renormalize the deviation back to the initial norm (e.g., cite Bennetin et al.
+1976, Tabor 1989).
+The following pseudocode outlines this procedure:\footnote{see Gary for Python implementation?}\\
+\begin{algorithmic}[1]
+\State {\bf define} orbital integration timestep, $h$, and number of steps, $K$
+\State {\bf define} initial norm of deviation vector, $\delta_0$, to be sufficiently small
+\State {\bf define} renormalization integration period, $\tau \ll hK$ 
+\For{each timestep when integrating the main orbit, $\bs{w}(t)$}
+\State step forward the orbit and deviation vector orbit by one timestep, $t_{i-1} \rightarrow t_i$
+\If {a normalization timestep}
+\State measure and store the length of the deviation vector, $\delta_i = \|\delta \bs{w}_i\|$
+\State renormalize the length of the deviation vector, $\delta \bs{w}_i = \delta \bs{w}_i (\delta_0/\delta_i)$
+\EndIf
+\EndFor 
+\end{algorithmic}
+The FTMLE after a given number of timesteps, $N$, is then estimated as
+\begin{equation}
+	l_N = \frac{1}{N\tau}\sum_i^N \ln \, \delta_i
+\end{equation}
+and the MLE is
+\begin{equation}
+	\lambda_{\rm max} = \lim_{N\rightarrow \infty} l_N.
+\end{equation}
+[Some more practical issues with implementation] [Figure showing $l_N$ for regular, chaotic orbit?]
 
-\subsection{Numerically computing the Hessian in action-space}
-The Hessian can be computed for an orbit in a given potential provided with a numerical scheme for computing the actions, angles, and frequencies of an orbit in an arbitrary potential.
-This is a non-trivial task, and we use a custom implementation of the method presented in \citep{sanders14} to compute these quantities.
-The procedure for computing the Hessian is as follows:
-\begin{enumerate}
-	\item numerically integrate the orbit of interest in the test potential;
-	\item fit a \emph{toy potential} \citep[see][]{sanders14} to the orbit, and create a 3D grid of actions in this toy potential;
-	\item transform the actions and angles of the grid from the toy potential to ordinary phase-space;
-	\item integrate the grid of initial conditions in the test potential;
-	\item solve for actions, angles, and frequencies of this grid;
-	\item use the (now deformed) grid in action-space to linearly interpolate a regular grid about the orbit of interest;
-	\item use a finite-difference scheme to compute first derivatives of the frequency vectors ($D_{ij} = \partial \Omega_i / \partial J_j$).
-\end{enumerate}
-Once we compute the Hessian, we may solve for the eigenvalues and eigenvectors of the matrix.
+[Copy in text fomr ``behavior of MLE'']
 
-An alternate approach uses the debris particles themselves to solve for the components of the Hessian at position (in action-space) of the progenitor orbit.
-Equation \ref{eq:deltafreq} demonstrates the relationship between the components of the relative frequency vector to the components of the relative action vector (through the Hessian) for a single particle.
-We now consider a case where we have a set of $K$ particles that we know should obey the equations
-\begin{equation}
-	\Delta\Omega_i^{(k)} = D_{ij}\Delta J_j^{(k)}.
-\end{equation}
-This is analogous to the classic problem in linear algebra of solving an overconstrained set of linear equations, $\mathbf{y} = \mathbf{A}\mathbf{x}$.
-Here, $\mathbf{y}$ is the length-$3K$ (column) vector of frequencies for each star, $\mathbf{x}$ is a length-9 (column) vector containing the ``unravelled'' values of the Hessian, and $\mathbf{A}$ is a $3K\times 9$ matrix containing the relative actions for all stars:
-\begin{align}
-	\mathbf{y} &= \left[ \begin{array}{c} 
-					\Delta\Omega_1^{(1)} \\ \Delta\Omega_2^{(1)} \\ \Delta\Omega_3^{(1)} \\ \vdots \\ 
-					\Delta\Omega_1^{(K)} \\ \Delta\Omega_2^{(K)} \\ \Delta\Omega_3^{(K)}
-				\end{array} \right]\\
-	\mathbf{A} &= \begin{bmatrix} 
-					\Delta J_1^{(1)} & \Delta J_2^{(1)} & \Delta J_3^{(1)} & 0 & 0 & 0 & 0 & 0 & 0  \\
-					0 & 0 & 0 & \Delta J_1^{(1)} & \Delta J_2^{(1)} & \Delta J_3^{(1)} & 0 & 0 & 0  \\
-					0 & 0 & 0 & 0 & 0 & 0 & \Delta J_1^{(1)} & \Delta J_2^{(1)} & \Delta J_3^{(1)}  \\
-					&&&&\vdots \\
-					\Delta J_1^{(K)} & \Delta J_2^{(K)} & \Delta J_3^{(K)} & 0 & 0 & 0 & 0 & 0 & 0  \\
-					0 & 0 & 0 & \Delta J_1^{(K)} & \Delta J_2^{(K)} & \Delta J_3^{(K)} & 0 & 0 & 0  \\
-					0 & 0 & 0 & 0 & 0 & 0 & \Delta J_1^{(K)} & \Delta J_2^{(K)} & \Delta J_3^{(K)}  \\
-			       \end{bmatrix}\\
-	\mathbf{x} &= \left[ \begin{array}{c} D_{11} \\ D_{12} \\ \vdots \\ D_{33} \end{array} \right].
-\end{align}
-The parameter vector, $\mathbf{x}$, is found trivially with ordinary least squares.
+[Computationally, expensive to compute Lyapunov exponent for sticky or mildly chaotic orbits]
 
-% \begin{bmatrix} xz & xw \\ yz & yw \end{bmatrix}
+\subsection{Numerical estimation of fundamental frequencies}
 
-\section{Experiments}\label{sec:experiments}
+Bounded, regular orbits in triaxial potentials will have three \emph{fundamental frequencies}, $\bs{\Omega}$, that determine the periodic behavior of motion (Eq.~\ref{eq:aafreq}).
 
-\subsection{Comparing Hessian solvers}
+[Figure: reproduce some result?]
+
+or
+[Figure: for same orbits as in Lyapunov section, show frequency diffusion?]
+
+\subsection{Particle balls}
+
+Need a fast way to characterize morphology of local orbits after finite time
+
+Can't run N-body for many tens of thousands of initial conditions
+
+Explain simplistic approach
+
+[Figure: Compare N-body simulation to balls of particles -- captures the essence]
+
+Cite papers that use Lagrange-point stripping methods, Rewinder paper
+
+Like only considering debris that comes off at a single pericentric passage
+
+\section{Experiments}
+
+\subsection{Frequency map}
+
+We generate frequency maps at X energies for tube orbits started around the major and intermediate axes of the potential (figures)
+
+Explain qualitative differences between the two -- e.g., Y-Z plane will have many more stochastic tubes because intermediate axis
+
+\subsection{Morphology}
+
+\subsubsection{Entropy evolution}
+
+\subsubsection{Density evolution}
+
+Compare mean density or some other metric for the balls, show that decreases faster for chaotic orbit -- chaotic mixing dominates over phase-mixing
+
+Power law (Helmi and White)
+
+\subsubsection{Adding a disk?}
 
 \section{Discussion}\label{sec:discussion}
+
+\subsection{Potential}
+
+Time dependence
+
+Radial dependence of axis ratios
+
+Why not disk + bulge? What do we expect when we add? See later section for test of this...
 

### 2015-02-08 ms.tex
filling in more text

@@
 \begin{align}
-	\Phi(u,\phi,\theta) \approx \frac{v_c^2}{\ln2 - 1/2}\left[F_1(u) + \frac{1}{2}(e_b^2 + e_c^2)F_2(u) + \frac{1}{2} [(e_b\sin\theta \sin\phi)^2 + (e_c\cos\theta)^2] F_3(u) \right]
+	\Phi(u,\phi,\theta) &\approx \frac{v_c^2}{A}\left[F_1(u) + \frac{1}{2}(e_b^2 + e_c^2)F_2(u) + \frac{1}{2} [(e_b\sin\theta \sin\phi)^2 + (e_c\cos\theta)^2] F_3(u) \right]\\
+	A &= \left(\ln2 - \frac{1}{2}\right) + \left(\ln2-\frac{3}{4}\right) (e_b^2 + e_c^2)
 \end{align}
-where $v_c$ is the circular velocity at the scale radius, $r_s$ (for the spherical case), $e_b = \sqrt{1 - (b/a)^2}$, and $e_c = \sqrt{1 - (c/a)^2}$.
+where $v_c$ is the circular velocity at the scale radius, $r_s$, along the $x$-axis, $e_b = \sqrt{1 - (b/a)^2}$, and $e_c = \sqrt{1 - (c/a)^2}$.
 The functions $F_i(u)$ are given in the appendix of \cite{lee03}.
-We chose $r_s=20~{\rm kpc}$ and $v_c \approx 200~{\rm km}~{\rm s}^{-1}$ , which [...] (Figure~\ref{fig:potential}).
+We chose $r_s=20~{\rm kpc}$ and $v_c = 200~{\rm km}~{\rm s}^{-1}$ to [...].
+Figure~\ref{fig:potential} shows the rotation curve of this potential along the three principal axes.
 
@@
 \includegraphics[width=0.75\textwidth]{figures/potential.pdf}
-\caption{Rotation curve along $x$, $y$, and $z$ axes for the triaxial, NFW potential } \label{fig:orbit-tree}
+\caption{Rotation curve along $x$, $y$, and $z$ axes for the triaxial, NFW potential } \label{fig:potential}
 \end{center}
@@
 
-[see discussion section for adding terms (disk, time evolution) to potential]
+This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the effect of chaos in the halos of such galaxies.
+Real galactic potentials will have a significant component due to the disk and bulge and may have twists (cite), radially changing axis ratios (cite), significant substructure (cite), or time dependence (either from bulk rotation, mass growth, mergers, etc.; cite).
+We expect all of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for more [...] and a few simple tests that demonstrate this expectation.
 
@@
 
-[Use a blah blah integrator...]
+We use the Dormand-Prince 8th-order Runge-Kutta scheme (cite Dormand 1981) to integrate orbits in the above potential.
+In particular, we use the Python wrapper to the FORTRAN routines provided in the Scipy package (cite?).
+For all orbits, we ensure that energy is conserved to $\Delta E/E_0 \leq 10^{-9}$ by the end of integration.
+Unless otherwise specified, orbits are integrated for $\sim$200 orbital periods with 500 steps per orbital period.
+Some box orbits require more internal integration steps in order to conserve energy to the required level.
 
@@
 
+Real disk will be self-consistent with halo, so this isn't really a fair test, but still...
+
 \acknowledgements

### 2015-02-08 ms.tex
added a bunch of text for numerical methods

@@
 \newcommand{\ident}{\mathbb{1}}
+\newcommand{\inttime}{t_{\rm int}}
 
@@
 
+[ Need to talk about how it is, in general, hard to find transformation to angle-action coords ]
+
 \section{Numerical methods}
@@
 
-Bounded, regular orbits in triaxial potentials will have three \emph{fundamental frequencies}, $\bs{\Omega}$, that determine the periodic behavior of motion (Eq.~\ref{eq:aafreq}).
+[Compare Lyapunov time to freq. diffusion time?]
+
+Bounded, regular orbits in a triaxial potential will have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion (Eq.~\ref{eq:aafreq}).
+This implies that for any canonical coordinate, $q$, the motion can be decomposed as
+\begin{equation}
+	q(t) = \sum^\infty_k a_k e^{i \omega_k t}
+\end{equation}
+where these frequencies, $\omega_k$ are linear, integer combinations of the fundamental frequencies
+\begin{equation}
+	\omega_k = \bs{n}_k \cdot \bs{\Omega}
+\end{equation}
+and the $a_k$ are complex amplitudes.
+\cite{laskar??} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hanning-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
+It has been shown that using the Hanning filter makes the accuracy in determining the frequencies scale as $\inttime^{-4}$ \citep{laskar98}, where $\inttime$ is the integration time.
+We follow \cite{valluri98} and refer to this method as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) even though it differs slightly from the original method described in (cite Laskar).
+
+If an orbit is chaotic, the motion can no longer be expressed in terms of a single set of fundamental frequencies because the actions and frequencies of the orbit change with time.
+For a weakly chaotic orbit, the orbit may act consistently periodic over windows of time.
+NAFF will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change character (possibly) smoothly with time.
+For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically over time.
+The frequencies picked out by NAFF for such orbits will therefore represent the average periodic nature of the orbit, and these will dramatically if comparing two consecutive integration periods.
+Therefore, the frequency diffusion rate per orbit, $\mathcal{R}_\Omega$, is a good predictor for how chaotic an orbit is:
+\begin{equation}
+	\mathcal{R}_\Omega = \max_{\bs{a}}\,2(\bs{\Omega}_{T_1} - \bs{\Omega}_{T_2}) / \inttime
+\end{equation}
+where $\bs{\Omega}_{T_1}$ and $\bs{\Omega}_{T_2}$ are the fundamental frequencies computed from two consecutive sections of orbit, $\inttime$ is the total integration time (over both sections), and $\bs{a}$ are the amplitudes corresponding to the frequencies.
+The frequency diffusion rate measured in the frequency with the largest amplitude is better correlated with the strength of chaos in an orbit relative to, e.g., the mean or maximum rate (cite valluri recent work?).
+
+We have found that for ``sticky'' or mildly chaotic orbits, reliably computing the Lyapunov exponent can require integration times equivalent to thousands of orbital periods.
+In comparison, owing to the precision of NAFF, detecting small frequency diffusion rates can be done with only hundreds of orbital periods.
+[example?
+comparison?]
+
+NAFF has been applied to many problems in Solar System and Galactic dynamics, especially in studying the nature of orbits in triaxial systems.
+For more details about the method, see, for example, (cite ...).
+The original procedure was written in FORTRAN, but we have implemented and tested our own implementation in Python.
+This code is publicly available through the Gary package.\footnote{link to repo}
 
 [Figure: reproduce some result?]
-
 or
@@
 \subsection{Particle balls}
+
+Our goal is to characterize the morphology of tidal debris along orbits generic to triaxial potentials.
+For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{pearson15, fardal??} [TODO: should I put sarahs paper here even though we didn't say this explicitly?], however, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depends on the degree of stochasticity.
+Further, quantifying the degree of stochasticity with the frequency diffusion rate and Lyapunov exponent represent local measures of chaos and it is unclear how this generalizes to the small but finite spreads in orbital characteristics common to debris that is tidally disrupted from globular clusters or dwarf satellite galaxies.
+To answer this question requires a massive exploration of parameter space -- both in potential and in orbits -- and thus would require a prohibitive number of N-body simulations to truly analyze the disruption process in all situations.
+
+We therefore take a simplified approach that does reproduce [...] N-body simulations for several test cases.
+For a given set of orbital initial conditions, we find the coordinates at the nearest pericenter, initialize a small ``ball'' of test particle orbits around this position, then follow the orbits of all of these test particles for some time.
+For a given mass scale, the ball particle position and velocity offsets are assumed to be Normal distributed away from the parent orbit with spreads set by the tidal radius in position and velocity scale in velocity.
+If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and is drawn from
+\begin{align}
+	\delta\bs{x}_i &\sim \mathcal{N}(0, r_{\rm tide})\\
+	\delta\bs{v}_i &\sim \mathcal{N}(0, \sigma_v)\\
+	r_{\rm tide} &=|\bs{x}_0| \left(\frac{m}{3M_{\rm enc}(|\bs{x}_0|)}\right)^{1/3} \\
+	\sigma_v &= |\bs{v}_0|\left(\frac{m}{3M_{\rm enc}(|\bs{x}_0|)}\right)^{1/3} 
+\end{align}
+where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, and $m$ is the mass scale of the debris.
+
+For this work, we take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass to estimate the above scales.
 

### 2015-02-11 ms.tex
paper updates

@@
 For computational efficiency, many authors instead express the triaxiality in the form of the potential, however this can lead to unphysical situations where the density becomes negative.
-\citet{lee03} derive a perturbative expansion of the potential integral and show that the expansion is accurate even for modest axis ratios (e.g., the median values shown above).
+\citet{leesuto03} derive a perturbative expansion of the potential integral and show that the expansion is accurate even for modest axis ratios (e.g., the median values shown above).
 
-For this work, we use the triaxial potential expression from \citet{lee03}, parametrized in a slightly different manner.
+For this work, we use the triaxial potential expression from \citet{leesuto03}, parametrized in a slightly different manner.
 In terms of spherical coordinates\footnote{(radius, azimuth, colatitude)} with the radius normalized by the scale radius, $u = r/r_s$
@@
 where $v_c$ is the circular velocity at the scale radius, $r_s$, along the $x$-axis, $e_b = \sqrt{1 - (b/a)^2}$, and $e_c = \sqrt{1 - (c/a)^2}$.
-The functions $F_i(u)$ are given in the appendix of \cite{lee03}.
+The functions $F_i(u)$ are given in the appendix of \cite{leesuto03}.
 We chose $r_s=20~{\rm kpc}$ and $v_c = 200~{\rm km}~{\rm s}^{-1}$ to [...].
@@
 where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, and $m$ is the mass scale of the debris.
-
 For this work, we take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass to estimate the above scales.
@@
 Like only considering debris that comes off at a single pericentric passage
+
+We follow \citet[][hereafter MV96]{merritt96} and quantify the degree of mixing by computing the ``distance'' between the configuration space density of the ensemble of particles at a given time, $\rho(\bs{x},t)$, with that of a distribution that is uniform over the energy hypersurface, $\rho_{E_0}(\bs{x})$, where $E_0$ is the energy of the parent orbit and hypersurface of interest.
+Our method differs in a few ways from that of MV96.
+First, MV96 use isoenergy ensembles of particles for following the density because that is the only way to guarantee that, if chaotic, nearly all orbits in the ensemble will lie on chaotic orbits.
+We instead use ensembles with small ($\approx$1\%) spreads in energy, but we still expect that such an ensemble around a parent chaotic orbit will reach a fully mixed state.
+The final distribution, rather than fill the energy hypersurface, will instead have some thickness around this surface.
+Around a typical regular orbit, each individual orbit will fill its torus ergodically -- that is, a surface with lower dimension than the energy hypersurface -- and thus, a small enough ensemble will never get ``close'' to the fully mixed state as defined above because it will remain confined into a subspace on the energy surface.
+In practice, only strongly chaotic systems will approach $\rho_{E_0}(\bs{x})$; the rate will depend on how stochastic the parent orbit is (e.g., as measured by the Lyapunov time or frequency diffusion rate), and also on how ``trapped'' the chaotic region is (a small chaotic region wedged between stable tori will take longer to appear chaotic, even if internally the frequency diffusion rate is large).
+However the difference in configuration space density between an ensemble undergoing chaotic mixing versus an ensemble around a regular orbit may be detectable over a much shorter timescale.
+The fully-mixed, normalized, configuration-space density that fills a given energy hypersurface, $E_0$, is
+\begin{align}
+	\rho_{E_0}(\bs{x}) &= \int \dd^3v \, \rho_0 \, \delta(H - E_0)\\
+	&= \int \dd^3v \, \rho_0 \, \delta\left(|\bs{v}|^2/2 + \Phi(\bs{x}) - E_0\right)\\
+	&= 4\pi \int \dd v \, v^2 \, \rho_0 \, \delta\left(v^2/2 + \Phi(\bs{x}) - E_0\right)\\
+	&= A \sqrt{E_0 - \Phi(\bs{x})}
+\end{align}
+where $A$ is a constant determined such that $\int \dd^3\bs{x} \, \rho_{E_0}(\bs{x}) = 1$, and $\delta(\cdot)$ is the Dirac delta function.
+
+Second, we compute the Kullback-Leibler divergence \citep[KL divergence;][]{kullback51} by estimating the configuration space density $\rho(\bs{x},t_j)$ using kernel density estimation (KDE) with the ensemble of particle positions at time $t_j$, then use Monte Carlo integration to evaluate the KL divergence rather than evaluate the integral over a three dimensional grid.
+The KL divergence, $D_{\rm KL}$, is a measure of the relative Shannon entropy between two normalized density functions, $p(x)$ and $q(x)$,
+\begin{equation}
+	D_{\rm KL} = \int \dd x \, p(x) \ln \frac{p(x)}{q(x)}
+\end{equation}
+We would like to compute 
+\begin{align}
+	D_{\rm KL}(t_j) &= \int \dd^3 \bs{x} \, \rho(\bs{x},t_j) \ln \frac{\rho(\bs{x},t_j)}{\rho_{E_0}(\bs{x})}
+\end{align}
+however, we only have an estimate of the density $\rho_{\rm KDE}(\bs{x},t_j) \approx \rho(\bs{x},t_j)$. Using samples from this KDE estimated density, $\bs{x}_i$, the KL divergence is
+\begin{align}
+	D_{\rm KL}(t_j) &\approx \frac{1}{N} \sum_i \ln \frac{\rho_{\rm KDE}(\bs{x}_i,t_j)}{\rho_{E_0}(\bs{x}_i)}\\
+	\bs{x}_i &\sim \rho_{\rm KDE}(\bs{x},t_j)\label{eq:sampled_from}
+\end{align}
+where Equation~\ref{eq:sampled_from} indicates that the $\bs{x}_i$ are samples from the KDE density.
+For any computation of the density estimate, we use 10-fold cross-validation to find the optimal kernel bandwidth, $\sigma_{\rm KDE}$, by maximizing the likelihood of withheld particles.
+Figure~\ref{fig:DKL} shows $D_{\rm KL}$ as a function of time for a regular (black) and chaotic (blue) orbit.
+The chaotic orbit ensemble mixes much faster than the regular orbit ensemble.
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=0.75\textwidth]{figures/D_KL.pdf}
+\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:DKL}
+\end{center}
+\end{figure*}
 
@@
 \acknowledgements
+The authors wish to thank ...
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 
 This research made use of Astropy, a community-developed core \texttt{Python} package for Astronomy \citep{astropy13}.
-This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance. \\
+This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance.
 

### 2015-02-11 ms.tex
write write write

@@
 \renewcommand\algorithmicdo{}
+\usepackage{mathrsfs}
 
@@
 %\newcommand{\jac}{\bs{\rm J}}
-\newcommand{\jac}{\bs{G}}
+%\newcommand{\jac}{\bs{G}}
+\newcommand{\jac}{\mathscr{J}}
 
@@
 
-For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (d.o.f.), the equations of motion are most simply expressed in angle-action coordinates.
+\hrule
+
+For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (dof), the equations of motion are most simply expressed in angle-action coordinates.
 In these coordinates, the actions, $J_n$, are integrals of the motion, and the conjugate angle variables, $\theta_n$, increase linearly with time
@@
 \end{align}
-where $n=1,2,...N$, and $\Omega_n$ are the fundamental frequencies of the orbit.
-Owing to the involutive property of the actions and the periodic nature of the angle variables, the topology of angle-action space is equivalent to that of the surface of an $N$-dimensional torus, and thus orbits are often referred to in terms of \emph{orbital tori}; each set of actions, $(\act_1,\act_2,...\act_N)$, uniquely determine the torus associated with a given orbit.
+where $i=1,2,...N$, and $\Omega_n$ are the fundamental frequencies of the orbit.
+Owing to the involutive property of the actions and the periodic nature of the angle variables, the topology of angle-action space is equivalent to that of the surface of an $N$-dimensional torus, and thus orbits are often referred to in terms of \emph{orbital tori}: each set of actions, $(\act_1,\act_2,...\act_N)$, uniquely determine the torus associated with a given orbit and the angles specify the location on the surface of the $N$-torus.
 
-If galactic halos are triaxial, their gravitational potentials will almost certainly not be globally integrable.
-For a given halo, if the potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
+If, to first order, galactic halos are triaxial, galactic gravitational potentials will almost certainly not be globally integrable.
+If a triaxial potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
 \begin{equation}
@@
 \end{equation}
-then a large number of regular orbits will survive the perturbation (e.g., the KAM theorem; cite Arnold)\footnote{3-vectors are represented with bold symbols}.
+then a large number of regular orbits will survive the perturbation (by way of the KAM theorem)\footnote{3-vectors are represented with bold symbols, e.g., $\bs{J}=(J_1,J_2,J_3)$.}.
 The tori that survive will in general be separated by regions of irregular or chaotic motion, and thus any transformations to angle-action coordinates can only be defined locally.
-The types of motion generic to typical triaxial systems can be split into several sub-categories (summarized in Figure~\ref{fig:orbit-tree}).
+
+The types of motion generic to triaxial systems can be split into several sub-categories (summarized in Figure~\ref{fig:orbit-tree}).
 Regular motion can be classified by the number of resonance relations obeyed \citep[e.g.,][]{lichtenberg83, valluri98}: \emph{conditionally periodic} orbits obey no resonance relations, \emph{uni-resonant} orbits obey a single relation of the form $\bs{m}\cdot\bs{\Omega}=0$, and \emph{bi-resonant} orbits obey two resonance relations, $\bs{m}\cdot\bs{\Omega}=0$ and $\bs{n}\cdot\bs{\Omega}=0$, where $\bs{m}$ and $\bs{n}$ are integer vectors.
@@
 
-Resonant orbits are dynamically important in systems with several degrees of freedom as they determine the structure of orbit-space.
-Stable resonant orbits generate families of orbits that behave or appear similarly, whereas unstable resonances typically generate regions of stochasticity.
-Not all orbits can be trivially associated with resonances, however, due to the infinitely complex, hierarchical structure of stable resonances.
-In autonomous, two d.o.f.
-systems, stochastic regions are generally surrounded and separated by regular motion, but in three or more d.o.f.
-the chaotic regions will connect and overlap; this is the origin of the phenomenon of Arnold diffusion \citep{arnold79}.
-The implication of Arnold diffusion is that chaotic orbits may radically change shape by means of traversing the web of unstable resonances and are strictly only bounded to their own energy hypersurface.
-Another way to understand this phenomenon is by means of orbital tori: in two dimensions, if a given orbital torus is destroyed due to perturbations, the orbit is still bounded between the surrounding tori due to energy conservation.
-In three dimensions, the extra degree of freedom allows chaotic orbits to ``escape'' confinement.
+Resonant orbits are rare \citep{merritt99} but dynamically important in systems with several degrees of freedom as they determine the structure of orbit-space.
+Stable, resonant orbits generate families of orbits that may behave or appear similarly, whereas unstable resonances are typically associated with regions of chaos.
+However not all orbits can be trivially associated with resonances due to the infinitely complex, hierarchical structure of resonances (though the importance or ``size'' of a resonance falls off inversely with the norm of its integer vector, $|\bs{m}|^{-\alpha}$, with $\alpha > 0$).
+In autonomous, two dof systems, stochastic regions are generally surrounded and separated by regular motion, thus chaotic orbits are confined.
+In three or more dof the chaotic regions will connect and overlap.
+This is the origin of Arnold diffusion \citep{arnold64}, which allows chaotic orbits to radically change shape by traversing the web of unstable resonances and are only strictly bounded to their own energy hypersurface.
+This phenomenon is perhaps easier to visualize in action-space: in two dof systems, if a given set of tori are destroyed due to potential perturbations, an orbit in this now chaotic region is still bounded between the surrounding tori due to energy conservation.
+In three dimensions, the extra degree of freedom allows chaotic orbits to ``escape'' this confinement.
 
-For small perturbations, generically, many regular orbits survive and only small chaotic islands are introduced.
-As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed, then the uni-resonant tori are destroyed, and finally the bi-resonant tori -- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.
-The irregular or chaotic orbits that fill the parts of action-space where tori have been destroyed are loosely classified by being either \emph{strongly} or \emph{weakly} chaotic.
-Qualitatively, strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale.
-There are many methods for classifying and measuring the degree of stochasticity of an orbit.
+For small potential perturbations, many regular orbits survive and only small chaotic islands are introduced.
+As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed.
+Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori -- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.
+However, even orbits that are strictly chaotic may behave regularly over timescales much longer than, e.g., the dynamical time of the system.
+Chaotic can therefore be loosely classified by being either \emph{strongly} or \emph{weakly} chaotic: strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale (we quantify this distinction in Section~\ref{sec:}).
+This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{}) or frequency diffusion time (Section~\ref{}), and to the mixing rate of small ensembles of orbits (Section~\ref{}).
 
-[ Need to talk about how it is, in general, hard to find transformation to angle-action coords ]
+[summarizing paragraph connecting everything, mention mixing, relation to tidal streams]
 
@@
 
-The density distributions within dark matter halos formed in cosmological N-body simulations are typically triaxial (e.g., millenium, aquarius, via lactea).
+The density distributions within dark matter halos formed in cosmological N-body simulations are typically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}.
 \citet{jing02} found that a triaxial generalization of the classic NFW density profile \citep{navarro96} generates excellent fits to their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of the many halos formed in their simulations.
@@
 For computational efficiency, many authors instead express the triaxiality in the form of the potential, however this can lead to unphysical situations where the density becomes negative.
-\citet{leesuto03} derive a perturbative expansion of the potential integral and show that the expansion is accurate even for modest axis ratios (e.g., the median values shown above).
+\citet{leesuto03} derive a perturbative expansion of the potential integral for a triaxial NFW density and show that the expansion is accurate even for modest axis ratios (e.g., the median values shown above).
 
 For this work, we use the triaxial potential expression from \citet{leesuto03}, parametrized in a slightly different manner.
-In terms of spherical coordinates\footnote{(radius, azimuth, colatitude)} with the radius normalized by the scale radius, $u = r/r_s$
+In terms of spherical coordinates\footnote{$(r,\phi,\theta)$ = (radius, azimuth, colatitude)} with the radius normalized by the scale radius, $u = r/r_s$
 \begin{align}
@@
 The functions $F_i(u)$ are given in the appendix of \cite{leesuto03}.
-We chose $r_s=20~{\rm kpc}$ and $v_c = 200~{\rm km}~{\rm s}^{-1}$ to [...].
+We chose $r_s=20~{\rm kpc}$ and $v_c = 200~{\rm km}~{\rm s}^{-1}$ as arbitrary choices but motivated by the halo concentration distributions found by \cite{jing02}.
 Figure~\ref{fig:potential} shows the rotation curve of this potential along the three principal axes.
@@
 This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the effect of chaos in the halos of such galaxies.
-Real galactic potentials will have a significant component due to the disk and bulge and may have twists (cite), radially changing axis ratios (cite), significant substructure (cite), or time dependence (either from bulk rotation, mass growth, mergers, etc.; cite).
-We expect all of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for more [...] and a few simple tests that demonstrate this expectation.
+Real galactic potentials will have a significant component due to the disk and bulge and may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{veraciro11}, significant substructure \citep{zemp09}, or time dependence (either from bulk rotation, mass growth, mergers, etc.; cite).
+We expect all of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple examples that demonstrate this expectation.
 
@@
 We use the Dormand-Prince 8th-order Runge-Kutta scheme (cite Dormand 1981) to integrate orbits in the above potential.
-In particular, we use the Python wrapper to the FORTRAN routines provided in the Scipy package (cite?).
+In particular, we use the Python wrapper to the FORTRAN routines provided in the Scipy package \citep{scipy}.
 For all orbits, we ensure that energy is conserved to $\Delta E/E_0 \leq 10^{-9}$ by the end of integration.
@@
 
-[Compare Lyapunov time to freq. diffusion time?]
-
 Bounded, regular orbits in a triaxial potential will have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion (Eq.~\ref{eq:aafreq}).
-This implies that for any canonical coordinate, $q$, the motion can be decomposed as
+This implies that for any canonical coordinate, $q$, the motion in this coordinate can be decomposed as
 \begin{equation}
@@
 \end{equation}
-where these frequencies, $\omega_k$ are linear, integer combinations of the fundamental frequencies
+where these frequencies, $\omega_k$, are linear, integer combinations of the fundamental frequencies
 \begin{equation}
@@
 and the $a_k$ are complex amplitudes.
-\cite{laskar??} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hanning-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
-It has been shown that using the Hanning filter makes the accuracy in determining the frequencies scale as $\inttime^{-4}$ \citep{laskar98}, where $\inttime$ is the integration time.
-We follow \cite{valluri98} and refer to this method as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) even though it differs slightly from the original method described in (cite Laskar).
+\cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hanning-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
+It has been shown that using the Hanning filter makes the accuracy in determining the frequencies scale as $\inttime^{-4}$ \citep{laskar99}, where $\inttime$ is the integration time.
+We follow \cite{valluri98} and refer to this method as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) even though it differs slightly from the original method from Laskar.
 
 If an orbit is chaotic, the motion can no longer be expressed in terms of a single set of fundamental frequencies because the actions and frequencies of the orbit change with time.
-For a weakly chaotic orbit, the orbit may act consistently periodic over windows of time.
+For a weakly chaotic orbit, the orbit may act consistently periodic over long windows of time.
 NAFF will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change character (possibly) smoothly with time.
-For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically over time.
-The frequencies picked out by NAFF for such orbits will therefore represent the average periodic nature of the orbit, and these will dramatically if comparing two consecutive integration periods.
+For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two separate sections of orbit.
+The frequencies picked out by NAFF for such orbits will therefore represent the average periodic nature of the orbit over a given period, and these will vary dramatically if comparing two consecutive integration periods.
 Therefore, the frequency diffusion rate per orbit, $\mathcal{R}_\Omega$, is a good predictor for how chaotic an orbit is:
@@
 
-NAFF has been applied to many problems in Solar System and Galactic dynamics, especially in studying the nature of orbits in triaxial systems.
-For more details about the method, see, for example, (cite ...).
-The original procedure was written in FORTRAN, but we have implemented and tested our own implementation in Python.
-This code is publicly available through the Gary package.\footnote{link to repo}
+NAFF has been applied to many problems in Solar System and Galactic dynamics, especially in the study of of orbits in triaxial systems.
+For more details about the method, see \cite{papaphillipou, laskar, etc.}.
+We have implemented and tested our own implementation of NAFF in Python.
+This code is publicly available through the Gary package.\footnote{link to repo} Figure~\ref{fig:logfreqs} shows a test of the implementation in which we reproduce the frequency map at a fixed energy of an axisymmetric, logarithmic potential \cite[][pg.
+260, Figure~3.45]{binneytremaine}.
+Plotted are the frequency ratios [...].
+In such a map, stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
 
-[Figure: reproduce some result?]
-or
-[Figure: for same orbits as in Lyapunov section, show frequency diffusion?]
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=0.75\textwidth]{figures/log-freqmap.png}
+\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:logfreqs}
+\end{center}
+\end{figure*}
 
-\subsection{Particle balls}
+[Figure: for same orbits as in Lyapunov section, show frequency diffusion by computing frequencies in rolling windows?]
+
+\subsection{Test particle ensembles}
 
 Our goal is to characterize the morphology of tidal debris along orbits generic to triaxial potentials.
-For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{pearson15, fardal??} [TODO: should I put sarahs paper here even though we didn't say this explicitly?], however, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depends on the degree of stochasticity.
-Further, quantifying the degree of stochasticity with the frequency diffusion rate and Lyapunov exponent represent local measures of chaos and it is unclear how this generalizes to the small but finite spreads in orbital characteristics common to debris that is tidally disrupted from globular clusters or dwarf satellite galaxies.
-To answer this question requires a massive exploration of parameter space -- both in potential and in orbits -- and thus would require a prohibitive number of N-body simulations to truly analyze the disruption process in all situations.
+For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{pearson15, fardal??} [TODO: should I put sarahs paper here even though we didn't say this explicitly?], however, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
+Further, quantifying the degree of stochasticity with, for example, the frequency diffusion rate represents a \emph{local} measure of chaos and it is unclear how this generalizes to the small but finite spreads in orbital characteristics common to debris that is tidally disrupted from globular clusters or dwarf satellite galaxies.
+To answer this question would require a massive exploration of parameter space -- both in potential and in orbits -- and thus would require a prohibitive number of N-body simulations to truly analyze the disruption process in all situations.
 
-We therefore take a simplified approach that does reproduce [...] N-body simulations for several test cases.
-For a given set of orbital initial conditions, we find the coordinates at the nearest pericenter, initialize a small ``ball'' of test particle orbits around this position, then follow the orbits of all of these test particles for some time.
-For a given mass scale, the ball particle position and velocity offsets are assumed to be Normal distributed away from the parent orbit with spreads set by the tidal radius in position and velocity scale in velocity.
+We therefore take a simplified approach and consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event of a low-mass progenitor.
+For a given set of orbital initial conditions -- the parent orbit -- we find the position of the nearest pericenter, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all of these test particles for some evolution time, $T_{\rm ev}$.
+For a given mass scale, the ensemble particle position and velocity offsets are assumed to be Normal-distributed away from the parent orbit with spreads set by the tidal radius in position and velocity scale in velocity \citep[e.g.,][]{johnston98, apw14}.
 If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and is drawn from
@@
 \end{align}
-where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, and $m$ is the mass scale of the debris.
+where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, and $m$ is the mass scale of the debris, and $|\cdot |$ is the Euclidean norm.
 For this work, we take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass to estimate the above scales.
 
-Need a fast way to characterize morphology of local orbits after finite time
-
-Can't run N-body for many tens of thousands of initial conditions
-
-Explain simplistic approach
-
-[Figure: Compare N-body simulation to balls of particles -- captures the essence]
-
-Cite papers that use Lagrange-point stripping methods, Rewinder paper
-
-Like only considering debris that comes off at a single pericentric passage
-
-We follow \citet[][hereafter MV96]{merritt96} and quantify the degree of mixing by computing the ``distance'' between the configuration space density of the ensemble of particles at a given time, $\rho(\bs{x},t)$, with that of a distribution that is uniform over the energy hypersurface, $\rho_{E_0}(\bs{x})$, where $E_0$ is the energy of the parent orbit and hypersurface of interest.
+As the particle ensembles evolve in time, we follow \citet[][hereafter MV96]{merritt96} and quantify the degree of mixing of the ensemble by computing the ``distance'' between the configuration space density of the ensemble at a given time, $\rho(\bs{x},t)$, with that of a distribution that is uniform over the energy hypersurface, $\rho_{E_0}(\bs{x})$, where $E_0$ is the energy of the parent orbit and thus hypersurface of interest.
 Our method differs in a few ways from that of MV96.
@@
 
-Second, we compute the Kullback-Leibler divergence \citep[KL divergence;][]{kullback51} by estimating the configuration space density $\rho(\bs{x},t_j)$ using kernel density estimation (KDE) with the ensemble of particle positions at time $t_j$, then use Monte Carlo integration to evaluate the KL divergence rather than evaluate the integral over a three dimensional grid.
+Second, we compute the Kullback-Leibler divergence \citep[KL divergence;][]{kullback51} by estimating the configuration space density $\rho(\bs{x},t_j)$ using kernel density estimation (KDE)\footnote{We use an implementation from the Python package scikit-learn \citep{scikitlearn}.} with the ensemble of particle positions at time $t_j$, then use Monte Carlo integration to evaluate the KL divergence rather than evaluate the integral over a three dimensional grid.
 The KL divergence, $D_{\rm KL}$, is a measure of the relative Shannon entropy between two normalized density functions, $p(x)$ and $q(x)$,
@@
 \begin{align}
-	D_{\rm KL}(t_j) &\approx \frac{1}{N} \sum_i \ln \frac{\rho_{\rm KDE}(\bs{x}_i,t_j)}{\rho_{E_0}(\bs{x}_i)}\\
+	D_{\rm KL}(t_j) &\approx \frac{1}{N} \sum_i^N \ln \frac{\rho_{\rm KDE}(\bs{x}_i,t_j)}{\rho_{E_0}(\bs{x}_i)}\\
 	\bs{x}_i &\sim \rho_{\rm KDE}(\bs{x},t_j)\label{eq:sampled_from}
@@
 
-\subsubsection{Entropy evolution}
+\subsubsection{Minimum KLD map}
+Look at minimum value of $D_{\rm KL}$ over $T_{\rm ev}$, use that as metric of
 
-\subsubsection{Density evolution}
-
-Compare mean density or some other metric for the balls, show that decreases faster for chaotic orbit -- chaotic mixing dominates over phase-mixing
-
-Power law (Helmi and White)
-
-\subsubsection{Adding a disk?}
+\subsection{Adding a disk?}
 

### 2015-02-13 ms.tex
clarify bt08 plot
added writing to experiments

@@
 \usepackage{mathrsfs}
+\usepackage{subfigure}
 
@@
 \newcommand{\inttime}{t_{\rm int}}
+\newcommand{\fdrate}{\mathcal{R}_\Omega}
 
@@
 \end{align}
-where $v_c$ is the circular velocity at the scale radius, $r_s$, along the $x$-axis, $e_b = \sqrt{1 - (b/a)^2}$, and $e_c = \sqrt{1 - (c/a)^2}$.
+where $e_b = \sqrt{1 - (b/a)^2}$, $e_c = \sqrt{1 - (c/a)^2}$, and $v_c$ is the circular velocity at the scale radius, $r_s$, for the spherical case.
 The functions $F_i(u)$ are given in the appendix of \cite{leesuto03}.
-We chose $r_s=20~{\rm kpc}$ and $v_c = 200~{\rm km}~{\rm s}^{-1}$ as arbitrary choices but motivated by the halo concentration distributions found by \cite{jing02}.
+We chose $r_s=40~{\rm kpc}$ and $v_c = 150~{\rm km}~{\rm s}^{-1}$ by taking the mean halo concentration found by \cite{jing02} for a ${\rm M}_{vir} \approx 10^{12}~\msun$ halo, $c_e\approx5$, and assuming $R_{vir}\approx200~{\rm kpc}$.
 Figure~\ref{fig:potential} shows the rotation curve of this potential along the three principal axes.
@@
 
-\subsection{Orbit integration}
+\subsection{Orbit integration}\label{sec:integration}
 
@@
 The frequencies picked out by NAFF for such orbits will therefore represent the average periodic nature of the orbit over a given period, and these will vary dramatically if comparing two consecutive integration periods.
-Therefore, the frequency diffusion rate per orbit, $\mathcal{R}_\Omega$, is a good predictor for how chaotic an orbit is:
+Therefore, the frequency diffusion rate per orbit, $\fdrate$, is a good predictor for how chaotic an orbit is:
 \begin{equation}
-	\mathcal{R}_\Omega = \max_{\bs{a}}\,2(\bs{\Omega}_{T_1} - \bs{\Omega}_{T_2}) / \inttime
+	\fdrate = \max_{\bs{a}}\,2(\bs{\Omega}_{T_1} - \bs{\Omega}_{T_2}) / \inttime \label{eq:fdrate}
 \end{equation}
@@
 260, Figure~3.45]{binneytremaine}.
-Plotted are the frequency ratios [...].
+Plotted are the (Cartesian) frequency ratios recovered for a grid of iso-energy, box orbits integrated in the potential $\Phi(x,y,z) = \frac{1}{2}\ln\left(x^2 + (y/0.9)^2 + (z/0.7)^2 + 0.1\right)$.
 In such a map, stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
@@
 
-\subsection{Frequency map}
+\subsection{Orbital properties: frequencies and frequency diffusion}
 
-We generate frequency maps at X energies for tube orbits started around the major and intermediate axes of the potential (figures)
+We start by studying the properties of isoenergy orbits in the potential described above.
+We generate grids of initial condition along the $xz$ ($y=0$) and $yz$ ($x=0$) planes\footnote{Recall that $x$ is the major, $y$ the intermediate, and $z$ the minor axis.} all with the same energy, $E=XXX$.
+For the $xz$ orbits, we fix $v_x = v_z = 0$, and compute $v_y$ from the energy.
+Similarly, for the $yz$ orbits, we fix $v_y = v_z = 0$, and compute $v_x$ from the energy.
+We integrate the orbits as described in Section~\ref{sec:integration} and compute the fundamental frequencies of each orbit with NAFF.
+Figure~\ref{fig:freqmap} shows frequency maps for the $xz$ (left) and $yz$ (right) grids.
+For short-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $z$ axis, $\Omega_\phi$, and the oscillation along the $z$ axis, $\Omega_z$.
+For long-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $x$ axis, $\Omega_\phi$, and the oscillation along the $x$ axis, $\Omega_x$.
+Box orbits are not shown.
+Already evident in these panels is that many of the orbits are regular: the transformed, even-spaced sampling of the initial conditions is preserved in the frequency transformation, implying that the frequencies for many orbits do not change appreciably.
+Also apparent is resonant trapping,  which appear as over-dense, linear features in these frequency maps.
 
-Explain qualitative differences between the two -- e.g., Y-Z plane will have many more stochastic tubes because intermediate axis
+The majority of the orbits generated on these grids are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic.
+Tube orbits are generally less stochastic than box orbits -- the other major class of orbits in triaxial potentials -- which tend to plunge deep into the inner regions of the potential.
+Thin tidal streams are thought to form along tube orbits rather than box orbits because of the fast disruption expected for stellar systems on radially plunging orbits.
+The major classes of stable tube orbits circulate about either the major or minor axis; the most prominent orbits are the short-axis and outer long-axis tubes, but these grids do generate some inner long-axis tubes and box orbits.
+Orbits that try to circulate about the intermediate axis tend to be highly irregular and do not preserve their sense of circulation for long periods of time.
+NAFF recovers fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL95][]{papaphillippou96}.
+PL96 show that a good choice of coordinates for tube orbits are the Poincar\'e symplectic polar coordinates, a set of canonical coordinates similar to cylindrical coordinates.
+For approximately XX\% of the orbits, NAFF fails to recover the frequencies (similar difficulties were reported in \cite{valluri98} and \cite{papaphillippou96} for recovering certain tube orbit frequencies).
 
-\subsection{Morphology}
+\begin{figure}[h]
+\centering
+	\includegraphics[width=0.45\columnwidth] {figures/freqmap_xz.png}
+	\includegraphics[width=0.45\columnwidth] {figures/freqmap_yz.png}
+	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+	\label{fig:freqmap}
+\end{figure}
 
-\subsubsection{Minimum KLD map}
+For each orbit, we also compute the frequency diffusion rate per orbit, $\fdrate$ (Equation~\ref{eq:fdrate}) by computing the frequencies for the first half and second half of the orbital time series separately.
+This rate is only a measure of the time-averaged diffusion rate because the frequencies are computed for long segments of orbit, but \cite{valluri??} [...].
+Figure~\ref{fig:diffusionmap} shows maps of the frequency diffusion for the $xz$ (left) and $yz$ (right) grids: plotted are the initial conditions for all orbits colored by $\log_{10}(\fdrate)$ (white points are strongly chaotic, black are regular).
+The rich structure of unstable resonances apparent in these frequency diffusion maps highlight the accuracy of NAFF.
+It seems that tube orbits in these potentials are mostly regular or only mildly chaotic, however islands of stronger chaos do appear, especially at the intersections of unstable resonances.
+The bright, unstable regions in these maps are a visualization of the Arnold web at this energy.
+It is evident that these unstable regions are interconnected and therefore a tube orbit that starts, for example, in the lower left-hand region of one of these maps could  theoretically end up as a box orbit (along the zero-velocity curve) at some point in it its history (but the timescale for this to happen is likely prohibitively long).
+
+\begin{figure}[h]
+\centering
+	\includegraphics[width=0.45\columnwidth] {figures/diffusion_map_xz.png}
+	\includegraphics[width=0.45\columnwidth] {figures/diffusion_map_yz.png}
+	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+	\label{fig:diffusionmap}
+\end{figure}
+
+Figure~\ref{fig:xx} shows a sample of regular orbits from these two grids, plotted in projection.
+The left column are orbits from the $xz$ grid and the right column shows orbits from the $yz$ grid.
+Orbits near a resonance are labeled with the integer vector associated with the nearest resonance.
+Figures~\ref{fig:chaotic1} and \ref{fig:chaotic2} show two chaotic orbits, one with a large $\fdrate$ and one with a small $\fdrate$ (labeled in the figure titles).
+The figures show projections of the orbit in three consecutive sections of orbit with $\sim$50 periods shown per panel.
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/chaotic_orbit1.png}
+\caption{Just a placeholder figure that has a similar shape to what I will put here} \label{fig:chaotic1}
+\end{center}
+\end{figure*}
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/chaotic_orbit2.png}
+\caption{Just a placeholder figure that has a similar shape to what I will put here} \label{fig:chaotic2}
+\end{center}
+\end{figure*}
+
+\subsection{Ensemble properties: mixing of nearby orbits}
+
 Look at minimum value of $D_{\rm KL}$ over $T_{\rm ev}$, use that as metric of

### 2015-02-14 ms.tex
intro noise

@@
 
-Adding third degree of freedom opens up many new orbits / resonances / etc. -- Valluri and Merritt work...
+On the largest scales ($\gtrsim$30~kpc for a Milky-Way-like galaxy) in the halos of galaxies, the gravitational potentials are thought to be triaxial.
+However, despite suggestive evidence, this fundamental prediction from LCDM has not been conclusively confirmed from observations.
+Around other galaxies, it is generally hard to measure the 3D shape of [..] because they are seen in projection.
+From our position within the Milky Way, our 3D view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii.
+Unfortunately, the halo also has a low density of tracers.
+Luckily, however, many of the halo stars are likely associated with debris stripped from in-falling satellite galaxies.
 
-The halos of galaxies are filled with substructure...
-As a satellite galaxy or globular cluster orbits within a parent galaxy, its mass is eroded due to the tidal forces of the host galaxy.
+As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded due to the tidal forces of the host galaxy potential.
+The phase-space density and therefore the observable properties of the tidal debris are sensitive to the internal progenitor properties, orbit of the progenitor, and underlying mass distribution of the host galaxy.
+The morphological evolution of the debris depends on the spread of orbital properties (e.g., actions) of the debris and the orbit of the progenitor system, both of which are determined by the shape and radial profile of the gravitational potential of the host galaxy.
+In the halo of the Milky Way, tidal streams of stars are found over a large range of distances.
+By simultaneously modeling the observed density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution of the Milky Way on these large scales.
+
+Tidal streams are generally modeled as 1D structures with simple internal progenitor dynamics and density evolution. 
+Pearson et al.
+(2014) tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in an oblate and a triaxial potential using \emph{Streakline} models (K\"upper et al.
+(XXXX)) :
+		\item in the oblate potential, a thin model stream is easily found that reproduces the observed stellar density morphology of the stream;
+		\item in the triaxial potential, the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness of the stream, given the observational constraints of the present-day position and velocity of the cluster.
+This result suggests that the shape of the host galaxy potential directly influences the morphology of the resulting debris, and that other morphological classes exist between thin, 1D streams and diffuse, 2D shells.
+To understand the formation of these fans and the implications of this work we must first understand the complex orbital structure of triaxial systems.
+
+In this work, we study the effect of orbital classes that are prevalent in triaxial systems (e.g., stochastic, near-resonant, resonant, and quasiperiodic) on the resulting morphology of tidal streams.
+
 For (typically) many orbits, these tidally stripped stars remain spatially coherent.
@@
 
-\hrule
+Adding third degree of freedom opens up many new orbits / resonances / etc. -- Valluri and Merritt work...
+
+The halos of galaxies are filled with substructure...
 
@@
 	\theta_n(t) &= \Omega_n t + \theta_n(0)\\
-	\dot{\theta}_n &= \frac{\partial H}{\partial \act_n} = \Omega_n(\act_1,...,\act_n)\label{eq:aafreq}
+	\dot{\theta}_n &= \frac{\partial H}{\partial \act_n} = \Omega_n(\act_1,\act_2,...,\act_n)\label{eq:aafreq}
 \end{align}
@@
 \end{equation}
-and the $a_k$ are complex amplitudes.
+and the $a_k$ are complex amplitudes and $\bs{n}_k$ is an integer vector.
 \cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hanning-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
@@
 
-\subsection{Orbital properties: frequencies and frequency diffusion}
+\subsection{Orbital properties: frequencies, frequency diffusion, and Lyapunov exponents}
 
@@
 It is evident that these unstable regions are interconnected and therefore a tube orbit that starts, for example, in the lower left-hand region of one of these maps could  theoretically end up as a box orbit (along the zero-velocity curve) at some point in it its history (but the timescale for this to happen is likely prohibitively long).
+Figure~\ref{fig:diffusionrates} shows histograms of $\log_{10}(\fdrate)$ for all orbits in each initial condition grid for box and tube orbits separately.
+[...1-2 sentences discussion this (need to wait to see final figure)]
 
@@
 \end{figure}
+
+\begin{figure}[h]
+\centering
+	\includegraphics[width=0.45\columnwidth] {figures/diffusion_rate_hist_xz.pdf}
+	\includegraphics[width=0.45\columnwidth] {figures/diffusion_rate_hist_yz.pdf}
+	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc. Separate box from tube orbits.} 
+	\label{fig:diffusionrates}
+\end{figure}
+
+[Lyapunov exponent grids...]
 
@@
 
-\subsection{Ensemble properties: mixing of nearby orbits}
+\subsection{Ensemble properties: mixing of debris}
 
-Look at minimum value of $D_{\rm KL}$ over $T_{\rm ev}$, use that as metric of
+We measure the state of mixing of the debris as described above: we compute the KL divergence between the configuration-space density of the ensemble and the configuration-space density expected if the ensemble particles were uniformly distributed over the parent orbit energy hypersurface.
+
+a:
+This measure can be computed at any time -- [we show two different maps of this quantity: one after 10 Gyr, one after 15 orbits?]
+
+b: 
+From the KL divergence, we compute a mixing time by measuring the slope or power law index?
+
+[Measure mixing time from KLD(t), plot freq. diffusion time vs. mixing time, show we can predict morphology from orbital properties of progenitor?]
+
+\subsection{Observable consequences of mixing}
+
+Only effectively looking at debris disrupted from a single pericenter -- debris near progenitor constantly repopulated as more and more stuff disrupts out
+
+Streams should be shorter? How to test this? Length of stream in angle at isodensity contour?
 
@@
 \section{Discussion}\label{sec:discussion}
+
+Recovering the potential of triaxial systems is not trivial. Danger of orbit fitting because chaotic orbits can reproduce morphologies of regular orbits.
 
@@
 \acknowledgements
-The authors wish to thank ...
+The authors wish to thank Sarah Pearson, Andreas K\"upper, David Hogg, David Merritt, \emph{Stream Team}, etc...
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 

### 2015-02-14 ms.tex
more intro work

@@
 
-On the largest scales ($\gtrsim$30~kpc for a Milky-Way-like galaxy) in the halos of galaxies, the gravitational potentials are thought to be triaxial.
-However, despite suggestive evidence, this fundamental prediction from LCDM has not been conclusively confirmed from observations.
-Around other galaxies, it is generally hard to measure the 3D shape of [..] because they are seen in projection.
-From our position within the Milky Way, our 3D view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii.
-Unfortunately, the halo also has a low density of tracers.
+On the largest scales ($\gtrsim$30~kpc for a Milky-Way-like galaxy) in the halos of galaxies, gravitational potentials are thought to be triaxial.
+However, despite suggestive evidence, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively confirmed from observations.
+Around other galaxies, it is generally hard to measure the 3D mass profile because they are seen in projection.
+From our position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii.
+Unfortunately, the Milky Way halo has a low density of visible tracers.
 Luckily, however, many of the halo stars are likely associated with debris stripped from in-falling satellite galaxies.
@@
 As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded due to the tidal forces of the host galaxy potential.
-The phase-space density and therefore the observable properties of the tidal debris are sensitive to the internal progenitor properties, orbit of the progenitor, and underlying mass distribution of the host galaxy.
+If the orbit of such an object is regular, only mildly eccentric, and stays far from sharp density features (e.g., the disk), the disruption is a steady process and the mass is slowly stretched into long streams of matter.
+The phase-space density and therefore the observable properties of these tidal streams are sensitive to the internal progenitor properties, orbit of the progenitor, and underlying mass distribution of the host galaxy.
 The morphological evolution of the debris depends on the spread of orbital properties (e.g., actions) of the debris and the orbit of the progenitor system, both of which are determined by the shape and radial profile of the gravitational potential of the host galaxy.
-In the halo of the Milky Way, tidal streams of stars are found over a large range of distances.
-By simultaneously modeling the observed density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution of the Milky Way on these large scales.
+In the halo of the Milky Way, tidal streams of stars are found over a large range of distances ($\sim$10-100~kpc).
+By simultaneously modeling the observed phase-space density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution of the Milky Way on these large scales.
 
-Tidal streams are generally modeled as 1D structures with simple internal progenitor dynamics and density evolution. 
-Pearson et al.
-(2014) tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in an oblate and a triaxial potential using \emph{Streakline} models (K\"upper et al.
-(XXXX)) :
-		\item in the oblate potential, a thin model stream is easily found that reproduces the observed stellar density morphology of the stream;
-		\item in the triaxial potential, the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness of the stream, given the observational constraints of the present-day position and velocity of the cluster.
-This result suggests that the shape of the host galaxy potential directly influences the morphology of the resulting debris, and that other morphological classes exist between thin, 1D streams and diffuse, 2D shells.
-To understand the formation of these fans and the implications of this work we must first understand the complex orbital structure of triaxial systems.
+Tidal streams are generally modeled as inherently 1D structures formed on regular orbits with simple internal progenitor dynamics and density evolution.
+Methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to direct $N$-body simulations.
+All methods have been tested in some way on simulated observations of data and authors have mainly demonstrated their ability to recover parameters of analytic, static potential forms.
+Many of the methods have been [...] in axisymmetric or otherwise simple potentials but claim to extend to more complex, generalized potentials (e.g., triaxial of multi-component potentials).
 
-In this work, we study the effect of orbital classes that are prevalent in triaxial systems (e.g., stochastic, near-resonant, resonant, and quasiperiodic) on the resulting morphology of tidal streams.
+\citet{pearson15} tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} models \citep{kupper12}.
+They use the observed SDSS number density of stars and a limited number of radial velocities measured for stream members to fit their models to the data.
+In the oblate potential (a three-component, bulge, disk, and spherical halo potential), a thin model stream is easily found that reproduces the observed stellar density morphology of the stream.
+In the triaxial potential (the potential fit to the Sagitarrius stream by \cite{law10}), the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness of the stream, given the observational constraints of the present-day position and velocity of the cluster.
+This result suggests that the shape of the host galaxy potential directly influences the morphology of the resulting debris.
+In order to extending tidal stream models to more generic potentials, it is important to understand why this stream-fanning occurs and whether this and other strange debris morphologies are generic to more complex potentials.
 
-For (typically) many orbits, these tidally stripped stars remain spatially coherent.
-Of particular interest are cold, young dynamical structures --- ``tidal streams'' -- which provide useful information about the mass distributions at large distances in the galaxies within which they orbit.
-[At these distances, disk / classic baryonic components can't help you, mostly dark matter]
-The formation and morphology of tidal streams has been studied extensively in a variety of assumed potential forms. [Stream formation in oblate, spherical potentials, action-angle, etc.]
-
-Adding third degree of freedom opens up many new orbits / resonances / etc. -- Valluri and Merritt work...
-
-The halos of galaxies are filled with substructure...
-
+One obvious difference between the two potentials considered in \citet{pearson15} is the extra symmetry of the oblate potential.
 For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (dof), the equations of motion are most simply expressed in angle-action coordinates.

### 2015-02-16 ms.tex
werqin

@@
 \altaffiltext{\adrn}{To whom correspondence should be addressed: adrn@astro.columbia.edu}
-\altaffiltext{\mich}{ }
+\altaffiltext{\mich}{Department of Astronomy, 
+			   University of Michigan,
+			   Ann Arbor, MI 48109, USA}
 
@@
 On the largest scales ($\gtrsim$30~kpc for a Milky-Way-like galaxy) in the halos of galaxies, gravitational potentials are thought to be triaxial.
-However, despite suggestive evidence, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively confirmed from observations.
+Despite suggestive evidence, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively confirmed from observations.
 Around other galaxies, it is generally hard to measure the 3D mass profile because they are seen in projection.
-From our position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii.
-Unfortunately, the Milky Way halo has a low density of visible tracers.
-Luckily, however, many of the halo stars are likely associated with debris stripped from in-falling satellite galaxies.
+From the Earth's position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii.
+Unfortunately, the Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely associated with debris stripped from in-falling satellite galaxies.
 
@@
 The phase-space density and therefore the observable properties of these tidal streams are sensitive to the internal progenitor properties, orbit of the progenitor, and underlying mass distribution of the host galaxy.
-The morphological evolution of the debris depends on the spread of orbital properties (e.g., actions) of the debris and the orbit of the progenitor system, both of which are determined by the shape and radial profile of the gravitational potential of the host galaxy.
+The morphological evolution of the debris depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are set by the shape and radial profile of the gravitational potential of the host galaxy.
 In the halo of the Milky Way, tidal streams of stars are found over a large range of distances ($\sim$10-100~kpc).
-By simultaneously modeling the observed phase-space density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution of the Milky Way on these large scales.
+By simultaneously modeling the observed phase-space density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution of the Milky Way on these scales.
 
-Tidal streams are generally modeled as inherently 1D structures formed on regular orbits with simple internal progenitor dynamics and density evolution.
+Tidal streams are generally modeled as 1D structures formed on regular orbits with simple internal progenitor dynamics and density evolution.
 Methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to direct $N$-body simulations.
 All methods have been tested in some way on simulated observations of data and authors have mainly demonstrated their ability to recover parameters of analytic, static potential forms.
-Many of the methods have been [...] in axisymmetric or otherwise simple potentials but claim to extend to more complex, generalized potentials (e.g., triaxial of multi-component potentials).
+Many of the methods have been [...] in axisymmetric or otherwise simple potentials but claim to extend to more complex, generalized potentials (e.g., triaxial or multi-component potentials).
 
-\citet{pearson15} tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} models \citep{kupper12}.
-They use the observed SDSS number density of stars and a limited number of radial velocities measured for stream members to fit their models to the data.
+\citet{pearson15} tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} models \citep{kuepper12}.
+They use the observed SDSS number density of stars and a limited number of radial velocities measured for stream members to fit model streams to the data.
 In the oblate potential (a three-component, bulge, disk, and spherical halo potential), a thin model stream is easily found that reproduces the observed stellar density morphology of the stream.
-In the triaxial potential (the potential fit to the Sagitarrius stream by \cite{law10}), the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness of the stream, given the observational constraints of the present-day position and velocity of the cluster.
+In the triaxial potential (the potential from \cite{law10}, a three component, bulge, disk, and triaxial halo fit to Sagitarrius stream data), the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness of the stream given the observational constraints of the present-day position and velocity of the cluster.
 This result suggests that the shape of the host galaxy potential directly influences the morphology of the resulting debris.
-In order to extending tidal stream models to more generic potentials, it is important to understand why this stream-fanning occurs and whether this and other strange debris morphologies are generic to more complex potentials.
+In order to extend tidal stream models to more generic potentials, it is important to understand why this stream-fanning occurs and whether this and other strange debris morphologies are generic to more complex potentials.
 
@@
 Owing to the involutive property of the actions and the periodic nature of the angle variables, the topology of angle-action space is equivalent to that of the surface of an $N$-dimensional torus, and thus orbits are often referred to in terms of \emph{orbital tori}: each set of actions, $(\act_1,\act_2,...\act_N)$, uniquely determine the torus associated with a given orbit and the angles specify the location on the surface of the $N$-torus.
+If a global transformation to angle-action coordinates exists, the potential is said to be ``integrable.''
 
-If, to first order, galactic halos are triaxial, galactic gravitational potentials will almost certainly not be globally integrable.
+If to first order galactic halos are triaxial, galactic gravitational potentials will almost certainly not be globally integrable.
 If a triaxial potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
@@
 The tori that survive will in general be separated by regions of irregular or chaotic motion, and thus any transformations to angle-action coordinates can only be defined locally.
+Analytic transformations to angle-action coordinates are only known for a handful of simple, integrable potentials.
+Much effort has gone in to developing general methods for transforming either from position-velocity to angle-action coordinates, $(\bs{x},\bs{v})\rightarrow(\bs{\theta},\bs{J})$ \citep{many}, or vice-versa, $(\bs{\theta},\bs{J})\rightarrow(\bs{x},\bs{v})$ \citep{many}.
 
@@
 Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori -- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.
-However, even orbits that are strictly chaotic may behave regularly over timescales much longer than, e.g., the dynamical time of the system.
+However, even orbits that are strictly chaotic may behave regularly over timescales much longer than, for example, the dynamical time of the system.
 Chaotic can therefore be loosely classified by being either \emph{strongly} or \emph{weakly} chaotic: strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale (we quantify this distinction in Section~\ref{sec:}).
-This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{}) or frequency diffusion time (Section~\ref{}), and to the mixing rate of small ensembles of orbits (Section~\ref{}).
+This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{}) or frequency diffusion time (Section~\ref{}).
+In Section~\ref{} we look at how these relate to the the mixing rate of small ensembles of orbits.
 
@@
 All of these parameters are specified in terms of the \emph{density}; for orbit analysis, we need to determine the form of the potential in terms of these parameters, which, in general, requires numerical integration of the density at each position of interest.
-For computational efficiency, many authors instead express the triaxiality in the form of the potential, however this can lead to unphysical situations where the density becomes negative.
+For computational efficiency, many authors instead express the triaxiality in the form of the potential, but this can lead to unphysical situations where the density becomes negative.
 \citet{leesuto03} derive a perturbative expansion of the potential integral for a triaxial NFW density and show that the expansion is accurate even for modest axis ratios (e.g., the median values shown above).
@@
 Real galactic potentials will have a significant component due to the disk and bulge and may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{veraciro11}, significant substructure \citep{zemp09}, or time dependence (either from bulk rotation, mass growth, mergers, etc.; cite).
-We expect all of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple examples that demonstrate this expectation.
+We expect all of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
 
@@
 
-We use the Dormand-Prince 8th-order Runge-Kutta scheme (cite Dormand 1981) to integrate orbits in the above potential.
+We use the Dormand-Prince 8th-order Runge-Kutta scheme (TODO: cite Dormand 1981) to integrate orbits in the above potential.
 In particular, we use the Python wrapper to the FORTRAN routines provided in the Scipy package \citep{scipy}.
@@
 1976, Tabor 1989).
-The following pseudocode outlines this procedure:\footnote{see Gary for Python implementation?}\\
-\begin{algorithmic}[1]
-\State {\bf define} orbital integration timestep, $h$, and number of steps, $K$
-\State {\bf define} initial norm of deviation vector, $\delta_0$, to be sufficiently small
-\State {\bf define} renormalization integration period, $\tau \ll hK$ 
-\For{each timestep when integrating the main orbit, $\bs{w}(t)$}
-\State step forward the orbit and deviation vector orbit by one timestep, $t_{i-1} \rightarrow t_i$
-\If {a normalization timestep}
-\State measure and store the length of the deviation vector, $\delta_i = \|\delta \bs{w}_i\|$
-\State renormalize the length of the deviation vector, $\delta \bs{w}_i = \delta \bs{w}_i (\delta_0/\delta_i)$
-\EndIf
-\EndFor 
-\end{algorithmic}
+%The following pseudocode outlines this procedure:\footnote{see Gary for Python implementation?}\\
+%\begin{algorithmic}[1]
+%\State {\bf define} orbital integration timestep, $h$, and number of steps, $K$
+%\State {\bf define} initial norm of deviation vector, $\delta_0$, to be sufficiently small
+%\State {\bf define} renormalization integration period, $\tau \ll hK$ 
+%\For{each timestep when integrating the main orbit, $\bs{w}(t)$}
+%\State step forward the orbit and deviation vector orbit by one timestep, $t_{i-1} \rightarrow t_i$
+%\If {a normalization timestep}
+%\State measure and store the length of the deviation vector, $\delta_i = \|\delta \bs{w}_i\|$
+%\State renormalize the length of the deviation vector, $\delta \bs{w}_i = \delta \bs{w}_i (\delta_0/\delta_i)$
+%\EndIf
+%\EndFor 
+%\end{algorithmic}
 The FTMLE after a given number of timesteps, $N$, is then estimated as
@@
 NAFF has been applied to many problems in Solar System and Galactic dynamics, especially in the study of of orbits in triaxial systems.
-For more details about the method, see \cite{papaphillipou, laskar, etc.}.
+For more details about the method, see \cite{papaphilippou96, laskar, etc.}.
 We have implemented and tested our own implementation of NAFF in Python.
@@
 Our goal is to characterize the morphology of tidal debris along orbits generic to triaxial potentials.
-For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{pearson15, fardal??} [TODO: should I put sarahs paper here even though we didn't say this explicitly?], however, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
+For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{pearson15, fardal14} [TODO: should I put sarahs paper here even though we didn't say this explicitly?], however, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
 Further, quantifying the degree of stochasticity with, for example, the frequency diffusion rate represents a \emph{local} measure of chaos and it is unclear how this generalizes to the small but finite spreads in orbital characteristics common to debris that is tidally disrupted from globular clusters or dwarf satellite galaxies.
@@
 Orbits that try to circulate about the intermediate axis tend to be highly irregular and do not preserve their sense of circulation for long periods of time.
-NAFF recovers fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL95][]{papaphillippou96}.
+NAFF recovers fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL96][]{papaphilippou96}.
 PL96 show that a good choice of coordinates for tube orbits are the Poincar\'e symplectic polar coordinates, a set of canonical coordinates similar to cylindrical coordinates.
-For approximately XX\% of the orbits, NAFF fails to recover the frequencies (similar difficulties were reported in \cite{valluri98} and \cite{papaphillippou96} for recovering certain tube orbit frequencies).
+For approximately XX\% of the orbits, NAFF fails to recover the frequencies (similar difficulties were reported in \cite{valluri98} and \cite{papaphilippou96} for recovering certain tube orbit frequencies).
 

### 2015-02-19 ms.tex
hyperref
intro end

@@
 \usepackage{subfigure}
+\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
 
@@
 \begin{document}
+
+%TODO:
+%The classical way to detect chaotic motion is
+%the computation of the Lyapunov exponents. We
+%claim that for weakly chaotic solutions, the frequency
+%analysis gives information more rapidly
+%and more precisely. The reason is that the frequency
+%analysis is an instantaneous observation
+%of the fundamental frequencies which are essential
+%parameters of the dynamical system, while
+%the usual computation of Lyapunov exponents
+%acts like a mean and needs a very long time span
+%to provide reliable results (for references and
+%discussion on Lyapunov exponents, see refs. [3,
+%12].
 
@@
 However, even orbits that are strictly chaotic may behave regularly over timescales much longer than, for example, the dynamical time of the system.
-Chaotic can therefore be loosely classified by being either \emph{strongly} or \emph{weakly} chaotic: strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale (we quantify this distinction in Section~\ref{sec:}).
-This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{}) or frequency diffusion time (Section~\ref{}).
-In Section~\ref{} we look at how these relate to the the mixing rate of small ensembles of orbits.
+Chaotic can therefore be loosely classified by being either \emph{strongly} or \emph{weakly} chaotic: strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale (we quantify this distinction with the experiments of Section~\ref{sec:experiments}).
+This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{sec:lyap}) or frequency diffusion time (Section~\ref{sec:naff}).
 
-[summarizing paragraph connecting everything, mention mixing, relation to tidal streams]
+These local measures quantify the divergence of infinitesimally nearby orbits or the rate of frequency diffusion along an orbit.
+For an ensemble of orbits with finite spreads in energy or frequencies (e.g., tidal debris), each orbit will have a slightly different characteristic timescale, but we still expect mixing time of the ensemble to correlate with local measures of chaos of the ``parent'' orbit of the ensemble (Section~\ref{sec:ensemble}).
+The evolution of tidal debris will depend on the ensemble mixing rate of the parent orbit: chaotic mixing will cause the density of the debris to fall off more quickly than the typical power-law behavior expected for regular orbits \citep[e.g.,][]{merritt96, helmi99}.
+Since all known tidal streams have been discovered as kinematic overdensities of stars, this implies (1) [something about accretion history], (2) the long, thin tidal streams we see may represent a sampling from ``special'' orbits in the Galactic potential where the mixing time is low, and (3) previous assumptions that connect the length and size of streams to their evolution time may be invalid.
 
-\section{Numerical methods}
+This paper is organized as follows: In Section~\ref{sec:methods}, we outline the numerical methods used for the experiments of Section~\ref{sec:experiments}.
+In Section~\ref{sec:discussion}, we describe the connection between different measures of chaotic behavior, further discuss the implications of chaotic mixing, and comment on the applicability of the methods presented here to more realistic Galactic potentials and simulations.
+
+\section{Numerical methods}\label{sec:methods}
 
@@
 
-\subsection{Lyapunov exponents}
+\subsection{Lyapunov exponents} \label{sec:lyap}
 
@@
 
-\subsection{Numerical estimation of fundamental frequencies}
+\subsection{Numerical estimation of fundamental frequencies}\label{sec:naff}
 
@@
 
-\subsection{Test particle ensembles}
+\subsection{Test particle ensembles} \label{sec:ensemble}
 
@@
 
-\section{Experiments}
+\section{Experiments}\label{sec:experiments}
 
@@
 
-[Lyapunov exponent grids...]
+\begin{figure}[!h]
+\centering
+	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_xz.png}
+	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_yz.png}
+	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+	\label{fig:lyapmap}
+\end{figure}
+
+We also compute the Lyapunov exponents for all orbits in the grid by integrating their orbits for 250 orbital periods.
+Figure~\ref{fig:lyapmap} shows the initial condition grids again, but now the greyscale intensity corresponds to the logarithm of the Lyapunov time, $t_\lambda = \lambda_{\rm max}^{-1}$.
+These Lyapunov exponent maps show that, for a fixed integration time, only the most chaotic orbits are identified.
+NAFF is able to detect small differences in the frequencies and thus measure small diffusion rates with comparable integration times.
 
@@
 \acknowledgements
-The authors wish to thank Sarah Pearson, Andreas K\"upper, David Hogg, David Merritt, \emph{Stream Team}, etc...
+The authors wish to thank Sarah Pearson, Andreas K\"upper, David Hogg, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 

### 2015-02-20 ms.tex
lyapunov

@@
 
-\subsection{Potential choice}
+\subsection{Potential choice}\label{sec:potential}
 
@@
 \end{equation}
-[Some more practical issues with implementation] [Figure showing $l_N$ for regular, chaotic orbit?]
 
-[Copy in text fomr ``behavior of MLE'']
+For most regular orbits, deviations will grow close to linearly or as some power-law of time. 
+%As we have seen (Section~\ref{sec:introduction}), if the orbit is regular, there exists a local transformation to action-angle variables where the angle variables increase linearly with time, $\theta_n \propto \Omega_n t$.
+We can look at small variations around the orbit defined by the above equations:\footnote{Note the implied summation in repeated indices in the above expression.}
+%\begin{align}
+%	\frac{d (\delta \theta_n)}{dt} &= \frac{\partial \Omega_n}{\partial J_m} \delta J_m
+%\end{align}
+%These equations are trivially integrated:
+%\begin{align}
+%	\delta \theta_n(t) &= \delta \theta_n(0) + \left(\frac{\partial \Omega_n}{\partial J_m} \delta J_m \right) t
+%\end{align}
+%It's clear then that the norm of the deviation vector grows linearly with time:
+%\begin{align}
+%	\|\delta \bs{w}(t)\| &= \left[\sum_n (\delta \theta_n(t))^2 + \sum_n (\delta J_n)^2\right]^{1/2}\\
+%	&= \left[\sum_n \left(\delta \theta_n(0) + \frac{\partial \Omega_n}{\partial J_m} \delta J_m t\right)^2 + \sum_n (\delta J_n)^2\right]^{1/2}\\
+%	&\propto t.
+%\end{align}
+From Eq.~\ref{eq:lmax} and Eq.~\ref{eq:lyapmax} it is evident that any deviation vector that grows as a power law with time, $t^k$, will asymptote to 0 from the limit 
+\begin{equation}
+	\lambda_{\rm max} \propto \lim_{t\rightarrow \infty} k \frac{\ln t}{t} = 0.
+\end{equation}
+At long times, the implication is that the the numerically computed MLE should approach 0 like $t^{-1}$ for regular orbits.
+For chaotic orbits, the divergence is exponential, and the limit should converge to the rate of the exponential: the Lyapunov exponent.
 
-[Computationally, expensive to compute Lyapunov exponent for sticky or mildly chaotic orbits]
+In practice, the MLE is often estimated as the mean of $l_N$ after the summation diverges from power-law behavior.
+Figure~\ref{fig:lyap-orbits} shows $l_N$ computed for two nearby orbits in the potential of Section~\ref{sec:potential}.
+In physical units and Cartesian coordinates, the initial conditions for the two orbits are:
+\begin{align}
+	\bs{x}_{0}^{(1)} &= (27.85, 0, 23.35)~{\rm kpc}\\
+	\bs{v}_{0}^{(1)} &= (0, 115.37, 0)~{\rm km~s}^{-1}\\
+	\bs{x}_{0}^{(2)} &= (27.85, 0, 23.10)~{\rm kpc}\\
+	\bs{v}_{0}^{(2)} &= (0, 116.43, 0)~{\rm km~s}^{-1}.
+\end{align}
+Over-plotted on both panels are lines (dashed) that show $t^{-0.9}$ behavior.
+The MLE estimate continues towards zero for the regular orbit (top) but saturates for the chaotic orbit (bottom) and continues to increase at least until the end of our integration period.
+For mildly chaotic or ``sticky'' orbits, it can take integration periods of many hundreds or thousands of orbits to determine a reliable estimate of the Lyapunov exponent.
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=0.75\textwidth]{figures/lyap-orbits.pdf}
+\caption{ } \label{fig:lyap-orbits}
+\end{center}
+\end{figure*}
 

### 2015-02-22 ms.tex
minor changes from draft reading

@@
 
+% Highlight that stochastic tubes have not been studied much...
+
 \title{Tidal streams in triaxial systems}
@@
 	    Monica Valluri\altaffilmark{\mich}}
+\date{\centering \today}
 
@@
 On the largest scales ($\gtrsim$30~kpc for a Milky-Way-like galaxy) in the halos of galaxies, gravitational potentials are thought to be triaxial.
-Despite suggestive evidence, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively confirmed from observations.
+Despite suggestive evidence from a range of complimentary techniques, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively verified by observations.
 Around other galaxies, it is generally hard to measure the 3D mass profile because they are seen in projection.
@@
 The phase-space density and therefore the observable properties of these tidal streams are sensitive to the internal progenitor properties, orbit of the progenitor, and underlying mass distribution of the host galaxy.
-The morphological evolution of the debris depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are set by the shape and radial profile of the gravitational potential of the host galaxy.
+The morphological evolution of the debris depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also set by the shape and radial profile of the gravitational potential of the host galaxy.
 In the halo of the Milky Way, tidal streams of stars are found over a large range of distances ($\sim$10-100~kpc).
@@
 All methods have been tested in some way on simulated observations of data and authors have mainly demonstrated their ability to recover parameters of analytic, static potential forms.
-Many of the methods have been [...] in axisymmetric or otherwise simple potentials but claim to extend to more complex, generalized potentials (e.g., triaxial or multi-component potentials).
+Many of the methods have been shown to work in axisymmetric or otherwise simple potentials but claim to extend to more complex, generalized potentials (e.g., triaxial or multi-component potentials).
 
@@
 In order to extend tidal stream models to more generic potentials, it is important to understand why this stream-fanning occurs and whether this and other strange debris morphologies are generic to more complex potentials.
+This also has implications for modeling the accretion histories of galaxies.
+Debris morphology is often separated into two classes --- streams (usually from high angular momentum progenitor orbits) and shells (from radially plunging orbits) --- but in practice the distinction may not be as clear due to other mixing effects.
 
-One obvious difference between the two potentials considered in \citet{pearson15} is the extra symmetry of the oblate potential.
-For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (dof), the equations of motion are most simply expressed in angle-action coordinates.
+The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the oblate potential.
+For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (dof), the equations of motion are most simply expressed in angle-action coordinates \cite[e.g.,][]{goldstein80}.
 In these coordinates, the actions, $J_n$, are integrals of the motion, and the conjugate angle variables, $\theta_n$, increase linearly with time
 \begin{align}
+	\dot{\act}_n &= -\frac{\partial H}{\partial \theta_n} = 0\\
 	\act_n &= {\rm const.}\\
-	\dot{\act}_n &= -\frac{\partial H}{\partial \theta_n} = 0\\
-	\theta_n(t) &= \Omega_n t + \theta_n(0)\\
-	\dot{\theta}_n &= \frac{\partial H}{\partial \act_n} = \Omega_n(\act_1,\act_2,...,\act_n)\label{eq:aafreq}
+	\dot{\theta}_n &= \frac{\partial H}{\partial \act_n} = \Omega_n(\act_1,\act_2,...,\act_n)\label{eq:aafreq}\\
+	\theta_n(t) &= \Omega_n t + \theta_n(0)
 \end{align}
 where $i=1,2,...N$, and $\Omega_n$ are the fundamental frequencies of the orbit.
-Owing to the involutive property of the actions and the periodic nature of the angle variables, the topology of angle-action space is equivalent to that of the surface of an $N$-dimensional torus, and thus orbits are often referred to in terms of \emph{orbital tori}: each set of actions, $(\act_1,\act_2,...\act_N)$, uniquely determine the torus associated with a given orbit and the angles specify the location on the surface of the $N$-torus.
-If a global transformation to angle-action coordinates exists, the potential is said to be ``integrable.''
+The topology of angle-action space is equivalent to that of the surface of an $N$-dimensional torus, and thus orbits are often referred to in terms of \emph{orbital tori}: each set of actions, $(\act_1,\act_2,...\act_N)$, uniquely ``label'' the torus associated with a given orbit and the angles specify the location on the surface of the $N$-torus.
+If a global transformation to angle-action coordinates exists, the potential is said to be \emph{integrable}.
 
-If to first order galactic halos are triaxial, galactic gravitational potentials will almost certainly not be globally integrable.
-If a triaxial potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
+If galactic halos are triaxial, galactic gravitational potentials will almost certainly not be globally integrable.
+If a triaxial (three dof) potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
 \begin{equation}
@@
 Stable, resonant orbits generate families of orbits that may behave or appear similarly, whereas unstable resonances are typically associated with regions of chaos.
-However not all orbits can be trivially associated with resonances due to the infinitely complex, hierarchical structure of resonances (though the importance or ``size'' of a resonance falls off inversely with the norm of its integer vector, $|\bs{m}|^{-\alpha}$, with $\alpha > 0$).
+Not all orbits can be trivially associated with resonances due to the infinitely complex, hierarchical structure of resonances (though the importance or ``size'' of a resonance falls off inversely with the norm of its integer vector, $|\bs{m}|^{-\alpha}$, with $\alpha > 0$).
 In autonomous, two dof systems, stochastic regions are generally surrounded and separated by regular motion, thus chaotic orbits are confined.
@@
 However, even orbits that are strictly chaotic may behave regularly over timescales much longer than, for example, the dynamical time of the system.
-Chaotic can therefore be loosely classified by being either \emph{strongly} or \emph{weakly} chaotic: strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale (we quantify this distinction with the experiments of Section~\ref{sec:experiments}).
+Chaotic orbits can therefore be loosely classified by being either \emph{strongly} or \emph{weakly} chaotic: strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale (we quantify this distinction with the experiments of Section~\ref{sec:experiments}).
 This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{sec:lyap}) or frequency diffusion time (Section~\ref{sec:naff}).
@@
 \citet{jing02} found that a triaxial generalization of the classic NFW density profile \citep{navarro96} generates excellent fits to their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of the many halos formed in their simulations.
-They find median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that \citet{jing02} use the opposite notation so that $c$ is the major and $a$ is the minor axis.} They find significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).
+They find median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that \citet{jing02} use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
+They find significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).
 
@@
 
-For this work, we use the triaxial potential expression from \citet{leesuto03}, parametrized in a slightly different manner.
+In this work, we use the triaxial potential expression from \citet{leesuto03}, parametrized in a slightly different manner.
 In terms of spherical coordinates\footnote{$(r,\phi,\theta)$ = (radius, azimuth, colatitude)} with the radius normalized by the scale radius, $u = r/r_s$
@@
 The functions $F_i(u)$ are given in the appendix of \cite{leesuto03}.
-We chose $r_s=40~{\rm kpc}$ and $v_c = 150~{\rm km}~{\rm s}^{-1}$ by taking the mean halo concentration found by \cite{jing02} for a ${\rm M}_{vir} \approx 10^{12}~\msun$ halo, $c_e\approx5$, and assuming $R_{vir}\approx200~{\rm kpc}$.
+We chose $r_s=40~{\rm kpc}$ and $v_c = 150~{\rm km}~{\rm s}^{-1}$ by taking the mean halo concentration for a ${\rm M}_{vir} \approx 10^{12}~\msun$ halo, $c_e\approx5$, from \cite{jing02} and by assuming $R_{vir}\approx200~{\rm kpc}$.
 Figure~\ref{fig:potential} shows the rotation curve of this potential along the three principal axes.
@@
 This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the effect of chaos in the halos of such galaxies.
-Real galactic potentials will have a significant component due to the disk and bulge and may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{veraciro11}, significant substructure \citep{zemp09}, or time dependence (either from bulk rotation, mass growth, mergers, etc.; cite).
+Realistic Falactic potentials will have a significant component due to the disk and bulge and may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{veraciro11}, significant substructure \citep{zemp09}, or time dependence (either from bulk rotation, mass growth, mergers, etc.; cite).
 We expect all of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
@@
 For all orbits, we ensure that energy is conserved to $\Delta E/E_0 \leq 10^{-9}$ by the end of integration.
-Unless otherwise specified, orbits are integrated for $\sim$200 orbital periods with 500 steps per orbital period.
+Unless otherwise specified, orbits are integrated for $\sim$250 orbital periods with 500 steps per orbital period.
 Some box orbits require more internal integration steps in order to conserve energy to the required level.
@@
 
-The simplest and perhaps most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent of an orbit.
+The simplest and perhaps most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit.
 Defining the vector $\bs{w} = (q_1,...,q_N,p_1,...,p_N)$ for some set of canonical coordinates $(q_i,p_i)$, we can write Hamilton's equations as
@@
 with $N$-dimensional identity matrices $\ident$.
-If we consider a nearby position separated from $\bs{w}$ by an infinitesimal deviation $\delta \bs{w}$, such that $\bs{w}' = \bs{w} + \delta \bs{w}$.
+If we consider a nearby phase-space position, $\bs{w}'$, separated from $\bs{w}$ by an infinitesimal deviation, $\delta \bs{w}$, such that $\bs{w}' = \bs{w} + \delta \bs{w}$.
 We can expand to linear order in the deviation about the parent orbit and write the equations of motion for the deviation as follows
@@
 \end{align}
-where $\jac$ is the Jacobian of Eq.~\ref{eq:ham}, evaluated at the parent orbit.
+where $\jac$ is the Jacobian of Eq.~\ref{eq:ham}, evaluated at the parent orbit (the Hessian of the Hamiltonian).
 For chaotic orbits, the maximum eigenvalue of the solution matrix to Eq.~\ref{eq:deviate} is positive real, leading to exponential divergence of nearby orbits.
@@
 \end{equation}
-and the MLE is
+where $\delta_i=\|\delta \bs{w}(t_i)\|$ and the MLE is
 \begin{equation}
@@
 \end{equation}
-At long times, the implication is that the the numerically computed MLE should approach 0 like $t^{-1}$ for regular orbits.
+At long times, the numerically computed MLE should approach 0 as $t^{-1}$ for regular orbits.
 For chaotic orbits, the divergence is exponential, and the limit should converge to the rate of the exponential: the Lyapunov exponent.
@@
 \cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hanning-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
-It has been shown that using the Hanning filter makes the accuracy in determining the frequencies scale as $\inttime^{-4}$ \citep{laskar99}, where $\inttime$ is the integration time.
+It has been shown that using a Hanning filter makes the accuracy in determining the frequencies scale as $\inttime^{-4}$ instead of $\inttime^{-1}$ for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time.
+In each complex time series, once the frequency is found, the amplitude of the term is found by projecting back on to the function itself.
+The component is then subtracted from the time series, and this procedure continues iteratively until some breaking condition (for example, the amplitude of the next term is below some threshold).
 We follow \cite{valluri98} and refer to this method as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) even though it differs slightly from the original method from Laskar.
@@
 For a weakly chaotic orbit, the orbit may act consistently periodic over long windows of time.
-NAFF will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change character (possibly) smoothly with time.
+NAFF will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change character with time.
 For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two separate sections of orbit.
@@
 \end{equation}
-where $\bs{\Omega}_{T_1}$ and $\bs{\Omega}_{T_2}$ are the fundamental frequencies computed from two consecutive sections of orbit, $\inttime$ is the total integration time (over both sections), and $\bs{a}$ are the amplitudes corresponding to the frequencies.
-The frequency diffusion rate measured in the frequency with the largest amplitude is better correlated with the strength of chaos in an orbit relative to, e.g., the mean or maximum rate (cite valluri recent work?).
-
-We have found that for ``sticky'' or mildly chaotic orbits, reliably computing the Lyapunov exponent can require integration times equivalent to thousands of orbital periods.
-In comparison, owing to the precision of NAFF, detecting small frequency diffusion rates can be done with only hundreds of orbital periods.
-[example?
-comparison?]
+where $\bs{\Omega}_{T_1}$ and $\bs{\Omega}_{T_2}$ are the fundamental frequencies computed from two consecutive sections of orbit, $\inttime$ is the total integration time (over both sections), and the maximum is taken with relation to the corresponding amplitudes, $\bs{a}$, of the frequency components.
+The frequency diffusion rate as defined above (e.g., measured in the frequency with the largest amplitude) is better correlated with the strength of chaos in an orbit \citep{valluri??} but differs from the original definition of \cite{valluri98}.
 
@@
 260, Figure~3.45]{binneytremaine}.
-Plotted are the (Cartesian) frequency ratios recovered for a grid of iso-energy, box orbits integrated in the potential $\Phi(x,y,z) = \frac{1}{2}\ln\left(x^2 + (y/0.9)^2 + (z/0.7)^2 + 0.1\right)$.
+Plotted are the (Cartesian) frequency ratios recovered for a grid of iso-energy, box orbits integrated in the potential
+\begin{equation}
+	\Phi(x,y,z) = \frac{1}{2}\ln\left(x^2 + (y/0.9)^2 + (z/0.7)^2 + 0.1\right).
+\end{equation}
 In such a map, stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
@@
 
+We have found that for ``sticky'' or mildly chaotic orbits, reliably computing the Lyapunov exponent can require integration times equivalent to thousands of orbital periods.
+In comparison, owing to the precision of NAFF, detecting small frequency diffusion rates can be done with only hundreds of orbital periods.
+[example?
+comparison?]
 [Figure: for same orbits as in Lyapunov section, show frequency diffusion by computing frequencies in rolling windows?]
@@
 
-\subsection{Adding a disk?}
-
 \section{Discussion}\label{sec:discussion}

### 2015-02-22 ms.tex
added a figure, changed norm symbol

@@
 \begin{center}
-\includegraphics[width=0.75\textwidth]{figures/log-freqmap.png}
+\includegraphics[width=0.5\textwidth]{figures/log-freqmap.png}
 \caption{This is really just a placeholder that shows what I want to show here...} \label{fig:logfreqs}
@@
 We have found that for ``sticky'' or mildly chaotic orbits, reliably computing the Lyapunov exponent can require integration times equivalent to thousands of orbital periods.
-In comparison, owing to the precision of NAFF, detecting small frequency diffusion rates can be done with only hundreds of orbital periods.
-[example?
-comparison?]
-[Figure: for same orbits as in Lyapunov section, show frequency diffusion by computing frequencies in rolling windows?]
+In comparison, owing to the precision of NAFF, detecting small frequency diffusion rates can be done with only $\sim$100 orbital periods.
+Figure~\ref{fig:fdiff-orbits} shows the two strongest frequency components plotted against each other for the two orbits of Figure~\ref{fig:lyap-orbits}.
+The orbits were integrated for a total of 400 orbital periods and the frequencies were computed for 12 consecutive windows of orbit, where each section was 100 orbital periods in length.
+The frequencies are consistent for the regular orbit (left panel) whereas there is significant diffusion in the chaotic orbit (right panel).
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/freq-diff-orbits.pdf}
+\caption{ } \label{fig:fdiff-orbits}
+\end{center}
+\end{figure*}
 
@@
 
+% TODO: below, should I put sarahs paper here even though we didn't say this explicitly?
 Our goal is to characterize the morphology of tidal debris along orbits generic to triaxial potentials.
-For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{pearson15, fardal14} [TODO: should I put sarahs paper here even though we didn't say this explicitly?], however, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
+For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{fardal14, pearson15}, however, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
 Further, quantifying the degree of stochasticity with, for example, the frequency diffusion rate represents a \emph{local} measure of chaos and it is unclear how this generalizes to the small but finite spreads in orbital characteristics common to debris that is tidally disrupted from globular clusters or dwarf satellite galaxies.
@@
 	\delta\bs{v}_i &\sim \mathcal{N}(0, \sigma_v)\\
-	r_{\rm tide} &=|\bs{x}_0| \left(\frac{m}{3M_{\rm enc}(|\bs{x}_0|)}\right)^{1/3} \\
-	\sigma_v &= |\bs{v}_0|\left(\frac{m}{3M_{\rm enc}(|\bs{x}_0|)}\right)^{1/3} 
+	r_{\rm tide} &= \|\bs{x}_0\| \left(\frac{m}{3M_{\rm enc}(\|\bs{x}_0\|)}\right)^{1/3} \\
+	\sigma_v &= \|\bs{v}_0\|\left(\frac{m}{3M_{\rm enc}(\|\bs{x}_0\|)}\right)^{1/3} 
 \end{align}
-where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, and $m$ is the mass scale of the debris, and $|\cdot |$ is the Euclidean norm.
+where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, and $m$ is the mass scale of the debris, and $\|\cdot \|$ is the Euclidean norm.
 For this work, we take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass to estimate the above scales.
@@
 The final distribution, rather than fill the energy hypersurface, will instead have some thickness around this surface.
-Around a typical regular orbit, each individual orbit will fill its torus ergodically -- that is, a surface with lower dimension than the energy hypersurface -- and thus, a small enough ensemble will never get ``close'' to the fully mixed state as defined above because it will remain confined into a subspace on the energy surface.
+Around a typical regular orbit, each individual orbit will fill its torus ergodically --- a surface with lower dimension than the energy hypersurface --- and thus, a small enough ensemble will never reach the fully mixed state as defined above because it will remain confined into a subspace on the energy surface.
 In practice, only strongly chaotic systems will approach $\rho_{E_0}(\bs{x})$; the rate will depend on how stochastic the parent orbit is (e.g., as measured by the Lyapunov time or frequency diffusion rate), and also on how ``trapped'' the chaotic region is (a small chaotic region wedged between stable tori will take longer to appear chaotic, even if internally the frequency diffusion rate is large).
@@
 
-Time dependence
-
-Radial dependence of axis ratios
-
-Why not disk + bulge? What do we expect when we add? See later section for test of this...
+Why not disk + bulge? What do we expect when we add?
 
 Real disk will be self-consistent with halo, so this isn't really a fair test, but still...
+
+Other effects: radial dependence of axis ratios, time dependence, etc.
 

### 2015-02-23 ms.tex
minor kvj edits

@@
 
-On the largest scales ($\gtrsim$30~kpc for a Milky-Way-like galaxy) in the halos of galaxies, gravitational potentials are thought to be triaxial.
+On the largest scales in the halos of galaxies ($\gtrsim$30~kpc for a Milky-Way-like galaxy), gravitational potentials are thought to be triaxial.
 Despite suggestive evidence from a range of complimentary techniques, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively verified by observations.
@@
 From the Earth's position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii.
-Unfortunately, the Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely associated with debris stripped from in-falling satellite galaxies.
+Unfortunately, the Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely on non-random orbits and thus contain extra information (e.g., debris stripped from in-falling satellite galaxies).
 
@@
 Tidal streams are generally modeled as 1D structures formed on regular orbits with simple internal progenitor dynamics and density evolution.
-Methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to direct $N$-body simulations.
+Methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to N-body simulations.
 All methods have been tested in some way on simulated observations of data and authors have mainly demonstrated their ability to recover parameters of analytic, static potential forms.

### 2015-02-26 ms.tex
monicas comments

@@
 
-If galactic halos are triaxial, galactic gravitational potentials will almost certainly not be globally integrable.
+Galactic gravitational potentials will almost certainly not be globally integrable, especially if the dark matter halos of galaxies are triaxial.
 If a triaxial (three dof) potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
@@
 The types of motion generic to triaxial systems can be split into several sub-categories (summarized in Figure~\ref{fig:orbit-tree}).
-Regular motion can be classified by the number of resonance relations obeyed \citep[e.g.,][]{lichtenberg83, valluri98}: \emph{conditionally periodic} orbits obey no resonance relations, \emph{uni-resonant} orbits obey a single relation of the form $\bs{m}\cdot\bs{\Omega}=0$, and \emph{bi-resonant} orbits obey two resonance relations, $\bs{m}\cdot\bs{\Omega}=0$ and $\bs{n}\cdot\bs{\Omega}=0$, where $\bs{m}$ and $\bs{n}$ are integer vectors.
+Regular motion can be classified by the number of resonance relations obeyed \citep[e.g.,][]{lichtenberg83}: \emph{conditionally periodic} orbits obey no resonance relations, \emph{uni-resonant} orbits obey a single relation of the form $\bs{m}\cdot\bs{\Omega}=0$, and \emph{bi-resonant} orbits obey two resonance relations, $\bs{m}\cdot\bs{\Omega}=0$ and $\bs{n}\cdot\bs{\Omega}=0$, where $\bs{m}$ and $\bs{n}$ are integer vectors.
 
@@
 
-Resonant orbits are rare \citep{merritt99} but dynamically important in systems with several degrees of freedom as they determine the structure of orbit-space.
+Truly resonant orbits are rare \citep{merritt99}, but dynamically important in systems with several degrees of freedom as they determine the structure of orbit-space.
 Stable, resonant orbits generate families of orbits that may behave or appear similarly, whereas unstable resonances are typically associated with regions of chaos.
@@
 The evolution of tidal debris will depend on the ensemble mixing rate of the parent orbit: chaotic mixing will cause the density of the debris to fall off more quickly than the typical power-law behavior expected for regular orbits \citep[e.g.,][]{merritt96, helmi99}.
-Since all known tidal streams have been discovered as kinematic overdensities of stars, this implies (1) [something about accretion history], (2) the long, thin tidal streams we see may represent a sampling from ``special'' orbits in the Galactic potential where the mixing time is low, and (3) previous assumptions that connect the length and size of streams to their evolution time may be invalid.
+Since all known tidal streams have been discovered as kinematic overdensities of stars, this implies (1) [something about accretion history], (2) the long, thin tidal streams we see may represent a sampling from ``special'' orbits in the Galactic potential where the mixing time is long, and (3) previous assumptions that connect the length and size of streams to their evolution time may be invalid.
 
@@
 This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the effect of chaos in the halos of such galaxies.
-Realistic Falactic potentials will have a significant component due to the disk and bulge and may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{veraciro11}, significant substructure \citep{zemp09}, or time dependence (either from bulk rotation, mass growth, mergers, etc.; cite).
+Realistic Galactic potentials will have a significant component due to the disk and bulge and may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}.
 We expect all of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
@@
 We instead use ensembles with small ($\approx$1\%) spreads in energy, but we still expect that such an ensemble around a parent chaotic orbit will reach a fully mixed state.
-The final distribution, rather than fill the energy hypersurface, will instead have some thickness around this surface.
+The final distribution, in addition to filling the energy hypersurface, will also have some thickness around this surface.
 Around a typical regular orbit, each individual orbit will fill its torus ergodically --- a surface with lower dimension than the energy hypersurface --- and thus, a small enough ensemble will never reach the fully mixed state as defined above because it will remain confined into a subspace on the energy surface.

### 2015-03-02 ms.tex
restructuring

@@
 As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded due to the tidal forces of the host galaxy potential.
-If the orbit of such an object is regular, only mildly eccentric, and stays far from sharp density features (e.g., the disk), the disruption is a steady process and the mass is slowly stretched into long streams of matter.
-The phase-space density and therefore the observable properties of these tidal streams are sensitive to the internal progenitor properties, orbit of the progenitor, and underlying mass distribution of the host galaxy.
+If the orbit of such an object is regular, only mildly eccentric, and stays far from sharp density features (e.g., the disk), the disruption is a steady process and the mass is slowly stretched into long streams of matter: tidal (stellar) streams.
+The phase-space density and therefore the observable properties of these tidal streams are sensitive to the internal progenitor properties, orbit of the progenitor, and  mass distribution of the host galaxy.
 The morphological evolution of the debris depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also set by the shape and radial profile of the gravitational potential of the host galaxy.
@@
 
-\citet{pearson15} tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} models \citep{kuepper12}.
+One example of stream modeling in a multi-component (static, analytic) potential was done by \citet{pearson15}, who tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} models \citep{kuepper12}.
 They use the observed SDSS number density of stars and a limited number of radial velocities measured for stream members to fit model streams to the data.
@@
 This result suggests that the shape of the host galaxy potential directly influences the morphology of the resulting debris.
-In order to extend tidal stream models to more generic potentials, it is important to understand why this stream-fanning occurs and whether this and other strange debris morphologies are generic to more complex potentials.
-This also has implications for modeling the accretion histories of galaxies.
-Debris morphology is often separated into two classes --- streams (usually from high angular momentum progenitor orbits) and shells (from radially plunging orbits) --- but in practice the distinction may not be as clear due to other mixing effects.
+% In order to extend tidal stream models to more generic potentials, it is important to understand why this stream-fanning occurs and whether this and other strange debris morphologies are generic to more complex potentials.
 
 The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the oblate potential.
-For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (dof), the equations of motion are most simply expressed in angle-action coordinates \cite[e.g.,][]{goldstein80}.
-In these coordinates, the actions, $J_n$, are integrals of the motion, and the conjugate angle variables, $\theta_n$, increase linearly with time
-\begin{align}
-	\dot{\act}_n &= -\frac{\partial H}{\partial \theta_n} = 0\\
-	\act_n &= {\rm const.}\\
-	\dot{\theta}_n &= \frac{\partial H}{\partial \act_n} = \Omega_n(\act_1,\act_2,...,\act_n)\label{eq:aafreq}\\
-	\theta_n(t) &= \Omega_n t + \theta_n(0)
-\end{align}
-where $i=1,2,...N$, and $\Omega_n$ are the fundamental frequencies of the orbit.
-The topology of angle-action space is equivalent to that of the surface of an $N$-dimensional torus, and thus orbits are often referred to in terms of \emph{orbital tori}: each set of actions, $(\act_1,\act_2,...\act_N)$, uniquely ``label'' the torus associated with a given orbit and the angles specify the location on the surface of the $N$-torus.
-If a global transformation to angle-action coordinates exists, the potential is said to be \emph{integrable}.
+It is well known that the number of symmetries or the number of degrees of freedom of a potential plays a major role in determining the orbital structure within the potential.
+The peculiar shape of the particular triaxial potential also likely has significant chaotic regions within the potential.
+\citet{pearson15} tested the stochasticity of the orbit of the fanned debris progenitor by computing the Lyapunov exponent but found that the orbit is consistent with being regular over dynamically relevant timescales (many Hubble times).
+It has been shown qualitatively that along some strongly chaotic orbits, tidal streams form large, diffuse fans of debris \citep[e.g.,][]{fardal14}.
+However, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity; the result from \citet{pearson15} suggests that even a weakly chaotic progenitor orbit effects the density and therefore observability of tidal streams.
 
-Galactic gravitational potentials will almost certainly not be globally integrable, especially if the dark matter halos of galaxies are triaxial.
-If a triaxial (three dof) potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
-\begin{equation}
-	H(\bs{\theta}, \bs{\act}) = H_0(\bs{\act}) + \epsilon H_1(\bs{\theta}, \bs{\act})
-\end{equation}
-then a large number of regular orbits will survive the perturbation (by way of the KAM theorem)\footnote{3-vectors are represented with bold symbols, e.g., $\bs{J}=(J_1,J_2,J_3)$.}.
-The tori that survive will in general be separated by regions of irregular or chaotic motion, and thus any transformations to angle-action coordinates can only be defined locally.
-Analytic transformations to angle-action coordinates are only known for a handful of simple, integrable potentials.
-Much effort has gone in to developing general methods for transforming either from position-velocity to angle-action coordinates, $(\bs{x},\bs{v})\rightarrow(\bs{\theta},\bs{J})$ \citep{many}, or vice-versa, $(\bs{\theta},\bs{J})\rightarrow(\bs{x},\bs{v})$ \citep{many}.
+In this work, we analyze whether the final configuration-space morphology of simple tidal stream models is correlated with measures of chaos local to the orbit of the progenitor.
+The effect of chaos in the halos of galaxies has largely been neglected because the chaotic timescales estimated were thought to be many times the age of the universe, but recent work has shown that even weak chaos may be relevant for tidal stream morphology.
+Here we choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos as computed from single-orbit diagnostics, and compare these results to measures of the mixing time of finite-volume ensembles of orbits (analogs to tidal debris).
+Finally, we connect the ensemble mixing rate to observational properties, thus demonstrating that local orbital measures of chaos can be used to predict the [...].
 
-The types of motion generic to triaxial systems can be split into several sub-categories (summarized in Figure~\ref{fig:orbit-tree}).
-Regular motion can be classified by the number of resonance relations obeyed \citep[e.g.,][]{lichtenberg83}: \emph{conditionally periodic} orbits obey no resonance relations, \emph{uni-resonant} orbits obey a single relation of the form $\bs{m}\cdot\bs{\Omega}=0$, and \emph{bi-resonant} orbits obey two resonance relations, $\bs{m}\cdot\bs{\Omega}=0$ and $\bs{n}\cdot\bs{\Omega}=0$, where $\bs{m}$ and $\bs{n}$ are integer vectors.
+This paper is organized as follows: In Section~\ref{sec:methods}, we describe our choice of potential and method for numerical orbit integration.
+We review relevant nonlinear dynamics in Section~\ref{sec:nldreview}.
+We present our results in Section~\ref{sec:results} and discuss their implications in Section~\ref{sec:discussion}.
+We conclude in Section~\ref{sec:conclusions}.
 
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=0.5\textwidth]{figures/orbit-tree.pdf}
-\caption{} \label{fig:orbit-tree}
-\end{center}
-\end{figure*}
+% , we describe the connection between different measures of chaotic behavior, further discuss the implications of chaotic mixing, and comment on the applicability of the methods presented here to more realistic Galactic potentials and simulations.
 
-Truly resonant orbits are rare \citep{merritt99}, but dynamically important in systems with several degrees of freedom as they determine the structure of orbit-space.
-Stable, resonant orbits generate families of orbits that may behave or appear similarly, whereas unstable resonances are typically associated with regions of chaos.
-Not all orbits can be trivially associated with resonances due to the infinitely complex, hierarchical structure of resonances (though the importance or ``size'' of a resonance falls off inversely with the norm of its integer vector, $|\bs{m}|^{-\alpha}$, with $\alpha > 0$).
-In autonomous, two dof systems, stochastic regions are generally surrounded and separated by regular motion, thus chaotic orbits are confined.
-In three or more dof the chaotic regions will connect and overlap.
-This is the origin of Arnold diffusion \citep{arnold64}, which allows chaotic orbits to radically change shape by traversing the web of unstable resonances and are only strictly bounded to their own energy hypersurface.
-This phenomenon is perhaps easier to visualize in action-space: in two dof systems, if a given set of tori are destroyed due to potential perturbations, an orbit in this now chaotic region is still bounded between the surrounding tori due to energy conservation.
-In three dimensions, the extra degree of freedom allows chaotic orbits to ``escape'' this confinement.
-
-For small potential perturbations, many regular orbits survive and only small chaotic islands are introduced.
-As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed.
-Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori -- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.
-However, even orbits that are strictly chaotic may behave regularly over timescales much longer than, for example, the dynamical time of the system.
-Chaotic orbits can therefore be loosely classified by being either \emph{strongly} or \emph{weakly} chaotic: strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale (we quantify this distinction with the experiments of Section~\ref{sec:experiments}).
-This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{sec:lyap}) or frequency diffusion time (Section~\ref{sec:naff}).
-
-These local measures quantify the divergence of infinitesimally nearby orbits or the rate of frequency diffusion along an orbit.
-For an ensemble of orbits with finite spreads in energy or frequencies (e.g., tidal debris), each orbit will have a slightly different characteristic timescale, but we still expect mixing time of the ensemble to correlate with local measures of chaos of the ``parent'' orbit of the ensemble (Section~\ref{sec:ensemble}).
-The evolution of tidal debris will depend on the ensemble mixing rate of the parent orbit: chaotic mixing will cause the density of the debris to fall off more quickly than the typical power-law behavior expected for regular orbits \citep[e.g.,][]{merritt96, helmi99}.
-Since all known tidal streams have been discovered as kinematic overdensities of stars, this implies (1) [something about accretion history], (2) the long, thin tidal streams we see may represent a sampling from ``special'' orbits in the Galactic potential where the mixing time is long, and (3) previous assumptions that connect the length and size of streams to their evolution time may be invalid.
-
-This paper is organized as follows: In Section~\ref{sec:methods}, we outline the numerical methods used for the experiments of Section~\ref{sec:experiments}.
-In Section~\ref{sec:discussion}, we describe the connection between different measures of chaotic behavior, further discuss the implications of chaotic mixing, and comment on the applicability of the methods presented here to more realistic Galactic potentials and simulations.
+%[Does this go anywhere?]
+%Thin stellar streams form when the spreading of the debris is faster in some preferred direction \citep[one eigenvalue of the Hessian of the potential is large relative to others e.g.,][]{helmi99, sanders??}.
+This is generically the case for orbits with significant angular momentum in a spherical potential ($|L| > J_r$, orbits that we expect for form streams), but [...
+comments about expectations in more complex potentials] In a spherical or mildly axisymmetric potential, the debris can only really spread in a single dimension because somehow the constraint that ....
+Along chaotic orbits, Hessian not defined and frequencies diffuse, ...
 
@@
 We use the Dormand-Prince 8th-order Runge-Kutta scheme (TODO: cite Dormand 1981) to integrate orbits in the above potential.
-In particular, we use the Python wrapper to the FORTRAN routines provided in the Scipy package \citep{scipy}.
-For all orbits, we ensure that energy is conserved to $\Delta E/E_0 \leq 10^{-9}$ by the end of integration.
+For all orbits, we ensure that energy is conserved to $\Delta E/E_0 \leq 10^{-8}$ by the end of integration.
 Unless otherwise specified, orbits are integrated for $\sim$250 orbital periods with 500 steps per orbital period.
@@
 
+\section{Review of nonlinear dynamics}\label{sec:nldreview}
+
+For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (dof), the equations of motion are most simply expressed in angle-action coordinates \cite[e.g.,][]{goldstein80}.
+In these coordinates, the actions, $J_n$, are integrals of the motion, and the conjugate angle variables, $\theta_n$, increase linearly with time
+\begin{align}
+	\dot{\act}_n &= -\frac{\partial H}{\partial \theta_n} = 0\\
+	\act_n &= {\rm const.}\\
+	\dot{\theta}_n &= \frac{\partial H}{\partial \act_n} = \Omega_n(\act_1,\act_2,...,\act_n)\label{eq:aafreq}\\
+	\theta_n(t) &= \Omega_n t + \theta_n(0)
+\end{align}
+where $i=1,2,...N$, and $\Omega_n$ are the fundamental frequencies of the orbit.
+The topology of angle-action space is equivalent to that of the surface of an $N$-dimensional torus, and thus orbits are often referred to in terms of \emph{orbital tori}: each set of actions, $(\act_1,\act_2,...\act_N)$, uniquely ``label'' the torus associated with a given orbit and the angles specify the location on the surface of the $N$-torus.
+If a global transformation to angle-action coordinates exists, the potential is said to be \emph{integrable}.
+
+Galactic gravitational potentials will almost certainly not be globally integrable.
+Dark matter halos of galaxies are expected to be triaxial but even if they are close to axisymmetric or spherical, the superposition of components (e.g., disk, bulge, halo) will break global integrability.
+If a triaxial (three dof) potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
+\begin{equation}
+	H(\bs{\theta}, \bs{\act}) = H_0(\bs{\act}) + \epsilon H_1(\bs{\theta}, \bs{\act})
+\end{equation}
+then a large number of regular orbits will survive the perturbation (by way of the KAM theorem)\footnote{3-vectors are represented with bold symbols, e.g., $\bs{J}=(J_1,J_2,J_3)$.}.
+The tori that survive will in general be separated by regions of irregular or chaotic motion, and thus any transformations to angle-action coordinates can only be defined locally.
+Analytic transformations to angle-action coordinates are only known for a handful of simple, integrable potentials.
+Much effort has gone in to developing general methods for transforming either from position-velocity to angle-action coordinates, $(\bs{x},\bs{v})\rightarrow(\bs{\theta},\bs{J})$ \citep{many}, or vice-versa, $(\bs{\theta},\bs{J})\rightarrow(\bs{x},\bs{v})$ \citep{many}.
+
+The types of motion generic to triaxial systems can be split into several sub-categories (summarized in Figure~\ref{fig:orbit-tree}).
+Regular motion can be classified by the number of resonance relations obeyed \citep[e.g.,][]{lichtenberg83}: \emph{conditionally periodic} orbits obey no resonance relations, \emph{uni-resonant} orbits obey a single relation of the form $\bs{m}\cdot\bs{\Omega}=0$, and \emph{bi-resonant} orbits obey two resonance relations, $\bs{m}\cdot\bs{\Omega}=0$ and $\bs{n}\cdot\bs{\Omega}=0$, where $\bs{m}$ and $\bs{n}$ are integer vectors.
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{figures/orbit-tree.pdf}
+\caption{} \label{fig:orbit-tree}
+\end{center}
+\end{figure*}
+
+Truly resonant orbits are rare \citep{merritt99}, but dynamically important in systems with several degrees of freedom as they determine the structure of orbit-space.
+Stable, resonant orbits generate families of orbits that may behave or appear similarly, whereas unstable resonances are typically associated with regions of chaos.
+Not all orbits can be trivially associated with resonances due to the infinitely complex, hierarchical structure of resonances (though the importance or ``size'' of a resonance falls off inversely with the norm of its integer vector, $|\bs{m}|^{-\alpha}$, with $\alpha > 0$).
+In autonomous, two dof systems, stochastic regions are generally surrounded and separated by regular motion, thus chaotic orbits are confined.
+In three or more dof the chaotic regions will connect and overlap.
+This is the origin of Arnold diffusion \citep{arnold64}, which allows chaotic orbits to radically change shape by traversing the web of unstable resonances and are only strictly bounded to their own energy hypersurface.
+This phenomenon is perhaps easier to visualize in action-space: in two dof systems, if a given set of tori are destroyed due to potential perturbations, an orbit in this now chaotic region is still bounded between the surrounding tori due to energy conservation.
+In three dimensions, the extra degree of freedom allows chaotic orbits to ``escape'' this confinement.
+
+For small potential perturbations, many regular orbits survive and only small chaotic islands are introduced.
+As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed.
+Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori -- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.
+However, even orbits that are strictly chaotic may behave regularly over timescales much longer than, for example, the dynamical time of the system.
+Chaotic orbits can therefore be loosely classified by being either \emph{strongly} or \emph{weakly} chaotic: strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale.
+This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{sec:lyap}) or frequency diffusion time (Section~\ref{sec:naff}).
+
+These local measures quantify the divergence of infinitesimally nearby orbits or the rate of frequency diffusion along an orbit.
+For an ensemble of orbits with finite spreads in energy or frequencies (e.g., tidal debris), each orbit will have a slightly different characteristic timescale, but we still expect mixing time of the ensemble to correlate with local measures of chaos of the ``parent'' orbit of the ensemble (Section~\ref{sec:ensemble}).
+The evolution of tidal debris will depend on the ensemble mixing rate around the parent orbit: chaotic mixing will cause the density of the debris to fall off more quickly than the typical power-law behavior expected for regular orbits \citep[e.g.,][]{merritt96, helmi99}.
+
 \subsection{Lyapunov exponents} \label{sec:lyap}
 
+%The simplest and perhaps most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit.
+Defining the vector $\bs{w} = (q_1,...,q_N,p_1,...,p_N)$ for some set of canonical coordinates $(q_i,p_i)$, we can write Hamilton's equations as
+%\begin{equation}
+%	\dot{\bs{w}} = \bs{\mathcal{J}} \frac{\partial H}{\partial \bs{w}} = f(\bs{w}) \label{eq:ham}
+%\end{equation}
+%where $\bs{\mathcal{J}}$ is the $2N \times 2N$ canonical Poisson tensor (also called the symplectic matrix) defined by
+%\begin{equation}
+%	\bs{\mathcal{J}} = \left( \begin{array}{c:c} 0 & \ident \\ \hdashline -\ident & 0 \end{array} \right)
+%\end{equation}
+%with $N$-dimensional identity matrices $\ident$.
+If we consider a nearby phase-space position, $\bs{w}'$, separated from $\bs{w}$ by an infinitesimal deviation, $\delta \bs{w}$, such that $\bs{w}' = \bs{w} + \delta \bs{w}$.
+We can expand to linear order in the deviation about the parent orbit and write the equations of motion for the deviation as follows
+%\begin{align}
+%	\dot{\bs{w}}' &= f(\bs{w}')\\
+%	\dot{\bs{w}} + \dot{\delta \bs{w}} &= f(\bs{w} +  \delta \bs{w})\\
+%	\dot{\bs{w}} + \dot{\delta \bs{w}} &\approx f(\bs{w}) + \jac_{\bs{w}}\cdot\delta \bs{w} + \mathcal{O}(\delta \bs{w}^2)\\
+%	\dot{\delta \bs{w}} &\approx \jac_{\bs{w}} \cdot  \delta \bs{w}\label{eq:deviate}
+%\end{align}
+%where $\jac$ is the Jacobian of Eq.~\ref{eq:ham}, evaluated at the parent orbit (the Hessian of the Hamiltonian).
+For chaotic orbits, the maximum eigenvalue of the solution matrix to Eq.~\ref{eq:deviate} is positive real, leading to exponential divergence of nearby orbits.
+
+%The simplest and perhaps most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit.
+Defining the vector $\bs{w} = (q_1,...,q_N,p_1,...,p_N)$ for some set of canonical coordinates $(q_i,p_i)$, and a nearby phase-space position, $\bs{w}'$, separated from $\bs{w}$ by an infinitesimal deviation, $\delta \bs{w}$, such that $\bs{w}' = \bs{w} + \delta \bs{w}$, we can expand to linear order in the deviation about the parent orbit and write the equations of motion for the deviation as
+%\begin{align}
+%	\dot{\delta \bs{w}} &\approx \jac_{\bs{w}} \cdot  \delta \bs{w}\label{eq:deviate}
+%\end{align}
+%where $\jac$ is the Jacobian of Eq.~\ref{eq:ham}, evaluated at the parent orbit (the Hessian of the Hamiltonian).
+For chaotic orbits, the maximum eigenvalue of the solution matrix to Eq.~\ref{eq:deviate} is positive real, leading to exponential divergence of nearby orbits.
+
 The simplest and perhaps most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit.
-Defining the vector $\bs{w} = (q_1,...,q_N,p_1,...,p_N)$ for some set of canonical coordinates $(q_i,p_i)$, we can write Hamilton's equations as
+For some set of canonical coordinates $(q,p)$, we define the vector $w_i = (w_1,...w_{2N}) = (q_1,...,q_N,p_1,...,p_N)$ so we can write Hamilton's equations as
 \begin{equation}
-	\dot{\bs{w}} = \bs{\mathcal{J}} \frac{\partial H}{\partial \bs{w}} = f(\bs{w}) \label{eq:ham}
+	\dot{w}_i = \mathcal{J}_{ij}\,\frac{\partial H}{\partial w_j} \label{eq:ham}
 \end{equation}
-where $\bs{\mathcal{J}}$ is the $2N \times 2N$ canonical Poisson tensor (also called the symplectic matrix) defined by
+where $\mathcal{J}_{ij}$ is the $2N \times 2N$ canonical Poisson tensor (also called the symplectic matrix) defined by
 \begin{equation}
-	\bs{\mathcal{J}} = \left( \begin{array}{c:c} 0 & \ident \\ \hdashline -\ident & 0 \end{array} \right)
+	\mathcal{J}_{ij} = \left( \begin{array}{c:c} 0 & \ident \\ \hdashline -\ident & 0 \end{array} \right)
 \end{equation}
 with $N$-dimensional identity matrices $\ident$.
-If we consider a nearby phase-space position, $\bs{w}'$, separated from $\bs{w}$ by an infinitesimal deviation, $\delta \bs{w}$, such that $\bs{w}' = \bs{w} + \delta \bs{w}$.
-We can expand to linear order in the deviation about the parent orbit and write the equations of motion for the deviation as follows
+If we consider a nearby phase-space position, $w_i'$, separated from $w_i$ by an infinitesimal deviation, $\delta w_i$, such that $w_i' = w_i + \delta w_i$.
+We can expand to linear order in the deviation about the parent orbit and write the equations of motion for the deviation as
 \begin{align}
-	\dot{\bs{w}}' &= f(\bs{w}')\\
-	\dot{\bs{w}} + \dot{\delta \bs{w}} &= f(\bs{w} +  \delta \bs{w})\\
-	\dot{\bs{w}} + \dot{\delta \bs{w}} &\approx f(\bs{w}) + \jac_{\bs{w}}\cdot\delta \bs{w} + \mathcal{O}(\delta \bs{w}^2)\\
-	\dot{\delta \bs{w}} &\approx \jac_{\bs{w}} \cdot  \delta \bs{w}\label{eq:deviate}
+	\dot{\delta w_i} = \mathcal{J}_{ij} \frac{\partial^2 H}{\partial w_j \partial w_k} \delta w_k =  \mathcal{J}_{ij}\, D_{jk} \, \delta w_k \label{eq:deviate}
 \end{align}
-where $\jac$ is the Jacobian of Eq.~\ref{eq:ham}, evaluated at the parent orbit (the Hessian of the Hamiltonian).
+where $D_{jk}$ is the Hessian matrix evaluated at the parent orbit.
 For chaotic orbits, the maximum eigenvalue of the solution matrix to Eq.~\ref{eq:deviate} is positive real, leading to exponential divergence of nearby orbits.
@@
 For chaotic orbits, the divergence is exponential, and the limit should converge to the rate of the exponential: the Lyapunov exponent.
-
 In practice, the MLE is often estimated as the mean of $l_N$ after the summation diverges from power-law behavior.
-Figure~\ref{fig:lyap-orbits} shows $l_N$ computed for two nearby orbits in the potential of Section~\ref{sec:potential}.
-In physical units and Cartesian coordinates, the initial conditions for the two orbits are:
-\begin{align}
-	\bs{x}_{0}^{(1)} &= (27.85, 0, 23.35)~{\rm kpc}\\
-	\bs{v}_{0}^{(1)} &= (0, 115.37, 0)~{\rm km~s}^{-1}\\
-	\bs{x}_{0}^{(2)} &= (27.85, 0, 23.10)~{\rm kpc}\\
-	\bs{v}_{0}^{(2)} &= (0, 116.43, 0)~{\rm km~s}^{-1}.
-\end{align}
-Over-plotted on both panels are lines (dashed) that show $t^{-0.9}$ behavior.
-The MLE estimate continues towards zero for the regular orbit (top) but saturates for the chaotic orbit (bottom) and continues to increase at least until the end of our integration period.
-For mildly chaotic or ``sticky'' orbits, it can take integration periods of many hundreds or thousands of orbits to determine a reliable estimate of the Lyapunov exponent.
-
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=0.75\textwidth]{figures/lyap-orbits.pdf}
-\caption{ } \label{fig:lyap-orbits}
-\end{center}
-\end{figure*}
 
@@
 
-We have found that for ``sticky'' or mildly chaotic orbits, reliably computing the Lyapunov exponent can require integration times equivalent to thousands of orbital periods.
-In comparison, owing to the precision of NAFF, detecting small frequency diffusion rates can be done with only $\sim$100 orbital periods.
-Figure~\ref{fig:fdiff-orbits} shows the two strongest frequency components plotted against each other for the two orbits of Figure~\ref{fig:lyap-orbits}.
-The orbits were integrated for a total of 400 orbital periods and the frequencies were computed for 12 consecutive windows of orbit, where each section was 100 orbital periods in length.
-The frequencies are consistent for the regular orbit (left panel) whereas there is significant diffusion in the chaotic orbit (right panel).
+\subsection{Comparison of methods}
+
+We choose two nearby short-axis tube orbits in the potential described in Section~\ref{sec:potential} in order to demonstrate the two methods described above.
+The orbits were chosen from the grids of initial conditions described in later sections because they are very close in phase-space but display very different behavior.
+In particular, Orbit A is regular and Orbit B is stochastic.
+Figure~\ref{fig:two-orbits} shows the two orbits in projection -- Orbit A is shown in the left column and Orbit B the right.
+Two consecutive integration periods for each orbit are shown as the two rows -- note the change in appearance of Orbit B (the chaotic orbit, right column).
+The initial conditions for the two orbits in physical units and Cartesian coordinates are:
+\begin{align}
+	\bs{x}_{0}^{(A)} &= (27.85, 0, 23.35)~{\rm kpc}\\
+	\bs{v}_{0}^{(A)} &= (0, 115.37, 0)~{\rm km~s}^{-1}\\
+	\bs{x}_{0}^{(B)} &= (27.85, 0, 23.10)~{\rm kpc}\\
+	\bs{v}_{0}^{(B)} &= (0, 116.43, 0)~{\rm km~s}^{-1}.
+\end{align}
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/freq-diff-orbits.pdf}
+\includegraphics[width=0.9\textwidth]{figures/two-orbits.pdf}
+\caption{ } \label{fig:two-orbits}
+\end{center}
+\end{figure*}
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=0.75\textwidth]{figures/lyap-orbits.pdf}
+\caption{ } \label{fig:two-methods}
+\end{center}
+\end{figure*}
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=0.75\textwidth]{figures/freq-diff-orbits.pdf}
 \caption{ } \label{fig:fdiff-orbits}
@@
 
-\subsection{Test particle ensembles} \label{sec:ensemble}
+Figure~\ref{fig:two-methods} shows $l_N$ computed for the two orbits.
+Over-plotted on both panels are lines (dashed) that show $t^{-0.9}$ behavior.
+The MLE estimate continues towards zero for the regular orbit (top) but saturates for the chaotic orbit (bottom).
+For mildly chaotic or ``sticky'' orbits, it can take integration periods of many thousands of orbits to determine a reliable estimate of the Lyapunov exponent.
+In comparison, owing to the precision of NAFF, detecting small frequency diffusion rates can be done with only $\sim$100 orbital periods.
+Figure~\ref{fig:fdiff-orbits} shows the three fundamental frequencies computed for 12 consecutive windows of orbit (each section is $\sim$100 orbital periods in length) for Orbit A (left) and Orbit B (right).
+Dashed horizontal line is the estimate of the uncertainty in the numerical computation of the frequencies from Orbit A.
+The value of each frequency in the first window is subtracted from each subsequent window.
+The frequencies only show minor deviations for Orbit A (left panel) whereas there is significant diffusion in Orbit B (right panel).
 
-% TODO: below, should I put sarahs paper here even though we didn't say this explicitly?
-Our goal is to characterize the morphology of tidal debris along orbits generic to triaxial potentials.
-For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{fardal14, pearson15}, however, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
-Further, quantifying the degree of stochasticity with, for example, the frequency diffusion rate represents a \emph{local} measure of chaos and it is unclear how this generalizes to the small but finite spreads in orbital characteristics common to debris that is tidally disrupted from globular clusters or dwarf satellite galaxies.
-To answer this question would require a massive exploration of parameter space -- both in potential and in orbits -- and thus would require a prohibitive number of N-body simulations to truly analyze the disruption process in all situations.
+\section{Test particle ensembles} \label{sec:ensemble}
+
+The Lyapunov time and frequency diffusion rate are local measures of chaos but ...
 
@@
 For any computation of the density estimate, we use 10-fold cross-validation to find the optimal kernel bandwidth, $\sigma_{\rm KDE}$, by maximizing the likelihood of withheld particles.
-Figure~\ref{fig:DKL} shows $D_{\rm KL}$ as a function of time for a regular (black) and chaotic (blue) orbit.
-The chaotic orbit ensemble mixes much faster than the regular orbit ensemble.
+
+Figure~\ref{fig:DKL} shows $D_{\rm KL}$ as a function of time for a regular (black) and chaotic (blue) orbit. The chaotic orbit ensemble mixes much faster than the regular orbit ensemble.
 
@@
 
-\section{Experiments}\label{sec:experiments}
+\section{Results}\label{sec:results}
 
@@
 
+Our goal is to characterize the morphology of tidal debris along orbits generic to triaxial potentials.
+For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{fardal14, pearson15}, however, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
+Further, quantifying the degree of stochasticity with, for example, the frequency diffusion rate represents a \emph{local} measure of chaos and it is unclear how this generalizes to the small but finite spreads in orbital characteristics common to debris that is tidally disrupted from globular clusters or dwarf satellite galaxies.
+To answer this question would require a massive exploration of parameter space -- both in potential and in orbits -- and thus would require a prohibitive number of N-body simulations to truly analyze the disruption process in all situations.
+
 We measure the state of mixing of the debris as described above: we compute the KL divergence between the configuration-space density of the ensemble and the configuration-space density expected if the ensemble particles were uniformly distributed over the parent orbit energy hypersurface.
@@
 
+Since all known tidal streams have been discovered as kinematic overdensities of stars, this implies (1) [something about accretion history], (2) the long, thin tidal streams we see may represent a sampling from ``special'' orbits in the Galactic potential where the mixing time is long, and (3) previous assumptions that connect the length and size of streams to their evolution time may be invalid.
+
 Recovering the potential of triaxial systems is not trivial. Danger of orbit fitting because chaotic orbits can reproduce morphologies of regular orbits.
@@
 
+Moved from intro:
+This also has implications for modeling the accretion histories of galaxies.
+Debris morphology is often separated into two classes --- streams (usually from high angular momentum progenitor orbits) and shells (from radially plunging orbits) --- but in practice the distinction may not be as clear due to other mixing effects.
+
+\section{Conclusions}\label{sec:conclusions}
+For small ensembles, the frequency diffusion rate is a predictor of debris morphology.
+
 \acknowledgements
-The authors wish to thank Sarah Pearson, Andreas K\"upper, David Hogg, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
+The authors wish to thank Sarah Pearson, Andreas K\"upper, David Hogg, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 

### 2015-03-04 ms.tex
werq it

@@
 Here we choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos as computed from single-orbit diagnostics, and compare these results to measures of the mixing time of finite-volume ensembles of orbits (analogs to tidal debris).
-Finally, we connect the ensemble mixing rate to observational properties, thus demonstrating that local orbital measures of chaos can be used to predict the [...].
+Finally, we connect the ensemble mixing rate to observational properties of streams, thus demonstrating that local orbital measures of chaos are predictors of tidal stream morphology.
+This provides a strong case for developing a method to use the mere existence of the many thin streams found in the halo of our own Galaxy to rule out classes or shapes of possible potentials that fit the Milky Way.
 
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{figures/orbit-tree.pdf}
-\caption{} \label{fig:orbit-tree}
+\includegraphics[width=0.6\textwidth]{figures/orbit-tree.pdf}
+\caption{A rough classification of the types of orbits found in triaxial potentials.
+Regular orbits are classified by the number of resonance relations obeyed: conditionally periodic (\emph{none}), uni-resonant (\emph{one}), bi-resonant (\emph{two}).
+Irregular or chaotic orbits are simply separated into two classes: strongly and weakly chaotic, defined by the timescales of interest in a problem.
+When the characteristic chaotic timescale is much longer than, for example, the age of the universe, the orbit is only weakly chaotic.
+} \label{fig:orbit-tree}
 \end{center}
@@
 As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed.
-Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori -- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.
+Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori --- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.
 However, even orbits that are strictly chaotic may behave regularly over timescales much longer than, for example, the dynamical time of the system.
@@
 \begin{equation}
-	l_N = \frac{1}{N\tau}\sum_i^N \ln \, \delta_i
+	l_N = \frac{1}{N\tau}\sum_i^N \ln \, \delta_i \label{eq:ftmle}
 \end{equation}
@@
 \begin{equation}
-	\Phi(x,y,z) = \frac{1}{2}\ln\left(x^2 + (y/0.9)^2 + (z/0.7)^2 + 0.1\right).
+	\Phi(x,y,z) = \frac{1}{2}\ln\left(x^2 + (y/0.9)^2 + (z/0.7)^2 + 0.1\right). \label{eq:logpotential}
 \end{equation}
 In such a map, stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
+The regularity of the points in this map reflects the input grid of initial conditions.
+Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.
 
@@
 \includegraphics[width=0.5\textwidth]{figures/log-freqmap.png}
-\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:logfreqs}
+\caption{Ratio of cartesian frequencies for a grid of iso-energy box orbits integrated in the potential of Equation~\ref{eq:logpotential}.
+Stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
+The regularity of the points in this map reflects the input grid of initial conditions.
+Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.} \label{fig:logfreqs}
 \end{center}
@@
 
-\subsection{Comparison of methods}
+NAFF recovers the fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL96][]{papaphilippou96}.
+PL96 show that a good choice of coordinates for tube orbits are the Poincar\'e symplectic polar coordinates, a set of canonical coordinates similar to cylindrical coordinates.
+When computing the frequencies for tube orbits, we first rotated the orbits until the circulation is about the $z$-axis, transform to Poincar\'e polar coordinates, then use NAFF to measure the fundamental frequencies.
+We could equivalently use the Cartesian time series, but the convergence of terms is slower.
+We have tested that our implementation of NAFF returns the same frequencies in either case for a set of tube orbits.
+For box orbits, the motion is close to separable in each Cartesian component and we therefore use Cartesian coordinates for estimating frequencies for these orbits.
+
+\subsection{Comparison of methods}\label{sec:comparison}
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=0.9\textwidth]{figures/two-orbits.pdf}
+\caption{ Two orbits integrated in the triaxial potential described in Section~\ref{sec:potential}.
+Left panels (black) show two consecutive sections of orbit, where the integration time for each section (rows) is set to 100 times the maximum orbital period.
+Right panels (red) shows the same for the chaotic orbit.
+The appearance of the orbit in configuration space changes appreciably over timescales of tens of orbital periods.
+} \label{fig:two-orbits}
+\end{center}
+\end{figure*}
 
 We choose two nearby short-axis tube orbits in the potential described in Section~\ref{sec:potential} in order to demonstrate the two methods described above.
-The orbits were chosen from the grids of initial conditions described in later sections because they are very close in phase-space but display very different behavior.
-In particular, Orbit A is regular and Orbit B is stochastic.
-Figure~\ref{fig:two-orbits} shows the two orbits in projection -- Orbit A is shown in the left column and Orbit B the right.
-Two consecutive integration periods for each orbit are shown as the two rows -- note the change in appearance of Orbit B (the chaotic orbit, right column).
+The orbits were chosen from a grid of initial conditions in the $x$-$z$ plane (described in later sections) because they are very close in phase-space but display very different behavior.
+In particular, Orbit A is regular or at least weakly chaotic and Orbit B is strongly chaotic.
+Figure~\ref{fig:two-orbits} shows the two orbits in projection --- Orbit A is shown in the left column and Orbit B the right.
+Two consecutive integration periods for each orbit are shown in the two rows; note the change in appearance of Orbit B (right column).
 The initial conditions for the two orbits in physical units and Cartesian coordinates are:
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth]{figures/two-orbits.pdf}
-\caption{ } \label{fig:two-orbits}
+\includegraphics[width=0.62\textwidth]{figures/lyap-orbits.pdf}
+\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over 1000 orbital periods (the maximum orbital period) for Orbit A (left) and Orbit B (right).
+For Orbit A, the FTMLE asymptotes towards zero with a characteristic power-law slope.
+For Orbit B, the FTMLE begins to saturate to a constant; the Lyapunov exponent estimated for this orbit is $\lambda_{\rm max} \approx 0.07 \, {\rm Gyr}^{-1}$ corresponding to a Lyapunov time of $t_\lambda \approx 14 \, {Gyr}$.
+Over-plotted on both panels are lines (dashed, blue) that show $t^{-0.9}$ behavior.} \label{fig:lyap-orbits}
+\end{center}
+\end{figure*}
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=0.62\textwidth]{figures/freq-diff-orbits.pdf}
+\caption{ Difference in frequencies computed for successive sections of 100 orbital periods from the frequencies computed in the initial section for Orbit A (left) and Orbit B (right).
+The three lines in each panel are for the three fundamental frequencies.
+Dashed (red) horizontal line shows the estimate of the uncertainty for the frequencies computed for Orbit A.
+After many hundreds of orbital periods, the frequency deviation is not appreciable indicating again that the orbit is regular or only very weakly chaotic.
+The frequency deviations computed for Orbit B, however, show a clear increasing trend, indicating that the frequencies are diffusing with time.
+} \label{fig:fdiff-orbits} % The frequency diffusion rate defined in Equation~\ref{eq:fdrate} measures the initial slope of this trend.
 \end{center}
@@
 
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=0.75\textwidth]{figures/lyap-orbits.pdf}
-\caption{ } \label{fig:two-methods}
-\end{center}
-\end{figure*}
-
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=0.75\textwidth]{figures/freq-diff-orbits.pdf}
-\caption{ } \label{fig:fdiff-orbits}
-\end{center}
-\end{figure*}
-
-Figure~\ref{fig:two-methods} shows $l_N$ computed for the two orbits.
-Over-plotted on both panels are lines (dashed) that show $t^{-0.9}$ behavior.
-The MLE estimate continues towards zero for the regular orbit (top) but saturates for the chaotic orbit (bottom).
+Figure~\ref{fig:lyap-orbits} shows $l_N$ computed for the two orbits described above.
+The MLE estimate continues towards zero for the regular orbit (left) but saturates for the chaotic orbit (right).
 For mildly chaotic or ``sticky'' orbits, it can take integration periods of many thousands of orbits to determine a reliable estimate of the Lyapunov exponent.
@@
 
-The Lyapunov time and frequency diffusion rate are local measures of chaos but ...
+The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant \emph{local} to a given orbit.
+That is, they are measures of how infinitesimal deviations will diverge, on average, from some parent orbit of interest.
+Tidal debris is disrupted from progenitor systems with finite spreads in orbital properties (e.g., energy).
+For a disrupting, globular-cluster-scale progenitor, the typical energy dispersion of the debris is 0.1--1\% of the progenitor orbital energy, but for a dwarf-galaxy-scale progenitor, the dispersion can be 5--20\%.
+This raises an important consideration: do finite deviations diverge faster than predicted from local measures of chaos?
+Does the Lyapunov time or frequency rate predict the timescale over which a finite phase-space volume stays ``coherent''?
 
-We therefore take a simplified approach and consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event of a low-mass progenitor.
-For a given set of orbital initial conditions -- the parent orbit -- we find the position of the nearest pericenter, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all of these test particles for some evolution time, $T_{\rm ev}$.
-For a given mass scale, the ensemble particle position and velocity offsets are assumed to be Normal-distributed away from the parent orbit with spreads set by the tidal radius in position and velocity scale in velocity \citep[e.g.,][]{johnston98, apw14}.
-If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and is drawn from
+It is computationally intractable to run full N-body simulations of stellar system disruption for large grids of orbital initial conditions and thus we take a simplified approach for analyzing these questions.
+We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
+For a given set of orbital initial conditions --- the parent orbit --- we find the position of the nearest pericenter, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all of these test particles for some evolution time, $T_{\rm ev}$.
+The physical scale of the ensemble is set by the tidal radius in position and velocity scale in velocity and are therefore set by the mass scale of the progenitor \citep[e.g.,][]{johnston98, apw14}.
+If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
 \begin{align}
@@
 \end{align}
-where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, and $m$ is the mass scale of the debris, and $\|\cdot \|$ is the Euclidean norm.
-For this work, we take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass to estimate the above scales.
+where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the debris, and $\|\cdot \|$ is the Euclidean norm.
+For this work, we take $m=2\times10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass to estimate the above scales.
 
-As the particle ensembles evolve in time, we follow \citet[][hereafter MV96]{merritt96} and quantify the degree of mixing of the ensemble by computing the ``distance'' between the configuration space density of the ensemble at a given time, $\rho(\bs{x},t)$, with that of a distribution that is uniform over the energy hypersurface, $\rho_{E_0}(\bs{x})$, where $E_0$ is the energy of the parent orbit and thus hypersurface of interest.
+As the particle ensembles evolve in time, we follow a method similar to that of \citet[][hereafter MV96]{merritt96} and quantify the degree of mixing of the ensemble by computing the ``distance'' between the configuration space density of the ensemble at a given time, $\rho(\bs{x},t)$, with that of a distribution that is uniform over the energy hypersurface, $\rho_{E_0}(\bs{x})$, where $E_0$ is the energy of the parent orbit and thus hypersurface of interest.
 Our method differs in a few ways from that of MV96.
-First, MV96 use isoenergy ensembles of particles for following the density because that is the only way to guarantee that, if chaotic, nearly all orbits in the ensemble will lie on chaotic orbits.
-We instead use ensembles with small ($\approx$1\%) spreads in energy, but we still expect that such an ensemble around a parent chaotic orbit will reach a fully mixed state.
+First, MV96 use isoenergy ensembles of particles for following the density because this is the only way to guarantee that, if chaotic, nearly all orbits in the ensemble will lie on chaotic orbits.
+We instead use ensembles with small, finite spreads in energy, but we still expect that most ensemble particle orbits around a parent chaotic orbit will reach a fully mixed state.
 The final distribution, in addition to filling the energy hypersurface, will also have some thickness around this surface.
-Around a typical regular orbit, each individual orbit will fill its torus ergodically --- a surface with lower dimension than the energy hypersurface --- and thus, a small enough ensemble will never reach the fully mixed state as defined above because it will remain confined into a subspace on the energy surface.
-In practice, only strongly chaotic systems will approach $\rho_{E_0}(\bs{x})$; the rate will depend on how stochastic the parent orbit is (e.g., as measured by the Lyapunov time or frequency diffusion rate), and also on how ``trapped'' the chaotic region is (a small chaotic region wedged between stable tori will take longer to appear chaotic, even if internally the frequency diffusion rate is large).
+Around a typical regular orbit, each individual orbit will fill its torus ergodically --- a surface with lower dimension than the energy hypersurface --- and thus, a small enough ensemble will never reach the fully mixed state as defined above because it will remain confined into a subspace of lower dimensionality.
+In practice, only strongly chaotic systems will get close to $\rho_{E_0}(\bs{x})$; we expect the rate to depend on how stochastic the parent orbit is (e.g., as measured by the Lyapunov time or frequency diffusion rate) and also on how ``trapped'' the chaotic region is (a small chaotic region wedged between stable tori will take longer to appear chaotic, even if internally the frequency diffusion rate is large).
 However the difference in configuration space density between an ensemble undergoing chaotic mixing versus an ensemble around a regular orbit may be detectable over a much shorter timescale.
@@
 	&= 4\pi \int \dd v \, v^2 \, \rho_0 \, \delta\left(v^2/2 + \Phi(\bs{x}) - E_0\right)\\
-	&= A \sqrt{E_0 - \Phi(\bs{x})}
+	&= A \sqrt{E_0 - \Phi(\bs{x})}\label{eq:mixeddensity}
 \end{align}
@@
 
-Second, we compute the Kullback-Leibler divergence \citep[KL divergence;][]{kullback51} by estimating the configuration space density $\rho(\bs{x},t_j)$ using kernel density estimation (KDE)\footnote{We use an implementation from the Python package scikit-learn \citep{scikitlearn}.} with the ensemble of particle positions at time $t_j$, then use Monte Carlo integration to evaluate the KL divergence rather than evaluate the integral over a three dimensional grid.
-The KL divergence, $D_{\rm KL}$, is a measure of the relative Shannon entropy between two normalized density functions, $p(x)$ and $q(x)$,
+For each ensemble, we compute the Kullback-Leibler divergence \citep[KL divergence;][]{kullback51} between the configuration-space density and the fully-mixed density (Equation~\ref{eq:mixeddensity}) by estimating the configuration space density $\rho(\bs{x},t_j)$ using kernel density estimation (KDE)\footnote{We use an implementation from the Python package scikit-learn \citep{scikitlearn}.} with the ensemble of particle positions at a given time $t_j$, We use Monte Carlo integration to evaluate the KL divergence rather than evaluate the integral over a three dimensional grid.
+The KL divergence, $D_{\rm KL}$, is a measure of the relative (Shannon) entropy between two normalized density functions, $p(x)$ and $q(x)$,
 \begin{equation}
@@
 For any computation of the density estimate, we use 10-fold cross-validation to find the optimal kernel bandwidth, $\sigma_{\rm KDE}$, by maximizing the likelihood of withheld particles.
+[TODO: previous sentence?]
 
-Figure~\ref{fig:DKL} shows $D_{\rm KL}$ as a function of time for a regular (black) and chaotic (blue) orbit. The chaotic orbit ensemble mixes much faster than the regular orbit ensemble.
+Figure~\ref{fig:DKL} shows $D_{\rm KL}$ as a function of time for the two orbits of Section~\ref{sec:comparison}.
+The chaotic orbit (Orbit B) ensemble mixes much faster than the regular orbit (Orbit A) ensemble, but note the timescale here is still over many tens of orbital periods.
+The KLD is related to the log-density of the debris so that the final difference in KLD shown in Figure~\ref{fig:DKL} corresponds to a factor of 10 difference in density.
+TODO: From the decay of the KLD, we measure a mixing rate, defined as [...] TODO: what is the mixing rate?
+etc...
 
@@
 We start by studying the properties of isoenergy orbits in the potential described above.
-We generate grids of initial condition along the $xz$ ($y=0$) and $yz$ ($x=0$) planes\footnote{Recall that $x$ is the major, $y$ the intermediate, and $z$ the minor axis.} all with the same energy, $E=XXX$.
-For the $xz$ orbits, we fix $v_x = v_z = 0$, and compute $v_y$ from the energy.
-Similarly, for the $yz$ orbits, we fix $v_y = v_z = 0$, and compute $v_x$ from the energy.
-We integrate the orbits as described in Section~\ref{sec:integration} and compute the fundamental frequencies of each orbit with NAFF.
-Figure~\ref{fig:freqmap} shows frequency maps for the $xz$ (left) and $yz$ (right) grids.
-For short-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $z$ axis, $\Omega_\phi$, and the oscillation along the $z$ axis, $\Omega_z$.
-For long-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $x$ axis, $\Omega_\phi$, and the oscillation along the $x$ axis, $\Omega_x$.
-Box orbits are not shown.
-Already evident in these panels is that many of the orbits are regular: the transformed, even-spaced sampling of the initial conditions is preserved in the frequency transformation, implying that the frequencies for many orbits do not change appreciably.
-Also apparent is resonant trapping,  which appear as over-dense, linear features in these frequency maps.
+We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy $E=XXX$.
+We fix $v_x = v_z = 0$, and compute $v_y$ from the energy.
+The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic.
+Tube orbits are generally less stochastic than box orbits --- the other major class of orbits in triaxial potentials --- which tend to plunge deep into the inner regions of the potential.
+Thin tidal streams are thought to form along tube orbits rather than box orbits because of the fast disruption expected for stellar systems on radially plunging orbits.
+The major classes of stable tube orbits circulate about either the major or minor axis.
+The most prominent orbits in this grid are the short-axis and outer long-axis tubes, but this grid generates all of the major orbit classes: short-axis tubes, long-axis inner tubes, long-axis outer tubes, stochastic intermediate-axis, and box orbits.
+[TODO: Stochastic tube orbits not well studied...because not very important here...]
 
-The majority of the orbits generated on these grids are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic.
-Tube orbits are generally less stochastic than box orbits -- the other major class of orbits in triaxial potentials -- which tend to plunge deep into the inner regions of the potential.
-Thin tidal streams are thought to form along tube orbits rather than box orbits because of the fast disruption expected for stellar systems on radially plunging orbits.
-The major classes of stable tube orbits circulate about either the major or minor axis; the most prominent orbits are the short-axis and outer long-axis tubes, but these grids do generate some inner long-axis tubes and box orbits.
-Orbits that try to circulate about the intermediate axis tend to be highly irregular and do not preserve their sense of circulation for long periods of time.
-NAFF recovers fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL96][]{papaphilippou96}.
-PL96 show that a good choice of coordinates for tube orbits are the Poincar\'e symplectic polar coordinates, a set of canonical coordinates similar to cylindrical coordinates.
+We integrate all orbits for $\approx$200 orbital periods (Section~\ref{sec:integration}) and use NAFF to compute the fundamental frequencies for the two consecutive sections of 100 orbital periods.
+Figure~\ref{fig:freqmap} (left) shows a map of the fundamental frequency ratios for all tube orbits from the input grid (we plot the frequencies estimated from the first 100 orbital periods).
+Already evident in this panel is that many of the orbits are regular: the transformed, even-spaced sampling of the initial conditions is preserved in the frequency transformation, implying that the frequencies for many orbits do not change appreciably.
+Also apparent is resonant trapping, which appear as over-dense, linear features.
 For approximately XX\% of the orbits, NAFF fails to recover the frequencies (similar difficulties were reported in \cite{valluri98} and \cite{papaphilippou96} for recovering certain tube orbit frequencies).
@@
 	\includegraphics[width=0.45\columnwidth] {figures/freqmap_xz.png}
-	\includegraphics[width=0.45\columnwidth] {figures/freqmap_yz.png}
-	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+	\includegraphics[width=0.45\columnwidth] {figures/diffusion_map_xz.png}
+	\caption{(\emph{Left} Frequency map of ...
+For short-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $z$ axis, $\Omega_\phi$, and the oscillation along the $z$ axis, $\Omega_z$.
+For long-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $x$ axis, $\Omega_\phi$, and the oscillation along the $x$ axis, $\Omega_x$.
+Box orbits are not shown.}
 	\label{fig:freqmap}
@@
 
-For each orbit, we also compute the frequency diffusion rate per orbit, $\fdrate$ (Equation~\ref{eq:fdrate}) by computing the frequencies for the first half and second half of the orbital time series separately.
-This rate is only a measure of the time-averaged diffusion rate because the frequencies are computed for long segments of orbit, but \cite{valluri??} [...].
-Figure~\ref{fig:diffusionmap} shows maps of the frequency diffusion for the $xz$ (left) and $yz$ (right) grids: plotted are the initial conditions for all orbits colored by $\log_{10}(\fdrate)$ (white points are strongly chaotic, black are regular).
-The rich structure of unstable resonances apparent in these frequency diffusion maps highlight the accuracy of NAFF.
-It seems that tube orbits in these potentials are mostly regular or only mildly chaotic, however islands of stronger chaos do appear, especially at the intersections of unstable resonances.
+For each orbit, we also compute the frequency diffusion rate per orbit, $\fdrate$ (Equation~\ref{eq:fdrate}), using the frequencies computed for the two integrated sections of orbit.
+This rate only measures the time-averaged diffusion because the frequencies are computed for long segments of orbit, but \cite{valluri??} [...].
+Figure~\ref{fig:freqmap} (right) shows the grid of initial conditions for all orbits colored by $\log_{10}(\fdrate)$ (white points are strongly chaotic, black are regular).
+A rich structure of resonances is apparent in this frequency diffusion map and highlights the accuracy of NAFF.
+From this map we again see that the tube orbits in this potential are mostly regular or only mildly chaotic, however islands of stronger chaos do appear, especially at the intersections of unstable resonances.
 The bright, unstable regions in these maps are a visualization of the Arnold web at this energy.
-It is evident that these unstable regions are interconnected and therefore a tube orbit that starts, for example, in the lower left-hand region of one of these maps could  theoretically end up as a box orbit (along the zero-velocity curve) at some point in it its history (but the timescale for this to happen is likely prohibitively long).
-Figure~\ref{fig:diffusionrates} shows histograms of $\log_{10}(\fdrate)$ for all orbits in each initial condition grid for box and tube orbits separately.
-[...1-2 sentences discussion this (need to wait to see final figure)]
-
-\begin{figure}[h]
-\centering
-	\includegraphics[width=0.45\columnwidth] {figures/diffusion_map_xz.png}
-	\includegraphics[width=0.45\columnwidth] {figures/diffusion_map_yz.png}
-	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
-	\label{fig:diffusionmap}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-	\includegraphics[width=0.45\columnwidth] {figures/diffusion_rate_hist_xz.pdf}
-	\includegraphics[width=0.45\columnwidth] {figures/diffusion_rate_hist_yz.pdf}
-	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc. Separate box from tube orbits.} 
-	\label{fig:diffusionrates}
-\end{figure}
+It is evident that these unstable regions are interconnected and therefore a tube orbit that starts, for example, in the lower left-hand region of one of these maps could theoretically end up as a box orbit at some point in it its history (but the timescale for this to happen is likely prohibitively long).
 
@@
 
+[TODO: move lyapunov stuff to above FDR stuff?]
 We also compute the Lyapunov exponents for all orbits in the grid by integrating their orbits for 250 orbital periods.
@@
 NAFF is able to detect small differences in the frequencies and thus measure small diffusion rates with comparable integration times.
-
-Figure~\ref{fig:xx} shows a sample of regular orbits from these two grids, plotted in projection.
-The left column are orbits from the $xz$ grid and the right column shows orbits from the $yz$ grid.
-Orbits near a resonance are labeled with the integer vector associated with the nearest resonance.
-Figures~\ref{fig:chaotic1} and \ref{fig:chaotic2} show two chaotic orbits, one with a large $\fdrate$ and one with a small $\fdrate$ (labeled in the figure titles).
-The figures show projections of the orbit in three consecutive sections of orbit with $\sim$50 periods shown per panel.
-
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/chaotic_orbit1.png}
-\caption{Just a placeholder figure that has a similar shape to what I will put here} \label{fig:chaotic1}
-\end{center}
-\end{figure*}
-
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/chaotic_orbit2.png}
-\caption{Just a placeholder figure that has a similar shape to what I will put here} \label{fig:chaotic2}
-\end{center}
-\end{figure*}
 
@@
 Further, quantifying the degree of stochasticity with, for example, the frequency diffusion rate represents a \emph{local} measure of chaos and it is unclear how this generalizes to the small but finite spreads in orbital characteristics common to debris that is tidally disrupted from globular clusters or dwarf satellite galaxies.
-To answer this question would require a massive exploration of parameter space -- both in potential and in orbits -- and thus would require a prohibitive number of N-body simulations to truly analyze the disruption process in all situations.
+To answer this question would require a massive exploration of parameter space --- both in potential and in orbits --- and thus would require a prohibitive number of N-body simulations to truly analyze the disruption process in all situations.
 
@@
 a:
-This measure can be computed at any time -- [we show two different maps of this quantity: one after 10 Gyr, one after 15 orbits?]
+This measure can be computed at any time --- [we show two different maps of this quantity: one after 10 Gyr, one after 15 orbits?]
 
@@
 
-Only effectively looking at debris disrupted from a single pericenter -- debris near progenitor constantly repopulated as more and more stuff disrupts out
+Only effectively looking at debris disrupted from a single pericenter --- debris near progenitor constantly repopulated as more and more stuff disrupts out
 
@@
 \section{Discussion}\label{sec:discussion}
+
+We have shown that local measures of chaos [...]
 

### 2015-03-08 ms.tex
comment out abstract

@@
 
-Tidal streams form from the steady disruption of stellar systems
-orbiting within the gravitational field of some parent galaxy. Many
-streams and debris structures have been discovered in the halo of the
-Milky Way and have been used to model the potential of the Galaxy. 
-However, few of these models have yet explored the properties of
-tidal debris in triaxial potentials.
-The existence of a variety of orbits, resonances, and chaotic regions in such potentials 
-suggest that the morphologies and dispersal timescales of
-debris could differ significantly from the simpler spherical and oblate cases.
-In this work we use a series of N-body simulations of stellar systems over a range 
-of masses of disruption in triaxial potentials to understand the influence of the nature 
-and types of orbits on debris morphologies.
-Our results suggest that the mere existence of the multitude of thin streams
-already known to orbit the Milky Way provides significant constraints on the
-classes of triaxial potentials that provide a good representation for its dark matter halo.
+%Tidal streams form from the steady disruption of stellar systems
+%orbiting within the gravitational field of some parent galaxy. Many
+%streams and debris structures have been discovered in the halo of the
+%Milky Way and have been used to model the potential of the Galaxy. 
+%However, few of these models have yet explored the properties of
+%tidal debris in triaxial potentials.
+%The existence of a variety of orbits, resonances, and chaotic regions in such potentials 
+%suggest that the morphologies and dispersal timescales of
+%debris could differ significantly from the simpler spherical and oblate cases.
+%In this work we use a series of N-body simulations of stellar systems over a range 
+%of masses of disruption in triaxial potentials to understand the influence of the nature 
+%and types of orbits on debris morphologies.
+%Our results suggest that the mere existence of the multitude of thin streams
+%already known to orbit the Milky Way provides significant constraints on the
+%classes of triaxial potentials that provide a good representation for its dark matter halo.
 
@@
 Figure~\ref{fig:DKL} shows $D_{\rm KL}$ as a function of time for the two orbits of Section~\ref{sec:comparison}.
-The chaotic orbit (Orbit B) ensemble mixes much faster than the regular orbit (Orbit A) ensemble, but note the timescale here is still over many tens of orbital periods.
-The KLD is related to the log-density of the debris so that the final difference in KLD shown in Figure~\ref{fig:DKL} corresponds to a factor of 10 difference in density.
-TODO: From the decay of the KLD, we measure a mixing rate, defined as [...] TODO: what is the mixing rate?
-etc...
+The chaotic orbit (Orbit B) ensemble mixes much faster than the regular orbit (Orbit A) ensemble.
+The KLD is related to the log-density of the debris so that even after XX, the difference in the KLD for the two orbits corresponds to a factor of 10 difference in density.
+Figure XX shows particle positions at apocenter after $\approx$10 Gyr of evolution for ensembles of 1000 particles generated around the two orbits.
+For visualization, the final positions are rotated to lie on the $x$ axis and the angular momentum vector of the parent orbit is aligned with the $z$ axis.
+
+From the decay of the KLD, we measure a mixing rate by [TODO: how?].
+
+ TODO: From the decay of the KLD, we measure a mixing rate, defined as [...] TODO: what is the mixing rate? etc...
 
@@
 \acknowledgements
-The authors wish to thank Sarah Pearson, Andreas K\"upper, David Hogg, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
+The authors wish to thank Sarah Pearson, Robyn Sanderson, Andreas K\"upper, David Hogg, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 

### 2015-03-22 ms.tex
added monte carlo blip
explain the projection of distribution function better

@@
 PL96 show that a good choice of coordinates for tube orbits are the Poincar\'e symplectic polar coordinates, a set of canonical coordinates similar to cylindrical coordinates.
-When computing the frequencies for tube orbits, we first rotated the orbits until the circulation is about the $z$-axis, transform to Poincar\'e polar coordinates, then use NAFF to measure the fundamental frequencies.
-We could equivalently use the Cartesian time series, but the convergence of terms is slower.
-We have tested that our implementation of NAFF returns the same frequencies in either case for a set of tube orbits.
+When computing the frequencies for tube orbits, we first rotate the orbits until the circulation is about the $z$-axis, transform to Poincar\'e polar coordinates, then use NAFF to measure the fundamental frequencies.
+We could equivalently use the Cartesian time series, but the convergence of terms is slower (the amplitudes of successive terms decrease slower for Cartesian coordinates).
+We have tested that our implementation of NAFF returns the same fundamental frequencies in either case for a set of tube orbits.
 For box orbits, the motion is close to separable in each Cartesian component and we therefore use Cartesian coordinates for estimating frequencies for these orbits.
@@
 
-The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant \emph{local} to a given orbit.
-That is, they are measures of how infinitesimal deviations will diverge, on average, from some parent orbit of interest.
+The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant \emph{local} to a given orbit --- that is, these quantities are measures of how infinitesimal deviations will diverge, on average, from some parent orbit of interest.
 Tidal debris is disrupted from progenitor systems with finite spreads in orbital properties (e.g., energy).
@@
 We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
-For a given set of orbital initial conditions --- the parent orbit --- we find the position of the nearest pericenter, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all of these test particles for some evolution time, $T_{\rm ev}$.
+For a given set of orbital initial conditions --- the ``parent'' orbit --- we find the position of the nearest pericenter, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all of these test particles for an integration time equal to 256 times the longest fundamental orbital period.
 The physical scale of the ensemble is set by the tidal radius in position and velocity scale in velocity and are therefore set by the mass scale of the progenitor \citep[e.g.,][]{johnston98, apw14}.
@@
 where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the debris, and $\|\cdot \|$ is the Euclidean norm.
-For this work, we take $m=2\times10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass to estimate the above scales.
+For this work, we take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass to estimate the above debris scales.
 
-As the particle ensembles evolve in time, we follow a method similar to that of \citet[][hereafter MV96]{merritt96} and quantify the degree of mixing of the ensemble by computing the ``distance'' between the configuration space density of the ensemble at a given time, $\rho(\bs{x},t)$, with that of a distribution that is uniform over the energy hypersurface, $\rho_{E_0}(\bs{x})$, where $E_0$ is the energy of the parent orbit and thus hypersurface of interest.
+As the particle ensembles evolve in time, we follow a method similar to that of \citet[][hereafter MV96]{merritt96} to quantify the degree of mixing of the ensemble by computing the ``distance'' between the configuration space density of the ensemble at a given time, $\rho(\bs{x},t)$, with that of a distribution that is uniformly distributed over the energy hypersurface, $\rho_{E_0}(\bs{x})$, where $E_0$ is the energy of the parent orbit and thus hypersurface of interest.
 Our method differs in a few ways from that of MV96.
 First, MV96 use isoenergy ensembles of particles for following the density because this is the only way to guarantee that, if chaotic, nearly all orbits in the ensemble will lie on chaotic orbits.
-We instead use ensembles with small, finite spreads in energy, but we still expect that most ensemble particle orbits around a parent chaotic orbit will reach a fully mixed state.
-The final distribution, in addition to filling the energy hypersurface, will also have some thickness around this surface.
-Around a typical regular orbit, each individual orbit will fill its torus ergodically --- a surface with lower dimension than the energy hypersurface --- and thus, a small enough ensemble will never reach the fully mixed state as defined above because it will remain confined into a subspace of lower dimensionality.
-In practice, only strongly chaotic systems will get close to $\rho_{E_0}(\bs{x})$; we expect the rate to depend on how stochastic the parent orbit is (e.g., as measured by the Lyapunov time or frequency diffusion rate) and also on how ``trapped'' the chaotic region is (a small chaotic region wedged between stable tori will take longer to appear chaotic, even if internally the frequency diffusion rate is large).
-However the difference in configuration space density between an ensemble undergoing chaotic mixing versus an ensemble around a regular orbit may be detectable over a much shorter timescale.
-The fully-mixed, normalized, configuration-space density that fills a given energy hypersurface, $E_0$, is
+We instead use ensembles with small but finite spreads in energy; we still expect that most ensemble particle orbits around a parent chaotic orbit will reach a fully mixed state but now the final distribution will have some thickness around this surface.
+Around a typical regular orbit, each individual orbit will fill its torus ergodically --- a surface with lower dimension than the energy hypersurface --- and thus, a small enough ensemble should never reach the fully mixed state as defined above because it will remain confined to a subspace of lower dimensionality.
+In practice, only strongly chaotic systems will get close to $\rho_{E_0}(\bs{x})$ in a reasonable amount of time; we expect the rate to depend on how stochastic the parent orbit is (e.g., as measured by the Lyapunov time or frequency diffusion rate) and also on how ``trapped'' the chaotic region is (a small chaotic region wedged between stable tori will take longer to appear chaotic, even if internally the frequency diffusion rate is large).
+However, the difference in configuration space density between an ensemble undergoing chaotic mixing versus an ensemble around a regular orbit may be detectable over a much shorter timescale.
+The fully-mixed, normalized, configuration-space density that fills a given energy hypersurface, $E_0$, is easily found by projecting the simple distribution function, $f$, into configuration space:
 \begin{align}
-	\rho_{E_0}(\bs{x}) &= \int \dd^3v \, \rho_0 \, \delta(H - E_0)\\
-	&= \int \dd^3v \, \rho_0 \, \delta\left(|\bs{v}|^2/2 + \Phi(\bs{x}) - E_0\right)\\
-	&= 4\pi \int \dd v \, v^2 \, \rho_0 \, \delta\left(v^2/2 + \Phi(\bs{x}) - E_0\right)\\
+	f(\bs{x}, \bs{v}) &= f_0 \, \delta(H(\bs{x}, \bs{v}) - E_0)\\
+	\rho_{E_0}(\bs{x}) &= \int \dd^3v \, f(\bs{x}, \bs{v})\\
+	&= \int \dd^3v \, f_0 \, \delta(H - E_0)\\
+	&= 4\pi \int \dd v \, v^2 \, f_0 \, \delta\left(v^2/2 + \Phi(\bs{x}) - E_0\right)\\
 	&= A \sqrt{E_0 - \Phi(\bs{x})}\label{eq:mixeddensity}
@@
 
+\begin{align}
+	E[\rho(\bs{x},t_j)] &= \int \rho(\bs{x},t_j) \, \rho(\bs{x},t_j) \, \dd \bs{x}\\
+	&\approx \frac{1}{N}\sum \rho(\bs{x}_i,t_j)
+\end{align}
+So we can actually estimate the mean density by evaluating at the particles and taking the mean.
+
 \section{Discussion}\label{sec:discussion}

### 2015-03-23 ms.tex
updated ensemble stuff

@@
 \begin{equation}
-	\fdrate = \max_{\bs{a}}\,2(\bs{\Omega}_{T_1} - \bs{\Omega}_{T_2}) / \inttime \label{eq:fdrate}
+	\fdrate = \max_{\bs{a}} \, (\bs{\Omega}_{T_1} - \bs{\Omega}_{T_2}) / N_T \label{eq:fdrate}
 \end{equation}
-where $\bs{\Omega}_{T_1}$ and $\bs{\Omega}_{T_2}$ are the fundamental frequencies computed from two consecutive sections of orbit, $\inttime$ is the total integration time (over both sections), and the maximum is taken with relation to the corresponding amplitudes, $\bs{a}$, of the frequency components.
+where $\bs{\Omega}_{T_1}$ and $\bs{\Omega}_{T_2}$ are the fundamental frequencies computed from two consecutive sections of orbit, $N_T$ is the number of orbital periods within a single section of orbit computed from the most significant fundamental frequency, and the maximum is taken with respect to the corresponding amplitudes, $\bs{a}$, of the frequency components.
 The frequency diffusion rate as defined above (e.g., measured in the frequency with the largest amplitude) is better correlated with the strength of chaos in an orbit \citep{valluri??} but differs from the original definition of \cite{valluri98}.
@@
 \end{equation}
-We would like to compute 
+Instead, we are computing the three dimensional integral
 \begin{align}
@@
 \end{align}
-however, we only have an estimate of the density $\rho_{\rm KDE}(\bs{x},t_j) \approx \rho(\bs{x},t_j)$. Using samples from this KDE estimated density, $\bs{x}_i$, the KL divergence is
+where we only have an estimate of the density $\rho_{\rm KDE}(\bs{x},t_j) \approx \rho(\bs{x},t_j)$. Using samples from this KDE estimated density, $\bs{x}_i$, the KL divergence is
 \begin{align}
@@
 For any computation of the density estimate, we use 10-fold cross-validation to find the optimal kernel bandwidth, $\sigma_{\rm KDE}$, by maximizing the likelihood of withheld particles.
-[TODO: previous sentence?]
 
 Figure~\ref{fig:DKL} shows $D_{\rm KL}$ as a function of time for the two orbits of Section~\ref{sec:comparison}.
-The chaotic orbit (Orbit B) ensemble mixes much faster than the regular orbit (Orbit A) ensemble.
-The KLD is related to the log-density of the debris so that even after XX, the difference in the KLD for the two orbits corresponds to a factor of 10 difference in density.
-Figure XX shows particle positions at apocenter after $\approx$10 Gyr of evolution for ensembles of 1000 particles generated around the two orbits.
-For visualization, the final positions are rotated to lie on the $x$ axis and the angular momentum vector of the parent orbit is aligned with the $z$ axis.
+The chaotic orbit (Orbit B, dashed line) ensemble mixes much faster than the regular orbit (Orbit A, solid line) ensemble.
+The KLD is related to the logarithm of the mean density of the debris so that a difference of 1 in the KLD roughly corresponds to a factor of 10 difference in mean density of the debris.
+Figure XX shows particle positions at apocenter after 2, 16, and 32 orbital periods of evolution for ensembles of 1000 particles generated around Orbit A and B.
+For visualization, the final positions are rotated to lie on the $x$ axis and the angular momentum vector of the parent orbit at the given time is aligned with the $z$ axis.
 
 From the decay of the KLD, we measure a mixing rate by [TODO: how?].
-
- TODO: From the decay of the KLD, we measure a mixing rate, defined as [...] TODO: what is the mixing rate? etc...
 
@@
 
-We start by studying the properties of isoenergy orbits in the potential described above.
+We start by studying the properties of individual orbits in the potential described above.
 We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy $E=XXX$.
@@
 The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic.
-Tube orbits are generally less stochastic than box orbits --- the other major class of orbits in triaxial potentials --- which tend to plunge deep into the inner regions of the potential.
+Tube orbits are generally less stochastic than box orbits --- the other major class of orbits in triaxial potentials --- which tend to plunge deep into the inner regions of the potential, however there are appreciable regions of stochastic tube orbits.
 Thin tidal streams are thought to form along tube orbits rather than box orbits because of the fast disruption expected for stellar systems on radially plunging orbits.
@@
 The most prominent orbits in this grid are the short-axis and outer long-axis tubes, but this grid generates all of the major orbit classes: short-axis tubes, long-axis inner tubes, long-axis outer tubes, stochastic intermediate-axis, and box orbits.
-[TODO: Stochastic tube orbits not well studied...because not very important here...]
 
-We integrate all orbits for $\approx$200 orbital periods (Section~\ref{sec:integration}) and use NAFF to compute the fundamental frequencies for the two consecutive sections of 100 orbital periods.
+We compute the Lyapunov exponents for all orbits in the grid by integrating the orbits for 1000 orbital periods and use the method described in Section~\ref{sec:lyap} to estimate the maximum Lyapunov exponent.
+Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $x--z$ plane) --- each pixel corresponds to an orbit, and the greyscale intensity is related to the logarithm of the Lyapunov time, $t_\lambda = \lambda_{\rm max}^{-1}$.
+This map shows that by numerically estimating the MLE over this fixed integration window, the majority of orbits appear to be regular or have long Lyapunov times.
+
+\begin{figure}[!h]
+\centering
+	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_xz.png}
+	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+	\label{fig:lyapmap}
+\end{figure}
+
+For all orbits, we also integrate instead for $\approx$200 orbital periods (Section~\ref{sec:integration}) and use NAFF to compute the fundamental frequencies for the two consecutive sections of 100 orbital periods.
 Figure~\ref{fig:freqmap} (left) shows a map of the fundamental frequency ratios for all tube orbits from the input grid (we plot the frequencies estimated from the first 100 orbital periods).
-Already evident in this panel is that many of the orbits are regular: the transformed, even-spaced sampling of the initial conditions is preserved in the frequency transformation, implying that the frequencies for many orbits do not change appreciably.
-Also apparent is resonant trapping, which appear as over-dense, linear features.
-For approximately XX\% of the orbits, NAFF fails to recover the frequencies (similar difficulties were reported in \cite{valluri98} and \cite{papaphilippou96} for recovering certain tube orbit frequencies).
+This panel demonstrates again that many of the orbits are regular: the transformed, even-spaced sampling of the initial conditions is preserved in the frequency transformation, implying that the frequencies for many orbits do not change appreciably.
+Also apparent is resonant trapping, which manifests as over-dense, linear features in the frequency map.
+For approximately 2\% of the orbits, NAFF fails to recover the frequencies (similar difficulties were reported in \cite{valluri98} and \cite{papaphilippou96} for recovering certain tube orbit frequencies).
 
@@
 For each orbit, we also compute the frequency diffusion rate per orbit, $\fdrate$ (Equation~\ref{eq:fdrate}), using the frequencies computed for the two integrated sections of orbit.
-This rate only measures the time-averaged diffusion because the frequencies are computed for long segments of orbit, but \cite{valluri??} [...].
 Figure~\ref{fig:freqmap} (right) shows the grid of initial conditions for all orbits colored by $\log_{10}(\fdrate)$ (white points are strongly chaotic, black are regular).
-A rich structure of resonances is apparent in this frequency diffusion map and highlights the accuracy of NAFF.
-From this map we again see that the tube orbits in this potential are mostly regular or only mildly chaotic, however islands of stronger chaos do appear, especially at the intersections of unstable resonances.
-The bright, unstable regions in these maps are a visualization of the Arnold web at this energy.
-It is evident that these unstable regions are interconnected and therefore a tube orbit that starts, for example, in the lower left-hand region of one of these maps could theoretically end up as a box orbit at some point in it its history (but the timescale for this to happen is likely prohibitively long).
+A rich structure of resonances appears in this frequency diffusion map that were not seen in the Lyapunov exponent map: this highlights the accuracy of NAFF.
+We again see that the tube orbits in this potential are mostly regular or only mildly chaotic, however islands of stronger chaos do appear, especially at the intersections of unstable resonances.
+The unstable (white) regions in these maps reveal the Arnold web of resonances at this energy.
+These unstable regions are interconnected and it is therefore possible for a tube orbit that starts, for example, in the lower left-hand region of one of these maps to end up as a box orbit at some point in it its history (but the timescale for this to happen is likely prohibitively long).
 
-\begin{figure}[!h]
-\centering
-	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_xz.png}
-	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_yz.png}
-	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
-	\label{fig:lyapmap}
-\end{figure}
-
-[TODO: move lyapunov stuff to above FDR stuff?]
-We also compute the Lyapunov exponents for all orbits in the grid by integrating their orbits for 250 orbital periods.
-Figure~\ref{fig:lyapmap} shows the initial condition grids again, but now the greyscale intensity corresponds to the logarithm of the Lyapunov time, $t_\lambda = \lambda_{\rm max}^{-1}$.
-These Lyapunov exponent maps show that, for a fixed integration time, only the most chaotic orbits are identified.
-NAFF is able to detect small differences in the frequencies and thus measure small diffusion rates with comparable integration times.
+The strongest chaotic regions appear in both the Lyapunov exponent and frequency diffusion rate (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqmap}, right), but the frequency diffusion rate is more accurate and thus more sensitive to mild chaos.
+It is unclear whether these regions will be dynamically important on realistic timescales, but we have confirmed from two independent methods that there are regions of strongly chaotic orbits even in this simple, triaxial potential.
 
@@
 Our goal is to characterize the morphology of tidal debris along orbits generic to triaxial potentials.
-For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{fardal14, pearson15}, however, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
+For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{fardal14, pearson15}, however, it is unknown how the observable properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
 Further, quantifying the degree of stochasticity with, for example, the frequency diffusion rate represents a \emph{local} measure of chaos and it is unclear how this generalizes to the small but finite spreads in orbital characteristics common to debris that is tidally disrupted from globular clusters or dwarf satellite galaxies.
@@
 
-We measure the state of mixing of the debris as described above: we compute the KL divergence between the configuration-space density of the ensemble and the configuration-space density expected if the ensemble particles were uniformly distributed over the parent orbit energy hypersurface.
+We measure the state of mixing of debris along the same grid of initial conditions as described above by computing the KL divergence (KLD) between the configuration-space density of the ensemble and the configuration-space density expected if the ensemble particles were uniformly distributed over the parent orbit energy hypersurface (Section~\ref{sec:ensemble}).
+From the time evolution of the KLD for a given ensemble around an orbit from the orbit grid, we estimate the mixing rate [TODO].
+Figure~\ref{fig:kldmap} shows the same grids of initial conditions as plotted in Figures~\ref{fig:lyapmap} and \ref{fig:freqmap}, but now the intensity of the pixels is set by the [TODO].
 
-a:
-This measure can be computed at any time --- [we show two different maps of this quantity: one after 10 Gyr, one after 15 orbits?]
-
-b: 
-From the KL divergence, we compute a mixing time by measuring the slope or power law index?
-
-[Measure mixing time from KLD(t), plot freq. diffusion time vs. mixing time, show we can predict morphology from orbital properties of progenitor?]
+\begin{figure}[!h]
+\centering
+	\includegraphics[width=0.4\columnwidth] {figures/kld-map.png}
+	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+	\label{fig:kldmap}
+\end{figure}
 
@@
 \section{Conclusions}\label{sec:conclusions}
-For small ensembles, the frequency diffusion rate is a predictor of debris morphology.
+We have shown that ...
 

### 2015-03-31 ms.tex
figures at end, adding to potential section

@@
 \newcommand{\inverse}[1]{{#1}^{-1}}
-\newcommand{\msun}{\mathrm{M}_\odot}
+\newcommand{\msun}{\ensuremath{\mathrm{M}_\odot}}
 \newcommand{\bs}[1]{\boldsymbol{#1}}
@@
 
-%\begin{figure*}[!h]
-%\begin{center}
-%\includegraphics[width=\textwidth]{PATH}
-%\caption{ CAPTION }\label{fig:FIGNAME}
-%\end{center}
-%\end{figure*}
-
 \section{Introduction}\label{sec:introduction}
@@
 As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded due to the tidal forces of the host galaxy potential.
-If the orbit of such an object is regular, only mildly eccentric, and stays far from sharp density features (e.g., the disk), the disruption is a steady process and the mass is slowly stretched into long streams of matter: tidal (stellar) streams.
+If the orbit of such an object is regular, only mildly eccentric, and stays far from sharp density features (e.g., the disk), the disruption is a steady process and the mass is slowly stretched into long streams of matter: stellar tidal streams.
 The phase-space density and therefore the observable properties of these tidal streams are sensitive to the internal progenitor properties, orbit of the progenitor, and  mass distribution of the host galaxy.
 The morphological evolution of the debris depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also set by the shape and radial profile of the gravitational potential of the host galaxy.
-In the halo of the Milky Way, tidal streams of stars are found over a large range of distances ($\sim$10-100~kpc).
+In the halo of the Milky Way, tidal streams of stars are found over a large range of distances ($\approx$10-100~kpc).
 By simultaneously modeling the observed phase-space density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution of the Milky Way on these scales.
@@
 It has been shown qualitatively that along some strongly chaotic orbits, tidal streams form large, diffuse fans of debris \citep[e.g.,][]{fardal14}.
-However, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity; the result from \citet{pearson15} suggests that even a weakly chaotic progenitor orbit effects the density and therefore observability of tidal streams.
+However, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity; the result from \citet{pearson15} suggests that even a weakly chaotic progenitor orbit affects the density and therefore observability of tidal streams.
 
@@
 
-The density distributions within dark matter halos formed in cosmological N-body simulations are typically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}.
-\citet{jing02} found that a triaxial generalization of the classic NFW density profile \citep{navarro96} generates excellent fits to their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of the many halos formed in their simulations.
-They find median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that \citet{jing02} use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
-They find significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).
+The density distributions within dark matter halos formed in cosmological N-body simulations are generically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}.
+With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo \citep[e.g.,][]{dubinski??, recent NiHao stuff}.
+It is less clear what happens to the outer halo where tidal streams are more readily found; for this reason, we ask what we would expect to happen to tidal streams if the halo in which they form is strongly triaxial.
+
+ \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the classic NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these halos.
+JS02 finds median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{??, NiHao} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
+JS02 finds significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).
 
@@
 
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=0.75\textwidth]{figures/potential.pdf}
-\caption{Rotation curve along $x$, $y$, and $z$ axes for the triaxial, NFW potential } \label{fig:potential}
-\end{center}
-\end{figure*}
+% Figure 1
+%\begin{figure*}[!h]
+%\begin{center}
+%\includegraphics[width=0.75\textwidth]{figures/potential.pdf}
+%\caption{Rotation curve along $x$, $y$, and $z$ axes for the triaxial, NFW potential } \label{fig:potential}
+%\end{center}
+%\end{figure*}
 
-This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the effect of chaos in the halos of such galaxies.
+This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies.
 Realistic Galactic potentials will have a significant component due to the disk and bulge and may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}.
@@
 
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{figures/orbit-tree.pdf}
-\caption{A rough classification of the types of orbits found in triaxial potentials.
+% Figure 2
+%\begin{figure*}[!h]
+%\begin{center}
+%\includegraphics[width=0.6\textwidth]{figures/orbit-tree.pdf}
+%\caption{A rough classification of the types of orbits found in triaxial potentials.
 Regular orbits are classified by the number of resonance relations obeyed: conditionally periodic (\emph{none}), uni-resonant (\emph{one}), bi-resonant (\emph{two}).
@@
 } \label{fig:orbit-tree}
-\end{center}
-\end{figure*}
+%\end{center}
+%\end{figure*}
 
@@
 
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=0.5\textwidth]{figures/log-freqmap.png}
-\caption{Ratio of cartesian frequencies for a grid of iso-energy box orbits integrated in the potential of Equation~\ref{eq:logpotential}.
+% Figure 3
+%\begin{figure*}[!h]
+%\begin{center}
+%\includegraphics[width=0.5\textwidth]{figures/log-freqmap.png}
+%\caption{Ratio of cartesian frequencies for a grid of iso-energy box orbits integrated in the potential of Equation~\ref{eq:logpotential}.
 Stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
@@
 Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.} \label{fig:logfreqs}
-\end{center}
-\end{figure*}
+%\end{center}
+%\end{figure*}
 
@@
 
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=0.9\textwidth]{figures/two-orbits.pdf}
-\caption{ Two orbits integrated in the triaxial potential described in Section~\ref{sec:potential}.
+% Figure 4
+%\begin{figure*}[!h]
+%\begin{center}
+%\includegraphics[width=0.9\textwidth]{figures/two-orbits.pdf}
+%\caption{ Two orbits integrated in the triaxial potential described in Section~\ref{sec:potential}.
 Left panels (black) show two consecutive sections of orbit, where the integration time for each section (rows) is set to 100 times the maximum orbital period.
@@
 } \label{fig:two-orbits}
-\end{center}
-\end{figure*}
+%\end{center}
+%\end{figure*}
 
@@
 
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=0.62\textwidth]{figures/lyap-orbits.pdf}
-\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over 1000 orbital periods (the maximum orbital period) for Orbit A (left) and Orbit B (right).
+% Figure 5, 6
+%\begin{figure*}[!h]
+%\begin{center}
+%\includegraphics[width=0.62\textwidth]{figures/lyap-orbits.pdf}
+%\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over 1000 orbital periods (the maximum orbital period) for Orbit A (left) and Orbit B (right).
 For Orbit A, the FTMLE asymptotes towards zero with a characteristic power-law slope.
@@
 Over-plotted on both panels are lines (dashed, blue) that show $t^{-0.9}$ behavior.} \label{fig:lyap-orbits}
-\end{center}
-\end{figure*}
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=0.62\textwidth]{figures/freq-diff-orbits.pdf}
-\caption{ Difference in frequencies computed for successive sections of 100 orbital periods from the frequencies computed in the initial section for Orbit A (left) and Orbit B (right).
+%\end{center}
+%\end{figure*}
+%
+%\begin{figure*}[!h]
+%\begin{center}
+%\includegraphics[width=0.62\textwidth]{figures/freq-diff-orbits.pdf}
+%\caption{ Difference in frequencies computed for successive sections of 100 orbital periods from the frequencies computed in the initial section for Orbit A (left) and Orbit B (right).
 The three lines in each panel are for the three fundamental frequencies.
@@
 } \label{fig:fdiff-orbits} % The frequency diffusion rate defined in Equation~\ref{eq:fdrate} measures the initial slope of this trend.
-\end{center}
-\end{figure*}
+%\end{center}
+%\end{figure*}
 
@@
 
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=0.75\textwidth]{figures/D_KL.pdf}
-\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:DKL}
-\end{center}
-\end{figure*}
-
 \section{Results}\label{sec:results}
@@
 
-\begin{figure}[!h]
-\centering
-	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_xz.png}
-	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
-	\label{fig:lyapmap}
-\end{figure}
+%\begin{figure}[!h]
+%\centering
+%	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_xz.png}
+%	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+%	\label{fig:lyapmap}
+%\end{figure}
 
@@
 
-\begin{figure}[!h]
-\centering
-	\includegraphics[width=0.4\columnwidth] {figures/kld-map.png}
-	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
-	\label{fig:kldmap}
-\end{figure}
+%\begin{figure}[!h]
+%\centering
+%	\includegraphics[width=0.4\columnwidth] {figures/kld-map.png}
+%	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+%	\label{fig:kldmap}
+%\end{figure}
 
@@
 \bibliography{refs}
+\clearpage
+
+% Figure 1
+\begin{figure*}[!p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/potential.pdf}
+\caption{Rotation curve along $x$, $y$, and $z$ axes for the triaxial, NFW potential } \label{fig:potential}
+\end{center}
+\end{figure*}
+
+% Figure 2
+\begin{figure*}[!p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/orbit-tree.pdf}
+\caption{A rough classification of the types of orbits found in triaxial potentials.
+Regular orbits are classified by the number of resonance relations obeyed: conditionally periodic (\emph{none}), uni-resonant (\emph{one}), bi-resonant (\emph{two}).
+Irregular or chaotic orbits are simply separated into two classes: strongly and weakly chaotic, defined by the timescales of interest in a problem.
+When the characteristic chaotic timescale is much longer than, for example, the age of the universe, the orbit is only weakly chaotic.
+} \label{fig:orbit-tree}
+\end{center}
+\end{figure*}
+
+% Figure 3
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/log-freqmap.png}
+\caption{Ratio of cartesian frequencies for a grid of iso-energy box orbits integrated in the potential of Equation~\ref{eq:logpotential}.
+Stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
+The regularity of the points in this map reflects the input grid of initial conditions.
+Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.} \label{fig:logfreqs}
+\end{center}
+\end{figure*}
+
+% Figure 4
+\begin{figure*}[!p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/two-orbits.pdf}
+\caption{ Two orbits integrated in the triaxial potential described in Section~\ref{sec:potential}.
+Left panels (black) show two consecutive sections of orbit, where the integration time for each section (rows) is set to 100 times the maximum orbital period.
+Right panels (red) shows the same for the chaotic orbit.
+The appearance of the orbit in configuration space changes appreciably over timescales of tens of orbital periods.
+} \label{fig:two-orbits}
+\end{center}
+\end{figure*}
+
+% Figure 5
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/lyap-orbits.pdf}
+\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over 1000 orbital periods (the maximum orbital period) for Orbit A (left) and Orbit B (right).
+For Orbit A, the FTMLE asymptotes towards zero with a characteristic power-law slope.
+For Orbit B, the FTMLE begins to saturate to a constant; the Lyapunov exponent estimated for this orbit is $\lambda_{\rm max} \approx 0.07 \, {\rm Gyr}^{-1}$ corresponding to a Lyapunov time of $t_\lambda \approx 14 \, {Gyr}$.
+Over-plotted on both panels are lines (dashed, blue) that show $t^{-0.9}$ behavior.} \label{fig:lyap-orbits}
+\end{center}
+\end{figure*}
+
+% Figure 6
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/freq-diff-orbits.pdf}
+\caption{ Difference in frequencies computed for successive sections of 100 orbital periods from the frequencies computed in the initial section for Orbit A (left) and Orbit B (right).
+The three lines in each panel are for the three fundamental frequencies.
+Dashed (red) horizontal line shows the estimate of the uncertainty for the frequencies computed for Orbit A.
+After many hundreds of orbital periods, the frequency deviation is not appreciable indicating again that the orbit is regular or only very weakly chaotic.
+The frequency deviations computed for Orbit B, however, show a clear increasing trend, indicating that the frequencies are diffusing with time.
+} \label{fig:fdiff-orbits} % The frequency diffusion rate defined in Equation~\ref{eq:fdrate} measures the initial slope of this trend.
+\end{center}
+\end{figure*}
+
+% Figure 7
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/D_KL.pdf}
+\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:DKL}
+\end{center}
+\end{figure*}
+
+% Figure 8
+\begin{figure}[p]
+\centering
+	\includegraphics[width=\columnwidth] {figures/lyap_map_xz.png}
+	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+	\label{fig:lyapmap}
+\end{figure}
+
+% Figure 9
+%\begin{figure}[!h]
+%\centering
+%	\includegraphics[width=0.4\columnwidth] {figures/kld-map.png}
+%	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+%	\label{fig:kldmap}
+%\end{figure}
 

### 2015-03-31 ms.tex
minor tweaks to text

@@
 This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies.
-Realistic Galactic potentials will have a significant component due to the disk and bulge and may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}.
+Realistic Galactic potentials will have a significant component due to the disk and bulge and may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11, NiHao}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}.
 We expect all of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
@@
 \begin{equation}
-	\omega_k = \bs{n}_k \cdot \bs{\Omega}
+	\omega_k = \bs{n}_k \cdot \bs{\Omega},
 \end{equation}
-and the $a_k$ are complex amplitudes and $\bs{n}_k$ is an integer vector.
+the $a_k$ are complex amplitudes, and the $\bs{n}_k$ are integer vectors.
 \cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hanning-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
@@
 \end{equation}
-where $\bs{\Omega}_{T_1}$ and $\bs{\Omega}_{T_2}$ are the fundamental frequencies computed from two consecutive sections of orbit, $N_T$ is the number of orbital periods within a single section of orbit computed from the most significant fundamental frequency, and the maximum is taken with respect to the corresponding amplitudes, $\bs{a}$, of the frequency components.
+where $\bs{\Omega}_{T_1}$ and $\bs{\Omega}_{T_2}$ are the fundamental frequencies computed from two consecutive sections of orbit, $N_T$ is the number of orbital periods within a single section of orbit computed from the most significant fundamental frequency, and the maximum is taken with respect to the corresponding amplitudes, $\bs{a}$, of the fundamental frequency components.
 The frequency diffusion rate as defined above (e.g., measured in the frequency with the largest amplitude) is better correlated with the strength of chaos in an orbit \citep{valluri??} but differs from the original definition of \cite{valluri98}.
+%; we have found that for extremely chaotic orbits --- where peaks in the FFT overlap --- NAFF fails to recover any frequencies
 
-NAFF has been applied to many problems in Solar System and Galactic dynamics, especially in the study of of orbits in triaxial systems.
+NAFF has been applied to many problems in Solar System and Galactic dynamics, especially in the study of orbits in triaxial systems.
 For more details about the method, see \cite{papaphilippou96, laskar, etc.}.
-We have implemented and tested our own implementation of NAFF in Python.
-This code is publicly available through the Gary package.\footnote{link to repo} Figure~\ref{fig:logfreqs} shows a test of the implementation in which we reproduce the frequency map at a fixed energy of an axisymmetric, logarithmic potential \cite[][pg.
+We have implemented and tested our own implementation of NAFF in \texttt{Python}.
+This code is publicly available through the \texttt{Gary} package.\footnote{link to repo} Figure~\ref{fig:logfreqs} shows a test of the implementation in which we reproduce the frequency map at a fixed energy of an axisymmetric, logarithmic potential \cite[][pg.
 260, Figure~3.45]{binneytremaine}.
@@
 
-NAFF recovers the fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL96][]{papaphilippou96}.
+NAFF recovers the fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL96;][]{papaphilippou96}.
 PL96 show that a good choice of coordinates for tube orbits are the Poincar\'e symplectic polar coordinates, a set of canonical coordinates similar to cylindrical coordinates.
-When computing the frequencies for tube orbits, we first rotate the orbits until the circulation is about the $z$-axis, transform to Poincar\'e polar coordinates, then use NAFF to measure the fundamental frequencies.
+When computing the frequencies for tube orbits, we first align the circulation about the $z$-axis through rotation, transform to Poincar\'e polar coordinates, then use NAFF to measure the fundamental frequencies.
 We could equivalently use the Cartesian time series, but the convergence of terms is slower (the amplitudes of successive terms decrease slower for Cartesian coordinates).
 We have tested that our implementation of NAFF returns the same fundamental frequencies in either case for a set of tube orbits.
-For box orbits, the motion is close to separable in each Cartesian component and we therefore use Cartesian coordinates for estimating frequencies for these orbits.
+For box orbits, the motion is close to separable in each Cartesian component and we therefore use Cartesian coordinates for estimating the frequencies for these orbits.
 
@@
 In comparison, owing to the precision of NAFF, detecting small frequency diffusion rates can be done with only $\sim$100 orbital periods.
+[TODO: axe this figure]
 Figure~\ref{fig:fdiff-orbits} shows the three fundamental frequencies computed for 12 consecutive windows of orbit (each section is $\sim$100 orbital periods in length) for Orbit A (left) and Orbit B (right).

### 2015-04-01 ms.tex
phrasing (are we not doing phrasing anymore???)

@@
 In comparison, owing to the precision of NAFF, detecting small frequency diffusion rates can be done with only $\sim$100 orbital periods.
-[TODO: axe this figure]
-Figure~\ref{fig:fdiff-orbits} shows the three fundamental frequencies computed for 12 consecutive windows of orbit (each section is $\sim$100 orbital periods in length) for Orbit A (left) and Orbit B (right).
+Figure~\ref{fig:fdiff-orbits} shows the three fundamental frequencies computed for a sliding window of 12 sections of orbit (each section is $\sim$100 orbital periods in length) for Orbit A (left) and Orbit B (right).
 Dashed horizontal line is the estimate of the uncertainty in the numerical computation of the frequencies from Orbit A.
@@
 This raises an important consideration: do finite deviations diverge faster than predicted from local measures of chaos?
-Does the Lyapunov time or frequency rate predict the timescale over which a finite phase-space volume stays ``coherent''?
+Does the Lyapunov time or frequency rate predict the timescale over which a finite phase-space volume (e.g., tidal debris) stays ``coherent''?
 
@@
 We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
-For a given set of orbital initial conditions --- the ``parent'' orbit --- we find the position of the nearest pericenter, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all of these test particles for an integration time equal to 256 times the longest fundamental orbital period.
+For a given set of orbital initial conditions --- the ``parent'' orbit --- we find the position of the nearest pericenter, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for an integration time equal to XXX times the longest fundamental orbital period.
 The physical scale of the ensemble is set by the tidal radius in position and velocity scale in velocity and are therefore set by the mass scale of the progenitor \citep[e.g.,][]{johnston98, apw14}.
@@
 
-As the particle ensembles evolve in time, we follow a method similar to that of \citet[][hereafter MV96]{merritt96} to quantify the degree of mixing of the ensemble by computing the ``distance'' between the configuration space density of the ensemble at a given time, $\rho(\bs{x},t)$, with that of a distribution that is uniformly distributed over the energy hypersurface, $\rho_{E_0}(\bs{x})$, where $E_0$ is the energy of the parent orbit and thus hypersurface of interest.
+[TODO: wait, actually, they differ in a log -- why doesn't this change things more? I guess this is saying like, mean(log(dens)) ~ log(mean(dens))??? confuzzled]
+As the particle ensembles evolve in time, we follow the mean configuration-space density of each to quantify the degree of mixing of the ensembles.
+This is similar to the method of \citet{merritt96} who measured the evolution of the relative entropy between the configuration-space density of isoenergy ensembles, $\rho(\bs{x},t)$, and the density of an ensemble that is uniformly distributed over the energy hypersurface, $\rho_{E_0}(\bs{x})$.
+In our case, the
+
+ [but isoenergy so constant...yadda yadda].
+
+to quantify the degree of mixing of the ensemble by computing the ``distance'' between the configuration space density of the ensemble at a given time, $\rho(\bs{x},t)$, with that of a distribution that is uniformly distributed over the energy hypersurface, $\rho_{E_0}(\bs{x})$, where $E_0$ is the energy of the parent orbit and thus hypersurface of interest.
 Our method differs in a few ways from that of MV96.

### 2015-04-14 ms.tex
add authors

@@
 \usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
+\usepackage[yyyymmdd]{datetime}
+\renewcommand{\dateseparator}{-}
 
@@
 	    Kathryn V. Johnston\altaffilmark{\colum},
-	    Monica Valluri\altaffilmark{\mich}}
+	    Monica Valluri\altaffilmark{\mich},
+	    Sarah Pearson\altaffilmark{\colum}}
+%	    Andreas H. W. K\"upper\altaffilmark{\colum}}
 \date{\centering \today}
@@
 \acknowledgements
-The authors wish to thank Sarah Pearson, Robyn Sanderson, Andreas K\"upper, David Hogg, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
+The authors wish to thank  Andreas K\"upper, Robyn Sanderson, David Hogg, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 

### 2015-05-11 ms.tex
intro tweaks

@@
 	    Monica Valluri\altaffilmark{\mich},
-	    Sarah Pearson\altaffilmark{\colum}}
-%	    Andreas H. W. K\"upper\altaffilmark{\colum}}
+	    Sarah Pearson\altaffilmark{\colum},
+	    Andreas H. W. K\"upper\altaffilmark{\colum},
+	    David W. Hogg\altaffilmark{\nyu,\cds,\mpia}}
 \date{\centering \today}
@@
 \newcommand{\mich}{3}
+\newcommand{\nyu}{4}
+\newcommand{\cds}{5}
+\newcommand{\mpia}{6}
 \altaffiltext{\colum}{Department of Astronomy, 
@@
 			   Ann Arbor, MI 48109, USA}
+\altaffiltext{\nyu}{Center for Cosmology and Particle Physics,
+                      Department of Physics, New York University,
+                      4 Washington Place, New York, NY, 10003, USA}
+\altaffiltext{\cds}{Center for Data Science,
+                      New York University,
+                      4 Washington Place, New York, NY, 10003, USA}
+\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
+                     K\"onigstuhl 17, D-69117 Heidelberg, Germany}
 
@@
 
-%Tidal streams form from the steady disruption of stellar systems
-%orbiting within the gravitational field of some parent galaxy. Many
-%streams and debris structures have been discovered in the halo of the
-%Milky Way and have been used to model the potential of the Galaxy. 
-%However, few of these models have yet explored the properties of
-%tidal debris in triaxial potentials.
-%The existence of a variety of orbits, resonances, and chaotic regions in such potentials 
-%suggest that the morphologies and dispersal timescales of
-%debris could differ significantly from the simpler spherical and oblate cases.
-%In this work we use a series of N-body simulations of stellar systems over a range 
-%of masses of disruption in triaxial potentials to understand the influence of the nature 
-%and types of orbits on debris morphologies.
-%Our results suggest that the mere existence of the multitude of thin streams
-%already known to orbit the Milky Way provides significant constraints on the
-%classes of triaxial potentials that provide a good representation for its dark matter halo.
+% Context
+The large-scale, smooth components of galactic gravitational potentials are thought to be triaxial in shape.
+Triaxial potentials that match observational constraints and measurements of the radial profile and axis ratios of dark matter halos in cosmological simulations are almost certainly not globally integrable.
+An appreciable number of chaotic orbits are therefore expected in galactic halos.
+It is yet unclear whether chaotic regions of phase-space will affect the density evolution of stellar tidal streams, which form in the halos of galaxies from the disruption of stellar systems.
+% Aims
+Many numerical methods exist for assessing the degree of chaos for a given orbit (e.g., the Lyapunov exponent, the frequency diffusion rate). 
+These indicators, which are local to a single orbit, often predict a timescale over which chaos will be dynamically important.
+In this paper, we aim to compare these timescales to the density evolution of the finite-volume ensembles of orbits.
+% Methods: 
+We derp derp derp.
+% Results:
+We find that mean indicators such as the Lyapunov time or the (long term) frequency diffusion rate of a given ``parent'' orbit are not good predictors for the short-term density evolution of ensembles of orbits around this parent orbit.
+Instead, we find that the variance of the fundamental frequencies of the ensemble orbits over short timescales is a stronger indicator of the density evolution.
+% Conclusions:
 
@@
 
-On the largest scales in the halos of galaxies ($\gtrsim$30~kpc for a Milky-Way-like galaxy), gravitational potentials are thought to be triaxial.
-Despite suggestive evidence from a range of complimentary techniques, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively verified by observations.
+The dark matter halos of galaxies are thought to be triaxial in shape.
+Despite suggestive evidence from a range of complimentary observational methods, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively verified.
 Around other galaxies, it is generally hard to measure the 3D mass profile because they are seen in projection.
 From the Earth's position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii.
-Unfortunately, the Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely on non-random orbits and thus contain extra information (e.g., debris stripped from in-falling satellite galaxies).
+The Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely on non-random orbits and thus contain extra information (e.g., debris stripped from in-falling satellite galaxies).
 
 As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded due to the tidal forces of the host galaxy potential.
-If the orbit of such an object is regular, only mildly eccentric, and stays far from sharp density features (e.g., the disk), the disruption is a steady process and the mass is slowly stretched into long streams of matter: stellar tidal streams.
+If the orbit of such an object is regular, only mildly eccentric, and stays far from sharp density features (e.g., the disk), the disruption is a steady process and the mass is slowly stretched into long streams of matter known as stellar tidal streams.
 The phase-space density and therefore the observable properties of these tidal streams are sensitive to the internal progenitor properties, orbit of the progenitor, and mass distribution of the host galaxy.
+Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies (cite many).
+Around the Milky Way, streams are found over a large range of distances ($\approx$10-100~kpc).
 The morphological evolution of the debris depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also set by the shape and radial profile of the gravitational potential of the host galaxy.
-In the halo of the Milky Way, tidal streams of stars are found over a large range of distances ($\approx$10-100~kpc).
-By simultaneously modeling the observed phase-space density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution of the Milky Way on these scales.
+By simultaneously modeling the observed phase-space density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution around the Milky Way.
 
 Tidal streams are generally modeled as 1D structures formed on regular orbits with simple internal progenitor dynamics and density evolution.
-Methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to N-body simulations.
-All methods have been tested in some way on simulated observations of data and authors have mainly demonstrated their ability to recover parameters of analytic, static potential forms.
+Methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to N-body simulations \citep[see, e.g., the introduction of][]{kuepper15}.
+All methods have been tested in some way on simulated observations of data and these tests typically demonstrate the recovery of parameters for analytic, static potential forms.
 Many of the methods have been shown to work in axisymmetric or otherwise simple potentials but claim to extend to more complex, generalized potentials (e.g., triaxial or multi-component potentials).
@@
 They use the observed SDSS number density of stars and a limited number of radial velocities measured for stream members to fit model streams to the data.
-In the oblate potential (a three-component, bulge, disk, and spherical halo potential), a thin model stream is easily found that reproduces the observed stellar density morphology of the stream.
-In the triaxial potential (the potential from \cite{law10}, a three component, bulge, disk, and triaxial halo fit to Sagitarrius stream data), the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness of the stream given the observational constraints of the present-day position and velocity of the cluster.
-This result suggests that the shape of the host galaxy potential directly influences the morphology of the resulting debris.
+In the oblate potential (a three-component bulge+disk+spherical halo potential), a thin model stream is easily found that reproduces the observed stellar density morphology of the stream.
+In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness of the stream given the observational constraints of the present-day position and velocity of the cluster.
 % In order to extend tidal stream models to more generic potentials, it is important to understand why this stream-fanning occurs and whether this and other strange debris morphologies are generic to more complex potentials.
@@
 The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the oblate potential.
-It is well known that the number of symmetries or the number of degrees of freedom of a potential plays a major role in determining the orbital structure within the potential.
-The peculiar shape of the particular triaxial potential also likely has significant chaotic regions within the potential.
-\citet{pearson15} tested the stochasticity of the orbit of the fanned debris progenitor by computing the Lyapunov exponent but found that the orbit is consistent with being regular over dynamically relevant timescales (many Hubble times).
-It has been shown qualitatively that along some strongly chaotic orbits, tidal streams form large, diffuse fans of debris \citep[e.g.,][]{fardal14}.
-However, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity; the result from \citet{pearson15} suggests that even a weakly chaotic progenitor orbit affects the density and therefore observability of tidal streams.
+It is well known that the number of degrees of freedom of a potential plays a critical role in determining the orbit structure in the potential.
+Potentials (Hamiltonians) with more than two degrees of freedom generically contain significant chaotic regions.
+\citet{pearson15} tested the stochasticity of the orbit of the progenitor that produced fanned debris by computing the Lyapunov exponent but found that the orbit is consistent with being regular over dynamically relevant timescales (many Hubble times).
+It has been shown previously that along some strongly chaotic orbits, tidal streams form large, diffuse fans of debris \citep[e.g.,][]{fardal14}.
+However, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
+The result from \citet{pearson15} suggests that even a weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.
 
+TODO: this paragraph needs to be reworked...
 In this work, we analyze whether the final configuration-space morphology of simple tidal stream models is correlated with measures of chaos local to the orbit of the progenitor.
 The effect of chaos in the halos of galaxies has largely been neglected because the chaotic timescales estimated were thought to be many times the age of the universe, but recent work has shown that even weak chaos may be relevant for tidal stream morphology.
-Here we choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos as computed from single-orbit diagnostics, and compare these results to measures of the mixing time of finite-volume ensembles of orbits (analogs to tidal debris).
+In this Article, we choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos for grids of constant-energy orbits as computed from single-orbit diagnostics, and compare these results to measures of the mixing time of finite-volume ensembles of orbits (analogs to tidal debris).
 Finally, we connect the ensemble mixing rate to observational properties of streams, thus demonstrating that local orbital measures of chaos are predictors of tidal stream morphology.
@@
 With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo \citep[e.g.,][]{dubinski??, recent NiHao stuff}.
-It is less clear what happens to the outer halo where tidal streams are more readily found; for this reason, we ask what we would expect to happen to tidal streams if the halo in which they form is strongly triaxial.
+It is less clear what happens to the outer halo where tidal streams are more readily found.
+%; for this reason, we ask what we would expect to happen to tidal streams if the halo in which they form is strongly triaxial.
 
@@
 The functions $F_i(u)$ are given in the appendix of \cite{leesuto03}.
-We chose $r_s=40~{\rm kpc}$ and $v_c = 150~{\rm km}~{\rm s}^{-1}$ by taking the mean halo concentration for a ${\rm M}_{vir} \approx 10^{12}~\msun$ halo, $c_e\approx5$, from \cite{jing02} and by assuming $R_{vir}\approx200~{\rm kpc}$.
-Figure~\ref{fig:potential} shows the rotation curve of this potential along the three principal axes.
+We chose $r_s=20~{\rm kpc}$ and $v_c = 175~{\rm km}~{\rm s}^{-1}$ by taking the mean halo concentration for a ${\rm M}_{vir} \approx 10^{12}~\msun$ halo, $c_e\approx5$, from \cite{jing02} and by assuming $R_{vir}\approx200~{\rm kpc}$.
+Figure~\ref{fig:potential} shows equipotential contours of the potential in projection.
 
@@
 This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies.
-Realistic Galactic potentials will have a significant component due to the disk and bulge and may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11, NiHao}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}.
-We expect all of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
+Realistic Galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11, NiHao}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}.
+We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
 
@@
 
-We use the Dormand-Prince 8th-order Runge-Kutta scheme (TODO: cite Dormand 1981) to integrate orbits in the above potential.
-For all orbits, we ensure that energy is conserved to $\Delta E/E_0 \leq 10^{-8}$ by the end of integration.
-Unless otherwise specified, orbits are integrated for $\sim$250 orbital periods with 500 steps per orbital period.
-Some box orbits require more internal integration steps in order to conserve energy to the required level.
+We use the Dormand-Prince 8th-order Runge-Kutta scheme \citep{prince81} to integrate orbits in the above potential.
+For all orbits we ensure that energy is conserved to $\Delta E/E_0 \leq 10^{-8}$ by the end of integration, however most orbits conserve energy to $\Delta E/E_0 \approx 10^{-13}$.
+Unless otherwise specified, the integration timesteps are chosen so that there are 512 steps per strongest orbital period component.
 
@@
 For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (dof), the equations of motion are most simply expressed in angle-action coordinates \cite[e.g.,][]{goldstein80}.
-In these coordinates, the actions, $J_n$, are integrals of the motion, and the conjugate angle variables, $\theta_n$, increase linearly with time
+In these coordinates, the actions, $J_n$, are (isolating) integrals of the motion, and the conjugate angle variables, $\theta_n$, increase linearly with time
 \begin{align}
@@
 \subsection{Numerical estimation of fundamental frequencies}\label{sec:naff}
+
+[Motivate why we use frequencies vs. actions vs. Lyapunov exponent -- can evaluate small-time movement vs. average?]
 
@@
 
-\begin{figure}[h]
-\centering
-	\includegraphics[width=0.45\columnwidth] {figures/freqmap_xz.png}
-	\includegraphics[width=0.45\columnwidth] {figures/diffusion_map_xz.png}
-	\caption{(\emph{Left} Frequency map of ...
+%\begin{figure}[h]
+%\centering
+%	\includegraphics[width=0.45\columnwidth] {figures/freqmap_xz.png}
+%	\includegraphics[width=0.45\columnwidth] {figures/diffusion_map_xz.png}
+%	\caption{(\emph{Left} Frequency map of ...
 For short-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $z$ axis, $\Omega_\phi$, and the oscillation along the $z$ axis, $\Omega_z$.
@@
 Box orbits are not shown.}
-	\label{fig:freqmap}
-\end{figure}
+%	\label{fig:freqmap}
+%\end{figure}
 
@@
 % Figure 2
-\begin{figure*}[!p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/orbit-tree.pdf}
-\caption{A rough classification of the types of orbits found in triaxial potentials.
+%\begin{figure*}[!p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/orbit-tree.pdf}
+%\caption{A rough classification of the types of orbits found in triaxial potentials.
 Regular orbits are classified by the number of resonance relations obeyed: conditionally periodic (\emph{none}), uni-resonant (\emph{one}), bi-resonant (\emph{two}).
@@
 } \label{fig:orbit-tree}
-\end{center}
-\end{figure*}
+%\end{center}
+%\end{figure*}
 
 % Figure 3
-\begin{figure*}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/log-freqmap.png}
-\caption{Ratio of cartesian frequencies for a grid of iso-energy box orbits integrated in the potential of Equation~\ref{eq:logpotential}.
+%\begin{figure*}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/log-freqmap.png}
+%\caption{Ratio of cartesian frequencies for a grid of iso-energy box orbits integrated in the potential of Equation~\ref{eq:logpotential}.
 Stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
@@
 Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.} \label{fig:logfreqs}
-\end{center}
-\end{figure*}
-
-% Figure 4
-\begin{figure*}[!p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/two-orbits.pdf}
-\caption{ Two orbits integrated in the triaxial potential described in Section~\ref{sec:potential}.
+%\end{center}
+%\end{figure*}
+%
+%% Figure 4
+%\begin{figure*}[!p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/two-orbits.pdf}
+%\caption{ Two orbits integrated in the triaxial potential described in Section~\ref{sec:potential}.
 Left panels (black) show two consecutive sections of orbit, where the integration time for each section (rows) is set to 100 times the maximum orbital period.
@@
 } \label{fig:two-orbits}
-\end{center}
-\end{figure*}
-
-% Figure 5
-\begin{figure*}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/lyap-orbits.pdf}
-\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over 1000 orbital periods (the maximum orbital period) for Orbit A (left) and Orbit B (right).
+%\end{center}
+%\end{figure*}
+%
+%% Figure 5
+%\begin{figure*}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/lyap-orbits.pdf}
+%\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over 1000 orbital periods (the maximum orbital period) for Orbit A (left) and Orbit B (right).
 For Orbit A, the FTMLE asymptotes towards zero with a characteristic power-law slope.
@@
 Over-plotted on both panels are lines (dashed, blue) that show $t^{-0.9}$ behavior.} \label{fig:lyap-orbits}
-\end{center}
-\end{figure*}
-
-% Figure 6
-\begin{figure*}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/freq-diff-orbits.pdf}
-\caption{ Difference in frequencies computed for successive sections of 100 orbital periods from the frequencies computed in the initial section for Orbit A (left) and Orbit B (right).
+%\end{center}
+%\end{figure*}
+%
+%% Figure 6
+%\begin{figure*}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/freq-diff-orbits.pdf}
+%\caption{ Difference in frequencies computed for successive sections of 100 orbital periods from the frequencies computed in the initial section for Orbit A (left) and Orbit B (right).
 The three lines in each panel are for the three fundamental frequencies.
@@
 } \label{fig:fdiff-orbits} % The frequency diffusion rate defined in Equation~\ref{eq:fdrate} measures the initial slope of this trend.
-\end{center}
-\end{figure*}
-
-% Figure 7
-\begin{figure*}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/D_KL.pdf}
-\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:DKL}
-\end{center}
-\end{figure*}
-
-% Figure 8
-\begin{figure}[p]
-\centering
-	\includegraphics[width=\columnwidth] {figures/lyap_map_xz.png}
-	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
-	\label{fig:lyapmap}
-\end{figure}
-
-% Figure 9
-%\begin{figure}[!h]
+%\end{center}
+%\end{figure*}
+%
+%% Figure 7
+%\begin{figure*}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/D_KL.pdf}
+%\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:DKL}
+%\end{center}
+%\end{figure*}
+%
+%% Figure 8
+%\begin{figure}[p]
 %\centering
-%	\includegraphics[width=0.4\columnwidth] {figures/kld-map.png}
+%	\includegraphics[width=\columnwidth] {figures/lyap_map_xz.png}
 %	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
-%	\label{fig:kldmap}
+%	\label{fig:lyapmap}
 %\end{figure}
+%
+%% Figure 9
+%%\begin{figure}[!h]
+%%\centering
+%%	\includegraphics[width=0.4\columnwidth] {figures/kld-map.png}
+%%	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+%%	\label{fig:kldmap}
+%%\end{figure}
 

### 2015-05-12 ms.tex
todos and new nonlinear dynamics section

@@
 \begin{document}
-
-%TODO:
-%The classical way to detect chaotic motion is
-%the computation of the Lyapunov exponents. We
-%claim that for weakly chaotic solutions, the frequency
-%analysis gives information more rapidly
-%and more precisely. The reason is that the frequency
-%analysis is an instantaneous observation
-%of the fundamental frequencies which are essential
-%parameters of the dynamical system, while
-%the usual computation of Lyapunov exponents
-%acts like a mean and needs a very long time span
-%to provide reliable results (for references and
-%discussion on Lyapunov exponents, see refs. [3,
-%12].
-
-% Highlight that stochastic tubes have not been studied much...
 
@@
 
-% Figure 1
-%\begin{figure*}[!h]
-%\begin{center}
-%\includegraphics[width=0.75\textwidth]{figures/potential.pdf}
-%\caption{Rotation curve along $x$, $y$, and $z$ axes for the triaxial, NFW potential } \label{fig:potential}
-%\end{center}
-%\end{figure*}
-
 This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies.
@@
 
-For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (dof), the equations of motion are most simply expressed in angle-action coordinates \cite[e.g.,][]{goldstein80}.
+% TODO: move all except chaotic diffusion stuff to appendix?
+
+An orbit in an $N$ degree of freedom (dof) Hamiltonian, $H$, is a set of $2N$ quasi-periodic time series, 
+\begin{equation}
+(w_1(t),...,w_{2N}(t)) = (q_1(t),...,q_{N}(t),p_1(t),...,p_{N}(t)) \label{eq:coords}
+\end{equation}
+where $q_i$ and $p_i$ are conjugate coordinates in the sense that
+\begin{align}
+	\dot{p}_i &= -\frac{\partial H}{\partial q_i}\\
+	\dot{q}_i &= \frac{\partial H}{\partial p_i}
+\end{align}
+for all $t$. The motion in any component, $w_i(t)$, can be represented as a Fourier sum,
+\begin{equation}
+	w_i(t) = \sum_k^\infty a_k \, e^{i\,\omega_k\,t}; \, (a_k \in \mathbb{C}).
+\end{equation}
+
+A \emph{regular orbit} is a set of such time series that can be transformed to a special set of canonical coordinates known as angle-action coordinates.
+In these coordinates, the position variables are angles, $\boldsymbol{\theta}$, that increase linearly with time with rates set by $N$ constant, fundamental frequencies, $\boldsymbol{\Omega} = (\Omega_1, ..., \Omega_N)$.
+The frequencies in the Fourier sum of any individual component of motion (the $\omega_k$ in Equation~1) are just linear, integer combinations of the fundamental frequencies, $\boldsymbol{\Omega}$ --- that is, for a regular orbit, any Fourier component frequency may be written
+\begin{align}
+	\omega_k &= \boldsymbol{n_k} \cdot \boldsymbol{\Omega}; \, (\boldsymbol{n}_k \in \mathbb{Z}^3).
+\end{align}
+The conjugate momentum coordinates --- the actions,  $\boldsymbol{J}$ --- are constants of motion. Even stronger, the actions are isolating integrals and any pair are in involution such that
+\begin{equation}
+	[J_i, J_j] = 0
+\end{equation}
+where $[\cdot,\cdot]$ is the Poisson bracket.
+This implies that for an $N$ dof system, a regular orbit has $N$ independent constants of motion and the motion is therefore restricted to an $N$-dimensional manifold embedded in the 2$N$ dimensional phase space.
+The conjugate angle variables define an orthogonal basis, and the topology of angle-action space is therefore the product of $N$ circles, an $N$-torus.
+Any regular orbit in an $N$ dof Hamiltonian can be understood as motion on the surface of an $N$-torus.
+Each set of actions, $(J_1,...,J_N)$, uniquely labels a torus, and regular orbits may be referred to in terms of their orbital tori.
+If a transformation to angle-action variables exists, the Hamiltonian may be written solely in terms of the actions,
+\begin{equation}
+	H = H(\boldsymbol{J})
+\end{equation}
+
+A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally --- that is, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position.
+The condition for global integrability is very restrictive and it seems that the likelihood that a Hamiltonian is globally integrable decreases as the number of degrees of freedom increase \citep[e.g.,][]{lichtenberg83}.
+
+Galactic potentials are almost certainly not globally integrable, but it is useful to understand the orbit structure in integrable systems before extending to more general potentials.
+Only one globally integrable triaxial mass model is known that generates motion that is qualitatively similar to that seen in Galaxies: the St\"ackel or ``Perfect Ellipsoid'' model \citep[e.g.,][]{kuzmin56, kuzmin73, deZeeuw85, cincotta06}.
+The four general classes of orbits in this potential are box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits.
+Unlike the box orbits, the tube orbits preserve a sense of rotation about either the long or short axis.
+
+Over long times, a generic orbit in this potential fills the surface of its 3-torus --- that is, the three fundamental frequencies of the orbit are incommensurable.
+If there exists a relation of the form
+\begin{equation}
+	\boldsymbol{n} \cdot \boldsymbol{\Omega} = n\, \Omega_1 + m\, \Omega_2 + l\, \Omega_3 = 0
+\end{equation}
+then the orbit is a resonant orbit and the effective dimensionality of the orbit is decreased to 2D.
+We refer to orbits that obey a single resonance relation as \emph{uni-resonant} orbits.
+If an additional resonance relation exists, the orbit is closed and becomes a 1D curve; we refer to such orbits as \emph{bi-resonant} orbits.
+The resonant structure of a potential --- the relative importance of particular resonance numbers --- is difficult to compute, but determines the global behavior of orbits in the potential.
+\cite{cincotta06} analytically compute the resonant structure in action-space for a St\"ackel potential.
+In frequency-space, the resonances are planes, but in action-space are generic 2D surfaces.
+
+The orbit structure of non-integrable potentials can be qualitatively understood by considering a Hamiltonian that is a small perturbation away from being globally integrable --- that is, a Hamiltonian that may be written
+\begin{equation}
+	H(\boldsymbol{J}, \boldsymbol{\theta}) = H_0(\boldsymbol{J}) + \epsilon \, H_1(\boldsymbol{J}, \boldsymbol{\theta})
+\end{equation}
+where $\epsilon$ is a small parameter that determines the perturbation strength \citep[a complete description of perturbation theory applied to nonlinear Hamiltonians is given in][]{lichtenberg83}.
+When $0 < |\epsilon| \ll 1$, resonant surfaces become ``thick'' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99} and the stable regions are surrounded by chaotic layers where stochastic motion occurs.
+Stochastic or chaotic orbits are characterized by apparently random behavior despite being generated by entirely deterministic systems.
+The stochastic layers form gaps in action space where tori are destroyed.
+The orbits that exist in these regions therefore can not be represented in angle-action coordinates, as their actions and frequencies evolve with time.
+
+Stochastic motion or the diffusion of an orbit in actions and frequencies may occur either \emph{across} the resonance layers or \emph{along} the stochastic layers (Arnold diffusion).
+For weakly chaotic orbits, motion across a resonance layer typically occurs with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
+Thus, if the resonance libration frequency is comparable to the orbital frequency, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
+However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can't change by large factors (unless there is resonance overlap and the motion is strongly chaotic).
+The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict, however, this has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}.
+For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; the implication of this connectivity is that an orbit may dramatically change its appearance as it navigates through the Arnold web, though the timescales for this to occur are typically many 1000s of orbital periods.
+
+When $\epsilon$ is large, there is no general prediction for the resulting behavior.
+It seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep{valluri, merritt, dehnen, etc.}.
+
+%For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (dof), the equations of motion are most simply expressed in angle-action coordinates \cite[e.g.,][]{goldstein80}.
 In these coordinates, the actions, $J_n$, are (isolating) integrals of the motion, and the conjugate angle variables, $\theta_n$, increase linearly with time
-\begin{align}
-	\dot{\act}_n &= -\frac{\partial H}{\partial \theta_n} = 0\\
-	\act_n &= {\rm const.}\\
-	\dot{\theta}_n &= \frac{\partial H}{\partial \act_n} = \Omega_n(\act_1,\act_2,...,\act_n)\label{eq:aafreq}\\
-	\theta_n(t) &= \Omega_n t + \theta_n(0)
-\end{align}
-where $i=1,2,...N$, and $\Omega_n$ are the fundamental frequencies of the orbit.
+%\begin{align}
+%	\dot{\act}_n &= -\frac{\partial H}{\partial \theta_n} = 0\\
+%	\act_n &= {\rm const.}\\
+%	\dot{\theta}_n &= \frac{\partial H}{\partial \act_n} = \Omega_n(\act_1,\act_2,...,\act_n)\label{eq:aafreq}\\
+%	\theta_n(t) &= \Omega_n t + \theta_n(0)
+%\end{align}
+%where $i=1,2,...N$, and $\Omega_n$ are the fundamental frequencies of the orbit.
 The topology of angle-action space is equivalent to that of the surface of an $N$-dimensional torus, and thus orbits are often referred to in terms of \emph{orbital tori}: each set of actions, $(\act_1,\act_2,...\act_N)$, uniquely ``label'' the torus associated with a given orbit and the angles specify the location on the surface of the $N$-torus.
 If a global transformation to angle-action coordinates exists, the potential is said to be \emph{integrable}.
-
-Galactic gravitational potentials will almost certainly not be globally integrable.
+%
+%Galactic gravitational potentials will almost certainly not be globally integrable.
 Dark matter halos of galaxies are expected to be triaxial but even if they are close to axisymmetric or spherical, the superposition of components (e.g., disk, bulge, halo) will break global integrability.
 If a triaxial (three dof) potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
-\begin{equation}
-	H(\bs{\theta}, \bs{\act}) = H_0(\bs{\act}) + \epsilon H_1(\bs{\theta}, \bs{\act})
-\end{equation}
-then a large number of regular orbits will survive the perturbation (by way of the KAM theorem)\footnote{3-vectors are represented with bold symbols, e.g., $\bs{J}=(J_1,J_2,J_3)$.}.
+%\begin{equation}
+%	H(\bs{\theta}, \bs{\act}) = H_0(\bs{\act}) + \epsilon H_1(\bs{\theta}, \bs{\act})
+%\end{equation}
+%then a large number of regular orbits will survive the perturbation (by way of the KAM theorem)\footnote{3-vectors are represented with bold symbols, e.g., $\bs{J}=(J_1,J_2,J_3)$.}.
 The tori that survive will in general be separated by regions of irregular or chaotic motion, and thus any transformations to angle-action coordinates can only be defined locally.
@@
 Much effort has gone in to developing general methods for transforming either from position-velocity to angle-action coordinates, $(\bs{x},\bs{v})\rightarrow(\bs{\theta},\bs{J})$ \citep{many}, or vice-versa, $(\bs{\theta},\bs{J})\rightarrow(\bs{x},\bs{v})$ \citep{many}.
-
-The types of motion generic to triaxial systems can be split into several sub-categories (summarized in Figure~\ref{fig:orbit-tree}).
+%
+%The types of motion generic to triaxial systems can be split into several sub-categories (summarized in Figure~\ref{fig:orbit-tree}).
 Regular motion can be classified by the number of resonance relations obeyed \citep[e.g.,][]{lichtenberg83}: \emph{conditionally periodic} orbits obey no resonance relations, \emph{uni-resonant} orbits obey a single relation of the form $\bs{m}\cdot\bs{\Omega}=0$, and \emph{bi-resonant} orbits obey two resonance relations, $\bs{m}\cdot\bs{\Omega}=0$ and $\bs{n}\cdot\bs{\Omega}=0$, where $\bs{m}$ and $\bs{n}$ are integer vectors.
-
-% Figure 2
-%\begin{figure*}[!h]
-%\begin{center}
-%\includegraphics[width=0.6\textwidth]{figures/orbit-tree.pdf}
-%\caption{A rough classification of the types of orbits found in triaxial potentials.
+%
+%% Figure 2
+%%\begin{figure*}[!h]
+%%\begin{center}
+%%\includegraphics[width=0.6\textwidth]{figures/orbit-tree.pdf}
+%%\caption{A rough classification of the types of orbits found in triaxial potentials.
 Regular orbits are classified by the number of resonance relations obeyed: conditionally periodic (\emph{none}), uni-resonant (\emph{one}), bi-resonant (\emph{two}).
@@
 } \label{fig:orbit-tree}
-%\end{center}
-%\end{figure*}
-
-Truly resonant orbits are rare \citep{merritt99}, but dynamically important in systems with several degrees of freedom as they determine the structure of orbit-space.
+%%\end{center}
+%%\end{figure*}
+%
+%Truly resonant orbits are rare \citep{merritt99}, but dynamically important in systems with several degrees of freedom as they determine the structure of orbit-space.
 Stable, resonant orbits generate families of orbits that may behave or appear similarly, whereas unstable resonances are typically associated with regions of chaos.
@@
 In three dimensions, the extra degree of freedom allows chaotic orbits to ``escape'' this confinement.
-
-For small potential perturbations, many regular orbits survive and only small chaotic islands are introduced.
+%
+%For small potential perturbations, many regular orbits survive and only small chaotic islands are introduced.
 As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed.
@@
 This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{sec:lyap}) or frequency diffusion time (Section~\ref{sec:naff}).
-
-These local measures quantify the divergence of infinitesimally nearby orbits or the rate of frequency diffusion along an orbit.
+%
+%These local measures quantify the divergence of infinitesimally nearby orbits or the rate of frequency diffusion along an orbit.
 For an ensemble of orbits with finite spreads in energy or frequencies (e.g., tidal debris), each orbit will have a slightly different characteristic timescale, but we still expect mixing time of the ensemble to correlate with local measures of chaos of the ``parent'' orbit of the ensemble (Section~\ref{sec:ensemble}).
 The evolution of tidal debris will depend on the ensemble mixing rate around the parent orbit: chaotic mixing will cause the density of the debris to fall off more quickly than the typical power-law behavior expected for regular orbits \citep[e.g.,][]{merritt96, helmi99}.
-
+%
 \subsection{Lyapunov exponents} \label{sec:lyap}
 
-%The simplest and perhaps most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit.
-Defining the vector $\bs{w} = (q_1,...,q_N,p_1,...,p_N)$ for some set of canonical coordinates $(q_i,p_i)$, we can write Hamilton's equations as
-%\begin{equation}
-%	\dot{\bs{w}} = \bs{\mathcal{J}} \frac{\partial H}{\partial \bs{w}} = f(\bs{w}) \label{eq:ham}
-%\end{equation}
-%where $\bs{\mathcal{J}}$ is the $2N \times 2N$ canonical Poisson tensor (also called the symplectic matrix) defined by
-%\begin{equation}
-%	\bs{\mathcal{J}} = \left( \begin{array}{c:c} 0 & \ident \\ \hdashline -\ident & 0 \end{array} \right)
-%\end{equation}
-%with $N$-dimensional identity matrices $\ident$.
-If we consider a nearby phase-space position, $\bs{w}'$, separated from $\bs{w}$ by an infinitesimal deviation, $\delta \bs{w}$, such that $\bs{w}' = \bs{w} + \delta \bs{w}$.
-We can expand to linear order in the deviation about the parent orbit and write the equations of motion for the deviation as follows
-%\begin{align}
-%	\dot{\bs{w}}' &= f(\bs{w}')\\
-%	\dot{\bs{w}} + \dot{\delta \bs{w}} &= f(\bs{w} +  \delta \bs{w})\\
-%	\dot{\bs{w}} + \dot{\delta \bs{w}} &\approx f(\bs{w}) + \jac_{\bs{w}}\cdot\delta \bs{w} + \mathcal{O}(\delta \bs{w}^2)\\
-%	\dot{\delta \bs{w}} &\approx \jac_{\bs{w}} \cdot  \delta \bs{w}\label{eq:deviate}
-%\end{align}
-%where $\jac$ is the Jacobian of Eq.~\ref{eq:ham}, evaluated at the parent orbit (the Hessian of the Hamiltonian).
-For chaotic orbits, the maximum eigenvalue of the solution matrix to Eq.~\ref{eq:deviate} is positive real, leading to exponential divergence of nearby orbits.
-
-%The simplest and perhaps most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit.
-Defining the vector $\bs{w} = (q_1,...,q_N,p_1,...,p_N)$ for some set of canonical coordinates $(q_i,p_i)$, and a nearby phase-space position, $\bs{w}'$, separated from $\bs{w}$ by an infinitesimal deviation, $\delta \bs{w}$, such that $\bs{w}' = \bs{w} + \delta \bs{w}$, we can expand to linear order in the deviation about the parent orbit and write the equations of motion for the deviation as
-%\begin{align}
-%	\dot{\delta \bs{w}} &\approx \jac_{\bs{w}} \cdot  \delta \bs{w}\label{eq:deviate}
-%\end{align}
-%where $\jac$ is the Jacobian of Eq.~\ref{eq:ham}, evaluated at the parent orbit (the Hessian of the Hamiltonian).
-For chaotic orbits, the maximum eigenvalue of the solution matrix to Eq.~\ref{eq:deviate} is positive real, leading to exponential divergence of nearby orbits.
+% TODO: move most of this to appendix? leave only discussion?
 
 The simplest and perhaps most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit.
-For some set of canonical coordinates $(q,p)$, we define the vector $w_i = (w_1,...w_{2N}) = (q_1,...,q_N,p_1,...,p_N)$ so we can write Hamilton's equations as
+Using our definition of $w$ from Equation~\ref{eq:coords}, we can write Hamilton's equations as
 \begin{equation}
@@
 
-[Motivate why we use frequencies vs. actions vs. Lyapunov exponent -- can evaluate small-time movement vs. average?]
+% TODO: move most to appendix, leave only discussion?
 
@@
 
-% Figure 3
-%\begin{figure*}[!h]
-%\begin{center}
-%\includegraphics[width=0.5\textwidth]{figures/log-freqmap.png}
-%\caption{Ratio of cartesian frequencies for a grid of iso-energy box orbits integrated in the potential of Equation~\ref{eq:logpotential}.
-Stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
-The regularity of the points in this map reflects the input grid of initial conditions.
-Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.} \label{fig:logfreqs}
-%\end{center}
-%\end{figure*}
-
 NAFF recovers the fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL96;][]{papaphilippou96}.
@@
 
+[Can see lots of cool structure, but rates are tiny! Why does this manifest itself over 10-20 periods?]
+
 \subsection{Comparison of methods}\label{sec:comparison}
-
-% Figure 4
-%\begin{figure*}[!h]
-%\begin{center}
-%\includegraphics[width=0.9\textwidth]{figures/two-orbits.pdf}
-%\caption{ Two orbits integrated in the triaxial potential described in Section~\ref{sec:potential}.
-Left panels (black) show two consecutive sections of orbit, where the integration time for each section (rows) is set to 100 times the maximum orbital period.
-Right panels (red) shows the same for the chaotic orbit.
-The appearance of the orbit in configuration space changes appreciably over timescales of tens of orbital periods.
-} \label{fig:two-orbits}
-%\end{center}
-%\end{figure*}
 
@@
 
-% Figure 5, 6
-%\begin{figure*}[!h]
-%\begin{center}
-%\includegraphics[width=0.62\textwidth]{figures/lyap-orbits.pdf}
-%\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over 1000 orbital periods (the maximum orbital period) for Orbit A (left) and Orbit B (right).
-For Orbit A, the FTMLE asymptotes towards zero with a characteristic power-law slope.
-For Orbit B, the FTMLE begins to saturate to a constant; the Lyapunov exponent estimated for this orbit is $\lambda_{\rm max} \approx 0.07 \, {\rm Gyr}^{-1}$ corresponding to a Lyapunov time of $t_\lambda \approx 14 \, {Gyr}$.
-Over-plotted on both panels are lines (dashed, blue) that show $t^{-0.9}$ behavior.} \label{fig:lyap-orbits}
-%\end{center}
-%\end{figure*}
-%
-%\begin{figure*}[!h]
-%\begin{center}
-%\includegraphics[width=0.62\textwidth]{figures/freq-diff-orbits.pdf}
-%\caption{ Difference in frequencies computed for successive sections of 100 orbital periods from the frequencies computed in the initial section for Orbit A (left) and Orbit B (right).
-The three lines in each panel are for the three fundamental frequencies.
-Dashed (red) horizontal line shows the estimate of the uncertainty for the frequencies computed for Orbit A.
-After many hundreds of orbital periods, the frequency deviation is not appreciable indicating again that the orbit is regular or only very weakly chaotic.
-The frequency deviations computed for Orbit B, however, show a clear increasing trend, indicating that the frequencies are diffusing with time.
-} \label{fig:fdiff-orbits} % The frequency diffusion rate defined in Equation~\ref{eq:fdrate} measures the initial slope of this trend.
-%\end{center}
-%\end{figure*}
-
 Figure~\ref{fig:lyap-orbits} shows $l_N$ computed for the two orbits described above.
@@
 
-\section{Test particle ensembles} \label{sec:ensemble}
+\section{Results}\label{sec:results}
+
+\subsection{Orbital properties: frequencies, frequency diffusion, and Lyapunov exponents}
+
+We start by studying the properties of individual orbits in the potential described above.
+We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy $E=XXX$.
+We fix $v_x = v_z = 0$, and compute $v_y$ from the energy.
+The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic.
+Tube orbits are generally less stochastic than box orbits --- the other major class of orbits in triaxial potentials --- which tend to plunge deep into the inner regions of the potential, however there are appreciable regions of stochastic tube orbits.
+Thin tidal streams are thought to form along tube orbits rather than box orbits because of the fast disruption expected for stellar systems on radially plunging orbits.
+The major classes of stable tube orbits circulate about either the major or minor axis.
+The most prominent orbits in this grid are the short-axis and outer long-axis tubes, but this grid generates all of the major orbit classes: short-axis tubes, long-axis inner tubes, long-axis outer tubes, stochastic intermediate-axis, and box orbits.
+
+We compute the Lyapunov exponents for all orbits in the grid by integrating the orbits for 1000 orbital periods and use the method described in Section~\ref{sec:lyap} to estimate the maximum Lyapunov exponent.
+Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $x--z$ plane) --- each pixel corresponds to an orbit, and the greyscale intensity is related to the logarithm of the Lyapunov time, $t_\lambda = \lambda_{\rm max}^{-1}$.
+This map shows that by numerically estimating the MLE over this fixed integration window, the majority of orbits appear to be regular or have long Lyapunov times.
+
+%\begin{figure}[!h]
+%\centering
+%	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_xz.png}
+%	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
+%	\label{fig:lyapmap}
+%\end{figure}
+
+For all orbits, we also integrate instead for $\approx$200 orbital periods (Section~\ref{sec:integration}) and use NAFF to compute the fundamental frequencies for the two consecutive sections of 100 orbital periods.
+Figure~\ref{fig:freqmap} (left) shows a map of the fundamental frequency ratios for all tube orbits from the input grid (we plot the frequencies estimated from the first 100 orbital periods).
+This panel demonstrates again that many of the orbits are regular: the transformed, even-spaced sampling of the initial conditions is preserved in the frequency transformation, implying that the frequencies for many orbits do not change appreciably.
+Also apparent is resonant trapping, which manifests as over-dense, linear features in the frequency map.
+For approximately 2\% of the orbits, NAFF fails to recover the frequencies (similar difficulties were reported in \cite{valluri98} and \cite{papaphilippou96} for recovering certain tube orbit frequencies).
+
+%\begin{figure}[h]
+%\centering
+%	\includegraphics[width=0.45\columnwidth] {figures/freqmap_xz.png}
+%	\includegraphics[width=0.45\columnwidth] {figures/diffusion_map_xz.png}
+%	\caption{(\emph{Left} Frequency map of ...
+For short-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $z$ axis, $\Omega_\phi$, and the oscillation along the $z$ axis, $\Omega_z$.
+For long-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $x$ axis, $\Omega_\phi$, and the oscillation along the $x$ axis, $\Omega_x$.
+Box orbits are not shown.}
+%	\label{fig:freqmap}
+%\end{figure}
+
+For each orbit, we also compute the frequency diffusion rate per orbit, $\fdrate$ (Equation~\ref{eq:fdrate}), using the frequencies computed for the two integrated sections of orbit.
+Figure~\ref{fig:freqmap} (right) shows the grid of initial conditions for all orbits colored by $\log_{10}(\fdrate)$ (white points are strongly chaotic, black are regular).
+A rich structure of resonances appears in this frequency diffusion map that were not seen in the Lyapunov exponent map: this highlights the accuracy of NAFF.
+We again see that the tube orbits in this potential are mostly regular or only mildly chaotic, however islands of stronger chaos do appear, especially at the intersections of unstable resonances.
+The unstable (white) regions in these maps reveal the Arnold web of resonances at this energy.
+These unstable regions are interconnected and it is therefore possible for a tube orbit that starts, for example, in the lower left-hand region of one of these maps to end up as a box orbit at some point in it its history (but the timescale for this to happen is likely prohibitively long).
+
+The strongest chaotic regions appear in both the Lyapunov exponent and frequency diffusion rate (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqmap}, right), but the frequency diffusion rate is more accurate and thus more sensitive to mild chaos.
+It is unclear whether these regions will be dynamically important on realistic timescales, but we have confirmed from two independent methods that there are regions of strongly chaotic orbits even in this simple, triaxial potential.
+
+\subsection{Ensemble properties: mixing of debris}
+
+% TODO: with 3 N-body simulations of regular, weakly chaotic, strongly chaotic, show that ensembles reproduce qualitative behavior
 
@@
 
-\section{Results}\label{sec:results}
-
-\subsection{Orbital properties: frequencies, frequency diffusion, and Lyapunov exponents}
-
-We start by studying the properties of individual orbits in the potential described above.
-We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy $E=XXX$.
-We fix $v_x = v_z = 0$, and compute $v_y$ from the energy.
-The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic.
-Tube orbits are generally less stochastic than box orbits --- the other major class of orbits in triaxial potentials --- which tend to plunge deep into the inner regions of the potential, however there are appreciable regions of stochastic tube orbits.
-Thin tidal streams are thought to form along tube orbits rather than box orbits because of the fast disruption expected for stellar systems on radially plunging orbits.
-The major classes of stable tube orbits circulate about either the major or minor axis.
-The most prominent orbits in this grid are the short-axis and outer long-axis tubes, but this grid generates all of the major orbit classes: short-axis tubes, long-axis inner tubes, long-axis outer tubes, stochastic intermediate-axis, and box orbits.
-
-We compute the Lyapunov exponents for all orbits in the grid by integrating the orbits for 1000 orbital periods and use the method described in Section~\ref{sec:lyap} to estimate the maximum Lyapunov exponent.
-Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $x--z$ plane) --- each pixel corresponds to an orbit, and the greyscale intensity is related to the logarithm of the Lyapunov time, $t_\lambda = \lambda_{\rm max}^{-1}$.
-This map shows that by numerically estimating the MLE over this fixed integration window, the majority of orbits appear to be regular or have long Lyapunov times.
-
-%\begin{figure}[!h]
-%\centering
-%	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_xz.png}
-%	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
-%	\label{fig:lyapmap}
-%\end{figure}
-
-For all orbits, we also integrate instead for $\approx$200 orbital periods (Section~\ref{sec:integration}) and use NAFF to compute the fundamental frequencies for the two consecutive sections of 100 orbital periods.
-Figure~\ref{fig:freqmap} (left) shows a map of the fundamental frequency ratios for all tube orbits from the input grid (we plot the frequencies estimated from the first 100 orbital periods).
-This panel demonstrates again that many of the orbits are regular: the transformed, even-spaced sampling of the initial conditions is preserved in the frequency transformation, implying that the frequencies for many orbits do not change appreciably.
-Also apparent is resonant trapping, which manifests as over-dense, linear features in the frequency map.
-For approximately 2\% of the orbits, NAFF fails to recover the frequencies (similar difficulties were reported in \cite{valluri98} and \cite{papaphilippou96} for recovering certain tube orbit frequencies).
-
-%\begin{figure}[h]
-%\centering
-%	\includegraphics[width=0.45\columnwidth] {figures/freqmap_xz.png}
-%	\includegraphics[width=0.45\columnwidth] {figures/diffusion_map_xz.png}
-%	\caption{(\emph{Left} Frequency map of ...
-For short-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $z$ axis, $\Omega_\phi$, and the oscillation along the $z$ axis, $\Omega_z$.
-For long-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $x$ axis, $\Omega_\phi$, and the oscillation along the $x$ axis, $\Omega_x$.
-Box orbits are not shown.}
-%	\label{fig:freqmap}
-%\end{figure}
-
-For each orbit, we also compute the frequency diffusion rate per orbit, $\fdrate$ (Equation~\ref{eq:fdrate}), using the frequencies computed for the two integrated sections of orbit.
-Figure~\ref{fig:freqmap} (right) shows the grid of initial conditions for all orbits colored by $\log_{10}(\fdrate)$ (white points are strongly chaotic, black are regular).
-A rich structure of resonances appears in this frequency diffusion map that were not seen in the Lyapunov exponent map: this highlights the accuracy of NAFF.
-We again see that the tube orbits in this potential are mostly regular or only mildly chaotic, however islands of stronger chaos do appear, especially at the intersections of unstable resonances.
-The unstable (white) regions in these maps reveal the Arnold web of resonances at this energy.
-These unstable regions are interconnected and it is therefore possible for a tube orbit that starts, for example, in the lower left-hand region of one of these maps to end up as a box orbit at some point in it its history (but the timescale for this to happen is likely prohibitively long).
-
-The strongest chaotic regions appear in both the Lyapunov exponent and frequency diffusion rate (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqmap}, right), but the frequency diffusion rate is more accurate and thus more sensitive to mild chaos.
-It is unclear whether these regions will be dynamically important on realistic timescales, but we have confirmed from two independent methods that there are regions of strongly chaotic orbits even in this simple, triaxial potential.
-
-\subsection{Ensemble properties: mixing of debris}
-
 Our goal is to characterize the morphology of tidal debris along orbits generic to triaxial potentials.
@@
 
+\subsection{Frequency variance}\label{sec:freqvar}
+
+Both the Lyapunov exponent and the frequency diffusion rate measure mean, long-term rates of chaotic diffusion.
+As discussed in Section~\ref{sec:nldreview}, frequency diffusion may occur either across a resonance layer or along a stochastic layer.
+If a weakly chaotic orbit is confined (by other nearby, non-overlapping resonances or stable regions), the mean drift of an orbit in frequency space will be small if computed over timescales long compared to the orbital time but short compared to the Arnold diffusion time.
+For the potential considered in this paper, these are precisely the timescales for which the frequency diffusion rate of the previous section were computed ($\sim$100's of orbital periods).
+This window length was chosen in order to have a strong enough signal in the frequency spectrum of chaotic orbits to measure the most significant frequencies using NAFF.
+
+In this section, we instead wish to resolve the short-time behavior of the frequency diffusion (corresponding to motion across resonance layers).
+For each orbit, we now instead use a rolling window to analyze the evolution of the frequency spectrum on shorter times.
+We use a window with a width of 50 orbital periods and shift the window by half an orbital period between each calculation of the most significant frequencies.
+[Why this window width?
+Stride?]
+
 %\begin{figure}[!h]
@@
 %\end{figure}
-
-\subsection{Observable consequences of mixing}
-
-Only effectively looking at debris disrupted from a single pericenter --- debris near progenitor constantly repopulated as more and more stuff disrupts out
-
-Streams should be shorter? How to test this? Length of stream in angle at isodensity contour?
-
-\begin{align}
-	E[\rho(\bs{x},t_j)] &= \int \rho(\bs{x},t_j) \, \rho(\bs{x},t_j) \, \dd \bs{x}\\
-	&\approx \frac{1}{N}\sum \rho(\bs{x}_i,t_j)
-\end{align}
-So we can actually estimate the mean density by evaluating at the particles and taking the mean.
 
@@
 \acknowledgements
-The authors wish to thank  Andreas K\"upper, Robyn Sanderson, David Hogg, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
+The authors wish to thank ... Robyn Sanderson, David Hogg, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 
@@
 % Figure 1
-\begin{figure*}[!p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/potential.pdf}
-\caption{Rotation curve along $x$, $y$, and $z$ axes for the triaxial, NFW potential } \label{fig:potential}
-\end{center}
-\end{figure*}
+%\begin{figure*}[!p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/potential.pdf}
+%\caption{Rotation curve along $x$, $y$, and $z$ axes for the triaxial, NFW potential } \label{fig:potential}
+%\end{center}
+%\end{figure*}
 

### 2015-05-13 ms.tex
made appendix

@@
 
-\title{Tidal streams in triaxial systems}
+\title{Chaotic diffusion of tidal streams}
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
@@
 
-This paper is organized as follows: In Section~\ref{sec:methods}, we describe our choice of potential and method for numerical orbit integration.
-We review relevant nonlinear dynamics in Section~\ref{sec:nldreview}.
+This paper is organized as follows: We review relevant nonlinear dynamics in Section~\ref{sec:nldreview}.
+In Section~\ref{sec:methods}, we describe our choice of potential and method for numerical orbit integration.
 We present our results in Section~\ref{sec:results} and discuss their implications in Section~\ref{sec:discussion}.
@@
 Along chaotic orbits, Hessian not defined and frequencies diffuse, ...
-
-\section{Numerical methods}\label{sec:methods}
-
-\subsection{Potential choice}\label{sec:potential}
-
-The density distributions within dark matter halos formed in cosmological N-body simulations are generically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}.
-With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo \citep[e.g.,][]{dubinski??, recent NiHao stuff}.
-It is less clear what happens to the outer halo where tidal streams are more readily found.
-%; for this reason, we ask what we would expect to happen to tidal streams if the halo in which they form is strongly triaxial.
-
- \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the classic NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these halos.
-JS02 finds median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{??, NiHao} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
-JS02 finds significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).
-
-All of these parameters are specified in terms of the \emph{density}; for orbit analysis, we need to determine the form of the potential in terms of these parameters, which, in general, requires numerical integration of the density at each position of interest.
-For computational efficiency, many authors instead express the triaxiality in the form of the potential, but this can lead to unphysical situations where the density becomes negative.
-\citet{leesuto03} derive a perturbative expansion of the potential integral for a triaxial NFW density and show that the expansion is accurate even for modest axis ratios (e.g., the median values shown above).
-
-In this work, we use the triaxial potential expression from \citet{leesuto03}, parametrized in a slightly different manner.
-In terms of spherical coordinates\footnote{$(r,\phi,\theta)$ = (radius, azimuth, colatitude)} with the radius normalized by the scale radius, $u = r/r_s$
-\begin{align}
-	\Phi(u,\phi,\theta) &\approx \frac{v_c^2}{A}\left[F_1(u) + \frac{1}{2}(e_b^2 + e_c^2)F_2(u) + \frac{1}{2} [(e_b\sin\theta \sin\phi)^2 + (e_c\cos\theta)^2] F_3(u) \right]\\
-	A &= \left(\ln2 - \frac{1}{2}\right) + \left(\ln2-\frac{3}{4}\right) (e_b^2 + e_c^2)
-\end{align}
-where $e_b = \sqrt{1 - (b/a)^2}$, $e_c = \sqrt{1 - (c/a)^2}$, and $v_c$ is the circular velocity at the scale radius, $r_s$, for the spherical case.
-The functions $F_i(u)$ are given in the appendix of \cite{leesuto03}.
-We chose $r_s=20~{\rm kpc}$ and $v_c = 175~{\rm km}~{\rm s}^{-1}$ by taking the mean halo concentration for a ${\rm M}_{vir} \approx 10^{12}~\msun$ halo, $c_e\approx5$, from \cite{jing02} and by assuming $R_{vir}\approx200~{\rm kpc}$.
-Figure~\ref{fig:potential} shows equipotential contours of the potential in projection.
-
-This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies.
-Realistic Galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11, NiHao}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}.
-We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
-
-\subsection{Orbit integration}\label{sec:integration}
-
-We use the Dormand-Prince 8th-order Runge-Kutta scheme \citep{prince81} to integrate orbits in the above potential.
-For all orbits we ensure that energy is conserved to $\Delta E/E_0 \leq 10^{-8}$ by the end of integration, however most orbits conserve energy to $\Delta E/E_0 \approx 10^{-13}$.
-Unless otherwise specified, the integration timesteps are chosen so that there are 512 steps per strongest orbital period component.
 
@@
 %
+
+\section{Numerical methods}\label{sec:methods}
+
+\subsection{Potential choice}\label{sec:potential}
+
+The density distributions within dark matter halos formed in cosmological N-body simulations are generically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}.
+With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo \citep[e.g.,][]{dubinski??, recent NiHao stuff}.
+It is less clear what happens to the outer halo where tidal streams are more readily found.
+%; for this reason, we ask what we would expect to happen to tidal streams if the halo in which they form is strongly triaxial.
+
+ \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the classic NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these halos.
+JS02 finds median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{??, NiHao} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
+JS02 finds significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).
+
+All of these parameters are specified in terms of the \emph{density}; for orbit analysis, we need to determine the form of the potential in terms of these parameters, which, in general, requires numerical integration of the density at each position of interest.
+For computational efficiency, many authors instead express the triaxiality in the form of the potential, but this can lead to unphysical situations where the density becomes negative.
+\citet{leesuto03} derive a perturbative expansion of the potential integral for a triaxial NFW density and show that the expansion is accurate even for modest axis ratios (e.g., the median values shown above).
+
+In this work, we use the triaxial potential expression from \citet{leesuto03}, parametrized in a slightly different manner.
+In terms of spherical coordinates\footnote{$(r,\phi,\theta)$ = (radius, azimuth, colatitude)} with the radius normalized by the scale radius, $u = r/r_s$
+\begin{align}
+	\Phi(u,\phi,\theta) &\approx \frac{v_c^2}{A}\left[F_1(u) + \frac{1}{2}(e_b^2 + e_c^2)F_2(u) + \frac{1}{2} [(e_b\sin\theta \sin\phi)^2 + (e_c\cos\theta)^2] F_3(u) \right]\\
+	A &= \left(\ln2 - \frac{1}{2}\right) + \left(\ln2-\frac{3}{4}\right) (e_b^2 + e_c^2)
+\end{align}
+where $e_b = \sqrt{1 - (b/a)^2}$, $e_c = \sqrt{1 - (c/a)^2}$, and $v_c$ is the circular velocity at the scale radius, $r_s$, for the spherical case.
+The functions $F_i(u)$ are given in the appendix of \cite{leesuto03}.
+We chose $r_s=20~{\rm kpc}$ and $v_c = 175~{\rm km}~{\rm s}^{-1}$ by taking the mean halo concentration for a ${\rm M}_{vir} \approx 10^{12}~\msun$ halo, $c_e\approx5$, from \cite{jing02} and by assuming $R_{vir}\approx200~{\rm kpc}$.
+Figure~\ref{fig:potential} shows equipotential contours of the potential in projection.
+
+This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies.
+Realistic Galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11, NiHao}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}.
+We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
+
+\subsection{Orbit integration}\label{sec:integration}
+
+We use the Dormand-Prince 8th-order Runge-Kutta scheme \citep{prince81} to integrate orbits in the above potential.
+For all orbits we ensure that energy is conserved to $\Delta E/E_0 \leq 10^{-8}$ by the end of integration, however most orbits conserve energy to $\Delta E/E_0 \approx 10^{-13}$.
+Unless otherwise specified, the integration timesteps are chosen so that there are 512 steps per strongest orbital period component.
+
 \subsection{Lyapunov exponents} \label{sec:lyap}
@@
 The simplest and perhaps most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit.
-Using our definition of $w$ from Equation~\ref{eq:coords}, we can write Hamilton's equations as
-\begin{equation}
-	\dot{w}_i = \mathcal{J}_{ij}\,\frac{\partial H}{\partial w_j} \label{eq:ham}
-\end{equation}
-where $\mathcal{J}_{ij}$ is the $2N \times 2N$ canonical Poisson tensor (also called the symplectic matrix) defined by
-\begin{equation}
-	\mathcal{J}_{ij} = \left( \begin{array}{c:c} 0 & \ident \\ \hdashline -\ident & 0 \end{array} \right)
-\end{equation}
-with $N$-dimensional identity matrices $\ident$.
-If we consider a nearby phase-space position, $w_i'$, separated from $w_i$ by an infinitesimal deviation, $\delta w_i$, such that $w_i' = w_i + \delta w_i$.
-We can expand to linear order in the deviation about the parent orbit and write the equations of motion for the deviation as
-\begin{align}
-	\dot{\delta w_i} = \mathcal{J}_{ij} \frac{\partial^2 H}{\partial w_j \partial w_k} \delta w_k =  \mathcal{J}_{ij}\, D_{jk} \, \delta w_k \label{eq:deviate}
-\end{align}
-where $D_{jk}$ is the Hessian matrix evaluated at the parent orbit.
-For chaotic orbits, the maximum eigenvalue of the solution matrix to Eq.~\ref{eq:deviate} is positive real, leading to exponential divergence of nearby orbits.
-
-Computing the Lyapunov spectrum for a given orbit is often not necessary if one is only interested in characterizing the degree of chaos.
-Instead, it is often more efficient to compute the maximum Lyapunov exponent by estimating the finite-time maximum Lyapunov exponent (FTMLE), defined as
-\begin{equation}
-	l_{\rm max}(t) = \frac{1}{t}\ln \frac{\|\delta \bs{w}(t)\|}{\|\delta \bs{w}_0\|}.\label{eq:lmax}
-\end{equation}
-The maximum Lyapunov exponent is the limit as $t\rightarrow \infty$ of the FTMLE
-\begin{equation}
-	\lambda_{\rm max} = \lim_{t\rightarrow\infty}l_{\rm max}(t). \label{eq:lyapmax}
-\end{equation}
-Numerically computing this quantity is not trivial because (1) obviously the limit to infinity is not possible and (2) the norm of the deviation vector $\|\delta \bs{w}(t)\|$ is expected to increase exponentially for chaotic orbits, leading to numerical problems.
-To circumvent these issues, it is sufficient to instead start a nearby orbit with some small initial deviation with norm $\delta_0$, integrate for a sufficiently small amount of time, $\tau$, then renormalize the deviation back to the initial norm (e.g., cite Bennetin et al.
-1976, Tabor 1989).
-%The following pseudocode outlines this procedure:\footnote{see Gary for Python implementation?}\\
-%\begin{algorithmic}[1]
-%\State {\bf define} orbital integration timestep, $h$, and number of steps, $K$
-%\State {\bf define} initial norm of deviation vector, $\delta_0$, to be sufficiently small
-%\State {\bf define} renormalization integration period, $\tau \ll hK$ 
-%\For{each timestep when integrating the main orbit, $\bs{w}(t)$}
-%\State step forward the orbit and deviation vector orbit by one timestep, $t_{i-1} \rightarrow t_i$
-%\If {a normalization timestep}
-%\State measure and store the length of the deviation vector, $\delta_i = \|\delta \bs{w}_i\|$
-%\State renormalize the length of the deviation vector, $\delta \bs{w}_i = \delta \bs{w}_i (\delta_0/\delta_i)$
-%\EndIf
-%\EndFor 
-%\end{algorithmic}
-The FTMLE after a given number of timesteps, $N$, is then estimated as
-\begin{equation}
-	l_N = \frac{1}{N\tau}\sum_i^N \ln \, \delta_i \label{eq:ftmle}
-\end{equation}
-where $\delta_i=\|\delta \bs{w}(t_i)\|$ and the MLE is
-\begin{equation}
-	\lambda_{\rm max} = \lim_{N\rightarrow \infty} l_N.
-\end{equation}
-
-For most regular orbits, deviations will grow close to linearly or as some power-law of time. 
-%As we have seen (Section~\ref{sec:introduction}), if the orbit is regular, there exists a local transformation to action-angle variables where the angle variables increase linearly with time, $\theta_n \propto \Omega_n t$.
-We can look at small variations around the orbit defined by the above equations:\footnote{Note the implied summation in repeated indices in the above expression.}
-%\begin{align}
-%	\frac{d (\delta \theta_n)}{dt} &= \frac{\partial \Omega_n}{\partial J_m} \delta J_m
-%\end{align}
-%These equations are trivially integrated:
-%\begin{align}
-%	\delta \theta_n(t) &= \delta \theta_n(0) + \left(\frac{\partial \Omega_n}{\partial J_m} \delta J_m \right) t
-%\end{align}
-%It's clear then that the norm of the deviation vector grows linearly with time:
-%\begin{align}
-%	\|\delta \bs{w}(t)\| &= \left[\sum_n (\delta \theta_n(t))^2 + \sum_n (\delta J_n)^2\right]^{1/2}\\
-%	&= \left[\sum_n \left(\delta \theta_n(0) + \frac{\partial \Omega_n}{\partial J_m} \delta J_m t\right)^2 + \sum_n (\delta J_n)^2\right]^{1/2}\\
-%	&\propto t.
-%\end{align}
-From Eq.~\ref{eq:lmax} and Eq.~\ref{eq:lyapmax} it is evident that any deviation vector that grows as a power law with time, $t^k$, will asymptote to 0 from the limit 
-\begin{equation}
-	\lambda_{\rm max} \propto \lim_{t\rightarrow \infty} k \frac{\ln t}{t} = 0.
-\end{equation}
-At long times, the numerically computed MLE should approach 0 as $t^{-1}$ for regular orbits.
-For chaotic orbits, the divergence is exponential, and the limit should converge to the rate of the exponential: the Lyapunov exponent.
-In practice, the MLE is often estimated as the mean of $l_N$ after the summation diverges from power-law behavior.
 
@@
 \bibliography{refs}
-\clearpage
+
+\appendix
+\section{Lyapunov exponents}
+
+%Using the definition of $w$ from Equation~\ref{eq:coords}, we can write Hamilton's equations as \footnote{A sum is implied with any repeated indices.}
+%\begin{equation}
+%	\dot{w}_i = \mathcal{J}_{ik}\,\frac{\partial H}{\partial w_k} \label{eq:ham}
+%\end{equation}
+%where $\mathcal{J}_{ik}$ is the $2N \times 2N$ canonical Poisson tensor (also called the symplectic matrix) defined by
+%\begin{equation}
+%	\mathcal{J}_{ik} = \left( \begin{array}{c:c} 0 & \ident \\ \hdashline -\ident & 0 \end{array} \right)
+%\end{equation}
+%with $N$-dimensional identity matrices $\ident$.
+We will consider a nearby phase-space position, $w_i'$, separated from $w_i$ by an infinitesimal deviation, $\delta w_i$, such that $w_i' = w_i + \delta w_i$.
+We can expand to linear order in the deviation about the parent orbit and write the equations of motion for the deviation as
+%\begin{align}
+%	\dot{\delta w_i} &= \mathcal{J}_{ik} \, \frac{\partial^2 H}{\partial w_k \partial w_m} \, \delta w_m =  \mathcal{J}_{ik}\, D_{km} \, \delta w_m \\
+%	&= A_{im} \, \delta w_m
+%\end{align}
+%where $D_{km}$ is the Hessian matrix evaluated at the parent orbit. The general solution for this equation is
+%\begin{equation}
+%	\delta w_i(t) = B_{im}(t) \, \delta w_i(0)
+%\end{equation}
+%where $\delta w_i(0)$ are the initial conditions for the deviation vector and $B_{im}$
+% For chaotic orbits, the maximum eigenvalue of the solution matrix to Eq.~\ref{eq:deviate} is positive real, leading to exponential divergence of nearby orbits.
+
+If one is only interested in characterizing the degree of chaos, computing the full Lyapunov spectrum for an orbit is often not necessary.
+It is often sufficient to compute an estimate of the maximum Lyapunov exponent by estimating the finite-time maximum Lyapunov exponent (FTMLE).
+Using the definition of $\boldsymbol{w}$ from Equation~\ref{eq:coords}, consider an orbit that is a small deviation away from the parent orbit, $\boldsymbol{w}' = \boldsymbol{w} + \delta\boldsymbol{w}$.
+If the parent orbit is chaotic, the norm of the infinitesimal deviation should grow exponentially with time with some characteristic rate, $\lambda$,
+\begin{equation}
+	\|\delta\boldsymbol{w}(t)\| = e^{\lambda \, t} \, \|\delta\boldsymbol{w}_0\|
+\end{equation}
+\citep[see, e.g.,][]{who?}. From this expression, we see that
+\begin{equation}
+	\lambda(t) = \frac{1}{t}\ln \frac{\|\delta \bs{w}(t)\|}{\|\delta \bs{w}_0\|} \label{eq:mle}
+\end{equation}
+where the maximum Lyapunov exponent is the limit as $t\rightarrow \infty$,
+\begin{equation}
+	\lambda_{\rm max} = \lim_{t\rightarrow\infty}\lambda(t). \label{eq:lmax}
+\end{equation}
+Numerically computing this quantity is not trivial because (1) obviously the limit to infinity is not possible and (2) the norm of the deviation vector $\|\delta \bs{w}(t)\|$ is expected to increase exponentially for chaotic orbits, leading to nonlinear evolution of the deviation and numerical problems.
+To circumvent these issues, it is sufficient to instead start a nearby orbit with some small initial deviation with norm $\delta_0$, integrate for a sufficiently small amount of time, $\tau$, then renormalize the deviation back to the initial norm (e.g., TODO: cite Bennetin et al.
+1976).
+There is no general way to determine $\tau$ except to perform convergence tests.
+%The following pseudocode outlines this procedure:\footnote{see Gary for Python implementation?}\\
+%\begin{algorithmic}[1]
+%\State {\bf define} orbital integration timestep, $h$, and number of steps, $K$
+%\State {\bf define} initial norm of deviation vector, $\delta_0$, to be sufficiently small
+%\State {\bf define} renormalization integration period, $\tau \ll hK$ 
+%\For{each timestep when integrating the main orbit, $\bs{w}(t)$}
+%\State step forward the orbit and deviation vector orbit by one timestep, $t_{i-1} \rightarrow t_i$
+%\If {a normalization timestep}
+%\State measure and store the length of the deviation vector, $\delta_i = \|\delta \bs{w}_i\|$
+%\State renormalize the length of the deviation vector, $\delta \bs{w}_i = \delta \bs{w}_i (\delta_0/\delta_i)$
+%\EndIf
+%\EndFor 
+%\end{algorithmic}
+The FTMLE after a given number of timesteps, $N$, is then estimated as
+\begin{equation}
+	\lambda_N = \frac{1}{t_N}\sum_i^N \ln \frac{\|\delta \bs{w}(t_i)\|}{\|\delta \bs{w}_0\|} \label{eq:ftmle}
+\end{equation}
+and the MLE is estimated after a very long time to approximate the limit of Equation~\ref{eq:lmax}.
+
+For most regular orbits, deviations will grow linearly or as a power-law of time.
+As we have seen in Section~\ref{sec:nldreview}, if the orbit is regular, there exists a local transformation to action-angle variables where the angle variables increase linearly with time, $\theta_i \propto \Omega_i t$.
+We can look at small variations around the angle-space orbit,
+\begin{align}
+	\frac{d (\delta \theta_i)}{dt} &= \frac{\partial \Omega_i}{\partial J_k} \delta J_k
+\end{align}
+These equations are trivially integrated:
+\begin{align}
+	\delta \theta_i(t) &= \delta \theta_i(0) + \left(\frac{\partial \Omega_i}{\partial J_k} \delta J_k \right) t
+\end{align}
+It's clear then that the norm of the deviation vector grows linearly with time:
+\begin{align}
+	\|\delta \bs{w}(t)\| &= \left[\sum_i (\delta \theta_i(t))^2 + \sum_i (\delta J_i)^2\right]^{1/2}\\
+	&= \left[\sum_i \left(\delta \theta_i(0) + \frac{\partial \Omega_i}{\partial J_k} \delta J_k t\right)^2 + \sum_i (\delta J_i)^2\right]^{1/2}\\
+	&\propto t.
+\end{align}
+From Equations~\ref{eq:mle} and \ref{eq:lmax} it is evident that any deviation vector that grows as a power law with time, $t^k$, will asymptote to 0 from the limit 
+\begin{equation}
+	\lambda_{\rm max} \propto \lim_{t\rightarrow \infty} k \frac{\ln t}{t} = 0.
+\end{equation}
+At long times, the numerically computed MLE for a regular orbit should approach 0 as $t^{-1}$.
+For chaotic orbits, the divergence is exponential, and the limit should converge to the rate of the exponential: the Lyapunov exponent.
+In practice, the MLE is often estimated as the mean of $\lambda_N$ after the summation diverges from power-law behavior.
+For weakly chaotic orbits, reliable computation of the MLE may take integration of thousands of orbital periods.
 

### 2015-05-15 ms.tex
more to appendix and cleaning up numbers and text of results

@@
 \usepackage{amssymb,amsmath,amsbsy}
+\usepackage{booktabs}
 \usepackage{bbold}
-\usepackage{arydshln}
-\usepackage{algpseudocode}
-\renewcommand\algorithmicthen{}
-\renewcommand\algorithmicdo{}
 \usepackage{mathrsfs}
@@
 \begin{equation}
-	w_i(t) = \sum_k^\infty a_k \, e^{i\,\omega_k\,t}; \, (a_k \in \mathbb{C}).
+	w_i(t) = \sum_k^\infty a_k \, e^{i\,\omega_k\,t}; \, (a_k \in \mathbb{C}). \label{eq:fourier}
 \end{equation}
@@
 \begin{align}
-	\omega_k &= \boldsymbol{n_k} \cdot \boldsymbol{\Omega}; \, (\boldsymbol{n}_k \in \mathbb{Z}^3).
+	\omega_k &= \boldsymbol{n_k} \cdot \boldsymbol{\Omega}; \, (\boldsymbol{n}_k \in \mathbb{Z}^3). \label{eq:fourierfreq}
 \end{align}
@@
 We chose $r_s=20~{\rm kpc}$ and $v_c = 175~{\rm km}~{\rm s}^{-1}$ by taking the mean halo concentration for a ${\rm M}_{vir} \approx 10^{12}~\msun$ halo, $c_e\approx5$, from \cite{jing02} and by assuming $R_{vir}\approx200~{\rm kpc}$.
-Figure~\ref{fig:potential} shows equipotential contours of the potential in projection.
+Figure~\ref{fig:potential} shows equipotential contours of this potential in projection.
 
@@
 We use the Dormand-Prince 8th-order Runge-Kutta scheme \citep{prince81} to integrate orbits in the above potential.
+Specifically, we use a \texttt{Python} wrapper over the \texttt{C} implementation by \cite{hairer93}.
 For all orbits we ensure that energy is conserved to $\Delta E/E_0 \leq 10^{-8}$ by the end of integration, however most orbits conserve energy to $\Delta E/E_0 \approx 10^{-13}$.
-Unless otherwise specified, the integration timesteps are chosen so that there are 512 steps per strongest orbital period component.
+Unless otherwise specified the integration timesteps are chosen so that there are 512 steps per strongest orbital period component, but the integrator uses adaptive stepping between each main step in order to satisfy a specified tolerance (we set the absolute tolerance to $\approx$100 times machine precision, $\texttt{atol} = 10^{-13}$).
 
@@
 
-% TODO: move most of this to appendix? leave only discussion?
+The most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit.
+The MLE measures the mean rate of divergence of two infinitesimally separated orbits and is only strictly defined in terms of a limit that goes to infinite time.
+Thus, we can never truly compute the MLE and it can take integration for many thousands of orbital periods to compute a converged numerical approximation of the MLE for a moderately chaotic orbit.
+In this work, we use the algorithm introduced by \cite{wolf85} for computing the MLE (for more a more detailed description of this algorithm, see Appendix~\ref{sec:lyapapdx}).
 
-The simplest and perhaps most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit.
+The MLE, $\lambda_{\rm max}$, is interpreted as a rate that determines the nature of exponential divergence of infinitesimally close chaotic orbits.
+It is therefore useful to consider the corresponding $e$-folding time,
+\begin{equation}
+	t_{\rm \lambda} = \frac{1}{\lambda_{\rm max}}.
+\end{equation}
+We will use this as the prediction from the Lyapunov exponent for the timescale over which chaos should be dynamically important for a given orbit.
 
-\subsection{Numerical estimation of fundamental frequencies}\label{sec:naff}
+\subsection{Numerical approximation of fundamental frequencies}\label{sec:naff}
 
-% TODO: move most to appendix, leave only discussion?
+Bounded, regular orbits in a triaxial potential will have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion.
+The motion in any canonical coordinate can therefore be decomposed as a Fourier sum (Equation~\ref{eq:fourier}) where the Fourier frequencies are linear, integer combinations of the fundamental frequencies (Equation~\ref{eq:fourierfreq}).
+\cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hamming-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
+It has been shown that using a filter makes the accuracy in determining the frequencies converge much faster than the typical $\inttime^{-1}$ expected for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time.
+Once the strongest frequencies of the time series are found, the spectrum must be searched for the three fundamental frequencies.
 
-Bounded, regular orbits in a triaxial potential will have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion (Eq.~\ref{eq:aafreq}).
-This implies that for any canonical coordinate, $q$, the motion in this coordinate can be decomposed as
-\begin{equation}
-	q(t) = \sum^\infty_k a_k e^{i \omega_k t}
-\end{equation}
-where these frequencies, $\omega_k$, are linear, integer combinations of the fundamental frequencies
-\begin{equation}
-	\omega_k = \bs{n}_k \cdot \bs{\Omega},
-\end{equation}
-the $a_k$ are complex amplitudes, and the $\bs{n}_k$ are integer vectors.
-\cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hanning-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
-It has been shown that using a Hanning filter makes the accuracy in determining the frequencies scale as $\inttime^{-4}$ instead of $\inttime^{-1}$ for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time.
-In each complex time series, once the frequency is found, the amplitude of the term is found by projecting back on to the function itself.
-The component is then subtracted from the time series, and this procedure continues iteratively until some breaking condition (for example, the amplitude of the next term is below some threshold).
-We follow \cite{valluri98} and refer to this method as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) even though it differs slightly from the original method from Laskar.
+This method is referred to as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) and has been used extensively in planetary dynamics \citep{} and galaxy dynamics \citep{valluri98}, especially in the study of orbits in triaxial systems.
+We have implemented and tested our own open-source implementation of NAFF in \texttt{Python}.\footnote{This code is publicly available through the \texttt{Gary} package, \url{http://...}} Though our implementation differs slightly from the original definition and from that used in \cite{valluri98}, we still refer to the method as NAFF.
+For more details about NAFF, see Appendix~\ref{sec:naffapdx} or \cite{papaphilippou96, laskar, etc.}.
 
 If an orbit is chaotic, the motion can no longer be expressed in terms of a single set of fundamental frequencies because the actions and frequencies of the orbit change with time.
-For a weakly chaotic orbit, the orbit may act consistently periodic over long windows of time.
+For a weakly chaotic orbit, the orbit may appear consistently periodic over long windows of time.
 NAFF will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change character with time.
 For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two separate sections of orbit.
-The frequencies picked out by NAFF for such orbits will therefore represent the average periodic nature of the orbit over a given period, and these will vary dramatically if comparing two consecutive integration periods.
-Therefore, the frequency diffusion rate per orbit, $\fdrate$, is a good predictor for how chaotic an orbit is:
+The frequencies picked out by NAFF for such orbits will therefore represent the average periodic nature of the orbit over a given period.
+We define the frequency diffusion rate per orbit, $\fdrate$, as
 \begin{equation}
-	\fdrate = \max_{\bs{a}} \, (\bs{\Omega}_{T_1} - \bs{\Omega}_{T_2}) / N_T \label{eq:fdrate}
+	\fdrate = \max_{\bs{a}} \, (\bs{\Omega}_{T_2} - \bs{\Omega}_{T_1}) / N_T \label{eq:fdrate}
 \end{equation}
@@
 The frequency diffusion rate as defined above (e.g., measured in the frequency with the largest amplitude) is better correlated with the strength of chaos in an orbit \citep{valluri??} but differs from the original definition of \cite{valluri98}.
-%; we have found that for extremely chaotic orbits --- where peaks in the FFT overlap --- NAFF fails to recover any frequencies
 
-NAFF has been applied to many problems in Solar System and Galactic dynamics, especially in the study of orbits in triaxial systems.
-For more details about the method, see \cite{papaphilippou96, laskar, etc.}.
-We have implemented and tested our own implementation of NAFF in \texttt{Python}.
-This code is publicly available through the \texttt{Gary} package.\footnote{link to repo} Figure~\ref{fig:logfreqs} shows a test of the implementation in which we reproduce the frequency map at a fixed energy of an axisymmetric, logarithmic potential \cite[][pg.
-260, Figure~3.45]{binneytremaine}.
-Plotted are the (Cartesian) frequency ratios recovered for a grid of iso-energy, box orbits integrated in the potential
+From this rate, we can compute a timescale over which we would expect order-unity changes to the fundamental frequencies of an orbit,
 \begin{equation}
-	\Phi(x,y,z) = \frac{1}{2}\ln\left(x^2 + (y/0.9)^2 + (z/0.7)^2 + 0.1\right). \label{eq:logpotential}
+	t_\Omega = \frac{2 \pi}{\bs{\Omega}_{T_1}} \frac{1}{\fdrate}.
 \end{equation}
-In such a map, stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
-The regularity of the points in this map reflects the input grid of initial conditions.
-Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.
 
-NAFF recovers the fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL96;][]{papaphilippou96}.
-PL96 show that a good choice of coordinates for tube orbits are the Poincar\'e symplectic polar coordinates, a set of canonical coordinates similar to cylindrical coordinates.
-When computing the frequencies for tube orbits, we first align the circulation about the $z$-axis through rotation, transform to Poincar\'e polar coordinates, then use NAFF to measure the fundamental frequencies.
-We could equivalently use the Cartesian time series, but the convergence of terms is slower (the amplitudes of successive terms decrease slower for Cartesian coordinates).
-We have tested that our implementation of NAFF returns the same fundamental frequencies in either case for a set of tube orbits.
-For box orbits, the motion is close to separable in each Cartesian component and we therefore use Cartesian coordinates for estimating the frequencies for these orbits.
+% =====================================================================
+%	Results
+%
+\section{Results}
 
-[Can see lots of cool structure, but rates are tiny! Why does this manifest itself over 10-20 periods?]
+In previous work, we found that tidal debris from a simulated globular cluster (Palomar 5) on an orbit with a long Lyapunov time can form large ``fans'' of debris similar to those seen for progenitor systems on strongly chaotic orbits \citep{pearson15, fardal14}.
+Here, to illustrate the methods described above, we repeat the calculation of the Lyapunov exponent for the Pal 5, stream-fanning orbit of \cite{pearson15} with a longer integration time and compute the frequency diffusion rate.
+From these, we measure the Lyapunov time and frequency diffusion time.
 
-\subsection{Comparison of methods}\label{sec:comparison}
+Figure~\ref{fig:pal5} shows the finite-time Lyapunov exponent computed for the Pal 5 orbit --- shaded region shows the ``converged'' part of the estimate that is used to compute the maximum Lyapunov exponent.
+By inverting the rate, we find $t_\lambda = XX~{\rm Gyr}$, consistent with our finding in \cite{pearson15}.
+We then compute the frequency diffusion rate (Equation~\ref{eq:fdrate}) using two consecutive integration windows with length $\approx$50 orbital periods.
+We find $\fdrate = 10^{-XX}$, which corresponds to a frequency diffusion time, $t_\Omega = XX~{\rm Gyr}$.
+Though this orbit is computed in a different potential than will be used below, this demonstrates the apparent discrepancy between the timescales over which chaos is estimated to be relevant (XX's of Gyrs) and the integration time over which stream-fanning becomes apparent ($<8$ Gyr).
+Below, we [...] try to understand by looking at grids of orbits...
 
-We choose two nearby short-axis tube orbits in the potential described in Section~\ref{sec:potential} in order to demonstrate the two methods described above.
-The orbits were chosen from a grid of initial conditions in the $x$-$z$ plane (described in later sections) because they are very close in phase-space but display very different behavior.
-In particular, Orbit A is regular or at least weakly chaotic and Orbit B is strongly chaotic.
-Figure~\ref{fig:two-orbits} shows the two orbits in projection --- Orbit A is shown in the left column and Orbit B the right.
-Two consecutive integration periods for each orbit are shown in the two rows; note the change in appearance of Orbit B (right column).
-The initial conditions for the two orbits in physical units and Cartesian coordinates are:
-\begin{align}
-	\bs{x}_{0}^{(A)} &= (27.85, 0, 23.35)~{\rm kpc}\\
-	\bs{v}_{0}^{(A)} &= (0, 115.37, 0)~{\rm km~s}^{-1}\\
-	\bs{x}_{0}^{(B)} &= (27.85, 0, 23.10)~{\rm kpc}\\
-	\bs{v}_{0}^{(B)} &= (0, 116.43, 0)~{\rm km~s}^{-1}.
-\end{align}
+\subsection{Part I: Lyapunov and frequency diffusion times}\label{sec:results1}
 
-Figure~\ref{fig:lyap-orbits} shows $l_N$ computed for the two orbits described above.
-The MLE estimate continues towards zero for the regular orbit (left) but saturates for the chaotic orbit (right).
-For mildly chaotic or ``sticky'' orbits, it can take integration periods of many thousands of orbits to determine a reliable estimate of the Lyapunov exponent.
-In comparison, owing to the precision of NAFF, detecting small frequency diffusion rates can be done with only $\sim$100 orbital periods.
-Figure~\ref{fig:fdiff-orbits} shows the three fundamental frequencies computed for a sliding window of 12 sections of orbit (each section is $\sim$100 orbital periods in length) for Orbit A (left) and Orbit B (right).
-Dashed horizontal line is the estimate of the uncertainty in the numerical computation of the frequencies from Orbit A.
-The value of each frequency in the first window is subtracted from each subsequent window.
-The frequencies only show minor deviations for Orbit A (left panel) whereas there is significant diffusion in Orbit B (right panel).
-
-\section{Results}\label{sec:results}
-
-\subsection{Orbital properties: frequencies, frequency diffusion, and Lyapunov exponents}
-
-We start by studying the properties of individual orbits in the potential described above.
-We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy $E=XXX$.
+We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy (per unit mass) chosen to span a range of distances comparable to the scale radius of the potential (Section~\ref{sec:potential}), $E=-(397.2~{\rm km~s}^{-1})^2$ (in physical units).
 We fix $v_x = v_z = 0$, and compute $v_y$ from the energy.
@@
 
-We compute the Lyapunov exponents for all orbits in the grid by integrating the orbits for 1000 orbital periods and use the method described in Section~\ref{sec:lyap} to estimate the maximum Lyapunov exponent.
-Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $x--z$ plane) --- each pixel corresponds to an orbit, and the greyscale intensity is related to the logarithm of the Lyapunov time, $t_\lambda = \lambda_{\rm max}^{-1}$.
-This map shows that by numerically estimating the MLE over this fixed integration window, the majority of orbits appear to be regular or have long Lyapunov times.
+To compute the Lyapunov exponents, we integrate all orbits in the grid for 1000 orbital periods and use the method described in Section~\ref{sec:lyap}.
+Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $xz$ plane) --- each pixel corresponds to an orbit, and the logarithm of the Lyapunov time (in units of Gyr) is mapped to greyscale intensity.
+The darker pixels have shorter Lyapunov times and are more chaotic.
+By numerically estimating the MLE over this fixed integration window, the majority of orbits appear to be regular because they have exceedingly long Lyapunov times.
 
-%\begin{figure}[!h]
-%\centering
-%	\includegraphics[width=0.4\columnwidth] {figures/lyap_map_xz.png}
-%	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
-%	\label{fig:lyapmap}
-%\end{figure}
+For each orbit, we also separately integrate for $\approx$100 orbital periods and use NAFF to compute the fundamental frequencies for the two consecutive sections of 50 orbital periods.
+Figure~\ref{fig:freqdiff} shows the same grid of initial conditions, but now the greyscale intensity is set by the logarithm of the frequency diffusion time.
+The darker pixels have shorter frequency diffusion times and are more chaotic.
+This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential.
+This map highlights the accuracy of NAFF --- weak chaos is detected over $\approx$100 orbital periods, compared to the many thousands of orbits it would take to detect such features with the maximum Lyapunov exponent.
+The tube orbits in this potential are mostly regular or only mildly chaotic --- the largest regular regions are associated with the short-axis and long-axis tube orbits --- however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.
 
-For all orbits, we also integrate instead for $\approx$200 orbital periods (Section~\ref{sec:integration}) and use NAFF to compute the fundamental frequencies for the two consecutive sections of 100 orbital periods.
-Figure~\ref{fig:freqmap} (left) shows a map of the fundamental frequency ratios for all tube orbits from the input grid (we plot the frequencies estimated from the first 100 orbital periods).
-This panel demonstrates again that many of the orbits are regular: the transformed, even-spaced sampling of the initial conditions is preserved in the frequency transformation, implying that the frequencies for many orbits do not change appreciably.
-Also apparent is resonant trapping, which manifests as over-dense, linear features in the frequency map.
-For approximately 2\% of the orbits, NAFF fails to recover the frequencies (similar difficulties were reported in \cite{valluri98} and \cite{papaphilippou96} for recovering certain tube orbit frequencies).
-
-%\begin{figure}[h]
-%\centering
-%	\includegraphics[width=0.45\columnwidth] {figures/freqmap_xz.png}
-%	\includegraphics[width=0.45\columnwidth] {figures/diffusion_map_xz.png}
-%	\caption{(\emph{Left} Frequency map of ...
-For short-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $z$ axis, $\Omega_\phi$, and the oscillation along the $z$ axis, $\Omega_z$.
-For long-axis tubes, the frequencies plotted are the radial frequency, $\Omega_r$, the azimuthal frequency around the $x$ axis, $\Omega_\phi$, and the oscillation along the $x$ axis, $\Omega_x$.
-Box orbits are not shown.}
-%	\label{fig:freqmap}
-%\end{figure}
-
-For each orbit, we also compute the frequency diffusion rate per orbit, $\fdrate$ (Equation~\ref{eq:fdrate}), using the frequencies computed for the two integrated sections of orbit.
-Figure~\ref{fig:freqmap} (right) shows the grid of initial conditions for all orbits colored by $\log_{10}(\fdrate)$ (white points are strongly chaotic, black are regular).
-A rich structure of resonances appears in this frequency diffusion map that were not seen in the Lyapunov exponent map: this highlights the accuracy of NAFF.
-We again see that the tube orbits in this potential are mostly regular or only mildly chaotic, however islands of stronger chaos do appear, especially at the intersections of unstable resonances.
-The unstable (white) regions in these maps reveal the Arnold web of resonances at this energy.
-These unstable regions are interconnected and it is therefore possible for a tube orbit that starts, for example, in the lower left-hand region of one of these maps to end up as a box orbit at some point in it its history (but the timescale for this to happen is likely prohibitively long).
-
-The strongest chaotic regions appear in both the Lyapunov exponent and frequency diffusion rate (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqmap}, right), but the frequency diffusion rate is more accurate and thus more sensitive to mild chaos.
+The strongest chaotic regions appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}), but the timescales [...].
+Figure~\ref{fig:lyapvfreqdiff} shows the Lyapunov time plotted against the frequency diffusion time [...].
+frequency diffusion rate is more accurate and thus more sensitive to mild chaos.
 It is unclear whether these regions will be dynamically important on realistic timescales, but we have confirmed from two independent methods that there are regions of strongly chaotic orbits even in this simple, triaxial potential.
 
-\subsection{Ensemble properties: mixing of debris}
+The frequency diffusion time grid reveals the intricate resonant structure of the potential considered, but the majority of the orbits have predicted chaotic timescales that correspond to many hundreds or thousands of orbital times.
+In the next section, we compute the mixing rates for finite-volume ensembles of orbits and compare the density evolution of the ensembles to the predictions from the single-orbit chaos diagnostics of this section.
+
+\subsection{Part II: Ensemble properties and mixing} \label{sec:results2}
 
@@
 This raises an important consideration: do finite deviations diverge faster than predicted from local measures of chaos?
-Does the Lyapunov time or frequency rate predict the timescale over which a finite phase-space volume (e.g., tidal debris) stays ``coherent''?
+Does the Lyapunov or frequency diffusion time predict the timescale over which a finite phase-space volume (e.g., tidal debris) stays ``coherent''?
+
+We start by considering three particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular orbit, a weakly chaotic orbit, and a strongly chaotic orbit.
+The orbits were chosen at random from a single column of the orbit grid and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}.
+[...
+N-body simulations, expectation that only strongly chaotic orbit show strange morph, yadda]
+
+\begin{table*}[ht]
+\begin{center}
+	\begin{tabular}{c | c c c c c c c c }
+		ID & $x$ & $y$ & $z$ & $v_x$ & $v_y$ & $v_z$ & $t_\lambda$ & $t_\Omega$ \\\toprule
+		R & 10. & 0. & 0. & 0. & 0. & 0. & XX & YY\\
+		\midrule
+		W & 10. & 0. & 0. & 0. & 0. & 0. & XX & YY\\
+		\midrule
+		S & 10. & 0. & 0. & 0. & 0. & 0. & XX & YY\\
+		\bottomrule
+		\end{tabular}
+	\caption{TODO. \label{tbl:three-orbits}}
+\end{center}
+\end{table*}
+
+% TODO: table with initial conditions
 
@@
 
-\subsection{Frequency variance}\label{sec:freqvar}
+\subsection{Part III: Short-time frequency evolution}\label{sec:results3}
 
@@
 \appendix
-\section{Lyapunov exponents}
+\section{Lyapunov exponents} \label{sec:lyapapdx}
 
@@
 
+\section{Numerical Approximation of Fundamental Frequencies}\label{sec:naffapdx}
+
+Figure~\ref{fig:logfreqs} shows a test of the implementation in which we reproduce the frequency map at a fixed energy of an axisymmetric, logarithmic potential \cite[][pg.
+260, Figure~3.45]{binneytremaine}.
+Plotted are the (Cartesian) frequency ratios recovered for a grid of iso-energy, box orbits integrated in the potential
+\begin{equation}
+	\Phi(x,y,z) = \frac{1}{2}\ln\left(x^2 + (y/0.9)^2 + (z/0.7)^2 + 0.1\right). \label{eq:logpotential}
+\end{equation}
+In such a map, stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
+The regularity of the points in this map reflects the input grid of initial conditions.
+Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.
+
+NAFF recovers the fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL96;][]{papaphilippou96}.
+PL96 show that a good choice of coordinates for tube orbits are the Poincar\'e symplectic polar coordinates, a set of canonical coordinates similar to cylindrical coordinates.
+When computing the frequencies for tube orbits, we first align the circulation about the $z$-axis through rotation, transform to Poincar\'e polar coordinates, then use NAFF to measure the fundamental frequencies.
+We could equivalently use the Cartesian time series, but the convergence of terms is slower (the amplitudes of successive terms decrease slower for Cartesian coordinates).
+We have tested that our implementation of NAFF returns the same fundamental frequencies in either case for a set of tube orbits.
+For box orbits, the motion is close to separable in each Cartesian component and we therefore use Cartesian coordinates for estimating the frequencies for these orbits.
+
 % Figure 1
-%\begin{figure*}[!p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/potential.pdf}
-%\caption{Rotation curve along $x$, $y$, and $z$ axes for the triaxial, NFW potential } \label{fig:potential}
-%\end{center}
-%\end{figure*}
+% TODO: should I show density? cool 3D render of density?
+\begin{figure*}[!p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/potential.pdf}
+\caption{Equipotential contours for the triaxial NFW potential considered in this work. There are eight contour levels evenly spaced and linear in the value of the potential. } \label{fig:potential}
+\end{center}
+\end{figure*}
 
 % Figure 2
-%\begin{figure*}[!p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/orbit-tree.pdf}
-%\caption{A rough classification of the types of orbits found in triaxial potentials.
-Regular orbits are classified by the number of resonance relations obeyed: conditionally periodic (\emph{none}), uni-resonant (\emph{one}), bi-resonant (\emph{two}).
-Irregular or chaotic orbits are simply separated into two classes: strongly and weakly chaotic, defined by the timescales of interest in a problem.
-When the characteristic chaotic timescale is much longer than, for example, the age of the universe, the orbit is only weakly chaotic.
-} \label{fig:orbit-tree}
-%\end{center}
-%\end{figure*}
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/lyap-orbits.pdf}
+\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over YYYY orbital periods for the stream-fanning Palomar 5 orbit of \cite{pearson15}.
+The FTMLE saturates to the Lyapunov exponent in the shaded region and is estimated to be $\lambda_{\rm max} \approx XXX \, {\rm Gyr}^{-1}$, corresponding to a Lyapunov time of $t_\lambda \approx XX \, {Gyr}$.
+Over-plotted on both panels are lines (dashed, blue) that show $t^{-0.9}$ behavior.} \label{fig:pal5}
+\end{center}
+\end{figure*}
 
 % Figure 3
-%\begin{figure*}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/log-freqmap.png}
-%\caption{Ratio of cartesian frequencies for a grid of iso-energy box orbits integrated in the potential of Equation~\ref{eq:logpotential}.
-Stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
-The regularity of the points in this map reflects the input grid of initial conditions.
-Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.} \label{fig:logfreqs}
-%\end{center}
-%\end{figure*}
-%
-%% Figure 4
-%\begin{figure*}[!p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/two-orbits.pdf}
-%\caption{ Two orbits integrated in the triaxial potential described in Section~\ref{sec:potential}.
-Left panels (black) show two consecutive sections of orbit, where the integration time for each section (rows) is set to 100 times the maximum orbital period.
-Right panels (red) shows the same for the chaotic orbit.
-The appearance of the orbit in configuration space changes appreciably over timescales of tens of orbital periods.
-} \label{fig:two-orbits}
-%\end{center}
-%\end{figure*}
-%
-%% Figure 5
-%\begin{figure*}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/lyap-orbits.pdf}
-%\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over 1000 orbital periods (the maximum orbital period) for Orbit A (left) and Orbit B (right).
-For Orbit A, the FTMLE asymptotes towards zero with a characteristic power-law slope.
-For Orbit B, the FTMLE begins to saturate to a constant; the Lyapunov exponent estimated for this orbit is $\lambda_{\rm max} \approx 0.07 \, {\rm Gyr}^{-1}$ corresponding to a Lyapunov time of $t_\lambda \approx 14 \, {Gyr}$.
-Over-plotted on both panels are lines (dashed, blue) that show $t^{-0.9}$ behavior.} \label{fig:lyap-orbits}
-%\end{center}
-%\end{figure*}
-%
-%% Figure 6
-%\begin{figure*}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/freq-diff-orbits.pdf}
-%\caption{ Difference in frequencies computed for successive sections of 100 orbital periods from the frequencies computed in the initial section for Orbit A (left) and Orbit B (right).
-The three lines in each panel are for the three fundamental frequencies.
-Dashed (red) horizontal line shows the estimate of the uncertainty for the frequencies computed for Orbit A.
-After many hundreds of orbital periods, the frequency deviation is not appreciable indicating again that the orbit is regular or only very weakly chaotic.
-The frequency deviations computed for Orbit B, however, show a clear increasing trend, indicating that the frequencies are diffusing with time.
-} \label{fig:fdiff-orbits} % The frequency diffusion rate defined in Equation~\ref{eq:fdrate} measures the initial slope of this trend.
-%\end{center}
-%\end{figure*}
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/lyap_map.png}
+\caption{  } \label{fig:lyapmap} 
+\end{center}
+\end{figure*}
 %
@@
 
+% Figure ??
+%\begin{figure*}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/log-freqmap.png}
+%\caption{Ratio of cartesian frequencies for a grid of iso-energy box orbits integrated in the potential of Equation~\ref{eq:logpotential}.
+Stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
+The regularity of the points in this map reflects the input grid of initial conditions.
+Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.} \label{fig:logfreqs}
+%\end{center}
+%\end{figure*}
+
 \end{document}

### 2015-05-16 ms.tex
table of numbers

@@
 
-% TODO: with 3 N-body simulations of regular, weakly chaotic, strongly chaotic, show that ensembles reproduce qualitative behavior
-
 The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant \emph{local} to a given orbit --- that is, these quantities are measures of how infinitesimal deviations will diverge, on average, from some parent orbit of interest.
@@
 
-We start by considering three particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular orbit, a weakly chaotic orbit, and a strongly chaotic orbit.
-The orbits were chosen at random from a single column of the orbit grid and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}.
-[...
-N-body simulations, expectation that only strongly chaotic orbit show strange morph, yadda]
+We start by considering three particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular orbit (R), a weakly chaotic orbit (W), and a strongly chaotic orbit (S).
+The orbits were chosen at random from a single column of the orbit grid based on their frequency diffusion time, and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}.
+We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of tidal stream formation from globular-cluster-mass progenitor systems on these three orbits.
+[TODO: mass scale 2.5E4 and mass contained within??] For each simulation, the progenitor consists of a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
+We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter.
+Figure~\ref{fig:nbodysims} shows the final particle distributions rotated so that the angular momentum of the progenitor orbit is aligned with the $z$-axis.
+A thin, cold stream forms on the regular orbit, a more diffuse stream on the mildly chaotic orbit, and a ``fanned'' stream on the strongly chaotic orbit.
+Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit, it is surprising that stream formed on this orbit appears to be more diffuse than the regular orbit stream.
 
@@
 \begin{center}
-	\begin{tabular}{c | c c c c c c c c }
-		ID & $x$ & $y$ & $z$ & $v_x$ & $v_y$ & $v_z$ & $t_\lambda$ & $t_\Omega$ \\\toprule
-		R & 10. & 0. & 0. & 0. & 0. & 0. & XX & YY\\
+	\begin{tabular}{c | c c c c c c | c c }
+		{\bf ID} & $\bs{x}$ & $\bs{y}$ & $\bs{z}$ & $\bs{v_x}$ & $\bs{v_y}$ & $\bs{v_z}$ & $\bs{t_\lambda}$ & $\bs{t_\Omega}$ \\\toprule
+		R & 27.30 & 0 & 1.10 & 0 & 183.61 & 0 & XX & YY\\
 		\midrule
-		W & 10. & 0. & 0. & 0. & 0. & 0. & XX & YY\\
+		W & 27.30 & 0 & 20.30 & 0 & 121.20 & 0 & XX & YY\\
 		\midrule
-		S & 10. & 0. & 0. & 0. & 0. & 0. & XX & YY\\
+		S & 27.30 & 0 & 21.90 & 0 & 111.22 & 0 & XX & YY\\
 		\bottomrule
 		\end{tabular}
-	\caption{TODO. \label{tbl:three-orbits}}
+	\caption{Orbits from the $xz$ grid with a range of chaotic timescales --- R is a regular orbit, W a weakly chaotic orbit, and S a strongly chaotic orbit.
+Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$, and times ($t_\lambda$, $t_\Omega$) in Gyr.
+\label{tbl:three-orbits}}
 \end{center}
@@
 
-% TODO: table with initial conditions
-
-It is computationally intractable to run full N-body simulations of stellar system disruption for large grids of orbital initial conditions and thus we take a simplified approach for analyzing these questions.
+It is computationally intractable to run full N-body simulations for the large grid of orbital initial conditions used for the maps of the previous section and we therefore take a simplified approach for studying how finite-volume debris spreads along each of these orbits.
 We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
@@
 where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the debris, and $\|\cdot \|$ is the Euclidean norm.
-For this work, we take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass to estimate the above debris scales.
+We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales.
+
+Over long times, a chaotic orbit will ergodically mix over the surface of its energy hypersurface (dimensionality 5 in a triaxial potential) whereas a (non-resonant) regular orbit will mix over the surface of its 3-torus.
+
+ We instead use ensembles with small spreads in energy so that we expect that most ensemble particle orbits around a parent chaotic orbit will reach a fully mixed state but now the final distribution will have some thickness around this surface.
+Around a typical regular orbit, each individual orbit will fill its torus ergodically --- a surface with lower dimension than the energy hypersurface --- and thus, a small enough ensemble should never reach the fully mixed state as defined above because it will remain confined to a subspace of lower dimensionality.
 
@@
 \end{figure*}
-%
-%% Figure 7
-%\begin{figure*}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/D_KL.pdf}
-%\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:DKL}
-%\end{center}
-%\end{figure*}
-%
-%% Figure 8
-%\begin{figure}[p]
-%\centering
-%	\includegraphics[width=\columnwidth] {figures/lyap_map_xz.png}
-%	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
-%	\label{fig:lyapmap}
-%\end{figure}
-%
-%% Figure 9
-%%\begin{figure}[!h]
-%%\centering
-%%	\includegraphics[width=0.4\columnwidth] {figures/kld-map.png}
-%%	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
-%%	\label{fig:kldmap}
-%%\end{figure}
 
-% Figure ??
-%\begin{figure*}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/log-freqmap.png}
-%\caption{Ratio of cartesian frequencies for a grid of iso-energy box orbits integrated in the potential of Equation~\ref{eq:logpotential}.
-Stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
-The regularity of the points in this map reflects the input grid of initial conditions.
-Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.} \label{fig:logfreqs}
-%\end{center}
-%\end{figure*}
+% Figure 4
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/fdiff_map.png}
+\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:freqdiff}
+\end{center}
+\end{figure*}
+
+% Figure 5
+\begin{figure*}[p]
+\begin{center}
+%\includegraphics[width=\textwidth]{figures/lyap_vs_fdiff.png}
+\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:lyapvfreqdiff}
+\end{center}
+\end{figure*}
+
+% Figure 6
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/nbody.png}
+\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:nbodysims}
+\end{center}
+\end{figure*}
 

### 2015-05-16 ms.tex
add parameter to freq diff time

@@
 The frequencies picked out by NAFF for such orbits will therefore represent the average periodic nature of the orbit over a given period.
-We define the frequency diffusion rate per orbit, $\fdrate$, as
+We define the fractional frequency diffusion rate per orbit, $\fdrate$, as
 \begin{equation}
-	\fdrate = \max_{\bs{a}} \, (\bs{\Omega}_{T_2} - \bs{\Omega}_{T_1}) / N_T \label{eq:fdrate}
+	\fdrate = \max_{\bs{a}} \, \left(\frac{\bs{\Omega}_{T_2} - \bs{\Omega}_{T_1}}{\bs{\Omega}_{T_1}}\right) \times \frac{1}{N_T}\label{eq:fdrate}
 \end{equation}
@@
 
-From this rate, we can compute a timescale over which we would expect order-unity changes to the fundamental frequencies of an orbit,
+From this rate, we can compute a timescale for the diffusion of the fundamental frequencies of an orbit.
+For a small ensemble of orbits, the relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
+We therefore define the frequency diffusion time with an unspecified parameter, $f$,
 \begin{equation}
-	t_\Omega = \frac{2 \pi}{\bs{\Omega}_{T_1}} \frac{1}{\fdrate}.
+	t_\Omega = \frac{2 \pi}{\bs{\Omega}_{T_1}} \frac{f}{\fdrate}.
 \end{equation}
+When $f = 1$, this is the timescale over which we expect order-unity changes to the frequencies, whereas $f = 0.01$ corresponds to $1\%$ changes.
 

### 2015-05-16 ms.tex
nld review and moved some to results 3

@@
 
-Over long times, a generic orbit in this potential fills the surface of its 3-torus --- that is, the three fundamental frequencies of the orbit are incommensurable.
+Over long times, a generic, \emph{conditionally-periodic} orbit in this potential fills the surface of its 3-torus --- that is, the three fundamental frequencies of the orbit are incommensurable.
 If there exists a relation of the form
@@
 The resonant structure of a potential --- the relative importance of particular resonance numbers --- is difficult to compute, but determines the global behavior of orbits in the potential.
-\cite{cincotta06} analytically compute the resonant structure in action-space for a St\"ackel potential.
 In frequency-space, the resonances are planes, but in action-space are generic 2D surfaces.
+% \cite{cincotta06} analytically compute the resonant structure in action-space for a St\"ackel potential.
 
-The orbit structure of non-integrable potentials can be qualitatively understood by considering a Hamiltonian that is a small perturbation away from being globally integrable --- that is, a Hamiltonian that may be written
+The orbit structure of non-integrable potentials can be understood by considering a Hamiltonian that is a small perturbation away from being globally integrable --- that is, a Hamiltonian that may be written
 \begin{equation}
@@
 \end{equation}
-where $\epsilon$ is a small parameter that determines the perturbation strength \citep[a complete description of perturbation theory applied to nonlinear Hamiltonians is given in][]{lichtenberg83}.
+where $H_0(\bs{J})$ represents an integrable Hamiltonian, and $\epsilon$ is a small parameter that determines the perturbation strength \citep[a complete description of perturbation theory applied to nonlinear Hamiltonians is given in][]{lichtenberg83}.
 When $0 < |\epsilon| \ll 1$, resonant surfaces become ``thick'' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99} and the stable regions are surrounded by chaotic layers where stochastic motion occurs.
@@
 The orbits that exist in these regions therefore can not be represented in angle-action coordinates, as their actions and frequencies evolve with time.
-
-Stochastic motion or the diffusion of an orbit in actions and frequencies may occur either \emph{across} the resonance layers or \emph{along} the stochastic layers (Arnold diffusion).
-For weakly chaotic orbits, motion across a resonance layer typically occurs with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
-Thus, if the resonance libration frequency is comparable to the orbital frequency, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
-However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can't change by large factors (unless there is resonance overlap and the motion is strongly chaotic).
-The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict, however, this has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}.
-For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; the implication of this connectivity is that an orbit may dramatically change its appearance as it navigates through the Arnold web, though the timescales for this to occur are typically many 1000s of orbital periods.
-
-When $\epsilon$ is large, there is no general prediction for the resulting behavior.
-It seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep{valluri, merritt, dehnen, etc.}.
-
-%For an integrable Hamiltonian, $H$, with $N$ degrees of freedom (dof), the equations of motion are most simply expressed in angle-action coordinates \cite[e.g.,][]{goldstein80}.
-In these coordinates, the actions, $J_n$, are (isolating) integrals of the motion, and the conjugate angle variables, $\theta_n$, increase linearly with time
-%\begin{align}
-%	\dot{\act}_n &= -\frac{\partial H}{\partial \theta_n} = 0\\
-%	\act_n &= {\rm const.}\\
-%	\dot{\theta}_n &= \frac{\partial H}{\partial \act_n} = \Omega_n(\act_1,\act_2,...,\act_n)\label{eq:aafreq}\\
-%	\theta_n(t) &= \Omega_n t + \theta_n(0)
-%\end{align}
-%where $i=1,2,...N$, and $\Omega_n$ are the fundamental frequencies of the orbit.
-The topology of angle-action space is equivalent to that of the surface of an $N$-dimensional torus, and thus orbits are often referred to in terms of \emph{orbital tori}: each set of actions, $(\act_1,\act_2,...\act_N)$, uniquely ``label'' the torus associated with a given orbit and the angles specify the location on the surface of the $N$-torus.
-If a global transformation to angle-action coordinates exists, the potential is said to be \emph{integrable}.
-%
-%Galactic gravitational potentials will almost certainly not be globally integrable.
-Dark matter halos of galaxies are expected to be triaxial but even if they are close to axisymmetric or spherical, the superposition of components (e.g., disk, bulge, halo) will break global integrability.
-If a triaxial (three dof) potential is close to integrable such that the Hamiltonian can be written in terms of a small perturbation, $\epsilon$, away from some integrable Hamiltonian, $H_0$,
-%\begin{equation}
-%	H(\bs{\theta}, \bs{\act}) = H_0(\bs{\act}) + \epsilon H_1(\bs{\theta}, \bs{\act})
-%\end{equation}
-%then a large number of regular orbits will survive the perturbation (by way of the KAM theorem)\footnote{3-vectors are represented with bold symbols, e.g., $\bs{J}=(J_1,J_2,J_3)$.}.
-The tori that survive will in general be separated by regions of irregular or chaotic motion, and thus any transformations to angle-action coordinates can only be defined locally.
-Analytic transformations to angle-action coordinates are only known for a handful of simple, integrable potentials.
-Much effort has gone in to developing general methods for transforming either from position-velocity to angle-action coordinates, $(\bs{x},\bs{v})\rightarrow(\bs{\theta},\bs{J})$ \citep{many}, or vice-versa, $(\bs{\theta},\bs{J})\rightarrow(\bs{x},\bs{v})$ \citep{many}.
-%
-%The types of motion generic to triaxial systems can be split into several sub-categories (summarized in Figure~\ref{fig:orbit-tree}).
-Regular motion can be classified by the number of resonance relations obeyed \citep[e.g.,][]{lichtenberg83}: \emph{conditionally periodic} orbits obey no resonance relations, \emph{uni-resonant} orbits obey a single relation of the form $\bs{m}\cdot\bs{\Omega}=0$, and \emph{bi-resonant} orbits obey two resonance relations, $\bs{m}\cdot\bs{\Omega}=0$ and $\bs{n}\cdot\bs{\Omega}=0$, where $\bs{m}$ and $\bs{n}$ are integer vectors.
-%
-%% Figure 2
-%%\begin{figure*}[!h]
-%%\begin{center}
-%%\includegraphics[width=0.6\textwidth]{figures/orbit-tree.pdf}
-%%\caption{A rough classification of the types of orbits found in triaxial potentials.
-Regular orbits are classified by the number of resonance relations obeyed: conditionally periodic (\emph{none}), uni-resonant (\emph{one}), bi-resonant (\emph{two}).
-Irregular or chaotic orbits are simply separated into two classes: strongly and weakly chaotic, defined by the timescales of interest in a problem.
-When the characteristic chaotic timescale is much longer than, for example, the age of the universe, the orbit is only weakly chaotic.
-} \label{fig:orbit-tree}
-%%\end{center}
-%%\end{figure*}
-%
-%Truly resonant orbits are rare \citep{merritt99}, but dynamically important in systems with several degrees of freedom as they determine the structure of orbit-space.
-Stable, resonant orbits generate families of orbits that may behave or appear similarly, whereas unstable resonances are typically associated with regions of chaos.
-Not all orbits can be trivially associated with resonances due to the infinitely complex, hierarchical structure of resonances (though the importance or ``size'' of a resonance falls off inversely with the norm of its integer vector, $|\bs{m}|^{-\alpha}$, with $\alpha > 0$).
-In autonomous, two dof systems, stochastic regions are generally surrounded and separated by regular motion, thus chaotic orbits are confined.
-In three or more dof the chaotic regions will connect and overlap.
-This is the origin of Arnold diffusion \citep{arnold64}, which allows chaotic orbits to radically change shape by traversing the web of unstable resonances and are only strictly bounded to their own energy hypersurface.
-This phenomenon is perhaps easier to visualize in action-space: in two dof systems, if a given set of tori are destroyed due to potential perturbations, an orbit in this now chaotic region is still bounded between the surrounding tori due to energy conservation.
-In three dimensions, the extra degree of freedom allows chaotic orbits to ``escape'' this confinement.
-%
-%For small potential perturbations, many regular orbits survive and only small chaotic islands are introduced.
+For small values of $\epsilon$, many regular orbits survive and only small chaotic islands are introduced.
 As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed.
 Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori --- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.
-However, even orbits that are strictly chaotic may behave regularly over timescales much longer than, for example, the dynamical time of the system.
-Chaotic orbits can therefore be loosely classified by being either \emph{strongly} or \emph{weakly} chaotic: strongly chaotic orbits behave irregularly over a few ($\sim$10) dynamical times, whereas weakly chaotic orbits behave regularly over this timescale.
-This rough classification is tied to local measures, such as the Lyapunov time (Section~\ref{sec:lyap}) or frequency diffusion time (Section~\ref{sec:naff}).
+
+When $\epsilon$ is large, there is no general prediction for the resulting behavior, however it seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep{valluri, merritt, dehnen, etc.}.
+We therefore expect a large number of regular orbits will survive --- the so-called KAM tori \citep{} --- however the tori that survive will be separated by regions of chaotic motion.
+Any transformations to angle-action coordinates must be defined locally due to the destruction of tori caused by the onset of chaotic motion.
 %
-%These local measures quantify the divergence of infinitesimally nearby orbits or the rate of frequency diffusion along an orbit.
-For an ensemble of orbits with finite spreads in energy or frequencies (e.g., tidal debris), each orbit will have a slightly different characteristic timescale, but we still expect mixing time of the ensemble to correlate with local measures of chaos of the ``parent'' orbit of the ensemble (Section~\ref{sec:ensemble}).
-The evolution of tidal debris will depend on the ensemble mixing rate around the parent orbit: chaotic mixing will cause the density of the debris to fall off more quickly than the typical power-law behavior expected for regular orbits \citep[e.g.,][]{merritt96, helmi99}.
-%
+%Analytic transformations to angle-action coordinates are only known for a handful of simple, integrable potentials.
+Much effort has gone in to developing general methods for transforming either from position-velocity to angle-action coordinates, $(\bs{x},\bs{v})\rightarrow(\bs{\theta},\bs{J})$ \citep{many}, or vice-versa, $(\bs{\theta},\bs{J})\rightarrow(\bs{x},\bs{v})$ \citep{many}.
 
@@
 This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies.
-Realistic Galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11, NiHao}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}.
+Realistic galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11, NiHao}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}.
 We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
@@
 Specifically, we use a \texttt{Python} wrapper over the \texttt{C} implementation by \cite{hairer93}.
-For all orbits we ensure that energy is conserved to $\Delta E/E_0 \leq 10^{-8}$ by the end of integration, however most orbits conserve energy to $\Delta E/E_0 \approx 10^{-13}$.
+For all orbits we ensure that energy is conserved to $|\Delta E/E_0| \leq 10^{-8}$ by the end of integration, however most orbits conserve energy to a part in $\approx$$10^{-13}$.
 Unless otherwise specified the integration timesteps are chosen so that there are 512 steps per strongest orbital period component, but the integrator uses adaptive stepping between each main step in order to satisfy a specified tolerance (we set the absolute tolerance to $\approx$100 times machine precision, $\texttt{atol} = 10^{-13}$).
@@
 We then compute the frequency diffusion rate (Equation~\ref{eq:fdrate}) using two consecutive integration windows with length $\approx$50 orbital periods.
-We find $\fdrate = 10^{-XX}$, which corresponds to a frequency diffusion time, $t_\Omega = XX~{\rm Gyr}$.
+We find $\fdrate = 10^{-XX}$, which corresponds to a $1\%$ frequency diffusion time, $t_\Omega = XX~{\rm Gyr}$ (the spread in frequencies for a $10^4~\msun$ globular cluster is $\approx$1\%).
 Though this orbit is computed in a different potential than will be used below, this demonstrates the apparent discrepancy between the timescales over which chaos is estimated to be relevant (XX's of Gyrs) and the integration time over which stream-fanning becomes apparent ($<8$ Gyr).
@@
 \subsection{Part III: Short-time frequency evolution}\label{sec:results3}
+
+Stochastic motion or the diffusion of an orbit in actions and frequencies may occur either \emph{across} the resonance layers or \emph{along} the stochastic layers (Arnold diffusion).
+For weakly chaotic orbits, motion across a resonance layer typically occurs with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
+Thus, if the resonance libration frequency is comparable to the orbital frequency, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
+However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can't change by large factors (unless there is resonance overlap and the motion is strongly chaotic).
+The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict, however, this has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}.
+For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; the implication of this connectivity is that an orbit may dramatically change its appearance as it navigates through the Arnold web, though the timescales for this to occur are typically many 1000s of orbital periods.
 

### 2015-05-18 ms.tex
short time evolution

@@
 
-The MLE, $\lambda_{\rm max}$, is interpreted as a rate that determines the nature of exponential divergence of infinitesimally close chaotic orbits.
+The MLE, $\lambda_{\rm max}$, is interpreted as a rate that sets the exponential divergence of infinitesimally close chaotic orbits.
 It is therefore useful to consider the corresponding $e$-folding time,
@@
 The strongest chaotic regions appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}), but the timescales [...].
-Figure~\ref{fig:lyapvfreqdiff} shows the Lyapunov time plotted against the frequency diffusion time [...].
 frequency diffusion rate is more accurate and thus more sensitive to mild chaos.
@@
 Figure~\ref{fig:nbodysims} shows the final particle distributions rotated so that the angular momentum of the progenitor orbit is aligned with the $z$-axis.
-A thin, cold stream forms on the regular orbit, a more diffuse stream on the mildly chaotic orbit, and a ``fanned'' stream on the strongly chaotic orbit.
+A thin stream forms on the regular orbit (left column), a more diffuse stream on the mildly chaotic orbit (middle column), and a ``fanned'' stream on the strongly chaotic orbit (right column).
 Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit, it is surprising that stream formed on this orbit appears to be more diffuse than the regular orbit stream.
@@
 We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
-For a given set of orbital initial conditions --- the ``parent'' orbit --- we find the position of the nearest pericenter, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for an integration time equal to XXX times the longest fundamental orbital period.
+For a given set of orbital initial conditions --- the ``parent'' orbit --- we find the position of the nearest pericenter, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for an integration time equal to 20 orbital periods.
 The physical scale of the ensemble is set by the tidal radius in position and velocity scale in velocity and are therefore set by the mass scale of the progenitor \citep[e.g.,][]{johnston98, apw14}.
@@
 
-Over long times, a chaotic orbit will ergodically mix over the surface of its energy hypersurface (dimensionality 5 in a triaxial potential) whereas a (non-resonant) regular orbit will mix over the surface of its 3-torus.
+Small ensembles of (conditionally-periodic) regular orbits will reach a fully phase-mixed state in a timescale  $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in frequency-space of the ensemble --- each individual orbit will fill the surface of its torus ergodically, but the ensemble will only mix due to sheering from slight variations in their fundamental frequencies \citep{merritt96, helmi99}.
+Ensembles of chaotic orbits behave quite differently.
+Over long times, a chaotic orbit in an autonomous Hamiltonian will ergodically mix over the energy hypersurface (a 5D surface, compared to the 3D stable tori).
+Generically, a small ensemble of chaotic orbits will lose their coherence much faster than for regular orbits, however computing the rate and amount of spreading is not a fully understood problem.
+Certain ensembles may be trapped between resonances for long times and thus remain in a state of pseudo-coherence for long times, whereas others may rapidly diffuse \citep[see, e.g.,][]{merritt96}.
+This phenomenon is generic to potentials that are not globally stochastic so that there are regions of stability and chaos intermingled throughout phase-space.
 
- We instead use ensembles with small spreads in energy so that we expect that most ensemble particle orbits around a parent chaotic orbit will reach a fully mixed state but now the final distribution will have some thickness around this surface.
-Around a typical regular orbit, each individual orbit will fill its torus ergodically --- a surface with lower dimension than the energy hypersurface --- and thus, a small enough ensemble should never reach the fully mixed state as defined above because it will remain confined to a subspace of lower dimensionality.
-
-[TODO: wait, actually, they differ in a log -- why doesn't this change things more? I guess this is saying like, mean(log(dens)) ~ log(mean(dens))??? confuzzled]
-As the particle ensembles evolve in time, we follow the mean configuration-space density of each to quantify the degree of mixing of the ensembles.
-This is similar to the method of \citet{merritt96} who measured the evolution of the relative entropy between the configuration-space density of isoenergy ensembles, $\rho(\bs{x},t)$, and the density of an ensemble that is uniformly distributed over the energy hypersurface, $\rho_{E_0}(\bs{x})$.
-In our case, the
-
- [but isoenergy so constant...yadda yadda].
-
-to quantify the degree of mixing of the ensemble by computing the ``distance'' between the configuration space density of the ensemble at a given time, $\rho(\bs{x},t)$, with that of a distribution that is uniformly distributed over the energy hypersurface, $\rho_{E_0}(\bs{x})$, where $E_0$ is the energy of the parent orbit and thus hypersurface of interest.
-Our method differs in a few ways from that of MV96.
-First, MV96 use isoenergy ensembles of particles for following the density because this is the only way to guarantee that, if chaotic, nearly all orbits in the ensemble will lie on chaotic orbits.
-We instead use ensembles with small but finite spreads in energy; we still expect that most ensemble particle orbits around a parent chaotic orbit will reach a fully mixed state but now the final distribution will have some thickness around this surface.
-Around a typical regular orbit, each individual orbit will fill its torus ergodically --- a surface with lower dimension than the energy hypersurface --- and thus, a small enough ensemble should never reach the fully mixed state as defined above because it will remain confined to a subspace of lower dimensionality.
-In practice, only strongly chaotic systems will get close to $\rho_{E_0}(\bs{x})$ in a reasonable amount of time; we expect the rate to depend on how stochastic the parent orbit is (e.g., as measured by the Lyapunov time or frequency diffusion rate) and also on how ``trapped'' the chaotic region is (a small chaotic region wedged between stable tori will take longer to appear chaotic, even if internally the frequency diffusion rate is large).
-However, the difference in configuration space density between an ensemble undergoing chaotic mixing versus an ensemble around a regular orbit may be detectable over a much shorter timescale.
-The fully-mixed, normalized, configuration-space density that fills a given energy hypersurface, $E_0$, is easily found by projecting the simple distribution function, $f$, into configuration space:
-\begin{align}
-	f(\bs{x}, \bs{v}) &= f_0 \, \delta(H(\bs{x}, \bs{v}) - E_0)\\
-	\rho_{E_0}(\bs{x}) &= \int \dd^3v \, f(\bs{x}, \bs{v})\\
-	&= \int \dd^3v \, f_0 \, \delta(H - E_0)\\
-	&= 4\pi \int \dd v \, v^2 \, f_0 \, \delta\left(v^2/2 + \Phi(\bs{x}) - E_0\right)\\
-	&= A \sqrt{E_0 - \Phi(\bs{x})}\label{eq:mixeddensity}
-\end{align}
-where $A$ is a constant determined such that $\int \dd^3\bs{x} \, \rho_{E_0}(\bs{x}) = 1$, and $\delta(\cdot)$ is the Dirac delta function.
-
-For each ensemble, we compute the Kullback-Leibler divergence \citep[KL divergence;][]{kullback51} between the configuration-space density and the fully-mixed density (Equation~\ref{eq:mixeddensity}) by estimating the configuration space density $\rho(\bs{x},t_j)$ using kernel density estimation (KDE)\footnote{We use an implementation from the Python package scikit-learn \citep{scikitlearn}.} with the ensemble of particle positions at a given time $t_j$, We use Monte Carlo integration to evaluate the KL divergence rather than evaluate the integral over a three dimensional grid.
-The KL divergence, $D_{\rm KL}$, is a measure of the relative (Shannon) entropy between two normalized density functions, $p(x)$ and $q(x)$,
-\begin{equation}
-	D_{\rm KL} = \int \dd x \, p(x) \ln \frac{p(x)}{q(x)}
-\end{equation}
-Instead, we are computing the three dimensional integral
-\begin{align}
-	D_{\rm KL}(t_j) &= \int \dd^3 \bs{x} \, \rho(\bs{x},t_j) \ln \frac{\rho(\bs{x},t_j)}{\rho_{E_0}(\bs{x})}
-\end{align}
-where we only have an estimate of the density $\rho_{\rm KDE}(\bs{x},t_j) \approx \rho(\bs{x},t_j)$. Using samples from this KDE estimated density, $\bs{x}_i$, the KL divergence is
-\begin{align}
-	D_{\rm KL}(t_j) &\approx \frac{1}{N} \sum_i^N \ln \frac{\rho_{\rm KDE}(\bs{x}_i,t_j)}{\rho_{E_0}(\bs{x}_i)}\\
-	\bs{x}_i &\sim \rho_{\rm KDE}(\bs{x},t_j)\label{eq:sampled_from}
-\end{align}
-where Equation~\ref{eq:sampled_from} indicates that the $\bs{x}_i$ are samples from the KDE density.
-For any computation of the density estimate, we use 10-fold cross-validation to find the optimal kernel bandwidth, $\sigma_{\rm KDE}$, by maximizing the likelihood of withheld particles.
-
-Figure~\ref{fig:DKL} shows $D_{\rm KL}$ as a function of time for the two orbits of Section~\ref{sec:comparison}.
-The chaotic orbit (Orbit B, dashed line) ensemble mixes much faster than the regular orbit (Orbit A, solid line) ensemble.
-The KLD is related to the logarithm of the mean density of the debris so that a difference of 1 in the KLD roughly corresponds to a factor of 10 difference in mean density of the debris.
-Figure XX shows particle positions at apocenter after 2, 16, and 32 orbital periods of evolution for ensembles of 1000 particles generated around Orbit A and B.
-For visualization, the final positions are rotated to lie on the $x$ axis and the angular momentum vector of the parent orbit at the given time is aligned with the $z$ axis.
-
-From the decay of the KLD, we measure a mixing rate by [TODO: how?].
-
-Our goal is to characterize the morphology of tidal debris along orbits generic to triaxial potentials.
-For example, it has been shown that along some chaotic orbits, tidal streams form large, diffuse ``fans'' of debris \citep[e.g.][]{fardal14, pearson15}, however, it is unknown how the observable properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
-Further, quantifying the degree of stochasticity with, for example, the frequency diffusion rate represents a \emph{local} measure of chaos and it is unclear how this generalizes to the small but finite spreads in orbital characteristics common to debris that is tidally disrupted from globular clusters or dwarf satellite galaxies.
-To answer this question would require a massive exploration of parameter space --- both in potential and in orbits --- and thus would require a prohibitive number of N-body simulations to truly analyze the disruption process in all situations.
-
-We measure the state of mixing of debris along the same grid of initial conditions as described above by computing the KL divergence (KLD) between the configuration-space density of the ensemble and the configuration-space density expected if the ensemble particles were uniformly distributed over the parent orbit energy hypersurface (Section~\ref{sec:ensemble}).
-From the time evolution of the KLD for a given ensemble around an orbit from the orbit grid, we estimate the mixing rate [TODO].
-Figure~\ref{fig:kldmap} shows the same grids of initial conditions as plotted in Figures~\ref{fig:lyapmap} and \ref{fig:freqmap}, but now the intensity of the pixels is set by the [TODO].
+We find that chaotic mixing of small orbit ensembles is detectable over short times, even when the predicted chaotic timescales (from the Lyapunov and frequency diffusion times) are long.
+For each ensemble of orbits, we compute the Figure~\ref{fig:ensemblemap} shows the grid of initial conditions described in Section~\ref{fig:results1}, but now the greyscale indicates the [...].
+After just 20 orbital periods, much of the structure apparent in the frequency diffusion time grid (Figure~\ref{fig:freqdiff}) is visible in this grid, even though [...].
 
@@
 
-Stochastic motion or the diffusion of an orbit in actions and frequencies may occur either \emph{across} the resonance layers or \emph{along} the stochastic layers (Arnold diffusion).
+Why does chaos manifest itself after just 20 orbital periods around orbits with predicted chaotic timescales equal to hundreds or thousands of orbital periods?
+To understand this, we first return to a review of nonlinear dynamics.
+Stochastic motion or chaotic diffusion of an orbit in actions and frequencies may occur either \emph{across} a resonance layers or \emph{along} the stochastic layers that surround the resonances in mildly stochastic potentials (Arnold diffusion).
 For weakly chaotic orbits, motion across a resonance layer typically occurs with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
-Thus, if the resonance libration frequency is comparable to the orbital frequency, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
+Thus, if the resonance libration frequency is comparable (within a factor of a few) to the orbital frequency, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
 However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can't change by large factors (unless there is resonance overlap and the motion is strongly chaotic).
 The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict, however, this has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}.
-For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; the implication of this connectivity is that an orbit may dramatically change its appearance as it navigates through the Arnold web, though the timescales for this to occur are typically many 1000s of orbital periods.
+For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; an orbit that ergodically mixes over its energy hypersurface must traverse this web, though the timescales typical for this phenomenon are many thousands of orbital periods.
 
 Both the Lyapunov exponent and the frequency diffusion rate measure mean, long-term rates of chaotic diffusion.
-As discussed in Section~\ref{sec:nldreview}, frequency diffusion may occur either across a resonance layer or along a stochastic layer.
-If a weakly chaotic orbit is confined (by other nearby, non-overlapping resonances or stable regions), the mean drift of an orbit in frequency space will be small if computed over timescales long compared to the orbital time but short compared to the Arnold diffusion time.
-For the potential considered in this paper, these are precisely the timescales for which the frequency diffusion rate of the previous section were computed ($\sim$100's of orbital periods).
-This window length was chosen in order to have a strong enough signal in the frequency spectrum of chaotic orbits to measure the most significant frequencies using NAFF.
+If a weakly chaotic orbit is confined (by other nearby, non-overlapping resonances or stable regions), the mean drift of an orbit in frequency space may be small if computed over timescales long compared to the orbital time but short compared to the Arnold diffusion time.
+For the potential considered in this paper, it turns out that these are precisely the timescales for which the frequency diffusion rate of the previous section were computed ($\approx$100's of orbital periods).
+This window length was chosen in order to have a strong enough signal in the frequency spectrum of chaotic orbits to measure the most significant frequencies using NAFF, however, it erases the small-scale variation of the frequencies.
 
-In this section, we instead wish to resolve the short-time behavior of the frequency diffusion (corresponding to motion across resonance layers).
-For each orbit, we now instead use a rolling window to analyze the evolution of the frequency spectrum on shorter times.
-We use a window with a width of 50 orbital periods and shift the window by half an orbital period between each calculation of the most significant frequencies.
-[Why this window width?
-Stride?]
+As a demonstration, we consider the three orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:three-orbits}).
+To resolve the short-time behavior of the frequency diffusion (corresponding to motion across resonance layers) we compute the frequencies in a series of rolling windows for ensembles of  orbits centered on the three parent orbits.
+We create the ensembles with the same procedure as described in Section~\ref{sec:results2} but only sample 100 orbits.
+We use a window with a width equal to 50 orbital periods and shift the window by half an orbital period between each calculation of the most significant frequencies (for a smaller window size, NAFF fails to recover the frequencies for the strongly chaotic orbit).
+Figure~\ref{fig:three-orbits-freqs} shows plots of the two strongest frequencies: the frequencies for each ensemble orbit computed in the first window are shown as green circles, whereas the frequencies in the final window are shown as red squares.
+The frequencies computed for all intermediate windows are shown as black points.
+The trajectory...
+chaotic evolution of frequencies...blah..
+TODO
 
-%\begin{figure}[!h]
-%\centering
-%	\includegraphics[width=0.4\columnwidth] {figures/kld-map.png}
-%	\caption{Not the final figures. Need to re-make with same axis limits, remove labels, put in same PDF file...etc.} 
-%	\label{fig:kldmap}
-%\end{figure}
+Chaotic ensembles quickly expand to fill allowed volume, then ...
+
+Figure~\ref{fig:freqvar_map} ...
 
@@
 \acknowledgements
-The authors wish to thank ... Robyn Sanderson, David Hogg, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
+The authors wish to thank ... Robyn Sanderson, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 
@@
 For box orbits, the motion is close to separable in each Cartesian component and we therefore use Cartesian coordinates for estimating the frequencies for these orbits.
+[TODO: Demo figure of log potential???
+convince out implementation is good?]
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/lyap-orbits.pdf}
+\includegraphics[width=\textwidth]{figures/pal5-lyap.png}
 \caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over YYYY orbital periods for the stream-fanning Palomar 5 orbit of \cite{pearson15}.
 The FTMLE saturates to the Lyapunov exponent in the shaded region and is estimated to be $\lambda_{\rm max} \approx XXX \, {\rm Gyr}^{-1}$, corresponding to a Lyapunov time of $t_\lambda \approx XX \, {Gyr}$.
-Over-plotted on both panels are lines (dashed, blue) that show $t^{-0.9}$ behavior.} \label{fig:pal5}
+} \label{fig:pal5}
 \end{center}
@@
 \begin{center}
-%\includegraphics[width=\textwidth]{figures/lyap_vs_fdiff.png}
-\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:lyapvfreqdiff}
+\includegraphics[width=\textwidth]{figures/nbody.png}
+\caption{Particle positions from the final snapshots of N-body simulations of a globular-cluster-mass satellite disrupting in a triaxial NFW potential (Section~\ref{sec:potential}).
+Each satellite is evolved for 20 orbital periods, and begins and ends at apocenter.
+For visualization, the positions have been aligned so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
+Top row shows the positions projected onto the resulting $xy$ plane, bottom row shows the positions projected onto the $yz$ plane.
+From left to right, the columns indicate that the progenitor is on a regular, mildly chaotic, or strongly chaotic orbit.
+Even on the mildly chaotic orbit --- where the predicted chaotic timescale is XX Gyr or XX orbital periods --- the resulting tidal stream is more diffuse than the regular case.
+} \label{fig:nbodysims}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/nbody.png}
-\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:nbodysims}
+\includegraphics[width=\textwidth]{figures/ensembles.png}
+\caption{TODO: these look like they have been evolved on slightly different orbits or for slightly different amount of time? } \label{fig:ensembles}
+\end{center}
+\end{figure*}
+
+% Figure 7
+\clearpage
+\begin{figure*}[p]
+\begin{center}
+%\includegraphics[width=\textwidth]{figures/ensemble_map.png}
+\caption{ } \label{fig:ensemblemap}
+\end{center}
+\end{figure*}
+
+% Figure 8
+\clearpage
+\begin{figure*}[p]
+\begin{center}
+%\includegraphics[width=\textwidth]{figures/}
+\caption{ } \label{fig:three-orbits-freqs}
+\end{center}
+\end{figure*}
+
+% Figure 9
+\clearpage
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/freqvar.png}
+\caption{ } \label{fig:freqvar_map}
 \end{center}

### 2015-05-28 ms.tex
writing writing writing

@@
 \newcommand{\inverse}[1]{{#1}^{-1}}
+\newcommand{\mean}[1]{\left< #1 \right>}
 \newcommand{\msun}{\ensuremath{\mathrm{M}_\odot}}
@@
 \newcommand{\inttime}{t_{\rm int}}
-\newcommand{\fdrate}{\mathcal{R}_\Omega}
+%\newcommand{\fdrate}{\mathcal{R}_\Omega}
 
@@
 The dark matter halos of galaxies are thought to be triaxial in shape.
-Despite suggestive evidence from a range of complimentary observational methods, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively verified.
+However, despite suggestive evidence from a range of complimentary observational methods, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively verified.
 Around other galaxies, it is generally hard to measure the 3D mass profile because they are seen in projection.
 From the Earth's position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii.
-The Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely on non-random orbits and thus contain extra information (e.g., debris stripped from in-falling satellite galaxies).
+The Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely on non-random orbits and thus contain extra information (e.g., debris stripped from stellar systems).
 
@@
 The phase-space density and therefore the observable properties of these tidal streams are sensitive to the internal progenitor properties, orbit of the progenitor, and mass distribution of the host galaxy.
-Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies (cite many).
-Around the Milky Way, streams are found over a large range of distances ($\approx$10-100~kpc).
-The morphological evolution of the debris depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also set by the shape and radial profile of the gravitational potential of the host galaxy.
+Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies that span a large range of distances \citep[$\approx$10--100~kpc][]{cite many}.
+The morphological evolution of tidal debris depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also determined by the shape and radial profile of the gravitational potential of the host galaxy.
 By simultaneously modeling the observed phase-space density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution around the Milky Way.
@@
 All methods have been tested in some way on simulated observations of data and these tests typically demonstrate the recovery of parameters for analytic, static potential forms.
-Many of the methods have been shown to work in axisymmetric or otherwise simple potentials but claim to extend to more complex, generalized potentials (e.g., triaxial or multi-component potentials).
+Many of the methods have been shown to work in axisymmetric or otherwise simple potentials but may extend to more complex, generalized potentials (e.g., triaxial or multi-component potentials).
 
-One example of stream modeling in a multi-component (static, analytic) potential was done by \citet{pearson15}, who tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} models \citep{kuepper12}.
-They use the observed SDSS number density of stars and a limited number of radial velocities measured for stream members to fit model streams to the data.
+One example of stream modeling in a multi-component (static, analytic) potential was done by \citet{pearson15}, who tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} \citep{kuepper12} and N-body models.
+They use the observed SDSS number density of stream stars, a limited number of radial velocities for stream members, and the sky position, distance, radial velocity, and proper motion of the cluster itself to fit model streams to the data.
 In the oblate potential (a three-component bulge+disk+spherical halo potential), a thin model stream is easily found that reproduces the observed stellar density morphology of the stream.
 In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness of the stream given the observational constraints of the present-day position and velocity of the cluster.
-% In order to extend tidal stream models to more generic potentials, it is important to understand why this stream-fanning occurs and whether this and other strange debris morphologies are generic to more complex potentials.
 
-The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the oblate potential.
-It is well known that the number of degrees of freedom of a potential plays a critical role in determining the orbit structure in the potential.
-Potentials (Hamiltonians) with more than two degrees of freedom generically contain significant chaotic regions.
-\citet{pearson15} tested the stochasticity of the orbit of the progenitor that produced fanned debris by computing the Lyapunov exponent but found that the orbit is consistent with being regular over dynamically relevant timescales (many Hubble times).
-It has been shown previously that along some strongly chaotic orbits, tidal streams form large, diffuse fans of debris \citep[e.g.,][]{fardal14}.
-However, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
-The result from \citet{pearson15} suggests that even a weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.
+The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the triaxial potential.
+It is well known that the number of degrees of freedom of a potential plays a critical role in determining the orbit structure of the potential --- Hamiltonians with more than two degrees of freedom generically contain significant chaotic regions.
+\citet{pearson15} tested the stochasticity of the orbit of the progenitor that produced fanned debris by computing the Lyapunov exponent along this orbit but found that it is consistent with being regular over dynamically relevant timescales (many Hubble times).
+It has been shown previously that along some strongly chaotic orbits, tidal streams do form large, diffuse fans of debris \citep[e.g.,][]{fardal14}, however it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
+The result from \citet{pearson15} suggests that even weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.
 
-TODO: this paragraph needs to be reworked...
-In this work, we analyze whether the final configuration-space morphology of simple tidal stream models is correlated with measures of chaos local to the orbit of the progenitor.
-The effect of chaos in the halos of galaxies has largely been neglected because the chaotic timescales estimated were thought to be many times the age of the universe, but recent work has shown that even weak chaos may be relevant for tidal stream morphology.
-In this Article, we choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos for grids of constant-energy orbits as computed from single-orbit diagnostics, and compare these results to measures of the mixing time of finite-volume ensembles of orbits (analogs to tidal debris).
-Finally, we connect the ensemble mixing rate to observational properties of streams, thus demonstrating that local orbital measures of chaos are predictors of tidal stream morphology.
-This provides a strong case for developing a method to use the mere existence of the many thin streams found in the halo of our own Galaxy to rule out classes or shapes of possible potentials that fit the Milky Way.
+In this work, we study the effect of chaotic diffusion on tidal stream morphology.
+We choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos as computed from single-orbit diagnostics for grids of constant-energy orbits, and compare these results to measures of the density evolution of finite-volume ensembles of orbits (meant to mimic tidal debris).
+We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles due to the nature of chaotic diffusion.
+This result supports a reevaluation of the importance of chaos in galactic halos and implies that the amount of chaos in a given potential may have significant consequences for the observability and survivability of thin, cold tidal streams.
 
-This paper is organized as follows: We review relevant nonlinear dynamics in Section~\ref{sec:nldreview}.
-In Section~\ref{sec:methods}, we describe our choice of potential and method for numerical orbit integration.
-We present our results in Section~\ref{sec:results} and discuss their implications in Section~\ref{sec:discussion}.
-We conclude in Section~\ref{sec:conclusions}.
-
-% , we describe the connection between different measures of chaotic behavior, further discuss the implications of chaotic mixing, and comment on the applicability of the methods presented here to more realistic Galactic potentials and simulations.
-
-%[Does this go anywhere?]
-%Thin stellar streams form when the spreading of the debris is faster in some preferred direction \citep[one eigenvalue of the Hessian of the potential is large relative to others e.g.,][]{helmi99, sanders??}.
-This is generically the case for orbits with significant angular momentum in a spherical potential ($|L| > J_r$, orbits that we expect for form streams), but [...
-comments about expectations in more complex potentials] In a spherical or mildly axisymmetric potential, the debris can only really spread in a single dimension because somehow the constraint that ....
-Along chaotic orbits, Hessian not defined and frequencies diffuse, ...
+This paper is organized as follows.
+We review relevant nonlinear dynamics in Section~\ref{sec:nldreview}.
+In Section~\ref{sec:methods}, we describe our choice of potential, method for numerical orbit integration, and introduce the chaos indicators used in this work.
+Our results are split into three subsections: in Section~\ref{sec:results1} we present iso-energy grids of orbits and discuss the orbit classes and chaotic timescales present in our potential; in Section~\ref{sec:results2} we study the density evolution of small ensembles of orbits around each orbit of the previous section; and, in Section~\ref{sec:results3} we describe the behavior of chaotic diffusion and use this to explain how chaos is relevant for tidal streams over short times.
+We discuss the implications of our results in Section~\ref{sec:discussion}, and conclude in Section~\ref{sec:conclusions}.
 
 \section{Review of nonlinear dynamics}\label{sec:nldreview}
-
-% TODO: move all except chaotic diffusion stuff to appendix?
 
@@
 \end{align}
-for all $t$. The motion in any component, $w_i(t)$, can be represented as a Fourier sum,
+for all $t$. If bounded, the motion in any component, $w_i(t)$, can be represented as a Fourier sum,
 \begin{equation}
-	w_i(t) = \sum_k^\infty a_k \, e^{i\,\omega_k\,t}; \, (a_k \in \mathbb{C}). \label{eq:fourier}
+	w_i(t) = \sum_k^\infty a_k \, e^{i\,\omega_k\,t} \label{eq:fourier}
 \end{equation}
+where the $a_k$ are complex amplitudes.
 
@@
 In these coordinates, the position variables are angles, $\boldsymbol{\theta}$, that increase linearly with time with rates set by $N$ constant, fundamental frequencies, $\boldsymbol{\Omega} = (\Omega_1, ..., \Omega_N)$.
-The frequencies in the Fourier sum of any individual component of motion (the $\omega_k$ in Equation~1) are just linear, integer combinations of the fundamental frequencies, $\boldsymbol{\Omega}$ --- that is, for a regular orbit, any Fourier component frequency may be written
+The frequency of a Fourier component in the sum for any individual component of motion (the $\omega_k$ in Equation~\ref{eq:fourier}) are just linear, integer combinations of the fundamental frequencies, $\boldsymbol{\Omega}$ --- that is, for a regular orbit, any Fourier component frequency may be written
 \begin{align}
-	\omega_k &= \boldsymbol{n_k} \cdot \boldsymbol{\Omega}; \, (\boldsymbol{n}_k \in \mathbb{Z}^3). \label{eq:fourierfreq}
-\end{align}
-The conjugate momentum coordinates --- the actions,  $\boldsymbol{J}$ --- are constants of motion. Even stronger, the actions are isolating integrals and any pair are in involution such that
+	\omega_k &= \boldsymbol{n_k} \cdot \boldsymbol{\Omega} \label{eq:fourierfreq}
+\end{align} %; \, (\boldsymbol{n}_k \in \mathbb{Z}^3)
+where $\boldsymbol{n}_k$ is a vector of $N$ integers.
+The conjugate momentum coordinates --- the actions,  $\boldsymbol{J}$ --- are constants of motion.
+Even stronger, the actions are isolating integrals and any pair are in involution such that
 \begin{equation}
@@
 This implies that for an $N$ dof system, a regular orbit has $N$ independent constants of motion and the motion is therefore restricted to an $N$-dimensional manifold embedded in the 2$N$ dimensional phase space.
-The conjugate angle variables define an orthogonal basis, and the topology of angle-action space is therefore the product of $N$ circles, an $N$-torus.
-Any regular orbit in an $N$ dof Hamiltonian can be understood as motion on the surface of an $N$-torus.
-Each set of actions, $(J_1,...,J_N)$, uniquely labels a torus, and regular orbits may be referred to in terms of their orbital tori.
-If a transformation to angle-action variables exists, the Hamiltonian may be written solely in terms of the actions,
-\begin{equation}
-	H = H(\boldsymbol{J})
-\end{equation}
+In fact, the topology of angle-action space is toroidal and any regular orbit in an $N$ dof Hamiltonian can be understood as motion on the surface of an $N$-torus.
+Each set of actions, $(J_1,...,J_N)$ (or frequencies), uniquely labels a torus, and regular orbits are sometimes referred to in terms of their orbital tori.
 
-A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally --- that is, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position.
+A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally --- for example, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position.
 The condition for global integrability is very restrictive and it seems that the likelihood that a Hamiltonian is globally integrable decreases as the number of degrees of freedom increase \citep[e.g.,][]{lichtenberg83}.
+In a globally integrable potential, the Hamiltonian may be written solely in terms of the actions, $H = H(\boldsymbol{J})$.
+Galactic potentials are almost certainly not globally integrable, but it is useful to understand the orbit structure in integrable systems before extending to more general potentials.
+Only one globally integrable triaxial mass model is known that generates motion that is qualitatively similar to that seen in Galaxies: the St\"ackel or ``Perfect Ellipsoid'' model \citep[e.g.,][]{kuzmin73, deZeeuw85}.
+The four general classes of orbits in this potential are box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits.
+The tube orbits preserve a sense of rotation about either the long or short axis and are therefore centrophobic, whereas the box orbits may approach arbitrarily close to the center of the potential and are centrophilic.
 
-Galactic potentials are almost certainly not globally integrable, but it is useful to understand the orbit structure in integrable systems before extending to more general potentials.
-Only one globally integrable triaxial mass model is known that generates motion that is qualitatively similar to that seen in Galaxies: the St\"ackel or ``Perfect Ellipsoid'' model \citep[e.g.,][]{kuzmin56, kuzmin73, deZeeuw85, cincotta06}.
-The four general classes of orbits in this potential are box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits.
-Unlike the box orbits, the tube orbits preserve a sense of rotation about either the long or short axis.
-
-Over long times, a generic, \emph{conditionally-periodic} orbit in this potential fills the surface of its 3-torus --- that is, the three fundamental frequencies of the orbit are incommensurable.
-If there exists a relation of the form
-\begin{equation}
-	\boldsymbol{n} \cdot \boldsymbol{\Omega} = n\, \Omega_1 + m\, \Omega_2 + l\, \Omega_3 = 0
-\end{equation}
-then the orbit is a resonant orbit and the effective dimensionality of the orbit is decreased to 2D.
-We refer to orbits that obey a single resonance relation as \emph{uni-resonant} orbits.
-If an additional resonance relation exists, the orbit is closed and becomes a 1D curve; we refer to such orbits as \emph{bi-resonant} orbits.
+The frequencies of a generic, non-resonant orbit are incommensurable --- that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} Conditionally-periodic orbits uniformly cover the surface of their orbital torus.
+If instead there exists a relation of the form $\boldsymbol{n} \cdot \boldsymbol{\Omega} = 0$ then the orbit is a resonant orbit.
+For resonant orbits, the existence of an additional \emph{local} constant of motion reduces the dimensionality of the orbit.
+We refer to orbits that obey a single resonance relation as \emph{uni-resonant} orbits, and if an additional resonance relation exists, \emph{bi-resonant}.
 The resonant structure of a potential --- the relative importance of particular resonance numbers --- is difficult to compute, but determines the global behavior of orbits in the potential.
-In frequency-space, the resonances are planes, but in action-space are generic 2D surfaces.
-% \cite{cincotta06} analytically compute the resonant structure in action-space for a St\"ackel potential.
+For a triaxial potential the resonances define planes in frequency-space, but resonance surfaces in action-space are generic 2D surfaces.
 
@@
 where $H_0(\bs{J})$ represents an integrable Hamiltonian, and $\epsilon$ is a small parameter that determines the perturbation strength \citep[a complete description of perturbation theory applied to nonlinear Hamiltonians is given in][]{lichtenberg83}.
-When $0 < |\epsilon| \ll 1$, resonant surfaces become ``thick'' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99} and the stable regions are surrounded by chaotic layers where stochastic motion occurs.
-Stochastic or chaotic orbits are characterized by apparently random behavior despite being generated by entirely deterministic systems.
-The stochastic layers form gaps in action space where tori are destroyed.
-The orbits that exist in these regions therefore can not be represented in angle-action coordinates, as their actions and frequencies evolve with time.
+When $0 < |\epsilon| \ll 1$, resonant surfaces become ``thick'' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99} and the stable regions are surrounded by stochastic layers where chaotic motion occurs.
+Chaotic orbits are characterized by exhibiting random behavior despite being generated by entirely deterministic equations of motion.
+Stochastic layers form gaps in action space where tori are destroyed; the orbits that exist in these regions therefore can not be represented in angle-action coordinates, as their actions and frequencies are not constant and diffuse chaotically with time.
 For small values of $\epsilon$, many regular orbits survive and only small chaotic islands are introduced.
@@
 When $\epsilon$ is large, there is no general prediction for the resulting behavior, however it seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep{valluri, merritt, dehnen, etc.}.
-We therefore expect a large number of regular orbits will survive --- the so-called KAM tori \citep{} --- however the tori that survive will be separated by regions of chaotic motion.
+We therefore expect a large number of regular orbits will survive --- the so-called KAM tori --- however the tori that survive will be separated by regions of chaotic motion.
 Any transformations to angle-action coordinates must be defined locally due to the destruction of tori caused by the onset of chaotic motion.
+[...
+how and where to end this section?]
 %
@@
 The density distributions within dark matter halos formed in cosmological N-body simulations are generically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}.
-With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo \citep[e.g.,][]{dubinski??, recent NiHao stuff}.
+With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo \citep[e.g.,][]{dubinski??, butsky15}.
 It is less clear what happens to the outer halo where tidal streams are more readily found.
@@
 
- \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the classic NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these halos.
-JS02 finds median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{??, NiHao} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
+ \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these halos.
+JS02 finds median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{millenium, hmm, butsky15} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
 JS02 finds significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).
@@
 This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies.
-Realistic galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11, NiHao}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}.
+Realistic galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors \citep{romanowsky98}, radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11,butsky15}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.; see, e.g.,][]{bailin05}.
 We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
@@
 
-\subsection{Numerical approximation of fundamental frequencies}\label{sec:naff}
+\subsection{Frequency diffusion rate}\label{sec:naff}
 
-Bounded, regular orbits in a triaxial potential will have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion.
+Bounded, regular orbits in a triaxial potential have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion.
 The motion in any canonical coordinate can therefore be decomposed as a Fourier sum (Equation~\ref{eq:fourier}) where the Fourier frequencies are linear, integer combinations of the fundamental frequencies (Equation~\ref{eq:fourierfreq}).
 \cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hamming-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
-It has been shown that using a filter makes the accuracy in determining the frequencies converge much faster than the typical $\inttime^{-1}$ expected for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time.
-Once the strongest frequencies of the time series are found, the spectrum must be searched for the three fundamental frequencies.
+It has been shown that using a filtered FFT makes the accuracy in determining the frequencies converge much faster than the typical $\inttime^{-1}$ expected for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time.
+The FFT is iteratively searched for the strongest frequency component, which is subtracted from the spectrum after each iteration.
+This procedure generates a table of frequencies for each component of motion which must then be searched for the three fundamental frequencies.
 
-This method is referred to as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) and has been used extensively in planetary dynamics \citep{} and galaxy dynamics \citep{valluri98}, especially in the study of orbits in triaxial systems.
+This method is referred to as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) and has been used extensively in planetary dynamics \citep[e.g.,][]{laskar93b, laskar96} and galaxy dynamics \citep{papaphilippou98, valluri98}, especially in the study of orbits in triaxial systems.
 We have implemented and tested our own open-source implementation of NAFF in \texttt{Python}.\footnote{This code is publicly available through the \texttt{Gary} package, \url{http://...}} Though our implementation differs slightly from the original definition and from that used in \cite{valluri98}, we still refer to the method as NAFF.
-For more details about NAFF, see Appendix~\ref{sec:naffapdx} or \cite{papaphilippou96, laskar, etc.}.
+For more details about NAFF, see Appendix~\ref{sec:naffapdx} or, e.g., \cite{laskar88, laskar93, papaphilippou96}.
 
-If an orbit is chaotic, the motion can no longer be expressed in terms of a single set of fundamental frequencies because the actions and frequencies of the orbit change with time.
+If an orbit is chaotic, the motion can no longer be expanded in terms of a single set of fundamental frequencies.
 For a weakly chaotic orbit, the orbit may appear consistently periodic over long windows of time.
-NAFF will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change character with time.
+NAFF will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change location and amplitude with time.
 For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two separate sections of orbit.
 The frequencies picked out by NAFF for such orbits will therefore represent the average periodic nature of the orbit over a given period.
-We define the fractional frequency diffusion rate per orbit, $\fdrate$, as
+We define the fractional frequency diffusion rate per orbital period, $\mathcal{R}$, in the $i$th fundamental frequency as
 \begin{equation}
-	\fdrate = \max_{\bs{a}} \, \left(\frac{\bs{\Omega}_{T_2} - \bs{\Omega}_{T_1}}{\bs{\Omega}_{T_1}}\right) \times \frac{1}{N_T}\label{eq:fdrate}
+	\mathcal{R}_i = \frac{\Omega_i^{(2)} - \Omega_{i}^{(1)}}{\Omega_{i}^{(1)} \, N} \label{eq:fdrate}
 \end{equation}
-where $\bs{\Omega}_{T_1}$ and $\bs{\Omega}_{T_2}$ are the fundamental frequencies computed from two consecutive sections of orbit, $N_T$ is the number of orbital periods within a single section of orbit computed from the most significant fundamental frequency, and the maximum is taken with respect to the corresponding amplitudes, $\bs{a}$, of the fundamental frequency components.
-The frequency diffusion rate as defined above (e.g., measured in the frequency with the largest amplitude) is better correlated with the strength of chaos in an orbit \citep{valluri??} but differs from the original definition of \cite{valluri98}.
-
-From this rate, we can compute a timescale for the diffusion of the fundamental frequencies of an orbit.
-For a small ensemble of orbits, the relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
-We therefore define the frequency diffusion time with an unspecified parameter, $f$,
+where $\Omega_i^{(1)}$ and $\Omega_i^{(2)}$ is the $i$th fundamental frequency computed in each of two consecutive sections of orbit and $N$ is the number of orbital periods, ${\rm T}$, in a single section of orbit.
+From this rate, we can compute a timescale; for a small ensemble of orbits (e.g., tidal debris), the relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
+We therefore define the frequency diffusion time (in units of number of orbital periods) with an unspecified parameter, $f$,
 \begin{equation}
-	t_\Omega = \frac{2 \pi}{\bs{\Omega}_{T_1}} \frac{f}{\fdrate}.
+	t_\Omega(f) = f \, (\max_{a_i} \, \mathcal{R}_i)^{-1} \label{eq:fdtime}
 \end{equation}
-When $f = 1$, this is the timescale over which we expect order-unity changes to the frequencies, whereas $f = 0.01$ corresponds to $1\%$ changes.
+where the maximum is taken with respect to the corresponding amplitudes, $a_i$, of the fundamental frequency components \citep[see][]{valluri12}.
+When $f = 1$, this is the number of orbital periods over which we expect order-unity changes to the frequencies, whereas $f = 0.01$ corresponds to $1\%$ changes.
 
@@
 In previous work, we found that tidal debris from a simulated globular cluster (Palomar 5) on an orbit with a long Lyapunov time can form large ``fans'' of debris similar to those seen for progenitor systems on strongly chaotic orbits \citep{pearson15, fardal14}.
-Here, to illustrate the methods described above, we repeat the calculation of the Lyapunov exponent for the Pal 5, stream-fanning orbit of \cite{pearson15} with a longer integration time and compute the frequency diffusion rate.
-From these, we measure the Lyapunov time and frequency diffusion time.
+Here, to illustrate the methods described above, we repeat the calculation of the Lyapunov exponent for the Pal 5, stream-fanning orbit of \cite{pearson15}.
+We use a longer integration time (equivalent to 10000 orbital periods) for the Lyapunov time calculation and additionally compute the frequency diffusion time.
 
-Figure~\ref{fig:pal5} shows the finite-time Lyapunov exponent computed for the Pal 5 orbit --- shaded region shows the ``converged'' part of the estimate that is used to compute the maximum Lyapunov exponent.
-By inverting the rate, we find $t_\lambda = XX~{\rm Gyr}$, consistent with our finding in \cite{pearson15}.
-We then compute the frequency diffusion rate (Equation~\ref{eq:fdrate}) using two consecutive integration windows with length $\approx$50 orbital periods.
-We find $\fdrate = 10^{-XX}$, which corresponds to a $1\%$ frequency diffusion time, $t_\Omega = XX~{\rm Gyr}$ (the spread in frequencies for a $10^4~\msun$ globular cluster is $\approx$1\%).
-Though this orbit is computed in a different potential than will be used below, this demonstrates the apparent discrepancy between the timescales over which chaos is estimated to be relevant (XX's of Gyrs) and the integration time over which stream-fanning becomes apparent ($<8$ Gyr).
-Below, we [...] try to understand by looking at grids of orbits...
+Figure~\ref{fig:pal5} shows the finite-time Lyapunov exponent computed for the Pal 5 orbit --- the shaded region shows the ``converged'' part of the estimate that is used to compute the maximum Lyapunov exponent.
+By inverting the rate, we find $t_\lambda \approx 12~{\rm Gyr} \approx 40\,{\rm T}$, shorter than our finding in \cite{pearson15} (from a shorter integration time) but still longer (by a factor of two) than the N-body simulations that display stream fanning.
+This discrepancy highlights the difficulty in reliably computing the Lyapunov exponent as it can take many thousands of orbital periods to converge.
+We compute the frequency diffusion rate (Equation~\ref{eq:fdrate}) using two consecutive integration windows with length $\approx$50 orbital periods.
+From this, we compute a frequency diffusion time $t_\Omega = 35~{\rm T}$ by assuming $f=0.005$ (the spread in frequencies in a $10^4~\msun$ globular cluster on this orbit is $\approx$0.5\%).
+Though this orbit is computed in a different potential than will be used below, this demonstrates the apparent discrepancy between the timescales over which chaos is estimated to be relevant and the integration time over which stream-fanning becomes apparent.
+% Below, we [...] try to understand by looking at grids of orbits...
 
@@
 Tube orbits are generally less stochastic than box orbits --- the other major class of orbits in triaxial potentials --- which tend to plunge deep into the inner regions of the potential, however there are appreciable regions of stochastic tube orbits.
-Thin tidal streams are thought to form along tube orbits rather than box orbits because of the fast disruption expected for stellar systems on radially plunging orbits.
+Thin tidal streams are thought to form along tube orbits rather than box orbits because of the faster disruption expected for stellar systems on radially plunging orbits.
 The major classes of stable tube orbits circulate about either the major or minor axis.
-The most prominent orbits in this grid are the short-axis and outer long-axis tubes, but this grid generates all of the major orbit classes: short-axis tubes, long-axis inner tubes, long-axis outer tubes, stochastic intermediate-axis, and box orbits.
+The most numerous orbits in this grid are the short-axis and outer long-axis tubes, but this grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits).
 
-To compute the Lyapunov exponents, we integrate all orbits in the grid for 1000 orbital periods and use the method described in Section~\ref{sec:lyap}.
-Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $xz$ plane) --- each pixel corresponds to an orbit, and the logarithm of the Lyapunov time (in units of Gyr) is mapped to greyscale intensity.
+To compute the Lyapunov exponents, we integrate all orbits in the grid for 1000 orbital periods [TODO: re-run for 5000 or 10000?] and use the method described in Section~\ref{sec:lyap}.
+Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $xz$ plane) --- each pixel corresponds to an orbit, and the logarithm of the Lyapunov time (in units of Gyr [TODO: convert to num.
+periods?]) is mapped to greyscale intensity.
 The darker pixels have shorter Lyapunov times and are more chaotic.
-By numerically estimating the MLE over this fixed integration window, the majority of orbits appear to be regular because they have exceedingly long Lyapunov times.
+By fixing the integration time (in units of orbital periods), the MLE cannot detect weak chaos and this the majority of orbits appear to be regular because they have exceedingly long Lyapunov times (all white points have $t_\lambda \gtrsim XX$).
 
@@
 The darker pixels have shorter frequency diffusion times and are more chaotic.
-This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential.
-This map highlights the accuracy of NAFF --- weak chaos is detected over $\approx$100 orbital periods, compared to the many thousands of orbits it would take to detect such features with the maximum Lyapunov exponent.
+This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential and highlights the accuracy of NAFF --- weak chaos is detectable over just 10's of orbital periods, compared to the many thousands of orbits it would take to detect such features with the maximum Lyapunov exponent.
 The tube orbits in this potential are mostly regular or only mildly chaotic --- the largest regular regions are associated with the short-axis and long-axis tube orbits --- however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.
 
-The strongest chaotic regions appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}), but the timescales [...].
-frequency diffusion rate is more accurate and thus more sensitive to mild chaos.
-It is unclear whether these regions will be dynamically important on realistic timescales, but we have confirmed from two independent methods that there are regions of strongly chaotic orbits even in this simple, triaxial potential.
-
-The frequency diffusion time grid reveals the intricate resonant structure of the potential considered, but the majority of the orbits have predicted chaotic timescales that correspond to many hundreds or thousands of orbital times.
-In the next section, we compute the mixing rates for finite-volume ensembles of orbits and compare the density evolution of the ensembles to the predictions from the single-orbit chaos diagnostics of this section.
+The strongest chaotic regions appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}).
+The resonant structure is revealed in the frequency diffusion map from integrations of just 100 orbital periods, though the majority of the orbits have estimated chaotic timescales much longer than 100 orbital periods.
+In the next section, we analyze the density evolution of finite-volume ensembles of orbits around each orbit in the above grids and compare the density evolution of the ensembles to the single-orbit chaos indicators computed in this section.
 
@@
 
-The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant \emph{local} to a given orbit --- that is, these quantities are measures of how infinitesimal deviations will diverge, on average, from some parent orbit of interest.
+The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant \emph{local} to a given orbit --- that is, these quantities are measures of how infinitesimal deviations will diverge on average from some parent orbit of interest, or of how long it takes for the frequencies to change by some amount.
 Tidal debris is disrupted from progenitor systems with finite spreads in orbital properties (e.g., energy).
 For a disrupting, globular-cluster-scale progenitor, the typical energy dispersion of the debris is 0.1--1\% of the progenitor orbital energy, but for a dwarf-galaxy-scale progenitor, the dispersion can be 5--20\%.
-This raises an important consideration: do finite deviations diverge faster than predicted from local measures of chaos?
-Does the Lyapunov or frequency diffusion time predict the timescale over which a finite phase-space volume (e.g., tidal debris) stays ``coherent''?
+Does the Lyapunov or frequency diffusion time predict the timescale over which a finite phase-space volume (e.g., tidal debris) stays coherent?
 
@@
 The orbits were chosen at random from a single column of the orbit grid based on their frequency diffusion time, and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}.
-We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of tidal stream formation from globular-cluster-mass progenitor systems on these three orbits.
-[TODO: mass scale 2.5E4 and mass contained within??] For each simulation, the progenitor consists of a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
+We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on these three orbits.
+The initial mass of the clusters is [TODO: 2.5E4 contained within??].
+For each simulation, the progenitor is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
 We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter.
@@
 		{\bf ID} & $\bs{x}$ & $\bs{y}$ & $\bs{z}$ & $\bs{v_x}$ & $\bs{v_y}$ & $\bs{v_z}$ & $\bs{t_\lambda}$ & $\bs{t_\Omega}$ \\\toprule
-		R & 27.30 & 0 & 1.10 & 0 & 183.61 & 0 & XX & YY\\
+		R & 24.90 & 0 & 6.90 & 0 & 190.76 & 0 & XX & YY\\
 		\midrule
-		W & 27.30 & 0 & 20.30 & 0 & 121.20 & 0 & XX & YY\\
+		W & 24.90 & 0 & 19.70 & 0 & 137.19 & 0 & XX & YY\\
 		\midrule
-		S & 27.30 & 0 & 21.90 & 0 & 111.22 & 0 & XX & YY\\
+		S & 24.90 & 0 & 26.10 & 0 & 93.11 & 0 & XX & YY\\
 		\bottomrule
@@
 	\caption{Orbits from the $xz$ grid with a range of chaotic timescales --- R is a regular orbit, W a weakly chaotic orbit, and S a strongly chaotic orbit.
-Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$, and times ($t_\lambda$, $t_\Omega$) in Gyr.
+Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$, and times ($t_\lambda$, $t_\Omega$) in number of orbital periods.
 \label{tbl:three-orbits}}
@@
 
-It is computationally intractable to run full N-body simulations for the large grid of orbital initial conditions used for the maps of the previous section and we therefore take a simplified approach for studying how finite-volume debris spreads along each of these orbits.
+It is computationally intractable to run full N-body simulations for the large grid of orbital initial conditions of the previous section and we therefore take a simplified approach for studying how finite-volume debris spreads along each of these orbits.
 We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
-For a given set of orbital initial conditions --- the ``parent'' orbit --- we find the position of the nearest pericenter, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for an integration time equal to 20 orbital periods.
-The physical scale of the ensemble is set by the tidal radius in position and velocity scale in velocity and are therefore set by the mass scale of the progenitor \citep[e.g.,][]{johnston98, apw14}.
+For a given set of orbital initial conditions --- the ``parent'' orbit --- we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time.
+The physical scale of the ensemble is set by the tidal radius in position and the velocity scale in velocity and are therefore set by the mass of the progenitor \citep[e.g.,][]{johnston98, apw14}.
 If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
@@
 We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales.
+Figure~\ref{fig:ensembles} shows such ensembles initialized around the three orbits described above and evolved for 16 orbital periods (same as the N-body simulations); the morphology of the ensembles is visually similar to the ``oldest'' (first stripped) debris in the N-body simulations.
 
@@
 Over long times, a chaotic orbit in an autonomous Hamiltonian will ergodically mix over the energy hypersurface (a 5D surface, compared to the 3D stable tori).
-Generically, a small ensemble of chaotic orbits will lose their coherence much faster than for regular orbits, however computing the rate and amount of spreading is not a fully understood problem.
+Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits, however analytically computing the rate and amount of spreading is very difficult even for simple potentials \cite[see, e.g.,][]{chirikov??}.
 Certain ensembles may be trapped between resonances for long times and thus remain in a state of pseudo-coherence for long times, whereas others may rapidly diffuse \citep[see, e.g.,][]{merritt96}.
-This phenomenon is generic to potentials that are not globally stochastic so that there are regions of stability and chaos intermingled throughout phase-space.
+This phenomenon is generic to potentials that are not globally stochastic in wich regions of stability and chaos are intermingled throughout phase-space.
 
-We find that chaotic mixing of small orbit ensembles is detectable over short times, even when the predicted chaotic timescales (from the Lyapunov and frequency diffusion times) are long.
-For each ensemble of orbits, we compute the Figure~\ref{fig:ensemblemap} shows the grid of initial conditions described in Section~\ref{fig:results1}, but now the greyscale indicates the [...].
-After just 20 orbital periods, much of the structure apparent in the frequency diffusion time grid (Figure~\ref{fig:freqdiff}) is visible in this grid, even though [...].
+We find that chaotic mixing of small orbit ensembles is detectable over short times, even when the predicted chaotic timescales (from the Lyapunov and frequency diffusion times) are long and therefore that the mean, single-orbit chaos indicators are not well-suited for determining the importance of chaotic diffusion on tidal stream evolution.
+For example, Figure~\ref{fig:ensemble-density} shows the evolution of the configuration-space density for ensembles evolved around the three orbits (R, W, S) described above.
+At each time step, we use kernel density estimation (KDE)\footnote{We use an Epanechnikov kernel with a bandwidth set to 1 kpc.
+We use an implementation from the Python package \texttt{scikit-learn} \citep{scikitlearn}.} with the ensemble of particle positions to estimate the configuration-space density field.
+We evaluate the density at the positions of each particle, $\rho_i$, and compare to the mean initial density, $\mean{\rho_0}$.
+Figure~\ref{fig:ensemble-density} shows histograms of the logarithm of $\rho_i/\mean{\rho_0}$ at three different times: after evolution for 4 (left), 8 (middle), and 16 (right) orbital periods.
+Both the mean density and shape of the density distribution changes over time and the nature of the evolution depends on the strength of chaos along the orbit.
+The regular orbit distribution stays relatively compact indicating that the density stays fairly uniformly high throughout the stream.
+For the weakly chaotic orbit, already after 8 orbits the density distribution shows a large tail to smaller densities --- this is ``fanned'' debris --- but maintains a significant number of particles in relatively high densities.
+The strongly chaotic orbit evolves much faster to low density.
+Stars with separations much larger than the bandwidth have unresolved density measurements --- this is why there appears to be a pile-up of stars at $\log_{10}\left[\rho_i / \mean{\rho_0}\right] \approx -3$.
+TODO: do we need to adjust bandwidth as the density evolves?
+
+The distributions all start with a peak at high density and a tail to lower density but evolve to the inverse state with a peak at low density and a tail to higher densities --- the skewness of a distribution is sensitive to exactly this property.
+Figure~\ref{fig:density-skewness} shows the mean density (top row) and skewness (bottom row) computed at 256 evenly-spaced intervals from the initial ensemble distribution to the distribution after 16 orbital periods.
+Both quantities (mean density and skewness) are correlated with the strength of chaos surrounding the parent orbit.
+
+Motivated by the apparent detection of chaos using these ensemble measures even around the weakly chaotic orbit, we compute the final skewness and mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
+For all ensembles, we integrate the orbits for 16 (parent) orbital periods and store the mean density and skewness at the end of integration for each.
+Figures~\ref{fig:ensemblemap-meandensity} and \ref{fig:ensemblemap-skewness} again show the grid of initial conditions (Section~\ref{sec:results1}), but now the greyscale indicates the final mean density or skewness, respectively.
+[TODO: some statement about the structure here, compare to freq.
+diff time map -- even though freq.
+diff time are long, still resolve structure in these maps.]
 
@@
 
-Why does chaos manifest itself after just 20 orbital periods around orbits with predicted chaotic timescales equal to hundreds or thousands of orbital periods?
-To understand this, we first return to a review of nonlinear dynamics.
-Stochastic motion or chaotic diffusion of an orbit in actions and frequencies may occur either \emph{across} a resonance layers or \emph{along} the stochastic layers that surround the resonances in mildly stochastic potentials (Arnold diffusion).
-For weakly chaotic orbits, motion across a resonance layer typically occurs with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
+Why does chaos manifest itself after just 16 orbital periods around orbits with predicted chaotic timescales equal to hundreds or thousands of orbital periods?
+To understand this, we first return to the concepts of nonlinear dynamics.
+In near-integrable potentials, chaotic diffusion of an orbit in actions and frequencies occurs both across resonance layers (a sort of stochastic libration) and along the stochastic layers that surround the resonances (Arnold diffusion).
+For weakly chaotic orbits, motion across a resonance layer can occur with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
 Thus, if the resonance libration frequency is comparable (within a factor of a few) to the orbital frequency, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
@@
 If a weakly chaotic orbit is confined (by other nearby, non-overlapping resonances or stable regions), the mean drift of an orbit in frequency space may be small if computed over timescales long compared to the orbital time but short compared to the Arnold diffusion time.
-For the potential considered in this paper, it turns out that these are precisely the timescales for which the frequency diffusion rate of the previous section were computed ($\approx$100's of orbital periods).
+For the potential considered in this paper, it turns out that these are comparable to the timescales for which the frequency diffusion rate of the previous section were computed ($\approx$100-1000 orbital periods).
 This window length was chosen in order to have a strong enough signal in the frequency spectrum of chaotic orbits to measure the most significant frequencies using NAFF, however, it erases the small-scale variation of the frequencies.
 
-As a demonstration, we consider the three orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:three-orbits}).
-To resolve the short-time behavior of the frequency diffusion (corresponding to motion across resonance layers) we compute the frequencies in a series of rolling windows for ensembles of  orbits centered on the three parent orbits.
-We create the ensembles with the same procedure as described in Section~\ref{sec:results2} but only sample 100 orbits.
-We use a window with a width equal to 50 orbital periods and shift the window by half an orbital period between each calculation of the most significant frequencies (for a smaller window size, NAFF fails to recover the frequencies for the strongly chaotic orbit).
-Figure~\ref{fig:three-orbits-freqs} shows plots of the two strongest frequencies: the frequencies for each ensemble orbit computed in the first window are shown as green circles, whereas the frequencies in the final window are shown as red squares.
-The frequencies computed for all intermediate windows are shown as black points.
-The trajectory...
-chaotic evolution of frequencies...blah..
-TODO
+As a demonstration of the small-scale frequency modulation, we consider the three orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:three-orbits}).
+To resolve the short-time behavior of the frequency diffusion (corresponding to motion across resonance layers) we compute the frequencies in a series of rolling windows along each orbit.
+We use a window with a width equal to 35 orbital periods and shift the window by half an orbital period between each calculation of the most significant frequencies --- this is the smallest window size we could choose before NAFF begins to consistently fail to recover the frequencies for the strongly chaotic orbit.
+Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time --- plotted are the percent deviations of the frequency ratios from the median value over all windows.
+The difference between the first window (green +) and the last window (red x) is 50 orbital periods.
+The frequencies of the weakly chaotic orbit (panel labeled $W$) evolve quickly but are bounded to a small volume with fractional size $\approx$0.3\% (presumably by nearby resonant surfaces).
+This is comparable to the frequency spread in globular cluster tidal debris (0.1--0.5\%).
 
-Chaotic ensembles quickly expand to fill allowed volume, then ...
+We see now where the discrepancy between chaotic timescales and the observable effects of chaos arise in tidal debris: even if the large-scale diffusion of frequencies is slow, an orbit may explore a small region of frequency space over much shorter times.
+The frequency diffusion time (Equation~\ref{eq:fdtime}) is an estimate of the time over which the mean value of the frequencies evolves, however the \emph{variance} of the frequencies over short times is dynamically relevant for small ensembles.
+This small-scale variability is insignificant for the evolution of global structure in, for example, an elliptical galaxy, but is signifiant for the morphological evolution of tidal debris with small spreads in frequencies.
 
-Figure~\ref{fig:freqvar_map} ...
+We now repeat the above analysis of the short-time evolution of frequencies, but now for ensembles of orbits around each of the three parent orbits.
+[TODO: a new metric, measuring the frequency variance, Figure~\ref{fig:freqvar_map} ...]
 
@@
 
-We have shown that local measures of chaos [...]
-
-Since all known tidal streams have been discovered as kinematic overdensities of stars, this implies (1) [something about accretion history], (2) the long, thin tidal streams we see may represent a sampling from ``special'' orbits in the Galactic potential where the mixing time is long, and (3) previous assumptions that connect the length and size of streams to their evolution time may be invalid.
+Since all known tidal streams have been discovered as kinematic over-densities of stars, this implies (1) [something about accretion history], (2) the long, thin tidal streams we see may represent a sampling from ``special'' orbits in the Galactic potential where the mixing time is long, and (3) previous assumptions that connect the length and size of streams to their evolution time may be invalid.
 
@@
 \includegraphics[width=\textwidth]{figures/pal5-lyap.png}
-\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over YYYY orbital periods for the stream-fanning Palomar 5 orbit of \cite{pearson15}.
+\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over XXX orbital periods for the stream-fanning Palomar 5 orbit of \cite{pearson15}.
 The FTMLE saturates to the Lyapunov exponent in the shaded region and is estimated to be $\lambda_{\rm max} \approx XXX \, {\rm Gyr}^{-1}$, corresponding to a Lyapunov time of $t_\lambda \approx XX \, {Gyr}$.
-} \label{fig:pal5}
+[TODO: add labels, remake with better font, remove blue line, add axis labels, add shaded region, check on all numbers for paragraph] } \label{fig:pal5}
 \end{center}
@@
 \includegraphics[width=\textwidth]{figures/lyap_map.png}
-\caption{  } \label{fig:lyapmap} 
+\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
+Each pixel in this image represents a single orbit, and the greyscale corresponds to the logarithm of the Lyapunov time (in units of orbital periods).
+Chaotic orbits with $t_\lambda \gtrsim XX~{\rm T}$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
+[TODO: remove Gyr and express in terms of orbital periods, add title, run Lyapunov grid for 5000 orbital periods instead of 1000?]} \label{fig:lyapmap}
 \end{center}
@@
 \includegraphics[width=\textwidth]{figures/fdiff_map.png}
-\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:freqdiff}
+\caption{The same grid of initial conditions as Figure~\ref{fig:lyapmap}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
+The frequencies are computed in two consecutive windows, each of which has length equal to 50 orbital periods.
+The frequencies are measured precisely so that small changes in the frequencies can be detected over just $\approx$10s of orbits.
+[TODO: remove Gyr and express in terms of orbital periods, add title] } \label{fig:freqdiff}
 \end{center}
@@
 \caption{Particle positions from the final snapshots of N-body simulations of a globular-cluster-mass satellite disrupting in a triaxial NFW potential (Section~\ref{sec:potential}).
-Each satellite is evolved for 20 orbital periods, and begins and ends at apocenter.
-For visualization, the positions have been aligned so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
-Top row shows the positions projected onto the resulting $xy$ plane, bottom row shows the positions projected onto the $yz$ plane.
-From left to right, the columns indicate that the progenitor is on a regular, mildly chaotic, or strongly chaotic orbit.
-Even on the mildly chaotic orbit --- where the predicted chaotic timescale is XX Gyr or XX orbital periods --- the resulting tidal stream is more diffuse than the regular case.
-} \label{fig:nbodysims}
+Each satellite is evolved for 16 orbital periods, and begins and ends at apocenter.
+For visualization, the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
+Top row shows the positions projected onto the resulting $xy$ plane (the instantaneous orbital plane of the progenitor at the end of the simulation), bottom row shows the positions projected onto the $yz$ plane.
+From left to right, the columns indicate that the progenitor is on a regular (R), weakly chaotic (W), or strongly chaotic (S) orbit.
+Even on the weakly chaotic orbit --- where the predicted chaotic timescale is XX orbital periods --- the resulting tidal stream displays stream-fanning and the ``oldest'' debris is more diffuse than in the regular case.
+[TODO: suptitle 'N-body simulations'?]} \label{fig:nbodysims}
 \end{center}
@@
 \includegraphics[width=\textwidth]{figures/ensembles.png}
-\caption{TODO: these look like they have been evolved on slightly different orbits or for slightly different amount of time? } \label{fig:ensembles}
+\caption{Final particle positions after integrating unbound, globular-cluster-sized ensembles of orbits generated around each of the three N-body orbits (Figure~\ref{fig:nbodysims}).
+The ensembles each contain 1000 particles and begin at pericenter rather than apocenter because they are meant to represent debris disrupted during a single (the first) pericentric passage.
+The particle positions qualitatively match the morphology of the ``oldest'' debris from the N-body simulations, though there are many more particles in these ensembles than are disrupted during the first pericentric passage of the simulations.
+[TODO: suptitle 'Orbit ensembles'?] } \label{fig:ensembles}
 \end{center}
@@
 
-% Figure 7
+% Figure ??
 \clearpage
@@
 \begin{center}
-%\includegraphics[width=\textwidth]{figures/ensemble_map.png}
-\caption{ } \label{fig:ensemblemap}
+\includegraphics[width=\textwidth]{figures/ensemble-density.png}
+\caption{ Distributions of the values of the estimated configuration-space density field evaluated at the positions of the ensemble particles after 4 (left), 8 (middle), and 16 (right) orbital periods.
+Colors represent the three orbits of Figures~\ref{fig:nbodysims} and \ref{fig:ensembles}: the regular, weakly chaotic, and strongly chaotic orbits.
+The density distribution of the regular orbit ensemble stays compact and decreases slowly.
+The density of the weakly chaotic ensemble initially evolves quickly to lower density, but then broadens.
+The strongly chaotic ensemble evolves fast to a state of low density.
+[TODO: can't distinguish colors when printed...]} \label{fig:ensemble-density}
 \end{center}
@@
 
-% Figure 8
+% Figure ??
 \clearpage
@@
 \begin{center}
-%\includegraphics[width=\textwidth]{figures/}
+\includegraphics[width=\textwidth]{figures/ensemble-density-skewness.png}
+\caption{ } \label{fig:density-skewness}
+\end{center}
+\end{figure*}
+
+% Figure ??
+\clearpage
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/ensemblemap-meandensity.png}
+\caption{ } \label{fig:ensemblemap-meandensity}
+\end{center}
+\end{figure*}
+
+% Figure ??
+\clearpage
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/ensemblemap-skewness.png}
+\caption{ } \label{fig:ensemblemap-skewness}
+\end{center}
+\end{figure*}
+
+% Figure ??
+\clearpage
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/freq-evolution.png}
 \caption{ } \label{fig:three-orbits-freqs}
@@
 
-% Figure 9
+% Figure ??
 \clearpage

### 2015-06-08 ms.tex
added table of parameters for potential

@@
 
-\title{Chaotic diffusion of tidal streams}
+\title{Chaotic mixing of tidal debris}
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
@@
 
+In order to explore the question of how chaos manifests over timescales much shorter than that predicted from generic chaos indicators, we must first understand the structure and behavior of orbits in gravitational potentials.
+
+\subsection{Orbits in integrable potentials}
+
 An orbit in an $N$ degree of freedom (dof) Hamiltonian, $H$, is a set of $2N$ quasi-periodic time series, 
@@
 The four general classes of orbits in this potential are box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits.
-The tube orbits preserve a sense of rotation about either the long or short axis and are therefore centrophobic, whereas the box orbits may approach arbitrarily close to the center of the potential and are centrophilic.
+The tube orbits preserve a sense of rotation about either the long or short axis and are therefore centrophobic, whereas the box orbits may approach arbitrarily close to the center of the potential (centrophilic) and may change the sense of rotation about any axis.
 
@@
 For a triaxial potential the resonances define planes in frequency-space, but resonance surfaces in action-space are generic 2D surfaces.
+
+\subsection{Orbits in non-integrable potentials}
+
+[TODO: take cartoons from talk and include here for description]
 
@@
 how and where to end this section?]
-%
-%Analytic transformations to angle-action coordinates are only known for a handful of simple, integrable potentials.
-Much effort has gone in to developing general methods for transforming either from position-velocity to angle-action coordinates, $(\bs{x},\bs{v})\rightarrow(\bs{\theta},\bs{J})$ \citep{many}, or vice-versa, $(\bs{\theta},\bs{J})\rightarrow(\bs{x},\bs{v})$ \citep{many}.
+
+\subsection{The behavior of chaotic orbits in non-integrable potentials}
+
+[??]
 
 \section{Numerical methods}\label{sec:methods}
+
+Our goal is to map the orbit structure of arbitrary potentials, with an emphasis on identifying the chaotic orbits and and understanding the evolution of these orbits over short times.
+In this section, we describe the methods we will use to detect and quantify the strength of chaos for large grids of orbits.
 
@@
 \begin{align}
-	\Phi(u,\phi,\theta) &\approx \frac{v_c^2}{A}\left[F_1(u) + \frac{1}{2}(e_b^2 + e_c^2)F_2(u) + \frac{1}{2} [(e_b\sin\theta \sin\phi)^2 + (e_c\cos\theta)^2] F_3(u) \right]\\
+	\Phi(u,\phi,\theta) &\approx \frac{v_c^2}{A}\left[F_1(u) + \frac{1}{2}(e_b^2 + e_c^2)F_2(u) + \frac{1}{2} [(e_b\sin\theta \sin\phi)^2 + (e_c\cos\theta)^2] F_3(u) \right]\label{eq:potential}\\
 	A &= \left(\ln2 - \frac{1}{2}\right) + \left(\ln2-\frac{3}{4}\right) (e_b^2 + e_c^2)
@@
 We chose $r_s=20~{\rm kpc}$ and $v_c = 175~{\rm km}~{\rm s}^{-1}$ by taking the mean halo concentration for a ${\rm M}_{vir} \approx 10^{12}~\msun$ halo, $c_e\approx5$, from \cite{jing02} and by assuming $R_{vir}\approx200~{\rm kpc}$.
-Figure~\ref{fig:potential} shows equipotential contours of this potential in projection.
+Figure~\ref{fig:potential} shows equipotential contours of this potential in projection, and Table~\ref{tbl:potential} summarizes the potential parameters.
+
+\begin{table*}[ht]
+\begin{center}
+	\begin{tabular}{ c  c }
+	         Parameter & Value\\\toprule
+		$v_c$ & 175~km~s$^{-1}$\\
+		$r_s$ & 20~kpc\\
+		$a$ & 1\\
+		$b$ & 0.77\\
+		$c$ & 0.55\\
+		\bottomrule
+		\end{tabular}
+	\caption{Summary of potential parameters for the triaxial NFW potential (Equation~\ref{eq:potential}) used in this work.
+Triaxiality is introduced in the density to ensure the density is physical at all radii.
+Velocity scale, scale radius, and axis ratios are chosen to match the median halo parameters for a $M_{\rm vir} \approx 10^{12}~\msun$ halo from \citep{jing02}.
+\label{tbl:potential}}
+\end{center}
+\end{table*}
 
@@
 
-As a demonstration of the small-scale frequency modulation, we consider the three orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:three-orbits}).
+As a demonstration of the small-scale frequency modulation, we consider again the three orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:three-orbits}).
+Figure~\ref{fig:fdiff-zooms} shows zoomed-reruns of the frequency diffusion time map of Figure~\ref{fig:freqdiff} centered on the three orbits --- the pixel positions are now shown relative to the three parent orbits.
+The frequency diffusion times shown in these grids were computed with two consecutive windows of length 35 orbital periods (shorter than the windows used in the previous, large grid of Figure~\ref{fig:freqdiff}).
+The resonant structure around the mildly chaotic orbit (white regions in panel W) bound the orbit to a small region, whereas for the strongly chaotic orbit (S), the entire region surrounding the orbit is strongly chaotic.
 To resolve the short-time behavior of the frequency diffusion (corresponding to motion across resonance layers) we compute the frequencies in a series of rolling windows along each orbit.
@@
 \begin{center}
+\includegraphics[width=1.1\textwidth]{figures/fdiff_zooms.png}
+\caption{ } \label{fig:fdiff-zooms}
+\end{center}
+\end{figure*}
+
+% Figure ??
+\clearpage
+\begin{figure*}[p]
+\begin{center}
 \includegraphics[width=\textwidth]{figures/freq-evolution.png}

### 2015-06-08 ms.tex
incorporating Kathryn's comments

@@
 
-\subsection{The behavior of chaotic orbits in non-integrable potentials}
+\subsection{The behavior of chaotic orbits in non-integrable potentials}\label{sec:behavior-chaotic}
 
@@
 
-In previous work, we found that tidal debris from a simulated globular cluster (Palomar 5) on an orbit with a long Lyapunov time can form large ``fans'' of debris similar to those seen for progenitor systems on strongly chaotic orbits \citep{pearson15, fardal14}.
-Here, to illustrate the methods described above, we repeat the calculation of the Lyapunov exponent for the Pal 5, stream-fanning orbit of \cite{pearson15}.
-We use a longer integration time (equivalent to 10000 orbital periods) for the Lyapunov time calculation and additionally compute the frequency diffusion time.
-
-Figure~\ref{fig:pal5} shows the finite-time Lyapunov exponent computed for the Pal 5 orbit --- the shaded region shows the ``converged'' part of the estimate that is used to compute the maximum Lyapunov exponent.
-By inverting the rate, we find $t_\lambda \approx 12~{\rm Gyr} \approx 40\,{\rm T}$, shorter than our finding in \cite{pearson15} (from a shorter integration time) but still longer (by a factor of two) than the N-body simulations that display stream fanning.
-This discrepancy highlights the difficulty in reliably computing the Lyapunov exponent as it can take many thousands of orbital periods to converge.
-We compute the frequency diffusion rate (Equation~\ref{eq:fdrate}) using two consecutive integration windows with length $\approx$50 orbital periods.
-From this, we compute a frequency diffusion time $t_\Omega = 35~{\rm T}$ by assuming $f=0.005$ (the spread in frequencies in a $10^4~\msun$ globular cluster on this orbit is $\approx$0.5\%).
-Though this orbit is computed in a different potential than will be used below, this demonstrates the apparent discrepancy between the timescales over which chaos is estimated to be relevant and the integration time over which stream-fanning becomes apparent.
-% Below, we [...] try to understand by looking at grids of orbits...
+In Section~\ref{sec:results1}, we generate grids of orbits in the potential described above to map the orbit structure of the potential.
+We classify each orbit in terms of the strength of chaos along the orbit as computed using the methods of Section~\ref{sec:methods}.
+With this initial classification, in Section~\ref{sec:results2} we follow the evolution of ensembles of orbits generated around each orbit in the initial grid.
+We find that the configuration-space density of ensembles around weakly chaotic orbits evolve faster (e.g., in mean density) than expected given the timescales over which chaos is computed to be relevant for the parent orbits.
+In Section~\ref{sec:results3}, we explain this phenomenon in the context of how chaotic diffusion occurs (e.g., Section~\ref{sec:sec:behavior-chaotic}).
 
@@
 
-% Figure 2
-\begin{figure*}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/pal5-lyap.png}
-\caption{ FTMLE (Equation~\ref{eq:ftmle}) computed over XXX orbital periods for the stream-fanning Palomar 5 orbit of \cite{pearson15}.
-The FTMLE saturates to the Lyapunov exponent in the shaded region and is estimated to be $\lambda_{\rm max} \approx XXX \, {\rm Gyr}^{-1}$, corresponding to a Lyapunov time of $t_\lambda \approx XX \, {Gyr}$.
-[TODO: add labels, remake with better font, remove blue line, add axis labels, add shaded region, check on all numbers for paragraph] } \label{fig:pal5}
-\end{center}
-\end{figure*}
-
-% Figure 3
+% Figure ??
 \begin{figure*}[p]
@@
 
-% Figure 4
+% Figure ??
 \begin{figure*}[p]
@@
 
-% Figure 5
+% Figure ??
 \begin{figure*}[p]
@@
 
-% Figure 6
+% Figure ??
 \begin{figure*}[p]

### 2015-06-09 ms.tex
very important: added todos

@@
 \newcommand{\inttime}{t_{\rm int}}
-%\newcommand{\fdrate}{\mathcal{R}_\Omega}
+
+\newcommand{\project}[1]{\textsl{#1}}
+\newcommand{\superfreq}{\project{SuperFreq}}
+
+% TO DO
+\newcommand{\todo}[2]{{\color{red} TODO: (\MakeUppercase{#1}) #2}}
 
@@
 
-[TODO: take cartoons from talk and include here for description]
+\todo{APW}{take cartoons from talk and include here for description}
 
@@
 
-We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy (per unit mass) chosen to span a range of distances comparable to the scale radius of the potential (Section~\ref{sec:potential}), $E=-(397.2~{\rm km~s}^{-1})^2$ (in physical units).
+We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy (per unit mass) chosen to span a range of distances comparable to the scale radius of the potential ($E=-(397.2~{\rm km~s}^{-1})^2$ in physical units; see Table~\ref{tbl:potential}).
 We fix $v_x = v_z = 0$, and compute $v_y$ from the energy.
@@
 The major classes of stable tube orbits circulate about either the major or minor axis.
-The most numerous orbits in this grid are the short-axis and outer long-axis tubes, but this grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits).
+This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits), but the most numerous orbits in this grid are the short-axis and outer long-axis tubes.
 
-To compute the Lyapunov exponents, we integrate all orbits in the grid for 1000 orbital periods [TODO: re-run for 5000 or 10000?] and use the method described in Section~\ref{sec:lyap}.
-Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $xz$ plane) --- each pixel corresponds to an orbit, and the logarithm of the Lyapunov time (in units of Gyr [TODO: convert to num.
-periods?]) is mapped to greyscale intensity.
+To compute the Lyapunov exponents, we integrate all orbits in the grid for 1000 orbital periods \todo{apw}{re-run for 5000 or 10000 periods?} and use the method described in Section~\ref{sec:lyap}.
+Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $xz$ plane) --- each pixel corresponds to an orbit, and the logarithm of the Lyapunov time (in units of Gyr \todo{apw}{ convert to num.
+periods?}) is mapped to greyscale intensity.
 The darker pixels have shorter Lyapunov times and are more chaotic.
-By fixing the integration time (in units of orbital periods), the MLE cannot detect weak chaos and this the majority of orbits appear to be regular because they have exceedingly long Lyapunov times (all white points have $t_\lambda \gtrsim XX$).
+By fixing the integration time (in units of orbital periods), the MLE cannot detect weak chaos and the majority of orbits appear to be regular because they have exceedingly long Lyapunov times (all white points have $t_\lambda \gtrsim XX$).
 
-For each orbit, we also separately integrate for $\approx$100 orbital periods and use NAFF to compute the fundamental frequencies for the two consecutive sections of 50 orbital periods.
+For each orbit, we also separately integrate for $\approx$80 orbital periods and use \superfreq\ to compute the fundamental frequencies for the two consecutive sections of 40 orbital periods.
+We have found that this is the shortest integration window for which we are able to consistently and accurately recover the frequencies using \superfreq.
+
  Figure~\ref{fig:freqdiff} shows the same grid of initial conditions, but now the greyscale intensity is set by the logarithm of the frequency diffusion time.
@@
 We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on these three orbits.
-The initial mass of the clusters is [TODO: 2.5E4 contained within??].
+The initial mass of the clusters is \todo{kvj}{2.5E4 \msun contained within what again?
+35 scale radii?
+but this is plummer??}.
 For each simulation, the progenitor is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
@@
 Stars with separations much larger than the bandwidth have unresolved density measurements --- this is why there appears to be a pile-up of stars at $\log_{10}\left[\rho_i / \mean{\rho_0}\right] \approx -3$.
-TODO: do we need to adjust bandwidth as the density evolves?
+\todo{apw}{do we need to adjust bandwidth as the density evolves?}
 
@@
 Figures~\ref{fig:ensemblemap-meandensity} and \ref{fig:ensemblemap-skewness} again show the grid of initial conditions (Section~\ref{sec:results1}), but now the greyscale indicates the final mean density or skewness, respectively.
-[TODO: some statement about the structure here, compare to freq.
+\todo{apw}{talk about the structure here, compare to freq.
 diff time map -- even though freq.
-diff time are long, still resolve structure in these maps.]
+diff time are long, still resolve structure in these maps.}
 
@@
 We now repeat the above analysis of the short-time evolution of frequencies, but now for ensembles of orbits around each of the three parent orbits.
-[TODO: a new metric, measuring the frequency variance, Figure~\ref{fig:freqvar_map} ...]
+\todo{apw}{a new metric, measuring the frequency variance, Figure~\ref{fig:freqvar_map} ...}
 
@@
 [TODO: Demo figure of log potential???
-convince out implementation is good?]
+convince our implementation is good?]
 
@@
 Chaotic orbits with $t_\lambda \gtrsim XX~{\rm T}$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
-[TODO: remove Gyr and express in terms of orbital periods, add title, run Lyapunov grid for 5000 orbital periods instead of 1000?]} \label{fig:lyapmap}
+\todo{apw}{remove Gyr and express in terms of orbital periods, add title, run Lyapunov grid for 5000 orbital periods instead of 1000?}} \label{fig:lyapmap}
 \end{center}
@@
 The frequencies are measured precisely so that small changes in the frequencies can be detected over just $\approx$10s of orbits.
-[TODO: remove Gyr and express in terms of orbital periods, add title] } \label{fig:freqdiff}
+\todo{apw}{remove Gyr and express in terms of orbital periods, add title} } \label{fig:freqdiff}
 \end{center}
@@
 Even on the weakly chaotic orbit --- where the predicted chaotic timescale is XX orbital periods --- the resulting tidal stream displays stream-fanning and the ``oldest'' debris is more diffuse than in the regular case.
-[TODO: suptitle 'N-body simulations'?]} \label{fig:nbodysims}
+\todo{apw}{suptitle 'N-body simulations'?}} \label{fig:nbodysims}
 \end{center}
@@
 The particle positions qualitatively match the morphology of the ``oldest'' debris from the N-body simulations, though there are many more particles in these ensembles than are disrupted during the first pericentric passage of the simulations.
-[TODO: suptitle 'Orbit ensembles'?] } \label{fig:ensembles}
+\todo{apw}{suptitle 'Orbit ensembles'?} } \label{fig:ensembles}
 \end{center}
@@
 The strongly chaotic ensemble evolves fast to a state of low density.
-[TODO: can't distinguish colors when printed...]} \label{fig:ensemble-density}
+\todo{apw}{can't distinguish colors when printed...}} \label{fig:ensemble-density}
 \end{center}

### 2015-06-09 ms.tex
change to superfreq

@@
 We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
+\todo{apw}{Cite this paper: hunter05 for disk crossings and chaos}
 
@@
 This method is referred to as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) and has been used extensively in planetary dynamics \citep[e.g.,][]{laskar93b, laskar96} and galaxy dynamics \citep{papaphilippou98, valluri98}, especially in the study of orbits in triaxial systems.
-We have implemented and tested our own open-source implementation of NAFF in \texttt{Python}.\footnote{This code is publicly available through the \texttt{Gary} package, \url{http://...}} Though our implementation differs slightly from the original definition and from that used in \cite{valluri98}, we still refer to the method as NAFF.
-For more details about NAFF, see Appendix~\ref{sec:naffapdx} or, e.g., \cite{laskar88, laskar93, papaphilippou96}.
+We have implemented and tested a version of this procedure in the \project{Python} programming language.
+Our implementation differs slightly from the original definition and from that used in \cite{valluri98}: we have found that using a higher order window function \citep[e.g.,][]{hunter02} of the form
+\begin{equation}
+	W(t/\inttime) = \frac{2^p \, (p!)^2}{(2p)!} \left( 1 + \cos \frac{\pi t}{\inttime}\right)^p
+\end{equation}
+with $p=2$ allows for more reliable determination of frequencies from strongly chaotic orbits.
+We refer to this slight modification of the algorithm as \superfreq\ and the code is open-source and publicly available on \project{GitHub} at \url{https://github.com/adrn/SuperFreq}.
+For more details about the algorithm, see Appendix~\ref{sec:naffapdx} or \cite{laskar88, laskar93, papaphilippou96}.
 
@@
 For a weakly chaotic orbit, the orbit may appear consistently periodic over long windows of time.
-NAFF will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change location and amplitude with time.
+\superfreq\ will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change location and amplitude with time.
 For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two separate sections of orbit.
-The frequencies picked out by NAFF for such orbits will therefore represent the average periodic nature of the orbit over a given period.
+The frequencies picked out by \superfreq\ for such orbits will therefore represent the average periodic nature of the orbit over a given period.
 We define the fractional frequency diffusion rate per orbital period, $\mathcal{R}$, in the $i$th fundamental frequency as
@@
 \end{equation}
-where $\Omega_i^{(1)}$ and $\Omega_i^{(2)}$ is the $i$th fundamental frequency computed in each of two consecutive sections of orbit and $N$ is the number of orbital periods, ${\rm T}$, in a single section of orbit.
-From this rate, we can compute a timescale; for a small ensemble of orbits (e.g., tidal debris), the relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
-We therefore define the frequency diffusion time (in units of number of orbital periods) with an unspecified parameter, $f$,
+where $\Omega_i^{(1)}$ and $\Omega_i^{(2)}$ is the $i$th fundamental frequency computed in each of two consecutive sections of orbit and $N$ is the number of orbital periods in a single section of orbit.
+By inverting this rate, we can compute the timescale over which we expect order-unity changes to the fundamental frequencies: the \emph{frequency diffusion time} is defined as
 \begin{equation}
-	t_\Omega(f) = f \, (\max_{a_i} \, \mathcal{R}_i)^{-1} \label{eq:fdtime}
+	t_\Omega = \, (\max_{a_i} \, \mathcal{R}_i)^{-1} \label{eq:fdtime}
 \end{equation}
-where the maximum is taken with respect to the corresponding amplitudes, $a_i$, of the fundamental frequency components \citep[see][]{valluri12}.
-When $f = 1$, this is the number of orbital periods over which we expect order-unity changes to the frequencies, whereas $f = 0.01$ corresponds to $1\%$ changes.
+where the maximum is taken with respect to the corresponding amplitudes, $a_i$, of the fundamental frequency components \citep[see][]{valluri12}, and the time is in units of number of orbital periods, ${\rm T}$.
+For a small ensemble of orbits (e.g., tidal debris), the relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
+We can estimate this timescale by multiplying the frequency diffusion time by a factor equal to the fractional spread in frequencies of the debris.
+For example, a globular cluster typically has $\approx$1\% spreads in fundamental frequencies, so by multiplying the frequency diffusion time by $f = 0.01$, we can estimate the time (number of orbital periods) over which we expect the frequencies to evolve by this amount.
 
@@
 We have found that this is the shortest integration window for which we are able to consistently and accurately recover the frequencies using \superfreq.
-
- Figure~\ref{fig:freqdiff} shows the same grid of initial conditions, but now the greyscale intensity is set by the logarithm of the frequency diffusion time.
+Figure~\ref{fig:freqdiff} shows the same grid of initial conditions, but now the greyscale intensity is set by the logarithm of the (order-unity) frequency diffusion time.
 The darker pixels have shorter frequency diffusion times and are more chaotic.
-This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential and highlights the accuracy of NAFF --- weak chaos is detectable over just 10's of orbital periods, compared to the many thousands of orbits it would take to detect such features with the maximum Lyapunov exponent.
+This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential and highlights the accuracy of \superfreq\ --- weak chaos (grey) is detectable over just 10's of orbital periods, compared to the many thousands of orbits it would take to detect such features with the maximum Lyapunov exponent.
 The tube orbits in this potential are mostly regular or only mildly chaotic --- the largest regular regions are associated with the short-axis and long-axis tube orbits --- however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.
@@
 The strongest chaotic regions appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}).
-The resonant structure is revealed in the frequency diffusion map from integrations of just 100 orbital periods, though the majority of the orbits have estimated chaotic timescales much longer than 100 orbital periods.
+The resonant structure is revealed in the frequency diffusion map from integrations of just 80 orbital periods, though the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods.
 In the next section, we analyze the density evolution of finite-volume ensembles of orbits around each orbit in the above grids and compare the density evolution of the ensembles to the single-orbit chaos indicators computed in this section.
@@
 Tidal debris is disrupted from progenitor systems with finite spreads in orbital properties (e.g., energy).
-For a disrupting, globular-cluster-scale progenitor, the typical energy dispersion of the debris is 0.1--1\% of the progenitor orbital energy, but for a dwarf-galaxy-scale progenitor, the dispersion can be 5--20\%.
+For a disrupting, globular-cluster-scale progenitor, the typical energy dispersion of the debris is 0.1--1\% of the progenitor orbital energy \citep[assuming masses of $10^4$--$10^5$~\msun;][]{johnston98}, but for a dwarf-galaxy-scale progenitor, the dispersion can be $\gtrsim$10\%.
 Does the Lyapunov or frequency diffusion time predict the timescale over which a finite phase-space volume (e.g., tidal debris) stays coherent?
@@
 For each simulation, the progenitor is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
-We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter.
+\todo{apw}{Label the ICs for the 3 orbits on Figures 3 and 4} We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter.
 Figure~\ref{fig:nbodysims} shows the final particle distributions rotated so that the angular momentum of the progenitor orbit is aligned with the $z$-axis.
@@
 For the potential considered in this paper, it turns out that these are comparable to the timescales for which the frequency diffusion rate of the previous section were computed ($\approx$100-1000 orbital periods).
-This window length was chosen in order to have a strong enough signal in the frequency spectrum of chaotic orbits to measure the most significant frequencies using NAFF, however, it erases the small-scale variation of the frequencies.
+This window length was chosen in order to have a strong enough signal in the frequency spectrum of chaotic orbits to measure the most significant frequencies using \superfreq\, however, it erases the small-scale variation of the frequencies.
 
@@
 To resolve the short-time behavior of the frequency diffusion (corresponding to motion across resonance layers) we compute the frequencies in a series of rolling windows along each orbit.
-We use a window with a width equal to 35 orbital periods and shift the window by half an orbital period between each calculation of the most significant frequencies --- this is the smallest window size we could choose before NAFF begins to consistently fail to recover the frequencies for the strongly chaotic orbit.
+We use a window with a width equal to 35 orbital periods and shift the window by half an orbital period between each calculation of the most significant frequencies --- this is the smallest window size we could choose before \superfreq\ begins to consistently fail to recover the frequencies for the strongly chaotic orbit.
 Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time --- plotted are the percent deviations of the frequency ratios from the median value over all windows.

### 2015-06-10 ms.tex
adding text

@@
 \renewcommand{\dateseparator}{-}
+\usepackage{lscape}
+\usepackage{rotating}
 
@@
 
-In order to explore the question of how chaos manifests over timescales much shorter than that predicted from generic chaos indicators, we must first understand the structure and behavior of orbits in gravitational potentials.
-
-\subsection{Orbits in integrable potentials}
-
+In order to explore the question of how chaos manifests over timescales much shorter than that predicted from generic chaos indicators, we must first understand the structure and behavior of orbits in complex gravitational potentials.
 An orbit in an $N$ degree of freedom (dof) Hamiltonian, $H$, is a set of $2N$ quasi-periodic time series,
@@
 This implies that for an $N$ dof system, a regular orbit has $N$ independent constants of motion and the motion is therefore restricted to an $N$-dimensional manifold embedded in the 2$N$ dimensional phase space.
-In fact, the topology of angle-action space is toroidal and any regular orbit in an $N$ dof Hamiltonian can be understood as motion on the surface of an $N$-torus.
+The topology of angle-action space is toroidal and any regular orbit in an $N$ dof Hamiltonian can be understood as motion on the surface of an $N$-torus.
 Each set of actions, $(J_1,...,J_N)$ (or frequencies), uniquely labels a torus, and regular orbits are sometimes referred to in terms of their orbital tori.
+
+\subsection{Orbits in integrable potentials}
 
@@
 
-The frequencies of a generic, non-resonant orbit are incommensurable --- that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} Conditionally-periodic orbits uniformly cover the surface of their orbital torus.
-If instead there exists a relation of the form $\boldsymbol{n} \cdot \boldsymbol{\Omega} = 0$ then the orbit is a resonant orbit.
+The frequencies of a generic, non-resonant orbit are incommensurable --- that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} These conditionally-periodic orbits uniformly cover the surface of an orbital torus.
+If instead there exists a relation of the form $\boldsymbol{n} \cdot \boldsymbol{\Omega} = 0$ the orbit is a resonant orbit.
 For resonant orbits, the existence of an additional \emph{local} constant of motion reduces the dimensionality of the orbit.
@@
 For a triaxial potential the resonances define planes in frequency-space, but resonance surfaces in action-space are generic 2D surfaces.
+In plots of frequency ratios, the resonances appear as lines; Figure~\ref{fig:cartoons} (left panel) shows a cartoon portrait of a portion of frequency-space for an integrable potential with resonance lines, non-resonant, and resonant orbits marked.
+For an integrable potential, all non-resonant orbits are regular.
 
-\subsection{Orbits in non-integrable potentials}
+\subsection{Orbits in near- and non-integrable potentials}
 
-\todo{APW}{take cartoons from talk and include here for description}
-
-The orbit structure of non-integrable potentials can be understood by considering a Hamiltonian that is a small perturbation away from being globally integrable --- that is, a Hamiltonian that may be written
+The orbit structure of near-integrable potentials can be understood by considering a Hamiltonian that is a small perturbation away from being globally integrable --- that is, a Hamiltonian that may be written
 \begin{equation}
@@
 where $H_0(\bs{J})$ represents an integrable Hamiltonian, and $\epsilon$ is a small parameter that determines the perturbation strength \citep[a complete description of perturbation theory applied to nonlinear Hamiltonians is given in][]{lichtenberg83}.
-When $0 < |\epsilon| \ll 1$, resonant surfaces become ``thick'' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99} and the stable regions are surrounded by stochastic layers where chaotic motion occurs.
+When $0 < |\epsilon| \ll 1$, resonant surfaces become ``thick'' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99}, and these layers regions are then generically surrounded by stochastic layers where chaotic motion occurs.
 Chaotic orbits are characterized by exhibiting random behavior despite being generated by entirely deterministic equations of motion.
 Stochastic layers form gaps in action space where tori are destroyed; the orbits that exist in these regions therefore can not be represented in angle-action coordinates, as their actions and frequencies are not constant and diffuse chaotically with time.
-For small values of $\epsilon$, many regular orbits survive and only small chaotic islands are introduced.
+
+Figure~\ref{fig:cartoons} (middle panel) shows a cartoon of frequency space for a near-integrable potential (a small perturbation away from the potential in the left panel).
+Much of the structure that was present in the integrable potential remains in the near-integrable case, but the differences are highlighted.
+Orbits in the resonant layers surrounding the resonances (grey) are near-resonant orbits that will librate around the resonance.
+Chaotic orbits in the stochastic layer (red) behave erratically depending on the surrounding resonance structure.
+If the stochastic layer is small and the chaotic orbit is therefore confined in some way, it may behave nearly regular for long periods of time.
+
+For small values of $\epsilon$, many regular orbits survive and only small chaotic regions are introduced, especially where resonances overlap.
 As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed.
-Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori --- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.
+Figure~\ref{fig:cartoons} (right panel) qualitatively shows this phenomenon --- the resonant and resonant-layer orbits may still be regular, but many or all of the conditionally-periodic orbits can be chaotic.
+As the perturbation strength increases, next the uni-resonant tori are destroyed, and finally the bi-resonant tori --- these are least susceptible to destruction from perturbations \cite[for a more quantitative illustration of this transition from integrability to global chaos, see Figure~9 in][]{valluri98}.
 
@@
 Any transformations to angle-action coordinates must be defined locally due to the destruction of tori caused by the onset of chaotic motion.
-[...
-how and where to end this section?]
 
@@
 
-[??]
+Chaotic orbits ...
+
+Small ensembles of (conditionally-periodic) regular orbits will reach a fully phase-mixed state in a timescale  $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in frequency-space of the ensemble --- each individual orbit will fill the surface of its torus ergodically, but the ensemble will only mix due to sheering from slight variations in their fundamental frequencies \citep{merritt96, helmi99}.
+Ensembles of chaotic orbits behave quite differently.
+Over long times, a chaotic orbit in an autonomous Hamiltonian will ergodically mix over the energy hypersurface (a 5D surface, compared to the 3D stable tori).
+Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits, however analytically computing the rate and amount of spreading is very difficult even for simple potentials \cite[see, e.g.,][]{chirikov??}.
+Certain ensembles may be trapped between resonances for long times and thus remain in a state of pseudo-coherence for long times, whereas others may rapidly diffuse \citep[see, e.g.,][]{merritt96}.
+This phenomenon is generic to potentials that are not globally stochastic in which regions of stability and chaos are intermingled throughout phase-space.
+
+\todo{apw}{concluding paragraph...}
 
@@
 
-Small ensembles of (conditionally-periodic) regular orbits will reach a fully phase-mixed state in a timescale  $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in frequency-space of the ensemble --- each individual orbit will fill the surface of its torus ergodically, but the ensemble will only mix due to sheering from slight variations in their fundamental frequencies \citep{merritt96, helmi99}.
-Ensembles of chaotic orbits behave quite differently.
-Over long times, a chaotic orbit in an autonomous Hamiltonian will ergodically mix over the energy hypersurface (a 5D surface, compared to the 3D stable tori).
-Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits, however analytically computing the rate and amount of spreading is very difficult even for simple potentials \cite[see, e.g.,][]{chirikov??}.
-Certain ensembles may be trapped between resonances for long times and thus remain in a state of pseudo-coherence for long times, whereas others may rapidly diffuse \citep[see, e.g.,][]{merritt96}.
-This phenomenon is generic to potentials that are not globally stochastic in wich regions of stability and chaos are intermingled throughout phase-space.
-
 We find that chaotic mixing of small orbit ensembles is detectable over short times, even when the predicted chaotic timescales (from the Lyapunov and frequency diffusion times) are long and therefore that the mean, single-orbit chaos indicators are not well-suited for determining the importance of chaotic diffusion on tidal stream evolution.
@@
 
-Since all known tidal streams have been discovered as kinematic over-densities of stars, this implies (1) [something about accretion history], (2) the long, thin tidal streams we see may represent a sampling from ``special'' orbits in the Galactic potential where the mixing time is long, and (3) previous assumptions that connect the length and size of streams to their evolution time may be invalid.
+\subsection{Existence of streams}
+
+We have considered here a simple triaxial gravitational potential chosen to mimic the median properties of dark matter halos formed in dark-matter-only simulations (or the large-scale properties of halos formed in simulations with baryonic effects).
+We have found that
+
+\subsection{Cautions}
+
+Since all known tidal streams have been discovered as spatial over-densities of stars, this implies (1) [something about accretion history], (2) the long, thin tidal streams we see may represent a sampling from ``special'' orbits in the Galactic potential where the mixing time is long, and (3) previous assumptions that connect the length and size of streams to their evolution time may be invalid.
 
@@
 
-\subsection{Potential}
-
+[discussion of potential choice?]
 Why not disk + bulge? What do we expect when we add?
@@
 
-\section{Conclusions}\label{sec:conclusions}
+\section{Summary and Conclusions}\label{sec:conclusions}
 We have shown that ...
@@
 
-% Figure 1
-% TODO: should I show density? cool 3D render of density?
+% Figure ??
+\begin{sidewaysfigure}[!p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/cartoons.pdf}
+\caption{} 
+\label{fig:cartoons}
+\end{center}
+\end{sidewaysfigure}
+
+% Figure ??
 \begin{figure*}[!p]
@@
 \includegraphics[width=\textwidth]{figures/potential.pdf}
-\caption{Equipotential contours for the triaxial NFW potential considered in this work. There are eight contour levels evenly spaced and linear in the value of the potential. } \label{fig:potential}
+\caption{Equipotential contours for the triaxial NFW potential considered in this work.
+There are eight contour levels evenly spaced and linear in the value of the potential.
+\todo{apw}{should I show density as well or instead?}} \label{fig:potential}
 \end{center}

### 2015-06-11 ms.tex
moving words around

@@
 
-Chaotic orbits ...
+Chaotic orbits have no orbital actions because they are only strictly bound to their energy hypersurface (if the potential is time-independent).
+The orbits therefore do not have a single set of fundamental frequencies, but rather the frequencies that describe the character of motion evolve with time.
+In near-integrable potentials, chaotic diffusion of an orbit in actions and frequencies occurs both across resonance layers (a sort of stochastic libration) and along the stochastic layers that surround the resonances (Arnold diffusion).
+For weakly chaotic orbits, motion across a resonance layer can occur with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
+Thus, if the resonance libration frequency is comparable (within a factor of a few) to the orbital frequency, motion across a resonance can modulate the frequency spectrum of an orbit over dynamically relevant timescaes.
+However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can't change by large factors (unless there is resonance overlap and the motion is strongly chaotic).
+The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict, however, this has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}.
+For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; an orbit that ergodically mixes over its energy hypersurface must traverse this web, though the timescales typical for this phenomenon are many thousands of orbital periods.
 
@@
 Why does chaos manifest itself after just 16 orbital periods around orbits with predicted chaotic timescales equal to hundreds or thousands of orbital periods?
-To understand this, we first return to the concepts of nonlinear dynamics.
-In near-integrable potentials, chaotic diffusion of an orbit in actions and frequencies occurs both across resonance layers (a sort of stochastic libration) and along the stochastic layers that surround the resonances (Arnold diffusion).
-For weakly chaotic orbits, motion across a resonance layer can occur with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
-Thus, if the resonance libration frequency is comparable (within a factor of a few) to the orbital frequency, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
-However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can't change by large factors (unless there is resonance overlap and the motion is strongly chaotic).
-The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict, however, this has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}.
-For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; an orbit that ergodically mixes over its energy hypersurface must traverse this web, though the timescales typical for this phenomenon are many thousands of orbital periods.
-
 Both the Lyapunov exponent and the frequency diffusion rate measure mean, long-term rates of chaotic diffusion.

### 2015-06-17 ms.tex
words yo

--- 2015-06-10 ms.tex
@@
 \renewcommand{\dateseparator}{-}
-\usepackage{lscape}
-\usepackage{rotating}
+%\usepackage{lscape}
+%\usepackage{rotating}
 
@@
 The four general classes of orbits in this potential are box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits.
-The tube orbits preserve a sense of rotation about either the long or short axis and are therefore centrophobic, whereas the box orbits may approach arbitrarily close to the center of the potential (centrophilic) and may change the sense of rotation about any axis.
+The tube orbits preserve a sense of rotation about either the long or short axis and are therefore centrophobic, whereas the box orbits may approach arbitrarily close to the center of the potential (centrophilic) and may change sense of rotation about any axis.
 
@@
 The resonant structure of a potential --- the relative importance of particular resonance numbers --- is difficult to compute, but determines the global behavior of orbits in the potential.
-For a triaxial potential the resonances define planes in frequency-space, but resonance surfaces in action-space are generic 2D surfaces.
-In plots of frequency ratios, the resonances appear as lines; Figure~\ref{fig:cartoons} (left panel) shows a cartoon portrait of a portion of frequency-space for an integrable potential with resonance lines, non-resonant, and resonant orbits marked.
+For a triaxial potential the resonances define planes in frequency-space, but resonance surfaces in action-space are generic surfaces.
+In plots of frequency ratios, the resonances appear as lines; Figure~\ref{fig:cartoons} (left panel) shows a cartoon portrait of a portion of frequency-space for an integrable potential with example resonance lines, non-resonant, and resonant orbits marked.
 For an integrable potential, all non-resonant orbits are regular.
@@
 \end{equation}
-where $H_0(\bs{J})$ represents an integrable Hamiltonian, and $\epsilon$ is a small parameter that determines the perturbation strength \citep[a complete description of perturbation theory applied to nonlinear Hamiltonians is given in][]{lichtenberg83}.
-When $0 < |\epsilon| \ll 1$, resonant surfaces become ``thick'' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99}, and these layers regions are then generically surrounded by stochastic layers where chaotic motion occurs.
+where $H_0(\bs{J})$ represents an integrable Hamiltonian, and $\epsilon$ is a small parameter that determines the perturbation strength \citep[a description of perturbation theory applied to nonlinear Hamiltonians is given in][]{lichtenberg83}.
+When $0 < |\epsilon| \ll 1$, resonant surfaces become ``thick'' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99}.
+These resonant layers are then generically surrounded by stochastic layers where chaotic motion occurs.
 Chaotic orbits are characterized by exhibiting random behavior despite being generated by entirely deterministic equations of motion.
@@
 Chaotic orbits in the stochastic layer (red) behave erratically depending on the surrounding resonance structure.
-If the stochastic layer is small and the chaotic orbit is therefore confined in some way, it may behave nearly regular for long periods of time.
+If the stochastic layer is small and the chaotic orbit is therefore confined, the orbit may behave nearly regular for long periods of time.
 
@@
 We therefore expect a large number of regular orbits will survive --- the so-called KAM tori --- however the tori that survive will be separated by regions of chaotic motion.
-Any transformations to angle-action coordinates must be defined locally due to the destruction of tori caused by the onset of chaotic motion.
+Any transformations to angle-action coordinates must be defined locally due to the destruction of tori due to chaotic motion.
 
@@
 
-Chaotic orbits ...
+Chaotic orbits have no orbital actions because they are only strictly bound to their energy hypersurface (if the potential is time-independent).
+The orbits therefore do not have a single set of fundamental frequencies, but rather the frequencies that describe the character of motion evolve with time.
+In near-integrable potentials, chaotic diffusion of an orbit in actions and frequencies occurs both across resonance layers (a sort of stochastic libration) and along the stochastic layers that surround the resonances (Arnold diffusion).
+For weakly chaotic orbits, motion across a resonance layer can occur with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
+Thus, if the resonance libration frequency is comparable (within a factor of a few) to the orbital frequency, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
+However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can't change by large factors (unless there is resonance overlap and the motion is strongly chaotic).
+
+The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict, however, this has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}.
+For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; an orbit that ergodically mixes over its energy hypersurface must traverse this web, though the timescales typical for this phenomenon are many thousands of orbital periods.
+For galactic timescales, Arnold diffusion is not expected to be relevant, however chaotic motion across resonances can occur over short times.
+[the volume allowed to explore over short times is small, but could be relevant for tidal debris...]
+
+\subsection{Mixing of orbit ensembles}
 
@@
 
-Our goal is to map the orbit structure of arbitrary potentials, with an emphasis on identifying the chaotic orbits and and understanding the evolution of these orbits over short times.
+Our goal is to map the orbit structure of arbitrary potentials, with an emphasis on identifying the chaotic orbits and understanding the evolution of these orbits over short times.
 In this section, we describe the methods we will use to detect and quantify the strength of chaos for large grids of orbits.
@@
 
- \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these halos.
+ \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution (dark-matter-only) N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these halos.
 JS02 finds median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{millenium, hmm, butsky15} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
@@
 \begin{equation}
-	t_{\rm \lambda} = \frac{1}{\lambda_{\rm max}}.
+	t_{\rm \lambda} = \frac{1}{\lambda_{\rm max}} \todo{apw}{divide by T?}.
 \end{equation}
@@
 The motion in any canonical coordinate can therefore be decomposed as a Fourier sum (Equation~\ref{eq:fourier}) where the Fourier frequencies are linear, integer combinations of the fundamental frequencies (Equation~\ref{eq:fourierfreq}).
-\cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hamming-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
+\cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
 It has been shown that using a filtered FFT makes the accuracy in determining the frequencies converge much faster than the typical $\inttime^{-1}$ expected for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time.
@@
 For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two separate sections of orbit.
-The frequencies picked out by \superfreq\ for such orbits will therefore represent the average periodic nature of the orbit over a given period.
-We define the fractional frequency diffusion rate per orbital period, $\mathcal{R}$, in the $i$th fundamental frequency as
+The frequencies picked out by \superfreq\ for such orbits will therefore represent the average periodic nature of the orbit over a given integration window.
+We define the fractional frequency diffusion rate per orbital period, $\mathcal{R}$, in the $k$th fundamental frequency as
 \begin{equation}
-	\mathcal{R}_i = \frac{\Omega_i^{(2)} - \Omega_{i}^{(1)}}{\Omega_{i}^{(1)} \, N} \label{eq:fdrate}
+	\mathcal{R}_k = \frac{\Omega_{k}^{(2)} - \Omega_{k}^{(1)}}{\Omega_{k}^{(1)} \, N} \label{eq:fdrate}
 \end{equation}
-where $\Omega_i^{(1)}$ and $\Omega_i^{(2)}$ is the $i$th fundamental frequency computed in each of two consecutive sections of orbit and $N$ is the number of orbital periods in a single section of orbit.
+where the upper index refers to which of the two consecutive sections of orbit and $N$ is the number of orbital periods in a single section of orbit.
 By inverting this rate, we can compute the timescale over which we expect order-unity changes to the fundamental frequencies: the \emph{frequency diffusion time} is defined as
 \begin{equation}
-	t_\Omega = \, (\max_{a_i} \, \mathcal{R}_i)^{-1} \label{eq:fdtime}
+	t_\Omega = \, (\max_{a_k} \, \mathcal{R}_k)^{-1} \label{eq:fdtime}
 \end{equation}
 where the maximum is taken with respect to the corresponding amplitudes, $a_i$, of the fundamental frequency components \citep[see][]{valluri12}, and the time is in units of number of orbital periods, ${\rm T}$.
+
 For a small ensemble of orbits (e.g., tidal debris), the relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
@@
 
-We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy (per unit mass) chosen to span a range of distances comparable to the scale radius of the potential ($E=-(397.2~{\rm km~s}^{-1})^2$ in physical units; see Table~\ref{tbl:potential}).
+We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy (per unit mass), $E$, chosen to span a range of distances comparable to the scale radius of the potential ($E=-(397.2~{\rm km~s}^{-1})^2$ in physical units; see Table~\ref{tbl:potential}).
 We fix $v_x = v_z = 0$, and compute $v_y$ from the energy.
 The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic.
-Tube orbits are generally less stochastic than box orbits --- the other major class of orbits in triaxial potentials --- which tend to plunge deep into the inner regions of the potential, however there are appreciable regions of stochastic tube orbits.
+Tube orbits are generally less stochastic than box orbits, which tend to plunge deep into the inner regions of the potential where the density profile is steep, however there are appreciable regions of stochastic tube orbits.
 Thin tidal streams are thought to form along tube orbits rather than box orbits because of the faster disruption expected for stellar systems on radially plunging orbits.
@@
 
-To compute the Lyapunov exponents, we integrate all orbits in the grid for 1000 orbital periods \todo{apw}{re-run for 5000 or 10000 periods?} and use the method described in Section~\ref{sec:lyap}.
+We integrate all orbits in the grid for 1000 orbital periods \todo{apw}{re-run for 5000 or 10000 periods?} and use the method described in Section~\ref{sec:lyap} to compute the Lyapunov exponents.
 Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $xz$ plane) --- each pixel corresponds to an orbit, and the logarithm of the Lyapunov time (in units of Gyr \todo{apw}{ convert to num.
@@
 By fixing the integration time (in units of orbital periods), the MLE cannot detect weak chaos and the majority of orbits appear to be regular because they have exceedingly long Lyapunov times (all white points have $t_\lambda \gtrsim XX$).
+\todo{apw}{describe other features in lyapunov grid?}
 
@@
 The tube orbits in this potential are mostly regular or only mildly chaotic --- the largest regular regions are associated with the short-axis and long-axis tube orbits --- however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.
+\todo{apw}{describe unstable resonances, etc.}
 
@@
 
-The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant \emph{local} to a given orbit --- that is, these quantities are measures of how infinitesimal deviations will diverge on average from some parent orbit of interest, or of how long it takes for the frequencies to change by some amount.
+The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant \emph{local} to a given orbit --- that is, these quantities are measures of how infinitesimal deviations will diverge on average from some parent orbit of interest, or of how long it takes for the frequencies of a single orbit to change by some amount.
 Tidal debris is disrupted from progenitor systems with finite spreads in orbital properties (e.g., energy).
@@
 	\delta\bs{v}_i &\sim \mathcal{N}(0, \sigma_v)\\
-	r_{\rm tide} &= \|\bs{x}_0\| \left(\frac{m}{3M_{\rm enc}(\|\bs{x}_0\|)}\right)^{1/3} \\
-	\sigma_v &= \|\bs{v}_0\|\left(\frac{m}{3M_{\rm enc}(\|\bs{x}_0\|)}\right)^{1/3} 
+	r_{\rm tide} &= \|\bs{x}_0\| \left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} \\
+	\sigma_v &= \|\bs{v}_0\|\left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} 
 \end{align}
-where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the debris, and $\|\cdot \|$ is the Euclidean norm.
+where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the debris, and $\|\cdot \|$ is the Euclidean norm.
 We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales.
-Figure~\ref{fig:ensembles} shows such ensembles initialized around the three orbits described above and evolved for 16 orbital periods (same as the N-body simulations); the morphology of the ensembles is visually similar to the ``oldest'' (first stripped) debris in the N-body simulations.
+Figure~\ref{fig:ensembles} shows final positions of test-particle0 ensembles initialized around the three orbits described above and evolved for 16 orbital periods (same as the N-body simulations); the morphology of the ensembles is visually similar to the ``oldest'' (first stripped) debris in the N-body simulations.
 
@@
 As a demonstration of the small-scale frequency modulation, we consider again the three orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:three-orbits}).
-Figure~\ref{fig:fdiff-zooms} shows zoomed-reruns of the frequency diffusion time map of Figure~\ref{fig:freqdiff} centered on the three orbits --- the pixel positions are now shown relative to the three parent orbits.
-The frequency diffusion times shown in these grids were computed with two consecutive windows of length 35 orbital periods (shorter than the windows used in the previous, large grid of Figure~\ref{fig:freqdiff}).
+Figure~\ref{fig:fdiff-zooms} shows zoomed regions of the frequency diffusion time map of Figure~\ref{fig:freqdiff} centered on the three orbits --- the pixel positions are now shown relative to the three parent orbits.
+The frequency diffusion times shown in these grids were computed with two consecutive windows of length 40 orbital periods.
 The resonant structure around the mildly chaotic orbit (white regions in panel W) bound the orbit to a small region, whereas for the strongly chaotic orbit (S), the entire region surrounding the orbit is strongly chaotic.
 To resolve the short-time behavior of the frequency diffusion (corresponding to motion across resonance layers) we compute the frequencies in a series of rolling windows along each orbit.
-We use a window with a width equal to 35 orbital periods and shift the window by half an orbital period between each calculation of the most significant frequencies --- this is the smallest window size we could choose before \superfreq\ begins to consistently fail to recover the frequencies for the strongly chaotic orbit.
+We use a window with a width equal to 40 orbital periods and shift the window by half an orbital period between each calculation of the most significant frequencies --- this is the smallest window size we could choose before \superfreq\ begins to consistently fail to recover the frequencies for the strongly chaotic orbit.
 Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time --- plotted are the percent deviations of the frequency ratios from the median value over all windows.
@@
 
-We see now where the discrepancy between chaotic timescales and the observable effects of chaos arise in tidal debris: even if the large-scale diffusion of frequencies is slow, an orbit may explore a small region of frequency space over much shorter times.
+We see now where the discrepancy between chaotic timescales and the observable effects of chaos arise in tidal debris: even if the large-scale diffusion of frequencies is slow, an orbit may explore a  region of frequency space over much shorter times.
 The frequency diffusion time (Equation~\ref{eq:fdtime}) is an estimate of the time over which the mean value of the frequencies evolves, however the \emph{variance} of the frequencies over short times is dynamically relevant for small ensembles.
@@
 
-We now repeat the above analysis of the short-time evolution of frequencies, but now for ensembles of orbits around each of the three parent orbits.
-\todo{apw}{a new metric, measuring the frequency variance, Figure~\ref{fig:freqvar_map} ...}
+We repeat the above analysis of the short-time evolution of frequencies for all orbits in the orbit grid used above (Section~\ref{results1}): for each orbit, we integrate for 80 orbital periods, but compute the frequencies now in a rolling window along this integration time.
+We fix the window width to 40 orbital periods and shift the window by 0.5 orbital periods between each computation of the frequencies.
+For each orbit, we have the three frequencies (and amplitudes), $k=1,2,3$, computed at times $j$; we compute the fractional amplitude of the frequency variation for the $i$th orbit as
+\begin{equation}
+	? = \max_{a_k} \left(\frac{\max_j \Omega_{ik} - \min_j \Omega_{ik}}{2 \, \mean{\Omega_{ik}}_j}\right) \label{eq:frac-freq-amp}
+\end{equation}
+where, as in Equation~\ref{eq:fdtime}, the outer maximum is taken with respect to the amplitude of the frequency in the power-spectrum, $a_k$, and the inner maximum and minimum are taken over time, $j$.
+This is a measure of the magnitude of the short-time frequency evolution for a single orbit.
+For tidal debris with small spreads in orbital properties, nearby orbits will have similar behavior; for example, the central marker in Figure~\ref{fig:fdiff-zooms} shows the size of a globular-cluster-like system in configuration space.
+When the magnitude of the short-time frequency variability is comparable to the spread in orbital frequencies within the debris, the debris effectively has a larger spread in frequencies (the effective frequency distribution is like the convolution of the initial spread with a distribution that scales with the scale-length of the short-time variation).
+Therefore, the debris will ``fan'' or spread out over a few dynamical times and appear dynamically hotter.
+
+Figure~\ref{fig:XX} shows the grid of initial conditions, now colored by the (fractional) amplitude of the short-time frequency variation (Equation~\ref{eq:frac-freq-amp}).
+When the fractional excursions of the frequencies are of order $10^{-3}$--$10^{-2}$ or larger, globular cluster debris will appear ``fluffier.'' Comparing this to the timescales of Figure~\ref{} and Figure~\ref{}, it is clear that by analyzing the small-scale variations of the frequencies of a progenitor orbit over a given window of time provides a better prediction of the resulting morphology of tidal debris.
 
@@
 We have considered here a simple triaxial gravitational potential chosen to mimic the median properties of dark matter halos formed in dark-matter-only simulations (or the large-scale properties of halos formed in simulations with baryonic effects).
-We have found that
+We have found that ``mean'' chaos indicators such as the Lyapunov exponent or frequency diffusion rate predict that chaos should be dynamically unimportant in the halos of galaxies for the majority of orbits (e.g., the predicted timescales are 100-1000s of times longer than the typical orbital times, at least in this simple potential).
+However, while the mean chaotic evolution of, for example, the fundamental frequencies of an orbit may be prohibitively long, the frequencies of even mildly chaotic orbits do evolve on a smaller scale over much shorter times.
+The magnitude of these variations depends on the resonant structure local to the orbit (which is highly potential-dependent), but this result is generic to any potential with resonances and regions of stochasticity.
+When this small-scale but short-time frequency evolution has an amplitude comparable to or larger than the intrinsic spread in orbital frequencies for tidal debris, the debris will appear dynamically hotter and diffuse faster than it would along a regular orbit.
+We have found that even in the simple potential used in this work (Section~\ref{sec:potential}), a large fraction of orbits would naively be considered weakly chaotic, but tidal streams formed on these orbits would appear morphologically different (e.g., more diffuse) than those formed on regular orbits.
 
-\subsection{Cautions}
+If this is the case for a single-component, static, smooth potential, the mere existence of thin tidal streams in the halo of the Milky Way either (1) provides useful information about the potential on these scales by, e.g., implying a large degree of regularity, or (2) indicates that the thin, long streams (GD-1, ??) are on regular orbits.
+These are not mutually exclusive --- in fact, if the streams are on regular orbits, this would be a powerful way to check or rule out possible potential models.
+The longest thin streams are most valuable for this effort because they have clearly evolved for a long time, but the debris remains compact in frequency space.
+For shorter thin streams it will be hard to decouple the unknown evolution time from the analysis.
 
-Since all known tidal streams have been discovered as spatial over-densities of stars, this implies (1) [something about accretion history], (2) the long, thin tidal streams we see may represent a sampling from ``special'' orbits in the Galactic potential where the mixing time is long, and (3) previous assumptions that connect the length and size of streams to their evolution time may be invalid.
+\subsection{Ophiuchus stream}
+
+The Ophiuchus stream \citep{bernard14, sesar15} appears to be a thin, short tidal stream (deprojected length $\approx$1.5--2 kpc) near the Galactic bulge with no apparent progenitor.
+At Galactocentric $R \approx $, $z \approx $, the orbit of this stream likely passes through the MW disk, feels the triaxiality of the Galactic bar \citep[e.g.,][]{wegg13, wegg15}, and has a short orbital period (relative to streams in the halo).
+It is possible that the observed debris is the last remnants of the recently disrupted progenitor system \citep{sesar15}, however if stars were disrupted on previous pericentric passages, this debris is likely ``fanned'' but still near the observed portion of the stream.
+The fanned debris will be much lower surface brightness and thus much more difficult to detect, but would be [...].
+
+\subsection{Potential recovery with streams}
 
@@
 
+\subsection{Potential choice}
 [discussion of potential choice?]
@@
 
-Moved from intro:
-This also has implications for modeling the accretion histories of galaxies.
-Debris morphology is often separated into two classes --- streams (usually from high angular momentum progenitor orbits) and shells (from radially plunging orbits) --- but in practice the distinction may not be as clear due to other mixing effects.
+The potential considered in this work is ``unrealistic'' for the potential of a Milky-Way-like galaxy in that it is static, smooth, and does not contain baryonic components.
+Hydrodynamic simulations of forming galactic disks in cosmological dark matter halos have shown that baryonic feedback and relaxation can significantly change the inner distribution of dark matter and either make the potential more spherical or oblate \citep{many}.
+
+Time dependence of the Galactic potential is difficult to include generically.
+While, for example, the spherically average mass profile of a dark matter halo may evolve predictably \citep[in cosmological simulations;][]{wechsler02, buist14}, the global shape may change due to accretion and phase-mixing of subhalos, bulk rotation, or collapse.
+Over the last $\approx$6 Gyr, however, the Milky Way may have had a relatively calm accretion history and therefore these effects may not be important.
+In future work we will explore the effect of steady mass-accretion on the orbit structure of more complex potentials
 
 \section{Summary and Conclusions}\label{sec:conclusions}
+
 We have shown that ...
@@
 \acknowledgements
-The authors wish to thank ... Robyn Sanderson, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
+The authors wish to thank Robyn Sanderson, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 
@@
 
-\section{Numerical Approximation of Fundamental Frequencies}\label{sec:naffapdx}
+\section{SuperFreq}\label{sec:naffapdx}
 
+\todo{apw}{do we need this test?}
 Figure~\ref{fig:logfreqs} shows a test of the implementation in which we reproduce the frequency map at a fixed energy of an axisymmetric, logarithmic potential \cite[][pg.
@@
 % Figure ??
-\begin{sidewaysfigure}[!p]
+%\begin{sidewaysfigure}[!p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/cartoons.pdf}
+%\caption{} 
+%\label{fig:cartoons}
+%\end{center}
+%\end{sidewaysfigure}
+\begin{figure*}[!p]
 \begin{center}
@@
 \end{center}
-\end{sidewaysfigure}
+\end{figure*}
 

### 2015-06-20 ms.tex
added new figure and captions

@@
 \newcommand{\inttime}{t_{\rm int}}
+\newcommand{\periods}{\ensuremath{{\rm T}}}
 
@@
 The MLE, $\lambda_{\rm max}$, is interpreted as a rate that sets the exponential divergence of infinitesimally close chaotic orbits.
-It is therefore useful to consider the corresponding $e$-folding time,
+It is therefore useful to consider the corresponding $e$-folding time by inverting the rate.
+We normalize the Lyapunov time by the maximum orbital period of a given orbit, $T_{\rm max}$, so that the time is in units of number of orbital periods
 \begin{equation}
-	t_{\rm \lambda} = \frac{1}{\lambda_{\rm max}} \todo{apw}{divide by T?}.
+	t_{\rm \lambda} = \frac{1}{\lambda_{\rm max} \, T_{\rm max}}
 \end{equation}
@@
 We find that the configuration-space density of ensembles around weakly chaotic orbits evolve faster (e.g., in mean density) than expected given the timescales over which chaos is computed to be relevant for the parent orbits.
-In Section~\ref{sec:results3}, we explain this phenomenon in the context of how chaotic diffusion occurs (e.g., Section~\ref{sec:sec:behavior-chaotic}).
+In Section~\ref{sec:results3}, we explain this phenomenon in the context of how chaotic diffusion occurs (e.g., Section~\ref{sec:behavior-chaotic}).
 
@@
 
-We integrate all orbits in the grid for 1000 orbital periods \todo{apw}{re-run for 5000 or 10000 periods?} and use the method described in Section~\ref{sec:lyap} to compute the Lyapunov exponents.
-Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $xz$ plane) --- each pixel corresponds to an orbit, and the logarithm of the Lyapunov time (in units of Gyr \todo{apw}{ convert to num.
-periods?}) is mapped to greyscale intensity.
+We integrate all orbits in the grid for 10000 orbital periods and use the method described in Section~\ref{sec:lyap} to compute the Lyapunov exponents.
+Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $xz$ plane) --- each pixel corresponds to an orbit, and the logarithm of the Lyapunov time (in units of number of orbital periods) is mapped to greyscale intensity.
 The darker pixels have shorter Lyapunov times and are more chaotic.
-By fixing the integration time (in units of orbital periods), the MLE cannot detect weak chaos and the majority of orbits appear to be regular because they have exceedingly long Lyapunov times (all white points have $t_\lambda \gtrsim XX$).
-\todo{apw}{describe other features in lyapunov grid?}
+By fixing the integration time (in units of orbital periods, $\periods$), the MLE cannot detect weak chaos and the majority of orbits appear to be regular because they have exceedingly long Lyapunov times (all white points have $t_\lambda \gtrsim 1000~\periods$).
 
@@
 The tube orbits in this potential are mostly regular or only mildly chaotic --- the largest regular regions are associated with the short-axis and long-axis tube orbits --- however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.
-\todo{apw}{describe unstable resonances, etc.}
 
+\todo{apw}{describe other features in lyapunov grid and unstable resonances, etc.}
 The strongest chaotic regions appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}).
@@
 For each simulation, the progenitor is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
-\todo{apw}{Label the ICs for the 3 orbits on Figures 3 and 4} We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter.
+We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter.
 Figure~\ref{fig:nbodysims} shows the final particle distributions rotated so that the angular momentum of the progenitor orbit is aligned with the $z$-axis.
@@
 For example, Figure~\ref{fig:ensemble-density} shows the evolution of the configuration-space density for ensembles evolved around the three orbits (R, W, S) described above.
-At each time step, we use kernel density estimation (KDE)\footnote{We use an Epanechnikov kernel with a bandwidth set to 1 kpc.
-We use an implementation from the Python package \texttt{scikit-learn} \citep{scikitlearn}.} with the ensemble of particle positions to estimate the configuration-space density field.
+At each time step, we use kernel density estimation (KDE)\footnote{We use an implementation from the Python package \texttt{scikit-learn} \citep{scikitlearn}.} with the ensemble of particle positions to estimate the configuration-space density field.
+We use an Epanechnikov kernel with an adaptive bandwidth: at each evaluation of the density, we use 10-fold cross-validation to find the optimal kernel bandwidth.
 We evaluate the density at the positions of each particle, $\rho_i$, and compare to the mean initial density, $\mean{\rho_0}$.
-Figure~\ref{fig:ensemble-density} shows histograms of the logarithm of $\rho_i/\mean{\rho_0}$ at three different times: after evolution for 4 (left), 8 (middle), and 16 (right) orbital periods.
-Both the mean density and shape of the density distribution changes over time and the nature of the evolution depends on the strength of chaos along the orbit.
-The regular orbit distribution stays relatively compact indicating that the density stays fairly uniformly high throughout the stream.
-For the weakly chaotic orbit, already after 8 orbits the density distribution shows a large tail to smaller densities --- this is ``fanned'' debris --- but maintains a significant number of particles in relatively high densities.
+Figure~\ref{fig:ensemble-density} shows histograms of the logarithm of $\rho_i/\mean{\rho_0}$ for each of the three orbit ensembles at three different times: the initial density distribution (labeled $0~\periods$), and after evolution for 4 and 16 orbital periods.
+The distributions all shift towards lower densities as time progresses, however other subtle effects are noticeable.
+The regular orbit distribution stays relatively compact at all times indicating that the density stays fairly uniformly high throughout the stream.
+For the weakly chaotic orbit, the distribution is somewhat wider with a larger tail to lower densities --- this is the debris that begins to fan out.
 The strongly chaotic orbit evolves much faster to low density.
-Stars with separations much larger than the bandwidth have unresolved density measurements --- this is why there appears to be a pile-up of stars at $\log_{10}\left[\rho_i / \mean{\rho_0}\right] \approx -3$.
-\todo{apw}{do we need to adjust bandwidth as the density evolves?}
 
-The distributions all start with a peak at high density and a tail to lower density but evolve to the inverse state with a peak at low density and a tail to higher densities --- the skewness of a distribution is sensitive to exactly this property.
-Figure~\ref{fig:density-skewness} shows the mean density (top row) and skewness (bottom row) computed at 256 evenly-spaced intervals from the initial ensemble distribution to the distribution after 16 orbital periods.
-Both quantities (mean density and skewness) are correlated with the strength of chaos surrounding the parent orbit.
+The density distributions do not change shape dramatically, even for the strongly chaotic ensemble; we therefore use the mean density as a metric for studying the density evolution of orbit ensembles.
+Figure~\ref{fig:densities} shows the mean density of the ensemble particles computed at 128 evenly-spaced intervals from the initial ensemble distribution to the distribution after 16 orbital periods.
+It is clear that the density of the regular orbit ensemble evolves slower than the weakly chaotic ensemble, and the strongly chaotic ensemble quickly becomes diffuse.
 
-Motivated by the apparent detection of chaos using these ensemble measures even around the weakly chaotic orbit, we compute the final skewness and mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
-For all ensembles, we integrate the orbits for 16 (parent) orbital periods and store the mean density and skewness at the end of integration for each.
-Figures~\ref{fig:ensemblemap-meandensity} and \ref{fig:ensemblemap-skewness} again show the grid of initial conditions (Section~\ref{sec:results1}), but now the greyscale indicates the final mean density or skewness, respectively.
+Motivated by the noticeable discrepancy between the regular and weakly chaotic orbit ensembles --- even though the chaotic timescale of the weakly chaotic ensemble parent orbit is thousands of orbital periods ---we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
+For all ensembles, we integrate the orbits for 16 (parent) orbital periods and store the mean density at the end of integration for each.
+Figure~\ref{fig:ensemblemap-meandensity} again shows the grid of initial conditions (Section~\ref{sec:results1}), but now the greyscale indicates the final mean density.
 \todo{apw}{talk about the structure here, compare to freq.
@@
 \includegraphics[width=\textwidth]{figures/cartoons.pdf}
-\caption{} 
+\caption{An illustrative demonstration of the orbit structure for integrable (left), near-integrable (middle), and non-integrable (right) potentials in terms of the ratios of the fundamental frequencies.
+In an integrable potential, only resonant and non-resonant orbits exist, and all orbits are regular (resonances appear as lines in frequency-ratio-space).
+If the potential is perturbed mildly, resonance layers form around the resonances that host near-resonant orbits which behave like resonant orbits but have an additional frequency corresponding to libration about the resonance.
+Where resonances overlap or near separatrices, stochastic layers form and orbits will be chaotic.
+For more strongly perturbed potentials, many of the non-resonant orbits may become chaotic but resonance layers may grow and still remain regular.
+}
 \label{fig:cartoons}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/lyap_map.png}
+\includegraphics[width=\textwidth, trim={0 1cm 0 0}]{figures/lyap_map.png}
 \caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
 Each pixel in this image represents a single orbit, and the greyscale corresponds to the logarithm of the Lyapunov time (in units of orbital periods).
-Chaotic orbits with $t_\lambda \gtrsim XX~{\rm T}$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
-\todo{apw}{remove Gyr and express in terms of orbital periods, add title, run Lyapunov grid for 5000 orbital periods instead of 1000?}} \label{fig:lyapmap}
+Orbits are integrated for a total of 10000 orbital periods.
+Chaotic orbits with $t_\lambda \gtrsim 1000~\periods$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
+Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S), weakly chaotic (W), and regular (R) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:lyapmap}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/fdiff_map.png}
+\includegraphics[width=\textwidth, trim={0 1cm 0 0}]{figures/fdiff_map.png}
 \caption{The same grid of initial conditions as Figure~\ref{fig:lyapmap}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
-The frequencies are computed in two consecutive windows, each of which has length equal to 50 orbital periods.
+The frequencies are computed in two consecutive windows, each of which has length equal to 40 orbital periods.
 The frequencies are measured precisely so that small changes in the frequencies can be detected over just $\approx$10s of orbits.
-\todo{apw}{remove Gyr and express in terms of orbital periods, add title} } \label{fig:freqdiff}
+Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S), weakly chaotic (W), and regular (R) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
 \end{center}
@@
 From left to right, the columns indicate that the progenitor is on a regular (R), weakly chaotic (W), or strongly chaotic (S) orbit.
-Even on the weakly chaotic orbit --- where the predicted chaotic timescale is XX orbital periods --- the resulting tidal stream displays stream-fanning and the ``oldest'' debris is more diffuse than in the regular case.
-\todo{apw}{suptitle 'N-body simulations'?}} \label{fig:nbodysims}
+Even on the weakly chaotic orbit --- where the predicted chaotic timescale is XX orbital periods --- the resulting tidal stream displays stream-fanning and the ``oldest'' debris is more diffuse than in the regular case.} \label{fig:nbodysims}
 \end{center}
@@
 The particle positions qualitatively match the morphology of the ``oldest'' debris from the N-body simulations, though there are many more particles in these ensembles than are disrupted during the first pericentric passage of the simulations.
-\todo{apw}{suptitle 'Orbit ensembles'?} } \label{fig:ensembles}
+} \label{fig:ensembles}
 \end{center}
@@
 \includegraphics[width=\textwidth]{figures/ensemble-density.png}
-\caption{ Distributions of the values of the estimated configuration-space density field evaluated at the positions of the ensemble particles after 4 (left), 8 (middle), and 16 (right) orbital periods.
-Colors represent the three orbits of Figures~\ref{fig:nbodysims} and \ref{fig:ensembles}: the regular, weakly chaotic, and strongly chaotic orbits.
+\caption{ Distributions of the values of the estimated configuration-space density field evaluated at the positions of the ensemble particles after 0, 4, and 16 orbital periods for each of the three orbits of Figures~\ref{fig:nbodysims} and \ref{fig:ensembles}.
 The density distribution of the regular orbit ensemble stays compact and decreases slowly.
-The density of the weakly chaotic ensemble initially evolves quickly to lower density, but then broadens.
-The strongly chaotic ensemble evolves fast to a state of low density.
-\todo{apw}{can't distinguish colors when printed...}} \label{fig:ensemble-density}
+The density of the weakly chaotic ensemble initially evolves faster to lower density and has a larger tail to lower densities.
+The strongly chaotic ensemble evolves fast to a state of low density.} \label{fig:ensemble-density}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-density-skewness.png}
-\caption{ } \label{fig:density-skewness}
+\includegraphics[width=\textwidth]{figures/ensemble-densities.png}
+\caption{This is an alternate visualization of the density evolution of orbit ensembles around the three orbits of Figure~\ref{fig:ensemble-density}, showing only the evolution of the mean density for the orbits (normalized by the initial density of the ensemble).}
+\label{fig:densities}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/ensemblemap-meandensity.png}
-\caption{ } \label{fig:ensemblemap-meandensity}
-\end{center}
-\end{figure*}
-
-% Figure ??
-\clearpage
-\begin{figure*}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/ensemblemap-skewness.png}
-\caption{ } \label{fig:ensemblemap-skewness}
+\includegraphics[width=\textwidth]{figures/ensemble-map.png}
+\caption{The same grid of initial conditions as Figure~\ref{fig:lyapmap} and other previous figures, but now we consider the density evolution of small ensembles of orbits around each of the orbits in this grid.
+We generate globular-cluster-scale ensembles of orbits around each ``parent'' orbit in the grid (Section~\ref{sec:results2}) and follow the density of the ensemble as a function of time using kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
+Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S), weakly chaotic (W), and regular (R) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).}
+\label{fig:ensemblemap-meandensity}
 \end{center}
@@
 \includegraphics[width=1.1\textwidth]{figures/fdiff_zooms.png}
-\caption{ } \label{fig:fdiff-zooms}
+\caption{Zooms of the frequency diffusion time map of Figure~\ref{fig:freqdiff} around the three orbits highlighted in that figure and used in Figures~\ref{fig:nbodysims}--\ref{fig:ensemble-density} --- as in Figure~\ref{fig:freqdiff}, each pixel represents an orbit, and the greyscale of the pixel is set by the frequency diffusion time.
+Around each orbit, we initialize a new grid of $\approx$40,000 orbits to resolve the small-scale resonant structure around each orbit.
+Each orbit in the sub-grid is integrated for 80 orbital periods and the frequency diffusion time is computed from the two consecutive sections of 40 orbital periods.
+The marker at the center indicates the scale (e.g., tidal radius) of a $M \approx 10^4~\msun$ globular cluster at the given position in the initial condition grid.
+The regular orbit (R) is near a resonance (white line), the weakly chaotic orbit (W) is surrounded by an intricate resonant structure and may be in a region of resonance overlap, and the strongly chaotic orbit (S) is embedded in a region of strong stochasticity.
+}
+\label{fig:fdiff-zooms}
 \end{center}
@@
 \includegraphics[width=\textwidth]{figures/freq-evolution.png}
-\caption{ } \label{fig:three-orbits-freqs}
+\caption{ } 
+\label{fig:three-orbits-freqs}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/freqvar.png}
-\caption{ } \label{fig:freqvar_map}
+\includegraphics[width=\textwidth]{figures/freqvar-map.png}
+\caption{ } 
+\label{fig:freqvar_map}
+\end{center}
+\end{figure*}
+
+% Figure ??
+\clearpage
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/freqvar-map-stretched.png}
+\caption{ } 
+\label{fig:freqvar_map_stretched}
 \end{center}

### 2015-06-21 ms.tex
more words

@@
 
-\todo{apw}{describe other features in lyapunov grid and unstable resonances, etc.}
-The strongest chaotic regions appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}).
-The resonant structure is revealed in the frequency diffusion map from integrations of just 80 orbital periods, though the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods.
+The strongest chaotic regions (black) appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}).
+Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map, especially where resonances overlap --- for example, near $(x,z) \approx(13,7)$ and $(x,z) \approx(31,13)$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}.
+The details of the resonant structure is revealed in the frequency diffusion map from integrations of just 80 orbital periods, though the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods.
 In the next section, we analyze the density evolution of finite-volume ensembles of orbits around each orbit in the above grids and compare the density evolution of the ensembles to the single-orbit chaos indicators computed in this section.
@@
 35 scale radii?
-but this is plummer??}.
+but this is a plummer sphere?}.
 For each simulation, the progenitor is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
@@
 The magnitude of these variations depends on the resonant structure local to the orbit (which is highly potential-dependent), but this result is generic to any potential with resonances and regions of stochasticity.
-When this small-scale but short-time frequency evolution has an amplitude comparable to or larger than the intrinsic spread in orbital frequencies for tidal debris, the debris will appear dynamically hotter and diffuse faster than it would along a regular orbit.
+When this small-scale but short-time frequency evolution has an amplitude comparable to or larger than the intrinsic spread in orbital frequencies for tidal debris, the debris will appear dynamically hotter and become diffuse faster than along a regular orbit.
 We have found that even in the simple potential used in this work (Section~\ref{sec:potential}), a large fraction of orbits would naively be considered weakly chaotic, but tidal streams formed on these orbits would appear morphologically different (e.g., more diffuse) than those formed on regular orbits.
 
-If this is the case for a single-component, static, smooth potential, the mere existence of thin tidal streams in the halo of the Milky Way either (1) provides useful information about the potential on these scales by, e.g., implying a large degree of regularity, or (2) indicates that the thin, long streams (GD-1, ??) are on regular orbits.
-These are not mutually exclusive --- in fact, if the streams are on regular orbits, this would be a powerful way to check or rule out possible potential models.
-The longest thin streams are most valuable for this effort because they have clearly evolved for a long time, but the debris remains compact in frequency space.
-For shorter thin streams it will be hard to decouple the unknown evolution time from the analysis.
+If this is the case for a single-component, static, smooth potential, the mere existence of thin tidal streams in the halo of the Milky Way either (1) provides useful information about the potential on these scales by, e.g., implying a large degree of regularity, or (2) indicates that the thin, long streams (e.g., GD-1) are on regular orbits.
+These are not mutually exclusive --- in fact, if the streams are on regular orbits, this would be a powerful way to check or rule out possible potential models by requiring that the progenitor orbits remain regular.
+The longest thin streams are most valuable for this effort because they have clearly evolved for a long time, but the debris remains compact.
+For shorter thin streams it will be hard to decouple the unknown evolution time from enhanced density evolution from chaos.
 
@@
 The Ophiuchus stream \citep{bernard14, sesar15} appears to be a thin, short tidal stream (deprojected length $\approx$1.5--2 kpc) near the Galactic bulge with no apparent progenitor.
-At Galactocentric $R \approx $, $z \approx $, the orbit of this stream likely passes through the MW disk, feels the triaxiality of the Galactic bar \citep[e.g.,][]{wegg13, wegg15}, and has a short orbital period (relative to streams in the halo).
-It is possible that the observed debris is the last remnants of the recently disrupted progenitor system \citep{sesar15}, however if stars were disrupted on previous pericentric passages, this debris is likely ``fanned'' but still near the observed portion of the stream.
-The fanned debris will be much lower surface brightness and thus much more difficult to detect, but would be [...].
+At Galactocentric $R \approx 1$ kpc, $z \approx 5$ kpc, the orbits of the stream stars likely pass through the MW disk, feel the triaxiality of the Galactic bar \citep[e.g.,][]{wegg13, wegg15}, and have short orbital periods (relative to streams in the halo).
+It is possible that the observed debris is the last remnants of the recently disrupted progenitor system \citep{sesar15}, however if a significant number of stars were disrupted on previous pericentric passages, this older debris may be ``fanned'' and still near the observed portion of the stream.
+The fanned debris would be much lower surface brightness and thus much more difficult to detect.
+However, the detection of such stars would confirm that enhanced density evolution is a dynamically relevant process for thin streams in real galaxies.
 
-\subsection{Potential recovery with streams}
+\subsection{The dangers of stream modeling}
 
-Recovering the potential of triaxial systems is not trivial. Danger of orbit fitting because chaotic orbits can reproduce morphologies of regular orbits.
+Tidal stream modeling is one of the most promising ways to constrain the 3D mass distribution around the Milky Way at distances of 10s to 100s of kpc.
+Methods that use tidal streams to infer properties of the Galactic potential typically operate by constructing models of the debris distribution using either the present-day phase-space density \citep[e.g.,][]{kuepper12, kuepper15}, time-of-disruption phase-space density \citep{apw14}, or the density in angle-action coordinates \citep[e.g.,][]{sanders14, bovy14}.
+All of these methods may fail for modeling globular-cluster streams on mildly chaotic orbits.
+
+Consider a globular cluster system that disrupts along a mildly chaotic orbit so that a significant portion of the resulting stream is more diffuse (lower surface brightness) and therefore unobservable due to observational biases.
+The high-density part of the stream may still appear thin (e.g., the recently disrupted debris in the middle panels of Figure~\ref{fig:nbodysims}).
+If the stream modeling focuses on reproducing the present-day density of the stream and the potential is restricted to an integrable, analytic form, the debris may be well fit by a particular orbit and evolution time in such a potential.
+However, the inferences about the potential or mass distribution may be biased significantly and the inferences about the evolution time and internal properties of the system will certainly be incorrect.
+
+[what happens in angle-action coordinates?]
+
+[what about opposite effect -- if potential model too flexible, can always find chaotic orbit to reproduce regular orbit?]
 
 \subsection{Potential choice}
+
+The potential considered in this work is ``unrealistic'' for the total potential of a Milky-Way-like galaxy in that it is static, smooth, and does not contain baryonic components.
+Hydrodynamic simulations of forming galactic disks in cosmological dark matter halos have shown that baryonic feedback and relaxation can significantly change the inner distribution of dark matter and either make the potential more spherical or oblate \citep{many}.
+However, the significance of baryonic relaxation or of time-dependence, triaxiality, and substructure on shaping the matter distribution within the Milky Way is largely unknown.
+We now consider each of these effects independently and comment on the expected consequences of these perturbations.
+
+\subsubsection{Baryonic components}
+
+\cite[][D08]{debattista08} and \cite[][V10]{valluri10} studied the orbit evolution induced by growing a baryonic disk in dark matter halos with various shapes and orientations (e.g., prolate, triaxial).
+In the simulations of D08, the disk component is artificially fixed along one of the principal axes of the halo potential and the mass of the disk is grown adiabatically.
+D08 found that (in general) the growth of a disk slowly deforms the orbits of mass tracers (e.g., dark matter particles) and preferentially populates tube orbits but does not result in a large degree of chaotic scattering.
+Consequently, the inner shape of the potential becomes more oblate or spherical due to the prevalence of tube orbits.
+
+If the inner halos of galaxies are indeed close to spherical or oblate, the majority of orbits will be regular and chaos will be less important, however this is far from conclusive.
+First, there are observational hints that the intrinsic alignment of galactic disks may be close to random \citep{??} (rather than fixed to symmetry planes of the host halo potential).
+Simulations since the work of D08 have shown that by ``freeing'' the disk component, the angular momentum vector of the disk can precess and wander erratically away or around the principal axes of the host halo \citep{debattista15}.
+It is unknown what effect this would have on the orbit structure of nearby halo orbits, but we naively expect that this would destabilize a significant number of orbits.
+Second, if the outer potential remains triaxial, the transition region between, e.g., oblate and triaxial may be populated by chaotic orbits.
+We have found from simple experiments in superposing potential components that transition regions between potential shapes induce significant amounts of chaos.
+Though superposing potentials is not realistic, it is at least suggestive that such configurations might be unstable.
+A more careful exploration of this phenomenon will be presented in future work.
+Lastly, the existence of a bulge or thin disk component may enhance chaos through a process similar to periodic forcing of the orbit.
+Passage through a thin disk results in a nearly instantaneous, impulsive forcing of the orbit at a frequency set by the circulation frequency of the orbit.
+This may act to increase the significance of chaos along disk-crossing orbits \citep[e.g.,][]{kandrup00}.
+This effect is likely unresolved in many simulations where the force and spatial resolution is insufficient to capture this effect.
+
+\subsubsection{Subhalos}
+
+noise in hamiltonian systems: \citep{kandrup00}
+
+\subsubsection{Time-dependence}
+
+Time dependence of the Galactic potential is difficult to include generically.
+While, for example, the spherically average mass profile of a dark matter halo may evolve predictably \citep[in cosmological simulations;][]{wechsler02, buist14}, the global shape may change due to accretion and phase-mixing of subhalos, bulk rotation, or collapse.
+Over the last $\approx$6 Gyr, however, the Milky Way may have had a relatively calm accretion history and therefore these effects may not be important.
+In future work we will explore the effect of steady mass-accretion on the orbit structure of more complex potentials
+
 [discussion of potential choice?]
@@
 
-The potential considered in this work is ``unrealistic'' for the potential of a Milky-Way-like galaxy in that it is static, smooth, and does not contain baryonic components.
-Hydrodynamic simulations of forming galactic disks in cosmological dark matter halos have shown that baryonic feedback and relaxation can significantly change the inner distribution of dark matter and either make the potential more spherical or oblate \citep{many}.
-
-Time dependence of the Galactic potential is difficult to include generically.
-While, for example, the spherically average mass profile of a dark matter halo may evolve predictably \citep[in cosmological simulations;][]{wechsler02, buist14}, the global shape may change due to accretion and phase-mixing of subhalos, bulk rotation, or collapse.
-Over the last $\approx$6 Gyr, however, the Milky Way may have had a relatively calm accretion history and therefore these effects may not be important.
-In future work we will explore the effect of steady mass-accretion on the orbit structure of more complex potentials
-
 \section{Summary and Conclusions}\label{sec:conclusions}
 
-We have shown that ...
+We have shown that chaos may have a significant effect on the density evolution of tidal debris.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-map.png}
+\includegraphics[width=\textwidth, trim={0 1cm 0 0}]{figures/ensemble-map.png}
 \caption{The same grid of initial conditions as Figure~\ref{fig:lyapmap} and other previous figures, but now we consider the density evolution of small ensembles of orbits around each of the orbits in this grid.
@@
 \includegraphics[width=\textwidth]{figures/freq-evolution.png}
-\caption{ } 
+\caption{Evolution of the fundamental frequency ratios, $r_{1,3} = \Omega_1 / \Omega_3$ and $r_{2,3} = \Omega_2 / \Omega_3$, computed over 256 orbital periods for the three orbits, R, W, S.
+Fundamental frequencies are computed with a window width of $\approx$40 orbital periods, and the window is shifted by one orbital period between each computation (each grey point represents the fundamental frequency ratios computed in a single window).
+Plotted are the per cent deviations of the frequency ratios from the value in the initial window --- the initial value is shown as a green plus sign and the final value is shown as a red x.
+For the regular orbit (R), the fractional variation is around $10^{-6}$ and thus all points overlap on this scale.
+The weakly chaotic orbit (W) displays frequency variations comparable to the spread in frequencies in globular-cluster-like tidal debris.
+The frequency variations of the strongly chaotic orbit (S) can be many 10s of per cent.}
 \label{fig:three-orbits-freqs}

### 2015-06-21 ms.tex
references and more text about potentials

@@
 
-noise in hamiltonian systems: \citep{kandrup00}
+Cosmological simulations predict that dark matter halos are filled with substructure in the form of dark matter subhalos.
+If they exist, these subhalos may account for up to $\approx$1-10\% of the mass of the dark matter \citep{??} and therefore may contribute significantly to and orbit in the large-scale potential of any galaxy.
+Gravitational scattering due to subhalo interactions has been studied, however in the halos of galaxies where dynamical times are long, the scattering cross-section for \emph{strong} encounters is very small \citep{??}.
+Instead, the collective effect of the subhalos may instead act as a noise term in the Hamiltonian of any halo orbit.
+This subhalo-induced colored noise --- which depends on the mass spectrum and distribution of subhalos --- may also act to simply increase the magnitude of chaos along orbits, and destabilize sufficiently non-resonant orbits \citep[see, e.g.,][]{kandrup00}.
 
@@
 
-Time dependence of the Galactic potential is difficult to include generically.
-While, for example, the spherically average mass profile of a dark matter halo may evolve predictably \citep[in cosmological simulations;][]{wechsler02, buist14}, the global shape may change due to accretion and phase-mixing of subhalos, bulk rotation, or collapse.
-Over the last $\approx$6 Gyr, however, the Milky Way may have had a relatively calm accretion history and therefore these effects may not be important.
-In future work we will explore the effect of steady mass-accretion on the orbit structure of more complex potentials
-
-[discussion of potential choice?]
-Why not disk + bulge? What do we expect when we add?
-
-Real disk will be self-consistent with halo, so this isn't really a fair test, but still...
-
-Other effects: radial dependence of axis ratios, time dependence, etc.
+Galaxies are certainly not static systems.
+To first order, galaxies grow in mass --- for example, the spherically averaged mass profile of dark matter halos evolves fairly predictably in cosmological simulations \citep{wechsler02, buist14} after some initial period of stochastic mass growth.
+The Milky Way has probably had a fairly calm accretion history over the last 6 Gyr and therefore the mass growth may be similar to that seen in simulations.
+This steady growth most likely does not alter the global structure or shape of the potential.
+However, we also know from simulations that figure rotation, baryonic feedback, and the accretion and phase-mixing of subhalos do perturb the global state of simulated halos.
+\cite{deibel11} showed that by adopting pattern speeds comparable to those found in cosmological simulations, figure rotation generally acts to destabilize orbits (rather than stabilize chaotic orbits in the equivalent static potential) and the resulting orbit structure is that most regular orbits are associated with resonances.
+In future work we will explore the effect of these time-dependent processes on the chaotic dispersal of tidal streams using live potentials from cosmological N-body simulations.
 
@@
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
-This research made use of Astropy, a community-developed core \texttt{Python} package for Astronomy \citep{astropy13}.
+This research made use of \project{Astropy}, a community-developed core \texttt{Python} package for Astronomy \citep{astropy13}.
 This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance.

### 2015-06-22 ms.tex
so many words

@@
 % Context
-The large-scale, smooth components of galactic gravitational potentials are thought to be triaxial in shape.
-Triaxial potentials that match observational constraints and measurements of the radial profile and axis ratios of dark matter halos in cosmological simulations are almost certainly not globally integrable.
-An appreciable number of chaotic orbits are therefore expected in galactic halos.
-It is yet unclear whether chaotic regions of phase-space will affect the density evolution of stellar tidal streams, which form in the halos of galaxies from the disruption of stellar systems.
-% Aims
-Many numerical methods exist for assessing the degree of chaos for a given orbit (e.g., the Lyapunov exponent, the frequency diffusion rate). 
-These indicators, which are local to a single orbit, often predict a timescale over which chaos will be dynamically important.
-In this paper, we aim to compare these timescales to the density evolution of the finite-volume ensembles of orbits.
-% Methods: 
-We derp derp derp.
-% Results:
-We find that mean indicators such as the Lyapunov time or the (long term) frequency diffusion rate of a given ``parent'' orbit are not good predictors for the short-term density evolution of ensembles of orbits around this parent orbit.
-Instead, we find that the variance of the fundamental frequencies of the ensemble orbits over short timescales is a stronger indicator of the density evolution.
-% Conclusions:
+%The large-scale, smooth components of galactic gravitational potentials are thought to be triaxial in shape.
+%Triaxial potentials that match observational constraints and measurements of the radial profile and axis ratios of dark matter halos in cosmological simulations are almost certainly not globally integrable.
+%An appreciable number of chaotic orbits are therefore expected in galactic halos.
+%It is yet unclear whether chaotic regions of phase-space will affect the density evolution of stellar tidal streams, which form in the halos of galaxies from the disruption of stellar systems.
+%% Aims
+%Many numerical methods exist for assessing the degree of chaos for a given orbit (e.g., the Lyapunov exponent, the frequency diffusion rate). 
+%These indicators, which are local to a single orbit, often predict a timescale over which chaos will be dynamically important.
+%In this paper, we aim to compare these timescales to the density evolution of the finite-volume ensembles of orbits.
+%% Methods: 
+%We derp derp derp.
+%% Results:
+%We find that mean indicators such as the Lyapunov time or the (long term) frequency diffusion rate of a given ``parent'' orbit are not good predictors for the short-term density evolution of ensembles of orbits around this parent orbit.
+%Instead, we find that the variance of the fundamental frequencies of the ensemble orbits over short timescales is a stronger indicator of the density evolution.
+%% Conclusions:
 
@@
 
-When $\epsilon$ is large, there is no general prediction for the resulting behavior, however it seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep{valluri, merritt, dehnen, etc.}.
+When $\epsilon$ is large, there is no general prediction for the resulting behavior, however it seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep[e.g.,][]{valluri98, merritt99}.
 We therefore expect a large number of regular orbits will survive --- the so-called KAM tori --- however the tori that survive will be separated by regions of chaotic motion.
@@
 
-The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict, however, this has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}.
+The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict.
+This has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}.
 For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; an orbit that ergodically mixes over its energy hypersurface must traverse this web, though the timescales typical for this phenomenon are many thousands of orbital periods.
+
 For galactic timescales, Arnold diffusion is not expected to be relevant, however chaotic motion across resonances can occur over short times.
-[the volume allowed to explore over short times is small, but could be relevant for tidal debris...]
+If a stochastic orbit is semi-bounded by surrounding resonances, the allowed volume in which it can explore (in frequency space) is often quite small so that the orbit may appear to be approximately regular.
+However, this small-scale evolution may be important for tidal debris which has small intrinsic spreads in frequencies.
 
@@
 
-Small ensembles of (conditionally-periodic) regular orbits will reach a fully phase-mixed state in a timescale  $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in frequency-space of the ensemble --- each individual orbit will fill the surface of its torus ergodically, but the ensemble will only mix due to sheering from slight variations in their fundamental frequencies \citep{merritt96, helmi99}.
+Ensembles of orbits stuck between resonances may quickly spread to fill the allowed volume via the small-scale but short-time chaotic diffusion processes discussed above.
+This occurs due to phase-mixing of the fundamental frequencies of the ensemble orbits themselves --- the frequencies are evolving and thus trace out an orbit through frequency space.
+This is in contrast to phase-mixing of regular orbits and long-term chaotic mixing of strongly chaotic orbits.
+
+A small ensemble of (conditionally-periodic) regular orbits will reach a fully phase-mixed state in a timescale  $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in frequency-space of the ensemble --- each individual orbit will fill the surface of its torus ergodically, but the ensemble will only mix due to sheering from slight variations in their fundamental frequencies \citep{merritt96, helmi99}.
 Ensembles of chaotic orbits behave quite differently.
@@
 Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits, however analytically computing the rate and amount of spreading is very difficult even for simple potentials \cite[see, e.g.,][]{chirikov??}.
-Certain ensembles may be trapped between resonances for long times and thus remain in a state of pseudo-coherence for long times, whereas others may rapidly diffuse \citep[see, e.g.,][]{merritt96}.
-This phenomenon is generic to potentials that are not globally stochastic in which regions of stability and chaos are intermingled throughout phase-space.
+Typically these rates are comparable or larger than the Arnold diffusion timescale and are thus very long.
 
-\todo{apw}{concluding paragraph...}
+In this work, we investigate the effect of the short-time but small-scale frequency evolution along weakly chaotic orbits and discuss how this effects our understanding of the coherence and density evolution of tidal streams.
 
@@
 We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
-\todo{apw}{Cite this paper: hunter05 for disk crossings and chaos}
 
@@
 
-The MLE, $\lambda_{\rm max}$, is interpreted as a rate that sets the exponential divergence of infinitesimally close chaotic orbits.
+The MLE, $\lambda_{\rm max}$, is interpreted as a rate that quantifies the exponential divergence of infinitesimally close chaotic orbits.
 It is therefore useful to consider the corresponding $e$-folding time by inverting the rate.
-We normalize the Lyapunov time by the maximum orbital period of a given orbit, $T_{\rm max}$, so that the time is in units of number of orbital periods
+We normalize the Lyapunov time by the maximum orbital period\footnote{The maximum of the three orbital periods corresponding to the three fundamental frequencies (see Section~\ref{sec:naff}).} of a given orbit, $T_{\rm max}$, so that the time is in units of number of orbital periods
 \begin{equation}
@@
 \cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
-It has been shown that using a filtered FFT makes the accuracy in determining the frequencies converge much faster than the typical $\inttime^{-1}$ expected for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time.
+It has been shown that the accuracy in determining the frequencies converges much faster using a filtered FFT compared to the typical $\inttime^{-1}$ expected for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time.
 The FFT is iteratively searched for the strongest frequency component, which is subtracted from the spectrum after each iteration.
@@
 \begin{equation}
-	W(t/\inttime) = \frac{2^p \, (p!)^2}{(2p)!} \left( 1 + \cos \frac{\pi t}{\inttime}\right)^p
+	W(\tau=t/\inttime) = \frac{2^p \, (p!)^2}{(2p)!} \left( 1 + \cos \pi \tau\right)^p
 \end{equation}
 with $p=2$ allows for more reliable determination of frequencies from strongly chaotic orbits.
-We refer to this slight modification of the algorithm as \superfreq\ and the code is open-source and publicly available on \project{GitHub} at \url{https://github.com/adrn/SuperFreq}.
+We refer to this slight modification of the algorithm as \superfreq\ \citep{superfreq} and the code is open-source and publicly available on \project{GitHub} at \url{https://github.com/adrn/SuperFreq}.
 For more details about the algorithm, see Appendix~\ref{sec:naffapdx} or \cite{laskar88, laskar93, papaphilippou96}.
@@
 \superfreq\ will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change location and amplitude with time.
-For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two separate sections of orbit.
+For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two consecutive sections of orbit.
 The frequencies picked out by \superfreq\ for such orbits will therefore represent the average periodic nature of the orbit over a given integration window.
@@
 \end{equation}
-where the upper index refers to which of the two consecutive sections of orbit and $N$ is the number of orbital periods in a single section of orbit.
+where the upper index refers to the two consecutive sections of orbit and $N$ is the number of orbital periods in a single section of orbit.
 By inverting this rate, we can compute the timescale over which we expect order-unity changes to the fundamental frequencies: the \emph{frequency diffusion time} is defined as
@@
 \end{equation}
-where the maximum is taken with respect to the corresponding amplitudes, $a_i$, of the fundamental frequency components \citep[see][]{valluri12}, and the time is in units of number of orbital periods, ${\rm T}$.
+where the maximum is taken with respect to the corresponding amplitudes, $a_k$, of the fundamental frequency components \citep[see][]{valluri12}, and the time is in units of number of orbital periods, ${\rm T}$.
 
@@
 We can estimate this timescale by multiplying the frequency diffusion time by a factor equal to the fractional spread in frequencies of the debris.
-For example, a globular cluster typically has $\approx$1\% spreads in fundamental frequencies, so by multiplying the frequency diffusion time by $f = 0.01$, we can estimate the time (number of orbital periods) over which we expect the frequencies to evolve by this amount.
+For example, a globular cluster typically has $\approx$1\% spreads in fundamental frequencies, so by multiplying the frequency diffusion time by $f = 0.01$, we can estimate the time (in number of orbital periods) over which we expect the frequencies to evolve by this amount.
 
@@
 
-We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy (per unit mass), $E$, chosen to span a range of distances comparable to the scale radius of the potential ($E=-(397.2~{\rm km~s}^{-1})^2$ in physical units; see Table~\ref{tbl:potential}).
+We generate an isoenergy grid of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy (per unit mass), $E$, chosen to span a range of distances comparable to the scale radius of the potential ($E=-(397.2~{\rm km~s}^{-1})^2$ in physical units; see Table~\ref{tbl:potential}).
 We fix $v_x = v_z = 0$, and compute $v_y$ from the energy.
 The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic.
-Tube orbits are generally less stochastic than box orbits, which tend to plunge deep into the inner regions of the potential where the density profile is steep, however there are appreciable regions of stochastic tube orbits.
-Thin tidal streams are thought to form along tube orbits rather than box orbits because of the faster disruption expected for stellar systems on radially plunging orbits.
+Tube orbits are generally less stochastic than box orbits, which tend to plunge deep into the inner regions of the potential where the density profile is steep, however there are appreciable numbers of stochastic tube orbits.
 The major classes of stable tube orbits circulate about either the major or minor axis.
 This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits), but the most numerous orbits in this grid are the short-axis and outer long-axis tubes.
+Thin tidal streams are thought to form along tube orbits rather than box orbits because of the faster disruption expected for stellar systems on radially plunging orbits.
 
@@
 We start by considering three particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular orbit (R), a weakly chaotic orbit (W), and a strongly chaotic orbit (S).
-The orbits were chosen at random from a single column of the orbit grid based on their frequency diffusion time, and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}.
+The orbits were chosen from a single column of the orbit grid based on their frequency diffusion time to sample a range of possible values, and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}.
 We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on these three orbits.
@@
 		{\bf ID} & $\bs{x}$ & $\bs{y}$ & $\bs{z}$ & $\bs{v_x}$ & $\bs{v_y}$ & $\bs{v_z}$ & $\bs{t_\lambda}$ & $\bs{t_\Omega}$ \\\toprule
-		R & 24.90 & 0 & 6.90 & 0 & 190.76 & 0 & XX & YY\\
+		R & 29.10 & 0 & 5.10 & 0 & 168.07 & 0 & $>10^3$ & $>10^8$\\
 		\midrule
-		W & 24.90 & 0 & 19.70 & 0 & 137.19 & 0 & XX & YY\\
+		W & 29.10 & 0 & 17.70 & 0 & 125.49 & 0 & $\approx55$ & $\approx8 \times 10^4$\\
 		\midrule
-		S & 24.90 & 0 & 26.10 & 0 & 93.11 & 0 & XX & YY\\
+		S & 29.10 & 0 & 22.10 & 0 & 99.45 & 0 & $\approx3.5$ & $\approx6 \times 10^3$\\
 		\bottomrule
@@
 Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$, and times ($t_\lambda$, $t_\Omega$) in number of orbital periods.
+Recall that the frequency diffusion time, $t_\Omega$, is the time over which we expect order-unity changes in the fundamental frequencies, hence why the timescales appear quite long.
 \label{tbl:three-orbits}}
@@
 \end{align}
-where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the debris, and $\|\cdot \|$ is the Euclidean norm.
+where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the ``progenitor,'' and $\|\cdot \|$ is the Euclidean norm.
 We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales.
-Figure~\ref{fig:ensembles} shows final positions of test-particle0 ensembles initialized around the three orbits described above and evolved for 16 orbital periods (same as the N-body simulations); the morphology of the ensembles is visually similar to the ``oldest'' (first stripped) debris in the N-body simulations.
+Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the three orbits described above and evolved for 16 orbital periods (same as the N-body simulations); the morphology of the ensembles is visually similar to the ``oldest'' (first stripped) debris in the N-body simulations.
 
@@
 
+\todo{apw}{These paragraphs need work, but waiting for results from the cluster...}
 The density distributions do not change shape dramatically, even for the strongly chaotic ensemble; we therefore use the mean density as a metric for studying the density evolution of orbit ensembles.
@@
 
-Motivated by the noticeable discrepancy between the regular and weakly chaotic orbit ensembles --- even though the chaotic timescale of the weakly chaotic ensemble parent orbit is thousands of orbital periods ---we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
+Motivated by the noticeable discrepancy between the regular and weakly chaotic orbit ensembles --- even though the chaotic timescale of the weakly chaotic ensemble parent orbit is thousands of orbital periods --- we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
 For all ensembles, we integrate the orbits for 16 (parent) orbital periods and store the mean density at the end of integration for each.
@@
 diff time map -- even though freq.
-diff time are long, still resolve structure in these maps.}
+diff time are long, can still resolve structure in these maps.}
 
@@
 The frequency diffusion times shown in these grids were computed with two consecutive windows of length 40 orbital periods.
-The resonant structure around the mildly chaotic orbit (white regions in panel W) bound the orbit to a small region, whereas for the strongly chaotic orbit (S), the entire region surrounding the orbit is strongly chaotic.
-To resolve the short-time behavior of the frequency diffusion (corresponding to motion across resonance layers) we compute the frequencies in a series of rolling windows along each orbit.
-We use a window with a width equal to 40 orbital periods and shift the window by half an orbital period between each calculation of the most significant frequencies --- this is the smallest window size we could choose before \superfreq\ begins to consistently fail to recover the frequencies for the strongly chaotic orbit.
-Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time --- plotted are the percent deviations of the frequency ratios from the median value over all windows.
+The resonant structure around the mildly chaotic orbit (white lines in panel W) bound the orbit to a small region, whereas for the strongly chaotic orbit (S), the entire region surrounding the orbit is strongly chaotic.
+To resolve the short-time behavior of the frequency diffusion (corresponding to motion across or around resonance layers) we compute the frequencies in a series of rolling windows along each orbit.
+We use a window with a width equal to 40 orbital periods and shift the window by an orbital period between each calculation of the most significant frequencies --- again, this is the smallest window size we could choose before \superfreq\ begins to occasionally fail to recover the frequencies for the strongly chaotic orbit.
+Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time --- plotted are the percent deviations of the frequency ratios from the initial value.
 The difference between the first window (green +) and the last window (red x) is 50 orbital periods.
@@
 This is comparable to the frequency spread in globular cluster tidal debris (0.1--0.5\%).
+%\todo{apw}{check numbers in previous 2 sentences}
 
@@
 
-We repeat the above analysis of the short-time evolution of frequencies for all orbits in the orbit grid used above (Section~\ref{results1}): for each orbit, we integrate for 80 orbital periods, but compute the frequencies now in a rolling window along this integration time.
+We repeat the above analysis of the short-time evolution of frequencies for all orbits in the orbit grid used above (Section~\ref{sec:results1}): for each orbit, we integrate for 80 orbital periods, but compute the frequencies now in a rolling window along this integration time.
 We fix the window width to 40 orbital periods and shift the window by 0.5 orbital periods between each computation of the frequencies.
-For each orbit, we have the three frequencies (and amplitudes), $k=1,2,3$, computed at times $j$; we compute the fractional amplitude of the frequency variation for the $i$th orbit as
+For each orbit, we have the three frequencies (and amplitudes), $k=1,2,3$, computed at times $j$.
+We compute the fractional amplitude of the frequency variation for a given orbit as
 \begin{equation}
-	? = \max_{a_k} \left(\frac{\max_j \Omega_{ik} - \min_j \Omega_{ik}}{2 \, \mean{\Omega_{ik}}_j}\right) \label{eq:frac-freq-amp}
+	A_\Omega = \max_{a_k} \left(\frac{\max_j \Omega^{(j)}_{k} - \min_j \Omega^{(j)}_{k}}{2 \, \mean{\Omega^{(j)}_{k}}_j}\right) \label{eq:frac-freq-amp}
 \end{equation}
@@
 For tidal debris with small spreads in orbital properties, nearby orbits will have similar behavior; for example, the central marker in Figure~\ref{fig:fdiff-zooms} shows the size of a globular-cluster-like system in configuration space.
-When the magnitude of the short-time frequency variability is comparable to the spread in orbital frequencies within the debris, the debris effectively has a larger spread in frequencies (the effective frequency distribution is like the convolution of the initial spread with a distribution that scales with the scale-length of the short-time variation).
+When the magnitude of the short-time frequency variability is comparable to the spread in orbital frequencies within the debris, the debris effectively has a larger spread in frequencies (the effective frequency distribution is the convolution of the initial spread with a distribution that scales with the scale-length of the short-time variation).
 Therefore, the debris will ``fan'' or spread out over a few dynamical times and appear dynamically hotter.
 
-Figure~\ref{fig:XX} shows the grid of initial conditions, now colored by the (fractional) amplitude of the short-time frequency variation (Equation~\ref{eq:frac-freq-amp}).
-When the fractional excursions of the frequencies are of order $10^{-3}$--$10^{-2}$ or larger, globular cluster debris will appear ``fluffier.'' Comparing this to the timescales of Figure~\ref{} and Figure~\ref{}, it is clear that by analyzing the small-scale variations of the frequencies of a progenitor orbit over a given window of time provides a better prediction of the resulting morphology of tidal debris.
+\todo{apw}{new figure and paragraph repeating this for an ensemble of orbits, with frequencies for each ensemble orbit computed in the windows.
+show that final distribution of frequencies is spread more for weakly chaotic orbit}
+
+Figure~\ref{fig:freqvar_map} shows the grid of initial conditions, now colored by the (fractional) amplitude of the short-time frequency variation, $A_\Omega$ (Equation~\ref{eq:frac-freq-amp}).
+When the fractional excursions of the frequencies are of order $10^{-3}$--$10^{-2}$ or larger, globular cluster debris will appear more diffuse due to enhanced density evolution from fast, small-scale chaotic diffusion.
+Comparing this to the timescales of Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}, it is clear that the small-scale variations of the frequencies of a progenitor orbit over a given window of time provides a better prediction of the resulting morphology of tidal debris.
 
@@
 
+\subsection{The progenitor mass scale}
+
+In this work, we have only considered low-mass progenitor systems such as globular clusters because the intrinsic spreads in fundamental frequencies are small (0.1--0.5\%).
+Small changes to the frequencies of the orbits of tidal debris disrupted from these progenitors due to chaotic processes will therefore cause observable changes to the real-space morphology of the debris.
+For more massive progenitor systems, the typical size and velocity dispersion of the debris will be larger and thus the debris morphology will be less sensitive to small changes in orbital properties.
+The mass-scale of the debris that will display enhanced density evolution depends on the magnitude of weak chaos, which depends on the orbit structure of a given potential.
+In potentials with more significant chaos, debris disrupted from more massive progenitor systems may also display stream-fanning.
+
+The mass of the progenitor sets the physical scale of the debris as it is disrupted.
+For more massive progenitors and therefore larger spreads in debris, different orbits may have different chaotic diffusion timescales that depend on the details of the disruption (e.g., the orbital phase of the pericenter, or the energy offset of the debris).
+This could potentially lead to interesting density variations along the tidal debris.
+
 \subsection{Ophiuchus stream}
@@
 Tidal stream modeling is one of the most promising ways to constrain the 3D mass distribution around the Milky Way at distances of 10s to 100s of kpc.
-Methods that use tidal streams to infer properties of the Galactic potential typically operate by constructing models of the debris distribution using either the present-day phase-space density \citep[e.g.,][]{kuepper12, kuepper15}, time-of-disruption phase-space density \citep{apw14}, or the density in angle-action coordinates \citep[e.g.,][]{sanders14, bovy14}.
-All of these methods may fail for modeling globular-cluster streams on mildly chaotic orbits.
+Methods that use tidal streams to infer properties of the Galactic potential typically operate by constructing models of the debris distribution using either the present-day phase-space density \citep[e.g.,][]{kuepper12, kuepper15, amorisco and others}, time-of-disruption phase-space density \citep{apw13, apw14}, or the density in angle-action coordinates \citep{sanders14, bovy14}.
+All of these methods may fail or produce uninterpretable results if modeling globular-cluster streams on mildly chaotic orbits.
 
@@
 The high-density part of the stream may still appear thin (e.g., the recently disrupted debris in the middle panels of Figure~\ref{fig:nbodysims}).
-If the stream modeling focuses on reproducing the present-day density of the stream and the potential is restricted to an integrable, analytic form, the debris may be well fit by a particular orbit and evolution time in such a potential.
-However, the inferences about the potential or mass distribution may be biased significantly and the inferences about the evolution time and internal properties of the system will certainly be incorrect.
+If the stream modeling focuses on reproducing the present-day density of the stream and the potential is restricted to an integrable, analytic form, the phase-space distribution of the debris may be well-reproduced by a progenitor on a particular orbit in such a potential.
+However, the inferences about the potential or mass distribution may be biased significantly and the inferences about the evolution time and internal properties of the system will certainly be incorrect --- for example, the evolution time may be biased to shorter times to reproduce the apparent shortness of the stream.
+With realistic observational uncertainties, it is likely the case that for any stream formed from a progenitor on a weakly chaotic orbit, there is a regular progenitor orbit in some other potential that produces a model stream equally consistent with the data.
 
-[what happens in angle-action coordinates?]
+Due to the often unknown or poorly understood selection biases involved in kinematic observations of stream stars (due to, e.g., survey limits, dust extinction, and stellar population variations), modeling the final phase-space densities of tidal streams will lead to uninterpretable inferences about the potential of the host galaxy.
+We argue that by instead performing the density modeling at the time of disruption \citep[e.g.,][]{apw14}, uncertainties in, for example, the length or width of the stream will not bias potential parameter inferences as significantly (see Price-Whelan et al.
+2015, in prep.).
+In either case, it is critical that potential modeling efforts with tidal streams allow enough freedom in the potential to explore all possible configurations, though this quickly becomes a high-dimensional, multi-modal, and therefore computationally prohibitive inference problem.
 
-[what about opposite effect -- if potential model too flexible, can always find chaotic orbit to reproduce regular orbit?]
+\todo{apw}{Need more coherent words about what happens when trying to use angle-action coordinates...}.
 
@@
 Passage through a thin disk results in a nearly instantaneous, impulsive forcing of the orbit at a frequency set by the circulation frequency of the orbit.
-This may act to increase the significance of chaos along disk-crossing orbits \citep[e.g.,][]{kandrup00}.
+This may act to increase the significance of chaos along disk-crossing orbits \citep[e.g.,][]{kandrup00, hunter05}.
 This effect is likely unresolved in many simulations where the force and spatial resolution is insufficient to capture this effect.
@@
 
-We have shown that chaos may have a significant effect on the density evolution of tidal debris.
+The large-scale shapes of galactic gravitational potentials are thought to be triaxial and therefore contain an appreciable number of chaotic orbits.
+Chaos is not thought to be dynamically important for shaping the global structure of galaxies because the predicted chaotic timescales such as the Lyapunov or frequency diffusion times for orbits in galactic-like potentials are very long (many 1000s of orbital periods).
+However, these timescales measure the mean evolution of orbital properties such as the fundamental frequencies and do not capture the full nature of chaotic orbit diffusion.
+
+We have numerically computed the magnitude of chaos for a large grid of iso-energy orbits in a triaxial gravitational potential using two independent methods that have been used extensively to classify and characterize chaotic orbits: the Lyapunov exponent and frequency diffusion rate.
+From each of these indicators, we compute a timescale over which chaos is likely to be important and find that the majority of orbits have chaotic timescales greater than 100s of orbital periods, however with the frequency diffusion rate we are still able to resolve weak chaos.
+We then study the density evolution of small ensembles of orbits generated around each orbit in the grid used for the previous experiment and find that along some orbits classified as weakly chaotic --- with chaotic timescales of 100s of orbital periods --- the orbit ensembles display enhanced density evolution and reach a lower overall density faster than orbit ensembles around nearby regular orbits.
+We propose an explanation for this apparent discrepancy by considering the nature of chaotic diffusion: the classical chaos indicators are most sensitive to the  slow, Arnold diffusion process that can cause large changes to orbital properties, but small-scale frequency evolution occurs over much shorter times as chaotic orbits stochastically diffuse across the stochastic layers that surround many resonances.
+We propose a new measure of chaos with tidal stream morphological evolution in mind: the amplitude of small-scale, fractional changes in frequencies over a given number of orbital periods along a stream progenitor orbit.
+When this amplitude is comparable to or larger than the typical spread in orbital properties of tidal debris from the progenitor, the phase-space density of the debris will evolve faster to a state of lower density relative to nearby regular orbits.
+
+\todo{apw}{Concluding paragraph that drives home the importance of this...}
+The timescales in this paper are not meant to be interpreted physically, as the potential is a very crude approximation to the total potential of a Milky-Way-like galaxy.
+The key point is that even when the mean chaos indicators predict a long chaotic timescale, streams may display enhanced density evolution over much shorter times.
+Tidal debris is sensitive to the small-scale, short-time evolution of orbital properties (e.g., the fundamental frequencies) induced by weak chaos.
 
@@
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
-This research made use of \project{Astropy}, a community-developed core \texttt{Python} package for Astronomy \citep{astropy13}.
+This research made use of \project{Astropy}, a community-developed core \project{Python} package for Astronomy \citep{astropy13}.
 This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance.
@@
 \section{Lyapunov exponents} \label{sec:lyapapdx}
+\todo{apw}{This section has not been edited carefully...}
 
@@
 
-\todo{apw}{do we need this test?}
+\todo{apw}{This section has not been edited carefully...}
+
 Figure~\ref{fig:logfreqs} shows a test of the implementation in which we reproduce the frequency map at a fixed energy of an axisymmetric, logarithmic potential \cite[][pg.
@@
 
-NAFF recovers the fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL96;][]{papaphilippou96}.
+\superfreq\ recovers the fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL96;][]{papaphilippou96}.
 PL96 show that a good choice of coordinates for tube orbits are the Poincar\'e symplectic polar coordinates, a set of canonical coordinates similar to cylindrical coordinates.
-When computing the frequencies for tube orbits, we first align the circulation about the $z$-axis through rotation, transform to Poincar\'e polar coordinates, then use NAFF to measure the fundamental frequencies.
+When computing the frequencies for tube orbits, we first align the circulation about the $z$-axis through rotation, transform to Poincar\'e polar coordinates, then use \superfreq\ to measure the fundamental frequencies.
 We could equivalently use the Cartesian time series, but the convergence of terms is slower (the amplitudes of successive terms decrease slower for Cartesian coordinates).
-We have tested that our implementation of NAFF returns the same fundamental frequencies in either case for a set of tube orbits.
+We have tested that our implementation of \superfreq\ returns the same fundamental frequencies in either case for a set of tube orbits.
 For box orbits, the motion is close to separable in each Cartesian component and we therefore use Cartesian coordinates for estimating the frequencies for these orbits.
-[TODO: Demo figure of log potential???
-convince our implementation is good?]
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/freqvar-map.png}
-\caption{ } 
+\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/freqvar-map.png}
+\caption{The same grid of initial conditions as in previous figures, but now each pixel is colored by the amplitude of short-time frequency variations, $A_\Omega$ (Equation~\ref{eq:frac-freq-amp}).
+The amplitude is a fractional measurement, so that a value of $10^{-2}$ corresponds to a 1\% change in frequencies.
+To compute the short-time frequency evolution, the frequencies are computed in a sliding window of width 40 orbital periods, and the window is shifted by half of an orbital period between each computation of the frequencies.
+The frequency evolution is computed for the equivalent of 16 orbital periods.
+The weakly chaotic orbit has $A_\Omega = XX$, corresponding to YY\% variations in the frequencies over a few orbital periods and hence the observed enhanced density evolution.
+The three orbits, R, W, S, are highlighted with square markers (S is top, W is middle, R is bottom).
+}
 \label{fig:freqvar_map}
@@
 
-% Figure ??
-\clearpage
-\begin{figure*}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/freqvar-map-stretched.png}
-\caption{ } 
-\label{fig:freqvar_map_stretched}
-\end{center}
-\end{figure*}
-
 \end{document}

### 2015-06-22 ms.tex
added todos

@@
 \begin{abstract}
+
+\todo{apw}{write the abstract...}
 
@@
 
-\todo{apw}{These paragraphs need work, but waiting for results from the cluster...}
+\todo{apw}{These paragraphs need work, but still waiting for numerical results from the cluster...}
 The density distributions do not change shape dramatically, even for the strongly chaotic ensemble; we therefore use the mean density as a metric for studying the density evolution of orbit ensembles.
@@
 \todo{apw}{new figure and paragraph repeating this for an ensemble of orbits, with frequencies for each ensemble orbit computed in the windows.
-show that final distribution of frequencies is spread more for weakly chaotic orbit}
+show that final distribution of frequencies is spread more for weakly chaotic orbit (still waiting for queued code to finish on the cluster...)}
 
@@
 
-\todo{apw}{Concluding paragraph that drives home the importance of this...}
+\todo{apw}{Concluding paragraph that drives home the importance of this work...}
 The timescales in this paper are not meant to be interpreted physically, as the potential is a very crude approximation to the total potential of a Milky-Way-like galaxy.
@@
 \begin{center}
-\includegraphics[width=\textwidth, trim={0 1cm 0 0}]{figures/lyap_map.png}
+\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/lyap_map.png}
 \caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
@@
 Chaotic orbits with $t_\lambda \gtrsim 1000~\periods$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
-Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S), weakly chaotic (W), and regular (R) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:lyapmap}
+Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S), weakly chaotic (W), and regular (R) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).
+\todo{apw}{Need better markers for the 3 orbits -- don't show up in greyscale (printed)}} \label{fig:lyapmap}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth, trim={0 1cm 0 0}]{figures/fdiff_map.png}
+\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/fdiff_map.png}
 \caption{The same grid of initial conditions as Figure~\ref{fig:lyapmap}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
@@
 The particle positions qualitatively match the morphology of the ``oldest'' debris from the N-body simulations, though there are many more particles in these ensembles than are disrupted during the first pericentric passage of the simulations.
-} \label{fig:ensembles}
+}
+\label{fig:ensembles}
 \end{center}
@@
 The density of the weakly chaotic ensemble initially evolves faster to lower density and has a larger tail to lower densities.
-The strongly chaotic ensemble evolves fast to a state of low density.} \label{fig:ensemble-density}
+The strongly chaotic ensemble evolves fast to a state of low density.
+\todo{apw}{add border to histograms?
+get rid of colors (maybe just light to dark grey?)}} \label{fig:ensemble-density}
 \end{center}
@@
 \includegraphics[width=\textwidth]{figures/ensemble-densities.png}
-\caption{This is an alternate visualization of the density evolution of orbit ensembles around the three orbits of Figure~\ref{fig:ensemble-density}, showing only the evolution of the mean density for the orbits (normalized by the initial density of the ensemble).}
+\caption{This is an alternate visualization of the density evolution of orbit ensembles around the three orbits of Figure~\ref{fig:ensemble-density}, showing only the evolution of the mean density for the orbits (normalized by the initial density of the ensemble).
+\todo{apw}{change linestyles so this survives being printed (-, --, -.)}}
 \label{fig:densities}
@@
 \begin{center}
-\includegraphics[width=\textwidth, trim={0 1cm 0 0}]{figures/ensemble-map.png}
-\caption{The same grid of initial conditions as Figure~\ref{fig:lyapmap} and other previous figures, but now we consider the density evolution of small ensembles of orbits around each of the orbits in this grid.
+\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/ensemble-map.png}
+\caption{\todo{all}{Ignore this for now!
+Waiting for results from latest run on cluster...}The same grid of initial conditions as Figure~\ref{fig:lyapmap} and other previous figures, but now we consider the density evolution of small ensembles of orbits around each of the orbits in this grid.
 We generate globular-cluster-scale ensembles of orbits around each ``parent'' orbit in the grid (Section~\ref{sec:results2}) and follow the density of the ensemble as a function of time using kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
-Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S), weakly chaotic (W), and regular (R) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).}
+Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S), weakly chaotic (W), and regular (R) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).
+\todo{apw}{remake once new run on the cluster is complete}}
 \label{fig:ensemblemap-meandensity}
@@
 The regular orbit (R) is near a resonance (white line), the weakly chaotic orbit (W) is surrounded by an intricate resonant structure and may be in a region of resonance overlap, and the strongly chaotic orbit (S) is embedded in a region of strong stochasticity.
-}
+\todo{apw}{Move colorbar to bottom or top?}}
 \label{fig:fdiff-zooms}
@@
 The frequency evolution is computed for the equivalent of 16 orbital periods.
-The weakly chaotic orbit has $A_\Omega = XX$, corresponding to YY\% variations in the frequencies over a few orbital periods and hence the observed enhanced density evolution.
+The weakly chaotic orbit has $A_\Omega \approx 0.2\%$ over just 16 orbital periods --- this explains the enhanced density evolution observed in the N-body simulations and ensemble integrations (Figures~\ref{fig:nbodysims}-\ref{fig:ensembles}).
 The three orbits, R, W, S, are highlighted with square markers (S is top, W is middle, R is bottom).

### 2015-06-22 ms.tex
added new figures

@@
 \begin{center}
+\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-regular.png}
+\caption{Evolution of the fundamental frequency ratios, $r_{1,3} = \Omega_1 / \Omega_3$ and $r_{2,3} = \Omega_2 / \Omega_3$, over 16 orbital periods for an ensemble of 1000 orbits generated around the regular orbit, R, with mass scale $m=10^4~\msun$ (Section~\ref{sec:results2}).
+Fundamental frequencies are computed with a window width of $\approx$40 orbital periods, and the window is shifted by one orbital period between each computation (each grey point represents the fundamental frequency ratios computed in a single window).
+Plotted are the per cent deviations of the frequency ratios from the value in the initial window for the parent orbit.
+Left panel shows frequencies of each ensemble orbit for first time window.
+Middle panel shows the evolution of the frequencies over 16 orbital periods.
+Right panel shows final frequency values.
+For this orbit (R), all panels are identical because the orbit is regular (the frequencies don't evolve).}
+\label{fig:regular-ensemble-freq-evolution}
+\end{center}
+\end{figure*}
+
+% Figure ??
+\clearpage
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-mildly_chaotic.png}
+\caption{Same as Figure~\ref{fig:regular-ensemble-freq-evolution}, but for the weakly chaotic orbit (W).
+Frequencies now display short-time variations (compare left to right panels).
+Points in middle panel show the trajectories of each ensemble orbit in frequency-space.}
+\label{fig:mildly_chaotic-ensemble-freq-evolution}
+\end{center}
+\end{figure*}
+
+% Figure ??
+\clearpage
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-chaotic.png}
+\caption{Same as Figures~\ref{fig:regular-ensemble-freq-evolution}--\ref{fig:mildly_chaotic-ensemble-freq-evolution}, but for the strongly chaotic orbit (S).
+Frequencies are now explore a huge volume representing $>$100\% deviations from the initial value of the parent orbit.
+Frequencies evolve quickly around resonances --- especially the $\bs{n}=(2,-3,0)$, $\bs{n}=(1,-1,0)$, and $\bs{n}=(1,0,1)$ resonances --- which appear as linear features in these panels.}
+\label{fig:chaotic-ensemble-freq-evolution}
+\end{center}
+\end{figure*}
+
+% Figure ??
+\clearpage
+\begin{figure*}[p]
+\begin{center}
 \includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/freqvar-map.png}

### 2015-06-22 ms.tex
words about new figures

@@
 This is a measure of the magnitude of the short-time frequency evolution for a single orbit.
-For tidal debris with small spreads in orbital properties, nearby orbits will have similar behavior; for example, the central marker in Figure~\ref{fig:fdiff-zooms} shows the size of a globular-cluster-like system in configuration space.
-When the magnitude of the short-time frequency variability is comparable to the spread in orbital frequencies within the debris, the debris effectively has a larger spread in frequencies (the effective frequency distribution is the convolution of the initial spread with a distribution that scales with the scale-length of the short-time variation).
-Therefore, the debris will ``fan'' or spread out over a few dynamical times and appear dynamically hotter.
+For tidal debris with small spreads in orbital properties, we expect nearby orbits to have similar behavior; for example, the central marker in Figure~\ref{fig:fdiff-zooms} shows the size of a globular-cluster-like system in configuration space.
+We therefore expect a small ensemble of orbits in frequency space to expand over short times around even weakly chaotic parent orbits and the debris will therefore appear dynamically hotter in real-space.
 
-\todo{apw}{new figure and paragraph repeating this for an ensemble of orbits, with frequencies for each ensemble orbit computed in the windows.
-show that final distribution of frequencies is spread more for weakly chaotic orbit (still waiting for queued code to finish on the cluster...)}
+We illustrate this phenomenon by repeating the above experiment for small ensembles of orbits around each parent orbit.
+As before, we compute the frequencies of each orbit in a series of overlapping windows to resolve the short-time evolution of the frequencies.
+Figures~\ref{fig:regular-ensemble-freq-evolution}--\ref{fig:chaotic-ensemble-freq-evolution} show the evolution of the frequencies for all orbits in each ensemble generated around the three orbits, R, W, and S.
+In all panels in each of these figures, whats plotted are the per cent deviations of each frequency ratio from the parent orbit frequency ratios computed in the first time window.
+The left panels in each figure show these per cent deviations for all ensemble orbits around each of the three orbits.
+Middle panels show the full evolution of each ensemble orbit over 16 orbital periods, and final panel shows the final per cent deviations for all ensemble orbits (after 16 orbital periods).
+Note that the axis scales are quite different between each figure --- the strongly chaotic orbit ensemble evolves quickly due to its proximity to many overlapping resonances.
 

### 2015-06-24 ms.tex
introduction take 2

@@
 As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded due to the tidal forces of the host galaxy potential.
-If the orbit of such an object is regular, only mildly eccentric, and stays far from sharp density features (e.g., the disk), the disruption is a steady process and the mass is slowly stretched into long streams of matter known as stellar tidal streams.
-The phase-space density and therefore the observable properties of these tidal streams are sensitive to the internal progenitor properties, orbit of the progenitor, and mass distribution of the host galaxy.
-Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies that span a large range of distances \citep[$\approx$10--100~kpc][]{cite many}.
-The morphological evolution of tidal debris depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also determined by the shape and radial profile of the gravitational potential of the host galaxy.
-By simultaneously modeling the observed phase-space density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution around the Milky Way.
+Along regular, mildly-eccentric orbits, mass is disrupted with small spreads in orbital properties (e.g., energy, angular momentum).
+Once the debris has evolved far enough from the progenitor system that the self-gravity of the progenitor can be ignored (usually a fast process relative to the orbital time), the stars evolve essentially as an ensemble of test particle orbits in the potential of the host system.
+The debris remains coherent as a tidal stream if the phase-mixing time-scale is long: a small ensemble of regular orbits reaches a fully phase-mixed state in a timescale $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in fundamental frequencies of the ensemble.
+The ensemble spreads due to sheering from slight variations in their fundamental frequencies which, for tidal streams, preferentially occurs along one dimension \citep{merritt96, helmi99}.
 
-Tidal streams are generally modeled as 1D structures formed on regular orbits with simple internal progenitor dynamics and density evolution.
-Methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to N-body simulations \citep[see, e.g., the introduction of][]{kuepper15}.
+The morphological (density) evolution of the tidal debris therefore depends only on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also determined by the shape and radial profile of the gravitational potential of the host galaxy.
+By modeling the observed phase-space density of stream stars along with the host galaxy potential it is hoped that we may infer the 3D mass distribution of the host.
+Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies; the known streams span a large range of distances --- from $\approx10$ to $100~kpc$ --- and progenitor masses --- from $\approx$10$^3$ to $\approx$10$^8~\msun$ in stellar mass --- \citep[][]{cite many}.
+There has been extensive work on developing methods to use data from these streams to measure properties of Milky Way's dark matter halo.
+These methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to N-body simulations \citep[see, e.g., the introduction of][]{kuepper15}.
 All methods have been tested in some way on simulated observations of data and these tests typically demonstrate the recovery of parameters for analytic, static potential forms.
-Many of the methods have been shown to work in axisymmetric or otherwise simple potentials but may extend to more complex, generalized potentials (e.g., triaxial or multi-component potentials).
+\todo{apw}{This is all just about the methods -- do I need sentences or a paragraph on results?}
+
+%At present most of these methods have been shown to work in axisymmetric or otherwise simple potentials through tests with mock data.
 
@@
 In the oblate potential (a three-component bulge+disk+spherical halo potential), a thin model stream is easily found that reproduces the observed stellar density morphology of the stream.
-In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness of the stream given the observational constraints of the present-day position and velocity of the cluster.
+In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness and curvature of the stream given the observational constraints of the present-day position and velocity of the cluster.
 
@@
 It is well known that the number of degrees of freedom of a potential plays a critical role in determining the orbit structure of the potential --- Hamiltonians with more than two degrees of freedom generically contain significant chaotic regions.
-\citet{pearson15} tested the stochasticity of the orbit of the progenitor that produced fanned debris by computing the Lyapunov exponent along this orbit but found that it is consistent with being regular over dynamically relevant timescales (many Hubble times).
-It has been shown previously that along some strongly chaotic orbits, tidal streams do form large, diffuse fans of debris \citep[e.g.,][]{fardal14}, however it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
+\citet{pearson15} tested the stochasticity of the orbit of the progenitor that produced `fanned' debris by computing the Lyapunov exponent along this orbit but found that it is consistent with being regular over dynamically relevant timescales (many Hubble times).
+It has been shown previously that along some strongly chaotic orbits, tidal streams do form large, diffuse `fans' of debris \citep[e.g.,][]{fardal14, ngan15}, however it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
 The result from \citet{pearson15} suggests that even weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.
 
-In this work, we study the effect of chaotic diffusion on tidal stream morphology.
+In this work, we study the effect of chaotic diffusion of individual orbits on the density evolution of tidal debris.
 We choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos as computed from single-orbit diagnostics for grids of constant-energy orbits, and compare these results to measures of the density evolution of finite-volume ensembles of orbits (meant to mimic tidal debris).
-We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles due to the nature of chaotic diffusion.
+We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles due to the chaotic diffusion of the individual orbits.
+The diffusion of the fundamental frequencies of the orbits introduces an extra, time-dependent dispersion to the ensembles and therefore reduces the timescale over which mixing becomes relevant (e.g., $\sigma_\Omega$ can grow with time).
 This result supports a reevaluation of the importance of chaos in galactic halos and implies that the amount of chaos in a given potential may have significant consequences for the observability and survivability of thin, cold tidal streams.

### 2015-06-24 ms.tex
remove ticks

@@
 All methods have been tested in some way on simulated observations of data and these tests typically demonstrate the recovery of parameters for analytic, static potential forms.
-\todo{apw}{This is all just about the methods -- do I need sentences or a paragraph on results?}
+\todo{readers}{This is all just about the methods -- do I need sentences or a paragraph on results?}
 
@@
 In the oblate potential (a three-component bulge+disk+spherical halo potential), a thin model stream is easily found that reproduces the observed stellar density morphology of the stream.
-In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness and curvature of the stream given the observational constraints of the present-day position and velocity of the cluster.
+In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically form large, two-dimensional `fans' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness and curvature of the stream given the observational constraints of the present-day position and velocity of the cluster.
 
@@
 Galactic potentials are almost certainly not globally integrable, but it is useful to understand the orbit structure in integrable systems before extending to more general potentials.
-Only one globally integrable triaxial mass model is known that generates motion that is qualitatively similar to that seen in Galaxies: the St\"ackel or ``Perfect Ellipsoid'' model \citep[e.g.,][]{kuzmin73, deZeeuw85}.
+Only one globally integrable triaxial mass model is known that generates motion that is qualitatively similar to that seen in Galaxies: the St\"ackel or `Perfect Ellipsoid' model \citep[e.g.,][]{kuzmin73, deZeeuw85}.
 The four general classes of orbits in this potential are box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits.
@@
 where $H_0(\bs{J})$ represents an integrable Hamiltonian, and $\epsilon$ is a small parameter that determines the perturbation strength \citep[a description of perturbation theory applied to nonlinear Hamiltonians is given in][]{lichtenberg83}.
-When $0 < |\epsilon| \ll 1$, resonant surfaces become ``thick'' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99}.
+When $0 < |\epsilon| \ll 1$, resonant surfaces become `thick' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99}.
 These resonant layers are then generically surrounded by stochastic layers where chaotic motion occurs.
@@
 
-This method is referred to as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) and has been used extensively in planetary dynamics \citep[e.g.,][]{laskar93b, laskar96} and galaxy dynamics \citep{papaphilippou98, valluri98}, especially in the study of orbits in triaxial systems.
+This method is referred to as `Numerical Approximation of Fundamental Frequencies' (NAFF) and has been used extensively in planetary dynamics \citep[e.g.,][]{laskar93b, laskar96} and galaxy dynamics \citep{papaphilippou98, valluri98}, especially in the study of orbits in triaxial systems.
 We have implemented and tested a version of this procedure in the \project{Python} programming language.
@@
 Figure~\ref{fig:nbodysims} shows the final particle distributions rotated so that the angular momentum of the progenitor orbit is aligned with the $z$-axis.
-A thin stream forms on the regular orbit (left column), a more diffuse stream on the mildly chaotic orbit (middle column), and a ``fanned'' stream on the strongly chaotic orbit (right column).
+A thin stream forms on the regular orbit (left column), a more diffuse stream on the mildly chaotic orbit (middle column), and a `fanned' stream on the strongly chaotic orbit (right column).
 Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit, it is surprising that stream formed on this orbit appears to be more diffuse than the regular orbit stream.
@@
 We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
-For a given set of orbital initial conditions --- the ``parent'' orbit --- we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time.
+For a given set of orbital initial conditions --- the `parent' orbit --- we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time.
 The physical scale of the ensemble is set by the tidal radius in position and the velocity scale in velocity and are therefore set by the mass of the progenitor \citep[e.g.,][]{johnston98, apw14}.
@@
 \end{align}
-where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the ``progenitor,'' and $\|\cdot \|$ is the Euclidean norm.
+where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the `progenitor,' and $\|\cdot \|$ is the Euclidean norm.
 We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales.
-Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the three orbits described above and evolved for 16 orbital periods (same as the N-body simulations); the morphology of the ensembles is visually similar to the ``oldest'' (first stripped) debris in the N-body simulations.
+Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the three orbits described above and evolved for 16 orbital periods (same as the N-body simulations); the morphology of the ensembles is visually similar to the `oldest' (first stripped) debris in the N-body simulations.
 
@@
 We have considered here a simple triaxial gravitational potential chosen to mimic the median properties of dark matter halos formed in dark-matter-only simulations (or the large-scale properties of halos formed in simulations with baryonic effects).
-We have found that ``mean'' chaos indicators such as the Lyapunov exponent or frequency diffusion rate predict that chaos should be dynamically unimportant in the halos of galaxies for the majority of orbits (e.g., the predicted timescales are 100-1000s of times longer than the typical orbital times, at least in this simple potential).
+We have found that `mean' chaos indicators such as the Lyapunov exponent or frequency diffusion rate predict that chaos should be dynamically unimportant in the halos of galaxies for the majority of orbits (e.g., the predicted timescales are 100-1000s of times longer than the typical orbital times, at least in this simple potential).
 However, while the mean chaotic evolution of, for example, the fundamental frequencies of an orbit may be prohibitively long, the frequencies of even mildly chaotic orbits do evolve on a smaller scale over much shorter times.
@@
 At Galactocentric $R \approx 1$ kpc, $z \approx 5$ kpc, the orbits of the stream stars likely pass through the MW disk, feel the triaxiality of the Galactic bar \citep[e.g.,][]{wegg13, wegg15}, and have short orbital periods (relative to streams in the halo).
-It is possible that the observed debris is the last remnants of the recently disrupted progenitor system \citep{sesar15}, however if a significant number of stars were disrupted on previous pericentric passages, this older debris may be ``fanned'' and still near the observed portion of the stream.
+It is possible that the observed debris is the last remnants of the recently disrupted progenitor system \citep{sesar15}, however if a significant number of stars were disrupted on previous pericentric passages, this older debris may be `fanned' and still near the observed portion of the stream.
 The fanned debris would be much lower surface brightness and thus much more difficult to detect.
@@
 
-The potential considered in this work is ``unrealistic'' for the total potential of a Milky-Way-like galaxy in that it is static, smooth, and does not contain baryonic components.
+The potential considered in this work is `unrealistic' for the total potential of a Milky-Way-like galaxy in that it is static, smooth, and does not contain baryonic components.
 Hydrodynamic simulations of forming galactic disks in cosmological dark matter halos have shown that baryonic feedback and relaxation can significantly change the inner distribution of dark matter and either make the potential more spherical or oblate \citep{many}.
@@
 First, there are observational hints that the intrinsic alignment of galactic disks may be close to random \citep{??} (rather than fixed to symmetry planes of the host halo potential).
-Simulations since the work of D08 have shown that by ``freeing'' the disk component, the angular momentum vector of the disk can precess and wander erratically away or around the principal axes of the host halo \citep{debattista15}.
+Simulations since the work of D08 have shown that by `freeing' the disk component, the angular momentum vector of the disk can precess and wander erratically away or around the principal axes of the host halo \citep{debattista15}.
 It is unknown what effect this would have on the orbit structure of nearby halo orbits, but we naively expect that this would destabilize a significant number of orbits.
@@
 
-\superfreq\ recovers the fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are ``close'' to the angle variables \cite[PL96;][]{papaphilippou96}.
+\superfreq\ recovers the fundamental frequencies for an orbit faster (with a fewer number of terms) when the coordinates used are `close' to the angle variables \cite[PL96;][]{papaphilippou96}.
 PL96 show that a good choice of coordinates for tube orbits are the Poincar\'e symplectic polar coordinates, a set of canonical coordinates similar to cylindrical coordinates.
@@
 From left to right, the columns indicate that the progenitor is on a regular (R), weakly chaotic (W), or strongly chaotic (S) orbit.
-Even on the weakly chaotic orbit --- where the predicted chaotic timescale is XX orbital periods --- the resulting tidal stream displays stream-fanning and the ``oldest'' debris is more diffuse than in the regular case.} \label{fig:nbodysims}
+Even on the weakly chaotic orbit --- where the predicted chaotic timescale is XX orbital periods --- the resulting tidal stream displays stream-fanning and the `oldest' debris is more diffuse than in the regular case.} \label{fig:nbodysims}
 \end{center}
@@
 The ensembles each contain 1000 particles and begin at pericenter rather than apocenter because they are meant to represent debris disrupted during a single (the first) pericentric passage.
-The particle positions qualitatively match the morphology of the ``oldest'' debris from the N-body simulations, though there are many more particles in these ensembles than are disrupted during the first pericentric passage of the simulations.
+The particle positions qualitatively match the morphology of the `oldest' debris from the N-body simulations, though there are many more particles in these ensembles than are disrupted during the first pericentric passage of the simulations.
 }
@@
 Waiting for results from latest run on cluster...}The same grid of initial conditions as Figure~\ref{fig:lyapmap} and other previous figures, but now we consider the density evolution of small ensembles of orbits around each of the orbits in this grid.
-We generate globular-cluster-scale ensembles of orbits around each ``parent'' orbit in the grid (Section~\ref{sec:results2}) and follow the density of the ensemble as a function of time using kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
+We generate globular-cluster-scale ensembles of orbits around each `parent' orbit in the grid (Section~\ref{sec:results2}) and follow the density of the ensemble as a function of time using kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
 Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S), weakly chaotic (W), and regular (R) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).

### 2015-06-24 ms.tex
incorporating comments from kvj and sp

@@
 
-In order to explore the question of how chaos manifests over timescales much shorter than that predicted from generic chaos indicators, we must first understand the structure and behavior of orbits in complex gravitational potentials.
+To explore the question of if and how chaos manifests in the density evolution of orbit ensembles over timescales much shorter than that predicted from generic chaos indicators, we must first understand the behavior of individual orbits in complex gravitational potentials and the orbital structures in the potential itself (i.e.
+the strength of resonances and chaos).
 An orbit in an $N$ degree of freedom (dof) Hamiltonian, $H$, is a set of $2N$ quasi-periodic time series,
@@
 \end{align}
-for all $t$. If bounded, the motion in any component, $w_i(t)$, can be represented as a Fourier sum,
+for all $t$. If bounded, the motion in any component, $w_i(t)$, can be described as a Fourier sum,
 \begin{equation}
@@
 
-A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally --- for example, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position.
+A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally --- for example, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position \citep[e.g.,][]{goldstein80}.
 The condition for global integrability is very restrictive and it seems that the likelihood that a Hamiltonian is globally integrable decreases as the number of degrees of freedom increase \citep[e.g.,][]{lichtenberg83}.
@@
 
-The frequencies of a generic, non-resonant orbit are incommensurable --- that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} These conditionally-periodic orbits uniformly cover the surface of an orbital torus.
-If instead there exists a relation of the form $\boldsymbol{n} \cdot \boldsymbol{\Omega} = 0$ the orbit is a resonant orbit.
-For resonant orbits, the existence of an additional \emph{local} constant of motion reduces the dimensionality of the orbit.
-We refer to orbits that obey a single resonance relation as \emph{uni-resonant} orbits, and if an additional resonance relation exists, \emph{bi-resonant}.
+The frequencies of a generic orbit are typically incommensurable --- that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} These conditionally-periodic orbits uniformly cover the surface of an orbital torus.
+If instead there exists a relation of the form $\boldsymbol{n} \cdot \boldsymbol{\Omega} = 0$ the orbit is referred to as a resonant orbit.
+Resonant orbits are confined to a surface with lower dimensionality than the surface of an orbital torus, depending on the number of resonance relations obeyed: we refer to orbits that obey a single resonance relation as \emph{uni-resonant} orbits, and if an additional resonance relation exists, \emph{bi-resonant}.
+Uni-resonant orbits in a triaxial potential are confined to a 2D surface, and bi-resonant orbits are closed 1D curves.
 The resonant structure of a potential --- the relative importance of particular resonance numbers --- is difficult to compute, but determines the global behavior of orbits in the potential.
@@
 In plots of frequency ratios, the resonances appear as lines; Figure~\ref{fig:cartoons} (left panel) shows a cartoon portrait of a portion of frequency-space for an integrable potential with example resonance lines, non-resonant, and resonant orbits marked.
-For an integrable potential, all non-resonant orbits are regular.
+For an integrable potential, all orbits are regular.
 
@@
 When $0 < |\epsilon| \ll 1$, resonant surfaces become `thick' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99}.
-These resonant layers are then generically surrounded by stochastic layers where chaotic motion occurs.
+These resonant layers are then generically surrounded by stochastic layers where chaotic motion occurs (in the vicinity of separatrices).
 Chaotic orbits are characterized by exhibiting random behavior despite being generated by entirely deterministic equations of motion.
-Stochastic layers form gaps in action space where tori are destroyed; the orbits that exist in these regions therefore can not be represented in angle-action coordinates, as their actions and frequencies are not constant and diffuse chaotically with time.
+Actions strictly do not exist for chaotic orbits and therefore chaos `destroys' orbital tori in certain regions of action-space --- these frequency spectrum of orbits in these regions evolve chaotically with time.
 
@@
 Much of the structure that was present in the integrable potential remains in the near-integrable case, but the differences are highlighted.
-Orbits in the resonant layers surrounding the resonances (grey) are near-resonant orbits that will librate around the resonance.
+Orbits in the resonant layers surrounding the resonances (grey) are near-resonant orbits that librate around the resonance.
 Chaotic orbits in the stochastic layer (red) behave erratically depending on the surrounding resonance structure.
@@
 
-For small values of $\epsilon$, many regular orbits survive and only small chaotic regions are introduced, especially where resonances overlap.
+For small values of $\epsilon$, many regular orbits survive and only small chaotic regions are introduced, especially where resonances overlap \citep[see][]{chirikov60}.
 As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed.
@@
 We therefore expect a large number of regular orbits will survive --- the so-called KAM tori --- however the tori that survive will be separated by regions of chaotic motion.
-Any transformations to angle-action coordinates must be defined locally due to the destruction of tori due to chaotic motion.
+Any transformations to angle-action coordinates must be defined local to each resonance region due to the destruction of tori and chaotic motion which lead to discontinuous changes in orbital properties.
 
@@
 Thus, if the resonance libration frequency is comparable (within a factor of a few) to the orbital frequency, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
-However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can't change by large factors (unless there is resonance overlap and the motion is strongly chaotic).
+However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can not change by large factors (unless there is resonance overlap and the motion is strongly chaotic).
 
 The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict.
-This has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}.
+This has been done analytically for simple potentials \citep[e.g.,][]{chirikov79}.
 For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; an orbit that ergodically mixes over its energy hypersurface must traverse this web, though the timescales typical for this phenomenon are many thousands of orbital periods.
 
-For galactic timescales, Arnold diffusion is not expected to be relevant, however chaotic motion across resonances can occur over short times.
+Arnold diffusion is not expected to be significant for most orbits over timescales relevant to galaxies (10s of orbits), however chaotic motion across resonances can occur over short times.
 If a stochastic orbit is semi-bounded by surrounding resonances, the allowed volume in which it can explore (in frequency space) is often quite small so that the orbit may appear to be approximately regular.
-However, this small-scale evolution may be important for tidal debris which has small intrinsic spreads in frequencies.
+However, this small-scale evolution may be important for tidal debris which can have small intrinsic spreads in frequencies.
 
-\subsection{Mixing of orbit ensembles}
+\subsection{Mixing of orbit ensembles}\label{sec:chaotic-mixing}
 
-Ensembles of orbits stuck between resonances may quickly spread to fill the allowed volume via the small-scale but short-time chaotic diffusion processes discussed above.
-This occurs due to phase-mixing of the fundamental frequencies of the ensemble orbits themselves --- the frequencies are evolving and thus trace out an orbit through frequency space.
-This is in contrast to phase-mixing of regular orbits and long-term chaotic mixing of strongly chaotic orbits.
+An ensemble of regular orbits (e.g., tidal debris) will phase-mix due to (small) differences in the fundamental frequencies of the orbits.
+Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits \cite[see, e.g.,][]{kandrup94, merritt96, kandrup03}, however this depends on the details of the resonant structure around the ensemble and are thus difficult to predict.
+For example,  ensembles of orbits `stuck' between resonances may quickly spread to fill the allowed volume via the short-time chaotic diffusion processes, but then the orbits must escape this confinement and diffuse through the Arnold web to reach a fully mixed state \citep{merritt96}.
 
-A small ensemble of (conditionally-periodic) regular orbits will reach a fully phase-mixed state in a timescale  $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in frequency-space of the ensemble --- each individual orbit will fill the surface of its torus ergodically, but the ensemble will only mix due to sheering from slight variations in their fundamental frequencies \citep{merritt96, helmi99}.
-Ensembles of chaotic orbits behave quite differently.
-Over long times, a chaotic orbit in an autonomous Hamiltonian will ergodically mix over the energy hypersurface (a 5D surface, compared to the 3D stable tori).
-Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits, however analytically computing the rate and amount of spreading is very difficult even for simple potentials \cite[see, e.g.,][]{chirikov??}.
-Typically these rates are comparable or larger than the Arnold diffusion timescale and are thus very long.
-
-In this work, we investigate the effect of the short-time but small-scale frequency evolution along weakly chaotic orbits and discuss how this effects our understanding of the coherence and density evolution of tidal streams.
+In this work, we investigate the consequences of this short-time but small-scale frequency evolution and hypothesize that this may explain the enhanced density evolution of tidal debris around weakly chaotic regions where chaotic timescales are predicted to be long \citep[e.g.,][]{pearson15}.
+We then discuss how this would affect our understanding of the coherence and density evolution of tidal streams.
 
@@
 
-Our goal is to map the orbit structure of arbitrary potentials, with an emphasis on identifying the chaotic orbits and understanding the evolution of these orbits over short times.
+Our goal is to map the orbit structure of arbitrary (galactic) potentials, with an emphasis on identifying the chaotic orbits and understanding the evolution of these ensembles of orbits over short times.
+In particular, we aim to understand how this chaos-enhanced density evolution can affect tidal stream morphology.
 In this section, we describe the methods we will use to detect and quantify the strength of chaos for large grids of orbits.
@@
 		\end{tabular}
-	\caption{Summary of potential parameters for the triaxial NFW potential (Equation~\ref{eq:potential}) used in this work.
-Triaxiality is introduced in the density to ensure the density is physical at all radii.
+	\caption{Summary of parameters for the triaxial NFW potential (Equation~\ref{eq:potential}) used in this work.
+Triaxiality is introduced in the density (rather than the potential) to ensure that the density is physical at all radii.
 Velocity scale, scale radius, and axis ratios are chosen to match the median halo parameters for a $M_{\rm vir} \approx 10^{12}~\msun$ halo from \citep{jing02}.
@@
 For a weakly chaotic orbit, the orbit may appear consistently periodic over long windows of time.
-\superfreq\ will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change location and amplitude with time.
+\superfreq\ will pick out a set of frequencies for chaotic orbits that correspond to the largest peaks in the power spectrum of the orbits, however these peaks will change location and amplitude with time.
 For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two consecutive sections of orbit.
@@
 In Section~\ref{sec:results1}, we generate grids of orbits in the potential described above to map the orbit structure of the potential.
-We classify each orbit in terms of the strength of chaos along the orbit as computed using the methods of Section~\ref{sec:methods}.
+We classify each orbit in terms of the strength of chaos along the orbit as computed using the Lyapnov and frequency diffusion times of Section~\ref{sec:methods}.
 With this initial classification, in Section~\ref{sec:results2} we follow the evolution of ensembles of orbits generated around each orbit in the initial grid.
@@
 This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits), but the most numerous orbits in this grid are the short-axis and outer long-axis tubes.
-Thin tidal streams are thought to form along tube orbits rather than box orbits because of the faster disruption expected for stellar systems on radially plunging orbits.
+Thin tidal streams are thought to form along tube orbits rather than box orbits because of the faster disruption expected for stellar systems on radially plunging orbits which makes them more diffuse.
+
+\todo{apw}{New figure showing just the grid and orbit classifications, labels for orbit types}
 
@@
 For each orbit, we also separately integrate for $\approx$80 orbital periods and use \superfreq\ to compute the fundamental frequencies for the two consecutive sections of 40 orbital periods.
-We have found that this is the shortest integration window for which we are able to consistently and accurately recover the frequencies using \superfreq.
-Figure~\ref{fig:freqdiff} shows the same grid of initial conditions, but now the greyscale intensity is set by the logarithm of the (order-unity) frequency diffusion time.
+We have found that this is the shortest integration window for which we are able to successfully recover frequencies using \superfreq\ for $>$99.9\% of the orbits.
+Figure~\ref{fig:freqdiff} shows the same grid of initial conditions as in Figure~\ref{fig:lyapmap}, but now the greyscale intensity is set by the logarithm of the (order-unity) frequency diffusion time.
 The darker pixels have shorter frequency diffusion times and are more chaotic.
@@
 The strongest chaotic regions (black) appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}).
-Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map, especially where resonances overlap --- for example, near $(x,z) \approx(13,7)$ and $(x,z) \approx(31,13)$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}.
-The details of the resonant structure is revealed in the frequency diffusion map from integrations of just 80 orbital periods, though the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods.
-In the next section, we analyze the density evolution of finite-volume ensembles of orbits around each orbit in the above grids and compare the density evolution of the ensembles to the single-orbit chaos indicators computed in this section.
+Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map, especially where resonances overlap --- for example, near $(x,z) \approx(13,7)~{\rm kpc}$ and $(x,z) \approx(31,13)~{\rm kpc}$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}.
+The details of the resonant structure is revealed in the frequency diffusion map from integrations of just 80 orbital periods.
+
+While there is rich structure and a significant number of weakly chaotic orbits, the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods and are thus not expected to be relevant for tidal stream evolution.
+In the next section, we analyze the density evolution of finite-volume ensembles of orbits around each orbit in the above grids in order to study the effect of weak chaos on tidal debris.
+We then compare the density evolution of the ensembles to the single-orbit chaos indicators computed in this section.
 
@@
 
-The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant \emph{local} to a given orbit --- that is, these quantities are measures of how infinitesimal deviations will diverge on average from some parent orbit of interest, or of how long it takes for the frequencies of a single orbit to change by some amount.
+The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant for a given orbit --- that is, these quantities are measures of how infinitesimal deviations will diverge on average from some parent orbit of interest, or of how long it takes for the frequencies of a single orbit to change by some amount.
 Tidal debris is disrupted from progenitor systems with finite spreads in orbital properties (e.g., energy).
 For a disrupting, globular-cluster-scale progenitor, the typical energy dispersion of the debris is 0.1--1\% of the progenitor orbital energy \citep[assuming masses of $10^4$--$10^5$~\msun;][]{johnston98}, but for a dwarf-galaxy-scale progenitor, the dispersion can be $\gtrsim$10\%.
-Does the Lyapunov or frequency diffusion time predict the timescale over which a finite phase-space volume (e.g., tidal debris) stays coherent?
+In this section, we ask whether  the Lyapunov or frequency diffusion time predict the timescale over which a finite phase-space volume (e.g., tidal debris) stays coherent.
+
+It is computationally intractable to run full N-body simulations for the large grid of orbital initial conditions of the previous section and we therefore take a simplified approach for studying how finite-volume debris spreads along each of these orbits.
+We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
+For a given set of orbital initial conditions --- the `parent' orbit --- we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time.
+The physical scale of the ensemble is set by the tidal radius in position and the velocity scale in velocity and are therefore set by the mass of the progenitor \citep[e.g.,][]{johnston98, apw14}.
+If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
+\begin{align}
+	\delta\bs{x}_i &\sim \mathcal{N}(0, r_{\rm tide})\\
+	\delta\bs{v}_i &\sim \mathcal{N}(0, \sigma_v)\\
+	r_{\rm tide} &= \|\bs{x}_0\| \left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} \\
+	\sigma_v &= \|\bs{v}_0\|\left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} 
+\end{align}
+where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the `progenitor,' and $\|\cdot \|$ is the Euclidean norm.
+We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales.
 
@@
 The orbits were chosen from a single column of the orbit grid based on their frequency diffusion time to sample a range of possible values, and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}.
-We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on these three orbits.
-The initial mass of the clusters is \todo{kvj}{2.5E4 \msun contained within what again?
-35 scale radii?
-but this is a plummer sphere?}.
-For each simulation, the progenitor is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
-We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter.
+Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the three orbits described above and evolved for 16 orbital periods.
+A thin `stream' forms on the regular orbit (left column), a more diffuse `stream' on the mildly chaotic orbit (middle column), and a `fanned' stream on the strongly chaotic orbit (right column).
+Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit, it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream.
+
+We verify that these orbit ensembles capture the nature of more realistic stream formation by running N-body simulations of globular-cluster-mass progenitor systems on these same three orbits.
+We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run the simulations, which we set up to run from apocenter-to-apocenter (rather than pericenter-to-apocenter as in the ensembles), but finish with the progenitor in the same location as the parent orbit in the ensemble evolution described above.
 Figure~\ref{fig:nbodysims} shows the final particle distributions rotated so that the angular momentum of the progenitor orbit is aligned with the $z$-axis.
-A thin stream forms on the regular orbit (left column), a more diffuse stream on the mildly chaotic orbit (middle column), and a `fanned' stream on the strongly chaotic orbit (right column).
-Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit, it is surprising that stream formed on this orbit appears to be more diffuse than the regular orbit stream.
+From a comparison of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims}, it is clear that the morphology of the ensembles is visually similar to the `oldest' (first stripped) debris in the N-body simulations.
 
@@
 
-It is computationally intractable to run full N-body simulations for the large grid of orbital initial conditions of the previous section and we therefore take a simplified approach for studying how finite-volume debris spreads along each of these orbits.
-We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
-For a given set of orbital initial conditions --- the `parent' orbit --- we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time.
-The physical scale of the ensemble is set by the tidal radius in position and the velocity scale in velocity and are therefore set by the mass of the progenitor \citep[e.g.,][]{johnston98, apw14}.
-If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
-\begin{align}
-	\delta\bs{x}_i &\sim \mathcal{N}(0, r_{\rm tide})\\
-	\delta\bs{v}_i &\sim \mathcal{N}(0, \sigma_v)\\
-	r_{\rm tide} &= \|\bs{x}_0\| \left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} \\
-	\sigma_v &= \|\bs{v}_0\|\left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} 
-\end{align}
-where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the `progenitor,' and $\|\cdot \|$ is the Euclidean norm.
-We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales.
-Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the three orbits described above and evolved for 16 orbital periods (same as the N-body simulations); the morphology of the ensembles is visually similar to the `oldest' (first stripped) debris in the N-body simulations.
-
-We find that chaotic mixing of small orbit ensembles is detectable over short times, even when the predicted chaotic timescales (from the Lyapunov and frequency diffusion times) are long and therefore that the mean, single-orbit chaos indicators are not well-suited for determining the importance of chaotic diffusion on tidal stream evolution.
-For example, Figure~\ref{fig:ensemble-density} shows the evolution of the configuration-space density for ensembles evolved around the three orbits (R, W, S) described above.
+Visual inspection of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims} suggests that chaotic mixing of small orbit ensembles affects the configuration-space evolution of an ensemble over short times, even when the predicted chaotic timescales (from the Lyapunov and frequency diffusion times) are long and therefore that the mean, single-orbit chaos indicators are not well-suited for determining the importance of chaotic diffusion on tidal stream evolution.
+As a quantitative measure of this enhanced density evolution, Figure~\ref{fig:ensemble-density} shows the evolution of the configuration-space density for ensembles evolved around the three orbits (R, W, S) described above.
 At each time step, we use kernel density estimation (KDE)\footnote{We use an implementation from the Python package \texttt{scikit-learn} \citep{scikitlearn}.} with the ensemble of particle positions to estimate the configuration-space density field.
@@
 
+\todo{apw}{motivate by saying we need a single measure to characterize ensemble evolution even though density evolution can be complex...can then map entire orbit grid}
 \todo{apw}{These paragraphs need work, but still waiting for numerical results from the cluster...}
@@
 diff time are long, can still resolve structure in these maps.}
+
+We conclude from these experiments that the degree of chaos is an indicator that ensembles of orbits will mix faster than regular phase-mixing, however the Lyapunov time and frequency diffusion time are not good predictors for the timescale over which this mixing will occur.
+To understand this discrepancy, we next explore why this occurs (Section~\ref{sec:results3}, in the context of Section~\ref{sec:chaotic-mixing}) and develop a new indicator that can be tested over a finite number of orbital periods to predict whether chaotic diffusion will enhance the mixing of orbit ensembles (Section~\ref{sec:results4}).
 
@@
 \todo{apw}{This section has not been edited carefully...}
+\todo{apw}{What do I mean by `more reliable' for $p=2$?}
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/nbody.png}
-\caption{Particle positions from the final snapshots of N-body simulations of a globular-cluster-mass satellite disrupting in a triaxial NFW potential (Section~\ref{sec:potential}).
-Each satellite is evolved for 16 orbital periods, and begins and ends at apocenter.
-For visualization, the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
-Top row shows the positions projected onto the resulting $xy$ plane (the instantaneous orbital plane of the progenitor at the end of the simulation), bottom row shows the positions projected onto the $yz$ plane.
-From left to right, the columns indicate that the progenitor is on a regular (R), weakly chaotic (W), or strongly chaotic (S) orbit.
-Even on the weakly chaotic orbit --- where the predicted chaotic timescale is XX orbital periods --- the resulting tidal stream displays stream-fanning and the `oldest' debris is more diffuse than in the regular case.} \label{fig:nbodysims}
+\includegraphics[width=\textwidth]{figures/ensembles.png}
+\caption{Final particle positions after integrating unbound, globular-cluster-sized ensembles of orbits generated around each of the three N-body orbits (Figure~\ref{fig:nbodysims}).
+The ensembles each contain 1000 particles and begin at pericenter rather than apocenter because they are meant to represent debris disrupted during a single (the first) pericentric passage.
+The particle positions qualitatively match the morphology of the `oldest' debris from the N-body simulations, though there are many more particles in these ensembles than are disrupted during the first pericentric passage of the simulations.
+}
+\label{fig:ensembles}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/ensembles.png}
-\caption{Final particle positions after integrating unbound, globular-cluster-sized ensembles of orbits generated around each of the three N-body orbits (Figure~\ref{fig:nbodysims}).
-The ensembles each contain 1000 particles and begin at pericenter rather than apocenter because they are meant to represent debris disrupted during a single (the first) pericentric passage.
-The particle positions qualitatively match the morphology of the `oldest' debris from the N-body simulations, though there are many more particles in these ensembles than are disrupted during the first pericentric passage of the simulations.
-}
-\label{fig:ensembles}
+\includegraphics[width=\textwidth]{figures/nbody.png}
+\caption{ \todo{apw}{clean this up}
+Length-scale of plummer sphere set to get 10\% mass loss so 1000 particles so ensemble tests reasonable...
+We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on these three orbits.
+The initial mass of the clusters is .
+For each simulation, the progenitor is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
+We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter.
+Particle positions from the final snapshots of N-body simulations of a globular-cluster-mass satellite disrupting in a triaxial NFW potential (Section~\ref{sec:potential}).
+Each satellite is evolved for 16 orbital periods, and begins and ends at apocenter.
+For visualization, the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
+Top row shows the positions projected onto the resulting $xy$ plane (the instantaneous orbital plane of the progenitor at the end of the simulation), bottom row shows the positions projected onto the $yz$ plane.
+From left to right, the columns indicate that the progenitor is on a regular (R), weakly chaotic (W), or strongly chaotic (S) orbit.
+Even on the weakly chaotic orbit --- where the predicted chaotic timescale is XX orbital periods --- the resulting tidal stream displays stream-fanning and the `oldest' debris is more diffuse than in the regular case.}
+\label{fig:nbodysims}
 \end{center}

### 2015-06-25 ms.tex
new results section

@@
 In Section~\ref{sec:methods}, we describe our choice of potential, method for numerical orbit integration, and introduce the chaos indicators used in this work.
-Our results are split into three subsections: in Section~\ref{sec:results1} we present iso-energy grids of orbits and discuss the orbit classes and chaotic timescales present in our potential; in Section~\ref{sec:results2} we study the density evolution of small ensembles of orbits around each orbit of the previous section; and, in Section~\ref{sec:results3} we describe the behavior of chaotic diffusion and use this to explain how chaos is relevant for tidal streams over short times.
+Our results are split into four subsections: in Section~\ref{sec:results1} we present iso-energy grids of orbits and discuss the orbit classes and chaotic timescales present in our potential; in Section~\ref{sec:results2} we study the density evolution of small ensembles of orbits around each orbit of the previous section; in Section~\ref{sec:results3} we describe the behavior of chaotic diffusion and use this to explain how chaos is relevant for tidal streams over short times; and in Section~\ref{sec:results4} we propose a new indicator for evaluating whether chaotic diffusion of orbits will be important for the morphological evolution of tidal debris.
 We discuss the implications of our results in Section~\ref{sec:discussion}, and conclude in Section~\ref{sec:conclusions}.
@@
 This small-scale variability is insignificant for the evolution of global structure in, for example, an elliptical galaxy, but is signifiant for the morphological evolution of tidal debris with small spreads in frequencies.
+
+\subsection{Part IV: Proposed indicator of chaotic density evolution}\label{sec:results4}
 

### 2015-06-25 ms.tex
redo disc and conc w kvj and sp comments

@@
 
-\section{Discussion}\label{sec:discussion}
+\section{Discussion: limitations and future work}\label{sec:discussion}
 
-\subsection{Existence of streams}
+% Something about this: The timescales in this paper are not meant to be interpreted physically, as the potential is a very crude approximation to the total potential of a Milky-Way-like galaxy.
+The key point is that even when the mean chaos indicators predict a long chaotic timescale, streams may display enhanced density evolution over much shorter times.
+Tidal debris is sensitive to the small-scale, short-time evolution of orbital properties (e.g., the fundamental frequencies) induced by weak chaos.
 
-We have considered here a simple triaxial gravitational potential chosen to mimic the median properties of dark matter halos formed in dark-matter-only simulations (or the large-scale properties of halos formed in simulations with baryonic effects).
-We have found that `mean' chaos indicators such as the Lyapunov exponent or frequency diffusion rate predict that chaos should be dynamically unimportant in the halos of galaxies for the majority of orbits (e.g., the predicted timescales are 100-1000s of times longer than the typical orbital times, at least in this simple potential).
-However, while the mean chaotic evolution of, for example, the fundamental frequencies of an orbit may be prohibitively long, the frequencies of even mildly chaotic orbits do evolve on a smaller scale over much shorter times.
-The magnitude of these variations depends on the resonant structure local to the orbit (which is highly potential-dependent), but this result is generic to any potential with resonances and regions of stochasticity.
-When this small-scale but short-time frequency evolution has an amplitude comparable to or larger than the intrinsic spread in orbital frequencies for tidal debris, the debris will appear dynamically hotter and become diffuse faster than along a regular orbit.
-We have found that even in the simple potential used in this work (Section~\ref{sec:potential}), a large fraction of orbits would naively be considered weakly chaotic, but tidal streams formed on these orbits would appear morphologically different (e.g., more diffuse) than those formed on regular orbits.
-
-If this is the case for a single-component, static, smooth potential, the mere existence of thin tidal streams in the halo of the Milky Way either (1) provides useful information about the potential on these scales by, e.g., implying a large degree of regularity, or (2) indicates that the thin, long streams (e.g., GD-1) are on regular orbits.
-These are not mutually exclusive --- in fact, if the streams are on regular orbits, this would be a powerful way to check or rule out possible potential models by requiring that the progenitor orbits remain regular.
-The longest thin streams are most valuable for this effort because they have clearly evolved for a long time, but the debris remains compact.
-For shorter thin streams it will be hard to decouple the unknown evolution time from enhanced density evolution from chaos.
+We have shown that the Lyapunov and frequency diffusion times are indicators of chaos and that the frequency diffusion time resolves the detailed resonant structure of gravitational potentials, but the timescales predicted do not capture the importance of chaos for the density evolution of ensembles of orbits meant to mimic tidal debris.
+We have shown that small-scale but fast chaotic diffusion of orbits can explain this enhanced mixing, and have proposed a new indicator to characterize the importance of this effect over given evolution times.
+In the sub-sections below, we discuss a few important limitations that remain for exploration in future work.
 
@@
 
-In this work, we have only considered low-mass progenitor systems such as globular clusters because the intrinsic spreads in fundamental frequencies are small (0.1--0.5\%).
+We have only considered low-mass progenitor systems such as globular clusters because the intrinsic spreads in fundamental frequencies are small (0.1--0.5\%).
 Small changes to the frequencies of the orbits of tidal debris disrupted from these progenitors due to chaotic processes will therefore cause observable changes to the real-space morphology of the debris.
@@
 The mass-scale of the debris that will display enhanced density evolution depends on the magnitude of weak chaos, which depends on the orbit structure of a given potential.
-In potentials with more significant chaos, debris disrupted from more massive progenitor systems may also display stream-fanning.
+In potentials with more significant chaos, debris disrupted from more massive progenitor systems may also display `stream-fanning.'
 
-The mass of the progenitor sets the physical scale of the debris as it is disrupted.
-For more massive progenitors and therefore larger spreads in debris, different orbits may have different chaotic diffusion timescales that depend on the details of the disruption (e.g., the orbital phase of the pericenter, or the energy offset of the debris).
-This could potentially lead to interesting density variations along the tidal debris.
+\subsection{Potential choice}
+
+The potential considered in this work is `unrealistic' for the total potential of a Milky-Way-like galaxy in that it is static, smooth, and does not contain baryonic components.
+Simulations of forming galactic disks in cosmological dark matter halos have shown that baryonic feedback and relaxation can significantly change the inner distribution of dark matter and either make the potential more spherical or oblate \citep{many}.
+However, the significance of baryonic relaxation or of time-dependence, triaxiality, and substructure on shaping the matter distribution within the Milky Way is largely unknown.
+Here we briefly summarize future directions for potential models:
+\begin{itemize}
+	\item \emph{Baryonic components}: \cite[][D08]{debattista08} and \cite[][V10]{valluri10} studied the orbit evolution induced by growing a baryonic disk in dark matter halos with various shapes and orientations (e.g., prolate, triaxial).
+In general, the authors find that the growth of a disk slowly deforms the orbits of mass tracers (e.g., dark matter particles) and preferentially populates tube orbits.
+Consequently, the inner shape of the potential becomes more oblate or spherical due to the prevalence of tube orbits.
+If the inner halos of galaxies are indeed close to spherical or oblate, the majority of orbits will be regular and chaos will be less important, however this is far from conclusive.
+We have found from simple experiments in superposing potential components that transition regions between potential shapes can actually enhance significant amounts of chaos.
+Though superposing potentials is not realistic, it is at least suggests that a more careful exploration of potential configurations is required to understand how complex, radius-dependent potential forms affect the amount and significance of chaos in galactic halos.
+
+\item \emph{Time dependence}: Galaxies are certainly not static systems.
+To first order, galaxies grow in mass --- for example, the spherically averaged mass profile of dark matter halos evolves fairly predictably in cosmological simulations \citep{wechsler02, buist14} after some initial period of stochastic mass growth.
+The Milky Way has probably had a fairly calm accretion history over the last 6 Gyr and therefore the mass growth may be similar to that seen in simulations.
+This steady growth most likely does not alter the global structure or shape of the potential.
+However, we also know from simulations that figure rotation, baryonic feedback, and the accretion and phase-mixing of subhalos do perturb the global state of simulated halos.
+\cite{deibel11} showed that by adopting pattern speeds comparable to those found in cosmological simulations, figure rotation generally acts to destabilize orbits (rather than stabilize chaotic orbits in the equivalent static potential) and the resulting orbit structure is that most regular orbits are associated with resonances.
+In future work we will explore the effect of these time-dependent processes on the chaotic dispersal of tidal streams using live potentials from cosmological N-body simulations.
+
+\item \emph{Substructure}: Cosmological simulations predict that dark matter halos are filled with substructure in the form of dark matter subhalos.
+If they exist, these subhalos may account for up to $\approx$1-10\% of the mass of the dark matter \citep{??} and therefore may contribute significantly to and orbit in the large-scale potential of any galaxy.
+Gravitational scattering due to subhalo interactions has been studied, however in the halos of galaxies where dynamical times are long, the scattering cross-section for \emph{strong} encounters is very small \citep{??}.
+Instead, the collective effect of the subhalos may instead act as a noise term in the Hamiltonian of any halo orbit.
+This subhalo-induced colored noise --- which depends on the mass spectrum and distribution of subhalos --- may also act to simply increase the magnitude of chaos along orbits, and destabilize sufficiently non-resonant orbits \citep[see, e.g.,][]{kandrup00}.
+
+\end{itemize}
+
+\subsection{Stream modeling}
+
+Tidal stream modeling is one of the most promising ways to constrain the 3D mass distribution around the Milky Way at distances of 10s to 100s of kpc.
+Methods that use tidal streams to infer properties of the Galactic potential typically operate by constructing models of the debris distribution using either the present-day phase-space density \citep[e.g.,][]{kuepper12, kuepper15, amorisco and others}, time-of-disruption phase-space density \citep{apw13, apw14}, or the density in angle-action coordinates \citep{sanders14, bovy14}.
+All of these methods may fail or produce uninterpretable results if modeling globular-cluster streams on mildly chaotic orbits.
+For each of these methods, it is important to understand the failures and biases introduced by ignoring chaotic orbit evolution.
 
@@
 The fanned debris would be much lower surface brightness and thus much more difficult to detect.
-However, the detection of such stars would confirm that enhanced density evolution is a dynamically relevant process for thin streams in real galaxies.
-
-\subsection{The dangers of stream modeling}
-
-Tidal stream modeling is one of the most promising ways to constrain the 3D mass distribution around the Milky Way at distances of 10s to 100s of kpc.
-Methods that use tidal streams to infer properties of the Galactic potential typically operate by constructing models of the debris distribution using either the present-day phase-space density \citep[e.g.,][]{kuepper12, kuepper15, amorisco and others}, time-of-disruption phase-space density \citep{apw13, apw14}, or the density in angle-action coordinates \citep{sanders14, bovy14}.
-All of these methods may fail or produce uninterpretable results if modeling globular-cluster streams on mildly chaotic orbits.
-
-Consider a globular cluster system that disrupts along a mildly chaotic orbit so that a significant portion of the resulting stream is more diffuse (lower surface brightness) and therefore unobservable due to observational biases.
-The high-density part of the stream may still appear thin (e.g., the recently disrupted debris in the middle panels of Figure~\ref{fig:nbodysims}).
-If the stream modeling focuses on reproducing the present-day density of the stream and the potential is restricted to an integrable, analytic form, the phase-space distribution of the debris may be well-reproduced by a progenitor on a particular orbit in such a potential.
-However, the inferences about the potential or mass distribution may be biased significantly and the inferences about the evolution time and internal properties of the system will certainly be incorrect --- for example, the evolution time may be biased to shorter times to reproduce the apparent shortness of the stream.
-With realistic observational uncertainties, it is likely the case that for any stream formed from a progenitor on a weakly chaotic orbit, there is a regular progenitor orbit in some other potential that produces a model stream equally consistent with the data.
-
-Due to the often unknown or poorly understood selection biases involved in kinematic observations of stream stars (due to, e.g., survey limits, dust extinction, and stellar population variations), modeling the final phase-space densities of tidal streams will lead to uninterpretable inferences about the potential of the host galaxy.
-We argue that by instead performing the density modeling at the time of disruption \citep[e.g.,][]{apw14}, uncertainties in, for example, the length or width of the stream will not bias potential parameter inferences as significantly (see Price-Whelan et al.
-2015, in prep.).
-In either case, it is critical that potential modeling efforts with tidal streams allow enough freedom in the potential to explore all possible configurations, though this quickly becomes a high-dimensional, multi-modal, and therefore computationally prohibitive inference problem.
-
-\todo{apw}{Need more coherent words about what happens when trying to use angle-action coordinates...}.
-
-\subsection{Potential choice}
-
-The potential considered in this work is `unrealistic' for the total potential of a Milky-Way-like galaxy in that it is static, smooth, and does not contain baryonic components.
-Hydrodynamic simulations of forming galactic disks in cosmological dark matter halos have shown that baryonic feedback and relaxation can significantly change the inner distribution of dark matter and either make the potential more spherical or oblate \citep{many}.
-However, the significance of baryonic relaxation or of time-dependence, triaxiality, and substructure on shaping the matter distribution within the Milky Way is largely unknown.
-We now consider each of these effects independently and comment on the expected consequences of these perturbations.
-
-\subsubsection{Baryonic components}
-
-\cite[][D08]{debattista08} and \cite[][V10]{valluri10} studied the orbit evolution induced by growing a baryonic disk in dark matter halos with various shapes and orientations (e.g., prolate, triaxial).
-In the simulations of D08, the disk component is artificially fixed along one of the principal axes of the halo potential and the mass of the disk is grown adiabatically.
-D08 found that (in general) the growth of a disk slowly deforms the orbits of mass tracers (e.g., dark matter particles) and preferentially populates tube orbits but does not result in a large degree of chaotic scattering.
-Consequently, the inner shape of the potential becomes more oblate or spherical due to the prevalence of tube orbits.
-
-If the inner halos of galaxies are indeed close to spherical or oblate, the majority of orbits will be regular and chaos will be less important, however this is far from conclusive.
-First, there are observational hints that the intrinsic alignment of galactic disks may be close to random \citep{??} (rather than fixed to symmetry planes of the host halo potential).
-Simulations since the work of D08 have shown that by `freeing' the disk component, the angular momentum vector of the disk can precess and wander erratically away or around the principal axes of the host halo \citep{debattista15}.
-It is unknown what effect this would have on the orbit structure of nearby halo orbits, but we naively expect that this would destabilize a significant number of orbits.
-Second, if the outer potential remains triaxial, the transition region between, e.g., oblate and triaxial may be populated by chaotic orbits.
-We have found from simple experiments in superposing potential components that transition regions between potential shapes induce significant amounts of chaos.
-Though superposing potentials is not realistic, it is at least suggestive that such configurations might be unstable.
-A more careful exploration of this phenomenon will be presented in future work.
-Lastly, the existence of a bulge or thin disk component may enhance chaos through a process similar to periodic forcing of the orbit.
-Passage through a thin disk results in a nearly instantaneous, impulsive forcing of the orbit at a frequency set by the circulation frequency of the orbit.
-This may act to increase the significance of chaos along disk-crossing orbits \citep[e.g.,][]{kandrup00, hunter05}.
-This effect is likely unresolved in many simulations where the force and spatial resolution is insufficient to capture this effect.
-
-\subsubsection{Subhalos}
-
-Cosmological simulations predict that dark matter halos are filled with substructure in the form of dark matter subhalos.
-If they exist, these subhalos may account for up to $\approx$1-10\% of the mass of the dark matter \citep{??} and therefore may contribute significantly to and orbit in the large-scale potential of any galaxy.
-Gravitational scattering due to subhalo interactions has been studied, however in the halos of galaxies where dynamical times are long, the scattering cross-section for \emph{strong} encounters is very small \citep{??}.
-Instead, the collective effect of the subhalos may instead act as a noise term in the Hamiltonian of any halo orbit.
-This subhalo-induced colored noise --- which depends on the mass spectrum and distribution of subhalos --- may also act to simply increase the magnitude of chaos along orbits, and destabilize sufficiently non-resonant orbits \citep[see, e.g.,][]{kandrup00}.
-
-\subsubsection{Time-dependence}
-
-Galaxies are certainly not static systems.
-To first order, galaxies grow in mass --- for example, the spherically averaged mass profile of dark matter halos evolves fairly predictably in cosmological simulations \citep{wechsler02, buist14} after some initial period of stochastic mass growth.
-The Milky Way has probably had a fairly calm accretion history over the last 6 Gyr and therefore the mass growth may be similar to that seen in simulations.
-This steady growth most likely does not alter the global structure or shape of the potential.
-However, we also know from simulations that figure rotation, baryonic feedback, and the accretion and phase-mixing of subhalos do perturb the global state of simulated halos.
-\cite{deibel11} showed that by adopting pattern speeds comparable to those found in cosmological simulations, figure rotation generally acts to destabilize orbits (rather than stabilize chaotic orbits in the equivalent static potential) and the resulting orbit structure is that most regular orbits are associated with resonances.
-In future work we will explore the effect of these time-dependent processes on the chaotic dispersal of tidal streams using live potentials from cosmological N-body simulations.
+The detection of this low-surface-brightness component would open up the possibility that enhanced density evolution due to chaos is a dynamically relevant process for thin streams in real galaxies, though \cite{carlberg15} has shown that it is also possible also to get short, high density segments of streams from streams formed on eccentric orbits.
 
@@
 
-The large-scale shapes of galactic gravitational potentials are thought to be triaxial and therefore contain an appreciable number of chaotic orbits.
+%The large-scale shapes of galactic gravitational potentials are thought to be triaxial and therefore contain an appreciable number of chaotic orbits.
 Chaos is not thought to be dynamically important for shaping the global structure of galaxies because the predicted chaotic timescales such as the Lyapunov or frequency diffusion times for orbits in galactic-like potentials are very long (many 1000s of orbital periods).
@@
 
-We have numerically computed the magnitude of chaos for a large grid of iso-energy orbits in a triaxial gravitational potential using two independent methods that have been used extensively to classify and characterize chaotic orbits: the Lyapunov exponent and frequency diffusion rate.
+We have considered here a simple triaxial gravitational potential chosen to mimic the median properties of dark matter halos formed in dark-matter-only simulations (or the large-scale properties of halos formed in simulations with baryonic effects).
+We have numerically computed the magnitude of chaos for a large grid of iso-energy orbits in this potential using two independent methods that have been used extensively to classify and characterize chaotic orbits: the Lyapunov exponent and frequency diffusion rate.
 From each of these indicators, we compute a timescale over which chaos is likely to be important and find that the majority of orbits have chaotic timescales greater than 100s of orbital periods, however with the frequency diffusion rate we are still able to resolve weak chaos.
-We then study the density evolution of small ensembles of orbits generated around each orbit in the grid used for the previous experiment and find that along some orbits classified as weakly chaotic --- with chaotic timescales of 100s of orbital periods --- the orbit ensembles display enhanced density evolution and reach a lower overall density faster than orbit ensembles around nearby regular orbits.
+We then study the density evolution of small ensembles of orbits generated around each orbit in the grid used for the previous experiment and find that along some orbits classified as weakly chaotic --- with chaotic timescales of 100s of orbital periods --- the orbit ensembles display enhanced density evolution and reach a lower overall density faster than orbit ensembles around nearby regular orbits (which mix due to phase-mixing alone).
 We propose an explanation for this apparent discrepancy by considering the nature of chaotic diffusion: the classical chaos indicators are most sensitive to the slow, Arnold diffusion process that can cause large changes to orbital properties, but small-scale frequency evolution occurs over much shorter times as chaotic orbits stochastically diffuse across the stochastic layers that surround many resonances.
@@
 
-\todo{apw}{Concluding paragraph that drives home the importance of this work...}
-The timescales in this paper are not meant to be interpreted physically, as the potential is a very crude approximation to the total potential of a Milky-Way-like galaxy.
-The key point is that even when the mean chaos indicators predict a long chaotic timescale, streams may display enhanced density evolution over much shorter times.
-Tidal debris is sensitive to the small-scale, short-time evolution of orbital properties (e.g., the fundamental frequencies) induced by weak chaos.
+If this is the case for a single-component, static, smooth potential, the mere existence of thin tidal streams in the halo of the Milky Way either (1) provides useful information about the potential on these scales by, e.g., implying a large degree of regularity, or (2) indicates that the thin, long streams (e.g., GD-1) are on regular orbits.
+These are not mutually exclusive --- in fact, if the streams are on regular orbits, this would be a powerful way to check or rule out possible potential models by requiring that the progenitor orbits remain regular.
+The longest thin streams are most valuable for this effort because they have clearly evolved for a long time, but the debris remains compact.
+For shorter thin streams it will be hard to decouple the unknown evolution time from enhanced density evolution from chaos.
+
+\todo{apw}{Last paragraph that drives home the importance of this work...too tired to finish this tonight...}
 

### 2015-06-25 ms.tex
kvj comments round2

@@
 Once the debris has evolved far enough from the progenitor system that the self-gravity of the progenitor can be ignored (usually a fast process relative to the orbital time), the stars evolve essentially as an ensemble of test particle orbits in the potential of the host system.
-The debris remains coherent as a tidal stream if the phase-mixing time-scale is long: a small ensemble of regular orbits reaches a fully phase-mixed state in a timescale $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in fundamental frequencies of the ensemble.
+The debris remains coherent as a tidal stream if the phase-mixing time-scale is long: a small ensemble of regular orbits reaches a fully phase-mixed state in a timescale $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in fundamental frequencies of the ensemble (tidal debris from a globular cluster typically has frequency spreads $\approx$0.1--1\%, so it can take hundreds to thousands of orbital periods to fully phase-mix).
 The ensemble spreads due to sheering from slight variations in their fundamental frequencies which, for tidal streams, preferentially occurs along one dimension \citep{merritt96, helmi99}.
@@
 By modeling the observed phase-space density of stream stars along with the host galaxy potential it is hoped that we may infer the 3D mass distribution of the host.
-Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies; the known streams span a large range of distances --- from $\approx10$ to $100~kpc$ --- and progenitor masses --- from $\approx$10$^3$ to $\approx$10$^8~\msun$ in stellar mass --- \citep[][]{cite many}.
+Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies; the known streams span a large range of distances --- from $\approx10$ to $100~{\rm kpc}$ --- and progenitor masses --- from $\approx$10$^3$ to $\approx$10$^8~\msun$ in stellar mass --- \citep[][]{cite many}.
 There has been extensive work on developing methods to use data from these streams to measure properties of Milky Way's dark matter halo.
@@
 In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically form large, two-dimensional `fans' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness and curvature of the stream given the observational constraints of the present-day position and velocity of the cluster.
+The result in \citet{pearson15} demonstrates that the morphology of a stream alone can rule out a potential.
+With an understanding of the circumstances that lead to the differences in stream morphology, this could be a powerful tool for rejecting potentials from positional information alone.
 
@@
 The result from \citet{pearson15} suggests that even weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.
+Understanding why this occurs and developing a measure to quantify the importance of this enhanced density evolution is a promising new direction to be explored further in this work.
 
-In this work, we study the effect of chaotic diffusion of individual orbits on the density evolution of tidal debris.
+In this work, we study the effect of chaotic diffusion of the fundamental frequencies of individual orbits on the density evolution of tidal debris.
 We choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos as computed from single-orbit diagnostics for grids of constant-energy orbits, and compare these results to measures of the density evolution of finite-volume ensembles of orbits (meant to mimic tidal debris).
 We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles due to the chaotic diffusion of the individual orbits.
-The diffusion of the fundamental frequencies of the orbits introduces an extra, time-dependent dispersion to the ensembles and therefore reduces the timescale over which mixing becomes relevant (e.g., $\sigma_\Omega$ can grow with time).
-This result supports a reevaluation of the importance of chaos in galactic halos and implies that the amount of chaos in a given potential may have significant consequences for the observability and survivability of thin, cold tidal streams.
+For a chaotic orbit, the frequency spectrum of the orbit evolves with time --- for a small ensemble, the spread in frequencies is therefore time-dependent, which could enhance phase-mixing.
+This idea supports a reevaluation of the importance of chaos in galactic halos and implies that the amount of chaos in a given potential may have significant consequences for the observability and survivability of thin, cold tidal streams.
+We propose a new measure for determining whether this small-scale frequency diffusion will be important for the morphological evolution of tidal debris with a given spread in frequencies or orbital properties over finite and dynamically relevant times.
 
@@
 We then study the density evolution of small ensembles of orbits generated around each orbit in the grid used for the previous experiment and find that along some orbits classified as weakly chaotic --- with chaotic timescales of 100s of orbital periods --- the orbit ensembles display enhanced density evolution and reach a lower overall density faster than orbit ensembles around nearby regular orbits (which mix due to phase-mixing alone).
-We propose an explanation for this apparent discrepancy by considering the nature of chaotic diffusion: the classical chaos indicators are most sensitive to the slow, Arnold diffusion process that can cause large changes to orbital properties, but small-scale frequency evolution occurs over much shorter times as chaotic orbits stochastically diffuse across the stochastic layers that surround many resonances.
+We explain this apparent discrepancy by considering the nature of chaotic diffusion: the classical chaos indicators are most sensitive to the slow, Arnold diffusion process that can cause large changes to orbital properties, but small-scale frequency evolution occurs over much shorter times as chaotic orbits stochastically diffuse across the stochastic layers that surround many resonances.
 We propose a new measure of chaos with tidal stream morphological evolution in mind: the amplitude of small-scale, fractional changes in frequencies over a given number of orbital periods along a stream progenitor orbit.
@@
 
-If this is the case for a single-component, static, smooth potential, the mere existence of thin tidal streams in the halo of the Milky Way either (1) provides useful information about the potential on these scales by, e.g., implying a large degree of regularity, or (2) indicates that the thin, long streams (e.g., GD-1) are on regular orbits.
+Our results provide a clear explanation of how and why the morphology of tidal streams alone can be used to constrain the potential of the host galaxy.
+The mere existence of thin tidal streams in the halo of the Milky Way either (1) provides useful information about the potential on these scales by, e.g., implying a large degree of regularity, or (2) indicates that the thin, long streams (e.g., GD-1) are on regular orbits.
 These are not mutually exclusive --- in fact, if the streams are on regular orbits, this would be a powerful way to check or rule out possible potential models by requiring that the progenitor orbits remain regular.
@@
 For shorter thin streams it will be hard to decouple the unknown evolution time from enhanced density evolution from chaos.
-
-\todo{apw}{Last paragraph that drives home the importance of this work...too tired to finish this tonight...}
+\todo{apw}{A sentence or two on broader implications / applications}
 
@@
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
+This work was partially supported by the National Science Foundation under Grant No.
+AST-1312196.
 This research made use of \project{Astropy}, a community-developed core \project{Python} package for Astronomy \citep{astropy13}.

### 2015-06-26 ms.tex
greyscale figures

@@
 \usepackage{mathrsfs}
-\usepackage{subfigure}
+\usepackage{graphicx} 
 \usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
@@
 \renewcommand{\dateseparator}{-}
-%\usepackage{lscape}
-%\usepackage{rotating}
 
@@
 All methods have been tested in some way on simulated observations of data and these tests typically demonstrate the recovery of parameters for analytic, static potential forms.
-\todo{readers}{This is all just about the methods -- do I need sentences or a paragraph on results?}
-
-%At present most of these methods have been shown to work in axisymmetric or otherwise simple potentials through tests with mock data.
+% internal question: This is all just about the methods -- do I need sentences or a paragraph on results?
 
@@
 \citet{pearson15} tested the stochasticity of the orbit of the progenitor that produced `fanned' debris by computing the Lyapunov exponent along this orbit but found that it is consistent with being regular over dynamically relevant timescales (many Hubble times).
-It has been shown previously that along some strongly chaotic orbits, tidal streams do form large, diffuse `fans' of debris \citep[e.g.,][]{fardal14, ngan15}, however it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
+It has been shown previously that along some strongly chaotic orbits, tidal streams do form large, diffuse `fans' of debris \citep[e.g.,][]{fardal14}, however it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
 The result from \citet{pearson15} suggests that even weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.
@@
 
-\todo{apw}{New figure showing just the grid and orbit classifications, labels for orbit types}
+Figure~\ref{fig:icmap} shows the grid of initial conditions in the $xz$ plane --- each pixel in this grid represents an orbit, and the major orbit families are labeled.
+Points are colored (black / white) whether the orbit acts like a tube orbit or a box orbit over $\approx$50 orbital periods.
+Figure~\ref{fig:icorbits} show sample orbits from each of the main, labeled orbit families of the initial condition grid.
 
 We integrate all orbits in the grid for 10000 orbital periods and use the method described in Section~\ref{sec:lyap} to compute the Lyapunov exponents.
-Figure~\ref{fig:lyapmap} shows the grid of initial conditions (in the $xz$ plane) --- each pixel corresponds to an orbit, and the logarithm of the Lyapunov time (in units of number of orbital periods) is mapped to greyscale intensity.
+Figure~\ref{fig:lyapmap} again shows the grid of initial conditions, but now the logarithm of the Lyapunov time (in units of number of orbital periods) is mapped to greyscale intensity.
 The darker pixels have shorter Lyapunov times and are more chaotic.
@@
 Our results provide a clear explanation of how and why the morphology of tidal streams alone can be used to constrain the potential of the host galaxy.
+The longest thin streams are most valuable for this effort because they have clearly evolved for a long time, but the debris remains compact.
+For shorter thin streams it will be hard to decouple the unknown evolution time from enhanced density evolution from chaos.
 The mere existence of thin tidal streams in the halo of the Milky Way either (1) provides useful information about the potential on these scales by, e.g., implying a large degree of regularity, or (2) indicates that the thin, long streams (e.g., GD-1) are on regular orbits.
 These are not mutually exclusive --- in fact, if the streams are on regular orbits, this would be a powerful way to check or rule out possible potential models by requiring that the progenitor orbits remain regular.
-The longest thin streams are most valuable for this effort because they have clearly evolved for a long time, but the debris remains compact.
-For shorter thin streams it will be hard to decouple the unknown evolution time from enhanced density evolution from chaos.
-\todo{apw}{A sentence or two on broader implications / applications}
+
+ \todo{apw}{A sentence on broader implications / applications -- mapping the orbit structure of the potential by looking at streams}
 
@@
 \includegraphics[width=\textwidth]{figures/potential.pdf}
-\caption{Equipotential contours for the triaxial NFW potential considered in this work.
-There are eight contour levels evenly spaced and linear in the value of the potential.
-\todo{apw}{should I show density as well or instead?}} \label{fig:potential}
+\caption{Equipotential contours for the triaxial NFW potential considered in this work. There are eight contour levels evenly spaced and linear in the value of the potential. } \label{fig:potential}
+\end{center}
+\end{figure*}
+
+% Figure ??
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/ic-map.png}
+\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
+Each pixel in this image represents a single orbit, and the shading (white/black) indicates the behavior of the orbit over 50 orbital periods: if the orbit preserves a sense of rotation around some axis --- that is, the angular momentum around some axis does not change sign --- the orbit is considered a tube orbit (white), otherwise the orbit is considered a box or chaotic orbit (black).
+The regions hosting the major orbit types are indicated by text labels.
+Figure~\ref{fig:icorbits} shows examples of each of these four orbit types.
+} \label{fig:icmap}
+\end{center}
+\end{figure*}
+
+% Figure ??
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/ic-orbits-T.png}
+\caption{ Visualization of the four major regular orbit types in a triaxial potential: outer long-axis tubes (O), inner long-axis tubes (I), short-axis tubes (S), and box or boxlets (B).
+Shown are projections of these sample orbits onto the $xy$ and $yz$ planes and meant only to illustrate the general behavior of these orbit classes.} \label{fig:icorbits}
 \end{center}
@@
 \includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/lyap_map.png}
-\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
-Each pixel in this image represents a single orbit, and the greyscale corresponds to the logarithm of the Lyapunov time (in units of orbital periods).
+\caption{ The same grid of initial conditions as Figure~\ref{fig:icmap}, but now each pixel is colored by the  logarithm of the Lyapunov time (in units of orbital periods).
 Orbits are integrated for a total of 10000 orbital periods.
@@
 \includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/fdiff_map.png}
-\caption{The same grid of initial conditions as Figure~\ref{fig:lyapmap}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
+\caption{The same grid of initial conditions as Figure~\ref{fig:icmap}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
 The frequencies are computed in two consecutive windows, each of which has length equal to 40 orbital periods.
@@
 \includegraphics[width=\textwidth]{figures/ensemble-density.png}
-\caption{ Distributions of the values of the estimated configuration-space density field evaluated at the positions of the ensemble particles after 0, 4, and 16 orbital periods for each of the three orbits of Figures~\ref{fig:nbodysims} and \ref{fig:ensembles}.
+\caption{ Distributions of the values of the estimated configuration-space density field evaluated at the positions of the ensemble particles after 0, 4, and 16 orbital periods for each of the three orbits of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims}.
 The density distribution of the regular orbit ensemble stays compact and decreases slowly.
 The density of the weakly chaotic ensemble initially evolves faster to lower density and has a larger tail to lower densities.
-The strongly chaotic ensemble evolves fast to a state of low density.
-\todo{apw}{add border to histograms?
-get rid of colors (maybe just light to dark grey?)}} \label{fig:ensemble-density}
+The strongly chaotic ensemble evolves fast to a state of low density.} \label{fig:ensemble-density}
 \end{center}
@@
 \caption{This is an alternate visualization of the density evolution of orbit ensembles around the three orbits of Figure~\ref{fig:ensemble-density}, showing only the evolution of the mean density for the orbits (normalized by the initial density of the ensemble).
-\todo{apw}{change linestyles so this survives being printed (-, --, -.)}}
+}
 \label{fig:densities}

### 2015-06-26 ms.tex
citations and references

@@
 %The large-scale, smooth components of galactic gravitational potentials are thought to be triaxial in shape. 
-%Triaxial potentials that match observational constraints and measurements of the radial profile and axis ratios of dark matter halos in cosmological simulations are almost certainly not globally integrable.
-%An appreciable number of chaotic orbits are therefore expected in galactic halos.
-%It is yet unclear whether chaotic regions of phase-space will affect the density evolution of stellar tidal streams, which form in the halos of galaxies from the disruption of stellar systems.
-%% Aims
-%Many numerical methods exist for assessing the degree of chaos for a given orbit (e.g., the Lyapunov exponent, the frequency diffusion rate). 
-%These indicators, which are local to a single orbit, often predict a timescale over which chaos will be dynamically important.
-%In this paper, we aim to compare these timescales to the density evolution of the finite-volume ensembles of orbits.
-%% Methods: 
-%We derp derp derp.
+%Triaxial potentials that match observational constraints and measurements of the radial profile and axis ratios of dark matter halos in cosmological simulations are almost certainly not globally integrable and therefore contain an appreciable number of chaotic orbits.
+%However, estimates of the timescale over which chaos is likely to be important (e.g., the Lyapunov time) suggest that chaos may not be dynamically relevant over physically meaningful times.
+% Aims:
+
+% Methods:
+
 %% Results:
-%We find that mean indicators such as the Lyapunov time or the (long term) frequency diffusion rate of a given ``parent'' orbit are not good predictors for the short-term density evolution of ensembles of orbits around this parent orbit.
-%Instead, we find that the variance of the fundamental frequencies of the ensemble orbits over short timescales is a stronger indicator of the density evolution.
-%% Conclusions:
+% We find that indicators such as the Lyapunov time or the frequency diffusion time of an orbit are not good predictors for the short-time density evolution of ensembles of orbits (e.g., tidal debris) around this parent orbit.
+% We propose a new measure for evaluating whether chaotic diffusion is important for tidal debris with a given spread in orbital properties over a given period of time: the variance of the fundamental frequencies of the ensemble orbits over short timescales is a stronger indicator of the density evolution.
+% Conclusions:
 
@@
 From the Earth's position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii.
-The Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely on non-random orbits and thus contain extra information (e.g., debris stripped from stellar systems).
+The Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely associated with debris stripped from stellar systems and thus contain extra information.
 
@@
 
-One example of stream modeling in a multi-component (static, analytic) potential was done by \citet{pearson15}, who tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} \citep{kuepper12} and N-body models.
-They use the observed SDSS number density of stream stars, a limited number of radial velocities for stream members, and the sky position, distance, radial velocity, and proper motion of the cluster itself to fit model streams to the data.
-In the oblate potential (a three-component bulge+disk+spherical halo potential), a thin model stream is easily found that reproduces the observed stellar density morphology of the stream.
-In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically form large, two-dimensional `fans' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness and curvature of the stream given the observational constraints of the present-day position and velocity of the cluster.
+One example of stream modeling in a multi-component (static, analytic) potential was done by \citet{pearson15}, who aimed to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} \citep{kuepper12} and N-body models.
+They used the observed SDSS number density of stream stars, a limited number of radial velocities for stream members, and the sky position, distance, radial velocity, and proper motion of the cluster itself to fit model streams to the data.
+In the oblate potential (a three-component bulge+disk+spherical halo potential), a thin model stream was easily found that reproduces the observed stellar density morphology of the stream.
+In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically formed large, two-dimensional `fans' of debris near the ends, and no physically reasonable progenitor orbits could be found that reproduced the observed thinness and curvature of the stream given the observational constraints of the present-day position and velocity of the cluster.
 The result in \citet{pearson15} demonstrates that the morphology of a stream alone can rule out a potential.
@@
 
-The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the triaxial potential.
+The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the axisymmetric potential.
 It is well known that the number of degrees of freedom of a potential plays a critical role in determining the orbit structure of the potential --- Hamiltonians with more than two degrees of freedom generically contain significant chaotic regions.
@@
 In a globally integrable potential, the Hamiltonian may be written solely in terms of the actions, $H = H(\boldsymbol{J})$.
-Galactic potentials are almost certainly not globally integrable, but it is useful to understand the orbit structure in integrable systems before extending to more general potentials.
+Galactic potentials are almost certainly not globally integrable but it is useful to understand the orbit structure in integrable systems before extending to more general potentials.
 Only one globally integrable triaxial mass model is known that generates motion that is qualitatively similar to that seen in Galaxies: the St\"ackel or `Perfect Ellipsoid' model \citep[e.g.,][]{kuzmin73, deZeeuw85}.
@@
 The density distributions within dark matter halos formed in cosmological N-body simulations are generically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}.
-With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo \citep[e.g.,][]{dubinski??, butsky15}.
-It is less clear what happens to the outer halo where tidal streams are more readily found.
-%; for this reason, we ask what we would expect to happen to tidal streams if the halo in which they form is strongly triaxial.
+With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo \citep[e.g.,][though in Milky Way-like galaxies, baryonic disks will add non-sphericity to the total potential]{dubinski94,kazantzidis04, bryan13, butsky15}.
+It is less clear what happens to the outer halo.
 
  \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution (dark-matter-only) N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these halos.
-JS02 finds median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{millenium, hmm, butsky15} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
-JS02 finds significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).
+JS02 find median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{bett07, veraciro11, butsky15} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
+JS02 find significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).
 
@@
 
-This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies.
+This potential is a simple (and unrealistic) model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies.
 Realistic galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors \citep{romanowsky98}, radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11,butsky15}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.; see, e.g.,][]{bailin05}.
@@
 This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits), but the most numerous orbits in this grid are the short-axis and outer long-axis tubes.
-Thin tidal streams are thought to form along tube orbits rather than box orbits because of the faster disruption expected for stellar systems on radially plunging orbits which makes them more diffuse.
+Thin tidal streams may preferentially form along tube orbits rather than box orbits because of the faster disruption and debris diffusion expected for stellar systems on radially plunging orbits.
 
@@
 
-\todo{apw}{motivate by saying we need a single measure to characterize ensemble evolution even though density evolution can be complex...can then map entire orbit grid}
-\todo{apw}{These paragraphs need work, but still waiting for numerical results from the cluster...}
-The density distributions do not change shape dramatically, even for the strongly chaotic ensemble; we therefore use the mean density as a metric for studying the density evolution of orbit ensembles.
-Figure~\ref{fig:densities} shows the mean density of the ensemble particles computed at 128 evenly-spaced intervals from the initial ensemble distribution to the distribution after 16 orbital periods.
-It is clear that the density of the regular orbit ensemble evolves slower than the weakly chaotic ensemble, and the strongly chaotic ensemble quickly becomes diffuse.
+The density evolution can be complex, however with a single measure that characterizes the ensemble density evolution we can map the entire grid of orbits used in previous sections (Figure~\ref{fig:icmap}).
+The density distributions do not change shape dramatically, even for the strongly chaotic ensemble.
+Figure~\ref{fig:densities} shows the evolution of the mean density of the ensemble particles computed at 256 evenly-spaced intervals from the initial ensemble distribution to the distribution after 16 orbital periods.
+It is clear that the density of the regular orbit ensemble evolves slower than the weakly chaotic ensemble, and the strongly chaotic ensemble quickly becomes diffuse; the mean density does not capture the full nature of the density evolution is apparently sensitive to the enhanced density evolution observed along the weakly chaotic orbit seen in Figure~\ref{fig:ensemble-density}.
 
 Motivated by the noticeable discrepancy between the regular and weakly chaotic orbit ensembles --- even though the chaotic timescale of the weakly chaotic ensemble parent orbit is thousands of orbital periods --- we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
-For all ensembles, we integrate the orbits for 16 (parent) orbital periods and store the mean density at the end of integration for each.
-Figure~\ref{fig:ensemblemap-meandensity} again shows the grid of initial conditions (Section~\ref{sec:results1}), but now the greyscale indicates the final mean density.
-\todo{apw}{talk about the structure here, compare to freq.
-diff time map -- even though freq.
-diff time are long, can still resolve structure in these maps.}
+For all ensembles, we integrate the orbits for 16 (parent) orbital periods and store the initial and final values of the mean, configuration-space density.
+Figure~\ref{fig:ensemblemap-meandensity} again shows the grid of initial conditions (Section~\ref{sec:results1}, Figure~\ref{fig:icmap}), but now the greyscale indicates the ratio of the final mean density to the initial density.
+Much of the structure that is visible in the frequency diffusion time map (Figure~\ref{fig:freqdiff}) is again visible in this map of the density evolution of orbit ensembles.
+However, it is surprising that these features are present: the chaotic timescales predicted from both the Lyapunov and frequency diffusion times were typically 100s to 1000s of orbital periods for many of these resonance features.
 
@@
 The potential considered in this work is `unrealistic' for the total potential of a Milky-Way-like galaxy in that it is static, smooth, and does not contain baryonic components.
-Simulations of forming galactic disks in cosmological dark matter halos have shown that baryonic feedback and relaxation can significantly change the inner distribution of dark matter and either make the potential more spherical or oblate \citep{many}.
+Simulations of forming galactic disks in cosmological dark matter halos have shown that baryonic feedback and relaxation can significantly change the inner distribution of dark matter and either make the potential more spherical or oblate \citep{dubinski94,kazantzidis04, bryan13, butsky15}.
 However, the significance of baryonic relaxation or of time-dependence, triaxiality, and substructure on shaping the matter distribution within the Milky Way is largely unknown.
@@
 \item \emph{Substructure}: Cosmological simulations predict that dark matter halos are filled with substructure in the form of dark matter subhalos.
-If they exist, these subhalos may account for up to $\approx$1-10\% of the mass of the dark matter \citep{??} and therefore may contribute significantly to and orbit in the large-scale potential of any galaxy.
+If they exist, these subhalos may account for up to $\approx$1-10\% of the mass of the dark matter \citep[e.g.,][]{diemand07} and therefore may contribute significantly to and orbit in the large-scale potential of any galaxy.
 Gravitational scattering due to subhalo interactions has been studied, however in the halos of galaxies where dynamical times are long, the scattering cross-section for \emph{strong} encounters is very small \citep{??}.
@@
 We have considered here a simple triaxial gravitational potential chosen to mimic the median properties of dark matter halos formed in dark-matter-only simulations (or the large-scale properties of halos formed in simulations with baryonic effects).
-We have numerically computed the magnitude of chaos for a large grid of iso-energy orbits in this potential using two independent methods that have been used extensively to classify and characterize chaotic orbits: the Lyapunov exponent and frequency diffusion rate.
+We have numerically computed the magnitude of chaos for a large grid of iso-energy orbits in this potential using two independent methods that have been used extensively to classify and characterize chaotic orbits: 1) the Lyapunov exponent and 2) the frequency diffusion rate.
 From each of these indicators, we compute a timescale over which chaos is likely to be important and find that the majority of orbits have chaotic timescales greater than 100s of orbital periods, however with the frequency diffusion rate we are still able to resolve weak chaos.
@@
 
+Our main conclusions are summarized as follows:
+\begin{enumerate}
+	\item the Lyapunov time and frequency diffusion time are powerful indicators of chaos, but do not capture the importance of small-scale chaotic diffusion for the density evolution of small ensembles of orbits (tidal debris);
+	\item tidal debris becomes diffuse and thus harder to observe on weakly chaotic orbits when the small-scale chaotic diffusion of the fundamental frequencies has a scale comparable to the internal spread in frequencies of the debris;
+	\item the details of the enhanced mixing along weakly chaotic orbits depends on the resonant structure of the potential;
+	\item the variance of the fundamental frequencies of an orbit over a given time window is a strong predictor of the importance of small-scale chaotic frequency diffusion on the resulting morphology of tidal debris.
+\end{enumerate}
+
 Our results provide a clear explanation of how and why the morphology of tidal streams alone can be used to constrain the potential of the host galaxy.
@@
 These are not mutually exclusive --- in fact, if the streams are on regular orbits, this would be a powerful way to check or rule out possible potential models by requiring that the progenitor orbits remain regular.
-
- \todo{apw}{A sentence on broader implications / applications -- mapping the orbit structure of the potential by looking at streams}
 
@@
 \section{Lyapunov exponents} \label{sec:lyapapdx}
-\todo{apw}{This section has not been edited carefully...}
 
@@
 %\end{equation}
-%where $\delta w_i(0)$ are the initial conditions for the deviation vector and $B_{im}$
+%where $\delta w_i(0)$ are the initial conditions for the deviation vector and $B_{im}$ is the solution matrix. 
 % For chaotic orbits, the maximum eigenvalue of the solution matrix to Eq.~\ref{eq:deviate} is positive real, leading to exponential divergence of nearby orbits.
@@
 If one is only interested in characterizing the degree of chaos, computing the full Lyapunov spectrum for an orbit is often not necessary.
-It is often sufficient to compute an estimate of the maximum Lyapunov exponent by estimating the finite-time maximum Lyapunov exponent (FTMLE).
+It is usually sufficient to compute an estimate of the maximum Lyapunov exponent by estimating the finite-time maximum Lyapunov exponent (FTMLE).
 Using the definition of $\boldsymbol{w}$ from Equation~\ref{eq:coords}, consider an orbit that is a small deviation away from the parent orbit, $\boldsymbol{w}' = \boldsymbol{w} + \delta\boldsymbol{w}$.
@@
 \end{equation}
-\citep[see, e.g.,][]{who?}. From this expression, we see that
+\citep[see, e.g.,][]{lichtenberg83,tabor89}. From this expression, we see that
 \begin{equation}
@@
 Numerically computing this quantity is not trivial because (1) obviously the limit to infinity is not possible and (2) the norm of the deviation vector $\|\delta \bs{w}(t)\|$ is expected to increase exponentially for chaotic orbits, leading to nonlinear evolution of the deviation and numerical problems.
-To circumvent these issues, it is sufficient to instead start a nearby orbit with some small initial deviation with norm $\delta_0$, integrate for a sufficiently small amount of time, $\tau$, then renormalize the deviation back to the initial norm (e.g., TODO: cite Bennetin et al.
-1976).
+To circumvent these issues, it is sufficient to instead start a nearby orbit with some small initial deviation with norm $\delta_0$, integrate for a sufficiently small amount of time, $\tau$, then renormalize the deviation back to the initial norm \citep{benettin76}.
 There is no general way to determine $\tau$ except to perform convergence tests.
@@
 \end{equation}
-and the MLE is estimated after a very long time to approximate the limit of Equation~\ref{eq:lmax}.
+where the $t_i$ are the times at which the renormalization occurs and the MLE is estimated after a very long time to approximate the limit of Equation~\ref{eq:lmax}.
 
@@
 \end{align}
-These equations are trivially integrated:
+These equations are easily integrated:
 \begin{align}
@@
 Chaotic orbits with $t_\lambda \gtrsim 1000~\periods$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
-Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S), weakly chaotic (W), and regular (R) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).
-\todo{apw}{Need better markers for the 3 orbits -- don't show up in greyscale (printed)}} \label{fig:lyapmap}
+Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:lyapmap}
 \end{center}
@@
 The frequencies are measured precisely so that small changes in the frequencies can be detected over just $\approx$10s of orbits.
-Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S), weakly chaotic (W), and regular (R) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
+Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
 \end{center}
@@
 \includegraphics[width=\textwidth]{figures/nbody.png}
-\caption{ \todo{apw}{clean this up}
-Length-scale of plummer sphere set to get 10\% mass loss so 1000 particles so ensemble tests reasonable...
-We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on these three orbits.
-The initial mass of the clusters is .
-For each simulation, the progenitor is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
+\caption{ We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on the three orbits of Figure~\ref{fig:ensembles}.
+The progenitor in each simulation is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
 We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter.
-Particle positions from the final snapshots of N-body simulations of a globular-cluster-mass satellite disrupting in a triaxial NFW potential (Section~\ref{sec:potential}).
-Each satellite is evolved for 16 orbital periods, and begins and ends at apocenter.
-For visualization, the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
-Top row shows the positions projected onto the resulting $xy$ plane (the instantaneous orbital plane of the progenitor at the end of the simulation), bottom row shows the positions projected onto the $yz$ plane.
-From left to right, the columns indicate that the progenitor is on a regular (R), weakly chaotic (W), or strongly chaotic (S) orbit.
-Even on the weakly chaotic orbit --- where the predicted chaotic timescale is XX orbital periods --- the resulting tidal stream displays stream-fanning and the `oldest' debris is more diffuse than in the regular case.}
+The mass of the progenitor is set to XX, and the length-scale of the Plummer sphere is set to 20~pc to get $\approx$10\% mass loss over the integration time.
+Panels show particle positions from the final snapshots of the simulations --- for visualization the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
+These simulations confirm that the test-particle orbit ensembles do (qualitatively) capture the nature of the early-stripped tidal debris.
+}
 \label{fig:nbodysims}
@@
 We generate globular-cluster-scale ensembles of orbits around each `parent' orbit in the grid (Section~\ref{sec:results2}) and follow the density of the ensemble as a function of time using kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
-Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S), weakly chaotic (W), and regular (R) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).
-\todo{apw}{remake once new run on the cluster is complete}}
+Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).}
 \label{fig:ensemblemap-meandensity}

### 2015-06-27 ms.tex
more refs

@@
 By modeling the observed phase-space density of stream stars along with the host galaxy potential it is hoped that we may infer the 3D mass distribution of the host.
-Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies; the known streams span a large range of distances --- from $\approx10$ to $100~{\rm kpc}$ --- and progenitor masses --- from $\approx$10$^3$ to $\approx$10$^8~\msun$ in stellar mass --- \citep[][]{cite many}.
+Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies; the known streams span a large range of distances --- from $\approx10$ to $100~{\rm kpc}$ --- and progenitor masses --- from $\approx$10$^3$ to $\approx$10$^8~\msun$ in stellar mass --- \citep[][]{ibata94, belokurov06,grillmair06a,grillmair06b}.
 There has been extensive work on developing methods to use data from these streams to measure properties of Milky Way's dark matter halo.
@@
 If they exist, these subhalos may account for up to $\approx$1-10\% of the mass of the dark matter \citep[e.g.,][]{diemand07} and therefore may contribute significantly to and orbit in the large-scale potential of any galaxy.
-Gravitational scattering due to subhalo interactions has been studied, however in the halos of galaxies where dynamical times are long, the scattering cross-section for \emph{strong} encounters is very small \citep{??}.
+Gravitational scattering due to subhalo interactions has been studied, however in the halos of galaxies where dynamical times are long, the scattering cross-section for \emph{strong} encounters is small.
 Instead, the collective effect of the subhalos may instead act as a noise term in the Hamiltonian of any halo orbit.
@@
 Tidal stream modeling is one of the most promising ways to constrain the 3D mass distribution around the Milky Way at distances of 10s to 100s of kpc.
-Methods that use tidal streams to infer properties of the Galactic potential typically operate by constructing models of the debris distribution using either the present-day phase-space density \citep[e.g.,][]{kuepper12, kuepper15, amorisco and others}, time-of-disruption phase-space density \citep{apw13, apw14}, or the density in angle-action coordinates \citep{sanders14, bovy14}.
+Methods that use tidal streams to infer properties of the Galactic potential typically operate by constructing models of the debris distribution using either the present-day phase-space density \citep[e.g.,][]{kuepper12, kuepper15, gibbons14}, time-of-disruption phase-space density \citep{apw13, apw14}, or the density in angle-action coordinates \citep{sanders14, bovy14}.
 All of these methods may fail or produce uninterpretable results if modeling globular-cluster streams on mildly chaotic orbits.
@@
 
-Our main conclusions are summarized as follows:
+Our main results and conclusions are summarized as follows:
 \begin{enumerate}
+	\item ``stream-fanning'' of tidal streams on weakly chaotic orbits \citep[as seen in simulations by][]{pearson15} occurs due to small-scale chaotic evolution of the fundamental frequencies of the debris star orbits;
 	\item the Lyapunov time and frequency diffusion time are powerful indicators of chaos, but do not capture the importance of small-scale chaotic diffusion for the density evolution of small ensembles of orbits (tidal debris);

### 2015-07-04 ms.tex
andreas and monica comments

@@
 \newcommand{\mich}{3}
-\newcommand{\nyu}{4}
-\newcommand{\cds}{5}
-\newcommand{\mpia}{6}
+\newcommand{\hubble}{4}
+\newcommand{\nyu}{5}
+\newcommand{\cds}{6}
+\newcommand{\mpia}{7}
 \altaffiltext{\colum}{Department of Astronomy, 
@@
 			   Ann Arbor, MI 48109, USA}
+\altaffiltext{\hubble}{Hubble fellow}
 \altaffiltext{\nyu}{Center for Cosmology and Particle Physics,
@@
 There has been extensive work on developing methods to use data from these streams to measure properties of Milky Way's dark matter halo.
-These methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to N-body simulations \citep[see, e.g., the introduction of][]{kuepper15}.
+These methods span a range of complexity from orbit-fitting, to \emph{Streakline} \citep{kuepper12} or `particle-spray' models \citep{TODOCambridgePaper}, to action-space density modeling \citep[e.g.,][]{sanders13, bovy14}, to N-body simulations \citep[e.g.,][]{law10}.
 All methods have been tested in some way on simulated observations of data and these tests typically demonstrate the recovery of parameters for analytic, static potential forms.
+%\citep[see the introduction of][]{kuepper15}
 % internal question: This is all just about the methods -- do I need sentences or a paragraph on results?
@@
 The result from \citet{pearson15} suggests that even weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.
-Understanding why this occurs and developing a measure to quantify the importance of this enhanced density evolution is a promising new direction to be explored further in this work.
+Understanding why this occurs and developing a measure to quantify the importance of this enhanced density evolution is a promising new direction to be explored further below.
 
-In this work, we study the effect of chaotic diffusion of the fundamental frequencies of individual orbits on the density evolution of tidal debris.
+In this \emph{Article}, we study the effect of chaotic diffusion of the fundamental frequencies of individual orbits on the density evolution of tidal debris.
 We choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos as computed from single-orbit diagnostics for grids of constant-energy orbits, and compare these results to measures of the density evolution of finite-volume ensembles of orbits (meant to mimic tidal debris).
@@
 In these coordinates, the position variables are angles, $\boldsymbol{\theta}$, that increase linearly with time with rates set by $N$ constant, fundamental frequencies, $\boldsymbol{\Omega} = (\Omega_1, ..., \Omega_N)$.
-The frequency of a Fourier component in the sum for any individual component of motion (the $\omega_k$ in Equation~\ref{eq:fourier}) are just linear, integer combinations of the fundamental frequencies, $\boldsymbol{\Omega}$ --- that is, for a regular orbit, any Fourier component frequency may be written
+The frequency of a Fourier component (the $\omega_k$ in Equation~\ref{eq:fourier}) for any individual component of motion are just linear, integer combinations of the fundamental frequencies, $\boldsymbol{\Omega}$ --- that is, for a regular orbit, any Fourier component frequency may be written
 \begin{align}
@@
 
-A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally --- for example, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position \citep[e.g.,][]{goldstein80}.
-The condition for global integrability is very restrictive and it seems that the likelihood that a Hamiltonian is globally integrable decreases as the number of degrees of freedom increase \citep[e.g.,][]{lichtenberg83}.
+A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating Integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally --- for example, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position \citep[e.g.,][]{goldstein80}.
+The condition for global integrability is very restrictive and the likelihood that a Hamiltonian is globally integrable decreases as the number of degrees of freedom increase \citep[e.g.,][]{lichtenberg83} and as the potential becomes more complex (i.e.
+containing multiple components with different shapes).
 In a globally integrable potential, the Hamiltonian may be written solely in terms of the actions, $H = H(\boldsymbol{J})$.
@@
 If instead there exists a relation of the form $\boldsymbol{n} \cdot \boldsymbol{\Omega} = 0$ the orbit is referred to as a resonant orbit.
-Resonant orbits are confined to a surface with lower dimensionality than the surface of an orbital torus, depending on the number of resonance relations obeyed: we refer to orbits that obey a single resonance relation as \emph{uni-resonant} orbits, and if an additional resonance relation exists, \emph{bi-resonant}.
+Resonant orbits are confined to a surface with lower dimensionality than the surface of an orbital torus, depending on the number of resonance relations obeyed; in a triaxial potential, orbits may obey either zero, one, or two resonance relations.
+We refer to orbits that obey a single resonance relation as \emph{uni-resonant} orbits, and if an additional resonance relation exists, \emph{bi-resonant}.
 Uni-resonant orbits in a triaxial potential are confined to a 2D surface, and bi-resonant orbits are closed 1D curves.
-The resonant structure of a potential --- the relative importance of particular resonance numbers --- is difficult to compute, but determines the global behavior of orbits in the potential.
-For a triaxial potential the resonances define planes in frequency-space, but resonance surfaces in action-space are generic surfaces.
+The resonant structure of a potential --- the relative importance of particular resonance integer vectors --- is difficult to compute, but determines the global behavior of orbits in the potential.
 In plots of frequency ratios, the resonances appear as lines; Figure~\ref{fig:cartoons} (left panel) shows a cartoon portrait of a portion of frequency-space for an integrable potential with example resonance lines, non-resonant, and resonant orbits marked.
@@
 
-Figure~\ref{fig:cartoons} (middle panel) shows a cartoon of frequency space for a near-integrable potential (a small perturbation away from the potential in the left panel).
+Figure~\ref{fig:cartoons} (middle panel) shows a cartoon of frequency space for a near-integrable potential (a small perturbation away from the potential in the left panel) assuming the resonances are stable.
 Much of the structure that was present in the integrable potential remains in the near-integrable case, but the differences are highlighted.
@@
 If the stochastic layer is small and the chaotic orbit is therefore confined, the orbit may behave nearly regular for long periods of time.
+If a resonance is unstable, all orbits associated with the resonance will be chaotic and this will drive orbits away from the resonance; unstable resonances form linear gaps in frequency-space (this is not shown in the cartoon).
 
 For small values of $\epsilon$, many regular orbits survive and only small chaotic regions are introduced, especially where resonances overlap \citep[see][]{chirikov60}.
-As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed.
-Figure~\ref{fig:cartoons} (right panel) qualitatively shows this phenomenon --- the resonant and resonant-layer orbits may still be regular, but many or all of the conditionally-periodic orbits can be chaotic.
-As the perturbation strength increases, next the uni-resonant tori are destroyed, and finally the bi-resonant tori --- these are least susceptible to destruction from perturbations \cite[for a more quantitative illustration of this transition from integrability to global chaos, see Figure~9 in][]{valluri98}.
+As the strength of the perturbation increases, eventually most tori associated with conditionally-periodic motion will be destroyed.
+Figure~\ref{fig:cartoons} (right panel) qualitatively shows this phenomenon --- the resonant and resonant-layer orbits may still be regular, but many or all of the conditionally-periodic orbits will be chaotic.
+As the perturbation strength increases, eventually the uni- and bi-resonant tori are also destroyed --- these are less susceptible to destruction from perturbations \cite[for a more quantitative illustration of this transition from integrability to global chaos, see Figure~9 in][]{valluri98}.
 
@@
 
-Chaotic orbits have no orbital actions because they are only strictly bound to their energy hypersurface (if the potential is time-independent).
+Chaotic orbits have no orbital actions and only conserve energy (if the potential is time-independent).
 The orbits therefore do not have a single set of fundamental frequencies, but rather the frequencies that describe the character of motion evolve with time.
@@
 For weakly chaotic orbits, motion across a resonance layer can occur with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
-Thus, if the resonance libration frequency is comparable (within a factor of a few) to the orbital frequency, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
+Thus, if the resonance libration frequency is small, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
 However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can not change by large factors (unless there is resonance overlap and the motion is strongly chaotic).
@@
 The density distributions within dark matter halos formed in cosmological N-body simulations are generically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}.
-With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo \citep[e.g.,][though in Milky Way-like galaxies, baryonic disks will add non-sphericity to the total potential]{dubinski94,kazantzidis04, bryan13, butsky15}.
+With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo and the mass and shape of the baryonic component \citep[e.g.,][though in Milky Way-like galaxies, baryonic disks will add non-sphericity to the total potential]{dubinski94,kazantzidis04, bryan13, butsky15}.
 It is less clear what happens to the outer halo.
@@
 		$c$ & 0.55\\
+		\midrule
+		$T_{abc}$ & 0.58\\
 		\bottomrule
@@
 Velocity scale, scale radius, and axis ratios are chosen to match the median halo parameters for a $M_{\rm vir} \approx 10^{12}~\msun$ halo from \citep{jing02}.
+The triaxiality parameter, $T_{abc} = \frac{a^2 - b^2}{a^2 - c^2}$, is also shown.
 \label{tbl:potential}}
@@
 
-In Section~\ref{sec:results1}, we generate grids of orbits in the potential described above to map the orbit structure of the potential.
+In Section~\ref{sec:results1}, we generate grids of orbits in the potential described above to map the orbital structure of the potential.
 We classify each orbit in terms of the strength of chaos along the orbit as computed using the Lyapnov and frequency diffusion times of Section~\ref{sec:methods}.
@@
 Figure~\ref{fig:icmap} shows the grid of initial conditions in the $xz$ plane --- each pixel in this grid represents an orbit, and the major orbit families are labeled.
-Points are colored (black / white) whether the orbit acts like a tube orbit or a box orbit over $\approx$50 orbital periods.
+Points are colored (black / white) depending on whether the orbit acts like a tube orbit or a box orbit over $\approx$50 orbital periods.
 Figure~\ref{fig:icorbits} show sample orbits from each of the main, labeled orbit families of the initial condition grid.
@@
 
-While there is rich structure and a significant number of weakly chaotic orbits, the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods and are thus not expected to be relevant for tidal stream evolution.
+While there is rich structure and a significant number of weakly chaotic orbits, the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods (thousands of Gigayears) and are thus not expected to be relevant for tidal stream evolution.
 In the next section, we analyze the density evolution of finite-volume ensembles of orbits around each orbit in the above grids in order to study the effect of weak chaos on tidal debris.
@@
 
-The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant for a given orbit --- that is, these quantities are measures of how infinitesimal deviations will diverge on average from some parent orbit of interest, or of how long it takes for the frequencies of a single orbit to change by some amount.
+The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant for a given orbit --- that is, these quantities are measures of how infinitesimal deviations will diverge on average from some parent orbit, or of how long it takes for the frequencies of a single orbit to change by some amount.
 Tidal debris is disrupted from progenitor systems with finite spreads in orbital properties (e.g., energy).
@@
 A thin `stream' forms on the regular orbit (left column), a more diffuse `stream' on the mildly chaotic orbit (middle column), and a `fanned' stream on the strongly chaotic orbit (right column).
-Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit, it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream.
+Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit ($t_\lambda \approx 55$, $t_\Omega \approx 8\times10^4$), it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream.
 
@@
 Figure~\ref{fig:densities} shows the evolution of the mean density of the ensemble particles computed at 256 evenly-spaced intervals from the initial ensemble distribution to the distribution after 16 orbital periods.
-It is clear that the density of the regular orbit ensemble evolves slower than the weakly chaotic ensemble, and the strongly chaotic ensemble quickly becomes diffuse; the mean density does not capture the full nature of the density evolution is apparently sensitive to the enhanced density evolution observed along the weakly chaotic orbit seen in Figure~\ref{fig:ensemble-density}.
+It is clear that the density of the regular orbit ensemble evolves slower than the weakly chaotic ensemble, and the strongly chaotic ensemble quickly becomes diffuse; \todo{apw}{WTF?} the mean density does not capture the full nature of the density evolution is apparently sensitive to the enhanced density evolution observed along the weakly chaotic orbit seen in Figure~\ref{fig:ensemble-density}.
 
@@
 Much of the structure that is visible in the frequency diffusion time map (Figure~\ref{fig:freqdiff}) is again visible in this map of the density evolution of orbit ensembles.
-However, it is surprising that these features are present: the chaotic timescales predicted from both the Lyapunov and frequency diffusion times were typically 100s to 1000s of orbital periods for many of these resonance features.
+However, it is surprising that these features are present: the chaotic timescales predicted from both the Lyapunov and frequency diffusion times were typically 100s to 1000s of orbital periods for many of the unstable resonance features (curved lines).
+\todo{apw}{Why don't these features show up?}
 
@@
 If a weakly chaotic orbit is confined (by other nearby, non-overlapping resonances or stable regions), the mean drift of an orbit in frequency space may be small if computed over timescales long compared to the orbital time but short compared to the Arnold diffusion time.
-For the potential considered in this paper, it turns out that these are comparable to the timescales for which the frequency diffusion rate of the previous section were computed ($\approx$100-1000 orbital periods).
-This window length was chosen in order to have a strong enough signal in the frequency spectrum of chaotic orbits to measure the most significant frequencies using \superfreq\, however, it erases the small-scale variation of the frequencies.
+This window length was chosen in order to have a strong enough signal in the frequency spectrum of chaotic orbits to measure the most significant frequencies using \superfreq\,, but it erases the small-scale variation of the frequencies.
+% For the potential considered in this paper, it seems that these are comparable to the timescales for which the frequency diffusion rate of the previous section were computed ($\approx$100-1000 orbital periods).
 
@@
 	A_\Omega = \max_{a_k} \left(\frac{\max_j \Omega^{(j)}_{k} - \min_j \Omega^{(j)}_{k}}{2 \, \mean{\Omega^{(j)}_{k}}_j}\right) \label{eq:frac-freq-amp}
-\end{equation}
+\end{equation}\todo{apw}{change to RMS?}
 where, as in Equation~\ref{eq:fdtime}, the outer maximum is taken with respect to the amplitude of the frequency in the power-spectrum, $a_k$, and the inner maximum and minimum are taken over time, $j$.
@@
 	\item \emph{Baryonic components}: \cite[][D08]{debattista08} and \cite[][V10]{valluri10} studied the orbit evolution induced by growing a baryonic disk in dark matter halos with various shapes and orientations (e.g., prolate, triaxial).
-In general, the authors find that the growth of a disk slowly deforms the orbits of mass tracers (e.g., dark matter particles) and preferentially populates tube orbits.
-Consequently, the inner shape of the potential becomes more oblate or spherical due to the prevalence of tube orbits.
+In general, the authors find that the growth of a disk slowly deforms the orbits of mass tracers (e.g., dark matter particles) and preferentially populates tube or other `round' orbits (i.e.
+even the chaotic and box orbits fill a fairly spherical or oblate volume).
+Consequently, the inner shape of the potential becomes more oblate or spherical.
 If the inner halos of galaxies are indeed close to spherical or oblate, the majority of orbits will be regular and chaos will be less important, however this is far from conclusive.
-We have found from simple experiments in superposing potential components that transition regions between potential shapes can actually enhance significant amounts of chaos.
-Though superposing potentials is not realistic, it is at least suggests that a more careful exploration of potential configurations is required to understand how complex, radius-dependent potential forms affect the amount and significance of chaos in galactic halos.
+We have found from simple experiments in superposing potential components that transition regions between potential shapes can significantly enhance the amount of chaos.
+Though superposing potentials is not realistic, it at least suggests that a more careful exploration of potential configurations is required to understand how complex, radius-dependent potential forms affect the amount and significance of chaos in galactic halos.
 
@@
 The authors wish to thank Robyn Sanderson, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
+
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
@@
 AST-1312196.
+MV is supported in part by the University of MichiganÕs Elizabeth Crosby Fund
+and the Office of the VP for Research and NASA ATP grant NNX15AK79G.
+AHWK would like to acknowledge support from NASA through Hubble Fellowship grant HST-HF-51323.01-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS 5-26555.
+
 This research made use of \project{Astropy}, a community-developed core \project{Python} package for Astronomy \citep{astropy13}.

### 2015-07-06 ms.tex
more refs, hogg comments, andreas comments

@@
 
-\title{Chaotic mixing of tidal debris}
+\title{Chaotic dispersal of tidal debris}
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn},
@@
 	    Sarah Pearson\altaffilmark{\colum},
-	    Andreas H. W. K\"upper\altaffilmark{\colum},
+	    Andreas H. W. K\"upper\altaffilmark{\colum,\hubble},
 	    David W. Hogg\altaffilmark{\nyu,\cds,\mpia}}
@@
 
-\todo{apw}{write the abstract...}
-
-% Context
-%The large-scale, smooth components of galactic gravitational potentials are thought to be triaxial in shape. 
-%Triaxial potentials that match observational constraints and measurements of the radial profile and axis ratios of dark matter halos in cosmological simulations are almost certainly not globally integrable and therefore contain an appreciable number of chaotic orbits.
-%However, estimates of the timescale over which chaos is likely to be important (e.g., the Lyapunov time) suggest that chaos may not be dynamically relevant over physically meaningful times.
-% Aims:
-
-% Methods:
-
-%% Results:
-% We find that indicators such as the Lyapunov time or the frequency diffusion time of an orbit are not good predictors for the short-time density evolution of ensembles of orbits (e.g., tidal debris) around this parent orbit.
-% We propose a new measure for evaluating whether chaotic diffusion is important for tidal debris with a given spread in orbital properties over a given period of time: the variance of the fundamental frequencies of the ensemble orbits over short timescales is a stronger indicator of the density evolution.
-% Conclusions:
+Several long, dynamically cold stellar streams have been observed around the Milky Way Galaxy, presumably formed from the tidal disruption of globular clusters.
+In integrable potentials---where all orbits are dynamically regular---tidal debris phase-mixes close to the orbit of the progenitor system. 
+However, cosmological simulations of structure formation suggest that the Milky Way's dark matter halo is expected not to be fully integrable; an appreciable fraction of orbits will be chaotic.
+This paper examines the influence of chaos on the phase-space morphology of cold tidal streams.
+We find very stark results:
+Streams in chaotic regions look very different from those in regular regions.
+We find that simulated streams (simulated using test particle ensembles of nearby orbits) are sensitive to chaos on a much shorter time-scale than any standard prediction (from the Lyapunov or frequency-diffusion times).
+For example, on a mildly chaotic orbit with a chaotic timescale predicted to be 50 to 100 orbital periods ($\approx$50 to 100 Gyr), the resulting stellar stream is, after just 16 orbits, substantially more diffuse than any formed on a nearby but regular orbit.
+We find that the enhanced diffusion of the stream stars can be estimated by looking at the variance in orbital frequencies along the parent (progenitor) orbit.
+Our results suggest that long, cold streams around our Galaxy must exist only on regular (or very nearly regular) orbits; they potentially provide a map of the regular regions of the Milky Way potential.
+This suggests a promising new direction for the use of tidal streams to constrain the distribution of dark matter around our Galaxy.
 
@@
 
-The dark matter halos of galaxies are thought to be triaxial in shape.
+The dark-matter haloes of galaxies are thought to be triaxial in shape with axis ratios and alignments that depend on radius.
 However, despite suggestive evidence from a range of complimentary observational methods, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively verified.
-Around other galaxies, it is generally hard to measure the 3D mass profile because they are seen in projection.
+Around other galaxies, it is generally hard to measure the 3D mass profile because informative tracers are rare and the haloes are seen in projection.
 From the Earth's position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii.
-The Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely associated with debris stripped from stellar systems and thus contain extra information.
+The Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely associated with debris stripped from stellar systems and thus contain additional information encoded by the formation mechanism.
 
-As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded due to the tidal forces of the host galaxy potential.
-Along regular, mildly-eccentric orbits, mass is disrupted with small spreads in orbital properties (e.g., energy, angular momentum).
+As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded by tidal forces from the host-galaxy potential.
+Along regular, mildly eccentric orbits, mass is disrupted with small spreads in orbital properties (e.g., energy, angular momentum).
 Once the debris has evolved far enough from the progenitor system that the self-gravity of the progenitor can be ignored (usually a fast process relative to the orbital time), the stars evolve essentially as an ensemble of test particle orbits in the potential of the host system.
 The debris remains coherent as a tidal stream if the phase-mixing time-scale is long: a small ensemble of regular orbits reaches a fully phase-mixed state in a timescale $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in fundamental frequencies of the ensemble (tidal debris from a globular cluster typically has frequency spreads $\approx$0.1--1\%, so it can take hundreds to thousands of orbital periods to fully phase-mix).
-The ensemble spreads due to sheering from slight variations in their fundamental frequencies which, for tidal streams, preferentially occurs along one dimension \citep{merritt96, helmi99}.
+The ensemble spreads due to shearing from slight variations in their fundamental frequencies, which, for tidal streams, preferentially occurs along one dimension \citep{merritt96, helmi99}.
 
-The morphological (density) evolution of the tidal debris therefore depends only on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also determined by the shape and radial profile of the gravitational potential of the host galaxy.
+The morphological (density) evolution of the tidal debris therefore depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also determined by the shape and radial profile of the gravitational potential of the host galaxy.
 By modeling the observed phase-space density of stream stars along with the host galaxy potential it is hoped that we may infer the 3D mass distribution of the host.
-Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies; the known streams span a large range of distances --- from $\approx10$ to $100~{\rm kpc}$ --- and progenitor masses --- from $\approx$10$^3$ to $\approx$10$^8~\msun$ in stellar mass --- \citep[][]{ibata94, belokurov06,grillmair06a,grillmair06b}.
+Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies; the known streams span a large range of distances --- from $\approx10$ to $100~{\rm kpc}$ --- and progenitor masses --- from $\approx$10$^3$ to $\approx$10$^8~\msun$ in stellar mass --- \citep[][]{ibata94,odenkirchen01,belokurov06,grillmair06a,grillmair06b,bonaca12}.
 There has been extensive work on developing methods to use data from these streams to measure properties of Milky Way's dark matter halo.
-These methods span a range of complexity from orbit-fitting, to \emph{Streakline} \citep{kuepper12} or `particle-spray' models \citep{TODOCambridgePaper}, to action-space density modeling \citep[e.g.,][]{sanders13, bovy14}, to N-body simulations \citep[e.g.,][]{law10}.
+These methods span a range of complexity from orbit-fitting, to \emph{Streakline} \citep{kuepper12} or `particle-spray' models \citep{gibbons14}, to action-space density modeling \citep[e.g.,][]{sanders14, bovy14}, to N-body simulations \citep[e.g.,][]{law10}.
 All methods have been tested in some way on simulated observations of data and these tests typically demonstrate the recovery of parameters for analytic, static potential forms.
@@
 
-One example of stream modeling in a multi-component (static, analytic) potential was done by \citet{pearson15}, who aimed to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} \citep{kuepper12} and N-body models.
-They used the observed SDSS number density of stream stars, a limited number of radial velocities for stream members, and the sky position, distance, radial velocity, and proper motion of the cluster itself to fit model streams to the data.
+One example of stream modeling in a multi-component (static, analytic) potential was presented by \citet{pearson15}, who aimed to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} \citep{kuepper12} and N-body models.
+They used the observed number density of stream stars, a limited number of radial velocities for stream members, and the sky position, distance, radial velocity, and proper motion of the cluster itself to fit model streams to the data.
 In the oblate potential (a three-component bulge+disk+spherical halo potential), a thin model stream was easily found that reproduces the observed stellar density morphology of the stream.
 In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically formed large, two-dimensional `fans' of debris near the ends, and no physically reasonable progenitor orbits could be found that reproduced the observed thinness and curvature of the stream given the observational constraints of the present-day position and velocity of the cluster.
-The result in \citet{pearson15} demonstrates that the morphology of a stream alone can rule out a potential.
+The result in \citet{pearson15} demonstrates that the morphology of a stream alone can be used to rule out a potential.
 With an understanding of the circumstances that lead to the differences in stream morphology, this could be a powerful tool for rejecting potentials from positional information alone.
@@
 The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the axisymmetric potential.
-It is well known that the number of degrees of freedom of a potential plays a critical role in determining the orbit structure of the potential --- Hamiltonians with more than two degrees of freedom generically contain significant chaotic regions.
+It is well known that the number of degrees of freedom of a potential plays a critical role in determining the orbit structure of the potential; Hamiltonians with more than two degrees of freedom generically contain significant chaotic regions.
 \citet{pearson15} tested the stochasticity of the orbit of the progenitor that produced `fanned' debris by computing the Lyapunov exponent along this orbit but found that it is consistent with being regular over dynamically relevant timescales (many Hubble times).
@@
 We choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos as computed from single-orbit diagnostics for grids of constant-energy orbits, and compare these results to measures of the density evolution of finite-volume ensembles of orbits (meant to mimic tidal debris).
-We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles due to the chaotic diffusion of the individual orbits.
-For a chaotic orbit, the frequency spectrum of the orbit evolves with time --- for a small ensemble, the spread in frequencies is therefore time-dependent, which could enhance phase-mixing.
-This idea supports a reevaluation of the importance of chaos in galactic halos and implies that the amount of chaos in a given potential may have significant consequences for the observability and survivability of thin, cold tidal streams.
+We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles through the chaotic diffusion of the individual orbits.
+For a chaotic orbit, the frequency spectrum of the orbit evolves with time: For a small ensemble, the spread in frequencies is therefore time-dependent, which could enhance phase-mixing.
+This idea supports a reevaluation of the importance of chaos in galactic haloes and implies that the amount of chaos in a given potential may have significant consequences for the observability and survivability of thin, cold tidal streams.
 We propose a new measure for determining whether this small-scale frequency diffusion will be important for the morphological evolution of tidal debris with a given spread in frequencies or orbital properties over finite and dynamically relevant times.
@@
 \begin{equation}
-	w_i(t) = \sum_k^\infty a_k \, e^{i\,\omega_k\,t} \label{eq:fourier}
+	w_i(t) = \sum_j^\infty a_j \, e^{i\,\omega_j\,t} \label{eq:fourier}
 \end{equation}
-where the $a_k$ are complex amplitudes.
+where the $a_j$ are complex amplitudes.
 
@@
 \begin{align}
-	\omega_k &= \boldsymbol{n_k} \cdot \boldsymbol{\Omega} \label{eq:fourierfreq}
+	\omega_j &= \boldsymbol{n_j} \cdot \boldsymbol{\Omega} \label{eq:fourierfreq}
 \end{align} %; \, (\boldsymbol{n}_k \in \mathbb{Z}^3)
@@
 
-The density distributions within dark matter halos formed in cosmological N-body simulations are generically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}.
+The density distributions within dark-matter haloes formed in cosmological N-body simulations are generically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}.
 With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo and the mass and shape of the baryonic component \citep[e.g.,][though in Milky Way-like galaxies, baryonic disks will add non-sphericity to the total potential]{dubinski94,kazantzidis04, bryan13, butsky15}.
@@
 
- \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution (dark-matter-only) N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these halos.
+ \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution (dark-matter-only) N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these haloes.
 JS02 find median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{bett07, veraciro11, butsky15} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
@@
 Velocity scale, scale radius, and axis ratios are chosen to match the median halo parameters for a $M_{\rm vir} \approx 10^{12}~\msun$ halo from \citep{jing02}.
-The triaxiality parameter, $T_{abc} = \frac{a^2 - b^2}{a^2 - c^2}$, is also shown.
+The triaxiality parameter, $T_{abc} = \frac{a^2 - b^2}{a^2 - c^2}$, is also given.
 \label{tbl:potential}}
@@
 
-This potential is a simple (and unrealistic) model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies.
+This potential is a simple (and unrealistic) model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the haloes of such galaxies.
 Realistic galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors \citep{romanowsky98}, radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11,butsky15}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.; see, e.g.,][]{bailin05}.
@@
 
-Figure~\ref{fig:icmap} shows the grid of initial conditions in the $xz$ plane --- each pixel in this grid represents an orbit, and the major orbit families are labeled.
+%Figure~\ref{fig:icmap} shows the grid of initial conditions in the $xz$ plane --- each pixel in this grid represents an orbit, and the major orbit families are labeled.
 Points are colored (black / white) depending on whether the orbit acts like a tube orbit or a box orbit over $\approx$50 orbital periods.
@@
 
-The density evolution can be complex, however with a single measure that characterizes the ensemble density evolution we can map the entire grid of orbits used in previous sections (Figure~\ref{fig:icmap}).
+The density evolution can be complex, however with a single measure that characterizes the ensemble density evolution we can map the entire grid of orbits used in previous sections (e.g., Figure~\ref{fig:lyapmap}).
 The density distributions do not change shape dramatically, even for the strongly chaotic ensemble.
@@
 For all ensembles, we integrate the orbits for 16 (parent) orbital periods and store the initial and final values of the mean, configuration-space density.
-Figure~\ref{fig:ensemblemap-meandensity} again shows the grid of initial conditions (Section~\ref{sec:results1}, Figure~\ref{fig:icmap}), but now the greyscale indicates the ratio of the final mean density to the initial density.
+Figure~\ref{fig:ensemblemap-meandensity} again shows the grid of initial conditions (Section~\ref{sec:results1}, but now the greyscale indicates the ratio of the final mean density to the initial density.
 Much of the structure that is visible in the frequency diffusion time map (Figure~\ref{fig:freqdiff}) is again visible in this map of the density evolution of orbit ensembles.
@@
 The potential considered in this work is `unrealistic' for the total potential of a Milky-Way-like galaxy in that it is static, smooth, and does not contain baryonic components.
-Simulations of forming galactic disks in cosmological dark matter halos have shown that baryonic feedback and relaxation can significantly change the inner distribution of dark matter and either make the potential more spherical or oblate \citep{dubinski94,kazantzidis04, bryan13, butsky15}.
+Simulations of forming galactic disks in cosmological dark-matter haloes have shown that baryonic feedback and relaxation can significantly change the inner distribution of dark matter and either make the potential more spherical or oblate \citep{dubinski94,kazantzidis04, bryan13, butsky15}.
 However, the significance of baryonic relaxation or of time-dependence, triaxiality, and substructure on shaping the matter distribution within the Milky Way is largely unknown.
@@
 \begin{itemize}
-	\item \emph{Baryonic components}: \cite[][D08]{debattista08} and \cite[][V10]{valluri10} studied the orbit evolution induced by growing a baryonic disk in dark matter halos with various shapes and orientations (e.g., prolate, triaxial).
+	\item \emph{Baryonic components}: \cite[][D08]{debattista08} and \cite[][V10]{valluri10} studied the orbit evolution induced by growing a baryonic disk in dark-matter haloes with various shapes and orientations (e.g., prolate, triaxial).
 In general, the authors find that the growth of a disk slowly deforms the orbits of mass tracers (e.g., dark matter particles) and preferentially populates tube or other `round' orbits (i.e.
@@
 Consequently, the inner shape of the potential becomes more oblate or spherical.
-If the inner halos of galaxies are indeed close to spherical or oblate, the majority of orbits will be regular and chaos will be less important, however this is far from conclusive.
+If the inner haloes of galaxies are indeed close to spherical or oblate, the majority of orbits will be regular and chaos will be less important, however this is far from conclusive.
 We have found from simple experiments in superposing potential components that transition regions between potential shapes can significantly enhance the amount of chaos.
-Though superposing potentials is not realistic, it at least suggests that a more careful exploration of potential configurations is required to understand how complex, radius-dependent potential forms affect the amount and significance of chaos in galactic halos.
+Though superposing potentials is not realistic, it at least suggests that a more careful exploration of potential configurations is required to understand how complex, radius-dependent potential forms affect the amount and significance of chaos in galactic haloes.
 
 \item \emph{Time dependence}: Galaxies are certainly not static systems.
-To first order, galaxies grow in mass --- for example, the spherically averaged mass profile of dark matter halos evolves fairly predictably in cosmological simulations \citep{wechsler02, buist14} after some initial period of stochastic mass growth.
+To first order, galaxies grow in mass --- for example, the spherically averaged mass profile of dark-matter haloes evolves fairly predictably in cosmological simulations \citep{wechsler02, buist14} after some initial period of stochastic mass growth.
 The Milky Way has probably had a fairly calm accretion history over the last 6 Gyr and therefore the mass growth may be similar to that seen in simulations.
 This steady growth most likely does not alter the global structure or shape of the potential.
-However, we also know from simulations that figure rotation, baryonic feedback, and the accretion and phase-mixing of subhalos do perturb the global state of simulated halos.
+However, we also know from simulations that figure rotation, baryonic feedback, and the accretion and phase-mixing of subhaloes do perturb the global state of simulated haloes.
 \cite{deibel11} showed that by adopting pattern speeds comparable to those found in cosmological simulations, figure rotation generally acts to destabilize orbits (rather than stabilize chaotic orbits in the equivalent static potential) and the resulting orbit structure is that most regular orbits are associated with resonances.
@@
 
-\item \emph{Substructure}: Cosmological simulations predict that dark matter halos are filled with substructure in the form of dark matter subhalos.
-If they exist, these subhalos may account for up to $\approx$1-10\% of the mass of the dark matter \citep[e.g.,][]{diemand07} and therefore may contribute significantly to and orbit in the large-scale potential of any galaxy.
-Gravitational scattering due to subhalo interactions has been studied, however in the halos of galaxies where dynamical times are long, the scattering cross-section for \emph{strong} encounters is small.
-Instead, the collective effect of the subhalos may instead act as a noise term in the Hamiltonian of any halo orbit.
-This subhalo-induced colored noise --- which depends on the mass spectrum and distribution of subhalos --- may also act to simply increase the magnitude of chaos along orbits, and destabilize sufficiently non-resonant orbits \citep[see, e.g.,][]{kandrup00}.
+\item \emph{Substructure}: Cosmological simulations predict that dark-matter haloes are filled with substructure in the form of dark matter subhaloes.
+If they exist, these subhaloes may account for up to $\approx$1-10\% of the mass of the dark matter \citep[e.g.,][]{diemand07} and therefore may contribute significantly to and orbit in the large-scale potential of any galaxy.
+Gravitational scattering due to subhalo interactions has been studied, however in the haloes of galaxies where dynamical times are long, the scattering cross-section for \emph{strong} encounters is small.
+Instead, the collective effect of the subhaloes may instead act as a noise term in the Hamiltonian of any halo orbit.
+This subhalo-induced colored noise --- which depends on the mass spectrum and distribution of subhaloes --- may also act to simply increase the magnitude of chaos along orbits, and destabilize sufficiently non-resonant orbits \citep[see, e.g.,][]{kandrup00}.
 
@@
 
-We have considered here a simple triaxial gravitational potential chosen to mimic the median properties of dark matter halos formed in dark-matter-only simulations (or the large-scale properties of halos formed in simulations with baryonic effects).
+We have considered here a simple triaxial gravitational potential chosen to mimic the median properties of dark-matter haloes formed in dark-matter-only simulations (or the large-scale properties of haloes formed in simulations with baryonic effects).
 We have numerically computed the magnitude of chaos for a large grid of iso-energy orbits in this potential using two independent methods that have been used extensively to classify and characterize chaotic orbits: 1) the Lyapunov exponent and 2) the frequency diffusion rate.
@@
 
-% Figure ??
-\begin{figure*}[p]
-\begin{center}
-\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/ic-map.png}
-\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
+%% Figure ??
+%\begin{figure*}[p]
+%\begin{center}
+%\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/ic-map.png}
+%\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
 Each pixel in this image represents a single orbit, and the shading (white/black) indicates the behavior of the orbit over 50 orbital periods: if the orbit preserves a sense of rotation around some axis --- that is, the angular momentum around some axis does not change sign --- the orbit is considered a tube orbit (white), otherwise the orbit is considered a box or chaotic orbit (black).
@@
 } \label{fig:icmap}
-\end{center}
-\end{figure*}
-
-% Figure ??
-\begin{figure*}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/ic-orbits-T.png}
-\caption{ Visualization of the four major regular orbit types in a triaxial potential: outer long-axis tubes (O), inner long-axis tubes (I), short-axis tubes (S), and box or boxlets (B).
+%\end{center}
+%\end{figure*}
+%
+%% Figure ??
+%\begin{figure*}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/ic-orbits-T.png}
+%\caption{ Visualization of the four major regular orbit types in a triaxial potential: outer long-axis tubes (O), inner long-axis tubes (I), short-axis tubes (S), and box or boxlets (B).
 Shown are projections of these sample orbits onto the $xy$ and $yz$ planes and meant only to illustrate the general behavior of these orbit classes.} \label{fig:icorbits}
-\end{center}
-\end{figure*}
+%\end{center}
+%\end{figure*}
 
@@
 \includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/lyap_map.png}
-\caption{ The same grid of initial conditions as Figure~\ref{fig:icmap}, but now each pixel is colored by the  logarithm of the Lyapunov time (in units of orbital periods).
+\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
+Each pixel in this image represents a single orbit, and each pixel is colored by the  logarithm of the Lyapunov time (in units of orbital periods).
 Orbits are integrated for a total of 10000 orbital periods.
@@
 \includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/fdiff_map.png}
-\caption{The same grid of initial conditions as Figure~\ref{fig:icmap}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
+\caption{The same grid of initial conditions as Figure~\ref{fig:lyapmap}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
 The frequencies are computed in two consecutive windows, each of which has length equal to 40 orbital periods.

### 2015-07-06 ms.tex
added todo

@@
 We then study the density evolution of small ensembles of orbits generated around each orbit in the grid used for the previous experiment and find that along some orbits classified as weakly chaotic --- with chaotic timescales of 100s of orbital periods --- the orbit ensembles display enhanced density evolution and reach a lower overall density faster than orbit ensembles around nearby regular orbits (which mix due to phase-mixing alone).
-We explain this apparent discrepancy by considering the nature of chaotic diffusion: the classical chaos indicators are most sensitive to the slow, Arnold diffusion process that can cause large changes to orbital properties, but small-scale frequency evolution occurs over much shorter times as chaotic orbits stochastically diffuse across the stochastic layers that surround many resonances.
+We explain this apparent discrepancy \todo{apw}{Hogg's comment is to clearly state the discrepancy (has to do with timescales)} by considering the nature of chaotic diffusion: the classical chaos indicators are most sensitive to the slow, Arnold diffusion process that can cause large changes to orbital properties, but small-scale frequency evolution occurs over much shorter times as chaotic orbits stochastically diffuse across the stochastic layers that surround many resonances.
 We propose a new measure of chaos with tidal stream morphological evolution in mind: the amplitude of small-scale, fractional changes in frequencies over a given number of orbital periods along a stream progenitor orbit.

### 2015-07-07 ms.tex
buncha stuff

@@
 \usepackage{graphicx} 
+\usepackage[caption=false]{subfig}
 \usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
@@
 We find that simulated streams (simulated using test particle ensembles of nearby orbits) are sensitive to chaos on a much shorter time-scale than any standard prediction (from the Lyapunov or frequency-diffusion times).
-For example, on a mildly chaotic orbit with a chaotic timescale predicted to be 50 to 100 orbital periods ($\approx$50 to 100 Gyr), the resulting stellar stream is, after just 16 orbits, substantially more diffuse than any formed on a nearby but regular orbit.
+For example, on a mildly chaotic orbit with a chaotic timescale predicted to be $>$1000 orbital periods ($>$1000 Gyr), the resulting stellar stream is, after just 16 orbits, substantially more diffuse than any formed on a nearby but regular orbit.
 We find that the enhanced diffusion of the stream stars can be estimated by looking at the variance in orbital frequencies along the parent (progenitor) orbit.
@@
 By modeling the observed phase-space density of stream stars along with the host galaxy potential it is hoped that we may infer the 3D mass distribution of the host.
-Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies; the known streams span a large range of distances --- from $\approx10$ to $100~{\rm kpc}$ --- and progenitor masses --- from $\approx$10$^3$ to $\approx$10$^8~\msun$ in stellar mass --- \citep[][]{ibata94,odenkirchen01,belokurov06,grillmair06a,grillmair06b,bonaca12}.
+Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies; the known streams span a large range of distances---from $\approx10$ to $100~{\rm kpc}$---and progenitor masses---from $\approx$10$^3$ to $\approx$10$^8~\msun$ in stellar mass---\citep[][]{ibata94,odenkirchen01,belokurov06,grillmair06a,grillmair06b,bonaca12}.
 There has been extensive work on developing methods to use data from these streams to measure properties of Milky Way's dark matter halo.
@@
 In these coordinates, the position variables are angles, $\boldsymbol{\theta}$, that increase linearly with time with rates set by $N$ constant, fundamental frequencies, $\boldsymbol{\Omega} = (\Omega_1, ..., \Omega_N)$.
-The frequency of a Fourier component (the $\omega_k$ in Equation~\ref{eq:fourier}) for any individual component of motion are just linear, integer combinations of the fundamental frequencies, $\boldsymbol{\Omega}$ --- that is, for a regular orbit, any Fourier component frequency may be written
+The frequency of a Fourier component (the $\omega_k$ in Equation~\ref{eq:fourier}) for any individual component of motion are just linear, integer combinations of the fundamental frequencies, $\boldsymbol{\Omega}$---that is, for a regular orbit, any Fourier component frequency may be written
 \begin{align}
@@
 where $\boldsymbol{n}_k$ is a vector of $N$ integers.
-The conjugate momentum coordinates --- the actions,  $\boldsymbol{J}$ --- are constants of motion.
+The conjugate momentum coordinates---the actions,  $\boldsymbol{J}$---are constants of motion.
 Even stronger, the actions are isolating integrals and any pair are in involution such that
@@
 
-A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating Integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally --- for example, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position \citep[e.g.,][]{goldstein80}.
+A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating Integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally---for example, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position \citep[e.g.,][]{goldstein80}.
 The condition for global integrability is very restrictive and the likelihood that a Hamiltonian is globally integrable decreases as the number of degrees of freedom increase \citep[e.g.,][]{lichtenberg83} and as the potential becomes more complex (i.e.
@@
 
-The frequencies of a generic orbit are typically incommensurable --- that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} These conditionally-periodic orbits uniformly cover the surface of an orbital torus.
+The frequencies of a generic orbit are typically incommensurable---that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} These conditionally-periodic orbits uniformly cover the surface of an orbital torus.
 If instead there exists a relation of the form $\boldsymbol{n} \cdot \boldsymbol{\Omega} = 0$ the orbit is referred to as a resonant orbit.
@@
 Uni-resonant orbits in a triaxial potential are confined to a 2D surface, and bi-resonant orbits are closed 1D curves.
-The resonant structure of a potential --- the relative importance of particular resonance integer vectors --- is difficult to compute, but determines the global behavior of orbits in the potential.
+The resonant structure of a potential---the relative importance of particular resonance integer vectors---is difficult to compute, but determines the global behavior of orbits in the potential.
 In plots of frequency ratios, the resonances appear as lines; Figure~\ref{fig:cartoons} (left panel) shows a cartoon portrait of a portion of frequency-space for an integrable potential with example resonance lines, non-resonant, and resonant orbits marked.
@@
 
-The orbit structure of near-integrable potentials can be understood by considering a Hamiltonian that is a small perturbation away from being globally integrable --- that is, a Hamiltonian that may be written
+The orbit structure of near-integrable potentials can be understood by considering a Hamiltonian that is a small perturbation away from being globally integrable---that is, a Hamiltonian that may be written
 \begin{equation}
@@
 Chaotic orbits are characterized by exhibiting random behavior despite being generated by entirely deterministic equations of motion.
-Actions strictly do not exist for chaotic orbits and therefore chaos `destroys' orbital tori in certain regions of action-space --- these frequency spectrum of orbits in these regions evolve chaotically with time.
+Actions strictly do not exist for chaotic orbits and therefore chaos `destroys' orbital tori in certain regions of action-space---these frequency spectrum of orbits in these regions evolve chaotically with time.
 
@@
 As the strength of the perturbation increases, eventually most tori associated with conditionally-periodic motion will be destroyed.
-Figure~\ref{fig:cartoons} (right panel) qualitatively shows this phenomenon --- the resonant and resonant-layer orbits may still be regular, but many or all of the conditionally-periodic orbits will be chaotic.
-As the perturbation strength increases, eventually the uni- and bi-resonant tori are also destroyed --- these are less susceptible to destruction from perturbations \cite[for a more quantitative illustration of this transition from integrability to global chaos, see Figure~9 in][]{valluri98}.
+Figure~\ref{fig:cartoons} (right panel) qualitatively shows this phenomenon---the resonant and resonant-layer orbits may still be regular, but many or all of the conditionally-periodic orbits will be chaotic.
+As the perturbation strength increases, eventually the uni- and bi-resonant tori are also destroyed---these are less susceptible to destruction from perturbations \cite[for a more quantitative illustration of this transition from integrability to global chaos, see Figure~9 in][]{valluri98}.
 
 When $\epsilon$ is large, there is no general prediction for the resulting behavior, however it seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep[e.g.,][]{valluri98, merritt99}.
-We therefore expect a large number of regular orbits will survive --- the so-called KAM tori --- however the tori that survive will be separated by regions of chaotic motion.
+We therefore expect a large number of regular orbits will survive---the so-called KAM tori---however the tori that survive will be separated by regions of chaotic motion.
 Any transformations to angle-action coordinates must be defined local to each resonance region due to the destruction of tori and chaotic motion which lead to discontinuous changes in orbital properties.
@@
 The motion in any canonical coordinate can therefore be decomposed as a Fourier sum (Equation~\ref{eq:fourier}) where the Fourier frequencies are linear, integer combinations of the fundamental frequencies (Equation~\ref{eq:fourierfreq}).
-\cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$).
-It has been shown that the accuracy in determining the frequencies converges much faster using a filtered FFT compared to the typical $\inttime^{-1}$ expected for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time.
-The FFT is iteratively searched for the strongest frequency component, which is subtracted from the spectrum after each iteration.
-This procedure generates a table of frequencies for each component of motion which must then be searched for the three fundamental frequencies.
-
+\cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses fast-Fourier transforms (FFTs) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$) to identify the frequencies.
 This method is referred to as `Numerical Approximation of Fundamental Frequencies' (NAFF) and has been used extensively in planetary dynamics \citep[e.g.,][]{laskar93b, laskar96} and galaxy dynamics \citep{papaphilippou98, valluri98}, especially in the study of orbits in triaxial systems.
 We have implemented and tested a version of this procedure in the \project{Python} programming language.
-Our implementation differs slightly from the original definition and from that used in \cite{valluri98}: we have found that using a higher order window function \citep[e.g.,][]{hunter02} of the form
-\begin{equation}
-	W(\tau=t/\inttime) = \frac{2^p \, (p!)^2}{(2p)!} \left( 1 + \cos \pi \tau\right)^p
-\end{equation}
-with $p=2$ allows for more reliable determination of frequencies from strongly chaotic orbits.
-We refer to this slight modification of the algorithm as \superfreq\ \citep{superfreq} and the code is open-source and publicly available on \project{GitHub} at \url{https://github.com/adrn/SuperFreq}.
-For more details about the algorithm, see Appendix~\ref{sec:naffapdx} or \cite{laskar88, laskar93, papaphilippou96}.
+Our implementation differs slightly from the original definition and from that used in \cite{valluri98}; we refer to this slight modification of the algorithm as \superfreq\ \citep{superfreq} and the code is open-source and publicly available on \project{GitHub} at \url{https://github.com/adrn/SuperFreq}.
+For more details about the algorithm and differences with previous work, see \cite{laskar88, laskar93, papaphilippou96} and Appendix~\ref{sec:naffapdx}.
 
@@
 
-%Figure~\ref{fig:icmap} shows the grid of initial conditions in the $xz$ plane --- each pixel in this grid represents an orbit, and the major orbit families are labeled.
+%Figure~\ref{fig:icmap} shows the grid of initial conditions in the $xz$ plane---each pixel in this grid represents an orbit, and the major orbit families are labeled.
 Points are colored (black / white) depending on whether the orbit acts like a tube orbit or a box orbit over $\approx$50 orbital periods.
@@
 The darker pixels have shorter frequency diffusion times and are more chaotic.
-This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential and highlights the accuracy of \superfreq\ --- weak chaos (grey) is detectable over just 10's of orbital periods, compared to the many thousands of orbits it would take to detect such features with the maximum Lyapunov exponent.
-The tube orbits in this potential are mostly regular or only mildly chaotic --- the largest regular regions are associated with the short-axis and long-axis tube orbits --- however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.
+This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential and highlights the accuracy of \superfreq\---weak chaos (grey) is detectable over just 10's of orbital periods, compared to the many thousands of orbits it would take to detect such features with the maximum Lyapunov exponent.
+The tube orbits in this potential are mostly regular or only mildly chaotic---the largest regular regions are associated with the short-axis and long-axis tube orbits---however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.
 
 The strongest chaotic regions (black) appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}).
-Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map, especially where resonances overlap --- for example, near $(x,z) \approx(13,7)~{\rm kpc}$ and $(x,z) \approx(31,13)~{\rm kpc}$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}.
+Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map, especially where resonances overlap---for example, near $(x,z) \approx(13,7)~{\rm kpc}$ and $(x,z) \approx(31,13)~{\rm kpc}$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}.
 The details of the resonant structure is revealed in the frequency diffusion map from integrations of just 80 orbital periods.
@@
 
-The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant for a given orbit --- that is, these quantities are measures of how infinitesimal deviations will diverge on average from some parent orbit, or of how long it takes for the frequencies of a single orbit to change by some amount.
+The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant for a given orbit---that is, these quantities are measures of how infinitesimal deviations will diverge on average from some parent orbit, or of how long it takes for the frequencies of a single orbit to change by some amount.
 Tidal debris is disrupted from progenitor systems with finite spreads in orbital properties (e.g., energy).
@@
 We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
-For a given set of orbital initial conditions --- the `parent' orbit --- we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time.
+For a given set of orbital initial conditions---the `parent' orbit---we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time.
 The physical scale of the ensemble is set by the tidal radius in position and the velocity scale in velocity and are therefore set by the mass of the progenitor \citep[e.g.,][]{johnston98, apw14}.
@@
 		\end{tabular}
-	\caption{Orbits from the $xz$ grid with a range of chaotic timescales --- R is a regular orbit, W a weakly chaotic orbit, and S a strongly chaotic orbit.
+	\caption{Orbits from the $xz$ grid with a range of chaotic timescales---R is a regular orbit, W a weakly chaotic orbit, and S a strongly chaotic orbit.
 Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$, and times ($t_\lambda$, $t_\Omega$) in number of orbital periods.
@@
 The regular orbit distribution stays relatively compact at all times indicating that the density stays fairly uniformly high throughout the stream.
-For the weakly chaotic orbit, the distribution is somewhat wider with a larger tail to lower densities --- this is the debris that begins to fan out.
+For the weakly chaotic orbit, the distribution is somewhat wider with a larger tail to lower densities---this is the debris that begins to fan out.
 The strongly chaotic orbit evolves much faster to low density.
@@
 
-Motivated by the noticeable discrepancy between the regular and weakly chaotic orbit ensembles --- even though the chaotic timescale of the weakly chaotic ensemble parent orbit is thousands of orbital periods --- we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
+Motivated by the noticeable discrepancy between the regular and weakly chaotic orbit ensembles---even though the chaotic timescale of the weakly chaotic ensemble parent orbit is thousands of orbital periods---we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
 For all ensembles, we integrate the orbits for 16 (parent) orbital periods and store the initial and final values of the mean, configuration-space density.
@@
 As a demonstration of the small-scale frequency modulation, we consider again the three orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:three-orbits}).
-Figure~\ref{fig:fdiff-zooms} shows zoomed regions of the frequency diffusion time map of Figure~\ref{fig:freqdiff} centered on the three orbits --- the pixel positions are now shown relative to the three parent orbits.
+Figure~\ref{fig:fdiff-zooms} shows zoomed regions of the frequency diffusion time map of Figure~\ref{fig:freqdiff} centered on the three orbits---the pixel positions are now shown relative to the three parent orbits.
 The frequency diffusion times shown in these grids were computed with two consecutive windows of length 40 orbital periods.
@@
 To resolve the short-time behavior of the frequency diffusion (corresponding to motion across or around resonance layers) we compute the frequencies in a series of rolling windows along each orbit.
-We use a window with a width equal to 40 orbital periods and shift the window by an orbital period between each calculation of the most significant frequencies --- again, this is the smallest window size we could choose before \superfreq\ begins to occasionally fail to recover the frequencies for the strongly chaotic orbit.
-Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time --- plotted are the percent deviations of the frequency ratios from the initial value.
+We use a window with a width equal to 40 orbital periods and shift the window by an orbital period between each calculation of the most significant frequencies---again, this is the smallest window size we could choose before \superfreq\ begins to occasionally fail to recover the frequencies for the strongly chaotic orbit.
+Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time---plotted are the percent deviations of the frequency ratios from the initial value.
 The difference between the first window (green +) and the last window (red x) is 50 orbital periods.
@@
 Middle panels show the full evolution of each ensemble orbit over 16 orbital periods, and final panel shows the final per cent deviations for all ensemble orbits (after 16 orbital periods).
-Note that the axis scales are quite different between each figure --- the strongly chaotic orbit ensemble evolves quickly due to its proximity to many overlapping resonances.
+Note that the axis scales are quite different between each figure---the strongly chaotic orbit ensemble evolves quickly due to its proximity to many overlapping resonances.
 
@@
 \item \emph{Time dependence}: Galaxies are certainly not static systems.
-To first order, galaxies grow in mass --- for example, the spherically averaged mass profile of dark-matter haloes evolves fairly predictably in cosmological simulations \citep{wechsler02, buist14} after some initial period of stochastic mass growth.
+To first order, galaxies grow in mass---for example, the spherically averaged mass profile of dark-matter haloes evolves fairly predictably in cosmological simulations \citep{wechsler02, buist14} after some initial period of stochastic mass growth.
 The Milky Way has probably had a fairly calm accretion history over the last 6 Gyr and therefore the mass growth may be similar to that seen in simulations.
@@
 Instead, the collective effect of the subhaloes may instead act as a noise term in the Hamiltonian of any halo orbit.
-This subhalo-induced colored noise --- which depends on the mass spectrum and distribution of subhaloes --- may also act to simply increase the magnitude of chaos along orbits, and destabilize sufficiently non-resonant orbits \citep[see, e.g.,][]{kandrup00}.
+This subhalo-induced colored noise---which depends on the mass spectrum and distribution of subhaloes---may also act to simply increase the magnitude of chaos along orbits, and destabilize sufficiently non-resonant orbits \citep[see, e.g.,][]{kandrup00}.
 
@@
 From each of these indicators, we compute a timescale over which chaos is likely to be important and find that the majority of orbits have chaotic timescales greater than 100s of orbital periods, however with the frequency diffusion rate we are still able to resolve weak chaos.
-We then study the density evolution of small ensembles of orbits generated around each orbit in the grid used for the previous experiment and find that along some orbits classified as weakly chaotic --- with chaotic timescales of 100s of orbital periods --- the orbit ensembles display enhanced density evolution and reach a lower overall density faster than orbit ensembles around nearby regular orbits (which mix due to phase-mixing alone).
+We then study the density evolution of small ensembles of orbits generated around each orbit in the grid used for the previous experiment and find that along some orbits classified as weakly chaotic---with chaotic timescales of 100s of orbital periods---the orbit ensembles display enhanced density evolution and reach a lower overall density faster than orbit ensembles around nearby regular orbits (which mix due to phase-mixing alone).
 We explain this apparent discrepancy \todo{apw}{Hogg's comment is to clearly state the discrepancy (has to do with timescales)} by considering the nature of chaotic diffusion: the classical chaos indicators are most sensitive to the slow, Arnold diffusion process that can cause large changes to orbital properties, but small-scale frequency evolution occurs over much shorter times as chaotic orbits stochastically diffuse across the stochastic layers that surround many resonances.
@@
 The mere existence of thin tidal streams in the halo of the Milky Way either (1) provides useful information about the potential on these scales by, e.g., implying a large degree of regularity, or (2) indicates that the thin, long streams (e.g., GD-1) are on regular orbits.
-These are not mutually exclusive --- in fact, if the streams are on regular orbits, this would be a powerful way to check or rule out possible potential models by requiring that the progenitor orbits remain regular.
+These are not mutually exclusive---in fact, if the streams are on regular orbits, this would be a powerful way to check or rule out possible potential models by requiring that the progenitor orbits remain regular.
 
@@
 
-\todo{apw}{This section has not been edited carefully...}
-\todo{apw}{What do I mean by `more reliable' for $p=2$?}
+\renewcommand{\thefigure}{\thesection.\arabic{figure}}
 
-Figure~\ref{fig:logfreqs} shows a test of the implementation in which we reproduce the frequency map at a fixed energy of an axisymmetric, logarithmic potential \cite[][pg.
-260, Figure~3.45]{binneytremaine}.
-Plotted are the (Cartesian) frequency ratios recovered for a grid of iso-energy, box orbits integrated in the potential
+The NAFF algorithm operates on numerically integrated orbital time-series (i.e.
+positions and velocities at a set of equally spaced times).
+Complex combinations of the phase-space coordinates of the orbit are Fourier-transformed using a standard FFT and the resultant spectrum is then searched for the strongest frequency component; this serves as an initial guess for the frequency of a particular Fourier component.
+Solving for the frequency that maximizes the Fourier integral of the FFT with a particular window function allows for a more accurate determination of the frequency.
+It has been shown that this accuracy converges much faster than the typical $\inttime^{-1}$ expected for a standard FFT by using a window function of the form
 \begin{equation}
-	\Phi(x,y,z) = \frac{1}{2}\ln\left(x^2 + (y/0.9)^2 + (z/0.7)^2 + 0.1\right). \label{eq:logpotential}
+	W(\tau=t/\inttime) = \frac{2^p \, (p!)^2}{(2p)!} \left( 1 + \cos \pi \tau\right)^p
 \end{equation}
-In such a map, stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities.
-The regularity of the points in this map reflects the input grid of initial conditions.
-Points that appear to be erratically scattered are chaotic orbits where the frequencies are changing with time.
+where $\inttime$ is the integration time \citep{laskar99}.
+Once the strongest frequency component is found, it is subtracted from the spectrum and the process is repeated iteratively.
+This procedure generates a table of frequencies for each component of motion which must then be searched for the three fundamental frequencies.
+The original implementations of NAFF have used $p=1$ \citep[e.g.,][]{laskar93, valluri98}, but we have chosen to use $p=2$.
+With a higher order window function, the central peak is broadened, but the amplitudes of the side lobes decrease faster \citep[see][]{hunter02}; we have found that this additional damping of the side lobes allows for more reliable determination of frequencies from strongly chaotic orbits.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth]{figures/log-freqmap.png}
+\caption{ A reproduction of Figure~3.45 from \citep{binneytremaine} as a validation of our frequency analysis code: Frequency ratios for 100000 isoenergy orbits integrated in a triaxial logarithmic potential.
+Linear features are resonances---stable resonances appear as dark lines, unstable resonances appear as linear gaps.} \label{fig:logfreqs}
+\end{center}
+\end{figure}
 
@@
 
+Figure~\ref{fig:logfreqs} shows a validation of our frequency analysis code in which we reproduce the frequency map at a fixed energy of an axisymmetric, logarithmic potential \cite[][pg.
+260, Figure~3.45]{binneytremaine}.
+Plotted are the (Cartesian) frequency ratios recovered for a grid of iso-energy, box orbits integrated in the potential
+\begin{equation}
+	\Phi(x,y,z) = \frac{1}{2}\ln\left(x^2 + (y/0.9)^2 + (z/0.7)^2 + 0.1\right). \label{eq:logpotential}
+\end{equation}
+Following \cite{binneytremaine}, we generate a grid of orbits on the equipotential surface $\Phi(x,y,z) = 0.5$ (we use a grid with 100000 orbits compared to their 10000 orbits).
+Each orbit is integrated for $\approx$40 orbital periods.
+In this figure, stable resonances appear as linear over-densities and unstable resonances appear as linear under-densities or gaps.
+The regularity of the points in this map reflects the input grid of initial conditions.
+Points that appear to be erratically scattered are chaotic orbits where the frequencies change with time.
+
+% ----------------------------------------------------------------------------------
+
+\renewcommand{\thefigure}{\arabic{figure}}
+\setcounter{figure}{0}
+
 % Figure ??
-%\begin{sidewaysfigure}[!p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/cartoons.pdf}
-%\caption{} 
-%\label{fig:cartoons}
-%\end{center}
-%\end{sidewaysfigure}
-\begin{figure*}[!p]
+\begin{figure}[!p]
 \begin{center}
@@
 \end{center}
-\end{figure*}
+\end{figure}
 
 % Figure ??
-\begin{figure*}[!p]
+\begin{figure}[!p]
 \begin{center}
@@
 \end{center}
-\end{figure*}
+\end{figure}
 
@@
 %\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
-Each pixel in this image represents a single orbit, and the shading (white/black) indicates the behavior of the orbit over 50 orbital periods: if the orbit preserves a sense of rotation around some axis --- that is, the angular momentum around some axis does not change sign --- the orbit is considered a tube orbit (white), otherwise the orbit is considered a box or chaotic orbit (black).
+Each pixel in this image represents a single orbit, and the shading (white/black) indicates the behavior of the orbit over 50 orbital periods: if the orbit preserves a sense of rotation around some axis---that is, the angular momentum around some axis does not change sign---the orbit is considered a tube orbit (white), otherwise the orbit is considered a box or chaotic orbit (black).
 The regions hosting the major orbit types are indicated by text labels.
@@
 % Figure ??
-\begin{figure*}[p]
+\begin{figure}[p]
+\centering
+    \subfloat{
+      \includegraphics[width=0.45\textwidth]{figures/pericenter-map.png}
+    }
+    \subfloat[Next sub-figure\label{subfig-2:dummy}]{
+      \includegraphics[width=0.45\textwidth]{figures/apocenter-map.png}
+    }
+\caption{\todo{apw}{yo}}
+\label{fig:coucou}
+\end{figure}
+
+% Figure ??
+\begin{figure}[p]
 \begin{center}
@@
 Chaotic orbits with $t_\lambda \gtrsim 1000~\periods$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
-Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:lyapmap}
+Three orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:lyapmap}
 \end{center}
-\end{figure*}
+\end{figure}
 
 % Figure ??
-\begin{figure*}[p]
+\begin{figure}[p]
 \begin{center}
@@
 The frequencies are measured precisely so that small changes in the frequencies can be detected over just $\approx$10s of orbits.
-Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
+Three orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
 \end{center}
-\end{figure*}
+\end{figure}
 
 % Figure ??
-\begin{figure*}[p]
+\begin{figure}[p]
 \begin{center}
@@
 \end{center}
-\end{figure*}
+\end{figure}
 
 % Figure ??
-\begin{figure*}[p]
+\begin{figure}[p]
 \begin{center}
@@
 The mass of the progenitor is set to XX, and the length-scale of the Plummer sphere is set to 20~pc to get $\approx$10\% mass loss over the integration time.
-Panels show particle positions from the final snapshots of the simulations --- for visualization the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
+Panels show particle positions from the final snapshots of the simulations---for visualization the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
 These simulations confirm that the test-particle orbit ensembles do (qualitatively) capture the nature of the early-stripped tidal debris.
@@
 \clearpage
-\begin{figure*}[p]
+\begin{figure}[p]
 \begin{center}
@@
 We generate globular-cluster-scale ensembles of orbits around each `parent' orbit in the grid (Section~\ref{sec:results2}) and follow the density of the ensemble as a function of time using kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
-Three orbits are highlighted with square markers --- from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).}
+Three orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).}
 \label{fig:ensemblemap-meandensity}
 \end{center}
-\end{figure*}
+\end{figure}
 
@@
 \clearpage
-\begin{figure*}[p]
+\begin{figure}[p]
 \begin{center}
 \includegraphics[width=1.1\textwidth]{figures/fdiff_zooms.png}
-\caption{Zooms of the frequency diffusion time map of Figure~\ref{fig:freqdiff} around the three orbits highlighted in that figure and used in Figures~\ref{fig:nbodysims}--\ref{fig:ensemble-density} --- as in Figure~\ref{fig:freqdiff}, each pixel represents an orbit, and the greyscale of the pixel is set by the frequency diffusion time.
+\caption{Zooms of the frequency diffusion time map of Figure~\ref{fig:freqdiff} around the three orbits highlighted in that figure and used in Figures~\ref{fig:nbodysims}--\ref{fig:ensemble-density}---as in Figure~\ref{fig:freqdiff}, each pixel represents an orbit, and the greyscale of the pixel is set by the frequency diffusion time.
 Around each orbit, we initialize a new grid of $\approx$40,000 orbits to resolve the small-scale resonant structure around each orbit.
@@
 Fundamental frequencies are computed with a window width of $\approx$40 orbital periods, and the window is shifted by one orbital period between each computation (each grey point represents the fundamental frequency ratios computed in a single window).
-Plotted are the per cent deviations of the frequency ratios from the value in the initial window --- the initial value is shown as a green plus sign and the final value is shown as a red x.
+Plotted are the per cent deviations of the frequency ratios from the value in the initial window---the initial value is shown as a green plus sign and the final value is shown as a red x.
 For the regular orbit (R), the fractional variation is around $10^{-6}$ and thus all points overlap on this scale.
@@
 Frequencies are now explore a huge volume representing $>$100\% deviations from the initial value of the parent orbit.
-Frequencies evolve quickly around resonances --- especially the $\bs{n}=(2,-3,0)$, $\bs{n}=(1,-1,0)$, and $\bs{n}=(1,0,1)$ resonances --- which appear as linear features in these panels.}
+Frequencies evolve quickly around resonances---especially the $\bs{n}=(2,-3,0)$, $\bs{n}=(1,-1,0)$, and $\bs{n}=(1,0,1)$ resonances---which appear as linear features in these panels.}
 \label{fig:chaotic-ensemble-freq-evolution}
 \end{center}
-\end{figure*}
+\end{figure}
 
@@
 The frequency evolution is computed for the equivalent of 16 orbital periods.
-The weakly chaotic orbit has $A_\Omega \approx 0.2\%$ over just 16 orbital periods --- this explains the enhanced density evolution observed in the N-body simulations and ensemble integrations (Figures~\ref{fig:nbodysims}-\ref{fig:ensembles}).
+The weakly chaotic orbit has $A_\Omega \approx 0.2\%$ over just 16 orbital periods---this explains the enhanced density evolution observed in the N-body simulations and ensemble integrations (Figures~\ref{fig:nbodysims}-\ref{fig:ensembles}).
 The three orbits, R, W, S, are highlighted with square markers (S is top, W is middle, R is bottom).

### 2015-07-11 ms.tex

@@
 Thin tidal streams may preferentially form along tube orbits rather than box orbits because of the faster disruption and debris diffusion expected for stellar systems on radially plunging orbits.
+Figure~\ref{fig:apoper} shows the grid of initial conditions in the $xz$ plane---each pixel in this grid represents an orbit, and the pixels are colored either by the median pericentric distance (left) or median apocentric distance (right) after 16 orbital periods.
 
@@
     }
-\caption{\todo{apw}{yo}}
-\label{fig:coucou}
+\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
+Each pixel in these panels represents a single orbit, and the shading of each pixel corresponds to the median pericentric distance (left) or median apocentric distance (right) computed over 16 orbital periods.
+The central black band in the left panel and white band in the right panel are tube orbits with apocenter-to-pericenter ratios close to one.
+}
+\label{fig:apoper}
 \end{figure}
@@
 \includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/lyap_map.png}
-\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
-Each pixel in this image represents a single orbit, and each pixel is colored by the  logarithm of the Lyapunov time (in units of orbital periods).
+\caption{ The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the logarithm of the Lyapunov time (in units of orbital periods).
 Orbits are integrated for a total of 10000 orbital periods.
@@
 \includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/fdiff_map.png}
-\caption{The same grid of initial conditions as Figure~\ref{fig:lyapmap}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
+\caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
 The frequencies are computed in two consecutive windows, each of which has length equal to 40 orbital periods.
@@
 \caption{\todo{all}{Ignore this for now!
-Waiting for results from latest run on cluster...}The same grid of initial conditions as Figure~\ref{fig:lyapmap} and other previous figures, but now we consider the density evolution of small ensembles of orbits around each of the orbits in this grid.
+Waiting for results from latest run on cluster...}The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now we consider the density evolution of small ensembles of orbits around each of the orbits in this grid.
 We generate globular-cluster-scale ensembles of orbits around each `parent' orbit in the grid (Section~\ref{sec:results2}) and follow the density of the ensemble as a function of time using kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
@@
 \includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/freqvar-map.png}
-\caption{The same grid of initial conditions as in previous figures, but now each pixel is colored by the amplitude of short-time frequency variations, $A_\Omega$ (Equation~\ref{eq:frac-freq-amp}).
+\caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the amplitude of short-time frequency variations, $A_\Omega$ (Equation~\ref{eq:frac-freq-amp}).
 The amplitude is a fractional measurement, so that a value of $10^{-2}$ corresponds to a 1\% change in frequencies.

### 2015-07-11 ms.tex
update ensemble description

@@
 
-For each orbit, we also separately integrate for $\approx$80 orbital periods and use \superfreq\ to compute the fundamental frequencies for the two consecutive sections of 40 orbital periods.
+For each orbit, we also separately integrate for $\approx$128 orbital periods and use \superfreq\ to compute the fundamental frequencies for the two consecutive sections of 64 orbital periods.
 We have found that this is the shortest integration window for which we are able to successfully recover frequencies using \superfreq\ for $>$99.9\% of the orbits.
@@
 Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map, especially where resonances overlap---for example, near $(x,z) \approx(13,7)~{\rm kpc}$ and $(x,z) \approx(31,13)~{\rm kpc}$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}.
-The details of the resonant structure is revealed in the frequency diffusion map from integrations of just 80 orbital periods.
+The details of the resonant structure is revealed in the frequency diffusion map from integrations of just 128 orbital periods.
 
 While there is rich structure and a significant number of weakly chaotic orbits, the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods (thousands of Gigayears) and are thus not expected to be relevant for tidal stream evolution.
-In the next section, we analyze the density evolution of finite-volume ensembles of orbits around each orbit in the above grids in order to study the effect of weak chaos on tidal debris.
+In the next section, we analyze the density evolution of finite-volume ensembles of orbits around each orbit in the above grids in order to compare the effect of ordinary phase-mixing of tidal debris with potentially enhanced mixing due to chaos.
 We then compare the density evolution of the ensembles to the single-orbit chaos indicators computed in this section.
@@
 The physical scale of the ensemble is set by the tidal radius in position and the velocity scale in velocity and are therefore set by the mass of the progenitor \citep[e.g.,][]{johnston98, apw14}.
-If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
+If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $\delta x_{k,i}$ and $\delta v_{k,i}$ are the $k$th components ($k=1,2,3$) of the position and velocity deviation vectors for the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
 \begin{align}
-	\delta\bs{x}_i &\sim \mathcal{N}(0, r_{\rm tide})\\
-	\delta\bs{v}_i &\sim \mathcal{N}(0, \sigma_v)\\
-	r_{\rm tide} &= \|\bs{x}_0\| \left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} \\
-	\sigma_v &= \|\bs{v}_0\|\left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} 
+	\delta x_{k,i} &\sim \mathcal{N}(0, r_{\rm tide}/\sqrt{3})\\
+	\delta v_{k,i} &\sim \mathcal{N}(0, \sigma_v/\sqrt{3})\\
+	r_{\rm tide} &= \|\bs{x}_0\| \left(\frac{m}{M(<\|\bs{x}_0\|)}\right)^{1/3} \\
+	\sigma_v &= \|\bs{v}_0\|\left(\frac{m}{M(<\|\bs{x}_0\|)}\right)^{1/3} 
 \end{align}
 where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the `progenitor,' and $\|\cdot \|$ is the Euclidean norm.
+These scalings are chosen so that the ensembles all start with approximately the same initial spread in frequencies (which also scales with $(m/M)^{1/3}$).
 We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales.
 
+\todo{apw}{This needs to be updated...}
 We start by considering three particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular orbit (R), a weakly chaotic orbit (W), and a strongly chaotic orbit (S).

### 2015-07-16 ms.tex
Added some todos

@@
 The result from \citet{pearson15} suggests that even weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.
-Understanding why this occurs and developing a measure to quantify the importance of this enhanced density evolution is a promising new direction to be explored further below.
+Understanding why this occurs and developing a measure to quantify the importance of this enhanced density evolution is a promising new direction to be explored further.
 
@@
 \subsection{Mixing of orbit ensembles}\label{sec:chaotic-mixing}
+
+\todo{apw}{talk about streams, sanders/bovy stuff in actions / frequencies.
+streams are 1D stuctures.
+small deviations orthogonal to main elongation, e.g., from small-scale chaotic diffusion can matter}
 
@@
 
+\todo{apw}{rename 'orbit structure' or 'orbit properties' and also show apocenter/pericenter map}
 \subsection{Part I: Lyapunov and frequency diffusion times}\label{sec:results1}
@@
 For a given set of orbital initial conditions---the `parent' orbit---we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time.
+\todo{apw}{We want to isolate phase mixing, so we want the ensembles to have the same spreads in frequencies...no variance introduced in the velocities}
 The physical scale of the ensemble is set by the tidal radius in position and the velocity scale in velocity and are therefore set by the mass of the progenitor \citep[e.g.,][]{johnston98, apw14}.
@@
 
-\todo{apw}{This needs to be updated...}
-We start by considering three particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular orbit (R), a weakly chaotic orbit (W), and a strongly chaotic orbit (S).
-The orbits were chosen from a single column of the orbit grid based on their frequency diffusion time to sample a range of possible values, and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}.
+We start by considering three particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular (non-resonant) orbit (R), a weakly chaotic orbit (W), and a strongly chaotic orbit (S).
+The orbits were chosen to be close on the orbit grid so that their orbital properties (e.g., apocenter, pericenter) are similar, and the degree of chaos is taken from the \todo{apw}{nonresonant regular, weakly chaotic, strongly chaotic}
+based on their frequency diffusion time to sample a range of possible values, and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}.
 Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the three orbits described above and evolved for 16 orbital periods.
@@
 Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit ($t_\lambda \approx 55$, $t_\Omega \approx 8\times10^4$), it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream.
+\todo{apw}{numbers}
 
@@
 The density evolution can be complex, however with a single measure that characterizes the ensemble density evolution we can map the entire grid of orbits used in previous sections (e.g., Figure~\ref{fig:lyapmap}).
-The density distributions do not change shape dramatically, even for the strongly chaotic ensemble.
 Figure~\ref{fig:densities} shows the evolution of the mean density of the ensemble particles computed at 256 evenly-spaced intervals from the initial ensemble distribution to the distribution after 16 orbital periods.
 It is clear that the density of the regular orbit ensemble evolves slower than the weakly chaotic ensemble, and the strongly chaotic ensemble quickly becomes diffuse; \todo{apw}{WTF?} the mean density does not capture the full nature of the density evolution is apparently sensitive to the enhanced density evolution observed along the weakly chaotic orbit seen in Figure~\ref{fig:ensemble-density}.
-
-Motivated by the noticeable discrepancy between the regular and weakly chaotic orbit ensembles---even though the chaotic timescale of the weakly chaotic ensemble parent orbit is thousands of orbital periods---we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
-For all ensembles, we integrate the orbits for 16 (parent) orbital periods and store the initial and final values of the mean, configuration-space density.
+\todo{apw}{check above paragraph, add comments about lines that show power law behavior}
+
+Motivated by the noticeable discrepancy between the regular and weakly chaotic orbit ensembles---even though the chaotic timescale of the weakly chaotic ensemble parent orbit is several times the integration period used above (32 orbital periods)---we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
+For all ensembles, we integrate the orbits for 32 (parent) orbital periods and store the initial and final values of the mean, configuration-space density.
 Figure~\ref{fig:ensemblemap-meandensity} again shows the grid of initial conditions (Section~\ref{sec:results1}, but now the greyscale indicates the ratio of the final mean density to the initial density.
 Much of the structure that is visible in the frequency diffusion time map (Figure~\ref{fig:freqdiff}) is again visible in this map of the density evolution of orbit ensembles.
-However, it is surprising that these features are present: the chaotic timescales predicted from both the Lyapunov and frequency diffusion times were typically 100s to 1000s of orbital periods for many of the unstable resonance features (curved lines).
-\todo{apw}{Why don't these features show up?}
+However, it is surprising that these features are present: the chaotic timescales predicted from both the Lyapunov and frequency diffusion times were typically 100 to 1000s of orbital periods for many of the unstable resonance features (curved lines).
+\todo{apw}{Why don't these features show up?!?}
 
@@
 
+\todo{apw}{rename frequency-space structure or something, show the frequency plots for the three ensembles}
 \subsection{Part III: Short-time frequency evolution}\label{sec:results3}
@@
 This procedure generates a table of frequencies for each component of motion which must then be searched for the three fundamental frequencies.
-The original implementations of NAFF have used $p=1$ \citep[e.g.,][]{laskar93, valluri98}, but we have chosen to use $p=2$.
+The original implementations of NAFF have used $p=1$ \citep[e.g.,][]{laskar93, valluri98}, but we have chosen to use $p=4$.
 With a higher order window function, the central peak is broadened, but the amplitudes of the side lobes decrease faster \citep[see][]{hunter02}; we have found that this additional damping of the side lobes allows for more reliable determination of frequencies from strongly chaotic orbits.

### 2015-07-17 ms.tex
remove figure add figure

@@
 
-\todo{apw}{talk about streams, sanders/bovy stuff in actions / frequencies.
-streams are 1D stuctures.
-small deviations orthogonal to main elongation, e.g., from small-scale chaotic diffusion can matter}
-
 An ensemble of regular orbits (e.g., tidal debris) will phase-mix due to (small) differences in the fundamental frequencies of the orbits.
-Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits \cite[see, e.g.,][]{kandrup94, merritt96, kandrup03}, however this depends on the details of the resonant structure around the ensemble and are thus difficult to predict.
+The frequency distributions of thin streams are generally close to one-dimensional---that is, one eigenvalue of the distribution of frequencies for an orbit ensemble will be much larger than the others because of the local shape of the Hessian of the potential \citep{sanders13, bovy14}.
+Phase-mixing generically leads to power-law decay of the mean density of the ensembles: initially, the density decreases linearly in time because of the large, nearly one-dimensional spread in frequencies, then may proceed as $t^{-2}$ to $t^{-3}$ depending on the other eigenvalues of the frequency-space distribution.
+
+Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits \cite[see, e.g.,][]{kandrup94, merritt96, kandrup03}, however this depends on the details of the resonant structure around the ensemble and the chaotic evolution of the orbit and is thus difficult to predict.
 For example, ensembles of orbits `stuck' between resonances may quickly spread to fill the allowed volume via the short-time chaotic diffusion processes, but then the orbits must escape this confinement and diffuse through the Arnold web to reach a fully mixed state \citep{merritt96}.
-
 In this work, we investigate the consequences of this short-time but small-scale frequency evolution and hypothesize that this may explain the enhanced density evolution of tidal debris around weakly chaotic regions where chaotic timescales are predicted to be long \citep[e.g.,][]{pearson15}.
@@
 The MLE, $\lambda_{\rm max}$, is interpreted as a rate that quantifies the exponential divergence of infinitesimally close chaotic orbits.
-It is therefore useful to consider the corresponding $e$-folding time by inverting the rate.
-We normalize the Lyapunov time by the maximum orbital period\footnote{The maximum of the three orbital periods corresponding to the three fundamental frequencies (see Section~\ref{sec:naff}).} of a given orbit, $T_{\rm max}$, so that the time is in units of number of orbital periods
-\begin{equation}
-	t_{\rm \lambda} = \frac{1}{\lambda_{\rm max} \, T_{\rm max}}
+It is therefore useful to consider the corresponding $e$-folding time by inverting the rate,
+\begin{equation}
+	t_{\rm \lambda} = \frac{1}{\lambda_{\rm max}}.
 \end{equation}
@@
 We have implemented and tested a version of this procedure in the \project{Python} programming language.
-Our implementation differs slightly from the original definition and from that used in \cite{valluri98}; we refer to this slight modification of the algorithm as \superfreq\ \citep{superfreq} and the code is open-source and publicly available on \project{GitHub} at \url{https://github.com/adrn/SuperFreq}.
-For more details about the algorithm and differences with previous work, see \cite{laskar88, laskar93, papaphilippou96} and Appendix~\ref{sec:naffapdx}.
+Our implementation differs slightly from the original definition and from that used in \cite{valluri98}; we refer to this slight modification of the algorithm as \superfreq\ \citep{superfreq} and the code is open-source and publicly available on \project{GitHub}.\footnote{\url{https://github.com/adrn/SuperFreq}} For more details about the algorithm and differences with previous work, see \cite{laskar88, laskar93, papaphilippou96} and Appendix~\ref{sec:naffapdx}.
 
@@
 The frequencies picked out by \superfreq\ for such orbits will therefore represent the average periodic nature of the orbit over a given integration window.
-We define the fractional frequency diffusion rate per orbital period, $\mathcal{R}$, in the $k$th fundamental frequency as
-\begin{equation}
-	\mathcal{R}_k = \frac{\Omega_{k}^{(2)} - \Omega_{k}^{(1)}}{\Omega_{k}^{(1)} \, N} \label{eq:fdrate}
-\end{equation}
-where the upper index refers to the two consecutive sections of orbit and $N$ is the number of orbital periods in a single section of orbit.
+Following previous work, we define the fractional frequency diffusion rate, $\mathcal{R}$, in the $k$th fundamental frequency as
+\begin{equation}
+	\mathcal{R}_k = \frac{\Omega_{k}^{(2)} - \Omega_{k}^{(1)}}{\Omega_{k}^{(1)}} \label{eq:fdrate}
+\end{equation}
+where the upper index refers to the two consecutive sections of orbit \citep{laskar93, valluri98, valluri12}.
 By inverting this rate, we can compute the timescale over which we expect order-unity changes to the fundamental frequencies: the \emph{frequency diffusion time} is defined as
@@
 \end{equation}
-where the maximum is taken with respect to the corresponding amplitudes, $a_k$, of the fundamental frequency components \citep[see][]{valluri12}, and the time is in units of number of orbital periods, ${\rm T}$.
+where the maximum is taken with respect to the corresponding amplitudes, $a_k$, of the fundamental frequency components \citep[see][]{valluri12}.
 
@@
 
-\todo{apw}{rename 'orbit structure' or 'orbit properties' and also show apocenter/pericenter map}
 \subsection{Part I: Lyapunov and frequency diffusion times}\label{sec:results1}
@@
 Thin tidal streams may preferentially form along tube orbits rather than box orbits because of the faster disruption and debris diffusion expected for stellar systems on radially plunging orbits.
-Figure~\ref{fig:apoper} shows the grid of initial conditions in the $xz$ plane---each pixel in this grid represents an orbit, and the pixels are colored either by the median pericentric distance (left) or median apocentric distance (right) after 16 orbital periods.
-
-%Figure~\ref{fig:icmap} shows the grid of initial conditions in the $xz$ plane---each pixel in this grid represents an orbit, and the major orbit families are labeled.
-Points are colored (black / white) depending on whether the orbit acts like a tube orbit or a box orbit over $\approx$50 orbital periods.
-Figure~\ref{fig:icorbits} show sample orbits from each of the main, labeled orbit families of the initial condition grid.
+Figure~\ref{fig:apoper} shows the grid of initial conditions in the $xz$ plane---each pixel in this grid represents an orbit, and the pixels are colored by the median pericentric distance (left) and median apocentric distance (right) over an integration time of 64 orbital periods.
 
 We integrate all orbits in the grid for 10000 orbital periods and use the method described in Section~\ref{sec:lyap} to compute the Lyapunov exponents.
-Figure~\ref{fig:lyapmap} again shows the grid of initial conditions, but now the logarithm of the Lyapunov time (in units of number of orbital periods) is mapped to greyscale intensity.
+Figure~\ref{fig:lyapmap} again shows the grid of initial conditions, but now the color corresponds to the logarithm of the Lyapunov time.
 The darker pixels have shorter Lyapunov times and are more chaotic.
-By fixing the integration time (in units of orbital periods, $\periods$), the MLE cannot detect weak chaos and the majority of orbits appear to be regular because they have exceedingly long Lyapunov times (all white points have $t_\lambda \gtrsim 1000~\periods$).
-
-For each orbit, we also separately integrate for $\approx$128 orbital periods and use \superfreq\ to compute the fundamental frequencies for the two consecutive sections of 64 orbital periods.
-We have found that this is the shortest integration window for which we are able to successfully recover frequencies using \superfreq\ for $>$99.9\% of the orbits.
-Figure~\ref{fig:freqdiff} shows the same grid of initial conditions as in Figure~\ref{fig:lyapmap}, but now the greyscale intensity is set by the logarithm of the (order-unity) frequency diffusion time.
+Because of the fixed the integration time, the MLE cannot detect weak chaos and the majority of orbits appear to be regular because they have exceedingly long Lyapunov times (all white points have $(t_\lambda/\periods) \gtrsim 1000$).
+
+\todo{apw}{remove kpc units?}
+For each orbit, we also separately integrate for $\approx$256 orbital periods and use \superfreq\ to compute the fundamental frequencies for the two consecutive sections of 128 orbital periods.
+We have chosen this window size so that we recover the frequencies for regular orbits with fractional error $\approx10^{-8}$.
+With this integration window we are able to successfully recover frequencies using for $>$99.9\% of the orbits \superfreq.
+Figure~\ref{fig:freqdiff} shows the same grid of initial conditions as in Figure~\ref{fig:lyapmap}, but now the greyscale intensity is set by the logarithm of the frequency diffusion time.
 The darker pixels have shorter frequency diffusion times and are more chaotic.
-This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential and highlights the accuracy of \superfreq\---weak chaos (grey) is detectable over just 10's of orbital periods, compared to the many thousands of orbits it would take to detect such features with the maximum Lyapunov exponent.
+This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential and highlights the accuracy of \superfreq\---weak chaos (grey) is detectable over just $\sim$100 orbital periods, compared to the many tens of thousands of orbits it would take to detect such features with the maximum Lyapunov exponent.
 The tube orbits in this potential are mostly regular or only mildly chaotic---the largest regular regions are associated with the short-axis and long-axis tube orbits---however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.
@@
 The strongest chaotic regions (black) appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}).
-Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map, especially where resonances overlap---for example, near $(x,z) \approx(13,7)~{\rm kpc}$ and $(x,z) \approx(31,13)~{\rm kpc}$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}.
-The details of the resonant structure is revealed in the frequency diffusion map from integrations of just 128 orbital periods.
-
-While there is rich structure and a significant number of weakly chaotic orbits, the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods (thousands of Gigayears) and are thus not expected to be relevant for tidal stream evolution.
+Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map, especially where resonances overlap---for example, near $(x_0,z_0) \approx(13,7)~{\rm kpc}$ and $(x_0,z_0) \approx(31,13)~{\rm kpc}$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}.
+The details of the resonant structure is revealed in the frequency diffusion map from integrations of just 256 orbital periods.
+
+While there is rich structure and a significant number of weakly chaotic orbits, the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods and are thus not expected to be relevant for tidal stream evolution.
 In the next section, we analyze the density evolution of finite-volume ensembles of orbits around each orbit in the above grids in order to compare the effect of ordinary phase-mixing of tidal debris with potentially enhanced mixing due to chaos.
@@
 We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
-For a given set of orbital initial conditions---the `parent' orbit---we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time.
-\todo{apw}{We want to isolate phase mixing, so we want the ensembles to have the same spreads in frequencies...no variance introduced in the velocities}
-The physical scale of the ensemble is set by the tidal radius in position and the velocity scale in velocity and are therefore set by the mass of the progenitor \citep[e.g.,][]{johnston98, apw14}.
-If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $\delta x_{k,i}$ and $\delta v_{k,i}$ are the $k$th components ($k=1,2,3$) of the position and velocity deviation vectors for the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
+For a given set of orbital initial conditions---the `parent' orbit---we integrate for 128 orbital periods, find the phase-space position of the minimum pericenter over this times, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time.
+We are interested in the degree to which chaos enhances the mixing rate of orbit ensembles and we therefore want to isolate out the effect of ordinary phase-mixing along regular orbits; all ensembles should start with the same initial spread in fundamental frequencies.
+We set the physical scale of the ensemble by the tidal radius in position and the velocity scale in velocity; the initial spread in fundamental frequencies will scale as $(m/M)^{1/3}$ like the tidal radius and velocity scale \citep[e.g.,][]{johnston98, apw14}.
+If $(\bs{x},\bs{v})$ are a set of initial conditions at pericenter for a parent orbit, then $\delta x_{k,i}$ and $\delta v_{k,i}$ are the $k$th components ($k=1,2,3$) of the position and velocity deviation vectors for the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
 \begin{align}
@@
 where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the `progenitor,' and $\|\cdot \|$ is the Euclidean norm.
-These scalings are chosen so that the ensembles all start with approximately the same initial spread in frequencies (which also scales with $(m/M)^{1/3}$).
 We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales.
-
-We start by considering three particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular (non-resonant) orbit (R), a weakly chaotic orbit (W), and a strongly chaotic orbit (S).
-The orbits were chosen to be close on the orbit grid so that their orbital properties (e.g., apocenter, pericenter) are similar, and the degree of chaos is taken from the \todo{apw}{nonresonant regular, weakly chaotic, strongly chaotic}
-based on their frequency diffusion time to sample a range of possible values, and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}.
-Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the three orbits described above and evolved for 16 orbital periods.
-A thin `stream' forms on the regular orbit (left column), a more diffuse `stream' on the mildly chaotic orbit (middle column), and a `fanned' stream on the strongly chaotic orbit (right column).
-Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit ($t_\lambda \approx 55$, $t_\Omega \approx 8\times10^4$), it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream.
-\todo{apw}{numbers}
+% we have verified this using regular orbits in a logarithmic potential...
+
+We start by considering four particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a strongly chaotic orbit (D).
+The orbits were chosen to be close on the orbit grid so that their orbital properties (e.g., apocenter, pericenter) are similar, but have different frequency diffusion times; all four orbits are long-axis tube orbits.
+Figure~\ref{fig:orbits} shows the orbits in projection over a long integration period (1024 orbital periods).
+The initial conditions and chaos diagnostics for each orbit are listed in Table~\ref{tbl:orbit-info}.
+Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the four orbits described above and evolved for 64 orbital periods.
+A thin stream forms on the near-resonant orbit (left column), a thin---but more two-dimensional--- stream forms on the non-resonant orbit (middle-left), a more diffuse stream forms on the weakly chaotic orbit with an interesting morphology (middle-right), and a `fanned' stream on the strongly chaotic orbit (right).
+Given the long Lyapunov and frequency diffusion times of the weakly chaotic orbit ($t_\lambda/\periods > 900$, $t_\Omega/\periods \approx 2\times10^5$), it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream.
 
@@
 	\begin{tabular}{c | c c c c c c | c c }
-		{\bf ID} & $\bs{x}$ & $\bs{y}$ & $\bs{z}$ & $\bs{v_x}$ & $\bs{v_y}$ & $\bs{v_z}$ & $\bs{t_\lambda}$ & $\bs{t_\Omega}$ \\\toprule
-		R & 29.10 & 0 & 5.10 & 0 & 168.07 & 0 & $>10^3$ & $>10^8$\\
+		{\bf ID} & $\bs{x}$ & $\bs{y}$ & $\bs{z}$ & $\bs{v_x}$ & $\bs{v_y}$ & $\bs{v_z}$ & $\bs{t_\lambda/\periods}$ & $\bs{t_\Omega/\periods}$ \\\toprule
+A & 17 & 0 & 26.06 & 0 & 126.25 & 0 & $>$10^8 & XX\\
 \midrule
-		W & 29.10 & 0 & 17.70 & 0 & 125.49 & 0 & $\approx55$ & $\approx8 \times 10^4$\\
+B & 17 & 0 & 23.03 & 0 & 148.61 & 0 & XX & XX\\
 \midrule
-		S & 29.10 & 0 & 22.10 & 0 & 99.45 & 0 & $\approx3.5$ & $\approx6 \times 10^3$\\
+C & 17 & 0 & 25.35 & 0 & 131.68 & 0 & XX & XX\\
+\midrule
+D & 17 & 0 & 28.69 & 0 & 104.54 & 0 & XX & XX\\
 		\bottomrule
 		\end{tabular}
-	\caption{Orbits from the $xz$ grid with a range of chaotic timescales---R is a regular orbit, W a weakly chaotic orbit, and S a strongly chaotic orbit.
-Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$, and times ($t_\lambda$, $t_\Omega$) in number of orbital periods.
+	\caption{Orbits from the $xz$ grid with a range of chaotic timescales---A is a regular, near-resonant orbit, B a regular, non-resonant orbit, C a weakly chaotic orbit, and S a strongly chaotic orbit.
+Initial conditions are given in physical units: Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$.
+Chaotic timescales are expressed ($t_\lambda$, $t_\Omega$) in number of orbital periods.
 Recall that the frequency diffusion time, $t_\Omega$, is the time over which we expect order-unity changes in the fundamental frequencies, hence why the timescales appear quite long.
-\label{tbl:three-orbits}}
+\label{tbl:orbit-info}}
 \end{center}
@@
 \end{enumerate}
+\todo{apw}{make it clear: this is not about how much chaos, just timescale discrepancy}
 
@@
 \begin{center}
+\includegraphics[width=\textwidth]{figures/four-orbits.png}
+\caption{Four representative orbits chosen from the orbit grid (e.g., Figure~\ref{fig:apoper}): a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a strongly chaotic orbit (D).
+All orbits are long-axis tubes with similar pericenters and apocenters.
+Orbits in this Figure were integrated for 1024 orbital periods and are shown in projection.}
+\label{fig:orbits}
+\end{center}
+\end{figure}
+
+% Figure ??
+\begin{figure}[p]
+\begin{center}
 \includegraphics[width=\textwidth]{figures/ensembles.png}
 \caption{Final particle positions after integrating unbound, globular-cluster-sized ensembles of orbits generated around each of the three N-body orbits (Figure~\ref{fig:nbodysims}).
-The ensembles each contain 1000 particles and begin at pericenter rather than apocenter because they are meant to represent debris disrupted during a single (the first) pericentric passage.
-The particle positions qualitatively match the morphology of the `oldest' debris from the N-body simulations, though there are many more particles in these ensembles than are disrupted during the first pericentric passage of the simulations.
+The ensembles each contain 1024 particles and are initialized at pericenter.
+The particle positions qualitatively match the morphology of the `oldest' debris from the corresponding N-body simulations (Figure~\ref{fig:nbodysims}).
 }
@@
 The progenitor in each simulation is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
-We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter.
-The mass of the progenitor is set to XX, and the length-scale of the Plummer sphere is set to 20~pc to get $\approx$10\% mass loss over the integration time.
+We start the progenitor systems at apocenter and evolve for $\approx$64 orbital periods so that each simulation finishes again at subsequent apocenter.
+The mass of the progenitor is set to $10^4~\msun$, and the length-scale of the Plummer sphere is set to 5~pc to get $\approx$50--75\% mass loss over the integration time.
 Panels show particle positions from the final snapshots of the simulations---for visualization the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
-These simulations confirm that the test-particle orbit ensembles do (qualitatively) capture the nature of the early-stripped tidal debris.
+These simulations confirm that the test-particle orbit ensembles (Figure~\ref{fig:ensembles}) do (qualitatively) capture the nature of the early-stripped tidal debris.
 }
@@
 % Figure ??
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-density.png}
-\caption{ Distributions of the values of the estimated configuration-space density field evaluated at the positions of the ensemble particles after 0, 4, and 16 orbital periods for each of the three orbits of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims}.
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/ensemble-density.png}
+%\caption{ Distributions of the values of the estimated configuration-space density field evaluated at the positions of the ensemble particles after 0, 4, and 16 orbital periods for each of the three orbits of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims}.
 The density distribution of the regular orbit ensemble stays compact and decreases slowly.
@@
 The strongly chaotic ensemble evolves fast to a state of low density.} \label{fig:ensemble-density}
-\end{center}
-\end{figure}
+%\end{center}
+%\end{figure}
 
@@
 \includegraphics[width=\textwidth]{figures/ensemble-densities.png}
-\caption{This is an alternate visualization of the density evolution of orbit ensembles around the three orbits of Figure~\ref{fig:ensemble-density}, showing only the evolution of the mean density for the orbits (normalized by the initial density of the ensemble).
+\caption{Evolution of the mean density (normalized by the initial density) of orbit ensembles around the four orbits (Figure~\ref{fig:orbits}) over 64 orbital periods.
+Over-plotted are various power-laws that qualitatively show the decay of the mean density for the ensembles: the mean density of the near-resonant orbit (A) follows nearly $t^{-2}$; the non-resonant orbit (B) is slowly transitioning from $t^{-2}$ to $t^{-3}$; the weakly chaotic orbit (C) initially follows $t^{-2}$ but then exponential decay takes over; the strongly chaotic orbit (D) is roughly exponential with fairly erratic density evolution.
 }

### 2015-07-17 ms.tex
removing physical units.. WIP

@@
 We find that simulated streams (simulated using test particle ensembles of nearby orbits) are sensitive to chaos on a much shorter time-scale than any standard prediction (from the Lyapunov or frequency-diffusion times).
-For example, on a mildly chaotic orbit with a chaotic timescale predicted to be $>$1000 orbital periods ($>$1000 Gyr), the resulting stellar stream is, after just 16 orbits, substantially more diffuse than any formed on a nearby but regular orbit.
-We find that the enhanced diffusion of the stream stars can be estimated by looking at the variance in orbital frequencies along the parent (progenitor) orbit.
+For example, on a mildly chaotic orbit with a chaotic timescale predicted to be $>$1000 orbital periods ($>$1000 Gyr), the resulting stellar stream is, after just 64 orbits, substantially more diffuse than any formed on a nearby but regular orbit.
+We find that the enhanced diffusion of the stream stars can be understood by looking at the variance in orbital frequencies of orbit ensembles centered around the parent (progenitor) orbit.
 Our results suggest that long, cold streams around our Galaxy must exist only on regular (or very nearly regular) orbits; they potentially provide a map of the regular regions of the Milky Way potential.
@@
 An ensemble of regular orbits (e.g., tidal debris) will phase-mix due to (small) differences in the fundamental frequencies of the orbits.
-The frequency distributions of thin streams are generally close to one-dimensional---that is, one eigenvalue of the distribution of frequencies for an orbit ensemble will be much larger than the others because of the local shape of the Hessian of the potential \citep{sanders13, bovy14}.
-Phase-mixing generically leads to power-law decay of the mean density of the ensembles: initially, the density decreases linearly in time because of the large, nearly one-dimensional spread in frequencies, then may proceed as $t^{-2}$ to $t^{-3}$ depending on the other eigenvalues of the frequency-space distribution.
-
-Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits \cite[see, e.g.,][]{kandrup94, merritt96, kandrup03}, however this depends on the details of the resonant structure around the ensemble and the chaotic evolution of the orbit and is thus difficult to predict.
-For example, ensembles of orbits `stuck' between resonances may quickly spread to fill the allowed volume via the short-time chaotic diffusion processes, but then the orbits must escape this confinement and diffuse through the Arnold web to reach a fully mixed state \citep{merritt96}.
-In this work, we investigate the consequences of this short-time but small-scale frequency evolution and hypothesize that this may explain the enhanced density evolution of tidal debris around weakly chaotic regions where chaotic timescales are predicted to be long \citep[e.g.,][]{pearson15}.
+The frequency distributions of thin streams are generally close to one-dimensional---that is, one eigenvalue of the distribution of frequencies for an orbit ensemble will be much larger than the others because of the local shape of the Hessian of the potential \citep{helmi99, sanders13a, bovy14}.
+Phase-mixing generically leads to power-law decay of the mean density of the ensembles: initially, the density decreases linearly in time because of the large, nearly one-dimensional spread in frequencies, then may proceed as $t^{-2}$ to $t^{-3}$ depending on the other eigenvalues of the frequency-space distribution \citep[e.g.,][]{helmi99}.
+
+Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits \cite[see, e.g.,][]{kandrup94, merritt96, kandrup03}, however this depends on the details of the resonant structure around the ensemble and the chaotic evolution of the individual orbits and is thus difficult to predict.
+For example, ensembles of orbits `stuck' between resonances may quickly spread to fill the allowed volume, but then the orbits must escape this confinement and diffuse through the Arnold web to reach a fully mixed state \citep{merritt96}.
+In this work, we investigate the consequences of short-time but small-scale frequency evolution and hypothesize that this may explain the enhanced density evolution of tidal debris around weakly chaotic regions where chaotic timescales are predicted to be long \citep[e.g.,][]{pearson15}.
 We then discuss how this would affect our understanding of the coherence and density evolution of tidal streams.
@@
 \begin{equation}
-	\mathcal{R}_k = \frac{\Omega_{k}^{(2)} - \Omega_{k}^{(1)}}{\Omega_{k}^{(1)}} \label{eq:fdrate}
-\end{equation}
-where the upper index refers to the two consecutive sections of orbit \citep{laskar93, valluri98, valluri12}.
+	\mathcal{R}_k = \frac{\Omega_{k}^{(2)} - \Omega_{k}^{(1)}}{\Omega_{k}^{(1)} \, \Delta t} \label{eq:fdrate}
+\end{equation}
+where the upper index refers to the two consecutive sections of orbit and $\Delta t$ is the length of each integration window \citep{laskar93, valluri98, valluri12}.
 By inverting this rate, we can compute the timescale over which we expect order-unity changes to the fundamental frequencies: the \emph{frequency diffusion time} is defined as
@@
 We can estimate this timescale by multiplying the frequency diffusion time by a factor equal to the fractional spread in frequencies of the debris.
-For example, a globular cluster typically has $\approx$1\% spreads in fundamental frequencies, so by multiplying the frequency diffusion time by $f = 0.01$, we can estimate the time (in number of orbital periods) over which we expect the frequencies to evolve by this amount.
+For example, a globular cluster typically has $\approx$0.1\% spreads in fundamental frequencies, so by multiplying the frequency diffusion time by $f = 10^{-3}$, we can estimate the time (in number of orbital periods) over which we expect the frequencies to evolve by this amount.
 
@@
 We fix $v_x = v_z = 0$, and compute $v_y$ from the energy.
+From here onwards, all lengths are given in units of scale radii, $r_s$, velocities in units of circular velocity, $v_c$, (see Table~\ref{tbl:potential}) and times in units of orbital periods, $\periods$.
 The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic.
@@
 
-\todo{apw}{remove kpc units?}
+\todo{apw}{numbers here on (units)}
 For each orbit, we also separately integrate for $\approx$256 orbital periods and use \superfreq\ to compute the fundamental frequencies for the two consecutive sections of 128 orbital periods.
@@
 
+\todo{apw}{All lengths / distances in units of $r_s$, all velocities in units of $v_c$?}
+
 \begin{table*}[ht]
@@
 	\begin{tabular}{c | c c c c c c | c c }
-		{\bf ID} & $\bs{x}$ & $\bs{y}$ & $\bs{z}$ & $\bs{v_x}$ & $\bs{v_y}$ & $\bs{v_z}$ & $\bs{t_\lambda/\periods}$ & $\bs{t_\Omega/\periods}$ \\\toprule
-A & 17 & 0 & 26.06 & 0 & 126.25 & 0 & $>$10^8 & XX\\
+		{\bf ID} & $\bs{x/r_s}$ & $\bs{y/r_s}$ & $\bs{z/r_s}$ & $\bs{v_x/v_c}$ & $\bs{v_y/v_c}$ & $\bs{v_z/v_c}$ & $\bs{t_\lambda/\periods}$ & $\bs{t_\Omega/\periods}$ \\\toprule
+A & 0.85 & 0 & 1.303 & 0 & 0.721 & 0 & $>700$ & $>10^7$\\
 \midrule
-B & 17 & 0 & 23.03 & 0 & 148.61 & 0 & XX & XX\\
+B & 0.85 & 0 & 1.152 & 0 & 0.849 & 0 & $>700$ & $>10^7$\\
 \midrule
-C & 17 & 0 & 25.35 & 0 & 131.68 & 0 & XX & XX\\
+C & 0.85 & 0 & 1.27 & 0 & 0.752 & 0 & $>700$ & $\approx 3\times10^5$\\
 \midrule
-D & 17 & 0 & 28.69 & 0 & 104.54 & 0 & XX & XX\\
+D & 0.85 & 0 & 1.434 & 0 & 0.597 & 0 & 8.14 & $\approx 2.5\times10^4$\\
 		\bottomrule
 		\end{tabular}
-	\caption{Orbits from the $xz$ grid with a range of chaotic timescales---A is a regular, near-resonant orbit, B a regular, non-resonant orbit, C a weakly chaotic orbit, and S a strongly chaotic orbit.
-Initial conditions are given in physical units: Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$.
+	\caption{Orbital initial conditions from the $xz$ grid for orbits with a range of chaotic timescales---A is a regular, near-resonant orbit, B a regular, non-resonant orbit, C a weakly chaotic orbit, and D a strongly chaotic orbit.
+Positions ($x$, $y$, $z$) are given in units of scale radii, $r_s$, and velocities ($v_x$, $v_y$, $v_z$) in units of scale velocity, $v_c$.
 Chaotic timescales are expressed ($t_\lambda$, $t_\Omega$) in number of orbital periods.
@@
     }
-    \subfloat[Next sub-figure\label{subfig-2:dummy}]{
+    \subfloat{
       \includegraphics[width=0.45\textwidth]{figures/apocenter-map.png}
@@
 \caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
-Each pixel in these panels represents a single orbit, and the shading of each pixel corresponds to the median pericentric distance (left) or median apocentric distance (right) computed over 16 orbital periods.
+All distances are normalized by the potential scale radius.
+Each pixel in these panels represents a single orbit, and the shading of each pixel corresponds to the median pericentric distance (left) or median apocentric distance (right) computed over 64 orbital periods.
 The central black band in the left panel and white band in the right panel are tube orbits with apocenter-to-pericenter ratios close to one.
@@
 Orbits are integrated for a total of 10000 orbital periods.
-Chaotic orbits with $t_\lambda \gtrsim 1000~\periods$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
-Three orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:lyapmap}
+Chaotic orbits with $t_\lambda/\periods \gtrsim 700$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
+Four orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (D; white), near-resonant (A; black), weakly chaotic (C; black), and non-resonant (B; black) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:lyapmap}
 \end{center}
@@
 The frequencies are measured precisely so that small changes in the frequencies can be detected over just $\approx$10s of orbits.
-Three orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
+Three orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
 \end{center}

### 2015-07-18 ms.tex
update captions

@@
 The frequencies are measured precisely so that small changes in the frequencies can be detected over just $\approx$10s of orbits.
-Three orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
+Four orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (D; white), near-resonant (A; black), weakly chaotic (C; black), and non-resonant (B; black) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/ensemble-map.png}
-\caption{\todo{all}{Ignore this for now!
-Waiting for results from latest run on cluster...}The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now we consider the density evolution of small ensembles of orbits around each of the orbits in this grid.
-We generate globular-cluster-scale ensembles of orbits around each `parent' orbit in the grid (Section~\ref{sec:results2}) and follow the density of the ensemble as a function of time using kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
-Three orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (S; white), weakly chaotic (W; black), and regular (R; black) orbits of Table~\ref{tbl:three-orbits} (see Section~\ref{sec:results2}).}
+\includegraphics[width=0.9\textwidth, trim={0 0.2cm 0 0}]{figures/ensemble-map.png}
+\caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now the greyscale intensity is set by the mean density of a of small ensembles of orbits after integration for 64 orbital periods.
+The orbit ensembles are generated around each parent orbit in the grid (Section~\ref{sec:results2}) and the mean density is computed kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
+Given the extremely long chaotic timescales of the weakly chaotic orbits in the upper half of the grid, it is surprising that any of the high-order resonant structure is visible in this map.
+Four orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (D; white), near-resonant (A; black), weakly chaotic (C; black), and non-resonant (B; black) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).}
 \label{fig:ensemblemap-meandensity}
@@
 \begin{center}
-\includegraphics[width=1.1\textwidth]{figures/fdiff_zooms.png}
-\caption{Zooms of the frequency diffusion time map of Figure~\ref{fig:freqdiff} around the three orbits highlighted in that figure and used in Figures~\ref{fig:nbodysims}--\ref{fig:ensemble-density}---as in Figure~\ref{fig:freqdiff}, each pixel represents an orbit, and the greyscale of the pixel is set by the frequency diffusion time.
-Around each orbit, we initialize a new grid of $\approx$40,000 orbits to resolve the small-scale resonant structure around each orbit.
-Each orbit in the sub-grid is integrated for 80 orbital periods and the frequency diffusion time is computed from the two consecutive sections of 40 orbital periods.
-The marker at the center indicates the scale (e.g., tidal radius) of a $M \approx 10^4~\msun$ globular cluster at the given position in the initial condition grid.
-The regular orbit (R) is near a resonance (white line), the weakly chaotic orbit (W) is surrounded by an intricate resonant structure and may be in a region of resonance overlap, and the strongly chaotic orbit (S) is embedded in a region of strong stochasticity.
-\todo{apw}{Move colorbar to bottom or top?}}
+\includegraphics[width=0.9\textwidth]{figures/fdiff_zooms.png}
+\caption{Zooms of the frequency diffusion time map of Figure~\ref{fig:freqdiff} around the four orbits highlighted in that figure and used in Figures~\ref{fig:orbits}--\ref{fig:densities}.
+As with Figure~\ref{fig:freqdiff}, each pixel represents an orbit, and the greyscale of the pixel is set by the frequency diffusion time.
+Around each orbit, we initialize a new grid of $\approx$16,000 orbits to resolve the small-scale resonant structure around each orbit.
+The marker at the center indicates the scale (tidal radius) of a $M \approx 10^4~\msun$ globular cluster at the given position in the initial condition grid.
+The near-resonant orbit (A) is coincident with a high order resonance (white line), the non-resonant orbit (B) is between resonances, the weakly chaotic orbit (C) is surrounded by intricate structures and may be in a region of resonance overlap, and the strongly chaotic orbit (D) is embedded in a region of strong stochasticity.}
 \label{fig:fdiff-zooms}
@@
 % Figure ??
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/freqvar-map.png}
-\caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the amplitude of short-time frequency variations, $A_\Omega$ (Equation~\ref{eq:frac-freq-amp}).
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/freqvar-map.png}
+%\caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the amplitude of short-time frequency variations, $A_\Omega$ (Equation~\ref{eq:frac-freq-amp}).
 The amplitude is a fractional measurement, so that a value of $10^{-2}$ corresponds to a 1\% change in frequencies.
@@
 }
-\label{fig:freqvar_map}
-\end{center}
-\end{figure}
+%\label{fig:freqvar_map}
+%\end{center}
+%\end{figure}
 

### 2015-07-22 ms.tex
more todos

@@
 
-\todo{apw}{rename frequency-space structure or something, show the frequency plots for the three ensembles}
 \subsection{Part III: Short-time frequency evolution}\label{sec:results3}
@@
 
-As a demonstration of the small-scale frequency modulation, we consider again the three orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:three-orbits}).
-Figure~\ref{fig:fdiff-zooms} shows zoomed regions of the frequency diffusion time map of Figure~\ref{fig:freqdiff} centered on the three orbits---the pixel positions are now shown relative to the three parent orbits.
-The frequency diffusion times shown in these grids were computed with two consecutive windows of length 40 orbital periods.
+\todo{apw}{clean up}
+As a demonstration of the small-scale frequency modulation, we consider again the four orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:orbit-info}).
 The resonant structure around the mildly chaotic orbit (white lines in panel W) bound the orbit to a small region, whereas for the strongly chaotic orbit (S), the entire region surrounding the orbit is strongly chaotic.
@@
 
-\subsection{Part IV: Proposed indicator of chaotic density evolution}\label{sec:results4}
-
-We repeat the above analysis of the short-time evolution of frequencies for all orbits in the orbit grid used above (Section~\ref{sec:results1}): for each orbit, we integrate for 80 orbital periods, but compute the frequencies now in a rolling window along this integration time.
-We fix the window width to 40 orbital periods and shift the window by 0.5 orbital periods between each computation of the frequencies.
-For each orbit, we have the three frequencies (and amplitudes), $k=1,2,3$, computed at times $j$.
-We compute the fractional amplitude of the frequency variation for a given orbit as
-\begin{equation}
-	A_\Omega = \max_{a_k} \left(\frac{\max_j \Omega^{(j)}_{k} - \min_j \Omega^{(j)}_{k}}{2 \, \mean{\Omega^{(j)}_{k}}_j}\right) \label{eq:frac-freq-amp}
-\end{equation}\todo{apw}{change to RMS?}
-where, as in Equation~\ref{eq:fdtime}, the outer maximum is taken with respect to the amplitude of the frequency in the power-spectrum, $a_k$, and the inner maximum and minimum are taken over time, $j$.
-This is a measure of the magnitude of the short-time frequency evolution for a single orbit.
-For tidal debris with small spreads in orbital properties, we expect nearby orbits to have similar behavior; for example, the central marker in Figure~\ref{fig:fdiff-zooms} shows the size of a globular-cluster-like system in configuration space.
-We therefore expect a small ensemble of orbits in frequency space to expand over short times around even weakly chaotic parent orbits and the debris will therefore appear dynamically hotter in real-space.
+We therefore expect a small ensemble of orbits in frequency space to expand over short times around even weakly chaotic parent orbits and the debris should therefore appear dynamically hotter in real-space.
+
+\todo{apw}{insert here the bit about the frequency distributions and chaotic evolution orthogonal to freq distributions}
 
@@
 Note that the axis scales are quite different between each figure---the strongly chaotic orbit ensemble evolves quickly due to its proximity to many overlapping resonances.
-
-Figure~\ref{fig:freqvar_map} shows the grid of initial conditions, now colored by the (fractional) amplitude of the short-time frequency variation, $A_\Omega$ (Equation~\ref{eq:frac-freq-amp}).
-When the fractional excursions of the frequencies are of order $10^{-3}$--$10^{-2}$ or larger, globular cluster debris will appear more diffuse due to enhanced density evolution from fast, small-scale chaotic diffusion.
-Comparing this to the timescales of Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}, it is clear that the small-scale variations of the frequencies of a progenitor orbit over a given window of time provides a better prediction of the resulting morphology of tidal debris.
 
@@
 Our main results and conclusions are summarized as follows:
+\todo{apw}{revisit this...}
 \begin{enumerate}
@@
 	\item the details of the enhanced mixing along weakly chaotic orbits depends on the resonant structure of the potential;
-	\item the variance of the fundamental frequencies of an orbit over a given time window is a strong predictor of the importance of small-scale chaotic frequency diffusion on the resulting morphology of tidal debris.
+	\item the variance of the fundamental frequencies of an orbit over a given time window may be a better predictor of the importance of small-scale chaotic frequency diffusion on the resulting morphology of tidal debris.
 \end{enumerate}
@@
 % Figure ??
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=0.9\textwidth]{figures/fdiff_zooms.png}
-\caption{Zooms of the frequency diffusion time map of Figure~\ref{fig:freqdiff} around the four orbits highlighted in that figure and used in Figures~\ref{fig:orbits}--\ref{fig:densities}.
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=0.9\textwidth]{figures/fdiff_zooms.png}
+%\caption{Zooms of the frequency diffusion time map of Figure~\ref{fig:freqdiff} around the four orbits highlighted in that figure and used in Figures~\ref{fig:orbits}--\ref{fig:densities}.
 As with Figure~\ref{fig:freqdiff}, each pixel represents an orbit, and the greyscale of the pixel is set by the frequency diffusion time.
@@
 The near-resonant orbit (A) is coincident with a high order resonance (white line), the non-resonant orbit (B) is between resonances, the weakly chaotic orbit (C) is surrounded by intricate structures and may be in a region of resonance overlap, and the strongly chaotic orbit (D) is embedded in a region of strong stochasticity.}
-\label{fig:fdiff-zooms}
-\end{center}
-\end{figure}
+%\label{fig:fdiff-zooms}
+%\end{center}
+%\end{figure}
 

### 2015-07-23 ms.tex
figure swap

@@
 We use a window with a width equal to 40 orbital periods and shift the window by an orbital period between each calculation of the most significant frequencies---again, this is the smallest window size we could choose before \superfreq\ begins to occasionally fail to recover the frequencies for the strongly chaotic orbit.
-Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time---plotted are the percent deviations of the frequency ratios from the initial value.
+Figure~\ref{fig:four-orbits-freqs} shows plots of the frequency ratios as computed in each window of time---plotted are the percent deviations of the frequency ratios from the initial value.
 The difference between the first window (green +) and the last window (red x) is 50 orbital periods.
@@
 The frequency variations of the strongly chaotic orbit (S) can be many 10s of per cent.}
-\label{fig:three-orbits-freqs}
-\end{center}
-\end{figure}
-
-% Figure ??
+\label{fig:four-orbits-freqs}
+\end{center}
+\end{figure}
+
 \clearpage
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-regular.png}
-\caption{Evolution of the fundamental frequency ratios, $r_{1,3} = \Omega_1 / \Omega_3$ and $r_{2,3} = \Omega_2 / \Omega_3$, over 16 orbital periods for an ensemble of 1000 orbits generated around the regular orbit, R, with mass scale $m=10^4~\msun$ (Section~\ref{sec:results2}).
+\includegraphics[width=\textwidth]{figures/ensemble-freqs.png}
+\caption{ } 
+\label{fig:ensemble-freqs}
+\end{center}
+\end{figure}
+
+% Figure ??
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-regular.png}
+%\caption{Evolution of the fundamental frequency ratios, $r_{1,3} = \Omega_1 / \Omega_3$ and $r_{2,3} = \Omega_2 / \Omega_3$, over 16 orbital periods for an ensemble of 1000 orbits generated around the regular orbit, R, with mass scale $m=10^4~\msun$ (Section~\ref{sec:results2}).
 Fundamental frequencies are computed with a window width of $\approx$40 orbital periods, and the window is shifted by one orbital period between each computation (each grey point represents the fundamental frequency ratios computed in a single window).
@@
 For this orbit (R), all panels are identical because the orbit is regular (the frequencies don't evolve).}
-\label{fig:regular-ensemble-freq-evolution}
-\end{center}
-\end{figure}
-
-% Figure ??
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-mildly_chaotic.png}
-\caption{Same as Figure~\ref{fig:regular-ensemble-freq-evolution}, but for the weakly chaotic orbit (W).
+%\label{fig:regular-ensemble-freq-evolution}
+%\end{center}
+%\end{figure}
+%
+%% Figure ??
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-mildly_chaotic.png}
+%\caption{Same as Figure~\ref{fig:regular-ensemble-freq-evolution}, but for the weakly chaotic orbit (W).
 Frequencies now display short-time variations (compare left to right panels).
 Points in middle panel show the trajectories of each ensemble orbit in frequency-space.}
-\label{fig:mildly_chaotic-ensemble-freq-evolution}
-\end{center}
-\end{figure}
-
-% Figure ??
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-chaotic.png}
-\caption{Same as Figures~\ref{fig:regular-ensemble-freq-evolution}--\ref{fig:mildly_chaotic-ensemble-freq-evolution}, but for the strongly chaotic orbit (S).
+%\label{fig:mildly_chaotic-ensemble-freq-evolution}
+%\end{center}
+%\end{figure}
+%
+%% Figure ??
+%\clearpage
+%\begin{figure}[p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-chaotic.png}
+%\caption{Same as Figures~\ref{fig:regular-ensemble-freq-evolution}--\ref{fig:mildly_chaotic-ensemble-freq-evolution}, but for the strongly chaotic orbit (S).
 Frequencies are now explore a huge volume representing $>$100\% deviations from the initial value of the parent orbit.
 Frequencies evolve quickly around resonances---especially the $\bs{n}=(2,-3,0)$, $\bs{n}=(1,-1,0)$, and $\bs{n}=(1,0,1)$ resonances---which appear as linear features in these panels.}
-\label{fig:chaotic-ensemble-freq-evolution}
-\end{center}
-\end{figure}
+%\label{fig:chaotic-ensemble-freq-evolution}
+%\end{center}
+%\end{figure}
 

### 2015-07-24 ms.tex
density evolution

@@
 Streams in chaotic regions look very different from those in regular regions.
-We find that simulated streams (simulated using test particle ensembles of nearby orbits) are sensitive to chaos on a much shorter time-scale than any standard prediction (from the Lyapunov or frequency-diffusion times).
-For example, on a mildly chaotic orbit with a chaotic timescale predicted to be $>$1000 orbital periods ($>$1000 Gyr), the resulting stellar stream is, after just 64 orbits, substantially more diffuse than any formed on a nearby but regular orbit.
+We find that simulated streams (simulated using test particle ensembles of nearby orbits) can be sensitive to chaos on a much shorter time-scale than any standard prediction (from the Lyapunov or frequency-diffusion times).
+For example, on a weakly chaotic orbit with a chaotic timescale predicted to be $>$1000 orbital periods ($>$1000 Gyr), the resulting stellar stream is, after just 64 orbits, substantially more diffuse than any formed on a nearby but regular orbit.
 We find that the enhanced diffusion of the stream stars can be understood by looking at the variance in orbital frequencies of orbit ensembles centered around the parent (progenitor) orbit.
@@
 
-An ensemble of regular orbits (e.g., tidal debris) will phase-mix due to (small) differences in the fundamental frequencies of the orbits.
+An ensemble of regular orbits (e.g., tidal debris) will phase-mix because of (small) differences in the fundamental frequencies of the orbits.
 The frequency distributions of thin streams are generally close to one-dimensional---that is, one eigenvalue of the distribution of frequencies for an orbit ensemble will be much larger than the others because of the local shape of the Hessian of the potential \citep{helmi99, sanders13a, bovy14}.
-Phase-mixing generically leads to power-law decay of the mean density of the ensembles: initially, the density decreases linearly in time because of the large, nearly one-dimensional spread in frequencies, then may proceed as $t^{-2}$ to $t^{-3}$ depending on the other eigenvalues of the frequency-space distribution \citep[e.g.,][]{helmi99}.
+Phase-mixing generically leads to power-law decay of the mean density of the ensembles: initially, the density decreases linearly in time because of the large, nearly one-dimensional spread in frequencies, then may proceed as $t^{-2}$ to $t^{-3}$ depending on relative sizes of the other eigenvalues of the frequency-space distribution \citep[e.g.,][]{helmi99}.
 
@@
 Realistic galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors \citep{romanowsky98}, radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11,butsky15}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.; see, e.g.,][]{bailin05}.
-We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.
+We expect inclusion of any of these effects to increase the complexity of the resonant structure and influence of chaos (see Section~\ref{sec:discussion}).
 
@@
 
-\todo{apw}{numbers here on (units)}
-For each orbit, we also separately integrate for $\approx$256 orbital periods and use \superfreq\ to compute the fundamental frequencies for the two consecutive sections of 128 orbital periods.
-We have chosen this window size so that we recover the frequencies for regular orbits with fractional error $\approx10^{-8}$.
+For each orbit, we also separately integrate for 256 orbital periods and use \superfreq\ to compute the fundamental frequencies for the two consecutive sections of 128 orbital periods.
+We have chosen this window size so that we recover the frequencies for regular orbits with fractional error $\approx10^{-8}$ \citep[we estimate the error in frequency recovery using the method described in][]{laskar93}.
 With this integration window we are able to successfully recover frequencies using for $>$99.9\% of the orbits \superfreq.
@@
 The darker pixels have shorter frequency diffusion times and are more chaotic.
-This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential and highlights the accuracy of \superfreq\---weak chaos (grey) is detectable over just $\sim$100 orbital periods, compared to the many tens of thousands of orbits it would take to detect such features with the maximum Lyapunov exponent.
+This map reveals the intersection of this particular energy hypersurface with the rich structure of resonant surfaces present in this potential and highlights the accuracy of \superfreq\---weak chaos (grey) is detectable over much shorter integration periods using frequency analysis, compared to the many tens of thousands of orbits it would take to detect such features with the maximum Lyapunov exponent.
 The tube orbits in this potential are mostly regular or only mildly chaotic---the largest regular regions are associated with the short-axis and long-axis tube orbits---however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.
@@
 The strongest chaotic regions (black) appear in both of the above grids (Figure~\ref{fig:lyapmap} and Figure~\ref{fig:freqdiff}).
-Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map, especially where resonances overlap---for example, near $(x_0,z_0) \approx(13,7)~{\rm kpc}$ and $(x_0,z_0) \approx(31,13)~{\rm kpc}$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}.
+Some of the weakly chaotic unstable resonances do appear in the Lyapunov time map---for example, near $(x_0,z_0) / r_s \approx(1.5,0.6)$ and $(x_0,z_0) /r_s \approx(0.6,0.4)$ where there is a slight hint of weak chaos (grey) in Figure~\ref{fig:lyapmap}.
 The details of the resonant structure is revealed in the frequency diffusion map from integrations of just 256 orbital periods.
-
 While there is rich structure and a significant number of weakly chaotic orbits, the majority of the orbits have estimated chaotic timescales corresponding to thousands of orbital periods and are thus not expected to be relevant for tidal stream evolution.
@@
 
-The Lyapunov time and frequency diffusion rate measure the timescales over which chaos is relevant for a given orbit---that is, these quantities are measures of how infinitesimal deviations will diverge on average from some parent orbit, or of how long it takes for the frequencies of a single orbit to change by some amount.
+The Lyapunov and frequency diffusion times measure the timescales over which chaos is relevant for a given orbit---that is, these quantities are measures of how infinitesimal deviations will diverge on average from some parent orbit, or of how long it takes for the frequencies of a single orbit to change by some amount.
 Tidal debris is disrupted from progenitor systems with finite spreads in orbital properties (e.g., energy).
-For a disrupting, globular-cluster-scale progenitor, the typical energy dispersion of the debris is 0.1--1\% of the progenitor orbital energy \citep[assuming masses of $10^4$--$10^5$~\msun;][]{johnston98}, but for a dwarf-galaxy-scale progenitor, the dispersion can be $\gtrsim$10\%.
+For a disrupting, globular-cluster-scale progenitor, the typical energy or frequency-space dispersion of the debris is 0.1--1\% of the progenitor orbital energy \citep[assuming masses of $10^4$--$10^5$~\msun;][]{johnston98}, but for a dwarf-galaxy-scale progenitor, the dispersion can be $\gtrsim$10\%.
 In this section, we ask whether  the Lyapunov or frequency diffusion time predict the timescale over which a finite phase-space volume (e.g., tidal debris) stays coherent.
@@
 We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event.
-For a given set of orbital initial conditions---the `parent' orbit---we integrate for 128 orbital periods, find the phase-space position of the minimum pericenter over this times, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time.
+For a given set of orbital initial conditions---the `parent' orbit---we integrate for 128 orbital periods, find the phase-space position of the minimum pericenter over this time, initialize a small ensemble of test particle orbits around this position, and integrate the orbits of all test particles for some integration time.
 We are interested in the degree to which chaos enhances the mixing rate of orbit ensembles and we therefore want to isolate out the effect of ordinary phase-mixing along regular orbits; all ensembles should start with the same initial spread in fundamental frequencies.
@@
 Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the four orbits described above and evolved for 64 orbital periods.
-A thin stream forms on the near-resonant orbit (left column), a thin---but more two-dimensional--- stream forms on the non-resonant orbit (middle-left), a more diffuse stream forms on the weakly chaotic orbit with an interesting morphology (middle-right), and a `fanned' stream on the strongly chaotic orbit (right).
+A thin stream forms on the near-resonant orbit (left column), a thin---but more two-dimensional---stream forms on the non-resonant orbit (middle-left), a more diffuse stream forms on the weakly chaotic orbit with an interesting morphology (middle-right), and a `fanned' stream on the strongly chaotic orbit (right).
 Given the long Lyapunov and frequency diffusion times of the weakly chaotic orbit ($t_\lambda/\periods > 900$, $t_\Omega/\periods \approx 2\times10^5$), it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream.
 
-We verify that these orbit ensembles capture the nature of more realistic stream formation by running N-body simulations of globular-cluster-mass progenitor systems on these same three orbits.
+We verify that these orbit ensembles capture the nature of more realistic stream formation by running N-body simulations of globular-cluster-mass progenitor systems on these same four orbits.
 We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run the simulations, which we set up to run from apocenter-to-apocenter (rather than pericenter-to-apocenter as in the ensembles), but finish with the progenitor in the same location as the parent orbit in the ensemble evolution described above.
@@
 From a comparison of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims}, it is clear that the morphology of the ensembles is visually similar to the `oldest' (first stripped) debris in the N-body simulations.
-
-\todo{apw}{All lengths / distances in units of $r_s$, all velocities in units of $v_c$?}
 
@@
 
-Visual inspection of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims} suggests that chaotic mixing of small orbit ensembles affects the configuration-space evolution of an ensemble over short times, even when the predicted chaotic timescales (from the Lyapunov and frequency diffusion times) are long and therefore that the mean, single-orbit chaos indicators are not well-suited for determining the importance of chaotic diffusion on tidal stream evolution.
-As a quantitative measure of this enhanced density evolution, Figure~\ref{fig:ensemble-density} shows the evolution of the configuration-space density for ensembles evolved around the three orbits (R, W, S) described above.
+Visual inspection of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims} suggests that chaotic mixing of small orbit ensembles affects the configuration-space evolution of an ensemble over short times, even when the predicted chaotic timescales (from the Lyapunov and frequency diffusion times) are long: The mean, single-orbit chaos indicators are not well-suited for determining the importance of chaotic diffusion on tidal stream evolution.
+As a quantitative measure of this enhanced density evolution, we compute the evolution of the mean configuration-space density for orbit ensembles evolved around the four parent orbits (A,B,C,D) described above.
 At each time step, we use kernel density estimation (KDE)\footnote{We use an implementation from the Python package \texttt{scikit-learn} \citep{scikitlearn}.} with the ensemble of particle positions to estimate the configuration-space density field.
@@
 We evaluate the density at the positions of each particle, $\rho_i$, and compare to the mean initial density, $\mean{\rho_0}$.
-Figure~\ref{fig:ensemble-density} shows histograms of the logarithm of $\rho_i/\mean{\rho_0}$ for each of the three orbit ensembles at three different times: the initial density distribution (labeled $0~\periods$), and after evolution for 4 and 16 orbital periods.
-The distributions all shift towards lower densities as time progresses, however other subtle effects are noticeable.
-The regular orbit distribution stays relatively compact at all times indicating that the density stays fairly uniformly high throughout the stream.
-For the weakly chaotic orbit, the distribution is somewhat wider with a larger tail to lower densities---this is the debris that begins to fan out.
-The strongly chaotic orbit evolves much faster to low density.
-
-The density evolution can be complex, however with a single measure that characterizes the ensemble density evolution we can map the entire grid of orbits used in previous sections (e.g., Figure~\ref{fig:lyapmap}).
-Figure~\ref{fig:densities} shows the evolution of the mean density of the ensemble particles computed at 256 evenly-spaced intervals from the initial ensemble distribution to the distribution after 16 orbital periods.
-It is clear that the density of the regular orbit ensemble evolves slower than the weakly chaotic ensemble, and the strongly chaotic ensemble quickly becomes diffuse; \todo{apw}{WTF?} the mean density does not capture the full nature of the density evolution is apparently sensitive to the enhanced density evolution observed along the weakly chaotic orbit seen in Figure~\ref{fig:ensemble-density}.
-\todo{apw}{check above paragraph, add comments about lines that show power law behavior}
-
-Motivated by the noticeable discrepancy between the regular and weakly chaotic orbit ensembles---even though the chaotic timescale of the weakly chaotic ensemble parent orbit is several times the integration period used above (32 orbital periods)---we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
-For all ensembles, we integrate the orbits for 32 (parent) orbital periods and store the initial and final values of the mean, configuration-space density.
+
+Figure~\ref{fig:densities} shows the mean density of the ensemble particles computed at 256 evenly-spaced intervals from the initial ensemble distribution to the distribution after 64 orbital periods.
+Over-laid on each panel are qualitative power-laws (i.e.
+not fit to the density evolution) that demonstrate the expected trends: After initial $t^{-1}$ decay, the mean density of the near-resonant orbit (A) should follow $t^-2$, whereas the non-resonant orbit (B) will transition from $t^{-2}$ to $t^{-3}$ after long times (and thus currently display an intermediate power-law index).
+Given the extremely long chaotic timescale for the weakly chaotic orbit (C), we would expect the mean density to follow a simple power-law over long times, however at $\approx$16 orbital periods the density clearly diverges and begins to follow a decaying exponential.
+The mean density along the strongly chaotic orbit (D) evolves stochastically but generally follows a decaying exponential.
+
+Motivated by the noticeable discrepancy between the final mean density between the two regular (A,B) and the weakly chaotic (C) orbit ensembles---even though the chaotic timescale of the weakly chaotic ensemble parent orbit is several times the integration period used above---we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}.
+For all ensembles, we integrate the orbits for 64 (parent) orbital periods and compute the initial and final values of the mean, configuration-space density.
 Figure~\ref{fig:ensemblemap-meandensity} again shows the grid of initial conditions (Section~\ref{sec:results1}, but now the greyscale indicates the ratio of the final mean density to the initial density.
-Much of the structure that is visible in the frequency diffusion time map (Figure~\ref{fig:freqdiff}) is again visible in this map of the density evolution of orbit ensembles.
-However, it is surprising that these features are present: the chaotic timescales predicted from both the Lyapunov and frequency diffusion times were typically 100 to 1000s of orbital periods for many of the unstable resonance features (curved lines).
-\todo{apw}{Why don't these features show up?!?}
-
-We conclude from these experiments that the degree of chaos is an indicator that ensembles of orbits will mix faster than regular phase-mixing, however the Lyapunov time and frequency diffusion time are not good predictors for the timescale over which this mixing will occur.
-To understand this discrepancy, we next explore why this occurs (Section~\ref{sec:results3}, in the context of Section~\ref{sec:chaotic-mixing}) and develop a new indicator that can be tested over a finite number of orbital periods to predict whether chaotic diffusion will enhance the mixing of orbit ensembles (Section~\ref{sec:results4}).
+Interestingly, much of the structure that is visible in the upper-half of the frequency diffusion time map (Figure~\ref{fig:freqdiff}) is again visible in this map of the density evolution of orbit ensembles.
+Many stable resonances appear as lighter, curved features in the upper portion of this grid where the density remains systematically higher, and regions of strong resonance overlap lead to weak chaos and systematically lower density.
+However, it is surprising that these features are present: The chaotic timescales predicted from both the Lyapunov and frequency diffusion times were typically 100 to 1000s of orbital periods for many of the unstable features around resonances.
+The timescales in the lower portion of the grid (where the orbits are primarily stable, short-axis tube orbits) are simply too long to detect density differences over 64 orbital periods even from weak chaos in this particular choice of potential.
+
+It is worth noting that the \todo{apw}{comment on failure of frequency measurement in resonance layer from aliasing but not in stochastic layer ...}
+
+We conclude from these experiments that the degree of chaos is an indicator that ensembles of orbits may mix faster than predicted from regular phase-mixing, however the Lyapunov time and frequency diffusion time are not good predictors for the timescale over which this mixing will occur.
+To understand this discrepancy, we next explore why this occurs (Section~\ref{sec:results3}, in the context of Section~\ref{sec:chaotic-mixing}).
 
@@
 
-% Figure ??
+% Check frequency analysis precision
 \clearpage
@@
 
+% Figure - All the subgrid experiments
 \clearpage

### 2015-07-25 ms.tex
adding bits about frequency distribution

@@
 
-It is worth noting that the \todo{apw}{comment on failure of frequency measurement in resonance layer from aliasing but not in stochastic layer ...}
+It is worth noting that the frequency diffusion time map predicts that the resonances are surrounded by thick stochastic layers, while in the density map the resonance layers appear to be stable (and thus lead to higher mean densities).
+This comes from a failure of frequency determination: Near resonances it can take extremely long integration periods to ensure accurate recovery of the frequencies due to aliasing \citep{laskar03}.
+However, at the edges of the resonance layers---where we expect to find stochastic layers and do see weakly chaotic motion---the frequency diffusion time is a robust indicator of chaos.
 
@@
 
-Why does chaos manifest itself after just 16 orbital periods around orbits with predicted chaotic timescales equal to hundreds or thousands of orbital periods?
+Why does chaos manifest itself after just 64 orbital periods around orbits with predicted chaotic timescales larger than thousands of orbital periods?
 Both the Lyapunov exponent and the frequency diffusion rate measure mean, long-term rates of chaotic diffusion.
-If a weakly chaotic orbit is confined (by other nearby, non-overlapping resonances or stable regions), the mean drift of an orbit in frequency space may be small if computed over timescales long compared to the orbital time but short compared to the Arnold diffusion time.
-This window length was chosen in order to have a strong enough signal in the frequency spectrum of chaotic orbits to measure the most significant frequencies using \superfreq\,, but it erases the small-scale variation of the frequencies.
+If a weakly chaotic orbit is confined (by other nearby, non-overlapping resonances or stable regions), the mean drift of an orbit in frequency space may be small if computed over timescales long compared to the orbital time but short compared to the Arnold diffusion time, however small-scale variation of the frequencies may occur over much shorter timescales.
+% This window length was chosen in order to have a strong enough signal in the frequency spectrum of chaotic orbits to measure the most significant frequencies using \superfreq\,, but it erases the small-scale variation of the frequencies.
 % For the potential considered in this paper, it seems that these are comparable to the timescales for which the frequency diffusion rate of the previous section were computed ($\approx$100-1000 orbital periods).
 
-\todo{apw}{clean up}
 As a demonstration of the small-scale frequency modulation, we consider again the four orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:orbit-info}).
-The resonant structure around the mildly chaotic orbit (white lines in panel W) bound the orbit to a small region, whereas for the strongly chaotic orbit (S), the entire region surrounding the orbit is strongly chaotic.
 To resolve the short-time behavior of the frequency diffusion (corresponding to motion across or around resonance layers) we compute the frequencies in a series of rolling windows along each orbit.
-We use a window with a width equal to 40 orbital periods and shift the window by an orbital period between each calculation of the most significant frequencies---again, this is the smallest window size we could choose before \superfreq\ begins to occasionally fail to recover the frequencies for the strongly chaotic orbit.
-Figure~\ref{fig:four-orbits-freqs} shows plots of the frequency ratios as computed in each window of time---plotted are the percent deviations of the frequency ratios from the initial value.
-The difference between the first window (green +) and the last window (red x) is 50 orbital periods.
-The frequencies of the weakly chaotic orbit (panel labeled $W$) evolve quickly but are bounded to a small volume with fractional size $\approx$0.3\% (presumably by nearby resonant surfaces).
-This is comparable to the frequency spread in globular cluster tidal debris (0.1--0.5\%).
-%\todo{apw}{check numbers in previous 2 sentences}
+We use a window with a width equal to 64 orbital periods and shift the window by an orbital period between each calculation of the most significant frequencies.
+Figure~\ref{fig:four-orbits-freqs} shows plots of the frequencies of each orbit computed in each window of time---plotted in projection are the percent deviations of the frequency from the initial value.
+The difference between the first window (green +) and the last window (red x) is 128 orbital periods.
+For the two regular orbits, (A) and (B), the frequencies vary only from numerical issues when recovering the frequencies ($\ll$1\%) and thus the initial and final value markers overlap.
+The frequencies of the weakly chaotic orbit (C) evolve quickly but are bounded to a small volume with fractional size $\approx$1\% (presumably by nearby resonant surfaces); this is larger than the frequency spread in globular cluster tidal debris (0.1--0.5\%).
+The frequencies of the strongly chaotic orbit (D) also evolve quickly but fill a larger volume.
 
@@
 We therefore expect a small ensemble of orbits in frequency space to expand over short times around even weakly chaotic parent orbits and the debris should therefore appear dynamically hotter in real-space.
-
-\todo{apw}{insert here the bit about the frequency distributions and chaotic evolution orthogonal to freq distributions}
-
-We illustrate this phenomenon by repeating the above experiment for small ensembles of orbits around each parent orbit.
-As before, we compute the frequencies of each orbit in a series of overlapping windows to resolve the short-time evolution of the frequencies.
+The effect is especially significant if the chaotic evolution of the frequencies occurs orthogonal to the largest eigenvector of the distribution of frequencies or the local Hessian of the Hamiltonian.
+We illustrate this phenomenon by computing the frequencies for each orbit in small ensembles of orbits around each of the four orbits (A,B,C,D).
+\todo{apw}{Repeat figure 12 for next window?} We compute the frequencies of each orbit in a series of overlapping windows to resolve the short-time evolution of the frequencies.
 Figures~\ref{fig:regular-ensemble-freq-evolution}--\ref{fig:chaotic-ensemble-freq-evolution} show the evolution of the frequencies for all orbits in each ensemble generated around the three orbits, R, W, and S.
@@
 
+\todo{apw}{cite Ngan 2015}
 \section{Discussion: limitations and future work}\label{sec:discussion}

### 2015-07-29 ms.tex
added some references

@@
 \citet{pearson15} tested the stochasticity of the orbit of the progenitor that produced `fanned' debris by computing the Lyapunov exponent along this orbit but found that it is consistent with being regular over dynamically relevant timescales (many Hubble times).
-It has been shown previously that along some strongly chaotic orbits, tidal streams do form large, diffuse `fans' of debris \citep[e.g.,][]{fardal14}, however it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
+It has been shown previously that along some strongly chaotic orbits, tidal streams do form large, diffuse `fans' of debris \citep[e.g.,][]{fardal14, ngan15}, however it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
 The result from \citet{pearson15} suggests that even weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.
@@
 In Section~\ref{sec:methods}, we describe our choice of potential, method for numerical orbit integration, and introduce the chaos indicators used in this work.
-Our results are split into four subsections: in Section~\ref{sec:results1} we present iso-energy grids of orbits and discuss the orbit classes and chaotic timescales present in our potential; in Section~\ref{sec:results2} we study the density evolution of small ensembles of orbits around each orbit of the previous section; in Section~\ref{sec:results3} we describe the behavior of chaotic diffusion and use this to explain how chaos is relevant for tidal streams over short times; and in Section~\ref{sec:results4} we propose a new indicator for evaluating whether chaotic diffusion of orbits will be important for the morphological evolution of tidal debris.
+Our results are split into four subsections: in Section~\ref{sec:results1} we present iso-energy grids of orbits and discuss the orbit classes and chaotic timescales present in our potential; in Section~\ref{sec:results2} we study the density evolution of small ensembles of orbits around each orbit of the previous section; in Section~\ref{sec:results3} we describe the behavior of chaotic diffusion and use this to explain how chaos is relevant for tidal streams over short times.
 We discuss the implications of our results in Section~\ref{sec:discussion}, and conclude in Section~\ref{sec:conclusions}.
@@
 Over-laid on each panel are qualitative power-laws (i.e.
-not fit to the density evolution) that demonstrate the expected trends: After initial $t^{-1}$ decay, the mean density of the near-resonant orbit (A) should follow $t^-2$, whereas the non-resonant orbit (B) will transition from $t^{-2}$ to $t^{-3}$ after long times (and thus currently display an intermediate power-law index).
+not fit to the density evolution) that demonstrate the expected trends: After initial $t^{-1}$ decay, the mean density of the near-resonant orbit (A) should follow $t^{-2}$, whereas the non-resonant orbit (B) will transition from $t^{-2}$ to $t^{-3}$ after long times (and thus currently display an intermediate power-law index).
 Given the extremely long chaotic timescale for the weakly chaotic orbit (C), we would expect the mean density to follow a simple power-law over long times, however at $\approx$16 orbital periods the density clearly diverges and begins to follow a decaying exponential.
@@
 We illustrate this phenomenon by computing the frequencies for each orbit in small ensembles of orbits around each of the four orbits (A,B,C,D).
-\todo{apw}{Repeat figure 12 for next window?} We compute the frequencies of each orbit in a series of overlapping windows to resolve the short-time evolution of the frequencies.
-Figures~\ref{fig:regular-ensemble-freq-evolution}--\ref{fig:chaotic-ensemble-freq-evolution} show the evolution of the frequencies for all orbits in each ensemble generated around the three orbits, R, W, and S.
-In all panels in each of these figures, whats plotted are the per cent deviations of each frequency ratio from the parent orbit frequency ratios computed in the first time window.
-The left panels in each figure show these per cent deviations for all ensemble orbits around each of the three orbits.
-Middle panels show the full evolution of each ensemble orbit over 16 orbital periods, and final panel shows the final per cent deviations for all ensemble orbits (after 16 orbital periods).
-Note that the axis scales are quite different between each figure---the strongly chaotic orbit ensemble evolves quickly due to its proximity to many overlapping resonances.
-
-\todo{apw}{cite Ngan 2015}
+
+We generate orbit ensembles with 128 orbits around the four orbits and integrate for two consecutive windows of 128 orbital periods.
+Figures~\ref{fig:ensemble-freqs0} and \ref{fig:ensemble-freqs1} show the distribution of frequencies for all orbits in the ensembles for the four orbits in each of the two consecutive windows.
+Around the near-resonant orbit (A), the frequency distribution is nearly one-dimensional, which explains the thin, one-dimensional morphology of the ensemble in the left column of Figure~\ref{fig:ensembles}.
+The non-resonant orbit ensemble (B) is `thicker'---the ratio of the two largest eigenvalues of this distribution is larger---and therefore in real space, the ensemble begins to spread over two dimensions, as is also evident in the final ensemble debris morphology in the middle-left panel of Figure~\ref{fig:ensembles}.
+Non-resonant orbits in axisymmetric or triaxial potentials will generically have thicker frequency distributions and the configuration-space density of typical tidal debris will therefore decrease faster in such potentials \citep{helmi99}.
+The weakly chaotic orbit ensemble (C) is clearly more diffuse than the two regular ensembles: As is evident in Figure~\ref{fig:four-orbits-freqs}, the small-scale chaotic evolution of the frequencies---though bounded---increases the variance of the frequency distribution.
+By comparing Figure~\ref{fig:ensemble-freqs0} and \ref{fig:ensemble-freqs1}, it appears that the frequency distribution gets more diffuse until it uniformly fills the allowed volume.
+The strongly chaotic orbit ensemble (D) already has a large spread in Figure~\ref{fig:ensemble-freqs0}, and in Figure~\ref{fig:ensemble-freqs1} many of the frequencies have fractional frequency differences relative to the parent orbit $\approx$5--10\% (off of the plot).
+
 \section{Discussion: limitations and future work}\label{sec:discussion}
@@
 Instead, the collective effect of the subhaloes may instead act as a noise term in the Hamiltonian of any halo orbit.
-This subhalo-induced colored noise---which depends on the mass spectrum and distribution of subhaloes---may also act to simply increase the magnitude of chaos along orbits, and destabilize sufficiently non-resonant orbits \citep[see, e.g.,][]{kandrup00}.
+This subhalo-induced heating---which depends on the mass spectrum and distribution of subhaloes---may also act to simply increase the magnitude of chaos along orbits, and destabilize sufficiently non-resonant orbits \citep[see, e.g.,][]{kandrup00}.
 
@@
 We then study the density evolution of small ensembles of orbits generated around each orbit in the grid used for the previous experiment and find that along some orbits classified as weakly chaotic---with chaotic timescales of 100s of orbital periods---the orbit ensembles display enhanced density evolution and reach a lower overall density faster than orbit ensembles around nearby regular orbits (which mix due to phase-mixing alone).
-We explain this apparent discrepancy \todo{apw}{Hogg's comment is to clearly state the discrepancy (has to do with timescales)} by considering the nature of chaotic diffusion: the classical chaos indicators are most sensitive to the slow, Arnold diffusion process that can cause large changes to orbital properties, but small-scale frequency evolution occurs over much shorter times as chaotic orbits stochastically diffuse across the stochastic layers that surround many resonances.
-We propose a new measure of chaos with tidal stream morphological evolution in mind: the amplitude of small-scale, fractional changes in frequencies over a given number of orbital periods along a stream progenitor orbit.
-When this amplitude is comparable to or larger than the typical spread in orbital properties of tidal debris from the progenitor, the phase-space density of the debris will evolve faster to a state of lower density relative to nearby regular orbits.
+We explain this discrepancy between the predicted chaotic timescale and the observed effects of chaos on tidal debris by considering the nature of chaotic diffusion: the classical chaos indicators are most sensitive to the slow, Arnold diffusion process that can cause large changes to orbital properties, but small-scale frequency evolution occurs over much shorter times as chaotic orbits stochastically diffuse across the stochastic layers that surround many resonances.
+The fundamental frequencies of a weakly chaotic orbit therefore vary over a small region (bounded by nearby stable resonances), and when this amplitude is comparable to or larger than the typical spread in frequencies of tidal debris from the progenitor, the phase-space density of the debris will evolve faster to a state of lower density relative to nearby regular orbits.
 
 Our main results and conclusions are summarized as follows:
-\todo{apw}{revisit this...}
 \begin{enumerate}
@@
 	\item the details of the enhanced mixing along weakly chaotic orbits depends on the resonant structure of the potential;
-	\item the variance of the fundamental frequencies of an orbit over a given time window may be a better predictor of the importance of small-scale chaotic frequency diffusion on the resulting morphology of tidal debris.
+	\item the covariance of the fundamental frequencies of an orbit over a given time window may be a better predictor of the importance of small-scale chaotic frequency diffusion on the resulting morphology of tidal debris.
 \end{enumerate}
-\todo{apw}{make it clear: this is not about how much chaos, just timescale discrepancy}
 
@@
 
-APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
+APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ DGE 11-44155.
 This work was partially supported by the National Science Foundation under Grant No.
@@
 The central black band in the left panel and white band in the right panel are tube orbits with apocenter-to-pericenter ratios close to one.
-}
+The four arrows are explained in Section~\ref{sec:results2} and the caption of Figure~\ref{fig:lyapmap}.}
 \label{fig:apoper}
@@
 Chaotic orbits with $t_\lambda/\periods \gtrsim 700$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
-Four orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (D; white), near-resonant (A; black), weakly chaotic (C; black), and non-resonant (B; black) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:lyapmap}
+Four orbits are pointed out with arrows---from top to bottom, these are the strongly chaotic (D), near-resonant (A), weakly chaotic (C), and non-resonant (B) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:lyapmap}
 \end{center}
@@
 The frequencies are measured precisely so that small changes in the frequencies can be detected over just $\approx$10s of orbits.
-Four orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (D; white), near-resonant (A; black), weakly chaotic (C; black), and non-resonant (B; black) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
+Four orbits are pointed out with arrows---from top to bottom, these are the strongly chaotic (D), near-resonant (A), weakly chaotic (C), and non-resonant (B) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
 \end{center}
@@
 Given the extremely long chaotic timescales of the weakly chaotic orbits in the upper half of the grid, it is surprising that any of the high-order resonant structure is visible in this map.
-Four orbits are highlighted with square markers---from top to bottom, these are the strongly chaotic (D; white), near-resonant (A; black), weakly chaotic (C; black), and non-resonant (B; black) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).}
+Four orbits are pointed out with arrows---from top to bottom, these are the strongly chaotic (D), near-resonant (A), weakly chaotic (C), and non-resonant (B) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).}
 \label{fig:ensemblemap-meandensity}
@@
 \includegraphics[width=\textwidth]{figures/freq-evolution.png}
-\caption{Evolution of the fundamental frequency ratios, $r_{1,3} = \Omega_1 / \Omega_3$ and $r_{2,3} = \Omega_2 / \Omega_3$, computed over 256 orbital periods for the three orbits, R, W, S.
-Fundamental frequencies are computed with a window width of $\approx$40 orbital periods, and the window is shifted by one orbital period between each computation (each grey point represents the fundamental frequency ratios computed in a single window).
-Plotted are the per cent deviations of the frequency ratios from the value in the initial window---the initial value is shown as a green plus sign and the final value is shown as a red x.
-For the regular orbit (R), the fractional variation is around $10^{-6}$ and thus all points overlap on this scale.
-The weakly chaotic orbit (W) displays frequency variations comparable to the spread in frequencies in globular-cluster-like tidal debris.
-The frequency variations of the strongly chaotic orbit (S) can be many 10s of per cent.}
+\caption{Evolution of the fundamental frequencies computed over a total of 256 orbital periods for the four orbits, A, B, C, D.
+Plotted are the per cent deviations of the frequencies computed in window relative to the value in the initial window---that is, if $j$ is the index of a given time window, $\delta \Omega_{1,j} = (\Omega_{1,j} - \Omega_{1,0})/\Omega_{1,0}$.
+The initial value is shown as a green plus sign and the final value is shown as a red x.
+The frequencies are computed with a window width of $\approx$128 orbital periods, and the window is shifted by one orbital period between each computation (each grey point represents the fundamental frequencies computed in a single window).
+For the regular orbits (A,B), the fractional variation is around $10^{-6}$ and thus all points overlap on this scale.
+The weakly chaotic orbit (C) displays frequency variations comparable to the spread in frequencies in globular-cluster-like tidal debris.
+The frequency variations of the strongly chaotic orbit (D) have a larger characteristic spread.}
 \label{fig:four-orbits-freqs}
@@
 
-% Figure - All the subgrid experiments
+% Figure - Ensemble frequencies
 \clearpage
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-freqs.png}
-\caption{ } 
-\label{fig:ensemble-freqs}
+\includegraphics[width=\textwidth]{figures/ensemble-freqs-0.png}
+\caption{Distributions of the fundamental frequencies for all orbits in 128-orbit ensembles generated around the four orbits, A, B, C, D.
+Plotted are the per cent deviations of the frequencies of each ensemble orbit relative to the frequencies of the parent orbit---that is, if $i$ is the index of a given orbit and $i=0$ is the parent orbit, $\delta \Omega_{1,i} = (\Omega_{1,i} - \Omega_{1,0})/\Omega_{1,0}$.
+All orbits are integrated for 128 orbital periods to compute the frequencies.
+The near-resonant ensemble frequency distribution (A) is nearly 1D, and thus the ensemble appears one-dimensional in configuration-space (Figure~\ref{fig:ensembles}).
+The two largest eigenvalues of the non-resonant ensemble frequency distribution (B) are closer, and thus the ensemble spreads quicker, leading to a two-dimensional spread of debris in configuration-space.
+Around both the weakly chaotic orbit (C) and strongly chaotic orbit (D), chaotic diffusion increases the spread of the debris significantly, leading to faster 3D spreading in configuration-space.}
+\label{fig:ensemble-freqs0}
+\end{center}
+\end{figure}
+
+% Figure - Ensemble frequencies
+\clearpage
+\begin{figure}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/ensemble-freqs-1.png}
+\caption{Same as Figure~\ref{fig:ensemble-freqs0} but after integrating the same orbits for another 128 orbital periods.
+The ensemble frequency distributions around the regular orbits (A,B) appear nearly identical, whereas both the weakly and strongly chaotic ensemble frequency distributions spread significantly.}
+\label{fig:ensemble-freqs1}
 \end{center}
@@
 The weakly chaotic orbit has $A_\Omega \approx 0.2\%$ over just 16 orbital periods---this explains the enhanced density evolution observed in the N-body simulations and ensemble integrations (Figures~\ref{fig:nbodysims}-\ref{fig:ensembles}).
-The three orbits, R, W, S, are highlighted with square markers (S is top, W is middle, R is bottom).
+The three orbits, R, W, S, are pointed out with arrows (S is top, W is middle, R is bottom).
 }

### 2015-07-30 ms.tex
sarah comments

@@
 Streams in chaotic regions look very different from those in regular regions.
-We find that simulated streams (simulated using test particle ensembles of nearby orbits) can be sensitive to chaos on a much shorter time-scale than any standard prediction (from the Lyapunov or frequency-diffusion times).
-For example, on a weakly chaotic orbit with a chaotic timescale predicted to be $>$1000 orbital periods ($>$1000 Gyr), the resulting stellar stream is, after just 64 orbits, substantially more diffuse than any formed on a nearby but regular orbit.
+We find that streams (simulated using test particle ensembles of nearby orbits) can be sensitive to chaos on a much shorter time-scale than any standard prediction (from the Lyapunov or frequency-diffusion times).
+For example, on a weakly chaotic orbit with a chaotic timescale predicted to be $>$1000 orbital periods ($>$1000 Gyr), the resulting stellar stream is, after just a few 10's of orbits, substantially more diffuse than any formed on a nearby but regular orbit.
 We find that the enhanced diffusion of the stream stars can be understood by looking at the variance in orbital frequencies of orbit ensembles centered around the parent (progenitor) orbit.
@@
 \citet{pearson15} tested the stochasticity of the orbit of the progenitor that produced `fanned' debris by computing the Lyapunov exponent along this orbit but found that it is consistent with being regular over dynamically relevant timescales (many Hubble times).
-It has been shown previously that along some strongly chaotic orbits, tidal streams do form large, diffuse `fans' of debris \citep[e.g.,][]{fardal14, ngan15}, however it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
+It has been shown previously that along some strongly chaotic orbits and in live cosmological haloes, tidal streams do form large, diffuse `fans' of debris \citep[e.g.,][]{fardal14, ngan15}, however it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity.
 The result from \citet{pearson15} suggests that even weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/log-freqmap.png}
+\includegraphics[width=\textwidth]{figures/log-freqmap.pdf}
 \caption{ A reproduction of Figure~3.45 from \citep{binneytremaine} as a validation of our frequency analysis code: Frequency ratios for 100000 isoenergy orbits integrated in a triaxial logarithmic potential.
@@
 %\begin{center}
-%\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/ic-map.png}
+%\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/ic-map.pdf}
 %\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
@@
 %\begin{center}
-%\includegraphics[width=\textwidth]{figures/ic-orbits-T.png}
+%\includegraphics[width=\textwidth]{figures/ic-orbits-T.pdf}
 %\caption{ Visualization of the four major regular orbit types in a triaxial potential: outer long-axis tubes (O), inner long-axis tubes (I), short-axis tubes (S), and box or boxlets (B).
@@
     \subfloat{
-      \includegraphics[width=0.45\textwidth]{figures/pericenter-map.png}
+      \includegraphics[width=0.45\textwidth]{figures/pericenter-map.pdf}
     }
     \subfloat{
-      \includegraphics[width=0.45\textwidth]{figures/apocenter-map.png}
+      \includegraphics[width=0.45\textwidth]{figures/apocenter-map.pdf}
     }
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/lyap_map.png}
+\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/lyap_map.pdf}
 \caption{ The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the logarithm of the Lyapunov time (in units of orbital periods).
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/fdiff_map.png}
+\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/fdiff_map.pdf}
 \caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/four-orbits.png}
+\includegraphics[width=\textwidth]{figures/four-orbits.pdf}
 \caption{Four representative orbits chosen from the orbit grid (e.g., Figure~\ref{fig:apoper}): a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a strongly chaotic orbit (D).
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/ensembles.png}
+\includegraphics[width=\textwidth]{figures/ensembles.pdf}
 \caption{Final particle positions after integrating unbound, globular-cluster-sized ensembles of orbits generated around each of the three N-body orbits (Figure~\ref{fig:nbodysims}).
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/nbody.png}
+\includegraphics[width=\textwidth]{figures/nbody.pdf}
 \caption{ We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on the three orbits of Figure~\ref{fig:ensembles}.
@@
 %\begin{center}
-%\includegraphics[width=\textwidth]{figures/ensemble-density.png}
+%\includegraphics[width=\textwidth]{figures/ensemble-density.pdf}
 %\caption{ Distributions of the values of the estimated configuration-space density field evaluated at the positions of the ensemble particles after 0, 4, and 16 orbital periods for each of the three orbits of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims}.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-densities.png}
+\includegraphics[width=\textwidth]{figures/ensemble-densities.pdf}
 \caption{Evolution of the mean density (normalized by the initial density) of orbit ensembles around the four orbits (Figure~\ref{fig:orbits}) over 64 orbital periods.
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth, trim={0 0.2cm 0 0}]{figures/ensemble-map.png}
+\includegraphics[width=0.9\textwidth, trim={0 0.2cm 0 0}]{figures/ensemble-map.pdf}
 \caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now the greyscale intensity is set by the mean density of a of small ensembles of orbits after integration for 64 orbital periods.
@@
 %\begin{center}
-%\includegraphics[width=0.9\textwidth]{figures/fdiff_zooms.png}
+%\includegraphics[width=0.9\textwidth]{figures/fdiff_zooms.pdf}
 %\caption{Zooms of the frequency diffusion time map of Figure~\ref{fig:freqdiff} around the four orbits highlighted in that figure and used in Figures~\ref{fig:orbits}--\ref{fig:densities}.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/freq-evolution.png}
+\includegraphics[width=\textwidth]{figures/freq-evolution.pdf}
 \caption{Evolution of the fundamental frequencies computed over a total of 256 orbital periods for the four orbits, A, B, C, D.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-freqs-0.png}
+\includegraphics[width=\textwidth]{figures/ensemble-freqs-0.pdf}
 \caption{Distributions of the fundamental frequencies for all orbits in 128-orbit ensembles generated around the four orbits, A, B, C, D.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-freqs-1.png}
+\includegraphics[width=\textwidth]{figures/ensemble-freqs-1.pdf}
 \caption{Same as Figure~\ref{fig:ensemble-freqs0} but after integrating the same orbits for another 128 orbital periods.
@@
 %\begin{center}
-%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-regular.png}
+%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-regular.pdf}
 %\caption{Evolution of the fundamental frequency ratios, $r_{1,3} = \Omega_1 / \Omega_3$ and $r_{2,3} = \Omega_2 / \Omega_3$, over 16 orbital periods for an ensemble of 1000 orbits generated around the regular orbit, R, with mass scale $m=10^4~\msun$ (Section~\ref{sec:results2}).
@@
 %\begin{center}
-%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-mildly_chaotic.png}
+%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-mildly_chaotic.pdf}
 %\caption{Same as Figure~\ref{fig:regular-ensemble-freq-evolution}, but for the weakly chaotic orbit (W).
@@
 %\begin{center}
-%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-chaotic.png}
+%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-chaotic.pdf}
 %\caption{Same as Figures~\ref{fig:regular-ensemble-freq-evolution}--\ref{fig:mildly_chaotic-ensemble-freq-evolution}, but for the strongly chaotic orbit (S).
@@
 %\begin{center}
-%\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/freqvar-map.png}
+%\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/freqvar-map.pdf}
 %\caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the amplitude of short-time frequency variations, $A_\Omega$ (Equation~\ref{eq:frac-freq-amp}).

### 2015-07-30 ms.tex
kvj final comments and xsede ack

@@
 
-The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the axisymmetric potential.
+The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the oblate potential.
 It is well known that the number of degrees of freedom of a potential plays a critical role in determining the orbit structure of the potential; Hamiltonians with more than two degrees of freedom generically contain significant chaotic regions.
@@
 We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles through the chaotic diffusion of the individual orbits.
-For a chaotic orbit, the frequency spectrum of the orbit evolves with time: For a small ensemble, the spread in frequencies is therefore time-dependent, which could enhance phase-mixing.
+For a chaotic orbit, the frequency spectrum of the orbit evolves with time: for a small ensemble, the spread in frequencies is therefore time-dependent, which could enhance phase-mixing.
 This idea supports a reevaluation of the importance of chaos in galactic haloes and implies that the amount of chaos in a given potential may have significant consequences for the observability and survivability of thin, cold tidal streams.
-We propose a new measure for determining whether this small-scale frequency diffusion will be important for the morphological evolution of tidal debris with a given spread in frequencies or orbital properties over finite and dynamically relevant times.
 
@@
 In Section~\ref{sec:methods}, we describe our choice of potential, method for numerical orbit integration, and introduce the chaos indicators used in this work.
-Our results are split into four subsections: in Section~\ref{sec:results1} we present iso-energy grids of orbits and discuss the orbit classes and chaotic timescales present in our potential; in Section~\ref{sec:results2} we study the density evolution of small ensembles of orbits around each orbit of the previous section; in Section~\ref{sec:results3} we describe the behavior of chaotic diffusion and use this to explain how chaos is relevant for tidal streams over short times.
+Our results are split into three subsections: in Section~\ref{sec:results1} we present iso-energy grids of orbits and discuss the orbit classes and chaotic timescales present in our potential; in Section~\ref{sec:results2} we study the density evolution of small ensembles of orbits around each orbit of the previous section; in Section~\ref{sec:results3} we describe the behavior of chaotic diffusion and use this to explain how chaos is relevant for tidal streams over short times.
 We discuss the implications of our results in Section~\ref{sec:discussion}, and conclude in Section~\ref{sec:conclusions}.
@@
 
-We integrate all orbits in the grid for 10000 orbital periods and use the method described in Section~\ref{sec:lyap} to compute the Lyapunov exponents.
-Figure~\ref{fig:lyapmap} again shows the grid of initial conditions, but now the color corresponds to the logarithm of the Lyapunov time.
+We integrate all orbits in the grid for 10000 orbital periods and use the method described in Section~\ref{sec:lyap} to compute the MLEs.
+Figure~\ref{fig:lyapmap} again shows the grid of initial conditions, but now the color corresponds to the logarithm of the inverse of the MLE (the Lyapunov time).
 The darker pixels have shorter Lyapunov times and are more chaotic.
@@
 Tidal debris is disrupted from progenitor systems with finite spreads in orbital properties (e.g., energy).
-For a disrupting, globular-cluster-scale progenitor, the typical energy or frequency-space dispersion of the debris is 0.1--1\% of the progenitor orbital energy \citep[assuming masses of $10^4$--$10^5$~\msun;][]{johnston98}, but for a dwarf-galaxy-scale progenitor, the dispersion can be $\gtrsim$10\%.
+For a disrupting, globular-cluster-scale progenitor, the typical energy or frequency-space dispersion of the debris is 0.1--1\% of the progenitor orbital energy \citep[assuming masses of $10^4$--$10^6$~\msun;][]{johnston98}, but for a dwarf-galaxy-scale progenitor, the dispersion can be $\sim$10\%.
 In this section, we ask whether  the Lyapunov or frequency diffusion time predict the timescale over which a finite phase-space volume (e.g., tidal debris) stays coherent.
@@
 We set the physical scale of the ensemble by the tidal radius in position and the velocity scale in velocity; the initial spread in fundamental frequencies will scale as $(m/M)^{1/3}$ like the tidal radius and velocity scale \citep[e.g.,][]{johnston98, apw14}.
-If $(\bs{x},\bs{v})$ are a set of initial conditions at pericenter for a parent orbit, then $\delta x_{k,i}$ and $\delta v_{k,i}$ are the $k$th components ($k=1,2,3$) of the position and velocity deviation vectors for the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
+If $(\bs{x}_0,\bs{v}_0)$ are a set of initial conditions at pericenter for a parent orbit, then $\delta x_{k,i}$ and $\delta v_{k,i}$ are the $k$th components ($k=1,2,3$) of the position and velocity deviation vectors for the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
 \begin{align}
@@
 Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the four orbits described above and evolved for 64 orbital periods.
-A thin stream forms on the near-resonant orbit (left column), a thin---but more two-dimensional---stream forms on the non-resonant orbit (middle-left), a more diffuse stream forms on the weakly chaotic orbit with an interesting morphology (middle-right), and a `fanned' stream on the strongly chaotic orbit (right).
+A thin stream forms on the near-resonant orbit (left column), a thin---but more two-dimensional---stream forms on the non-resonant orbit (middle-left), a more diffuse stream forms on the weakly chaotic orbit with a slightly `fanned' morphology (middle-right), and a two-dimensional, `fanned' stream on the strongly chaotic orbit (right).
 Given the long Lyapunov and frequency diffusion times of the weakly chaotic orbit ($t_\lambda/\periods > 900$, $t_\Omega/\periods \approx 2\times10^5$), it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream.
@@
 \begin{center}
-	\begin{tabular}{c | c c c c c c | c c }
+	\begin{tabular}{c | c c c c c c | c c}
 		{\bf ID} & $\bs{x/r_s}$ & $\bs{y/r_s}$ & $\bs{z/r_s}$ & $\bs{v_x/v_c}$ & $\bs{v_y/v_c}$ & $\bs{v_z/v_c}$ & $\bs{t_\lambda/\periods}$ & $\bs{t_\Omega/\periods}$ \\\toprule
-A & 0.85 & 0 & 1.303 & 0 & 0.721 & 0 & $>700$ & $>10^7$\\
+A & 0.85 & 0 & 1.303 & 0 & 0.721 & 0 & $>700$ & $>10^7$ \\
 \midrule
@@
 Interestingly, much of the structure that is visible in the upper-half of the frequency diffusion time map (Figure~\ref{fig:freqdiff}) is again visible in this map of the density evolution of orbit ensembles.
-Many stable resonances appear as lighter, curved features in the upper portion of this grid where the density remains systematically higher, and regions of strong resonance overlap lead to weak chaos and systematically lower density.
+Many features appear as lighter, curved features in the upper portion of this grid where the density remains systematically higher---these are the stable resonances of the potential.
+Darker regions have systematically lower densities and correspond to regions of resonance overlap where weakly chaotic orbits are found.
+
 However, it is surprising that these features are present: The chaotic timescales predicted from both the Lyapunov and frequency diffusion times were typically 100 to 1000s of orbital periods for many of the unstable features around resonances.
@@
 Figure~\ref{fig:four-orbits-freqs} shows plots of the frequencies of each orbit computed in each window of time---plotted in projection are the percent deviations of the frequency from the initial value.
-The difference between the first window (green +) and the last window (red x) is 128 orbital periods.
+The difference between the first window (blue +) and the last window (red x) is 128 orbital periods.
 For the two regular orbits, (A) and (B), the frequencies vary only from numerical issues when recovering the frequencies ($\ll$1\%) and thus the initial and final value markers overlap.
@@
 We have shown that the Lyapunov and frequency diffusion times are indicators of chaos and that the frequency diffusion time resolves the detailed resonant structure of gravitational potentials, but the timescales predicted do not capture the importance of chaos for the density evolution of ensembles of orbits meant to mimic tidal debris.
-We have shown that small-scale but fast chaotic diffusion of orbits can explain this enhanced mixing, and have proposed a new indicator to characterize the importance of this effect over given evolution times.
+We have shown that small-scale but fast chaotic diffusion of orbits can explain this enhanced mixing.
 In the sub-sections below, we discuss a few important limitations that remain for exploration in future work.
@@
 The mass-scale of the debris that will display enhanced density evolution depends on the magnitude of weak chaos, which depends on the orbit structure of a given potential.
-In potentials with more significant chaos, debris disrupted from more massive progenitor systems may also display `stream-fanning.'
+In potentials with more significant chaos, debris disrupted from more massive progenitor systems may also display `stream-fanning.' However, since the scale of the debris is larger it is more likely that some orbits will exist in chaotic regions.
 
@@
 The fanned debris would be much lower surface brightness and thus much more difficult to detect.
-The detection of this low-surface-brightness component would open up the possibility that enhanced density evolution due to chaos is a dynamically relevant process for thin streams in real galaxies, though \cite{carlberg15} has shown that it is also possible also to get short, high density segments of streams from streams formed on eccentric orbits.
+The detection of this low-surface-brightness component would open up the possibility that enhanced density evolution due to chaos is a dynamically relevant process for thin streams in real galaxies, though \cite{carlberg15} has shown that it is also possible also to get short, high density segments of streams from debris formed on eccentric orbits.
 
@@
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ DGE 11-44155.
-This work was partially supported by the National Science Foundation under Grant No.
-AST-1312196.
-MV is supported in part by the University of MichiganÕs Elizabeth Crosby Fund
-and the Office of the VP for Research and NASA ATP grant NNX15AK79G.
+KVJ and APW were partially supported by the National Science Foundation under Grant No.
+AST-1312196 and NASA grant NNX15AK78G.
+MV is supported in part by the University of MichiganÕs Elizabeth Crosby Fund and the Office of the VP for Research and NASA ATP grant NNX15AK79G.
 AHWK would like to acknowledge support from NASA through Hubble Fellowship grant HST-HF-51323.01-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS 5-26555.
@@
 This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance.
+This work used the Extreme Science and Engineering Discovery Environment (XSEDE), which is supported by National Science Foundation grant number ACI-1053575 \citep{xsede}.
 
@@
 Plotted are the per cent deviations of the frequencies computed in window relative to the value in the initial window---that is, if $j$ is the index of a given time window, $\delta \Omega_{1,j} = (\Omega_{1,j} - \Omega_{1,0})/\Omega_{1,0}$.
-The initial value is shown as a green plus sign and the final value is shown as a red x.
+The initial value is shown as a blue plus sign and the final value is shown as a red x.
 The frequencies are computed with a window width of $\approx$128 orbital periods, and the window is shifted by one orbital period between each computation (each grey point represents the fundamental frequencies computed in a single window).

### 2015-07-30 ms.tex
old figures removed
remove the trim

@@
 All methods have been tested in some way on simulated observations of data and these tests typically demonstrate the recovery of parameters for analytic, static potential forms.
-%\citep[see the introduction of][]{kuepper15}
-% internal question: This is all just about the methods -- do I need sentences or a paragraph on results?
 
@@
 If a weakly chaotic orbit is confined (by other nearby, non-overlapping resonances or stable regions), the mean drift of an orbit in frequency space may be small if computed over timescales long compared to the orbital time but short compared to the Arnold diffusion time, however small-scale variation of the frequencies may occur over much shorter timescales.
-% This window length was chosen in order to have a strong enough signal in the frequency spectrum of chaotic orbits to measure the most significant frequencies using \superfreq\,, but it erases the small-scale variation of the frequencies.
-% For the potential considered in this paper, it seems that these are comparable to the timescales for which the frequency diffusion rate of the previous section were computed ($\approx$100-1000 orbital periods).
 
@@
 
-%The large-scale shapes of galactic gravitational potentials are thought to be triaxial and therefore contain an appreciable number of chaotic orbits.
-Chaos is not thought to be dynamically important for shaping the global structure of galaxies because the predicted chaotic timescales such as the Lyapunov or frequency diffusion times for orbits in galactic-like potentials are very long (many 1000s of orbital periods).
-However, these timescales measure the mean evolution of orbital properties such as the fundamental frequencies and do not capture the full nature of chaotic orbit diffusion.
-
 We have considered here a simple triaxial gravitational potential chosen to mimic the median properties of dark-matter haloes formed in dark-matter-only simulations (or the large-scale properties of haloes formed in simulations with baryonic effects).
@@
 AST-1312196 and NASA grant NNX15AK78G.
-MV is supported in part by the University of MichiganÕs Elizabeth Crosby Fund and the Office of the VP for Research and NASA ATP grant NNX15AK79G.
+MV is supported in part by the University of Michigan’s Elizabeth Crosby Fund and the Office of the VP for Research and NASA ATP grant NNX15AK79G.
 AHWK would like to acknowledge support from NASA through Hubble Fellowship grant HST-HF-51323.01-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS 5-26555.
@@
 
-%% Figure ??
-%\begin{figure*}[p]
-%\begin{center}
-%\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/ic-map.pdf}
-%\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
-Each pixel in this image represents a single orbit, and the shading (white/black) indicates the behavior of the orbit over 50 orbital periods: if the orbit preserves a sense of rotation around some axis---that is, the angular momentum around some axis does not change sign---the orbit is considered a tube orbit (white), otherwise the orbit is considered a box or chaotic orbit (black).
-The regions hosting the major orbit types are indicated by text labels.
-Figure~\ref{fig:icorbits} shows examples of each of these four orbit types.
-} \label{fig:icmap}
-%\end{center}
-%\end{figure*}
-%
-%% Figure ??
-%\begin{figure*}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/ic-orbits-T.pdf}
-%\caption{ Visualization of the four major regular orbit types in a triaxial potential: outer long-axis tubes (O), inner long-axis tubes (I), short-axis tubes (S), and box or boxlets (B).
-Shown are projections of these sample orbits onto the $xy$ and $yz$ planes and meant only to illustrate the general behavior of these orbit classes.} \label{fig:icorbits}
-%\end{center}
-%\end{figure*}
-
 % Figure ??
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/lyap_map.pdf}
+\includegraphics[width=0.9\textwidth, trim={0 0 0 0}]{figures/lyap_map.pdf}
 \caption{ The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the logarithm of the Lyapunov time (in units of orbital periods).
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/fdiff_map.pdf}
+\includegraphics[width=0.9\textwidth, trim={0 0 0 0}]{figures/fdiff_map.pdf}
 \caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
@@
 % Figure ??
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/ensemble-density.pdf}
-%\caption{ Distributions of the values of the estimated configuration-space density field evaluated at the positions of the ensemble particles after 0, 4, and 16 orbital periods for each of the three orbits of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims}.
-The density distribution of the regular orbit ensemble stays compact and decreases slowly.
-The density of the weakly chaotic ensemble initially evolves faster to lower density and has a larger tail to lower densities.
-The strongly chaotic ensemble evolves fast to a state of low density.} \label{fig:ensemble-density}
-%\end{center}
-%\end{figure}
-
-% Figure ??
 \clearpage
@@
 \end{figure}
-
-% Figure ??
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=0.9\textwidth]{figures/fdiff_zooms.pdf}
-%\caption{Zooms of the frequency diffusion time map of Figure~\ref{fig:freqdiff} around the four orbits highlighted in that figure and used in Figures~\ref{fig:orbits}--\ref{fig:densities}.
-As with Figure~\ref{fig:freqdiff}, each pixel represents an orbit, and the greyscale of the pixel is set by the frequency diffusion time.
-Around each orbit, we initialize a new grid of $\approx$16,000 orbits to resolve the small-scale resonant structure around each orbit.
-The marker at the center indicates the scale (tidal radius) of a $M \approx 10^4~\msun$ globular cluster at the given position in the initial condition grid.
-The near-resonant orbit (A) is coincident with a high order resonance (white line), the non-resonant orbit (B) is between resonances, the weakly chaotic orbit (C) is surrounded by intricate structures and may be in a region of resonance overlap, and the strongly chaotic orbit (D) is embedded in a region of strong stochasticity.}
-%\label{fig:fdiff-zooms}
-%\end{center}
-%\end{figure}
 
@@
 
-% Figure ??
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-regular.pdf}
-%\caption{Evolution of the fundamental frequency ratios, $r_{1,3} = \Omega_1 / \Omega_3$ and $r_{2,3} = \Omega_2 / \Omega_3$, over 16 orbital periods for an ensemble of 1000 orbits generated around the regular orbit, R, with mass scale $m=10^4~\msun$ (Section~\ref{sec:results2}).
-Fundamental frequencies are computed with a window width of $\approx$40 orbital periods, and the window is shifted by one orbital period between each computation (each grey point represents the fundamental frequency ratios computed in a single window).
-Plotted are the per cent deviations of the frequency ratios from the value in the initial window for the parent orbit.
-Left panel shows frequencies of each ensemble orbit for first time window.
-Middle panel shows the evolution of the frequencies over 16 orbital periods.
-Right panel shows final frequency values.
-For this orbit (R), all panels are identical because the orbit is regular (the frequencies don't evolve).}
-%\label{fig:regular-ensemble-freq-evolution}
-%\end{center}
-%\end{figure}
-%
-%% Figure ??
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-mildly_chaotic.pdf}
-%\caption{Same as Figure~\ref{fig:regular-ensemble-freq-evolution}, but for the weakly chaotic orbit (W).
-Frequencies now display short-time variations (compare left to right panels).
-Points in middle panel show the trajectories of each ensemble orbit in frequency-space.}
-%\label{fig:mildly_chaotic-ensemble-freq-evolution}
-%\end{center}
-%\end{figure}
-%
-%% Figure ??
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=\textwidth]{figures/ensemble-freq-evolution-chaotic.pdf}
-%\caption{Same as Figures~\ref{fig:regular-ensemble-freq-evolution}--\ref{fig:mildly_chaotic-ensemble-freq-evolution}, but for the strongly chaotic orbit (S).
-Frequencies are now explore a huge volume representing $>$100\% deviations from the initial value of the parent orbit.
-Frequencies evolve quickly around resonances---especially the $\bs{n}=(2,-3,0)$, $\bs{n}=(1,-1,0)$, and $\bs{n}=(1,0,1)$ resonances---which appear as linear features in these panels.}
-%\label{fig:chaotic-ensemble-freq-evolution}
-%\end{center}
-%\end{figure}
-
-% Figure ??
-%\clearpage
-%\begin{figure}[p]
-%\begin{center}
-%\includegraphics[width=0.9\textwidth, trim={0 1cm 0 0}]{figures/freqvar-map.pdf}
-%\caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the amplitude of short-time frequency variations, $A_\Omega$ (Equation~\ref{eq:frac-freq-amp}).
-The amplitude is a fractional measurement, so that a value of $10^{-2}$ corresponds to a 1\% change in frequencies.
-To compute the short-time frequency evolution, the frequencies are computed in a sliding window of width 40 orbital periods, and the window is shifted by half of an orbital period between each computation of the frequencies.
-The frequency evolution is computed for the equivalent of 16 orbital periods.
-The weakly chaotic orbit has $A_\Omega \approx 0.2\%$ over just 16 orbital periods---this explains the enhanced density evolution observed in the N-body simulations and ensemble integrations (Figures~\ref{fig:nbodysims}-\ref{fig:ensembles}).
-The three orbits, R, W, S, are pointed out with arrows (S is top, W is middle, R is bottom).
-}
-%\label{fig:freqvar_map}
-%\end{center}
-%\end{figure}
-
 \end{document}

### 2015-09-14 ms.tex
remove date
working in ref response

@@
 \newcommand{\superfreq}{\project{SuperFreq}}
+\newcommand{\chchchanges}[1]{{\color{red} {#1}}}
 
@@
 \newcommand{\jac}{\mathscr{J}}
+
+\usepackage{color}
 
@@
 	    David W. Hogg\altaffilmark{\nyu,\cds,\mpia}}
-\date{\centering \today}
+%\date{\centering \today}
 
@@
 \end{equation}
-where $q_i$ and $p_i$ are conjugate coordinates in the sense that
+\chchchanges{where $q_k$ and $p_k$ are conjugate coordinates in the sense that
 \begin{align}
-	\dot{p}_i &= -\frac{\partial H}{\partial q_i}\\
-	\dot{q}_i &= \frac{\partial H}{\partial p_i}
+	\dot{p}_k &= -\frac{\partial H}{\partial q_k}\\
+	\dot{q}_k &= \frac{\partial H}{\partial p_k}
 \end{align}
-for all $t$. If bounded, the motion in any component, $w_i(t)$, can be described as a Fourier sum,
+for all $t$. If bounded, the motion in any component, $w_k(t)$, can be described as a Fourier sum,
 \begin{equation}
-	w_i(t) = \sum_j^\infty a_j \, e^{i\,\omega_j\,t} \label{eq:fourier}
+	w_k(t) = \sum_j^\infty a_{kj} \, e^{i\,\omega_j\,t} \label{eq:fourier}
 \end{equation}
-where the $a_j$ are complex amplitudes.
+where the $a_{kj}$ are complex amplitudes.}
 
@@
 Even stronger, the actions are isolating integrals and any pair are in involution such that
-\begin{equation}
-	[J_i, J_j] = 0
-\end{equation}
+\chchchanges{\begin{equation}
+	[J_\alpha, J_\beta] = 0
+\end{equation}}
 where $[\cdot,\cdot]$ is the Poisson bracket.
@@
 The topology of angle-action space is toroidal and any regular orbit in an $N$ dof Hamiltonian can be understood as motion on the surface of an $N$-torus.
-Each set of actions, $(J_1,...,J_N)$ (or frequencies), uniquely labels a torus, and regular orbits are sometimes referred to in terms of their orbital tori.
+\chchchanges{Each set of actions, $(J_1,...,J_N)$ (or frequencies), labels a torus,} and regular orbits are sometimes referred to in terms of their orbital tori \chchchanges{(see, e.g., Section 50 in \citealt{arnold78}, Section 10-5 in \citealt{goldstein80}, Section 3.5 in \citealt{binneytremaine})}.
 
@@
 Galactic potentials are almost certainly not globally integrable but it is useful to understand the orbit structure in integrable systems before extending to more general potentials.
-Only one globally integrable triaxial mass model is known that generates motion that is qualitatively similar to that seen in Galaxies: the St\"ackel or `Perfect Ellipsoid' model \citep[e.g.,][]{kuzmin73, deZeeuw85}.
+Only one globally integrable triaxial mass model is known that generates motion that is qualitatively similar to that seen in Galaxies: \chchchanges{the `Perfect Ellipsoid' model \citep[a special case of the St\"ackel model; see, e.g.,][]{kuzmin73, deZeeuw85}}.
 The four general classes of orbits in this potential are box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits.
@@
 
-The frequencies of a generic orbit are typically incommensurable---that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} These conditionally-periodic orbits uniformly cover the surface of an orbital torus.
+The frequencies of a generic orbit are typically incommensurable---that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} These \chchchanges{non-resonant} orbits uniformly cover the surface of an orbital torus.
 If instead there exists a relation of the form $\boldsymbol{n} \cdot \boldsymbol{\Omega} = 0$ the orbit is referred to as a resonant orbit.
@@
 These resonant layers are then generically surrounded by stochastic layers where chaotic motion occurs (in the vicinity of separatrices).
-Chaotic orbits are characterized by exhibiting random behavior despite being generated by entirely deterministic equations of motion.
-Actions strictly do not exist for chaotic orbits and therefore chaos `destroys' orbital tori in certain regions of action-space---these frequency spectrum of orbits in these regions evolve chaotically with time.
+\chchchanges{Chaotic orbits are not bound to the surface of a torus and instead fill a (2N-1)-dimensional volume where a continuous set of tori would exist in an unperturbed Hamiltonian, $H_0$.
+While the frequency spectrum of sub-sections of a regular orbit are indistinguishable (i.e.
+measured over a finite interval of time), the frequency spectra of sub-sections of a chaotic orbit will evolve stochastically.}
 
@@
 Much of the structure that was present in the integrable potential remains in the near-integrable case, but the differences are highlighted.
-Orbits in the resonant layers surrounding the resonances (grey) are near-resonant orbits that librate around the resonance.
+Orbits in the resonant layers surrounding the resonances (grey) are near-resonant orbits that librate around the resonance \chchchanges{and have finite thickness \citep[e.g.,][]{merritt99}}.
 Chaotic orbits in the stochastic layer (red) behave erratically depending on the surrounding resonance structure.
 If the stochastic layer is small and the chaotic orbit is therefore confined, the orbit may behave nearly regular for long periods of time.
-If a resonance is unstable, all orbits associated with the resonance will be chaotic and this will drive orbits away from the resonance; unstable resonances form linear gaps in frequency-space (this is not shown in the cartoon).
+\chchchanges{If the resonance in the unperturbed Hamiltonian is unstable, all orbits associated with the resonance will be chaotic; unstable resonances form linear gaps in frequency-space (this is not shown in the cartoon).
+Thus, only some resonances in the perturbed system will retain the signature of a resonance.}
 
-For small values of $\epsilon$, many regular orbits survive and only small chaotic regions are introduced, especially where resonances overlap \citep[see][]{chirikov60}.
-As the strength of the perturbation increases, eventually most tori associated with conditionally-periodic motion will be destroyed.
-Figure~\ref{fig:cartoons} (right panel) qualitatively shows this phenomenon---the resonant and resonant-layer orbits may still be regular, but many or all of the conditionally-periodic orbits will be chaotic.
+For small values of $\epsilon$, many regular orbits survive and only small chaotic regions are introduced, \chchchanges{especially in the vicinity of the intersection of two resonance lines \citep[commonly referred to as `resonance overlap'; see][]{chirikov60}}.
+As the strength of the perturbation increases, eventually most tori associated with \chchchanges{non-resonant} motion will be destroyed.
+Figure~\ref{fig:cartoons} (right panel) qualitatively shows this phenomenon---the resonant and resonant-layer orbits may still be regular, but many or all of the \chchchanges{non-resonant} orbits will be chaotic.
 As the perturbation strength increases, eventually the uni- and bi-resonant tori are also destroyed---these are less susceptible to destruction from perturbations \cite[for a more quantitative illustration of this transition from integrability to global chaos, see Figure~9 in][]{valluri98}.
@@
 Chaotic orbits have no orbital actions and only conserve energy (if the potential is time-independent).
-The orbits therefore do not have a single set of fundamental frequencies, but rather the frequencies that describe the character of motion evolve with time.
-In near-integrable potentials, chaotic diffusion of an orbit in actions and frequencies occurs both across resonance layers (a sort of stochastic libration) and along the stochastic layers that surround the resonances (Arnold diffusion).
-For weakly chaotic orbits, motion across a resonance layer can occur with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
+The orbits therefore do not have a single set of fundamental frequencies, \chchchanges{but rather the frequencies that describe the character of motion evolve with time.} In near-integrable potentials, \chchchanges{the frequencies of consecutive sub-sections of a chaotic orbit diffuse both around resonance layers\footnote{Note that during this diffusion, the frequencies never exactly hit those of a KAM torus but evolve stochastically around these discrete, stable tori \citep[cf.
+Figure 2 in][]{laskar99}.} (a sort of stochastic libration) and along the stochastic layers that surround the resonances (Arnold diffusion).} For weakly chaotic orbits, motion across a resonance layer can occur with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
 Thus, if the resonance libration frequency is small, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
@@
 Arnold diffusion is not expected to be significant for most orbits over timescales relevant to galaxies (10s of orbits), however chaotic motion across resonances can occur over short times.
-If a stochastic orbit is semi-bounded by surrounding resonances, the allowed volume in which it can explore (in frequency space) is often quite small so that the orbit may appear to be approximately regular.
-However, this small-scale evolution may be important for tidal debris which can have small intrinsic spreads in frequencies.
+\chchchanges{If a stochastic trajectory is surrounded by regular orbits, it may appear similar to a nearby regular orbit (the `lead` orbit) over long periods of time before escaping to another such semi-bounded region where it will change in appearance to `trace' a new lead orbit (a process by which it may eventually explore the whole Arnold web).
+However, two initially nearby trajectories may switch to  different lead orbits at different times, a process which is clearly important for the evolution of tidal debris.
+Additionally, if the volume of the surrounded region (e.g., in frequency space) is comparable to the characteristic spread of frequencies in the tidal debris, this small-scale evolution will be important for tidal debris.}
 
@@
 We chose $r_s=20~{\rm kpc}$ and $v_c = 175~{\rm km}~{\rm s}^{-1}$ by taking the mean halo concentration for a ${\rm M}_{vir} \approx 10^{12}~\msun$ halo, $c_e\approx5$, from \cite{jing02} and by assuming $R_{vir}\approx200~{\rm kpc}$.
-Figure~\ref{fig:potential} shows equipotential contours of this potential in projection, and Table~\ref{tbl:potential} summarizes the potential parameters.
+Figure~\ref{fig:potential} shows equipotential contours of this potential, and Table~\ref{tbl:potential} summarizes the potential parameters.
 
@@
 This potential is a simple (and unrealistic) model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the haloes of such galaxies.
-Realistic galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors \citep{romanowsky98}, radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11,butsky15}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.; see, e.g.,][]{bailin05}.
+Realistic galactic potentials will have a significant component due to the disk and bulge, \chchchanges{radially changing axis ratios or orientations \citep[e.g.,][]{romanowsky98, kazantzidis04,debattista08,veraciro11,butsky15}}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.; see, e.g.,][]{bailin05}.
 We expect inclusion of any of these effects to increase the complexity of the resonant structure and influence of chaos (see Section~\ref{sec:discussion}).
@@
 
-The most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit.
+The most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit \chchchanges{\citep{lyapunov92}}.
 The MLE measures the mean rate of divergence of two infinitesimally separated orbits and is only strictly defined in terms of a limit that goes to infinite time.

### 2015-09-14 ms.tex
ref response

@@
 \renewcommand{\dateseparator}{-}
+\usepackage[normalem]{ulem}
 
@@
 
-For a small ensemble of orbits (e.g., tidal debris), the relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
+For a small ensemble of orbits (e.g., tidal debris), \chchchanges{a} relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
 We can estimate this timescale by multiplying the frequency diffusion time by a factor equal to the fractional spread in frequencies of the debris.
@@
 We classify each orbit in terms of the strength of chaos along the orbit as computed using the Lyapnov and frequency diffusion times of Section~\ref{sec:methods}.
-With this initial classification, in Section~\ref{sec:results2} we follow the evolution of ensembles of orbits generated around each orbit in the initial grid.
+With this initial classification, in Section~\ref{sec:results2} we follow the evolution of ensembles of \chchchanges{trajectories} generated around each orbit in the initial grid.
 We find that the configuration-space density of ensembles around weakly chaotic orbits evolve faster (e.g., in mean density) than expected given the timescales over which chaos is computed to be relevant for the parent orbits.
@@
 The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic.
-Tube orbits are generally less stochastic than box orbits, which tend to plunge deep into the inner regions of the potential where the density profile is steep, however there are appreciable numbers of stochastic tube orbits.
-The major classes of stable tube orbits circulate about either the major or minor axis.
+\chchchanges{Chaotic orbits that trace tube orbits are generally less stochastic than chaotic orbits that trace box orbits, which tend to plunge deep into the inner regions of the potential where the density profile is steep.} The major classes of stable tube orbits circulate about either the major or minor axis.
 This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits), but the most numerous orbits in this grid are the short-axis and outer long-axis tubes.
@@
 For a given set of orbital initial conditions---the `parent' orbit---we integrate for 128 orbital periods, find the phase-space position of the minimum pericenter over this time, initialize a small ensemble of test particle orbits around this position, and integrate the orbits of all test particles for some integration time.
-We are interested in the degree to which chaos enhances the mixing rate of orbit ensembles and we therefore want to isolate out the effect of ordinary phase-mixing along regular orbits; all ensembles should start with the same initial spread in fundamental frequencies.
+We are interested in the degree to which chaos enhances the mixing rate of orbit ensembles and we therefore want to isolate out the effect of ordinary phase-mixing along regular orbits\chchchanges{\sout{; all ensembles should start with the same initial spread in fundamental frequencies}}.
 We set the physical scale of the ensemble by the tidal radius in position and the velocity scale in velocity; the initial spread in fundamental frequencies will scale as $(m/M)^{1/3}$ like the tidal radius and velocity scale \citep[e.g.,][]{johnston98, apw14}.
@@
 
-We start by considering four particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a strongly chaotic orbit (D).
+We start by considering four particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a \chchchanges{more} strongly chaotic orbit (D).
 The orbits were chosen to be close on the orbit grid so that their orbital properties (e.g., apocenter, pericenter) are similar, but have different frequency diffusion times; all four orbits are long-axis tube orbits.
-Figure~\ref{fig:orbits} shows the orbits in projection over a long integration period (1024 orbital periods).
+Figure~\ref{fig:orbits} shows the orbits in projection over \chchchanges{an integration period of 1024 orbital periods}.
 The initial conditions and chaos diagnostics for each orbit are listed in Table~\ref{tbl:orbit-info}.
 Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the four orbits described above and evolved for 64 orbital periods.
-A thin stream forms on the near-resonant orbit (left column), a thin---but more two-dimensional---stream forms on the non-resonant orbit (middle-left), a more diffuse stream forms on the weakly chaotic orbit with a slightly `fanned' morphology (middle-right), and a two-dimensional, `fanned' stream on the strongly chaotic orbit (right).
+A thin stream forms on the near-resonant orbit (left column), a thin---but more two-dimensional---stream forms on the non-resonant orbit (middle-left), a more diffuse stream forms on the weakly chaotic orbit with a slightly `fanned' morphology (middle-right), and a two-dimensional, `fanned' stream on the \chchchanges{more} strongly chaotic orbit (right).
 Given the long Lyapunov and frequency diffusion times of the weakly chaotic orbit ($t_\lambda/\periods > 900$, $t_\Omega/\periods \approx 2\times10^5$), it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream.
@@
 The non-resonant orbit ensemble (B) is `thicker'---the ratio of the two largest eigenvalues of this distribution is larger---and therefore in real space, the ensemble begins to spread over two dimensions, as is also evident in the final ensemble debris morphology in the middle-left panel of Figure~\ref{fig:ensembles}.
-Non-resonant orbits in axisymmetric or triaxial potentials will generically have thicker frequency distributions and the configuration-space density of typical tidal debris will therefore decrease faster in such potentials \citep{helmi99}.
+Non-resonant orbits in axisymmetric or triaxial potentials \chchchanges{will generically have frequency distributions that have more comparable spreads in each direction} and the configuration-space density of typical tidal debris will therefore decrease faster in such potentials \citep{helmi99}.
 The weakly chaotic orbit ensemble (C) is clearly more diffuse than the two regular ensembles: As is evident in Figure~\ref{fig:four-orbits-freqs}, the small-scale chaotic evolution of the frequencies---though bounded---increases the variance of the frequency distribution.
@@
 The fanned debris would be much lower surface brightness and thus much more difficult to detect.
-The detection of this low-surface-brightness component would open up the possibility that enhanced density evolution due to chaos is a dynamically relevant process for thin streams in real galaxies, though \cite{carlberg15} has shown that it is also possible also to get short, high density segments of streams from debris formed on eccentric orbits.
+The detection of this low-surface-brightness component would open up the possibility that enhanced density evolution due to chaos is a dynamically relevant process for thin streams in real galaxies, though \cite{carlberg15} has shown that it is also possible to get short, high density segments of streams from debris formed on eccentric orbits.
 
@@
 \includegraphics[width=\textwidth]{figures/four-orbits.pdf}
-\caption{Four representative orbits chosen from the orbit grid (e.g., Figure~\ref{fig:apoper}): a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a strongly chaotic orbit (D).
+\caption{Four representative orbits chosen from the orbit grid (e.g., Figure~\ref{fig:apoper}): a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a more strongly chaotic orbit (D).
 All orbits are long-axis tubes with similar pericenters and apocenters.

### 2015-09-21 ms.tex
andreas comments

@@
 In integrable potentials---where all orbits are dynamically regular---tidal debris phase-mixes close to the orbit of the progenitor system. 
-However, cosmological simulations of structure formation suggest that the Milky Way's dark matter halo is expected not to be fully integrable; an appreciable fraction of orbits will be chaotic.
+\chchchanges{However, the Milky Way's dark matter halo is expected not to be fully integrable; an appreciable fraction of orbits will be chaotic.}
 This paper examines the influence of chaos on the phase-space morphology of cold tidal streams.
-We find very stark results:
-Streams in chaotic regions look very different from those in regular regions.
-We find that streams (simulated using test particle ensembles of nearby orbits) can be sensitive to chaos on a much shorter time-scale than any standard prediction (from the Lyapunov or frequency-diffusion times).
+\chchchanges{Streams in chaotic regions look very different from those in regular regions.}
+\chchchanges{We find that streams can be sensitive to chaos on a much shorter time-scale than any standard prediction (from the Lyapunov or frequency-diffusion times).}
 For example, on a weakly chaotic orbit with a chaotic timescale predicted to be $>$1000 orbital periods ($>$1000 Gyr), the resulting stellar stream is, after just a few 10's of orbits, substantially more diffuse than any formed on a nearby but regular orbit.
@@
 Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies; the known streams span a large range of distances---from $\approx10$ to $100~{\rm kpc}$---and progenitor masses---from $\approx$10$^3$ to $\approx$10$^8~\msun$ in stellar mass---\citep[][]{ibata94,odenkirchen01,belokurov06,grillmair06a,grillmair06b,bonaca12}.
-There has been extensive work on developing methods to use data from these streams to measure properties of Milky Way's dark matter halo.
+There has been extensive work on developing methods to use data from these streams to measure properties of \chchchanges{the} Milky Way's dark matter halo.
 These methods span a range of complexity from orbit-fitting, to \emph{Streakline} \citep{kuepper12} or `particle-spray' models \citep{gibbons14}, to action-space density modeling \citep[e.g.,][]{sanders14, bovy14}, to N-body simulations \citep[e.g.,][]{law10}.
@@
 We have only considered low-mass progenitor systems such as globular clusters because the intrinsic spreads in fundamental frequencies are small (0.1--0.5\%).
-Small changes to the frequencies of the orbits of tidal debris disrupted from these progenitors due to chaotic processes will therefore cause observable changes to the real-space morphology of the debris.
+Small changes to the frequencies of the orbits of tidal debris \chchchanges{stripped} from these progenitors due to chaotic processes will therefore cause observable changes to the real-space morphology of the debris.
 For more massive progenitor systems, the typical size and velocity dispersion of the debris will be larger and thus the debris morphology will be less sensitive to small changes in orbital properties.
@@
 Tidal stream modeling is one of the most promising ways to constrain the 3D mass distribution around the Milky Way at distances of 10s to 100s of kpc.
-Methods that use tidal streams to infer properties of the Galactic potential typically operate by constructing models of the debris distribution using either the present-day phase-space density \citep[e.g.,][]{kuepper12, kuepper15, gibbons14}, time-of-disruption phase-space density \citep{apw13, apw14}, or the density in angle-action coordinates \citep{sanders14, bovy14}.
+Methods that use tidal streams to infer properties of the Galactic potential typically operate by constructing models of the debris distribution using either the present-day phase-space density \citep[e.g.,][]{varghese11, kuepper12, kuepper15, gibbons14}, time-of-disruption phase-space density \citep{apw13, apw14}, or the density in angle-action coordinates \citep{sanders14, bovy14}.
 All of these methods may fail or produce uninterpretable results if modeling globular-cluster streams on mildly chaotic orbits.
@@
 At Galactocentric $R \approx 1$ kpc, $z \approx 5$ kpc, the orbits of the stream stars likely pass through the MW disk, feel the triaxiality of the Galactic bar \citep[e.g.,][]{wegg13, wegg15}, and have short orbital periods (relative to streams in the halo).
-It is possible that the observed debris is the last remnants of the recently disrupted progenitor system \citep{sesar15}, however if a significant number of stars were disrupted on previous pericentric passages, this older debris may be `fanned' and still near the observed portion of the stream.
-The fanned debris would be much lower surface brightness and thus much more difficult to detect.
-The detection of this low-surface-brightness component would open up the possibility that enhanced density evolution due to chaos is a dynamically relevant process for thin streams in real galaxies, though \cite{carlberg15} has shown that it is also possible to get short, high density segments of streams from debris formed on eccentric orbits.
+It is possible that the observed debris is the last remnants of the recently disrupted progenitor system \citep{sesar15}, however if a significant number of stars were \chchchanges{stripped} on previous pericentric passages, this older debris may be `fanned' and still near the observed portion of the stream.
+\chchchanges{The fanned debris would have significantly lower surface brightness and thus would be more difficult to detect.} The detection of this low-surface-brightness component would open up the possibility that enhanced density evolution due to chaos is a dynamically relevant process for thin streams in real galaxies, though \cite{carlberg15} has shown that it is also possible to get short, high density segments of streams from debris formed on eccentric orbits.
 

### 2015-09-21 ms.tex
added density plot

@@
 % Figure ??
-\begin{figure}[!p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/potential.pdf}
-\caption{Equipotential contours for the triaxial NFW potential considered in this work. There are eight contour levels evenly spaced and linear in the value of the potential. } \label{fig:potential}
-\end{center}
+%\begin{figure}[!p]
+%\begin{center}
+%\includegraphics[width=\textwidth]{figures/potential.pdf}
+%\caption{Equipotential contours for the triaxial NFW potential considered in this work. There are eight contour levels evenly spaced and linear in the value of the potential. } \label{fig:potential}
+%\end{center}
+%\end{figure}
+\begin{figure}[p]
+\centering
+    \subfloat{
+      \includegraphics[width=0.45\textwidth]{figures/potential.pdf}
+    }
+    \subfloat{
+      \includegraphics[width=0.45\textwidth]{figures/density-contours.pdf}
+    }
+\caption{\chchchanges{Equipotential contours (left) and isodensity contours (right) for the triaxial NFW potential considered in this work.
+For the potential plot there are eight contour levels evenly spaced and linear in the value of the potential.
+For the density plot there are eight contour levels logarithmically spaced from $10^4~\msun~{\rm kpc}^{-3}$ to $10^7~\msun~{\rm kpc}^{-3}$.}}
+\label{fig:potential}
 \end{figure}

### 2015-09-22 ms.tex
more ref crap

@@
 We choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos as computed from single-orbit diagnostics for grids of constant-energy orbits, and compare these results to measures of the density evolution of finite-volume ensembles of orbits (meant to mimic tidal debris).
-We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles through the chaotic diffusion of the individual orbits.
+We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles through the chaotic diffusion of the \chchchanges{constituent} orbits.
 For a chaotic orbit, the frequency spectrum of the orbit evolves with time: for a small ensemble, the spread in frequencies is therefore time-dependent, which could enhance phase-mixing.
@@
 The orbits therefore do not have a single set of fundamental frequencies, \chchchanges{but rather the frequencies that describe the character of motion evolve with time.} In near-integrable potentials, \chchchanges{the frequencies of consecutive sub-sections of a chaotic orbit diffuse both around resonance layers\footnote{Note that during this diffusion, the frequencies never exactly hit those of a KAM torus but evolve stochastically around these discrete, stable tori \citep[cf.
-Figure 2 in][]{laskar99}.} (a sort of stochastic libration) and along the stochastic layers that surround the resonances (Arnold diffusion).} For weakly chaotic orbits, motion across a resonance layer can occur with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
-Thus, if the resonance libration frequency is small, motion across a resonance can modulate the frequency spectrum of an orbit over an orbital time.
-However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can not change by large factors (unless there is resonance overlap and the motion is strongly chaotic).
+Figure 2 in][]{laskar99}.} (a sort of stochastic libration) and along the stochastic layers that surround the resonances (Arnold diffusion).} For weakly chaotic orbits, motion \chchchanges{around} a resonance layer can occur with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
+Thus, if the resonance libration frequency is small, motion \chchchanges{around} a resonance can modulate the frequency spectrum of an orbit over an orbital time.
+However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can not change by large factors (\chchchanges{unless there are other nearby resonances and overlapping stochastic layers, in which case the motion may be strongly chaotic}).
 
@@
 The initial conditions and chaos diagnostics for each orbit are listed in Table~\ref{tbl:orbit-info}.
-Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the four orbits described above and evolved for 64 orbital periods.
+Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the four orbits described above and \chchchanges{integrated} for 64 orbital periods.
 A thin stream forms on the near-resonant orbit (left column), a thin---but more two-dimensional---stream forms on the non-resonant orbit (middle-left), a more diffuse stream forms on the weakly chaotic orbit with a slightly `fanned' morphology (middle-right), and a two-dimensional, `fanned' stream on the \chchchanges{more} strongly chaotic orbit (right).
@@
 \cite{deibel11} showed that by adopting pattern speeds comparable to those found in cosmological simulations, figure rotation generally acts to destabilize orbits (rather than stabilize chaotic orbits in the equivalent static potential) and the resulting orbit structure is that most regular orbits are associated with resonances.
+\chchchanges{The presence of and response to the Large and Small Magellanic Clouds may also introduce significant (time-dependent) perturbations to the global potential of the Milky Way \citep[e.g.,][]{besla10, gomez15}}.
 In future work we will explore the effect of these time-dependent processes on the chaotic dispersal of tidal streams using live potentials from cosmological N-body simulations.
@@
 The progenitor in each simulation is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
-We start the progenitor systems at apocenter and evolve for $\approx$64 orbital periods so that each simulation finishes again at subsequent apocenter.
+We start the progenitor systems at apocenter and \chchchanges{integrate} for $\approx$64 orbital periods so that each simulation finishes again at subsequent apocenter.
 The mass of the progenitor is set to $10^4~\msun$, and the length-scale of the Plummer sphere is set to 5~pc to get $\approx$50--75\% mass loss over the integration time.

### 2015-10-06 ms.tex
minor
andreas comments

@@
 
-A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating Integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally---for example, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position \citep[e.g.,][]{goldstein80}.
+A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally---for example, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position \citep[e.g.,][]{goldstein80}.
 The condition for global integrability is very restrictive and the likelihood that a Hamiltonian is globally integrable decreases as the number of degrees of freedom increase \citep[e.g.,][]{lichtenberg83} and as the potential becomes more complex (i.e.
@@
 We have shown that the Lyapunov and frequency diffusion times are indicators of chaos and that the frequency diffusion time resolves the detailed resonant structure of gravitational potentials, but the timescales predicted do not capture the importance of chaos for the density evolution of ensembles of orbits meant to mimic tidal debris.
-We have shown that small-scale but fast chaotic diffusion of orbits can explain this enhanced mixing.
+We have shown that small-scale but fast chaotic diffusion of \chchchanges{frequencies} can explain this enhanced mixing.
 In the sub-sections below, we discuss a few important limitations that remain for exploration in future work.

### 2015-10-06 ms.tex
monicas comments

@@
 Several long, dynamically cold stellar streams have been observed around the Milky Way Galaxy, presumably formed from the tidal disruption of globular clusters.
-In integrable potentials---where all orbits are dynamically regular---tidal debris phase-mixes close to the orbit of the progenitor system. 
+In integrable potentials---where all orbits are regular---tidal debris phase-mixes close to the orbit of the progenitor system. 
 \chchchanges{However, the Milky Way's dark matter halo is expected not to be fully integrable; an appreciable fraction of orbits will be chaotic.}
@@
 As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded by tidal forces from the host-galaxy potential.
-Along regular, mildly eccentric orbits, mass is disrupted with small spreads in orbital properties (e.g., energy, angular momentum).
+Along regular, mildly eccentric orbits, \chchchanges{mass stripped from the progenitor has a small spread} in orbital properties (e.g., energy, angular momentum).
 Once the debris has evolved far enough from the progenitor system that the self-gravity of the progenitor can be ignored (usually a fast process relative to the orbital time), the stars evolve essentially as an ensemble of test particle orbits in the potential of the host system.
@@
 The four general classes of orbits in this potential are box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits.
-The tube orbits preserve a sense of rotation about either the long or short axis and are therefore centrophobic, whereas the box orbits may approach arbitrarily close to the center of the potential (centrophilic) and may change sense of rotation about any axis.
+\chchchanges{Tube orbits have a non-zero time-averaged angular momentum about either the long or short axis and therefore never pass through the center of the potential (hence are centrophobic).
+Box orbits instead have a zero time-averaged angular momentum and hence have finite probability of passing through the center of the potential (hence are centrophilic).}
 
@@
 If the stochastic layer is small and the chaotic orbit is therefore confined, the orbit may behave nearly regular for long periods of time.
-\chchchanges{If the resonance in the unperturbed Hamiltonian is unstable, all orbits associated with the resonance will be chaotic; unstable resonances form linear gaps in frequency-space (this is not shown in the cartoon).
+\chchchanges{If the resonance in the unperturbed Hamiltonian is unstable, all orbits associated with the resonance will be chaotic; unstable resonances form linear gaps in frequency-space (this is not shown in the cartoon but are seen in Figure~\ref{fig:logfreqs}).
 Thus, only some resonances in the perturbed system will retain the signature of a resonance.}
@@
 Arnold diffusion is not expected to be significant for most orbits over timescales relevant to galaxies (10s of orbits), however chaotic motion across resonances can occur over short times.
-\chchchanges{If a stochastic trajectory is surrounded by regular orbits, it may appear similar to a nearby regular orbit (the `lead` orbit) over long periods of time before escaping to another such semi-bounded region where it will change in appearance to `trace' a new lead orbit (a process by which it may eventually explore the whole Arnold web).
-However, two initially nearby trajectories may switch to  different lead orbits at different times, a process which is clearly important for the evolution of tidal debris.
-Additionally, if the volume of the surrounded region (e.g., in frequency space) is comparable to the characteristic spread of frequencies in the tidal debris, this small-scale evolution will be important for tidal debris.}
+\chchchanges{If a stochastic trajectory is surrounded by regular orbits, it may be trapped around a regular parent orbit for long periods of time before escaping to another such semi-bounded region where it can become trapped around another parent orbit (a process by which it may eventually explore the whole Arnold web)---such orbits are commonly referred to as `sticky orbits.' Additionally, if the volume of the surrounded region in frequency space is comparable to the characteristic spread of frequencies in the tidal debris, this small-scale evolution will be important for tidal debris.}
 
@@
 The frequency distributions of thin streams are generally close to one-dimensional---that is, one eigenvalue of the distribution of frequencies for an orbit ensemble will be much larger than the others because of the local shape of the Hessian of the potential \citep{helmi99, sanders13a, bovy14}.
-Phase-mixing generically leads to power-law decay of the mean density of the ensembles: initially, the density decreases linearly in time because of the large, nearly one-dimensional spread in frequencies, then may proceed as $t^{-2}$ to $t^{-3}$ depending on relative sizes of the other eigenvalues of the frequency-space distribution \citep[e.g.,][]{helmi99}.
+Phase-mixing generically leads to power-law decay of the mean density of the ensembles: initially, the density decreases linearly in time because of the large, nearly one-dimensional spread in frequencies, then may proceed as $t^{-2}$ to $t^{-3}$ depending on relative sizes of the other eigenvalues of the frequency-space distribution \citep[e.g.,][]{helmi99, vogelsberger08}.
 
@@
 For example, ensembles of orbits `stuck' between resonances may quickly spread to fill the allowed volume, but then the orbits must escape this confinement and diffuse through the Arnold web to reach a fully mixed state \citep{merritt96}.
-In this work, we investigate the consequences of short-time but small-scale frequency evolution and hypothesize that this may explain the enhanced density evolution of tidal debris around weakly chaotic regions where chaotic timescales are predicted to be long \citep[e.g.,][]{pearson15}.
+\chchchanges{That is, while the orbit is struck, the small-scale variations effectively cause an increase in the variance of the frequency distribution, which would enhance mixing of the debris in configuration-space.} In this work, we investigate the consequences of short-time but small-scale frequency evolution and hypothesize that this may explain the enhanced density evolution of tidal debris around weakly chaotic regions where chaotic timescales are predicted to be long \citep[e.g.,][]{pearson15}.
 We then discuss how this would affect our understanding of the coherence and density evolution of tidal streams.
@@
 The density distributions within dark-matter haloes formed in cosmological N-body simulations are generically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}.
-With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo and the mass and shape of the baryonic component \citep[e.g.,][though in Milky Way-like galaxies, baryonic disks will add non-sphericity to the total potential]{dubinski94,kazantzidis04, bryan13, butsky15}.
-It is less clear what happens to the outer halo.
+With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo and the mass and shape of the baryonic component \citep[e.g.,][though in Milky Way-like galaxies, baryonic disks will add non-sphericity to the total potential]{dubinski94,kazantzidis04, debattista08, bryan13, butsky15}.
+It is less clear what happens to the outer halo \citep[e.g.,][]{zemp11, valluri13}.
 
  \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution (dark-matter-only) N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these haloes.
-JS02 find median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{bett07, veraciro11, butsky15} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
+JS02 find median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{bett07, veraciro11, butsky15, zhu15} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}.
 JS02 find significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).
@@
 
-For a small ensemble of orbits (e.g., tidal debris), \chchchanges{a} relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
+For a small ensemble of orbits (e.g., tidal debris), \chchchanges{a more} relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
 We can estimate this timescale by multiplying the frequency diffusion time by a factor equal to the fractional spread in frequencies of the debris.
@@
 We fix $v_x = v_z = 0$, and compute $v_y$ from the energy.
-From here onwards, all lengths are given in units of scale radii, $r_s$, velocities in units of circular velocity, $v_c$, (see Table~\ref{tbl:potential}) and times in units of orbital periods, $\periods$.
-The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic.
-\chchchanges{Chaotic orbits that trace tube orbits are generally less stochastic than chaotic orbits that trace box orbits, which tend to plunge deep into the inner regions of the potential where the density profile is steep.} The major classes of stable tube orbits circulate about either the major or minor axis.
-This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits), but the most numerous orbits in this grid are the short-axis and outer long-axis tubes.
+This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits).
+\chchchanges{The most numerous orbits in this grid are the short-axis and long-axis tubes that circulate about either the minor or major axis.
 Thin tidal streams may preferentially form along tube orbits rather than box orbits because of the faster disruption and debris diffusion expected for stellar systems on radially plunging orbits.
 Figure~\ref{fig:apoper} shows the grid of initial conditions in the $xz$ plane---each pixel in this grid represents an orbit, and the pixels are colored by the median pericentric distance (left) and median apocentric distance (right) over an integration time of 64 orbital periods.
+From here onwards, all lengths are given in units of scale radii, $r_s$, velocities in units of circular velocity, $v_c$, (see Table~\ref{tbl:potential}) and times in units of orbital periods, $\periods$.}
 
@@
 We have chosen this window size so that we recover the frequencies for regular orbits with fractional error $\approx10^{-8}$ \citep[we estimate the error in frequency recovery using the method described in][]{laskar93}.
-With this integration window we are able to successfully recover frequencies using for $>$99.9\% of the orbits \superfreq.
+With this integration window we are able to successfully recover frequencies for $>$99.9\% of the orbits \superfreq.
 Figure~\ref{fig:freqdiff} shows the same grid of initial conditions as in Figure~\ref{fig:lyapmap}, but now the greyscale intensity is set by the logarithm of the frequency diffusion time.
@@
 Instead, the collective effect of the subhaloes may instead act as a noise term in the Hamiltonian of any halo orbit.
-This subhalo-induced heating---which depends on the mass spectrum and distribution of subhaloes---may also act to simply increase the magnitude of chaos along orbits, and destabilize sufficiently non-resonant orbits \citep[see, e.g.,][]{kandrup00}.
+This subhalo-induced heating---which depends on the mass spectrum and distribution of subhaloes---may also act to simply increase the magnitude of chaos along orbits, and destabilize sufficiently non-resonant orbits \citep[see, e.g.,][]{kandrup00, siegalgaskins08}.
 
@@
 AST-1312196 and NASA grant NNX15AK78G.
-MV is supported in part by the University of Michigan’s Elizabeth Crosby Fund and the Office of the VP for Research and NASA ATP grant NNX15AK79G.
+MV is supported in part by the University of Michigan's Elizabeth Crosby Fund and the Office of the VP for Research and NASA ATP grant NNX15AK79G.
 AHWK would like to acknowledge support from NASA through Hubble Fellowship grant HST-HF-51323.01-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS 5-26555.
@@
 \caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now the greyscale intensity is set by the mean density of a of small ensembles of orbits after integration for 64 orbital periods.
-The orbit ensembles are generated around each parent orbit in the grid (Section~\ref{sec:results2}) and the mean density is computed kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
+The orbit ensembles are generated around each parent orbit in the grid (Section~\ref{sec:results2}) and the mean density is a kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
 Given the extremely long chaotic timescales of the weakly chaotic orbits in the upper half of the grid, it is surprising that any of the high-order resonant structure is visible in this map.

### 2015-10-06 ms.tex
put figs in text

@@
 
+% Figure ??
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/cartoons.pdf}
+\caption{An illustrative demonstration of the orbit structure for integrable (left), near-integrable (middle), and non-integrable (right) potentials in terms of the ratios of the fundamental frequencies.
+In an integrable potential, only resonant and non-resonant orbits exist, and all orbits are regular (resonances appear as lines in frequency-ratio-space).
+If the potential is perturbed mildly, resonance layers form around the resonances that host near-resonant orbits which behave like resonant orbits but have an additional frequency corresponding to libration about the resonance.
+Where resonances overlap or near separatrices, stochastic layers form and orbits will be chaotic.
+For more strongly perturbed potentials, many of the non-resonant orbits may become chaotic but resonance layers may grow and still remain regular.
+}
+\label{fig:cartoons}
+\end{center}
+\end{figure}
+
 \subsection{Orbits in near- and non-integrable potentials}
@@
 \citet{leesuto03} derive a perturbative expansion of the potential integral for a triaxial NFW density and show that the expansion is accurate even for modest axis ratios (e.g., the median values shown above).
+
+\begin{figure}[h]
+\centering
+    \subfloat{
+      \includegraphics[width=0.45\textwidth]{figures/potential.pdf}
+    }
+    \subfloat{
+      \includegraphics[width=0.45\textwidth]{figures/density-contours.pdf}
+    }
+\caption{\chchchanges{Equipotential contours (left) and isodensity contours (right) for the triaxial NFW potential considered in this work.
+For the potential plot there are eight contour levels evenly spaced and linear in the value of the potential.
+For the density plot there are eight contour levels logarithmically spaced from $10^4~\msun~{\rm kpc}^{-3}$ to $10^7~\msun~{\rm kpc}^{-3}$.}}
+\label{fig:potential}
+\end{figure}
 
@@
 
+% Figure ??
+\begin{figure}[h]%[p]
+\centering
+    \subfloat{
+      \includegraphics[width=0.45\textwidth]{figures/pericenter-map.pdf}
+    }
+    \subfloat{
+      \includegraphics[width=0.45\textwidth]{figures/apocenter-map.pdf}
+    }
+\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
+All distances are normalized by the potential scale radius.
+Each pixel in these panels represents a single orbit, and the shading of each pixel corresponds to the median pericentric distance (left) or median apocentric distance (right) computed over 64 orbital periods.
+The central black band in the left panel and white band in the right panel are tube orbits with apocenter-to-pericenter ratios close to one.
+The four arrows are explained in Section~\ref{sec:results2} and the caption of Figure~\ref{fig:lyapmap}.}
+\label{fig:apoper}
+\end{figure}
+
 We integrate all orbits in the grid for 10000 orbital periods and use the method described in Section~\ref{sec:lyap} to compute the MLEs.
@@
 Because of the fixed the integration time, the MLE cannot detect weak chaos and the majority of orbits appear to be regular because they have exceedingly long Lyapunov times (all white points have $(t_\lambda/\periods) \gtrsim 1000$).
+
+% Figure ??
+\begin{figure}[!h]%[p]
+\begin{center}
+\includegraphics[width=0.8\textwidth, trim={0 0 0 0}]{figures/lyap_map.pdf}
+\caption{ The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the logarithm of the Lyapunov time (in units of orbital periods).
+Orbits are integrated for a total of 10000 orbital periods.
+Chaotic orbits with $t_\lambda/\periods \gtrsim 700$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
+Four orbits are pointed out with arrows---from top to bottom, these are the strongly chaotic (D), near-resonant (A), weakly chaotic (C), and non-resonant (B) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:lyapmap}
+\end{center}
+\end{figure}
 
@@
 The tube orbits in this potential are mostly regular or only mildly chaotic---the largest regular regions are associated with the short-axis and long-axis tube orbits---however islands of stronger chaos do appear, especially at the intersections of resonances where resonance overlap occurs.
+
+% Figure ??
+\begin{figure}[!h]%[p]
+\begin{center}
+\includegraphics[width=0.8\textwidth, trim={0 0 0 0}]{figures/fdiff_map.pdf}
+\caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
+The frequencies are computed in two consecutive windows, each of which has length equal to 40 orbital periods.
+The frequencies are measured precisely so that small changes in the frequencies can be detected over just $\approx$10s of orbits.
+Four orbits are pointed out with arrows---from top to bottom, these are the strongly chaotic (D), near-resonant (A), weakly chaotic (C), and non-resonant (B) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
+\end{center}
+\end{figure}
 
@@
 
+% Figure ??
+\begin{figure}[h]%[p]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/four-orbits.pdf}
+\caption{Four representative orbits chosen from the orbit grid (e.g., Figure~\ref{fig:apoper}): a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a more strongly chaotic orbit (D).
+All orbits are long-axis tubes with similar pericenters and apocenters.
+Orbits in this Figure were integrated for 1024 orbital periods and are shown in projection.}
+\label{fig:orbits}
+\end{center}
+\end{figure}
+
+% Figure ??
+\begin{figure}[h]%[p]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/ensembles.pdf}
+\caption{Final particle positions after integrating unbound, globular-cluster-sized ensembles of orbits generated around each of the three N-body orbits (Figure~\ref{fig:nbodysims}).
+The ensembles each contain 1024 particles and are initialized at pericenter.
+The particle positions qualitatively match the morphology of the `oldest' debris from the corresponding N-body simulations (Figure~\ref{fig:nbodysims}).
+}
+\label{fig:ensembles}
+\end{center}
+\end{figure}
+
 We verify that these orbit ensembles capture the nature of more realistic stream formation by running N-body simulations of globular-cluster-mass progenitor systems on these same four orbits.
@@
 
+% Figure ??
+\begin{figure}[h]%[p]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/nbody.pdf}
+\caption{ We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on the three orbits of Figure~\ref{fig:ensembles}.
+The progenitor in each simulation is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
+We start the progenitor systems at apocenter and \chchchanges{integrate} for $\approx$64 orbital periods so that each simulation finishes again at subsequent apocenter.
+The mass of the progenitor is set to $10^4~\msun$, and the length-scale of the Plummer sphere is set to 5~pc to get $\approx$50--75\% mass loss over the integration time.
+Panels show particle positions from the final snapshots of the simulations---for visualization the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
+These simulations confirm that the test-particle orbit ensembles (Figure~\ref{fig:ensembles}) do (qualitatively) capture the nature of the early-stripped tidal debris.
+}
+\label{fig:nbodysims}
+\end{center}
+\end{figure}
+
 Visual inspection of Figures~\ref{fig:ensembles} and \ref{fig:nbodysims} suggests that chaotic mixing of small orbit ensembles affects the configuration-space evolution of an ensemble over short times, even when the predicted chaotic timescales (from the Lyapunov and frequency diffusion times) are long: The mean, single-orbit chaos indicators are not well-suited for determining the importance of chaotic diffusion on tidal stream evolution.
@@
 
+% Figure ??
+\clearpage
+\begin{figure}[h]%[p]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/ensemble-densities.pdf}
+\caption{Evolution of the mean density (normalized by the initial density) of orbit ensembles around the four orbits (Figure~\ref{fig:orbits}) over 64 orbital periods.
+Over-plotted are various power-laws that qualitatively show the decay of the mean density for the ensembles: the mean density of the near-resonant orbit (A) follows nearly $t^{-2}$; the non-resonant orbit (B) is slowly transitioning from $t^{-2}$ to $t^{-3}$; the weakly chaotic orbit (C) initially follows $t^{-2}$ but then exponential decay takes over; the strongly chaotic orbit (D) is roughly exponential with fairly erratic density evolution.
+}
+\label{fig:densities}
+\end{center}
+\end{figure}
+
 However, it is surprising that these features are present: The chaotic timescales predicted from both the Lyapunov and frequency diffusion times were typically 100 to 1000s of orbital periods for many of the unstable features around resonances.
@@
 To understand this discrepancy, we next explore why this occurs (Section~\ref{sec:results3}, in the context of Section~\ref{sec:chaotic-mixing}).
+
+% Figure ??
+\clearpage
+\begin{figure}[h]%[p]
+\begin{center}
+\includegraphics[width=0.8\textwidth, trim={0 0.2cm 0 0}]{figures/ensemble-map.pdf}
+\caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now the greyscale intensity is set by the mean density of a of small ensembles of orbits after integration for 64 orbital periods.
+The orbit ensembles are generated around each parent orbit in the grid (Section~\ref{sec:results2}) and the mean density is a kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
+Given the extremely long chaotic timescales of the weakly chaotic orbits in the upper half of the grid, it is surprising that any of the high-order resonant structure is visible in this map.
+Four orbits are pointed out with arrows---from top to bottom, these are the strongly chaotic (D), near-resonant (A), weakly chaotic (C), and non-resonant (B) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).}
+\label{fig:ensemblemap-meandensity}
+\end{center}
+\end{figure}
 
@@
 The frequency diffusion time (Equation~\ref{eq:fdtime}) is an estimate of the time over which the mean value of the frequencies evolves, however the \emph{variance} of the frequencies over short times is dynamically relevant for small ensembles.
-This small-scale variability is insignificant for the evolution of global structure in, for example, an elliptical galaxy, but is signifiant for the morphological evolution of tidal debris with small spreads in frequencies.
+This small-scale variability is insignificant for the evolution of global structure in, for example, an elliptical galaxy or in erasing substructure in the Solar neighborhood \citep[e.g.,][]{maffione15}, but is signifiant for the morphological evolution of tidal debris with small spreads in frequencies.
 
@@
 The strongly chaotic orbit ensemble (D) already has a large spread in Figure~\ref{fig:ensemble-freqs0}, and in Figure~\ref{fig:ensemble-freqs1} many of the frequencies have fractional frequency differences relative to the parent orbit $\approx$5--10\% (off of the plot).
+
+% Check frequency analysis precision
+\clearpage
+\begin{figure}[h]%[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/freq-evolution.pdf}
+\caption{Evolution of the fundamental frequencies computed over a total of 256 orbital periods for the four orbits, A, B, C, D.
+Plotted are the per cent deviations of the frequencies computed in window relative to the value in the initial window---that is, if $j$ is the index of a given time window, $\delta \Omega_{1,j} = (\Omega_{1,j} - \Omega_{1,0})/\Omega_{1,0}$.
+The initial value is shown as a blue plus sign and the final value is shown as a red x.
+The frequencies are computed with a window width of $\approx$128 orbital periods, and the window is shifted by one orbital period between each computation (each grey point represents the fundamental frequencies computed in a single window).
+For the regular orbits (A,B), the fractional variation is around $10^{-6}$ and thus all points overlap on this scale.
+The weakly chaotic orbit (C) displays frequency variations comparable to the spread in frequencies in globular-cluster-like tidal debris.
+The frequency variations of the strongly chaotic orbit (D) have a larger characteristic spread.}
+\label{fig:four-orbits-freqs}
+\end{center}
+\end{figure}
+
+% Figure - Ensemble frequencies
+\clearpage
+\begin{figure}[h]%[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/ensemble-freqs-0.pdf}
+\caption{Distributions of the fundamental frequencies for all orbits in 128-orbit ensembles generated around the four orbits, A, B, C, D.
+Plotted are the per cent deviations of the frequencies of each ensemble orbit relative to the frequencies of the parent orbit---that is, if $i$ is the index of a given orbit and $i=0$ is the parent orbit, $\delta \Omega_{1,i} = (\Omega_{1,i} - \Omega_{1,0})/\Omega_{1,0}$.
+All orbits are integrated for 128 orbital periods to compute the frequencies.
+The near-resonant ensemble frequency distribution (A) is nearly 1D, and thus the ensemble appears one-dimensional in configuration-space (Figure~\ref{fig:ensembles}).
+The two largest eigenvalues of the non-resonant ensemble frequency distribution (B) are closer, and thus the ensemble spreads quicker, leading to a two-dimensional spread of debris in configuration-space.
+Around both the weakly chaotic orbit (C) and strongly chaotic orbit (D), chaotic diffusion increases the spread of the debris significantly, leading to faster 3D spreading in configuration-space.}
+\label{fig:ensemble-freqs0}
+\end{center}
+\end{figure}
+
+% Figure - Ensemble frequencies
+\clearpage
+\begin{figure}[h]%[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/ensemble-freqs-1.pdf}
+\caption{Same as Figure~\ref{fig:ensemble-freqs0} but after integrating the same orbits for another 128 orbital periods.
+The ensemble frequency distributions around the regular orbits (A,B) appear nearly identical, whereas both the weakly and strongly chaotic ensemble frequency distributions spread significantly.}
+\label{fig:ensemble-freqs1}
+\end{center}
+\end{figure}
 
@@
 
+\setcounter{figure}{0}
 \begin{figure}
@@
 \renewcommand{\thefigure}{\arabic{figure}}
-\setcounter{figure}{0}
-
-% Figure ??
-\begin{figure}[!p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/cartoons.pdf}
-\caption{An illustrative demonstration of the orbit structure for integrable (left), near-integrable (middle), and non-integrable (right) potentials in terms of the ratios of the fundamental frequencies.
-In an integrable potential, only resonant and non-resonant orbits exist, and all orbits are regular (resonances appear as lines in frequency-ratio-space).
-If the potential is perturbed mildly, resonance layers form around the resonances that host near-resonant orbits which behave like resonant orbits but have an additional frequency corresponding to libration about the resonance.
-Where resonances overlap or near separatrices, stochastic layers form and orbits will be chaotic.
-For more strongly perturbed potentials, many of the non-resonant orbits may become chaotic but resonance layers may grow and still remain regular.
-}
-\label{fig:cartoons}
-\end{center}
-\end{figure}
 
@@
 %\end{figure}
-\begin{figure}[p]
-\centering
-    \subfloat{
-      \includegraphics[width=0.45\textwidth]{figures/potential.pdf}
-    }
-    \subfloat{
-      \includegraphics[width=0.45\textwidth]{figures/density-contours.pdf}
-    }
-\caption{\chchchanges{Equipotential contours (left) and isodensity contours (right) for the triaxial NFW potential considered in this work.
-For the potential plot there are eight contour levels evenly spaced and linear in the value of the potential.
-For the density plot there are eight contour levels logarithmically spaced from $10^4~\msun~{\rm kpc}^{-3}$ to $10^7~\msun~{\rm kpc}^{-3}$.}}
-\label{fig:potential}
-\end{figure}
-
-% Figure ??
-\begin{figure}[p]
-\centering
-    \subfloat{
-      \includegraphics[width=0.45\textwidth]{figures/pericenter-map.pdf}
-    }
-    \subfloat{
-      \includegraphics[width=0.45\textwidth]{figures/apocenter-map.pdf}
-    }
-\caption{ A grid of isoenergy orbits initialized on the $xz$ plane.
-All distances are normalized by the potential scale radius.
-Each pixel in these panels represents a single orbit, and the shading of each pixel corresponds to the median pericentric distance (left) or median apocentric distance (right) computed over 64 orbital periods.
-The central black band in the left panel and white band in the right panel are tube orbits with apocenter-to-pericenter ratios close to one.
-The four arrows are explained in Section~\ref{sec:results2} and the caption of Figure~\ref{fig:lyapmap}.}
-\label{fig:apoper}
-\end{figure}
-
-% Figure ??
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=0.9\textwidth, trim={0 0 0 0}]{figures/lyap_map.pdf}
-\caption{ The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the logarithm of the Lyapunov time (in units of orbital periods).
-Orbits are integrated for a total of 10000 orbital periods.
-Chaotic orbits with $t_\lambda/\periods \gtrsim 700$ appear regular because the integration time for each orbit is insufficient to resolve weak chaos.
-Four orbits are pointed out with arrows---from top to bottom, these are the strongly chaotic (D), near-resonant (A), weakly chaotic (C), and non-resonant (B) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:lyapmap}
-\end{center}
-\end{figure}
-
-% Figure ??
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=0.9\textwidth, trim={0 0 0 0}]{figures/fdiff_map.pdf}
-\caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now each pixel is colored by the logarithm of the frequency diffusion time (again in units of orbital periods).
-The frequencies are computed in two consecutive windows, each of which has length equal to 40 orbital periods.
-The frequencies are measured precisely so that small changes in the frequencies can be detected over just $\approx$10s of orbits.
-Four orbits are pointed out with arrows---from top to bottom, these are the strongly chaotic (D), near-resonant (A), weakly chaotic (C), and non-resonant (B) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).} \label{fig:freqdiff}
-\end{center}
-\end{figure}
-
-% Figure ??
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/four-orbits.pdf}
-\caption{Four representative orbits chosen from the orbit grid (e.g., Figure~\ref{fig:apoper}): a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a more strongly chaotic orbit (D).
-All orbits are long-axis tubes with similar pericenters and apocenters.
-Orbits in this Figure were integrated for 1024 orbital periods and are shown in projection.}
-\label{fig:orbits}
-\end{center}
-\end{figure}
-
-% Figure ??
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/ensembles.pdf}
-\caption{Final particle positions after integrating unbound, globular-cluster-sized ensembles of orbits generated around each of the three N-body orbits (Figure~\ref{fig:nbodysims}).
-The ensembles each contain 1024 particles and are initialized at pericenter.
-The particle positions qualitatively match the morphology of the `oldest' debris from the corresponding N-body simulations (Figure~\ref{fig:nbodysims}).
-}
-\label{fig:ensembles}
-\end{center}
-\end{figure}
-
-% Figure ??
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/nbody.pdf}
-\caption{ We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on the three orbits of Figure~\ref{fig:ensembles}.
-The progenitor in each simulation is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
-We start the progenitor systems at apocenter and \chchchanges{integrate} for $\approx$64 orbital periods so that each simulation finishes again at subsequent apocenter.
-The mass of the progenitor is set to $10^4~\msun$, and the length-scale of the Plummer sphere is set to 5~pc to get $\approx$50--75\% mass loss over the integration time.
-Panels show particle positions from the final snapshots of the simulations---for visualization the positions have been rotated so that the angular momentum of the progenitor at the final snapshot are aligned with the $z$-axis.
-These simulations confirm that the test-particle orbit ensembles (Figure~\ref{fig:ensembles}) do (qualitatively) capture the nature of the early-stripped tidal debris.
-}
-\label{fig:nbodysims}
-\end{center}
-\end{figure}
-
-% Figure ??
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-densities.pdf}
-\caption{Evolution of the mean density (normalized by the initial density) of orbit ensembles around the four orbits (Figure~\ref{fig:orbits}) over 64 orbital periods.
-Over-plotted are various power-laws that qualitatively show the decay of the mean density for the ensembles: the mean density of the near-resonant orbit (A) follows nearly $t^{-2}$; the non-resonant orbit (B) is slowly transitioning from $t^{-2}$ to $t^{-3}$; the weakly chaotic orbit (C) initially follows $t^{-2}$ but then exponential decay takes over; the strongly chaotic orbit (D) is roughly exponential with fairly erratic density evolution.
-}
-\label{fig:densities}
-\end{center}
-\end{figure}
-
-% Figure ??
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=0.9\textwidth, trim={0 0.2cm 0 0}]{figures/ensemble-map.pdf}
-\caption{The same grid of orbits as shown in Figure~\ref{fig:apoper}, but now the greyscale intensity is set by the mean density of a of small ensembles of orbits after integration for 64 orbital periods.
-The orbit ensembles are generated around each parent orbit in the grid (Section~\ref{sec:results2}) and the mean density is a kernel density estimation (KDE) with an adaptive Epanechnikov kernel.
-Given the extremely long chaotic timescales of the weakly chaotic orbits in the upper half of the grid, it is surprising that any of the high-order resonant structure is visible in this map.
-Four orbits are pointed out with arrows---from top to bottom, these are the strongly chaotic (D), near-resonant (A), weakly chaotic (C), and non-resonant (B) orbits of Table~\ref{tbl:orbit-info} (see Section~\ref{sec:results2}).}
-\label{fig:ensemblemap-meandensity}
-\end{center}
-\end{figure}
-
-% Check frequency analysis precision
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/freq-evolution.pdf}
-\caption{Evolution of the fundamental frequencies computed over a total of 256 orbital periods for the four orbits, A, B, C, D.
-Plotted are the per cent deviations of the frequencies computed in window relative to the value in the initial window---that is, if $j$ is the index of a given time window, $\delta \Omega_{1,j} = (\Omega_{1,j} - \Omega_{1,0})/\Omega_{1,0}$.
-The initial value is shown as a blue plus sign and the final value is shown as a red x.
-The frequencies are computed with a window width of $\approx$128 orbital periods, and the window is shifted by one orbital period between each computation (each grey point represents the fundamental frequencies computed in a single window).
-For the regular orbits (A,B), the fractional variation is around $10^{-6}$ and thus all points overlap on this scale.
-The weakly chaotic orbit (C) displays frequency variations comparable to the spread in frequencies in globular-cluster-like tidal debris.
-The frequency variations of the strongly chaotic orbit (D) have a larger characteristic spread.}
-\label{fig:four-orbits-freqs}
-\end{center}
-\end{figure}
-
-% Figure - Ensemble frequencies
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-freqs-0.pdf}
-\caption{Distributions of the fundamental frequencies for all orbits in 128-orbit ensembles generated around the four orbits, A, B, C, D.
-Plotted are the per cent deviations of the frequencies of each ensemble orbit relative to the frequencies of the parent orbit---that is, if $i$ is the index of a given orbit and $i=0$ is the parent orbit, $\delta \Omega_{1,i} = (\Omega_{1,i} - \Omega_{1,0})/\Omega_{1,0}$.
-All orbits are integrated for 128 orbital periods to compute the frequencies.
-The near-resonant ensemble frequency distribution (A) is nearly 1D, and thus the ensemble appears one-dimensional in configuration-space (Figure~\ref{fig:ensembles}).
-The two largest eigenvalues of the non-resonant ensemble frequency distribution (B) are closer, and thus the ensemble spreads quicker, leading to a two-dimensional spread of debris in configuration-space.
-Around both the weakly chaotic orbit (C) and strongly chaotic orbit (D), chaotic diffusion increases the spread of the debris significantly, leading to faster 3D spreading in configuration-space.}
-\label{fig:ensemble-freqs0}
-\end{center}
-\end{figure}
-
-% Figure - Ensemble frequencies
-\clearpage
-\begin{figure}[p]
-\begin{center}
-\includegraphics[width=\textwidth]{figures/ensemble-freqs-1.pdf}
-\caption{Same as Figure~\ref{fig:ensemble-freqs0} but after integrating the same orbits for another 128 orbital periods.
-The ensemble frequency distributions around the regular orbits (A,B) appear nearly identical, whereas both the weakly and strongly chaotic ensemble frequency distributions spread significantly.}
-\label{fig:ensemble-freqs1}
-\end{center}
-\end{figure}
 

### 2016-04-06 ms.tex
whitespace and finally putting in proofs edits

@@
-%\documentclass{emulateapj}
-\documentclass[letterpaper,12pt,preprint]{aastex}
+\documentclass{mn2e}
+%\documentclass[letterpaper,12pt,preprint]{aastex}
 
@@
 \newcommand{\superfreq}{\project{SuperFreq}}
-\newcommand{\chchchanges}[1]{{\color{red} {#1}}}
+%\newcommand{\chchchanges}[1]{{\color{red} {#1}}}
+\newcommand{\chchchanges}[1]{#1}
 
@@
 In integrable potentials---where all orbits are regular---tidal debris phase-mixes close to the orbit of the progenitor system.
-\chchchanges{However, the Milky Way's dark matter halo is expected not to be fully integrable; an appreciable fraction of orbits will be chaotic.}
+However, the Milky Way's dark matter halo is expected not to be fully integrable; an appreciable fraction of orbits will be chaotic.
 This paper examines the influence of chaos on the phase-space morphology of cold tidal streams.
-\chchchanges{Streams in chaotic regions look very different from those in regular regions.}
-\chchchanges{We find that streams can be sensitive to chaos on a much shorter time-scale than any standard prediction (from the Lyapunov or frequency-diffusion times).}
+\chchchanges{Streams even in weakly chaotic regions look very different from those in regular regions.}
+We find that streams can be sensitive to chaos on a much shorter time-scale than any standard prediction (from the Lyapunov or frequency-diffusion times).
 For example, on a weakly chaotic orbit with a chaotic timescale predicted to be $>$1000 orbital periods ($>$1000 Gyr), the resulting stellar stream is, after just a few 10's of orbits, substantially more diffuse than any formed on a nearby but regular orbit.
@@
 As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded by tidal forces from the host-galaxy potential.
-Along regular, mildly eccentric orbits, \chchchanges{mass stripped from the progenitor has a small spread} in orbital properties (e.g., energy, angular momentum).
+Along regular, mildly eccentric orbits, mass stripped from the progenitor has a small spread in orbital properties (e.g., energy, angular momentum).
 Once the debris has evolved far enough from the progenitor system that the self-gravity of the progenitor can be ignored (usually a fast process relative to the orbital time), the stars evolve essentially as an ensemble of test particle orbits in the potential of the host system.
@@
 Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies; the known streams span a large range of distances---from $\approx10$ to $100~{\rm kpc}$---and progenitor masses---from $\approx$10$^3$ to $\approx$10$^8~\msun$ in stellar mass---\citep[][]{ibata94,odenkirchen01,belokurov06,grillmair06a,grillmair06b,bonaca12}.
-There has been extensive work on developing methods to use data from these streams to measure properties of \chchchanges{the} Milky Way's dark matter halo.
+There has been extensive work on developing methods to use data from these streams to measure properties of the Milky Way's dark matter halo.
 These methods span a range of complexity from orbit-fitting, to \emph{Streakline} \citep{kuepper12} or `particle-spray' models \citep{gibbons14}, to action-space density modeling \citep[e.g.,][]{sanders14, bovy14}, to N-body simulations \citep[e.g.,][]{law10}.
@@
 We choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos as computed from single-orbit diagnostics for grids of constant-energy orbits, and compare these results to measures of the density evolution of finite-volume ensembles of orbits (meant to mimic tidal debris).
-We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles through the chaotic diffusion of the \chchchanges{constituent} orbits.
+We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles through the chaotic diffusion of the constituent orbits.
 For a chaotic orbit, the frequency spectrum of the orbit evolves with time: for a small ensemble, the spread in frequencies is therefore time-dependent, which could enhance phase-mixing.
@@
 \end{equation}
-\chchchanges{where $q_k$ and $p_k$ are conjugate coordinates in the sense that
+where $q_k$ and $p_k$ are conjugate coordinates in the sense that
 \begin{align}
@@
 \end{equation}
-where the $a_{kj}$ are complex amplitudes.}
+where the $a_{kj}$ are complex amplitudes.
 
@@
 Even stronger, the actions are isolating integrals and any pair are in involution such that
-\chchchanges{\begin{equation}
+\begin{equation}
 	[J_\alpha, J_\beta] = 0
-\end{equation}}
+\end{equation}
 where $[\cdot,\cdot]$ is the Poisson bracket.
@@
 The topology of angle-action space is toroidal and any regular orbit in an $N$ dof Hamiltonian can be understood as motion on the surface of an $N$-torus.
-\chchchanges{Each set of actions, $(J_1,...,J_N)$ (or frequencies), labels a torus,} and regular orbits are sometimes referred to in terms of their orbital tori \chchchanges{(see, e.g., Section 50 in \citealt{arnold78}, Section 10-5 in \citealt{goldstein80}, Section 3.5 in \citealt{binneytremaine})}.
+Each set of actions, $(J_1,...,J_N)$ (or frequencies), labels a torus, and regular orbits are sometimes referred to in terms of their orbital tori (see, e.g., Section 50 in \citealt{arnold78}, Section 10-5 in \citealt{goldstein80}, Section 3.5 in \citealt{binneytremaine}).
 
@@
 Galactic potentials are almost certainly not globally integrable but it is useful to understand the orbit structure in integrable systems before extending to more general potentials.
-Only one globally integrable triaxial mass model is known that generates motion that is qualitatively similar to that seen in Galaxies: \chchchanges{the `Perfect Ellipsoid' model \citep[a special case of the St\"ackel model; see, e.g.,][]{kuzmin73, deZeeuw85}}.
-The four general classes of orbits in this potential are box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits.
-\chchchanges{Tube orbits have a non-zero time-averaged angular momentum about either the long or short axis and therefore never pass through the center of the potential (hence are centrophobic).
-Box orbits instead have a zero time-averaged angular momentum and hence have finite probability of passing through the center of the potential (hence are centrophilic).}
+\chchchanges{For example, there are four general classes of orbits in the `Perfect Ellipsoid' potential \citep[an integrable triaxial potential and special case of the St\"ackel potential; see, e.g.,][]{kuzmin73, deZeeuw85}: box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits.
+Regular orbits in non-integrable triaxial potentials can typically still be identified with these classes.
+Tube orbits have a non-zero time-averaged angular momentum about either the long or short axis and therefore never pass through the center of the potential (hence are centrophobic).
+Box orbits instead have a zero time-averaged angular momentum and therefore have finite probability of passing through the center of the potential.
+These orbits are generally centrophilic, though some resonant box orbits are also centrophobic (e.g., banana, pretzel, fish).}
 
-The frequencies of a generic orbit are typically incommensurable---that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} These \chchchanges{non-resonant} orbits uniformly cover the surface of an orbital torus.
+The frequencies of a generic orbit are typically incommensurable---that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} These non-resonant orbits uniformly cover the surface of an orbital torus.
 If instead there exists a relation of the form $\boldsymbol{n} \cdot \boldsymbol{\Omega} = 0$ the orbit is referred to as a resonant orbit.
@@
 These resonant layers are then generically surrounded by stochastic layers where chaotic motion occurs (in the vicinity of separatrices).
-\chchchanges{Chaotic orbits are not bound to the surface of a torus and instead fill a (2N-1)-dimensional volume where a continuous set of tori would exist in an unperturbed Hamiltonian, $H_0$.
+Chaotic orbits are not bound to the surface of a torus and instead fill a (2N-1)-dimensional volume where a continuous set of tori would exist in an unperturbed Hamiltonian, $H_0$.
 While the frequency spectrum of sub-sections of a regular orbit are indistinguishable (i.e.
-measured over a finite interval of time), the frequency spectra of sub-sections of a chaotic orbit will evolve stochastically.}
+measured over a finite interval of time), the frequency spectra of sub-sections of a chaotic orbit will evolve stochastically.
 
@@
 Much of the structure that was present in the integrable potential remains in the near-integrable case, but the differences are highlighted.
-Orbits in the resonant layers surrounding the resonances (grey) are near-resonant orbits that librate around the resonance \chchchanges{and have finite thickness \citep[e.g.,][]{merritt99}}.
+Orbits in the resonant layers surrounding the resonances (grey) are near-resonant orbits that librate around the resonance and have finite thickness \citep[e.g.,][]{merritt99}.
 Chaotic orbits in the stochastic layer (red) behave erratically depending on the surrounding resonance structure.
 If the stochastic layer is small and the chaotic orbit is therefore confined, the orbit may behave nearly regular for long periods of time.
-\chchchanges{If the resonance in the unperturbed Hamiltonian is unstable, all orbits associated with the resonance will be chaotic; unstable resonances form linear gaps in frequency-space (this is not shown in the cartoon but are seen in Figure~\ref{fig:logfreqs}).
-Thus, only some resonances in the perturbed system will retain the signature of a resonance.}
+If the resonance in the unperturbed Hamiltonian is unstable, all orbits associated with the resonance will be chaotic; unstable resonances form linear gaps in frequency-space (this is not shown in the cartoon but are seen in Figure~\ref{fig:logfreqs}).
+Thus, only some resonances in the perturbed system will retain the signature of a resonance.
 
-For small values of $\epsilon$, many regular orbits survive and only small chaotic regions are introduced, \chchchanges{especially in the vicinity of the intersection of two resonance lines \citep[commonly referred to as `resonance overlap'; see][]{chirikov60}}.
-As the strength of the perturbation increases, eventually most tori associated with \chchchanges{non-resonant} motion will be destroyed.
-Figure~\ref{fig:cartoons} (right panel) qualitatively shows this phenomenon---the resonant and resonant-layer orbits may still be regular, but many or all of the \chchchanges{non-resonant} orbits will be chaotic.
+For small values of $\epsilon$, many regular orbits survive and only small chaotic regions are introduced, especially in the vicinity of the intersection of two resonance lines \citep[commonly referred to as `resonance overlap'; see][]{chirikov60}.
+As the strength of the perturbation increases, eventually most tori associated with non-resonant motion will be destroyed.
+Figure~\ref{fig:cartoons} (right panel) qualitatively shows this phenomenon---the resonant and resonant-layer orbits may still be regular, but many or all of the non-resonant orbits will be chaotic.
 As the perturbation strength increases, eventually the uni- and bi-resonant tori are also destroyed---these are less susceptible to destruction from perturbations \cite[for a more quantitative illustration of this transition from integrability to global chaos, see Figure~9 in][]{valluri98}.
@@
 When $\epsilon$ is large, there is no general prediction for the resulting behavior, however it seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep[e.g.,][]{valluri98, merritt99}.
-We therefore expect a large number of regular orbits will survive---the so-called KAM tori---however the tori that survive will be separated by regions of chaotic motion.
+We therefore expect a large number of regular orbits will survive---the so-called Kolmogorov–Arnold–Moser (KAM) tori---however the tori that survive will be separated by regions of chaotic motion.
 Any transformations to angle-action coordinates must be defined local to each resonance region due to the destruction of tori and chaotic motion which lead to discontinuous changes in orbital properties.
@@
 Chaotic orbits have no orbital actions and only conserve energy (if the potential is time-independent).
-The orbits therefore do not have a single set of fundamental frequencies, \chchchanges{but rather the frequencies that describe the character of motion evolve with time.} In near-integrable potentials, \chchchanges{the frequencies of consecutive sub-sections of a chaotic orbit diffuse both around resonance layers\footnote{Note that during this diffusion, the frequencies never exactly hit those of a KAM torus but evolve stochastically around these discrete, stable tori \citep[cf.
-Figure 2 in][]{laskar99}.} (a sort of stochastic libration) and along the stochastic layers that surround the resonances (Arnold diffusion).} For weakly chaotic orbits, motion \chchchanges{around} a resonance layer can occur with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
-Thus, if the resonance libration frequency is small, motion \chchchanges{around} a resonance can modulate the frequency spectrum of an orbit over an orbital time.
-However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can not change by large factors (\chchchanges{unless there are other nearby resonances and overlapping stochastic layers, in which case the motion may be strongly chaotic}).
+The orbits therefore do not have a single set of fundamental frequencies, but rather the frequencies that describe the character of motion evolve with time.
+In near-integrable potentials, the frequencies of consecutive sub-sections of a chaotic orbit diffuse both around resonance layers\footnote{Note that during this diffusion, the frequencies never exactly hit those of a KAM torus but evolve stochastically around these discrete, stable tori \citep[cf.
+Figure 2 in][]{laskar99}.
+(a sort of stochastic libration) and along the stochastic layers that surround the resonances (Arnold diffusion).} For weakly chaotic orbits, motion around a resonance layer can occur with a frequency close to the libration frequency of the nearby stable orbits in the resonance layer.
+Thus, if the resonance libration frequency is small, motion around a resonance can modulate the frequency spectrum of an orbit over an orbital time.
+However, the stochastic layers are often bounded in the direction orthogonal to the resonance by other stable, resonant regions so that the frequencies or actions can not change by large factors (unless there are other nearby resonances and overlapping stochastic layers, in which case the motion may be strongly chaotic).
 
@@
 Arnold diffusion is not expected to be significant for most orbits over timescales relevant to galaxies (10s of orbits), however chaotic motion across resonances can occur over short times.
-\chchchanges{If a stochastic trajectory is surrounded by regular orbits, it may be trapped around a regular parent orbit for long periods of time before escaping to another such semi-bounded region where it can become trapped around another parent orbit (a process by which it may eventually explore the whole Arnold web)---such orbits are commonly referred to as `sticky orbits.' Additionally, if the volume of the surrounded region in frequency space is comparable to the characteristic spread of frequencies in the tidal debris, this small-scale evolution will be important for tidal debris.}
+If a stochastic trajectory is surrounded by regular orbits, it may be trapped around a regular parent orbit for long periods of time before escaping to another such semi-bounded region where it can become trapped around another parent orbit (a process by which it may eventually explore the whole Arnold web)---such orbits are commonly referred to as `sticky orbits.' Additionally, if the volume of the surrounded region in frequency space is comparable to the characteristic spread of frequencies in the tidal debris, this small-scale evolution will be important for tidal debris.
 
@@
 For example, ensembles of orbits `stuck' between resonances may quickly spread to fill the allowed volume, but then the orbits must escape this confinement and diffuse through the Arnold web to reach a fully mixed state \citep{merritt96}.
-\chchchanges{That is, while the orbit is struck, the small-scale variations effectively cause an increase in the variance of the frequency distribution, which would enhance mixing of the debris in configuration-space.} In this work, we investigate the consequences of short-time but small-scale frequency evolution and hypothesize that this may explain the enhanced density evolution of tidal debris around weakly chaotic regions where chaotic timescales are predicted to be long \citep[e.g.,][]{pearson15}.
+That is, while the orbit is stuck, the small-scale variations effectively cause an increase in the variance of the frequency distribution, which would enhance mixing of the debris in configuration-space.
+In this work, we investigate the consequences of short-time but small-scale frequency evolution and hypothesize that this may explain the enhanced density evolution of tidal debris around weakly chaotic regions where chaotic timescales are predicted to be long \citep[e.g.,][]{pearson15}.
 We then discuss how this would affect our understanding of the coherence and density evolution of tidal streams.
@@
     }
-\caption{\chchchanges{Equipotential contours (left) and isodensity contours (right) for the triaxial NFW potential considered in this work.
+\caption{Equipotential contours (left) and isodensity contours (right) for the triaxial NFW potential considered in this work.
 For the potential plot there are eight contour levels evenly spaced and linear in the value of the potential.
-For the density plot there are eight contour levels logarithmically spaced from $10^4~\msun~{\rm kpc}^{-3}$ to $10^7~\msun~{\rm kpc}^{-3}$.}}
+For the density plot there are eight contour levels logarithmically spaced from $10^4~\msun~{\rm kpc}^{-3}$ to $10^7~\msun~{\rm kpc}^{-3}$.}
 \label{fig:potential}
@@
 This potential is a simple (and unrealistic) model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the haloes of such galaxies.
-Realistic galactic potentials will have a significant component due to the disk and bulge, \chchchanges{radially changing axis ratios or orientations \citep[e.g.,][]{romanowsky98, kazantzidis04,debattista08,veraciro11,butsky15}}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.; see, e.g.,][]{bailin05}.
+Realistic galactic potentials will have a significant component due to the disk and bulge, radially changing axis ratios or orientations \citep[e.g.,][]{romanowsky98, kazantzidis04,debattista08,veraciro11,butsky15}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.; see, e.g.,][]{bailin05}.
 We expect inclusion of any of these effects to increase the complexity of the resonant structure and influence of chaos (see Section~\ref{sec:discussion}).
@@
 
-The most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit \chchchanges{\citep{lyapunov92}}.
+The most well-known method for assessing chaotic motion is to analyze the Lyapunov spectrum or maximum Lyapunov exponent (MLE) of an orbit \citep{lyapunov92}.
 The MLE measures the mean rate of divergence of two infinitesimally separated orbits and is only strictly defined in terms of a limit that goes to infinite time.
@@
 
-For a small ensemble of orbits (e.g., tidal debris), \chchchanges{a more} relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
+For a small ensemble of orbits (e.g., tidal debris), a more relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble.
 We can estimate this timescale by multiplying the frequency diffusion time by a factor equal to the fractional spread in frequencies of the debris.
@@
 We classify each orbit in terms of the strength of chaos along the orbit as computed using the Lyapnov and frequency diffusion times of Section~\ref{sec:methods}.
-With this initial classification, in Section~\ref{sec:results2} we follow the evolution of ensembles of \chchchanges{trajectories} generated around each orbit in the initial grid.
+With this initial classification, in Section~\ref{sec:results2} we follow the evolution of ensembles of trajectories generated around each orbit in the initial grid.
 We find that the configuration-space density of ensembles around weakly chaotic orbits evolve faster (e.g., in mean density) than expected given the timescales over which chaos is computed to be relevant for the parent orbits.
@@
 This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits).
-\chchchanges{The most numerous orbits in this grid are the short-axis and long-axis tubes that circulate about either the minor or major axis.
+The most numerous orbits in this grid are the short-axis and long-axis tubes that circulate about either the minor or major axis.
 Thin tidal streams may preferentially form along tube orbits rather than box orbits because of the faster disruption and debris diffusion expected for stellar systems on radially plunging orbits.
 Figure~\ref{fig:apoper} shows the grid of initial conditions in the $xz$ plane---each pixel in this grid represents an orbit, and the pixels are colored by the median pericentric distance (left) and median apocentric distance (right) over an integration time of 64 orbital periods.
-From here onwards, all lengths are given in units of scale radii, $r_s$, velocities in units of circular velocity, $v_c$, (see Table~\ref{tbl:potential}) and times in units of orbital periods, $\periods$.}
+From here onwards, all lengths are given in units of scale radii, $r_s$, velocities in units of circular velocity, $v_c$, (see Table~\ref{tbl:potential}) and times in units of orbital periods, $\periods$.
 
@@
 For a given set of orbital initial conditions---the `parent' orbit---we integrate for 128 orbital periods, find the phase-space position of the minimum pericenter over this time, initialize a small ensemble of test particle orbits around this position, and integrate the orbits of all test particles for some integration time.
-We are interested in the degree to which chaos enhances the mixing rate of orbit ensembles and we therefore want to isolate out the effect of ordinary phase-mixing along regular orbits\chchchanges{\sout{; all ensembles should start with the same initial spread in fundamental frequencies}}.
+We are interested in the degree to which chaos enhances the mixing rate of orbit ensembles and we therefore want to isolate out the effect of ordinary phase-mixing along regular orbits.
 We set the physical scale of the ensemble by the tidal radius in position and the velocity scale in velocity; the initial spread in fundamental frequencies will scale as $(m/M)^{1/3}$ like the tidal radius and velocity scale \citep[e.g.,][]{johnston98, apw14}.
@@
 
-We start by considering four particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a \chchchanges{more} strongly chaotic orbit (D).
+We start by considering four particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular, near-resonant orbit (A), a regular, non-resonant orbit (B), a weakly chaotic orbit (C), and a more strongly chaotic orbit (D).
 The orbits were chosen to be close on the orbit grid so that their orbital properties (e.g., apocenter, pericenter) are similar, but have different frequency diffusion times; all four orbits are long-axis tube orbits.
-Figure~\ref{fig:orbits} shows the orbits in projection over \chchchanges{an integration period of 1024 orbital periods}.
+Figure~\ref{fig:orbits} shows the orbits in projection over an integration period of 1024 orbital periods.
 The initial conditions and chaos diagnostics for each orbit are listed in Table~\ref{tbl:orbit-info}.
-Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the four orbits described above and \chchchanges{integrated} for 64 orbital periods.
-A thin stream forms on the near-resonant orbit (left column), a thin---but more two-dimensional---stream forms on the non-resonant orbit (middle-left), a more diffuse stream forms on the weakly chaotic orbit with a slightly `fanned' morphology (middle-right), and a two-dimensional, `fanned' stream on the \chchchanges{more} strongly chaotic orbit (right).
+Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the four orbits described above and integrated for 64 orbital periods.
+A thin stream forms on the near-resonant orbit (left column), a thin---but more two-dimensional---stream forms on the non-resonant orbit (middle-left), a more diffuse stream forms on the weakly chaotic orbit with a slightly `fanned' morphology (middle-right), and a two-dimensional, `fanned' stream on the more strongly chaotic orbit (right).
 Given the long Lyapunov and frequency diffusion times of the weakly chaotic orbit ($t_\lambda/\periods > 900$, $t_\Omega/\periods \approx 2\times10^5$), it is surprising that the density evolution of the ensemble on this orbit appears to be more diffuse than the regular orbit stream.
@@
 The progenitor in each simulation is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking.
-We start the progenitor systems at apocenter and \chchchanges{integrate} for $\approx$64 orbital periods so that each simulation finishes again at subsequent apocenter.
+We start the progenitor systems at apocenter and integrate for $\approx$64 orbital periods so that each simulation finishes again at subsequent apocenter.
 The mass of the progenitor is set to $10^4~\msun$, and the length-scale of the Plummer sphere is set to 5~pc to get $\approx$50--75\% mass loss over the integration time.
@@
 The non-resonant orbit ensemble (B) is `thicker'---the ratio of the two largest eigenvalues of this distribution is larger---and therefore in real space, the ensemble begins to spread over two dimensions, as is also evident in the final ensemble debris morphology in the middle-left panel of Figure~\ref{fig:ensembles}.
-Non-resonant orbits in axisymmetric or triaxial potentials \chchchanges{will generically have frequency distributions that have more comparable spreads in each direction} and the configuration-space density of typical tidal debris will therefore decrease faster in such potentials \citep{helmi99}.
+Non-resonant orbits in axisymmetric or triaxial potentials will generically have frequency distributions that have more comparable spreads in each direction and the configuration-space density of typical tidal debris will therefore decrease faster in such potentials \citep{helmi99}.
 The weakly chaotic orbit ensemble (C) is clearly more diffuse than the two regular ensembles: As is evident in Figure~\ref{fig:four-orbits-freqs}, the small-scale chaotic evolution of the frequencies---though bounded---increases the variance of the frequency distribution.
@@
 We have shown that the Lyapunov and frequency diffusion times are indicators of chaos and that the frequency diffusion time resolves the detailed resonant structure of gravitational potentials, but the timescales predicted do not capture the importance of chaos for the density evolution of ensembles of orbits meant to mimic tidal debris.
-We have shown that small-scale but fast chaotic diffusion of \chchchanges{frequencies} can explain this enhanced mixing.
+We have shown that small-scale but fast chaotic diffusion of frequencies can explain this enhanced mixing.
 In the sub-sections below, we discuss a few important limitations that remain for exploration in future work.
@@
 We have only considered low-mass progenitor systems such as globular clusters because the intrinsic spreads in fundamental frequencies are small (0.1--0.5\%).
-Small changes to the frequencies of the orbits of tidal debris \chchchanges{stripped} from these progenitors due to chaotic processes will therefore cause observable changes to the real-space morphology of the debris.
+Small changes to the frequencies of the orbits of tidal debris stripped from these progenitors due to chaotic processes will therefore cause observable changes to the real-space morphology of the debris.
 For more massive progenitor systems, the typical size and velocity dispersion of the debris will be larger and thus the debris morphology will be less sensitive to small changes in orbital properties.
@@
 \cite{deibel11} showed that by adopting pattern speeds comparable to those found in cosmological simulations, figure rotation generally acts to destabilize orbits (rather than stabilize chaotic orbits in the equivalent static potential) and the resulting orbit structure is that most regular orbits are associated with resonances.
-\chchchanges{The presence of and response to the Large and Small Magellanic Clouds may also introduce significant (time-dependent) perturbations to the global potential of the Milky Way \citep[e.g.,][]{besla10, gomez15}}.
+The presence of and response to the Large and Small Magellanic Clouds may also introduce significant (time-dependent) perturbations to the global potential of the Milky Way \citep[e.g.,][]{besla10, gomez15}.
 In future work we will explore the effect of these time-dependent processes on the chaotic dispersal of tidal streams using live potentials from cosmological N-body simulations.
@@
 At Galactocentric $R \approx 1$ kpc, $z \approx 5$ kpc, the orbits of the stream stars likely pass through the MW disk, feel the triaxiality of the Galactic bar \citep[e.g.,][]{wegg13, wegg15}, and have short orbital periods (relative to streams in the halo).
-It is possible that the observed debris is the last remnants of the recently disrupted progenitor system \citep{sesar15}, however if a significant number of stars were \chchchanges{stripped} on previous pericentric passages, this older debris may be `fanned' and still near the observed portion of the stream.
-\chchchanges{The fanned debris would have significantly lower surface brightness and thus would be more difficult to detect.} The detection of this low-surface-brightness component would open up the possibility that enhanced density evolution due to chaos is a dynamically relevant process for thin streams in real galaxies, though \cite{carlberg15} has shown that it is also possible to get short, high density segments of streams from debris formed on eccentric orbits.
+It is possible that the observed debris is the last remnants of the recently disrupted progenitor system \citep{sesar15}, however if a significant number of stars were stripped on previous pericentric passages, this older debris may be `fanned' and still near the observed portion of the stream.
+The fanned debris would have significantly lower surface brightness and thus would be more difficult to detect.
+The detection of this low-surface-brightness component would open up the possibility that enhanced density evolution due to chaos is a dynamically relevant process for thin streams in real galaxies, though \cite{carlberg15} has shown that it is also possible to get short, high density segments of streams from debris formed on eccentric orbits.
 
@@
 The authors wish to thank Robyn Sanderson, Dan D'Orazio, David Merritt, and the \emph{Stream Team} for useful comments and discussion.
+\chchchanges{The authors additionally thank the referee, Walter Dehnen, for insightful comments that greatly improved this article.}
 

### 2016-04-06 ms.tex
arxiv resubmission

@@
-\documentclass{mn2e}
-%\documentclass[letterpaper,12pt,preprint]{aastex}
+%\documentclass{mn2e}
+\documentclass[letterpaper,12pt,preprint]{aastex}
 
@@
 
-The Ophiuchus stream \citep{bernard14, sesar15} appears to be a thin, short tidal stream (deprojected length $\approx$1.5--2 kpc) near the Galactic bulge with no apparent progenitor.
+The Ophiuchus stream \citep{bernard14, sesar15a} appears to be a thin, short tidal stream (deprojected length $\approx$1.5--2 kpc) near the Galactic bulge with no apparent progenitor.
 At Galactocentric $R \approx 1$ kpc, $z \approx 5$ kpc, the orbits of the stream stars likely pass through the MW disk, feel the triaxiality of the Galactic bar \citep[e.g.,][]{wegg13, wegg15}, and have short orbital periods (relative to streams in the halo).
-It is possible that the observed debris is the last remnants of the recently disrupted progenitor system \citep{sesar15}, however if a significant number of stars were stripped on previous pericentric passages, this older debris may be `fanned' and still near the observed portion of the stream.
+It is possible that the observed debris is the last remnants of the recently disrupted progenitor system \citep{sesar15a}, however if a significant number of stars were stripped on previous pericentric passages, this older debris may be `fanned' and still near the observed portion of the stream.
 The fanned debris would have significantly lower surface brightness and thus would be more difficult to detect.
